### 2020-04-23 docs/HLL_exp_bad/main.tex
latex source files

\documentclass{IEEEtran}
\usepackage{graphicx}
\usepackage{array}
\usepackage{url}
\usepackage{hyperref}
\title{\huge HyperLogLog, exponentially bad in Adversarial Setting}
\author{Mathilde Raynal, Kenneth G. Paterson}

\begin{document}

\maketitle

\IEEEtitleabstractindextext{%
\begin{abstract}
Computing the count of distinct elements in massive data sets is often necessary but memory or computationally expensive. The HyperLogLog (HLL) algorithm \cite{hll2} estimates a data stream’s cardinality while using significantly less memory at the expense of some accuracy. This trade off makes it a really attractive solution that can find its use in a wide range of applications, such as database management or network monitoring. Its recent popularity made it an interesting target for attackers, and vulnerabilities are starting to be acknowledged. This works puts itself in the continuity of \cite{hllvuln} and improves their attack on the HLL accuracy. We show that disclosing too much information about the HLL internals opens the door to estimate manipulation by dishonest users, that may result in an exponential reduction from the true cardinality.
\end{abstract}
}

\IEEEdisplaynontitleabstractindextext

\section{Introduction}
Obvious approaches to computing the cardinality of a big data set, such as sorting the elements or simply maintaining the set of unique elements seen, are impractical. HyperLogLog is the most widely used cardinality estimator nowadays, and the first reviews of its capabilities show a lack of security both in privacy and correctness. We move forward in exploiting the vulnerabilities of the sketch and show that an attacker can easily manipulate the HLL estimate with very small knowledge about the sketch and a limited number of computations.\\

The rest of the paper is organized as follows. Section 2 gives an overview of the HyperLogLog algorithm. Section 3 presents our attack and evaluates its impact under different adversarial models. Section 4 provides a conclusion and some ideas for future work.

\section{HyperLogLog}
\subsection{Algorithm}
HLL is a streaming algorithm based on the bit pattern observation that for a stream of randomly distributed numbers, if we observe a number $x$ with the maximum of $k$ leading 0 bits, the cardinality of the stream is likely to be 2$^{k+1}$. In practice, to guaranty uniform distribution of the inputs, they go through a hash function $h$ before being processed. This simple approach has the disadvantage of a large variance. In order to improve the estimate, we can use many estimators (buckets) instead of one and average the results. To map a value $h(x)$ to a bucket, \cite{loglog} suggests using the first few $n$ bits of $h(x)$ as an index, and then compute the longest sequence of leading 0s on the $l = len(h(x)) - n $ bits left. Estimators from all buckets are averaged with an harmonic mean and scaled by a constant $\alpha$, empirically computed to correct a systematic multiplicative bias. As result, the HyperLogLog algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using 1.5 kB of memory.

An interesting property of the HyperLogLog sketch is its ability to be merged in a lossless way. To combine the two buckets at index $i$, take the maximum estimator of the two and assign that to the matching bucket $i$ in the unioned HLL. Such a simple set union operation allows us to easily parallelize operations among multiple machines independently using the same hash function and the same number of buckets.

\subsection{Applications}
For completeness and motivation, we provide examples of concrete uses of HyperLogLog:
\begin{itemize}
    \item Cardinality estimation provides a significantly faster way for Facebook to compute statistics on their users. They use HyperLogLog to find out how many distinct people visited their website in the past week \cite{fbhll}.
    \item A solution to detecting DOS attacks using a sliding HyperLogLog \cite{slidinghll} (a variation which estimates the cardinality over a moving time window) is presented in \cite{portscanhll}. The sketch is used to count how many different ports are used as destination over all packets received, as an attempt to identify when a port scan attack is happening.
    \item \cite{flexswitch} implements the HyperLogLog on a light FlexSwitch to approximate the number of distinct packets in traffic flows and overcome the switch resource limits. The number of unique flows traversing a switch can then be used to provide a better congestion control protocol.
\end{itemize}

Regardless of the sensitivity of the applications, none of them are clear about their threat model, especially for the HyperLogLog. By not assessing the risks involved with the presence of an adversary targeting the HLL sketch, they may allow an attack like the one we present, breaking down the whole solution.

\subsection{Related Work}
The work of \cite{cardestprivacy} breaks the privacy of the HLL sketch and concludes that having the ability to test the insertion of a targeted item gives significant confidence about whether the item was originally present in the input data set. HLL sketches can be used by organizations to store and process location data, which is inherently sensitive, so loosing privacy would leak one user's location and potentially cause great harm. On the other hand, \cite{hllvuln} exploits a vulnerability of the HLL algorithm to manipulate the cardinality estimate, leading to a five-fold reduction from the true cardinality.

Both propose mitigations using a salted sketch, either as replacement, but at the cost of loosing mergeability, or in addition to an unsalted sketch, inducing a memory overhead. This might not be a convenient solution for some applications, they thus encourage further research.

\section{Vulnerabilities}
This section presents different adversarial settings and the corresponding attacks to manipulate the HLL estimate. The attacker's goal in our work is to bring the estimate down, as this would cause the most damage when looking at the applications of HLL. One can think about the attacker trying the hide from a detection system. Conveniently, it is easy to adopt the same strategy as the one we present for an attacker who is looking to artificially inflate the HLL estimate given a limited data set input. The impact over the cardinality estimate and the computational effort to be done by the attacker is analyzed afterwards.

\subsection{Adversarial setting}
We model 3 adversarial scenarios based on the attacker's knowledge and capabilities. For all of them, we assume that the sketch already contains some existing data from honest users to simulate a real life setup. Among the scenarios, S3 has the strongest assumptions as it combines the knowledge of attackers from S1 and S2.

S1: the attacker is able to test insertion of an item on a black-box snapshot of the sketch in use and check the effect on the cardinality estimate.

S2: the attacker has access to the details of the implementation, i.e., the number of buckets and the hash function $h$.

S3: the attacker has direct access to the sketch in use and all internals, i.e., the number of buckets, the hash function $h$, and also the sketch's buckets content. Those buckets hold the intermediate estimators updated by the honest user's data, that are averaged in a later step to compute the global cardinality estimator.\\

S1 and S2 are presented as M2 and M1 respectively in \cite{hllvuln}. We do not put our focus on S1 as this previous work fully exploit this scenario. The attack works as follow: ``after each insertion the cardinality is checked and the items which do not increment the HLL estimate are retained``. We start by presenting our attack in S2, providing much stronger results then the corresponding attack of \cite{hllvuln}, and then use the additional information the attacker has access to in S3 to make the attack even more efficient.

\subsection{Manipulating the estimate}
Under S2: The attacker has access to $h$ so he can adjust some field of the item $x$ he is trying to insert and keep only the one such that the $l$ rightmost bits of $h(x)$ start with a 1. Indeed, the number of leading 0 bits computed by the estimator of the destination bucket is always zero, so the estimator of this bucket is unlikely to be updated and the averaged approximated cardinality is left unchanged. As the attacker knows how many buckets are used, he can recover $n$ and $l$, and aim for the $n$th bit (first bit not used for bucket allocation) to be 1. For example, in the scenario of \cite{portscanhll}, the attacker could change the source port of his IP packet $x$ until he is satisfied with $h(x)$. \textbf{(what about repeated sampling?)}\\

Under S3: The attacker can adopt the same strategy but make the attack more efficient and reduce the sampling requirements with the additional information. Now that the attacker has the sketch, not only he knows the overall estimated cardinality but also the intermediate estimators computed per bucket. So, he can "allow" some leading 0 bits in the $l$ rightmost bits of $h(x)$ as long as it does not increment the bucket's estimate. In other words, if $h(x)$ is mapped to the $i$th bucket with estimated cardinality $2^{c_i + 1}$ (it observed a value with $c_i$ leading 0s), then, instead of the strict restriction of a 1 bit at the $n$th index, the attacker is satisfied when having at least one 1 bit in the $n$ to $n+c_i + 1$ bit slice of $h(x)$. In case the attacker hits an empty bucket (the estimator is 0), he must skip that bucket completely. Under those conditions, the estimator of the buckets are not updated, and there is no effect on the cardinality estimation of the HLL sketch.

\subsection{Impact}
Now that we have a way to exploit the vulnerability, we need to evaluate to which extent it can affect the estimate. We assume that $h$ is a "good" hash function, so any bit of $h(x)$ has a probability of $\frac{1}{2}$ to be 1 and $\frac{1}{2}$ to be 0.\\

Under S2, the attacker need to craft 2 $x$s on average to obtain a 1 bit at the $n$th position of $h(x)$ and successfully perform the attack. As we do not care about the remaining bits, the attacker can theoretically "hide" up to 2$^{len(h(x))-1}$ items into the sketch.\\

Under S3, let's compute the average work for the attacker to find an $x$ such that $h(x)$ meets the attack's requirements. We assume $h(x)$ is mapped to bucket $i$. The probability that $h(x)$ does not increment the estimator is the probability that $h(x)$ has less than $c_i$ leading 0s. So:\\
$P(h(x)$ does NOT update bucket $i| h(x)$ mapped to bucket $i)=P(h(x)$ has LESS than $c_i$ leading 0s$)=1-P(h(x)$ has strictly MORE than $c_i$ leading 0s$)=1-\frac{1}{2^{c_i+1}}$\\
Across all buckets, the attacker does not update any estimate with probability $P=
\sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i$ AND $h(x)$ does NOT update bucket $i) = \sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i)P(h(x)$ does NOT update bucket $i| h(x)$ mapped to bucket $i) = \sum_{i=1}^{2^n}\frac{1}{2^n} \times (1-\frac{1}{2^{c_i+1}}) = 1-\frac{1}{2}\times(\frac{1}{2^n}\times \sum_{i=1}^{2^n} {(2^{c_i})}^{-1}) = 1-\frac{1}{2}\times\frac{1}{H_c} = 1-(2H_c)^{-1}$, with $H_c$ the harmonic mean of the numbers $2^{c_1}, 2^{c_2}, ..., 2^{c_{2^n}}$.
So the attacker needs to craft on average $(1-(2H_c)^{-1})^{-1} x$s, and can theoretically "hide" up to 2$^{len(h(x))}\times (1-(2H_c)^{-1})$ items into the sketch.

\textit{NOTE: does it take into account the special case of empty bucket ? counters are initialized to $c_i=-oo$ (or -1 in the code???) thus $P(h(x)$ has LESS than $-oo$ leading $0s) = 0$ but it is hard to add it to the probabilities while staying clear.}\\
%2$^{len(h(x))}\times \frac{2H_c-1}{H_c}$ items into the sketch.\\

We assume here that there already are values coming from other honest users' data in the sketch, we now remove this assumption. Unfortunately, the attacker in S2 has no way to detect that he is attacking an empty sketch and cannot adapt his strategy, so all buckets hit will update once their estimator to $2^{0+1}=2$ items, leading to the sketch final approximation of the number of distinct elements to $\alpha2^{n+1}$ regardless of the amount of items the attacker will send. However, the attacker in S3 is able to notice that all buckets are empty and adapt his attack. As mentioned above, the attacker in S3 avoids empty buckets, but when dealing with an empty sketch, sacrifices must be made if the attacker wants to insert at least one item. To limit the detection, the attacker can target and fill one bucket only. This would fix $n+1$ bits of $h(x)$, allowing the attacker to add $2^{len(h(x))-n-1}$ elements while keeping the cardinality to 1.

\subsection{Experimental results}
There exist many implementations of the HyperLogLog algorithm available online. Among the repositories collecting more than 350 stars on github, we can cite \cite{clahll} and \cite{datasketch}. It is also featured in many frameworks, including Redis \cite{redis}, Google's BigQuery \cite{bigquery} and several products of Apache such as Spark \cite{spahll} \cite{spahll2} and Druid \cite{druhll}. It is interesting to point out that they all use a constant number of buckets and a fixed hash function, \textit{which is public or should be public according to Kerckhoffs's principle,} thus are potentially vulnerable to cardinality estimate manipulation when put in an adversarial setting.\\

As proof-of-concept, we implement our attack against the HLL algorithm from \cite{clahll}. The code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\textit{\url{https://github.com/PizzaWhisperer/HLLVuln}}}. Our items are 4 character long ASCII strings \textit{(7'311'616 distinct items)} and we choose $h$ to be the 32 bits Murmur3 hash function from \cite{murmur3code}. We create an HLL sketch using 256 buckets and initialize it with 1'000 random items representing an honest user's data.

We challenge the attacker to pick the largest possible number of items to add to the sketch while trying to keep the cardinality estimate as low as possible, either in the S2 or the S3 scenario. In a second time, in order to get results that can be compared to the work of \cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from the universe of possible strings, the attacker has to choose from a set of 250'000 random items that are given to him. Finally , for completeness, we also run the attacks against empty sketches. All attacks are ran 30 times and the outcomes are reported in Tab. \ref{table:tab1} and Tab. \ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the attacker and the approximation of the cardinality after adding the attacker's items.

Those empirical results confirm our impact analysis on several points. First, when attacking a sketch already containing some items, we can see that the attacker is able to hide approximately half of the items' set under S2, and around 83\%%( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$).
$\approx (1-\frac{1}{2\times H_c})$ with $H_c\approx\frac{1'000}{256}=3.9$, under S3.
Secondly, when attacking an empty sketch, the final cardinality estimate is always exactly 367 $=\alpha2^{n+1}$ with $\alpha = 0.72$ as specified by the Eq. (29) of \cite{hll2} or 1 in respectively S2 and S3.

To assess the efficiency of our attacks, we can confront our results from Tab. \ref{table:tab2} with the attack presented in the Section 5.1 of \cite{hllvuln}, where they target the Redis \cite{redishll} open-source implementation of HLL and perform the attack against an empty sketch. Given a set of 250'000 items, the attacker could choose ``74'390 distinct items and obtain an HLL estimate of only 15'780`` achieving a five-fold reduction from the true cardinality. In comparison, the attacker following our strategy is able to add 125'000 distinct items while keeping the estimate at 367, demonstrating a reduction by a factor of 340.

\begin{table}[h]
\caption{Result of the attacks over 30 iterations}
\begin{tabular}{| m{8.5em} | m{4em} | m{4em} | m{4em} | m{4em} |}
    \hline
    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S3} \\ \hline
    \textbf{Initial data?} & yes & no & yes & no \\ \hline
    \textbf{Original Est. Card.} & 1'000 & 0 & 1'005 & 0 \\ \hline
    \textbf{\# Items added} & 3'655'744 & 3'655'740 & 6'090'003 & 14'220 \\ \hline
    \textbf{Final Est. Card.} & 1'011 & 367 & 1'005 & 1 \\ \hline
\end{tabular}
\label{table:tab1}
\end{table}

\begin{table}[h]
\caption{Result of the attacks in the setting of \cite{hllvuln} over 30 iterations}
\begin{tabular}{| m{8.5em} | m{4em} | m{4em} | m{4em} | m{4em} |}
    \hline
    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S3} \\ \hline
    \textbf{Initial data?} & yes & no & yes & no \\ \hline
    \textbf{Original Est. Card.} & 1'004 & 0 & 998 & 0 \\ \hline
    \textbf{\# Items added} & 124'940 & 125'085 & 204'669 & 501 \\ \hline
    \textbf{Final Est. Card.} & 1'058 & 367 & 998 & 1 \\ \hline
\end{tabular}
\label{table:tab2}
\end{table}

\section{Conclusion}
The HyperLogLog algorithm is a elegant and efficient solution to the distinct count problem. Its simple structure makes it easy to code and use, as shows the growing number of available open-source implementations. Nonetheless, disclosing such low level details can be exploited by malicious users to manipulate the estimate of the sketch and break the system that is built on top. Our attack is trivial but powerful and should raise awareness on the security of HLL, such that product designers understand the risks they may encounter when exposed to an adversarial setting. Future work could consist in finding mitigations that do not rely on salt. For example, would the imposed ID setting (like Facebook) solve the issue but still be convenient for the various HLL applications? Also left for future work is to formalize the extension of our attack to the HLL variants, i.e., to show that it also breaks the HLL++ \cite{hllpratice} and sliding HLL \cite{slidinghll} algorithms.

\bibliographystyle{ieeetr}
\bibliography{ref.bib}

\end{document}

### 2020-04-24 docs/HLL_exp_bad/main.tex
abstract and intro

@@
 \usepackage{hyperref}
-\title{\huge HyperLogLog, exponentially bad in Adversarial Setting}
+
+%% TODO NOTES
+
+\usepackage{xcolor}
+\definecolor{oxygenorange}{HTML}{FFDD00}
+\usepackage[color=oxygenorange]{todonotes}
+\newcommand{\mathilde}[2][inline]{\todo[#1]{\textbf{Mathilde:} #2}\xspace}
+\newcommand{\kenny}[2][inline]{\todo[#1]{\textbf{Kenny:} #2}\xspace}
+
+\title{\huge HyperLogLog: Exponentially Bad in Adversarial Settings}
 \author{Mathilde Raynal, Kenneth G. Paterson}
@@
 \begin{abstract}
-Computing the count of distinct elements in massive data sets is often necessary but memory or computationally expensive.
-The HyperLogLog (HLL) algorithm \cite{hll2} estimates a data stream’s cardinality while using significantly less memory at the expense of some accuracy.
-This trade off makes it a really attractive solution that can find its use in a wide range of applications, such as database management or network monitoring.
-Its recent popularity made it an interesting target for attackers, and vulnerabilities are starting to be acknowledged.
-This works puts itself in the continuity of \cite{hllvuln} and improves their attack on the HLL accuracy.
-We show that disclosing too much information about the HLL internals opens the door to estimate manipulation by dishonest users, that may result in an exponential reduction from the true cardinality.
+Computing the count of distinct elements in massive data sets is a common task but naive approaches are memory-expensive.
+The HyperLogLog (HLL) algorithm (Durand and Flajolet, 2003) estimates a data stream's cardinality while using significantly less memory than a naive approach at the cost of some accuracy.
+This trade-off makes the HLL algorithm very attractive for a wide range of applications such as database management and network monitoring, where an exact count may not be needed.
+Recently, the HLL algorithm has started to be proposed for use in scenarios where the inputs may be adversarially generated, for example detection of network scanning attacks.
+This prompts an examination of the performance of the HLL algorithm in the face of adversarial inputs.
+We show that in such a setting, the HLL algorithm's estimate of cardinality can be exponentially bad: when an adversary has access to the internals of the HLL algorithm, an adversary who has some flexibility in choosing what inputs will be recorded can manipulate the cardinality estimate to be exponentially smaller than the true cardinality.
 \end{abstract}
@@
 \section{Introduction}
-Obvious approaches to computing the cardinality of a big data set, such as sorting the elements or simply maintaining the set of unique elements seen, are impractical.
-HyperLogLog is the most widely used cardinality estimator nowadays, and the first reviews of its capabilities show a lack of security both in privacy and correctness.
-We move forward in exploiting the vulnerabilities of the sketch and show that an attacker can easily manipulate the HLL estimate with very small knowledge about the sketch and a limited number of computations.\\
+Naive approaches to computing the cardinality of a big data set, such as sorting the elements or simply maintaining the set of unique elements seen, are impractical at scale.
+The HyperLogLog algorithm (HLL henceforth)~\cite{HLL } is today the most widely used cardinality estimator.
+It is increasingly used in settings where adversaries may have  incentives to manipulate the estimate made by the algorithm.
+For example,~\cite{} proposes its use for detecting network scanning attacks, while Facebook engineers have reported~\cite{} that they use HLL for estimating the number of active users.
+These applications demand a careful evaluation of the performance of HLL under adversarial input.
+
+Very recently, Reviriego and Ting~\cite{hllvuln} initiated work in this direction, showing that in specific attack settings, the cardinality estimate made by HLL could be modified through input selection.
+In this paper, we present a complete analysis of HLL in adversarial setting, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
+In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary can make HLL underestimate true cardinality by a factor this is exponential in the number of buckets used by the algorithm.
+
+\kenny{M: Please fill in missing references above.}
 
 The rest of the paper is organized as follows.
-Section 2 gives an overview of the HyperLogLog algorithm.
-Section 3 presents our attack and evaluates its impact under different adversarial models.
-Section 4 provides a conclusion and some ideas for future work.
+Section~\ref{sec:overview} gives an overview of the HyperLogLog algorithm.
+Section~\ref{sec:attacks} presents our different adversarial models.
+There, we also give attacks and evaluate their impacts under the different adversarial models.
+Section~\ref{sec:conclusions} provides our conclusions and some ideas for future work.
 
-\section{HyperLogLog}
+\section{HyperLogLog}\label{sec:overview}
 \subsection{Algorithm}
@@
 
-\section{Vulnerabilities}
+\section{Vulnerabilities}\label{sec:attacks}
 This section presents different adversarial settings and the corresponding attacks to manipulate the HLL estimate.
@@
 
-\section{Conclusion}
-The HyperLogLog algorithm is a elegant and efficient solution to the distinct count problem.
+\section{Conclusion}\label{sec:conclusions}
+The HyperLogLog algorithm is an elegant and efficient solution to the distinct count problem.
 Its simple structure makes it easy to code and use, as shows the growing number of available open-source implementations.

### 2020-04-24 docs/HLL_exp_bad/main.tex
work on section 2

@@
 
-\title{\huge HyperLogLog: Exponentially Bad in Adversarial Settings}
+\title{HyperLogLog: Exponentially Bad in Adversarial Settings}
 \author{Mathilde Raynal, Kenneth G. Paterson}
@@
 Naive approaches to computing the cardinality of a big data set, such as sorting the elements or simply maintaining the set of unique elements seen, are impractical at scale.
-The HyperLogLog algorithm (HLL henceforth)~\cite{HLL } is today the most widely used cardinality estimator.
+The HyperLogLog algorithm (HLL henceforth)~\cite{loglog,hll2} is today the most widely used cardinality estimator.
 It is increasingly used in settings where adversaries may have  incentives to manipulate the estimate made by the algorithm.
-For example,~\cite{} proposes its use for detecting network scanning attacks, while Facebook engineers have reported~\cite{} that they use HLL for estimating the number of active users.
+For example,~\cite{portscanhll} proposes its use for detecting network scanning attacks, while Facebook engineers have reported~\cite{fbhll} that they use HLL for estimating the number of active users.
 These applications demand a careful evaluation of the performance of HLL under adversarial input.
@@
 
-\kenny{M: Please fill in missing references above.}
-
-The rest of the paper is organized as follows.
+The rest of the paper is organised as follows.
 Section~\ref{sec:overview} gives an overview of the HyperLogLog algorithm.
@@
 
-\section{HyperLogLog}\label{sec:overview}
-\subsection{Algorithm}
-HLL is a streaming algorithm based on the bit pattern observation that for a stream of randomly distributed numbers, if we observe a number $x$ with the maximum of $k$ leading 0 bits, the cardinality of the stream is likely to be 2$^{k+1}$.
-In practice, to guaranty uniform distribution of the inputs, they go through a hash function $h$ before being processed.
-This simple approach has the disadvantage of a large variance.
-In order to improve the estimate, we can use many estimators (buckets) instead of one and average the results.
-To map a value $h(x)$ to a bucket, \cite{loglog} suggests using the first few $n$ bits of $h(x)$ as an index, and then compute the longest sequence of leading 0s on the $l = len(h(x)) - n $ bits left.
-Estimators from all buckets are averaged with an harmonic mean and scaled by a constant $\alpha$, empirically computed to correct a systematic multiplicative bias.
-As result, the HyperLogLog algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using 1.5 kB of memory.
+\section{Overview of HLL}\label{sec:overview}
+\subsection{The Algorithm}
+HLL~\cite{loglog,hll2} is a streaming algorithm based on the key observation that, for a stream of randomly distributed values represented as bit-strings of some fixed length, if we observe a value with the maximum of $k$ leading zero-bits, then the cardinality of the stream (i.e.\ the number of distinct values it contains) is likely to be on the order of $2^{k+1}$.
+In practice, to ensure a uniform distribution, the stream values are first processed by a hash function $h$ before the leading bits are inspected.
+So to estimate the cardinality of a stream, we need only store a representation of $k$, the length of the largest observed string of leading zero-bits for numbers in the stream.
+This provides a very compact mechanism for estimating the stream cardinality -- if the true cardinality is $N$, then just $\log\log(N)$ bits are needed to store the estimate.
 
-An interesting property of the HyperLogLog sketch is its ability to be merged in a lossless way.
-To combine the two buckets at index $i$, take the maximum estimator of the two and assign that to the matching bucket $i$ in the unioned HLL.
-Such a simple set union operation allows us to easily parallelize operations among multiple machines independently using the same hash function and the same number of buckets.
+This simple approach suffers from large variance in the estimation.
+In order to improve the estimate, we can use many estimators in parallel instead of one, and average the results.
+Each estimate is stored in its own bucket.
+To map a value $x$ to one of $2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$ (so that the output length of $h(\cdot)$ is $n+\ell$).
+Estimators from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ empirically computed to correct a systematic multiplicative bias.
+As result, HLL is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{add-ref-for-this-claim}.
+
+An interesting property of HLL is that it supports merging in a lossless way.
+To combine two buckets at index $i$ in two different HLL instances, take the maximum of the two bucket entries and assign that to the matching bucket $i$ in the merged HLL.
+Such a simple set union operation allows easy parallelisation operations among multiple machines independently, provided they use the same hash function and the same number of buckets.
 
 \subsection{Applications}
-For completeness and motivation, we provide examples of concrete uses of HyperLogLog:
+For completeness and motivation, we provide examples of concrete uses of HLL:
 \begin{itemize}
     \item Cardinality estimation provides a significantly faster way for Facebook to compute statistics on their users.
-They use HyperLogLog to find out how many distinct people visited their website in the past week \cite{fbhll}.
-    \item A solution to detecting DOS attacks using a sliding HyperLogLog \cite{slidinghll} (a variation which estimates the cardinality over a moving time window) is presented in \cite{portscanhll}.
-The sketch is used to count how many different ports are used as destination over all packets received, as an attempt to identify when a port scan attack is happening.
-    \item \cite{flexswitch} implements the HyperLogLog on a light FlexSwitch to approximate the number of distinct packets in traffic flows and overcome the switch resource limits.
+They use HLL to find out how many distinct people visited their website in the past week~\cite{fbhll}.
+    \item HLL has been proposed as a solution to detecting Denial-of-Service attacks~\cite{portscanhll} (this paper actually uses a variation on HLL called sliding HLL~\cite{slidinghll} which estimates the cardinality over a moving time window).
+Specifically, HLL is used to count how many different ports are used as destination port over all packets received, as an attempt to identify when a port scan attack is underway.
+        \item HLL has been implemented in network switches to approximate the number of distinct packets in traffic flows and overcome the switch resource limits~\cite{flexswitch}.
 The number of unique flows traversing a switch can then be used to provide a better congestion control protocol.
@@
 
-Regardless of the sensitivity of the applications, none of them are clear about their threat model, especially for the HyperLogLog.
-By not assessing the risks involved with the presence of an adversary targeting the HLL sketch, they may allow an attack like the one we present, breaking down the whole solution.
+Regardless of the sensitivity of the above examples, none of the relevant papers are clear about their threat model, especially for the HLL component.
+They are all potentially vulnerable to the attacks we present below.
+%By not assessing the risks involved with the presence of an adversary targeting the HLL sketch, they may allow an attack like the one we present, breaking down the whole solution.
 
 \subsection{Related Work}
-The work of \cite{cardestprivacy} breaks the privacy of the HLL sketch and concludes that having the ability to test the insertion of a targeted item gives significant confidence about whether the item was originally present in the input data set.
-HLL sketches can be used by organizations to store and process location data, which is inherently sensitive, so loosing privacy would leak one user's location and potentially cause great harm.
-On the other hand, \cite{hllvuln} exploits a vulnerability of the HLL algorithm to manipulate the cardinality estimate, leading to a five-fold reduction from the true cardinality.
 
-Both propose mitigations using a salted sketch, either as replacement, but at the cost of loosing mergeability, or in addition to an unsalted sketch, inducing a memory overhead.
-This might not be a convenient solution for some applications, they thus encourage further research.
+The work of~\cite{cardestprivacy} breaks the privacy of the HLL sketch and concludes that having the ability to test the insertion \kenny{what does this mean?} of a targeted item gives information about whether the item was originally present in the input data set.
+It is argued in~\cite{cardestprivacy} that HLL can be used by organizations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
+
+In~\cite{hllvuln}, Reviriego and Ting exploit a vulnerability of HLL to manipulate the cardinality estimate, leading to a five-fold reduction in the estimate compared to the true cardinality.
+
+Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted sketch, either as replacement (but at the cost of loosing mergability) or in addition to an unsalted sketch (inducing a memory overhead).
+%This might not be a convenient solution for some applications, they thus encourage further research.
 

### 2020-04-24 docs/HLL_exp_bad/main.tex
more work

@@
 
-\section{Vulnerabilities}\label{sec:attacks}
-This section presents different adversarial settings and the corresponding attacks to manipulate the HLL estimate.
-The attacker's goal in our work is to bring the estimate down, as this would cause the most damage when looking at the applications of HLL.
-One can think about the attacker trying the hide from a detection system.
-Conveniently, it is easy to adopt the same strategy as the one we present for an attacker who is looking to artificially inflate the HLL estimate given a limited data set input.
-The impact over the cardinality estimate and the computational effort to be done by the attacker is analyzed afterwards.
+\section{HLL in Adversarial Settings}\label{sec:attacks}
 
-\subsection{Adversarial setting}
-We model 3 adversarial scenarios based on the attacker's knowledge and capabilities.
-For all of them, we assume that the sketch already contains some existing data from honest users to simulate a real life setup.
-Among the scenarios, S3 has the strongest assumptions as it combines the knowledge of attackers from S1 and S2.
+This section presents different adversarial settings and corresponding attacks that manipulate the HLL cardinality estimate.
+The adversary's goal is to reduce the estimate as much as possible, since this would cause more damage when  considering the above-mentioned applications of HLL.
+It is easy to modify our attack strategies for an adversary who instead wishes to artificially inflate the HLL estimate given a limited input capability.
+We also analyze the impact on the HLL cardinality estimate and the computational effort required.
 
-S1: the attacker is able to test insertion of an item on a black-box snapshot of the sketch in use and check the effect on the cardinality estimate.
+Throughout, we assume the adversary has the ability to vary the contents of free fields in the input strings which will be inserted into the stream of values.
+For example it can manipulate IPID and other header fields in the context of IP packets.
+This enables the adversary to control the output of hash function $h$ and flexibly choose which items will be inserted into the HLL, and in which buckets.
 
-S2: the attacker has access to the details of the implementation, i.e., the number of buckets and the hash function $h$.
+\subsection{Adversarial Setting}
+We model three distinct adversarial scenarios based on the adversary's knowledge and capabilities.
+For all of them, we assume that the sketch already contains some existing data from honest users to simulate a real-life setting.
+Among the scenarios, S3 makes the strongest assumptions about the adversarial capabilities, since it combines the knowledge of adversarys from S1 and S2.
 
-S3: the attacker has direct access to the sketch in use and all internals, i.e., the number of buckets, the hash function $h$, and also the sketch's buckets content.
-Those buckets hold the intermediate estimators updated by the honest user's data, that are averaged in a later step to compute the global cardinality estimator.\\
+\begin{itemize}
+\item[S1:] the adversary has black-box access to the HLL in use -- that is, the adversary can ask for the HLL's cardinality estimate at any point in time, and so is able to check the effect on the cardinality estimate when inserting an item;
+\item[S2:] the adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use.
+\item[S3:] the adversary has direct access to the sketch in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL.
+Recall that these buckets hold the intermediate estimators that are averaged in a later step to compute the final cardinality estimate.
+\end{itemize}
 
-S1 and S2 are presented as M2 and M1 respectively in \cite{hllvuln}.
-We do not put our focus on S1 as this previous work fully exploit this scenario.
-The attack works as follow: ``after each insertion the cardinality is checked and the items which do not increment the HLL estimate are retained``.
-We start by presenting our attack in S2, providing much stronger results then the corresponding attack of \cite{hllvuln}, and then use the additional information the attacker has access to in S3 to make the attack even more efficient.
+Scenarios S1 and S2 are presented as M2 and M1, respectively, in~\cite{hllvuln}.
+We do not put our focus on S1 as this previous work in~\cite{hllvuln} fully exploits this scenario.
+Their attack works as follow: ``after each insertion the cardinality is checked and the items which do not increment the HLL estimate are retained''.
+It is unclear what this means in practice, since after insertion into the HLL, an item cannot usually be removed by the adversary.
+\kenny{Matilda, can you think about this sentence that I added?
+Do we need to rewrite S1?
+The scenario does not really make sense at all in practice.
+I think we might use this to beat up the other paper a bit more.}
 
-\subsection{Manipulating the estimate}
-Under S2: The attacker has access to $h$ so he can adjust some field of the item $x$ he is trying to insert and keep only the one such that the $l$ rightmost bits of $h(x)$ start with a 1.
-Indeed, the number of leading 0 bits computed by the estimator of the destination bucket is always zero, so the estimator of this bucket is unlikely to be updated and the averaged approximated cardinality is left unchanged.
-As the attacker knows how many buckets are used, he can recover $n$ and $l$, and aim for the $n$th bit (first bit not used for bucket allocation) to be 1.
-For example, in the scenario of \cite{portscanhll}, the attacker could change the source port of his IP packet $x$ until he is satisfied with $h(x)$.
-\textbf{(what about repeated sampling?)}\\
+We start by presenting our attack in scenario S2, providing much stronger results than the corresponding attack of \cite{hllvuln}, and then use the additional information the adversary has access to in scenario S3 to make the attack even more powerful.
 
-Under S3: The attacker can adopt the same strategy but make the attack more efficient and reduce the sampling requirements with the additional information.
-Now that the attacker has the sketch, not only he knows the overall estimated cardinality but also the intermediate estimators computed per bucket.
-So, he can "allow" some leading 0 bits in the $l$ rightmost bits of $h(x)$ as long as it does not increment the bucket's estimate.
-In other words, if $h(x)$ is mapped to the $i$th bucket with estimated cardinality $2^{c_i + 1}$ (it observed a value with $c_i$ leading 0s), then, instead of the strict restriction of a 1 bit at the $n$th index, the attacker is satisfied when having at least one 1 bit in the $n$ to $n+c_i + 1$ bit slice of $h(x)$.
-In case the attacker hits an empty bucket (the estimator is 0), he must skip that bucket completely.
-Under those conditions, the estimator of the buckets are not updated, and there is no effect on the cardinality estimation of the HLL sketch.
+\subsection{Manipulating the HLL Cardinality Estimate}
+
+\paragraph{Scenario S2:} The adversary knows $h$ so it can adjust fields in the item $x$ it is trying to insert and keep only those for which the $\ell$ rightmost bits of $h(x)$ start with a one bit.
+In this case, the number of leading zero bits computed by the estimator of the destination bucket is always zero, so the estimator of this bucket is unlikely to be updated and the averaged approximate cardinality is likely to be left unchanged.
+For example, in the scenario of \cite{portscanhll}, the adversary could change the source port of his IP packet until it is satisfied with its hash value.
+Given sufficient flexibility in the input $x$, the adversary can insert many items into the HLL without increasing the cardinality markedly, by repeatedly sampling from the input domain.
+
+\paragraph{Scenario S3:} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the intermediate estimators computed per bucket).
+So, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increment the current bucket estimate.
+In other words, if $h(x)$ is mapped to the $i$-th bucket with estimated cardinality $2^{c_i + 1}$ (because a value with $c_i$ leading zeroes was previously observed for this bucket), then, instead of the strict restriction of having a one bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a one bit in any of the first $c_i+1$ positions in this substring.
+In case the adversary hits an empty bucket (for which the estimator is 0), it must skip that bucket completely.
+Under these conditions, the estimators in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.
+A special case arises when the HLL is completely empty at the beginning of the attack; here the adversary is forced to increase the estimate in at least one bucket, but with sampling this can obviously be done in such a way as to increase the HLL estimate to $2$ in a single bucket, leaving all the other buckets untouched.
 
@@
 
-Under S2, the attacker need to craft 2 $x$s on average to obtain a 1 bit at the $n$th position of $h(x)$ and successfully perform the attack.
-As we do not care about the remaining bits, the attacker can theoretically "hide" up to 2$^{len(h(x))-1}$ items into the sketch.\\
+Under S2, the adversary need to craft 2 $x$s on average to obtain a 1 bit at the $n$th position of $h(x)$ and successfully perform the attack.
+As we do not care about the remaining bits, the adversary can theoretically "hide" up to 2$^{len(h(x))-1}$ items into the sketch.\\
 
-Under S3, let's compute the average work for the attacker to find an $x$ such that $h(x)$ meets the attack's requirements.
+Under S3, let's compute the average work for the adversary to find an $x$ such that $h(x)$ meets the attack's requirements.
 We assume $h(x)$ is mapped to bucket $i$.
@@
 $P(h(x)$ does NOT update bucket $i| h(x)$ mapped to bucket $i)=P(h(x)$ has LESS than $c_i$ leading 0s$)=1-P(h(x)$ has strictly MORE than $c_i$ leading 0s$)=1-\frac{1}{2^{c_i+1}}$\\
-Across all buckets, the attacker does not update any estimate with probability $P=
+Across all buckets, the adversary does not update any estimate with probability $P=
 \sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i$ AND $h(x)$ does NOT update bucket $i) = \sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i)P(h(x)$ does NOT update bucket $i| h(x)$ mapped to bucket $i) = \sum_{i=1}^{2^n}\frac{1}{2^n} \times (1-\frac{1}{2^{c_i+1}}) = 1-\frac{1}{2}\times(\frac{1}{2^n}\times \sum_{i=1}^{2^n} {(2^{c_i})}^{-1}) = 1-\frac{1}{2}\times\frac{1}{H_c} = 1-(2H_c)^{-1}$, with $H_c$ the harmonic mean of the numbers $2^{c_1}, 2^{c_2}, ..., 2^{c_{2^n}}$.
-So the attacker needs to craft on average $(1-(2H_c)^{-1})^{-1} x$s, and can theoretically "hide" up to 2$^{len(h(x))}\times (1-(2H_c)^{-1})$ items into the sketch.
+So the adversary needs to craft on average $(1-(2H_c)^{-1})^{-1} x$s, and can theoretically "hide" up to 2$^{len(h(x))}\times (1-(2H_c)^{-1})$ items into the sketch.
 
@@
 We assume here that there already are values coming from other honest users' data in the sketch, we now remove this assumption.
-Unfortunately, the attacker in S2 has no way to detect that he is attacking an empty sketch and cannot adapt his strategy, so all buckets hit will update once their estimator to $2^{0+1}=2$ items, leading to the sketch final approximation of the number of distinct elements to $\alpha2^{n+1}$ regardless of the amount of items the attacker will send.
-However, the attacker in S3 is able to notice that all buckets are empty and adapt his attack.
-As mentioned above, the attacker in S3 avoids empty buckets, but when dealing with an empty sketch, sacrifices must be made if the attacker wants to insert at least one item.
-To limit the detection, the attacker can target and fill one bucket only.
-This would fix $n+1$ bits of $h(x)$, allowing the attacker to add $2^{len(h(x))-n-1}$ elements while keeping the cardinality to 1.
+Unfortunately, the adversary in S2 has no way to detect that he is attacking an empty sketch and cannot adapt his strategy, so all buckets hit will update once their estimator to $2^{0+1}=2$ items, leading to the sketch final approximation of the number of distinct elements to $\alpha2^{n+1}$ regardless of the amount of items the adversary will send.
+However, the adversary in S3 is able to notice that all buckets are empty and adapt his attack.
+As mentioned above, the adversary in S3 avoids empty buckets, but when dealing with an empty sketch, sacrifices must be made if the adversary wants to insert at least one item.
+To limit the detection, the adversary can target and fill one bucket only.
+This would fix $n+1$ bits of $h(x)$, allowing the adversary to add $2^{len(h(x))-n-1}$ elements while keeping the cardinality to 1.
 
@@
 
-We challenge the attacker to pick the largest possible number of items to add to the sketch while trying to keep the cardinality estimate as low as possible, either in the S2 or the S3 scenario.
-In a second time, in order to get results that can be compared to the work of \cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from the universe of possible strings, the attacker has to choose from a set of 250'000 random items that are given to him.
+We challenge the adversary to pick the largest possible number of items to add to the sketch while trying to keep the cardinality estimate as low as possible, either in the S2 or the S3 scenario.
+In a second time, in order to get results that can be compared to the work of \cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose from a set of 250'000 random items that are given to him.
 Finally , for completeness, we also run the attacks against empty sketches.
@@
 \ref{table:tab1} and Tab.
-\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the attacker and the approximation of the cardinality after adding the attacker's items.
+\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary and the approximation of the cardinality after adding the adversary's items.
 
 Those empirical results confirm our impact analysis on several points.
-First, when attacking a sketch already containing some items, we can see that the attacker is able to hide approximately half of the items' set under S2, and around 83\%%( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$).
+First, when attacking a sketch already containing some items, we can see that the adversary is able to hide approximately half of the items' set under S2, and around 83\%%( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$).
 $\approx (1-\frac{1}{2\times H_c})$ with $H_c\approx\frac{1'000}{256}=3.9$, under S3.
@@
 \ref{table:tab2} with the attack presented in the Section 5.1 of \cite{hllvuln}, where they target the Redis \cite{redishll} open-source implementation of HLL and perform the attack against an empty sketch.
-Given a set of 250'000 items, the attacker could choose ``74'390 distinct items and obtain an HLL estimate of only 15'780`` achieving a five-fold reduction from the true cardinality.
-In comparison, the attacker following our strategy is able to add 125'000 distinct items while keeping the estimate at 367, demonstrating a reduction by a factor of 340.
+Given a set of 250'000 items, the adversary could choose ``74'390 distinct items and obtain an HLL estimate of only 15'780`` achieving a five-fold reduction from the true cardinality.
+In comparison, the adversary following our strategy is able to add 125'000 distinct items while keeping the estimate at 367, demonstrating a reduction by a factor of 340.
 

### 2020-04-24 docs/HLL_exp_bad/main.tex
typos

@@
 
-The work of~\cite{cardestprivacy} breaks the privacy of the HLL sketch and concludes that having the ability to test the insertion \kenny{what does this mean?} of a targeted item gives information about whether the item was originally present in the input data set.
+The work of~\cite{cardestprivacy} breaks the privacy of the HLL sketch and concludes that having the ability to test the insertion \kenny{what does this mean?
+see other comment later; the more I think about it, the more I think their paper is junk.} of a targeted item gives information about whether the item was originally present in the input data set.
 It is argued in~\cite{cardestprivacy} that HLL can be used by organizations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
@@
 It is unclear what this means in practice, since after insertion into the HLL, an item cannot usually be removed by the adversary.
-\kenny{Matilda, can you think about this sentence that I added?
+\kenny{Mathilde: can you think about this sentence that I added?
 Do we need to rewrite S1?

### 2020-04-24 docs/HLL_exp_bad/main.tex
attack impact

@@
 
-\paragraph{Scenario S2:} The adversary knows $h$ so it can adjust fields in the item $x$ it is trying to insert and keep only those for which the $\ell$ rightmost bits of $h(x)$ start with a one bit.
+\noindent\paragraph{Scenario S2:} The adversary knows $h$ so it can adjust fields in the item $x$ it is trying to insert and keep only those for which the $\ell$ rightmost bits of $h(x)$ start with a one bit.
 In this case, the number of leading zero bits computed by the estimator of the destination bucket is always zero, so the estimator of this bucket is unlikely to be updated and the averaged approximate cardinality is likely to be left unchanged.
@@
 
-\paragraph{Scenario S3:} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the intermediate estimators computed per bucket).
+\noindent\paragraph{Scenario S3:} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the intermediate estimators computed per bucket).
 So, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increment the current bucket estimate.
@@
 \subsection{Impact}
-Now that we have a way to exploit the vulnerability, we need to evaluate to which extent it can affect the estimate.
-We assume that $h$ is a "good" hash function, so any bit of $h(x)$ has a probability of $\frac{1}{2}$ to be 1 and $\frac{1}{2}$ to be 0.\\
+We now evaluate the impact of the above attacks on the HLL cardinality estimate.
+We assume that $h$ is a ``good" hash function, so that the bits of $h(x)$ are balanced (i.e.
+have a probability of $1/2$ of equalling 0 and 1).
 
-Under S2, the adversary need to craft 2 $x$s on average to obtain a 1 bit at the $n$th position of $h(x)$ and successfully perform the attack.
-As we do not care about the remaining bits, the adversary can theoretically "hide" up to 2$^{len(h(x))-1}$ items into the sketch.\\
+In scenario S2, the adversary needs to craft two distinct candidates for $x$ on average to obtain a one bit at the first position in the $\ell$ rightmost bits of $h(x)$, thus successfully performing the attack for a single insertion.
+Assuming the adversary has $t$ bits of flexibility in its choice of inputs, the adversary can expect to ``hide''  2$^{t-1}$ distinct items without increasing the HLL cardinality estimate at all.
 
-Under S3, let's compute the average work for the adversary to find an $x$ such that $h(x)$ meets the attack's requirements.
-We assume $h(x)$ is mapped to bucket $i$.
-The probability that $h(x)$ does not increment the estimator is the probability that $h(x)$ has less than $c_i$ leading 0s.
-So:\\
-$P(h(x)$ does NOT update bucket $i| h(x)$ mapped to bucket $i)=P(h(x)$ has LESS than $c_i$ leading 0s$)=1-P(h(x)$ has strictly MORE than $c_i$ leading 0s$)=1-\frac{1}{2^{c_i+1}}$\\
-Across all buckets, the adversary does not update any estimate with probability $P=
-\sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i$ AND $h(x)$ does NOT update bucket $i) = \sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i)P(h(x)$ does NOT update bucket $i| h(x)$ mapped to bucket $i) = \sum_{i=1}^{2^n}\frac{1}{2^n} \times (1-\frac{1}{2^{c_i+1}}) = 1-\frac{1}{2}\times(\frac{1}{2^n}\times \sum_{i=1}^{2^n} {(2^{c_i})}^{-1}) = 1-\frac{1}{2}\times\frac{1}{H_c} = 1-(2H_c)^{-1}$, with $H_c$ the harmonic mean of the numbers $2^{c_1}, 2^{c_2}, ..., 2^{c_{2^n}}$.
-So the adversary needs to craft on average $(1-(2H_c)^{-1})^{-1} x$s, and can theoretically "hide" up to 2$^{len(h(x))}\times (1-(2H_c)^{-1})$ items into the sketch.
+In scenario S3, we compute the average work required of the adversary to find an input $x$ such that $h(x)$ meets the attack's requirements.
+Let us assume for the moment that $c_i \ge 1$ for all buckets $i$.
+Assume $h(x)$ is mapped to bucket $i$.
+The probability that $h(x)$ does not increment the estimator is the probability that $h(x)$'s rightmost $\ell$ bits have $c_i$ or less leading zeroes.
+This is equal to $1-\frac{1}{2^{c_i+1}}$ assuming $h$ behaves like a random function.
+Averaging over all $2^n$ buckets (and assuming the bucket choice is uniformly random), the probability that input $x$ does not increment the estimator is given by:
+\begin{eqnarray*}
+\sum_{i=1}^{2^n}\frac{1}{2^n} \cdot (1-\frac{1}{2^{c_i+1}}) & = & 1-\frac{1}{2}\cdot (\frac{1}{2^n}\cdot \sum_{i=1}^{2^n} {(2^{c_i})}^{-1}) \\
+& = &1-(2H_c)^{-1} \\
+\end{eqnarray*}
+where $H_c$ is the harmonic mean of the counts $2^{c_1}, 2^{c_2}, \ldots, 2^{c_{2^n}}$.
 
-\textit{NOTE: does it take into account the special case of empty bucket ?
+%$P(h(x)$ does NOT update bucket $i| h(x)$ mapped to bucket $i)=P(h(x)$ has LESS than $c_i$ leading 0s$)=1-P(h(x)$ has strictly MORE than $c_i$ leading 0s$)=1-\frac{1}{2^{c_i+1}}$\\
+%Across all buckets, the adversary does not update any estimate with probability $P=
+%\sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i$ AND $h(x)$ does NOT update bucket $i) = \sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i)P(h(x)$ does NOT update bucket $i| h(x)$ mapped to bucket $i) = \sum_{i=1}^{2^n}\frac{1}{2^n} \times (1-\frac{1}{2^{c_i+1}}) = 1-\frac{1}{2}\times(\frac{1}{2^n}\times \sum_{i=1}^{2^n} {(2^{c_i})}^{-1}) = 1-\frac{1}{2}\times\frac{1}{H_c} = 1-(2H_c)^{-1}$, with $H_c$ the harmonic mean of the numbers $2^{c_1}, 2^{c_2}, ..., 2^{c_{2^n}}$.
+So the adversary needs to craft on average $(1-(2H_c)^{-1})^{-1}$ distinct candidates for $x$ in order to insert a single one, and, given $t$ bits of flexibility in its choice of inputs, can expect to ``hide" up to $(1-(2H_c)^{-1}) \cdot 2^t$ distinct items without increasing the HLL cardinality estimate at all.
+
+%\textit{NOTE: does it take into account the special case of empty bucket ?
 counters are initialized to $c_i=-oo$ (or -1 in the code???) thus $P(h(x)$ has LESS than $-oo$ leading $0s) = 0$ but it is hard to add it to the probabilities while staying clear.}\\
@@
 
-We assume here that there already are values coming from other honest users' data in the sketch, we now remove this assumption.
-Unfortunately, the adversary in S2 has no way to detect that he is attacking an empty sketch and cannot adapt his strategy, so all buckets hit will update once their estimator to $2^{0+1}=2$ items, leading to the sketch final approximation of the number of distinct elements to $\alpha2^{n+1}$ regardless of the amount of items the adversary will send.
-However, the adversary in S3 is able to notice that all buckets are empty and adapt his attack.
-As mentioned above, the adversary in S3 avoids empty buckets, but when dealing with an empty sketch, sacrifices must be made if the adversary wants to insert at least one item.
-To limit the detection, the adversary can target and fill one bucket only.
-This would fix $n+1$ bits of $h(x)$, allowing the adversary to add $2^{len(h(x))-n-1}$ elements while keeping the cardinality to 1.
+Note that the above analysis made the assumption that every bucket count $c_i$ was already at least 1.
+When some bucket counts are zero, the above analysis needs to be modified in order to avoid touching those buckets altogether.
+The modified analysis is quite straightforward.
+Suppose $r \le 2^n-1$ out of $2^n$ buckets have a count of zero.
+Then the cost of crafting an input $x$ becomes $(1-(2H_c)^{-1})^{-1} \cdot (1- \frac{r}{2^n})^{-1}$ trials, while the total number of items the adversary can expect to insert becomes $(1-(2H_c)^{-1}) \cdot (1- \frac{r}{2^n})\cdot 2^t$.
 
-\subsection{Experimental results}
+We assumed so far that there are already values coming from other honest users' data in the HLL.
+We now remove this assumption.
+Unfortunately, the adversary in scenario S2 has no way to detect that he is attacking an empty HLL and cannot adapt his strategy, so all buckets hit will update their estimator once to $2^{0+1}=2$ items, leading to a final cardinality approximation of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts.
+However, the adversary in S3 is able to detect that all buckets are empty and adapt his attack accordingly.
+The adversary can target and fill one bucket only.
+This requires fixing $n+1$ bits of $h(x)$ for each input $x$, allowing the adversary to ``hide'' $2^{t-n-1}$ elements while keeping the HLL cardinality estimate to 1.
+Other work/cardinality estimate trade-offs are of course possible.
+
+\subsection{Experimental Results}
 There exist many implementations of the HyperLogLog algorithm available online.

### 2020-04-24 docs/HLL_exp_bad/main.tex
tweaks

@@
 This prompts an examination of the performance of the HLL algorithm in the face of adversarial inputs.
-We show that in such a setting, the HLL algorithm's estimate of cardinality can be exponentially bad: when an adversary has access to the internals of the HLL algorithm, an adversary who has some flexibility in choosing what inputs will be recorded can manipulate the cardinality estimate to be exponentially smaller than the true cardinality.
+In this cautionary note, we show that in such a setting, the HLL algorithm's estimate of cardinality can be exponentially bad: when an adversary has access to the internals of the HLL algorithm, an adversary who has some flexibility in choosing what inputs will be recorded can manipulate the cardinality estimate to be exponentially smaller than the true cardinality.
 \end{abstract}
@@
 Each estimate is stored in its own bucket.
+The collection of buckets is referred to as the HLL \emph{sketch}.
 To map a value $x$ to one of $2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$ (so that the output length of $h(\cdot)$ is $n+\ell$).
@@
 An interesting property of HLL is that it supports merging in a lossless way.
-To combine two buckets at index $i$ in two different HLL instances, take the maximum of the two bucket entries and assign that to the matching bucket $i$ in the merged HLL.
-Such a simple set union operation allows easy parallelisation operations among multiple machines independently, provided they use the same hash function and the same number of buckets.
+To combine two buckets at index $i$ in two different HLL instances, take the maximum of the two bucket entries and assign that to the matching bucket $i$ in the merged HLL sketch.
+Such a simple set union operation allows easy parallelisation of operations among multiple machines independently, provided they use the same hash function and the same number of buckets.
 
@@
 They use HLL to find out how many distinct people visited their website in the past week~\cite{fbhll}.
-    \item HLL has been proposed as a solution to detecting Denial-of-Service attacks~\cite{portscanhll} (this paper actually uses a variation on HLL called sliding HLL~\cite{slidinghll} which estimates the cardinality over a moving time window).
+    \item HLL has been proposed as a solution to detecting Denial-of-Service attacks~\cite{portscanhll} (this paper actually uses a variation on HLL called \emph{sliding HLL}~\cite{slidinghll} which estimates the cardinality over a moving time window).
 Specifically, HLL is used to count how many different ports are used as destination port over all packets received, as an attempt to identify when a port scan attack is underway.
@@
 
-The work of~\cite{cardestprivacy} breaks the privacy of the HLL sketch and concludes that having the ability to test the insertion \kenny{what does this mean?
-see other comment later; the more I think about it, the more I think their paper is junk.} of a targeted item gives information about whether the item was originally present in the input data set.
-It is argued in~\cite{cardestprivacy} that HLL can be used by organizations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
+The work of~\cite{cardestprivacy} breaks the privacy of HLL and concludes that having the ability to test the insertion \kenny{what does this mean?
+see other comment later} of a targeted item gives information about whether the item was originally present in the input data set.
+It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
 
@@
 
-Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted sketch, either as replacement (but at the cost of loosing mergability) or in addition to an unsalted sketch (inducing a memory overhead).
+Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted sketch, either as replacement (but at the cost of losing mergability) or in addition to an unsalted sketch (inducing a memory overhead).
 %This might not be a convenient solution for some applications, they thus encourage further research.
@@
 For example it can manipulate IPID and other header fields in the context of IP packets.
-This enables the adversary to control the output of hash function $h$ and flexibly choose which items will be inserted into the HLL, and in which buckets.
+This enables the adversary to control the output of hash function $h$ and flexibly choose which items will be inserted into the HLL sketch, and in which buckets.
 
@@
 \item[S2:] the adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use.
-\item[S3:] the adversary has direct access to the sketch in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL.
+\item[S3:] the adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch.
 Recall that these buckets hold the intermediate estimators that are averaged in a later step to compute the final cardinality estimate.
@@
 The scenario does not really make sense at all in practice.
-I think we might use this to beat up the other paper a bit more.}
+I think we might use this to beat up the other paper a bit more.
+Maybe it also applies to the ``privacy'' paper.}
 
@@
 
-\noindent\paragraph{Scenario S3:} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the intermediate estimators computed per bucket).
+\noindent\paragraph{Scenario S3:} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values per bucket in the HLL sketch).
 So, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increment the current bucket estimate.
@@
 We now evaluate the impact of the above attacks on the HLL cardinality estimate.
-We assume that $h$ is a ``good" hash function, so that the bits of $h(x)$ are balanced (i.e.
-have a probability of $1/2$ of equalling 0 and 1).
+We assume that $h$ is a ``good" hash function, so that each of the bits of $h(x)$ can be regarded as being independent and uniformly random.
 
@@
 Assume $h(x)$ is mapped to bucket $i$.
-The probability that $h(x)$ does not increment the estimator is the probability that $h(x)$'s rightmost $\ell$ bits have $c_i$ or less leading zeroes.
-This is equal to $1-\frac{1}{2^{c_i+1}}$ assuming $h$ behaves like a random function.
-Averaging over all $2^n$ buckets (and assuming the bucket choice is uniformly random), the probability that input $x$ does not increment the estimator is given by:
+The probability that $h(x)$ does not increment the estimator is the probability that $h(x)$'s rightmost $\ell$ bits have $c_i$ or less leading zeroes; this probability is equal to $1-\frac{1}{2^{c_i+1}}$ assuming $h$ behaves like a random function.
+Averaging over all $2^n$ buckets (and using the fact that the bucket choice is uniformly random since the bits of $h$ are independent and uniformly random), the probability that input $x$ does not increment the estimator is given by:
 \begin{eqnarray*}
@@
 %\sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i$ AND $h(x)$ does NOT update bucket $i) = \sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i)P(h(x)$ does NOT update bucket $i| h(x)$ mapped to bucket $i) = \sum_{i=1}^{2^n}\frac{1}{2^n} \times (1-\frac{1}{2^{c_i+1}}) = 1-\frac{1}{2}\times(\frac{1}{2^n}\times \sum_{i=1}^{2^n} {(2^{c_i})}^{-1}) = 1-\frac{1}{2}\times\frac{1}{H_c} = 1-(2H_c)^{-1}$, with $H_c$ the harmonic mean of the numbers $2^{c_1}, 2^{c_2}, ..., 2^{c_{2^n}}$.
-So the adversary needs to craft on average $(1-(2H_c)^{-1})^{-1}$ distinct candidates for $x$ in order to insert a single one, and, given $t$ bits of flexibility in its choice of inputs, can expect to ``hide" up to $(1-(2H_c)^{-1}) \cdot 2^t$ distinct items without increasing the HLL cardinality estimate at all.
+So the adversary needs to craft on average $(1-(2H_c)^{-1})^{-1}$ distinct candidates for $x$ in order to insert a single candidate, and, given $t$ bits of flexibility in its choice of inputs, can expect to ``hide" up to $(1-(2H_c)^{-1}) \cdot 2^t$ distinct items without increasing the HLL cardinality estimate at all.
 
@@
 Note that the above analysis made the assumption that every bucket count $c_i$ was already at least 1.
-When some bucket counts are zero, the above analysis needs to be modified in order to avoid touching those buckets altogether.
+When some bucket counts are 0, the above analysis needs to be modified in order to avoid touching those buckets altogether.
 The modified analysis is quite straightforward.

### 2020-04-25 docs/HLL_exp_bad/main.tex
tweaks

@@
 We do not put our focus on S1 as this previous work in~\cite{hllvuln} fully exploits this scenario.
-Their attack works as follow: ``after each insertion the cardinality is checked and the items which do not increment the HLL estimate are retained''.
+Their attack works as follows: ``after each insertion the cardinality is checked and the items which do not increment the HLL estimate are retained''.
 It is unclear what this means in practice, since after insertion into the HLL, an item cannot usually be removed by the adversary.
@@
 \noindent\paragraph{Scenario S3:} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values per bucket in the HLL sketch).
-So, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increment the current bucket estimate.
+So, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
 In other words, if $h(x)$ is mapped to the $i$-th bucket with estimated cardinality $2^{c_i + 1}$ (because a value with $c_i$ leading zeroes was previously observed for this bucket), then, instead of the strict restriction of having a one bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a one bit in any of the first $c_i+1$ positions in this substring.
@@
 Assume $h(x)$ is mapped to bucket $i$.
-The probability that $h(x)$ does not increment the estimator is the probability that $h(x)$'s rightmost $\ell$ bits have $c_i$ or less leading zeroes; this probability is equal to $1-\frac{1}{2^{c_i+1}}$ assuming $h$ behaves like a random function.
-Averaging over all $2^n$ buckets (and using the fact that the bucket choice is uniformly random since the bits of $h$ are independent and uniformly random), the probability that input $x$ does not increment the estimator is given by:
+The probability that $h(x)$ does not increase the estimator is the probability that the  rightmost $\ell$ bits of $h(x)$ have $c_i$ or less leading zeroes; this probability is equal to $1-\frac{1}{2^{c_i+1}}$ assuming $h$ behaves like a random function.
+Averaging over all $2^n$ buckets (and using the fact that the bucket choice is uniformly random since the bits of $h$ are independent and uniformly random), the probability that input $x$ does not increase the estimator is given by:
 \begin{eqnarray*}

### 2020-04-25 docs/HLL_exp_bad/main.tex
tweaks

@@
 Computing the count of distinct elements in massive data sets is a common task but naive approaches are memory-expensive.
-The HyperLogLog (HLL) algorithm (Durand and Flajolet, 2003) estimates a data stream's cardinality while using significantly less memory than a naive approach at the cost of some accuracy.
+The HyperLogLog (HLL) algorithm (Flajolet \emph{et al.}, 2007) estimates a data stream's cardinality while using significantly less memory than a naive approach at the cost of some accuracy.
 This trade-off makes the HLL algorithm very attractive for a wide range of applications such as database management and network monitoring, where an exact count may not be needed.
@@
 In this paper, we present a complete analysis of HLL in adversarial setting, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
-In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary can make HLL underestimate true cardinality by a factor this is exponential in the number of buckets used by the algorithm.
+In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary can make HLL underestimate true cardinality by a factor that is exponential in the number of buckets used by the algorithm.
 
@@
 
-The work of~\cite{cardestprivacy} breaks the privacy of HLL and concludes that having the ability to test the insertion \kenny{what does this mean?
-see other comment later} of a targeted item gives information about whether the item was originally present in the input data set.
+The work of~\cite{cardestprivacy} studies the privacy properties of HLL in two different attack scenarios, referred to as \emph{insider} and \emph{external}.
+The attack target of~\cite{cardestprivacy} is different from ours, being concerned with loss of privacy, whereas we care about cardinality estimation.
 It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
+In Section~\ref{sec:attacks} we will discuss the two attack scenarios from~\cite{cardestprivacy} in more detail.
+In particular, we will show that an attacker who can insert items into the HLL and who can access the HLL cardinality estimate (via typical APIs in HLL implementations) as permitted in the external scenario can actually realise the same attacks as the insider attacker can, at least insofar as cardinality estimate manipulation is concerned.
 
 In~\cite{hllvuln}, Reviriego and Ting exploit a vulnerability of HLL to manipulate the cardinality estimate, leading to a five-fold reduction in the estimate compared to the true cardinality.
+They use an attack model that we consider to be quite artificial, since it gives the attacker the ability to test whether inserting an item \emph{would} increase the cardinality estimate, without actually having to insert the item.
+See further discussion in Section~\ref{sec:attacks}.
 
-Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted sketch, either as replacement (but at the cost of losing mergability) or in addition to an unsalted sketch (inducing a memory overhead).
+Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted sketch, either as replacement (but at the cost of losing mergeability) or in addition to an unsalted sketch (inducing a memory overhead).
 %This might not be a convenient solution for some applications, they thus encourage further research.
@@
 For all of them, we assume that the sketch already contains some existing data from honest users to simulate a real-life setting.
-Among the scenarios, S3 makes the strongest assumptions about the adversarial capabilities, since it combines the knowledge of adversarys from S1 and S2.
+Among the scenarios, S4 makes the strongest assumptions about the adversarial capabilities, since it combines the knowledge of adversaries from S1 and S2.
+\kenny{Not really!
+Because S1 is actually quite unrealistic, as we will argue below.}
 
 \begin{itemize}
-\item[S1:] the adversary has black-box access to the HLL in use -- that is, the adversary can ask for the HLL's cardinality estimate at any point in time, and so is able to check the effect on the cardinality estimate when inserting an item;
-\item[S2:] the adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use.
-\item[S3:] the adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch.
+\item[S1:] The adversary has black-box access to the HLL in use -- that is, the adversary can ask for the HLL's cardinality estimate at any point in time, and so is able to check the effect on the cardinality estimate when inserting an item.
+\kenny{TODO: clarify more what this is terms of a ''test oracle''; its stronger than described because the adverary can test the effect of an insertion without making the insertion!s}
+\item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''.
+In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack.
+\item[S3:] \kenny{TODO: The external attacker from~\cite{cardestprivacy}  described in terms of oracles.}
+\item[S4:] the adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch.
 Recall that these buckets hold the intermediate estimators that are averaged in a later step to compute the final cardinality estimate.
+Again, an S4 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack, but it can compute these values for itself (assuming its attack proceeds without input from other users being added to the HLL sketch in parallel with the attack).
 \end{itemize}
@@
 The scenario does not really make sense at all in practice.
-I think we might use this to beat up the other paper a bit more.
-Maybe it also applies to the ``privacy'' paper.}
+I think we might use this to beat up the other paper a bit more.}
 
-We start by presenting our attack in scenario S2, providing much stronger results than the corresponding attack of \cite{hllvuln}, and then use the additional information the adversary has access to in scenario S3 to make the attack even more powerful.
+\kenny{TODO: Show that S3 implies S4 $O(2^{n + \ell})$ precomputation and $O(\ell \cdot 2^{n})$ insert and read queries.
+So we.
+focus from now on S2 and S4 -- because S1 is unrealistic and S3 implies S4.}
+
+We start by presenting our attack in scenario S2, providing much stronger results than the corresponding attack of \cite{hllvuln}, and then use the additional information the adversary has access to in scenario S4 to make the attack even more powerful.
 
@@
 
-\noindent\paragraph{Scenario S3:} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values per bucket in the HLL sketch).
+\noindent\paragraph{Scenario S4:} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values per bucket in the HLL sketch).
 So, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
@@
 
-In scenario S3, we compute the average work required of the adversary to find an input $x$ such that $h(x)$ meets the attack's requirements.
+In scenario S4, we compute the average work required of the adversary to find an input $x$ such that $h(x)$ meets the attack's requirements.
 Let us assume for the moment that $c_i \ge 1$ for all buckets $i$.
@@
 Unfortunately, the adversary in scenario S2 has no way to detect that he is attacking an empty HLL and cannot adapt his strategy, so all buckets hit will update their estimator once to $2^{0+1}=2$ items, leading to a final cardinality approximation of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts.
-However, the adversary in S3 is able to detect that all buckets are empty and adapt his attack accordingly.
+However, the adversary in S4 is able to detect that all buckets are empty and adapt his attack accordingly.
 The adversary can target and fill one bucket only.
@@
 
-We challenge the adversary to pick the largest possible number of items to add to the sketch while trying to keep the cardinality estimate as low as possible, either in the S2 or the S3 scenario.
+We challenge the adversary to pick the largest possible number of items to add to the sketch while trying to keep the cardinality estimate as low as possible, either in the S2 or the S4 scenario.
 In a second time, in order to get results that can be compared to the work of \cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose from a set of 250'000 random items that are given to him.
@@
 First, when attacking a sketch already containing some items, we can see that the adversary is able to hide approximately half of the items' set under S2, and around 83\%%( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$).
-$\approx (1-\frac{1}{2\times H_c})$ with $H_c\approx\frac{1'000}{256}=3.9$, under S3.
+$\approx (1-\frac{1}{2\times H_c})$ with $H_c\approx\frac{1'000}{256}=3.9$, under S4.
 Secondly, when attacking an empty sketch, the final cardinality estimate is always exactly 367 $=\alpha2^{n+1}$ with $\alpha = 0.72$ as specified by the Eq.
-(29) of \cite{hll2} or 1 in respectively S2 and S3.
+(29) of \cite{hll2} or 1 in respectively S2 and S4.
 
@@
     \hline
-    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S3} \\ \hline
+    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
     \textbf{Initial data?} & yes & no & yes & no \\ \hline

### 2020-04-25 docs/HLL_exp_bad/main.tex
better attack analysis

@@
 This prompts an examination of the performance of the HLL algorithm in the face of adversarial inputs.
-In this cautionary note, we show that in such a setting, the HLL algorithm's estimate of cardinality can be exponentially bad: when an adversary has access to the internals of the HLL algorithm, an adversary who has some flexibility in choosing what inputs will be recorded can manipulate the cardinality estimate to be exponentially smaller than the true cardinality.
+In this cautionary note, we show that in such a setting, the HLL algorithm's estimate of cardinality can be exponentially bad: when an adversary has access to the internals of the HLL algorithm and has some flexibility in choosing what inputs will be recorded, it can manipulate the cardinality estimate to be exponentially smaller than the true cardinality.
 \end{abstract}
@@
 In scenario S2, the adversary needs to craft two distinct candidates for $x$ on average to obtain a one bit at the first position in the $\ell$ rightmost bits of $h(x)$, thus successfully performing the attack for a single insertion.
-Assuming the adversary has $t$ bits of flexibility in its choice of inputs, the adversary can expect to ``hide''  2$^{t-1}$ distinct items without increasing the HLL cardinality estimate at all.
+Assuming the adversary has $t$ bits of flexibility in its choice of inputs, the adversary can expect to ``hide''  2$^{t-1}$ distinct items without increasing the HLL cardinality estimate at all, under the assumption that all buckets in the HLL sketch already hold a value $c_i ge 1$.
+On the other hand, if some buckets are empty (but the adversary does not know which ones in the blind setting of S2), then we may expect a moderate increase in the HLL cardinality estimator.
+In the worst case, when all buckets are initially empty, the cardinality estimate will increase from 0 to at most $\alpha2^{n+1}$ (regardless of the number of items the adversary inserts).
 
@@
 
-Note that the above analysis made the assumption that every bucket count $c_i$ was already at least 1.
+Note that the above analysis for scenario S4 made the assumption that every bucket count $c_i$ was already at least 1.
 When some bucket counts are 0, the above analysis needs to be modified in order to avoid touching those buckets altogether.
@@
 Suppose $r \le 2^n-1$ out of $2^n$ buckets have a count of zero.
-Then the cost of crafting an input $x$ becomes $(1-(2H_c)^{-1})^{-1} \cdot (1- \frac{r}{2^n})^{-1}$ trials, while the total number of items the adversary can expect to insert becomes $(1-(2H_c)^{-1}) \cdot (1- \frac{r}{2^n})\cdot 2^t$.
+Then the cost of crafting an input $x$ becomes $(1-(2H_c)^{-1})^{-1} \cdot (1- \frac{r}{2^n})^{-1}$ trials, while the total number of items the adversary can expect to insert without increasing the HLL cardinality estimate becomes $(1-(2H_c)^{-1}) \cdot (1- \frac{r}{2^n})\cdot 2^t$.
+Of course, the adversary can relax its attack and allow some of the empty buckets to be hit.
+This increases the insertion rate at the cost of an increased final cardinality estimate.
 
@@
 We now remove this assumption.
-Unfortunately, the adversary in scenario S2 has no way to detect that he is attacking an empty HLL and cannot adapt his strategy, so all buckets hit will update their estimator once to $2^{0+1}=2$ items, leading to a final cardinality approximation of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts.
+The blind adversary in scenario S2 has no way to detect that he is attacking an empty HLL and cannot adapt his strategy, so all buckets that are hit will update their estimator once to $2^{0+1}=2$ items, leading to a final cardinality approximation of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts.
 However, the adversary in S4 is able to detect that all buckets are empty and adapt his attack accordingly.
-The adversary can target and fill one bucket only.
+The adversary can now target and fill one bucket only.
 This requires fixing $n+1$ bits of $h(x)$ for each input $x$, allowing the adversary to ``hide'' $2^{t-n-1}$ elements while keeping the HLL cardinality estimate to 1.
-Other work/cardinality estimate trade-offs are of course possible.
+Other work/cardinality-estimate trade-offs are of course possible.
+
+In both scenarios S2 and S4, we see that an adversary who has $t$ bits of flexibility in its inputs can expect to insert exponentially (in $t$) many inputs values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the cases where the HLL sketch has empty buckets in S2 or where the HLL sketch is empty to begin with in S2 and S4).
+The result is largely independent of the HLL parameters, but does depend on the extent to which the HLL is already filled, via a factor $1-(2H_c)^{-1}$ and the number of empty buckets $r$, via a factor $1- \frac{r}{2^n}$.
 
 \subsection{Experimental Results}
-There exist many implementations of the HyperLogLog algorithm available online.
-Among the repositories collecting more than 350 stars on github, we can cite \cite{clahll} and \cite{datasketch}.
-It is also featured in many frameworks, including Redis \cite{redis}, Google's BigQuery \cite{bigquery} and several products of Apache such as Spark \cite{spahll} \cite{spahll2} and Druid \cite{druhll}.
-It is interesting to point out that they all use a constant number of buckets and a fixed hash function, \textit{which is public or should be public according to Kerckhoffs's principle,} thus are potentially vulnerable to cardinality estimate manipulation when put in an adversarial setting.\\
+Many different implementations of HLL are available online.
+Among the repositories collecting more than 350 stars on GitHub, we can cite \cite{clahll} and \cite{datasketch}.
+HLL is also featured in many frameworks, including Redis \cite{redis}, Google's BigQuery \cite{bigquery} and several products of Apache such as Spark \cite{spahll,spahll2} and Druid \cite{druhll}.
+It is interesting to point out that all these implementations use a constant number of buckets and a fixed hash function.
+Thus they are vulnerable to our attacks in scenarios S2 and S4.
 
-As proof-of-concept, we implement our attack against the HLL algorithm from \cite{clahll}.
-The code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\textit{\url{https://github.com/PizzaWhisperer/HLLVuln}}}.
-Our items are 4 character long ASCII strings \textit{(7'311'616 distinct items)} and we choose $h$ to be the 32 bits Murmur3 hash function from \cite{murmur3code}.
-We create an HLL sketch using 256 buckets and initialize it with 1'000 random items representing an honest user's data.
+As a proof-of-concept, we implemented our attacks against the HLL implementation from \cite{clahll}.
+The attack code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\textit{\url{https://github.com/PizzaWhisperer/HLLVuln}}}.
+For simplicity, our items are 4-character long ASCII strings, giving us up to 7,311,616 distinct items that can be inserted.
+We choose $h$ to be the 32-bit Murmur3 hash function from \cite{murmur3code}.
+\kenny{Why do we get to choose $h$?
+Isn't it fixed by the specific implementation we are attacking?} We create an HLL sketch using $2^n = 256$ buckets and initialize it with 1,000 random items representing honest users' data.
 
-We challenge the adversary to pick the largest possible number of items to add to the sketch while trying to keep the cardinality estimate as low as possible, either in the S2 or the S4 scenario.
+We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of.
+while trying to keep the cardinality estimate as low as possible, either in the S2 or the S4 scenario.
 In a second time, in order to get results that can be compared to the work of \cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose from a set of 250'000 random items that are given to him.

### 2020-04-27 docs/HLL_exp_bad/main.tex
work on adv setting

@@
 Naive approaches to computing the cardinality of a big data set, such as sorting the elements or simply maintaining the set of unique elements seen, are impractical at scale.
-The HyperLogLog algorithm (HLL henceforth)~\cite{loglog,hll2} is today the most widely used cardinality estimator.
+The HyperLogLog algorithm (HLL henceforth)~\cite{hll} is today the most widely used cardinality estimator.
 It is increasingly used in settings where adversaries may have  incentives to manipulate the estimate made by the algorithm.
@@
 \subsection{The Algorithm}
-HLL~\cite{loglog,hll2} is a streaming algorithm based on the key observation that, for a stream of randomly distributed values represented as bit-strings of some fixed length, if we observe a value with the maximum of $k$ leading zero-bits, then the cardinality of the stream (i.e.\ the number of distinct values it contains) is likely to be on the order of $2^{k+1}$.
+HLL~\cite{hll} is a streaming algorithm based on the key observation that, for a stream of randomly distributed values represented as bit-strings of some fixed length, if we observe a value with the maximum of $k$ leading zero-bits, then the cardinality of the stream (i.e.\ the number of distinct values it contains) is likely to be on the order of $2^{k+1}$.
 In practice, to ensure a uniform distribution, the stream values are first processed by a hash function $h$ before the leading bits are inspected.
@@
 To map a value $x$ to one of $2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$ (so that the output length of $h(\cdot)$ is $n+\ell$).
-Estimators from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ empirically computed to correct a systematic multiplicative bias.
-As result, HLL is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{add-ref-for-this-claim}.
+Estimators from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ is empirically computed to correct a systematic multiplicative bias.
+As result, HLL is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
 
@@
 Specifically, HLL is used to count how many different ports are used as destination port over all packets received, as an attempt to identify when a port scan attack is underway.
-        \item HLL has been implemented in network switches to approximate the number of distinct packets in traffic flows and overcome the switch resource limits~\cite{flexswitch}.
+    \item HLL has been implemented in network soft-switches to approximate the number of distinct packets in traffic flows and overcome the switch resource limits~\cite{flexswitch}.
 The number of unique flows traversing a switch can then be used to provide a better congestion control protocol.
@@
 \subsection{Adversarial Setting}
-We model three distinct adversarial scenarios based on the adversary's knowledge and capabilities.
+We model four distinct adversarial scenarios based on the adversary's knowledge and capabilities.
 For all of them, we assume that the sketch already contains some existing data from honest users to simulate a real-life setting.
-Among the scenarios, S4 makes the strongest assumptions about the adversarial capabilities, since it combines the knowledge of adversaries from S1 and S2.
-\kenny{Not really!
-Because S1 is actually quite unrealistic, as we will argue below.}
+Among the scenarios, the \textit{insider} setup of \cite{cardestprivacy} that we present as S4 makes the strongest assumptions about the adversarial capabilities since we assume that the attacker owns the data structure and thus has an exhaustive view on the sketch.
+%combines the knowledge of adversaries from S1 and S2.
+\kenny{Not really! Because S1 is actually quite unrealistic, as we will argue below.}
 
 \begin{itemize}
-\item[S1:] The adversary has black-box access to the HLL in use -- that is, the adversary can ask for the HLL's cardinality estimate at any point in time, and so is able to check the effect on the cardinality estimate when inserting an item.
+\item[S1:] The adversary has access to a \textit{test oracle} provided by the sketch owner, which allows him to submit an item and learn the hypothetical effect of inserting it into the sketch.
+He is then able to determine if inserting this item \textit{would} increase the cardinality estimation.
+
+%The adversary is able to test the hypothetical effect of the insertion of an item on the cardinality estimation of the sketch.
+This can achieved via a \textit{test oracle} provided by the sketch owner, which allows an user to test the effect of an insertion on the cardinality estimate but without actually inserting the item, or if the attacker would gain access to a black-box replica of the sketch in use, i.e., buying the same switch and testing in the lab.
+%black-box access to the HLL in use -- that is, the adversary can ask for the HLL's cardinality estimate at any point in time, and so is able to check the effect on the cardinality estimate when inserting an item.
 \kenny{TODO: clarify more what this is terms of a ''test oracle''; its stronger than described because the adverary can test the effect of an insertion without making the insertion!s}
@@
 In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack.
-\item[S3:] \kenny{TODO: The external attacker from~\cite{cardestprivacy}  described in terms of oracles.}
-\item[S4:] the adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch.
+\mathilde{Would it make sense to add a sentence "This scenario would apply to~\cite{portscanhll}."?}
+\item[S3:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use.
+Additionally, he can access and interact with the sketch via an API provided by the sketch owner, allowing him to insert an item or ask for the HLL's cardinality estimate at any point in time.
+\kenny{TODO: The external attacker from~\cite{cardestprivacy}  described in terms of oracles.}
+\item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch.
 Recall that these buckets hold the intermediate estimators that are averaged in a later step to compute the final cardinality estimate.
 Again, an S4 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack, but it can compute these values for itself (assuming its attack proceeds without input from other users being added to the HLL sketch in parallel with the attack).
+\mathilde{I am not sure about the last sentence.
+In the \textit{insider} model of \cite{cardestprivacy} the attacker owns the sketch so has access to both statistics (cardinality estimation) and counters in real time.
+So if he receives data coming from another user / notices an update in the counters, he can check reiterate over the discarded items and check whether he can retain some more?
+I know the goal of \cite{cardestprivacy} is different and there it makes sense to consider the scenario where the owner is malicious but not necessarily for us.}
 \end{itemize}
 
-Scenarios S1 and S2 are presented as M2 and M1, respectively, in~\cite{hllvuln}.
-We do not put our focus on S1 as this previous work in~\cite{hllvuln} fully exploits this scenario.
-Their attack works as follows: ``after each insertion the cardinality is checked and the items which do not increment the HLL estimate are retained''.
-It is unclear what this means in practice, since after insertion into the HLL, an item cannot usually be removed by the adversary.
+Scenarios S1 and S2 are presented as M2 and M1, respectively, in~\cite{hllvuln} and S3 and S4 as the \textit{external} and \textit{internal} attack scenarios in \cite{cardestprivacy}.
+
+We start by highlighting the fact that S1 seems unrealistic in practice, as there is no reason why a sketch owner would provide such \textit{test oracle}.
+The authors of \cite{hllvuln} argue that the assumptions of S1 could also be fulfilled if the attacker gains access to a replica of the machine the HLL is implemented on.
+The adversary could run tests on it, sequentially inserting, asking for the cardinality estimation and resetting the device, retaining only the items which do not increment the HLL estimate.
+\mathilde{Todo: Add sentence This second approach does not look much more convincing/practical.}
+%This is unclear what this means in practice, since after insertion into the HLL sketch, an item cannot usually be removed by the adversary.
 \kenny{Mathilde: can you think about this sentence that I added?
@@
 
+Furthermore, we show that although \cite{cardestprivacy} presents S3 and S4 as detached scenarios, the attacker with the capabilities of S3 can obtain the knowledge of S4.
+By analyzing which inputs increase the cardinality estimation, an attacker can easily recover the contents of each bucket in the sketch.
+To guess the estimator of a targeted bucket, the attacker gradually inserts items with an incrementing number of leading zeros in the $\ell$ rightmost bits of $h(x)$, until the cardinality estimation of the sketch increases.
+Once a change occurs, the attacker knows that the maximum number of leading zeros observed in that bucket just got updated to a value that can be recovered by looking at the last inserted item.
+The attacker can determine all estimators with $O(2^{n + \ell})$ pre-computations and $O(\ell \cdot 2^{n})$ insert and cardinality estimation queries.
+The fuller the sketch the most work the attacker has to provide.
+
 \kenny{TODO: Show that S3 implies S4 $O(2^{n + \ell})$ precomputation and $O(\ell \cdot 2^{n})$ insert and read queries.
@@
 
-We start by presenting our attack in scenario S2, providing much stronger results than the corresponding attack of \cite{hllvuln}, and then use the additional information the adversary has access to in scenario S4 to make the attack even more powerful.
+As the assumptions of S1 appear unrealistic and an attacker can escalate from the knowledge of S3 to the one of S4, we focus on the scenarios S2 and S4.
+We first present our attack in scenario S2, providing much stronger results than the corresponding attack of \cite{hllvuln}, and then use the additional information the adversary has access to in scenario S4 to make the attack even more powerful.
 
@@
 
-\noindent\paragraph{Scenario S2:} The adversary knows $h$ so it can adjust fields in the item $x$ it is trying to insert and keep only those for which the $\ell$ rightmost bits of $h(x)$ start with a one bit.
+\noindent\paragraph{Scenario S2} The adversary knows $h$ so it can adjust fields in the item $x$ it is trying to insert and keep only those for which the $\ell$ rightmost bits of $h(x)$ start with a one bit.
 In this case, the number of leading zero bits computed by the estimator of the destination bucket is always zero, so the estimator of this bucket is unlikely to be updated and the averaged approximate cardinality is likely to be left unchanged.
-For example, in the scenario of \cite{portscanhll}, the adversary could change the source port of his IP packet until it is satisfied with its hash value.
+For example, in the scenario of \cite{portscanhll}, the adversary could change the source port of its IP packet until it is satisfied with its hash value.
 Given sufficient flexibility in the input $x$, the adversary can insert many items into the HLL without increasing the cardinality markedly, by repeatedly sampling from the input domain.
 
-\noindent\paragraph{Scenario S4:} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values per bucket in the HLL sketch).
+\noindent\paragraph{Scenario S4} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values per bucket in the HLL sketch).
 So, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
@@
 We now remove this assumption.
-The blind adversary in scenario S2 has no way to detect that he is attacking an empty HLL and cannot adapt his strategy, so all buckets that are hit will update their estimator once to $2^{0+1}=2$ items, leading to a final cardinality approximation of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts.
-However, the adversary in S4 is able to detect that all buckets are empty and adapt his attack accordingly.
+The blind adversary in scenario S2 has no way to detect that it is attacking an empty HLL and cannot adapt its strategy, so all buckets that are hit will update their estimator once to $2^{0+1}=2$ items, leading to a final cardinality approximation of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts.
+However, the adversary in S4 is able to detect that all buckets are empty and adapt its attack accordingly.
 The adversary can now target and fill one bucket only.
@@
 \kenny{Why do we get to choose $h$?
-Isn't it fixed by the specific implementation we are attacking?} We create an HLL sketch using $2^n = 256$ buckets and initialize it with 1,000 random items representing honest users' data.
+Isn't it fixed by the specific implementation we are attacking?} \mathilde{The library has a place holder for a 32 bit hash function but lets the user choose it.
+This was one of my concerns and I was wondering if it would make sense to attack \cite{datasketch} instead (hardcoded $h$, but when we discussed it, you said that due to Kerckhoff's principle, this hash function should be public so it does not matter that much.} We create an HLL sketch using $2^n = 256$ buckets and initialize it with 1,000 random items representing honest users' data.
 
@@
 Secondly, when attacking an empty sketch, the final cardinality estimate is always exactly 367 $=\alpha2^{n+1}$ with $\alpha = 0.72$ as specified by the Eq.
-(29) of \cite{hll2} or 1 in respectively S2 and S4.
+(29) of \cite{hll} or 1 in respectively S2 and S4.
 
@@
 \ref{table:tab2} with the attack presented in the Section 5.1 of \cite{hllvuln}, where they target the Redis \cite{redishll} open-source implementation of HLL and perform the attack against an empty sketch.
-Given a set of 250'000 items, the adversary could choose ``74'390 distinct items and obtain an HLL estimate of only 15'780`` achieving a five-fold reduction from the true cardinality.
+Given a set of 250'000 items, the adversary could choose ``74'390 distinct items and obtain an HLL estimate of only 15'780``~\cite{hllvuln} achieving a five-fold reduction from the true cardinality.
 In comparison, the adversary following our strategy is able to add 125'000 distinct items while keeping the estimate at 367, demonstrating a reduction by a factor of 340.
@@
     \hline
-    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
+    \textbf{Adv. Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
     \textbf{Initial data?} & yes & no & yes & no \\ \hline

### 2020-04-27 docs/HLL_exp_bad/main.tex
rm 'he' for attacker

@@
 \item[S1:] The adversary has access to a \textit{test oracle} provided by the sketch owner, which allows him to submit an item and learn the hypothetical effect of inserting it into the sketch.
-He is then able to determine if inserting this item \textit{would} increase the cardinality estimation.
+It is then able to determine if inserting this item \textit{would} increase the cardinality estimation.
 
@@
 \item[S3:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use.
-Additionally, he can access and interact with the sketch via an API provided by the sketch owner, allowing him to insert an item or ask for the HLL's cardinality estimate at any point in time.
+Additionally, it can access and interact with the sketch via an API provided by the sketch owner, allowing him to insert an item or ask for the HLL's cardinality estimate at any point in time.
 \kenny{TODO: The external attacker from~\cite{cardestprivacy}  described in terms of oracles.}
@@
 In the \textit{insider} model of \cite{cardestprivacy} the attacker owns the sketch so has access to both statistics (cardinality estimation) and counters in real time.
-So if he receives data coming from another user / notices an update in the counters, he can check reiterate over the discarded items and check whether he can retain some more?
+So if it receives data coming from another user / notices an update in the counters, it can check reiterate over the discarded items and check whether it can retain some more?
 I know the goal of \cite{cardestprivacy} is different and there it makes sense to consider the scenario where the owner is malicious but not necessarily for us.}
@@
 The attacker can determine all estimators with $O(2^{n + \ell})$ pre-computations and $O(\ell \cdot 2^{n})$ insert and cardinality estimation queries.
-The fuller the sketch the most work the attacker has to provide.
+\mathilde{todo: add sentence for The fuller the sketch the most work the attacker has to provide.}
 

### 2020-04-27 docs/HLL_exp_bad/main.tex
todo cmd
complete draft

--- 2020-04-27 docs/HLL_exp_bad/main.tex
@@
 \usepackage{hyperref}
+\usepackage{xspace}
 
@@
 \usepackage[color=oxygenorange]{todonotes}
-\newcommand{\mathilde}[2][inline]{\todo[#1]{\textbf{Mathilde:} #2}\xspace}
-\newcommand{\kenny}[2][inline]{\todo[#1]{\textbf{Kenny:} #2}\xspace}
+\newcommand{\mathilde}[1]{\todo[inline]{\textbf{Mathilde:} #1}\xspace}
+\newcommand{\kenny}[1]{\todo[inline]{\textbf{Kenny:} #1}\xspace}
 
@@
 In this paper, we present a complete analysis of HLL in adversarial setting, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
-In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary can make HLL underestimate true cardinality by a factor that is exponential in the number of buckets used by the algorithm.
+In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of $O(2^t)$ where $t$ is th number of bits of flexibility in the input).
 
@@
 For all of them, we assume that the sketch already contains some existing data from honest users to simulate a real-life setting.
-Among the scenarios, the \textit{insider} setup of \cite{cardestprivacy} that we present as S4 makes the strongest assumptions about the adversarial capabilities since we assume that the attacker owns the data structure and thus has an exhaustive view on the sketch.
+Among the scenarios, the \textit{insider} setup of \cite{cardestprivacy} that we present as S4 makes the strongest assumptions about the adversarial capabilities since it assumes that the attacker has a perfect view on the sketch.
 %combines the knowledge of adversaries from S1 and S2.
-\kenny{Not really! Because S1 is actually quite unrealistic, as we will argue below.}
+%\kenny{Not really! Because S1 is actually quite unrealistic, as we will argue below.}
 
 \begin{itemize}
-\item[S1:] The adversary has access to a \textit{test oracle} provided by the sketch owner, which allows him to submit an item and learn the hypothetical effect of inserting it into the sketch.
-He is then able to determine if inserting this item \textit{would} increase the cardinality estimation.
-
+\item[S1:] The adversary has access to a \textit{test oracle} provided by the sketch owner, which allows it to submit an item and learn the hypothetical effect of inserting it into the sketch.
+It is then able to determine if inserting this item \textit{would} increase the cardinality estimate, without actually inserting it.
 %The adversary is able to test the hypothetical effect of the insertion of an item on the cardinality estimation of the sketch.
@@
 %black-box access to the HLL in use -- that is, the adversary can ask for the HLL's cardinality estimate at any point in time, and so is able to check the effect on the cardinality estimate when inserting an item.
-\kenny{TODO: clarify more what this is terms of a ''test oracle''; its stronger than described because the adverary can test the effect of an insertion without making the insertion!s}
+%\kenny{TODO: clarify more what this is terms of a ''test oracle''; its stronger than described because the adversary can test the effect of an insertion without making the insertion!s}
 \item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''.
 In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack.
-\mathilde{Would it make sense to add a sentence "This scenario would apply to~\cite{portscanhll}."?}
+This kind of attack scenario is appropriate in the context of the port scanning attack of~\cite{portscanhll}.
 \item[S3:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use.
-Additionally, he can access and interact with the sketch via an API provided by the sketch owner, allowing him to insert an item or ask for the HLL's cardinality estimate at any point in time.
-\kenny{TODO: The external attacker from~\cite{cardestprivacy}  described in terms of oracles.}
-\item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch.
-Recall that these buckets hold the intermediate estimators that are averaged in a later step to compute the final cardinality estimate.
-Again, an S4 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack, but it can compute these values for itself (assuming its attack proceeds without input from other users being added to the HLL sketch in parallel with the attack).
-\mathilde{I am not sure about the last sentence.
-In the \textit{insider} model of \cite{cardestprivacy} the attacker owns the sketch so has access to both statistics (cardinality estimation) and counters in real time.
-So if he receives data coming from another user / notices an update in the counters, he can check reiterate over the discarded items and check whether he can retain some more?
-I know the goal of \cite{cardestprivacy} is different and there it makes sense to consider the scenario where the owner is malicious but not necessarily for us.}
+Additionally, it can access and interact with the sketch via an API provided by the sketch owner, allowing it to insert items into the HLL sketch and ask for the HLL's cardinality estimate at any point in time.
+\item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch.
+It then causes chosen inputs to be inserted into the HLL sketch.
+An S4 adversary can compute a lower bound on the HLL cardinality estimate based on its snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
 \end{itemize}
 
-Scenarios S1 and S2 are presented as M2 and M1, respectively, in~\cite{hllvuln} and S3 and S4 as the \textit{external} and \textit{internal} attack scenarios in \cite{cardestprivacy}.
+Scenarios S1 and S2 are presented as M2 and M1, respectively, in~\cite{hllvuln}.
+S3 is the \textit{external} attack scenario in~\cite{cardestprivacy}.
+S4 is a slightly weaker version of the \textit{insider} attack scenario in~\cite{cardestprivacy}, the difference being that an insider adversary in~\cite{cardestprivacy} has continuous access to the HLL sketch internals including all bucket values, while our adversary in S4 only has access to a snapshot of the HLL sketch internals at the start of its attack.
 
-We start by highlighting the fact that S1 seems unrealistic in practice, as there is no reason why a sketch owner would provide such \textit{test oracle}.
-The authors of \cite{hllvuln} argue that the assumptions of S1 could also be fulfilled if the attacker gains access to a replica of the machine the HLL is implemented on.
+We start by highlighting the fact that S1 seems unrealistic in practice, as there is no reason why a typical HLL implementation would provide such a test oracle.
+The authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker gains access to a replica of the machine on which the HLL is implemented.
 The adversary could run tests on it, sequentially inserting, asking for the cardinality estimation and resetting the device, retaining only the items which do not increment the HLL estimate.
-\mathilde{Todo: Add sentence This second approach does not look much more convincing/practical.}
-%This is unclear what this means in practice, since after insertion into the HLL sketch, an item cannot usually be removed by the adversary.
-\kenny{Mathilde: can you think about this sentence that I added?
-Do we need to rewrite S1?
-The scenario does not really make sense at all in practice.
-I think we might use this to beat up the other paper a bit more.}
+This may be possible under certain circumstances, but we do not consider it typical of HLL deployments.
 
-Furthermore, we show that although \cite{cardestprivacy} presents S3 and S4 as detached scenarios, the attacker with the capabilities of S3 can obtain the knowledge of S4.
-By analyzing which inputs increase the cardinality estimation, an attacker can easily recover the contents of each bucket in the sketch.
-To guess the estimator of a targeted bucket, the attacker gradually inserts items with an incrementing number of leading zeros in the $\ell$ rightmost bits of $h(x)$, until the cardinality estimation of the sketch increases.
-Once a change occurs, the attacker knows that the maximum number of leading zeros observed in that bucket just got updated to a value that can be recovered by looking at the last inserted item.
-The attacker can determine all estimators with $O(2^{n + \ell})$ pre-computations and $O(\ell \cdot 2^{n})$ insert and cardinality estimation queries.
-The fuller the sketch the most work the attacker has to provide.
+Furthermore, we show that although~\cite{cardestprivacy} presents S3 (external adversary) and the stronger version of our S4 (insider adversary) as two separate scenarios, an S3 adversary can easily infer the knowledge given to our S4 adversary in its snapshot.
+Indeed, an adversary can easily recover the contents of each bucket in the sketch simply by analysing which inputs increase the cardinality estimate.
+To find the estimate held in a targeted bucket, the attacker gradually inserts items $x$ with an incrementing number of leading zeros in the $\ell$ rightmost bits of $h(x)$ whilst holding the $n$ bits determining the bucket index constant, until the cardinality estimate of the sketch increases.
+Once a change occurs, the attacker knows that the maximum number of leading zeros observed in that bucket was just updated to a value that can be recovered by looking at the last inserted item.
+This assumes the attacker has sufficient flexibility in its choice of inputs $x$ so that the relevant bit conditions on $h(x)$ can be forced.
+The attacker can determine all bucket contents with, in the worst case, $O(2^{n + \ell})$ pre-computation and $O(\ell \cdot 2^{n})$ insert and cardinality estimate queries, and needs about $n + \ell$ bits of flexibility assuming $h$ behaves like a random function.
+The pre-computation and query complexity required can be much less if the HLL sketch is only moderately loaded.
 
-\kenny{TODO: Show that S3 implies S4 $O(2^{n + \ell})$ precomputation and $O(\ell \cdot 2^{n})$ insert and read queries.
-So we.
-focus from now on S2 and S4 -- because S1 is unrealistic and S3 implies S4.}
+A version of this attack can be carried out to elevate an external adversary to a full-strength insider adversary as per~\cite{cardestprivacy} (instead of our slightly weaker S4 adversary), since it can simply be done repeatedly each time the external adversary wants to know the complete internals of the HLL sketch.
+Obvious optimisations can be carried out to reduce the query complexity each time the attack is run (since the bucket contents can only increase over time, some queries are redundant; moreover, the pre-computation is only needed once and is moderate given typical HLL parameters, e.g.\ $n+\ell = 32$).
 
-As the assumptions of S1 appear unrealistic and an attacker can escalate from the knowledge of S3 to the one of S4, we focus on the scenarios S2 and S4.
-We first present our attack in scenario S2, providing much stronger results than the corresponding attack of \cite{hllvuln}, and then use the additional information the adversary has access to in scenario S4 to make the attack even more powerful.
+Since the assumptions of S1 appear to us to be unrealistic and since an attacker can escalate from S3 to S4, we focus on the scenarios S2 and S4 in what follows.
+We first present an attack in scenario S2, providing much stronger results than the corresponding attack of \cite{hllvuln} (in their equivalent scenario M1).
+We then use the additional information the adversary has access to in scenario S4 to make the attack even more powerful.
 
@@
 In scenario S2, the adversary needs to craft two distinct candidates for $x$ on average to obtain a one bit at the first position in the $\ell$ rightmost bits of $h(x)$, thus successfully performing the attack for a single insertion.
-Assuming the adversary has $t$ bits of flexibility in its choice of inputs, the adversary can expect to ``hide''  2$^{t-1}$ distinct items without increasing the HLL cardinality estimate at all, under the assumption that all buckets in the HLL sketch already hold a value $c_i ge 1$.
+Assuming the adversary has $t$ bits of flexibility in its choice of inputs, the adversary can expect to ``hide''  $2^{t-1}$ distinct items without increasing the HLL cardinality estimate at all, under the assumption that all buckets in the HLL sketch already hold a value $c_i \ge 1$.
 On the other hand, if some buckets are empty (but the adversary does not know which ones in the blind setting of S2), then we may expect a moderate increase in the HLL cardinality estimator.
 In the worst case, when all buckets are initially empty, the cardinality estimate will increase from 0 to at most $\alpha2^{n+1}$ (regardless of the number of items the adversary inserts).
+Notice that in all cases, provided $t$ is large enough, the adversary can keep the cardinality estimate linear in the total number of buckets ($2^n$) while the expected value of the true cardinality, $2^{t-1}$, can be made as large as the adversary pleases.
 
@@
 %\textit{NOTE: does it take into account the special case of empty bucket ?
-counters are initialized to $c_i=-oo$ (or -1 in the code???) thus $P(h(x)$ has LESS than $-oo$ leading $0s) = 0$ but it is hard to add it to the probabilities while staying clear.}\\
+counters are initialised to $c_i=-oo$ (or -1 in the code???) thus $P(h(x)$ has LESS than $-oo$ leading $0s) = 0$ but it is hard to add it to the probabilities while staying clear.}\\
 %2$^{len(h(x))}\times \frac{2H_c-1}{H_c}$ items into the sketch.\\
@@
 Of course, the adversary can relax its attack and allow some of the empty buckets to be hit.
-This increases the insertion rate at the cost of an increased final cardinality estimate.
+This increases the insertion rate at the cost of a small increase in the final cardinality estimate.
 
@@
 We now remove this assumption.
-The blind adversary in scenario S2 has no way to detect that it is attacking an empty HLL and cannot adapt its strategy, so all buckets that are hit will update their estimator once to $2^{0+1}=2$ items, leading to a final cardinality approximation of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts.
+The blind adversary in scenario S2 has no way to detect that it is attacking an empty HLL and cannot adapt its strategy, so all buckets that are hit will update their estimator once to $2^{0+1}=2$ items, leading to a final cardinality estimate of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts.
 However, the adversary in S4 is able to detect that all buckets are empty and adapt its attack accordingly.
@@
 This requires fixing $n+1$ bits of $h(x)$ for each input $x$, allowing the adversary to ``hide'' $2^{t-n-1}$ elements while keeping the HLL cardinality estimate to 1.
-Other work/cardinality-estimate trade-offs are of course possible.
+Other trade-offs between work and the cardinality estimate are of course possible.
 
 In both scenarios S2 and S4, we see that an adversary who has $t$ bits of flexibility in its inputs can expect to insert exponentially (in $t$) many inputs values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the cases where the HLL sketch has empty buckets in S2 or where the HLL sketch is empty to begin with in S2 and S4).
-The result is largely independent of the HLL parameters, but does depend on the extent to which the HLL is already filled, via a factor $1-(2H_c)^{-1}$ and the number of empty buckets $r$, via a factor $1- \frac{r}{2^n}$.
+The result is largely independent of the HLL parameters, but does depend on the extent to which the HLL sketch is already filled, via a factor $1-(2H_c)^{-1}$ and the number of empty buckets $r$, via a factor $1- \frac{r}{2^n}$.
 
@@
 For simplicity, our items are 4-character long ASCII strings, giving us up to 7,311,616 distinct items that can be inserted.
-We choose $h$ to be the 32-bit Murmur3 hash function from \cite{murmur3code}.
-\kenny{Why do we get to choose $h$?
+We choose $h$ to be the 32-bit Murmur3 hash function from \cite{murmur3code} (note that the library allows the hash function to be specified; we pick Murmur3  for our attack and consider it reasonable that the adversary should know the hash function in use).
+% \kenny{Why do we get to choose $h$?
 Isn't it fixed by the specific implementation we are attacking?} \mathilde{The library has a place holder for a 32 bit hash function but lets the user choose it.
-This was one of my concerns and I was wondering if it would make sense to attack \cite{datasketch} instead (hardcoded $h$, but when we discussed it, you said that due to Kerckhoff's principle, this hash function should be public so it does not matter that much.} We create an HLL sketch using $2^n = 256$ buckets and initialize it with 1,000 random items representing honest users' data.
+This was one of my concerns and I was wondering if it would make sense to attack \cite{datasketch} instead (hardcoded $h$, but when we discussed it, you said that due to Kerckhoff's principle, this hash function should be public so it does not matter that much.}
+We create an HLL sketch using $2^n = 256$ buckets and initialise it with 1,000 random items representing honest users' data.
 
-We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of.
-while trying to keep the cardinality estimate as low as possible, either in the S2 or the S4 scenario.
-In a second time, in order to get results that can be compared to the work of \cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose from a set of 250'000 random items that are given to him.
-Finally , for completeness, we also run the attacks against empty sketches.
-All attacks are ran 30 times and the outcomes are reported in Tab.
-\ref{table:tab1} and Tab.
-\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary and the approximation of the cardinality after adding the adversary's items.
+We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  7,311,616 possible inputs while trying to keep the cardinality estimate as low as possible, either in the S2 or the S4 scenario.
+As a second experiment, and in order to get results that can be directly compared to the work of~\cite{hllvuln}, we mounted our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose what to insert from a set of 250,000 random items that are given to him.
+Finally, for completeness, we also run the attacks against HLL sketches that are initially empty.
+All attacks were run 30 times and the outcomes averaged.
+Our results are reported in Tables~\ref{table:tab1} and~\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary, and the cardinality estimate after adding the adversary's items.
 
-Those empirical results confirm our impact analysis on several points.
-First, when attacking a sketch already containing some items, we can see that the adversary is able to hide approximately half of the items' set under S2, and around 83\%%( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$).
-$\approx (1-\frac{1}{2\times H_c})$ with $H_c\approx\frac{1'000}{256}=3.9$, under S4.
-Secondly, when attacking an empty sketch, the final cardinality estimate is always exactly 367 $=\alpha2^{n+1}$ with $\alpha = 0.72$ as specified by the Eq.
-(29) of \cite{hll} or 1 in respectively S2 and S4.
+These empirical results confirm our impact analysis on several points.
+First, when attacking a sketch already containing some items, we can see that the adversary is able to hide approximately half of the items in our starting set under S2, and around 83\% under S4.
+%( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$)
+Here $0.83 \approx (1-\frac{1}{2\cdot H_c})$ with $H_c\approx\frac{1000}{256}=3.9$.
+Secondly, when attacking an empty sketch, the final cardinality estimate is always exactly 367 $=\alpha2^{n+1}$ with $\alpha = 0.72$ as specified by equation (29) in~\cite{hll}, or 1, in respectively S2, S4.
 
-To assess the efficiency of our attacks, we can confront our results from Tab.
-\ref{table:tab2} with the attack presented in the Section 5.1 of \cite{hllvuln}, where they target the Redis \cite{redishll} open-source implementation of HLL and perform the attack against an empty sketch.
-Given a set of 250'000 items, the adversary could choose ``74'390 distinct items and obtain an HLL estimate of only 15'780``~\cite{hllvuln} achieving a five-fold reduction from the true cardinality.
-In comparison, the adversary following our strategy is able to add 125'000 distinct items while keeping the estimate at 367, demonstrating a reduction by a factor of 340.
+To assess the efficiency of our attacks, we can compare our results from Table~\ref{table:tab2} with the attack presented in the Section 5.1 of \cite{hllvuln}.
+In that paper, the authors target the Redis~\cite{redishll} open-source implementation of HLL and perform the attack against an empty sketch.
+Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780``~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality.
+In comparison, the adversary following our strategy is able to add 125,000 distinct items while keeping the cardinality estimate at 367, providing a reduction by a factor of 340 instead of 5.
 
 \begin{table}[h]
-\caption{Result of the attacks over 30 iterations}
+\caption{Attack results, averaged over 30 iterations.}
 \begin{tabular}{| m{8.5em} | m{4em} | m{4em} | m{4em} | m{4em} |}
     \hline
-    \textbf{Adv. Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
+    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
     \textbf{Initial data?} & yes & no & yes & no \\ \hline
-    \textbf{Original Est. Card.} & 1'000 & 0 & 1'005 & 0 \\ \hline
-    \textbf{\# Items added} & 3'655'744 & 3'655'740 & 6'090'003 & 14'220 \\ \hline
-    \textbf{Final Est. Card.} & 1'011 & 367 & 1'005 & 1 \\ \hline
+    \textbf{Original Card. Est.} & 1,000 & 0 & 1,005 & 0 \\ \hline
+    \textbf{\# Items added} & 3,655,744 & 3,655,740 & 6,090,003 & 14,220 \\ \hline
+    \textbf{Final Card. Est.} & 1'011 & 367 & 1,005 & 1 \\ \hline
 \end{tabular}
@@
 \begin{table}[h]
-\caption{Result of the attacks in the setting of \cite{hllvuln} over 30 iterations}
+\caption{Attack results, averaged over 30 iterations, in the setting of \cite{hllvuln}.}
 \begin{tabular}{| m{8.5em} | m{4em} | m{4em} | m{4em} | m{4em} |}
     \hline
-    \textbf{Adv. Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
+    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
     \textbf{Initial data?} & yes & no & yes & no \\ \hline
-    \textbf{Original Est. Card.} & 1'004 & 0 & 998 & 0 \\ \hline
-    \textbf{\# Items added} & 124'940 & 125'085 & 204'669 & 501 \\ \hline
-    \textbf{Final Est. Card.} & 1'058 & 367 & 998 & 1 \\ \hline
+    \textbf{Original Card. Est.} & 1,004 & 0 & 998 & 0 \\ \hline
+    \textbf{\# Items added} & 124,940 & 125,085 & 204,669 & 501 \\ \hline
+    \textbf{Final Card. Est.} & 1,058 & 367 & 998 & 1 \\ \hline
 \end{tabular}
@@
 \section{Conclusion}\label{sec:conclusions}
-The HyperLogLog algorithm is an elegant and efficient solution to the distinct count problem.
-Its simple structure makes it easy to code and use, as shows the growing number of available open-source implementations.
-Nonetheless, disclosing such low level details can be exploited by malicious users to manipulate the estimate of the sketch and break the system that is built on top.
-Our attack is trivial but powerful and should raise awareness on the security of HLL, such that product designers understand the risks they may encounter when exposed to an adversarial setting.
-Future work could consist in finding mitigations that do not rely on salt.
+The HyperLogLog algorithm is an elegant and efficient solution to the problem of estimating the cardinality of large sets.
+Its simple structure makes it easy to code and use, as shown by the growing number of available open-source implementations.
+Nonetheless, malicious users can manipulate the HLL cardinality estimate and thence break the security properties of systems relying on HLL.
+Our attacks are simple but powerful and should raise awareness of the limitations of HLL.
+Our analysis may assist software developers in understanding the risks they run when using HLL in adversarial settings.
+
+A formal analysis of using salting as a countermeasure to our attacks remains open.  
+%Future work could consist in finding mitigations that do not rely on salt.
 For example, would the imposed ID setting (like Facebook) solve the issue but still be convenient for the various HLL applications?
-Also left for future work is to formalize the extension of our attack to the HLL variants, i.e., to show that it also breaks the HLL++ \cite{hllpratice} and sliding HLL \cite{slidinghll} algorithms.
+Also left for future work is the task of extending our attacks to HLL variants such as HLL++~\cite{hllpratice} and sliding HLL~\cite{slidinghll}.
 

### 2020-04-27 docs/HLL_exp_bad/main.tex
adding comment on expts

@@
 
-To assess the efficiency of our attacks, we can compare our results from Table~\ref{table:tab2} with the attack presented in the Section 5.1 of \cite{hllvuln}.
+To assess the efficiency of our attacks, we can compare our results from Table~\ref{table:tab2} with the attack presented in Section 5.1 of~\cite{hllvuln}.
 In that paper, the authors target the Redis~\cite{redishll} open-source implementation of HLL and perform the attack against an empty sketch.
@@
 In comparison, the adversary following our strategy is able to add 125,000 distinct items while keeping the cardinality estimate at 367, providing a reduction by a factor of 340 instead of 5.
+\kenny{I think we can say more here -- isn't it true that the attack in Section 5.1 of~\cite{hllvuln} is in the M2 scenario?
+Which is our S1?
+Whereas the attack you describe here is more like an S2 attack, a much weaker attack model?}
 

### 2020-04-28 docs/HLL_exp_bad/main.tex
ref

@@
 Among the repositories collecting more than 350 stars on GitHub, we can cite \cite{clahll} and \cite{datasketch}.
-HLL is also featured in many frameworks, including Redis \cite{redis}, Google's BigQuery \cite{bigquery} and several products of Apache such as Spark \cite{spahll,spahll2} and Druid \cite{druhll}.
+HLL is also featured in many frameworks, including Redis \cite{redis}, Google's BigQuery \cite{bigquery}, Facebook's Airlift \cite{airlift} and several products of Apache such as Spark \cite{spahll} and Druid \cite{druhll}.
 It is interesting to point out that all these implementations use a constant number of buckets and a fixed hash function.
@@
 To assess the efficiency of our attacks, we can compare our results from Table~\ref{table:tab2} with the attack presented in Section 5.1 of~\cite{hllvuln}.
-In that paper, the authors target the Redis~\cite{redishll} open-source implementation of HLL and perform the attack against an empty sketch.
+In that paper, the authors target the Redis~\cite{redis} open-source implementation of HLL and perform the attack against an empty sketch.
 Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780``~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality.

### 2020-04-28 docs/HLL_exp_bad/main.tex
rect adv mod in exp sec

@@
 Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780``~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality.
+The authors present this setup as the M2 scenario, without providing any explanation.
+When looking at the open-source code, we could gain details on the implementation that any attacker could obtain, and thus argue that based on this knowledge, the adversarial scenario is actually M1 (S2 in this work), making the comparison relevant.
 In comparison, the adversary following our strategy is able to add 125,000 distinct items while keeping the cardinality estimate at 367, providing a reduction by a factor of 340 instead of 5.
-\kenny{I think we can say more here -- isn't it true that the attack in Section 5.1 of~\cite{hllvuln} is in the M2 scenario?
-Which is our S1?
-Whereas the attack you describe here is more like an S2 attack, a much weaker attack model?}
 

### 2020-04-29 docs/HLL_exp_bad/main.tex
changes

--- 2020-04-28 docs/HLL_exp_bad/main.tex
@@
 We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  7,311,616 possible inputs while trying to keep the cardinality estimate as low as possible, either in the S2 or the S4 scenario.
-As a second experiment, and in order to get results that can be directly compared to the work of~\cite{hllvuln}, we mounted our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose what to insert from a set of 250,000 random items that are given to him.
+As a second experiment, and in order to get results that can be directly compared to the work of~\cite{hllvuln}, we mounted our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose what to insert from a set of 250,000 random items that are given to it.
 Finally, for completeness, we also run the attacks against HLL sketches that are initially empty.

### 2020-04-29 docs/HLL_exp_bad/main.tex
more tweaks in expt section

--- 2020-04-29 docs/HLL_exp_bad/main.tex
@@
 
-Since the assumptions of S1 appear to us to be unrealistic and since an attacker can escalate from S3 to S4, we focus on the scenarios S2 and S4 in what follows.
+Since the assumptions of S1 appear to us to be unrealistic and since an attacker can escalate from S3 to S4, we focus on scenarios S2 and S4 in what follows.
 We first present an attack in scenario S2, providing much stronger results than the corresponding attack of \cite{hllvuln} (in their equivalent scenario M1).
@@
 We assumed so far that there are already values coming from other honest users' data in the HLL.
-We now remove this assumption.
-The blind adversary in scenario S2 has no way to detect that it is attacking an empty HLL and cannot adapt its strategy, so all buckets that are hit will update their estimator once to $2^{0+1}=2$ items, leading to a final cardinality estimate of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts.
-However, the adversary in S4 is able to detect that all buckets are empty and adapt its attack accordingly.
-The adversary can now target and fill one bucket only.
+We now remove this assumption and instead assume that the HLL sketch is empty at the beginning of our attacks.
+The blind adversary in scenario S2 has no way to detect that it is attacking an empty HLL and cannot adapt its strategy, so all buckets that are hit will update their estimator once to $2^{0+1}=2$ items, leading to a worst-case final cardinality estimate of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts.
+However, the adversary in S4 is able to detect that all buckets are empty and can adapt its attack accordingly.
+The adversary can now target and fill a single bucket only.
 This requires fixing $n+1$ bits of $h(x)$ for each input $x$, allowing the adversary to ``hide'' $2^{t-n-1}$ elements while keeping the HLL cardinality estimate to 1.
-Other trade-offs between work and the cardinality estimate are of course possible.
+Other trade-offs between work and the cardinality estimate are of course possible in scenario S4.
 
@@
 Many different implementations of HLL are available online.
-Among the repositories collecting more than 350 stars on GitHub, we can cite \cite{clahll} and \cite{datasketch}.
-HLL is also featured in many frameworks, including Redis \cite{redis}, Google's BigQuery \cite{bigquery}, Facebook's Airlift \cite{airlift} and several products of Apache such as Spark \cite{spahll} and Druid \cite{druhll}.
+Among the repositories collecting more than 350 stars on GitHub, we can cite~\cite{clahll} and~\cite{datasketch}.
+HLL is also featured in many frameworks, including Redis \cite{redis}, Google's BigQuery~\cite{bigquery}, Facebook's Airlift~\cite{airlift} and several products of Apache such as Spark~\cite{spahll} and Druid~\cite{druhll}.
 It is interesting to point out that all these implementations use a constant number of buckets and a fixed hash function.
@@
 
-As a proof-of-concept, we implemented our attacks against the HLL implementation from \cite{clahll}.
+As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}.
 The attack code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\textit{\url{https://github.com/PizzaWhisperer/HLLVuln}}}.
-For simplicity, our items are 4-character long ASCII strings, giving us up to 7,311,616 distinct items that can be inserted.
+For simplicity, our items are 4-character long ASCII strings, giving us up to $7,311,616= 2^{22.8}$ distinct items that can be inserted (hence we have $t=22.8$ in our attacks).
 We choose $h$ to be the 32-bit Murmur3 hash function from \cite{murmur3code} (note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
@@
 
-We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  7,311,616 possible inputs while trying to keep the cardinality estimate as low as possible, either in the S2 or the S4 scenario.
-As a second experiment, and in order to get results that can be directly compared to the work of~\cite{hllvuln}, we mounted our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose what to insert from a set of 250,000 random items that are given to it.
+We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $7,311,616$ possible inputs while trying to keep the cardinality estimate as low as possible, either in the S2 or the S4 scenario.
+As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mounted our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose what to insert from a set of 250,000 random items that are given to it.
 Finally, for completeness, we also run the attacks against HLL sketches that are initially empty.
@@
 These empirical results confirm our impact analysis on several points.
-First, when attacking a sketch already containing some items, we can see that the adversary is able to hide approximately half of the items in our starting set under S2, and around 83\% under S4.
+
+First, when attacking a sketch already containing some items, we can see that the adversary is on average able to hide about half of the items in our starting set under S2 (as predicted by our theoretical analysis), and about 83\% under S4.
 %( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$)
-Here $0.83 \approx (1-\frac{1}{2\cdot H_c})$ with $H_c\approx\frac{1000}{256}=3.9$.
-Secondly, when attacking an empty sketch, the final cardinality estimate is always exactly 367 $=\alpha2^{n+1}$ with $\alpha = 0.72$ as specified by equation (29) in~\cite{hll}, or 1, in respectively S2, S4.
+Here $0.83 \approx (1-\frac{1}{2\cdot H_c})$ with $H_c\approx\frac{1000}{256}=3.9$, so the results are also consistent with our theoretical analysis.
 
-To assess the efficiency of our attacks, we can compare our results from Table~\ref{table:tab2} with the attack presented in Section 5.1 of~\cite{hllvuln}.
+Secondly, when attacking an empty HLL sketch in S2, the final cardinality estimate is always exactly $367=\alpha2^{n+1}$ with $\alpha = 0.72$ as specified by equation (29) in~\cite{hll}.
+For S4, the final cardinality estimate is always exactly 1 (this comes from rounding up the estimated cardinality of $\alpha = 0.72$ that arises from our attack targeting a single bucket).
+Note that in S4 with an initially empty HLL sketch, we insert (on average) 14,220 items into the sketch from our candidate set of $7,311,616$ items; this is again consistent with our theoretical predictions of $2^{t-n-1} = 2^{22.8-9} = 2^{13.8} = 14,263$.
+Of course, we could insert many more items into an initially empty HLL sketch in S4 by relaxing the single bucket restriction.
+This would come at the price of allowing the cardinality estimate to increase beyond 1.
+
+We can also compare our results from Table~\ref{table:tab2} with results from the attack presented in Section 5.1 of~\cite{hllvuln}.
 In that paper, the authors target the Redis~\cite{redis} open-source implementation of HLL and perform the attack against an empty sketch.
-Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780``~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality.
-In comparison, the adversary following our strategy is able to add 125,000 distinct items while keeping the cardinality estimate at 367, providing a reduction by a factor of 340 instead of 5.
-\kenny{I think we can say more here -- isn't it true that the attack in Section 5.1 of~\cite{hllvuln} is in the M2 scenario?
-Which is our S1?
-Whereas the attack you describe here is more like an S2 attack, a much weaker attack model?}
+Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780''~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality.
+The authors present this attack as being in their M2 (our S1) scenario, without providing a detailed explanation.
+By looking at the Redis open-source code, we were able to obtain full details of the implementation that any attacker could obtain, and thus argue that based on this knowledge, the adversarial scenario is actually M1 (S2 in this work).
+This makes the comparison between the results of~\cite{hllvuln} and our S2 results appropriate.
+The S2 adversary following our strategy is able to add 125,000 distinct items while keeping the cardinality estimate at 367.
+Compared to~\cite{hllvuln}, we can insert twice as many items and reduce the cardinality estimate by a factor of 340 instead of 5.
 

### 2020-05-05 docs/HLL_exp_bad/main.tex
draft of S1

@@
 \begin{itemize}
-\item[S1:] The adversary has access to a \textit{test oracle} provided by the sketch owner, which allows it to submit an item and learn the hypothetical effect of inserting it into the sketch.
-It is then able to determine if inserting this item \textit{would} increase the cardinality estimate, without actually inserting it.
-%The adversary is able to test the hypothetical effect of the insertion of an item on the cardinality estimation of the sketch.
-This can achieved via a \textit{test oracle} provided by the sketch owner, which allows an user to test the effect of an insertion on the cardinality estimate but without actually inserting the item, or if the attacker would gain access to a black-box replica of the sketch in use, i.e., buying the same switch and testing in the lab.
-%black-box access to the HLL in use -- that is, the adversary can ask for the HLL's cardinality estimate at any point in time, and so is able to check the effect on the cardinality estimate when inserting an item.
-%\kenny{TODO: clarify more what this is terms of a ''test oracle''; its stronger than described because the adversary can test the effect of an insertion without making the insertion!s}
+\item[S1:] The adversary has access to an empty black-box replica of the sketch in use.
+He can use the API to insert elements into that shadow device, get the cardinality estimate, and reset the sketch.
+The attacker does not know the elements previously added by honest users nor the state.
+That is we can test and find elements that are unlikely to increase the HLL estimate on the shadow HLL and then insert the on the target HLL but we do not know the state of the target HLL.
+It would also be reasonable for the provider to transform the items/hash with  a salt e.g.
+with a different salt per customer.
+So one can have both an open source implementation and not know the hash.
 \item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''.
@@
 
-We start by highlighting the fact that S1 seems unrealistic in practice, as there is no reason why a typical HLL implementation would provide such a test oracle.
-The authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker gains access to a replica of the machine on which the HLL is implemented.
-The adversary could run tests on it, sequentially inserting, asking for the cardinality estimation and resetting the device, retaining only the items which do not increment the HLL estimate.
-This may be possible under certain circumstances, but we do not consider it typical of HLL deployments.
+We start by highlighting the fact that S1 seems unrealistic in practice.
+In a networking application, the authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker learns which machine is used to do the monitoring and buys the same device.
+It is reasonable to think that the provider may use a different salt per device, which would make any effect on the shadow device unexpected on the target and would break the whole attack presented in Section 3.2 of \cite{hllvuln}.
+We consider that the machines use the same exact internals.
+Assuming that such setup is possible, we show that it is actually possible for the attacker to learn the internals of the HLL sketch.
+The API must explicitly tell the customer the error of the sketch, and hence the number of buckets.
+We show that the attacker can learn the hash of a targeted item using the knowlwdege of S1, and thus end up in the S2 scenario.
+To learn $h(x)$, the attacker can insert the item $x$ on en empty sketch, and ask for the resulting cardinality, which we denote $c$.
+By solving:
+\begin{eqnarray*}
+    \alpha 2^{2n}(\sum_{j=1}^{2^n}2^{-c_j})^{-1}& = & c \\
+    \Leftrightarrow log_2(c)-log_2(\alpha m^2 - cm + c) &=& M[i].
+\end{eqnarray*}
+where  the attacker can learn the value of the updated estimator of the bucket $x$ was mapped to, and recover some bits of $h(x)$.
+This may be possible under certain circumstances, but we do not consider it typical of HLL deployments, and furthermore, it is easy for the attacker to escalate to the knowledge of S2.
 

### 2020-05-05 docs/HLL_exp_bad/main.tex
draft S1

@@
 He can use the API to insert elements into that shadow device, get the cardinality estimate, and reset the sketch.
-The attacker does not know the elements previously added by honest users nor the state.
-That is we can test and find elements that are unlikely to increase the HLL estimate on the shadow HLL and then insert the on the target HLL but we do not know the state of the target HLL.
-It would also be reasonable for the provider to transform the items/hash with  a salt e.g.
-with a different salt per customer.
-So one can have both an open source implementation and not know the hash.
+The attacker does not know the state of the targeted HLL sketch, i.e., the cardinality estimation nor elements previously added by other users.
 \item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''.
@@
 In a networking application, the authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker learns which machine is used to do the monitoring and buys the same device.
-It is reasonable to think that the provider may use a different salt per device, which would make any effect on the shadow device unexpected on the target and would break the whole attack presented in Section 3.2 of \cite{hllvuln}.
-We consider that the machines use the same exact internals.
-Assuming that such setup is possible, we show that it is actually possible for the attacker to learn the internals of the HLL sketch.
-The API must explicitly tell the customer the error of the sketch, and hence the number of buckets.
-We show that the attacker can learn the hash of a targeted item using the knowlwdege of S1, and thus end up in the S2 scenario.
-To learn $h(x)$, the attacker can insert the item $x$ on en empty sketch, and ask for the resulting cardinality, which we denote $c$.
-By solving:
+It is possible that the provider uses a different salt/key per device, which would make any pre-computation irrelevant, and considerably weaken the attack presented in Section 3.2 of \cite{hllvuln}.
+If we consider that the devices are perfect copies, we show that it is actually possible for the attacker to learn the internals of the HLL sketch.
+Precisely, the attacker can recover some bits of interest in the hash of targeted items, and infer the knowledge given to the adversary under S2.
+To do so, the attacker can insert the item $x$ on an empty sketch, ask for the resulting cardinality, which we denote $c$, and solve:
 \begin{eqnarray*}
     \alpha 2^{2n}(\sum_{j=1}^{2^n}2^{-c_j})^{-1}& = & c \\
-    \Leftrightarrow log_2(c)-log_2(\alpha m^2 - cm + c) &=& M[i].
+    \Leftrightarrow log_2(c)-log_2(\alpha 2^{2n} - c2^n + c) &=& c_i
 \end{eqnarray*}
-where  the attacker can learn the value of the updated estimator of the bucket $x$ was mapped to, and recover some bits of $h(x)$.
-This may be possible under certain circumstances, but we do not consider it typical of HLL deployments, and furthermore, it is easy for the attacker to escalate to the knowledge of S2.
+to learn $c_i$.
+$\alpha$ is a constant, $2^n$ is known as the provider must explicitly tell the customer the expected error of the estimation, hence the number of buckets, and $i$ is an arbitrary value representing the index of the bucket $x$ is mapped to (recovering $i$ is not useful).
+As $c_i$ denotes the $i$ bucket's content, updated by the insertion of $x$, $c_i-1$ is (an approximation of) the number of leading zeros in the $l$ rightmost bits of $h(x)$, which is the information extracted from $h(x)$ and used by the adversary in his attack under S2, as we detail in Section~\ref{sec:attacks}.
+%the value of the updated estimator of the bucket $x$ was mapped to, and recover some bits of $h(x)$.
+To summarize, we do not consider this scenario typical of HLL deployments, and even if such setup may be possible under certain circumstances, the attacker can infer the knowledge given in S2.
+
+\mathilde{Do you think that we should mention the idea behind the attack of \cite{hllvuln} Section 3.2 ?
+As they explained in their emails "The attacker can then test and find elements that do not increase the HLL estimate on the shadow HLL and thus will unlikely increase the cardinality of the target HLL when inserted."}
 

### 2020-05-12 docs/HLL_exp_bad/main.tex
modif S1

@@
 \begin{itemize}
-\item[S1:] The adversary has access to an empty black-box replica of the sketch in use.
-He can use the API to insert elements into that shadow device, get the cardinality estimate, and reset the sketch.
-The attacker does not know the state of the targeted HLL sketch, i.e., the cardinality estimation nor elements previously added by other users.
+\item[S1:] The adversary does not know the details of the HLL implementation but can access a copy of the HLL instance as an user.
+He can use the API to insert elements into this shadow sketch, get the cardinality estimate, and reset it to its native empty state.
+The attacker does not know the state of the targeted HLL sketch, meaning that he cannot query its cardinality estimation and does not have the list of elements previously inserted into it by other users.
 \item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''.

### 2020-05-15 docs/HLL_exp_bad/main.tex
authors
fix up

--- 2020-05-05 docs/HLL_exp_bad/main.tex
@@
-\documentclass{IEEEtran}
+\documentclass[11pt]{article}
 \usepackage{graphicx}
@@
 \usepackage{xspace}
+\usepackage{a4wide}
 
@@
 \title{HyperLogLog: Exponentially Bad in Adversarial Settings}
-\author{Mathilde Raynal, Kenneth G. Paterson}
+\author{Mathilde Raynal, Kenneth G. Paterson\\
+\vspace{3mm}
+Department of Computer Science, ETH Zurich\\
+{\tt mraynal@student.ethz.ch, kenny.paterson@inf.ethz.ch}}
 
@@
 
-\IEEEtitleabstractindextext{%
+%\IEEEtitleabstractindextext{%
 \begin{abstract}
@@
 \end{abstract}
-}
+%}
 
-\IEEEdisplaynontitleabstractindextext
+%\IEEEdisplaynontitleabstractindextext
 
@@
 \begin{itemize}
-\item[S1:] The adversary has access to an empty black-box replica of the sketch in use.
-He can use the API to insert elements into that shadow device, get the cardinality estimate, and reset the sketch.
-The attacker does not know the elements previously added by honest users nor the state.
-That is we can test and find elements that are unlikely to increase the HLL estimate on the shadow HLL and then insert the on the target HLL but we do not know the state of the target HLL.
-It would also be reasonable for the provider to transform the items/hash with  a salt e.g.
-with a different salt per customer.
-So one can have both an open source implementation and not know the hash.
+\item[S1:] The adversary does not know the details of the target HLL implementation but can access a copy of the HLL instance as a user.
+He can use the API to insert elements into this shadow sketch, get its cardinality estimate, and reset it to its native empty state.
+The attacker does not know the state of the targeted HLL sketch, meaning that he does not have the list of elements previously inserted into it by other users, nor can he obtain its cardinality estimate.
 \item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''.
@@
 We start by highlighting the fact that S1 seems unrealistic in practice.
-In a networking application, the authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker learns which machine is used to do the monitoring and buys the same device.
-It is reasonable to think that the provider may use a different salt per device, which would make any effect on the shadow device unexpected on the target and would break the whole attack presented in Section 3.2 of \cite{hllvuln}.
-We consider that the machines use the same exact internals.
-Assuming that such setup is possible, we show that it is actually possible for the attacker to learn the internals of the HLL sketch.
-The API must explicitly tell the customer the error of the sketch, and hence the number of buckets.
-We show that the attacker can learn the hash of a targeted item using the knowlwdege of S1, and thus end up in the S2 scenario.
-To learn $h(x)$, the attacker can insert the item $x$ on en empty sketch, and ask for the resulting cardinality, which we denote $c$.
-By solving:
+In a networking application, the authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker learns which machine is used to do the monitoring, buys the same device, and uses it as the shadow device.
+It is possible that the provider uses a different HLL salt or key per device.
+This would make any pre-computation with the shadow device irrelevant and considerably weaken the attack presented in Section 3.2 of \cite{hllvuln}.
+If, on the other hand, we assume that the devices are perfect copies of one another, then we show that it is actually possible for the attacker to predict the behaviour of the HLL sketch as elements are inserted into it, and effectively elevate itself to being an S2 adversary, without actually knowing the hash function $h$ in use.
+In detail, the attacker can recover some bits of interest in the hash values of targeted items, and infer the knowledge given to the adversary under S2.
+To do so, the attacker can insert the item $x$ on an empty sketch and ask for the resulting cardinality, denoted $c$.
+Note that inserting $x$ only touches one bucket of the HLL sketch; let that bucket have index $i$ (which is unknown to the attacker) and value $c_i$ after the insertion.
+From the formula used to compute the HLL sketch cardinality estimate $c$, we get the equation:
 \begin{eqnarray*}
-    \alpha 2^{2n}(\sum_{j=1}^{2^n}2^{-c_j})^{-1}& = & c \\
-    \Leftrightarrow log_2(c)-log_2(\alpha m^2 - cm + c) &=& M[i].
+   \log_2(c)-\log_2(\alpha 2^{2n} - c2^n + c) &=& c_i
 \end{eqnarray*}
-where  the attacker can learn the value of the updated estimator of the bucket $x$ was mapped to, and recover some bits of $h(x)$.
-This may be possible under certain circumstances, but we do not consider it typical of HLL deployments, and furthermore, it is easy for the attacker to escalate to the knowledge of S2.
+which enables the adversary to learn the value $c_i$ from $c$.
+(Here $\alpha$ is a constant, and $2^n$ is known since the provider must explicitly tell the customer the expected error of the estimation, hence the number of buckets.) Now $c_i-1$ is (an approximation of) the number of leading zeros in the $l$ rightmost bits of $h(x)$.
+This is precisely the information extracted from $h(x)$ and used by the adversary in his attack in scenario S2, as we detail in Section~\ref{sec:attacks}.
+%the value of the updated estimator of the bucket $x$ was mapped to, and recover some bits of $h(x)$.
+To summarize, we do not consider scenario S1 typical of HLL deployments, and even if such a setup may be possible under certain circumstances, the attacker can infer the information used in our attack in scenario S2.
+
+%\mathilde{Do you think that we should mention the idea behind the attack of \cite{hllvuln} Section 3.2 ?
+As they explained in their emails "The attacker can then test and find elements that do not increase the HLL estimate on the shadow HLL and thus will unlikely increase the cardinality of the target HLL when inserted."}
 
@@
 To find the estimate held in a targeted bucket, the attacker gradually inserts items $x$ with an incrementing number of leading zeros in the $\ell$ rightmost bits of $h(x)$ whilst holding the $n$ bits determining the bucket index constant, until the cardinality estimate of the sketch increases.
-Once a change occurs, the attacker knows that the maximum number of leading zeros observed in that bucket was just updated to a value that can be recovered by looking at the last inserted item.
+Once an increase occurs, the attacker knows that the maximum number of leading zeros observed in that bucket was just updated to a value that can be recovered by looking at the last inserted item.
 This assumes the attacker has sufficient flexibility in its choice of inputs $x$ so that the relevant bit conditions on $h(x)$ can be forced.
@@
 For example, in the scenario of \cite{portscanhll}, the adversary could change the source port of its IP packet until it is satisfied with its hash value.
-Given sufficient flexibility in the input $x$, the adversary can insert many items into the HLL without increasing the cardinality markedly, by repeatedly sampling from the input domain.
+Given sufficient flexibility in the input $x$, the adversary can insert many items into the HLL sketch without increasing the cardinality markedly, by repeatedly sampling from the input domain.
 

### 2020-05-25 docs/HLL_exp_bad/main.tex
fix up
pseudocode

@@
 Estimators from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ is empirically computed to correct a systematic multiplicative bias.
+On top of that, correction for small and big range are applied as shown in \ref{fig:hll}.
 As result, HLL is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
@@
 Such a simple set union operation allows easy parallelisation of operations among multiple machines independently, provided they use the same hash function and the same number of buckets.
+
+\begin{figure}[h]
+\centering
+\fbox{
+\begin{minipage}{40em}
+\begin{flushleft}
+\textit{Let $h : D \rightarrow \{0, 1\}^{32}$ hash data from $\mathcal{D}$ to binary $32$–bit words.\\
+Let $\rho(s)$ be the position of the leftmost $1$-bit of $s$: e.g., $\rho(1...) = 1, \rho(0001...) = 4, \rho(0^K) = K + 1.$\\}
+\textbf{define} $\alpha_{16} = 0.673$; $\alpha_{32} = 0.697$; $\alpha_{64} = 0.709$; $\alpha_{m} = 0.7213/(1+ 1.079/m)$ for $m \ge 128$;\\
+\vspace{0.5em}
+\textbf{Program} HyperLogLog (\textbf{input} $\mathcal{M}$ : multiset of items from domain $\mathcal{D}$).\\
+\textbf{assume} $m = 2^b$ with $b \in [4..16].$\\
+\textbf{initialize} a collection of $m$ registers, $M[1], ..., M[m]$, to $0$;\\
+\vspace{0.5em}
+\textbf{for} $v \in \mathcal{M}$ \textbf{do}\\
+    \hspace{2em} \textbf{set} $x := h(v);$\\
+    \hspace{2em} \textbf{set} $j = 1 + <x_1x_2...x_{b}>_2$; \hfill \textit{\{the binary address determined by the first $b$ bits of $x$\}}\\
+    \hspace{2em} \textbf{set} $w := x_{b+1}x_{b+2}...$;\\
+   \hspace{2em} \textbf{set} $M[j] :=$ max$(M[j], \rho(w))$;\\
+\vspace{0.2em}
+\textbf{compute} $E := \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$; \hfill \textit{\{the “raw” HyperLogLog estimate\}}\\
+\vspace{0.2em}
+\textbf{if} $E \le \frac{5}{2}m$ \textbf{then}\\
+    \hspace{2em}  \textbf{let} $V$ be the number of registers equal to $0$;\\
+    \hspace{2em}  \textbf{if} $V \neq 0$ \textbf{then set} $E^*:= m $log$(m/V)$ \textbf{else set} $E^*:= E$; \hfill \textit{\{small range correction\}}\\
+\textbf{if} $E \le \frac{1}{30}2^{32}$ \textbf{then}\\
+    \hspace{2em} \textbf{set} $E^*:= E$; \hfill \textit{\{intermediate range—no correction\}}\\
+\textbf{if} $E > \frac{1}{30}2^{32}$ \textbf{then}\\
+    \hspace{2em}  \textbf{set} $E^*:= -2^{32}$log$(1 - E/2^{32})$; \hfill \textit{\{large range correction\}}\\
+\textbf{return} \textit{cardinality estimate $E^*$ with typical relative error} $\pm1.04/\sqrt{m}$.
+\end{flushleft}
+\end{minipage}
+}
+\caption{The HyperLogLog Program}
+\label{fig:hll}
+\end{figure}
 
@@
 \begin{table}[h]
+\centering
 \caption{Attack results, averaged over 30 iterations.}
@@
 \begin{table}[h]
+\centering
 \caption{Attack results, averaged over 30 iterations, in the setting of \cite{hllvuln}.}
@@
 For example, would the imposed ID setting (like Facebook) solve the issue but still be convenient for the various HLL applications?
-Also left for future work is the task of extending our attacks to HLL variants such as HLL++~\cite{hllpratice} and sliding HLL~\cite{slidinghll}.
+Also left for future work is the task of extending our attacks to HLL variants such as redis alternative using maximum likelihood~\cite{newhll}, HLL++~\cite{hllpratice} and sliding HLL~\cite{slidinghll}.
 

### 2020-05-25 docs/HLL_exp_bad/main.tex
fix

@@
 Estimators from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ is empirically computed to correct a systematic multiplicative bias.
-On top of that, correction for small and big range are applied as shown in \ref{fig:hll}.
+Small and big cardinalities are handled differently to correct eventual distortions, as shown in Fig.
+\ref{fig:hll}.
+When the estimate is comparatively small with respect to $2^n$, the Hit Counting algorithm of Whang et al.~\cite{hitcounting} which looks at the number of empty bins is used to provide the estimator.
+A similar correction is applied for large cardinalities in order to take into account the likelihood of hashing collisions.
 As result, HLL is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
 \centering
@@
 \textbf{Program} HyperLogLog (\textbf{input} $\mathcal{M}$ : multiset of items from domain $\mathcal{D}$).\\
-\textbf{assume} $m = 2^b$ with $b \in [4..16].$\\
+\textbf{assume} $m = 2^n$ with $n \in [4..16].$\\
 \textbf{initialize} a collection of $m$ registers, $M[1], ..., M[m]$, to $0$;\\
@@
     \hspace{2em} \textbf{set} $x := h(v);$\\
-    \hspace{2em} \textbf{set} $j = 1 + <x_1x_2...x_{b}>_2$; \hfill \textit{\{the binary address determined by the first $b$ bits of $x$\}}\\
-    \hspace{2em} \textbf{set} $w := x_{b+1}x_{b+2}...$;\\
+    \hspace{2em} \textbf{set} $j = 1 + <x_1x_2...x_{n}>_2$; \hfill \textit{\{the binary address determined by the first $b$ bits of $x$\}}\\
+    \hspace{2em} \textbf{set} $w := x_{n+1}x_{n+2}...$;\\
    \hspace{2em} \textbf{set} $M[j] :=$ max$(M[j], \rho(w))$;\\

### 2020-08-19 docs/HLL_exp_bad/main.tex
Update from report

@@
 \documentclass[11pt]{article}
-\usepackage{graphicx}
-\usepackage{array}
-\usepackage{url}
-\usepackage{hyperref}
-\usepackage{xspace}
+\usepackage{graphicx, amsmath, amsthm, amssymb, setspace, array, url, hyperref, xspace}
 \usepackage{a4wide}
-
-%% TODO NOTES
-
-\usepackage{xcolor}
-\definecolor{oxygenorange}{HTML}{FFDD00}
-\usepackage[color=oxygenorange]{todonotes}
-\newcommand{\mathilde}[1]{\todo[inline]{\textbf{Mathilde:} #1}\xspace}
-\newcommand{\kenny}[1]{\todo[inline]{\textbf{Kenny:} #1}\xspace}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 
@@
 In this paper, we present a complete analysis of HLL in adversarial setting, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
-In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of $O(2^t)$ where $t$ is th number of bits of flexibility in the input).
+In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of $O(2^t)$ where $t$ is the number of bits of flexibility in the input).
 
@@
 Section~\ref{sec:overview} gives an overview of the HyperLogLog algorithm.
+A brief summary of the related works can be found in Section \ref{sec:related}.
 Section~\ref{sec:attacks} presents our different adversarial models.
 There, we also give attacks and evaluate their impacts under the different adversarial models.
-Section~\ref{sec:conclusions} provides our conclusions and some ideas for future work.
+The experimental results are discussed in Section~\ref{sec:exp} and Section~\ref{sec:conclusions} provides our conclusions and some ideas for future work.
 
@@
 In practice, to ensure a uniform distribution, the stream values are first processed by a hash function $h$ before the leading bits are inspected.
-So to estimate the cardinality of a stream, we need only store a representation of $k$, the length of the largest observed string of leading zero-bits for numbers in the stream.
+So to estimate the cardinality of a stream, we need only to store a representation of $k$, the length of the largest observed string of leading zero-bits for numbers in the stream.
 This provides a very compact mechanism for estimating the stream cardinality -- if the true cardinality is $N$, then just $\log\log(N)$ bits are needed to store the estimate.
@@
 In order to improve the estimate, we can use many estimators in parallel instead of one, and average the results.
-Each estimate is stored in its own bucket.
+Each estimate is stored in its own register, that we call a bucket.
 The collection of buckets is referred to as the HLL \emph{sketch}.
-To map a value $x$ to one of $2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$ (so that the output length of $h(\cdot)$ is $n+\ell$).
+To map a value $x$ to one of $2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$.
+The output length of $h(\cdot)$ is thus $n+\ell$, fixed to 32 in~\cite{hll}.
 Estimators from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ is empirically computed to correct a systematic multiplicative bias.
-Small and big cardinalities are handled differently to correct eventual distortions, as shown in Fig.
-\ref{fig:hll}.
-When the estimate is comparatively small with respect to $2^n$, the Hit Counting algorithm of Whang et al.~\cite{hitcounting} which looks at the number of empty bins is used to provide the estimator.
+
+Relatively small and big cardinalities are handled separately to avoid eventual distortions, as shown in Fig.~\ref{fig:hll}.
+When the raw cardinality estimate is below $\frac{5}{2}\times 2^n$, the Hit Counting algorithm of Whang et al.~\cite{hitcounting} is used to produce the final cardinality estimator, looking at the number of empty buckets (whose register is 0) instead of the number of leading zeros in the stream.
 A similar correction is applied for large cardinalities in order to take into account the likelihood of hashing collisions.
+Some works building on top of~\cite{hll} replace the Hit counting algorithm by other schemes and use a hash function with a larger output domain to reduce the chances of collision and thus get rid of the need for large cardinality range correction.
+For example, \cite{hllpratice} uses a 64-bit hash function and Linear Counting along with bias correction in the small cardinality range.
+In this report we target the original paper's implementation reported in Fig.~\ref{fig:hll} and leave for future work the analysis of its variations.
+
 As result, HLL is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
@@
 An interesting property of HLL is that it supports merging in a lossless way.
-To combine two buckets at index $i$ in two different HLL instances, take the maximum of the two bucket entries and assign that to the matching bucket $i$ in the merged HLL sketch.
+To combine two buckets at index $i$ in two different HLL instances, one can take the maximum of the two bucket entries and assign that to the matching bucket $i$ in the merged HLL sketch.
 Such a simple set union operation allows easy parallelisation of operations among multiple machines independently, provided they use the same hash function and the same number of buckets.
@@
 \centering
+     \makebox[\linewidth]{
 \fbox{
@@
 Let $\rho(s)$ be the position of the leftmost $1$-bit of $s$: e.g., $\rho(1...) = 1, \rho(0001...) = 4, \rho(0^K) = K + 1.$\\}
+\textbf{define} $m = 2^n$ with $n \in [4..16].$\\
 \textbf{define} $\alpha_{16} = 0.673$; $\alpha_{32} = 0.697$; $\alpha_{64} = 0.709$; $\alpha_{m} = 0.7213/(1+ 1.079/m)$ for $m \ge 128$;\\
 \vspace{0.5em}
-\textbf{Program} HyperLogLog (\textbf{input} $\mathcal{M}$ : multiset of items from domain $\mathcal{D}$).\\
-\textbf{assume} $m = 2^n$ with $n \in [4..16].$\\
+\textbf{Algorithm} HyperLogLog (\textbf{input} $\mathcal{M}$ : multiset of items from domain $\mathcal{D}$).\\
 \textbf{initialize} a collection of $m$ registers, $M[1], ..., M[m]$, to $0$;\\
@@
     \hspace{2em} \textbf{set} $x := h(v);$\\
-    \hspace{2em} \textbf{set} $j = 1 + <x_1x_2...x_{n}>_2$; \hfill \textit{\{the binary address determined by the first $b$ bits of $x$\}}\\
+    \hspace{2em} \textbf{set} $j = 1 + <x_1x_2...x_{n}>_2$; \hfill \textit{\{the binary address determined by the first $n$ bits of $x$\}}\\
     \hspace{2em} \textbf{set} $w := x_{n+1}x_{n+2}...$;\\
@@
     \hspace{2em}  \textbf{let} $V$ be the number of registers equal to $0$;\\
-    \hspace{2em}  \textbf{if} $V \neq 0$ \textbf{then set} $E^*:= m $log$(m/V)$ \textbf{else set} $E^*:= E$; \hfill \textit{\{small range correction\}}\\
+    \hspace{2em}  \textbf{if} $V \neq 0$ \textbf{then set} $E^*:= m $ln$(m/V)$ \textbf{else set} $E^*:= E$; \hfill \textit{\{small range correction\}}\\
 \textbf{if} $E \le \frac{1}{30}2^{32}$ \textbf{then}\\
@@
 \textbf{if} $E > \frac{1}{30}2^{32}$ \textbf{then}\\
-    \hspace{2em}  \textbf{set} $E^*:= -2^{32}$log$(1 - E/2^{32})$; \hfill \textit{\{large range correction\}}\\
+    \hspace{2em}  \textbf{set} $E^*:= -2^{32}$ln$(1 - E/2^{32})$; \hfill \textit{\{large range correction\}}\\
 \textbf{return} \textit{cardinality estimate $E^*$ with typical relative error} $\pm1.04/\sqrt{m}$.
@@
 \end{minipage}
-}
-\caption{The HyperLogLog Program}
+}}
+\caption{The HyperLogLog Algorithm from~\cite{hll}}
 \label{fig:hll}
@@
 They are all potentially vulnerable to the attacks we present below.
-%By not assessing the risks involved with the presence of an adversary targeting the HLL sketch, they may allow an attack like the one we present, breaking down the whole solution.
 
-\subsection{Related Work}
+\section{Related Work}\label{sec:related}
 
@@
 In~\cite{hllvuln}, Reviriego and Ting exploit a vulnerability of HLL to manipulate the cardinality estimate, leading to a five-fold reduction in the estimate compared to the true cardinality.
-They use an attack model that we consider to be quite artificial, since it gives the attacker the ability to test whether inserting an item \emph{would} increase the cardinality estimate, without actually having to insert the item.
+Their attack model can be considered to be quite artificial, since it gives the attacker the ability to test whether inserting an item \emph{would} increase the cardinality estimate, without actually having to insert the item.
+After some clarifications obtained from personal correspondence with the authors, we finally agree that such setup might be possible under certain circumstances.
 See further discussion in Section~\ref{sec:attacks}.
@@
 For example it can manipulate IPID and other header fields in the context of IP packets.
-This enables the adversary to control the output of hash function $h$ and flexibly choose which items will be inserted into the HLL sketch, and in which buckets.
+This enables the adversary to control the output of hash function $h$ and flexibly choose which items will be inserted into the HLL sketch.
 
-\subsection{Adversarial Setting}
+\subsection{Adversarial Setting}\label{sec:set}
 We model four distinct adversarial scenarios based on the adversary's knowledge and capabilities.
-For all of them, we assume that the sketch already contains some existing data from honest users to simulate a real-life setting.
-Among the scenarios, the \textit{insider} setup of \cite{cardestprivacy} that we present as S4 makes the strongest assumptions about the adversarial capabilities since it assumes that the attacker has a perfect view on the sketch.
-%combines the knowledge of adversaries from S1 and S2.
-%\kenny{Not really! Because S1 is actually quite unrealistic, as we will argue below.}
+To simulate real-life settings, we consider both cases when the sketch is shared and receives inputs from other honest users but also when inputs are under adversarial control only.
+It is safe to assume that the adversary knows the setting being used. 
+Those two settings differ mainly in the fact that the adversary is attacking a sketch that already contains other users' data in the former option, but is empty in the latter one.
+
+Among the scenarios, the \textit{insider} setup of~\cite{cardestprivacy} that we present as S4 makes the strongest assumptions about the adversarial capabilities since it assumes that the attacker has a perfect view on the sketch.
 
@@
 He can use the API to insert elements into this shadow sketch, get its cardinality estimate, and reset it to its native empty state.
-The attacker does not know the state of the targeted HLL sketch, meaning that he does not have the list of elements previously inserted into it by other users, nor can he obtain its cardinality estimate.
+The attacker does not know the state of the targeted HLL sketch, meaning that he does not have the list of elements previously inserted into it by other users.
 \item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''.
@@
 
-We start by highlighting the fact that S1 seems unrealistic in practice.
-In a networking application, the authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker learns which machine is used to do the monitoring, buys the same device, and uses it as the shadow device.
-It is possible that the provider uses a different HLL salt or key per device.
-This would make any pre-computation with the shadow device irrelevant and considerably weaken the attack presented in Section 3.2 of \cite{hllvuln}.
-If, on the other hand, we assume that the devices are perfect copies of one another, then we show that it is actually possible for the attacker to predict the behaviour of the HLL sketch as elements are inserted into it, and effectively elevate itself to being an S2 adversary, without actually knowing the hash function $h$ in use.
-In detail, the attacker can recover some bits of interest in the hash values of targeted items, and infer the knowledge given to the adversary under S2.
-To do so, the attacker can insert the item $x$ on an empty sketch and ask for the resulting cardinality, denoted $c$.
-Note that inserting $x$ only touches one bucket of the HLL sketch; let that bucket have index $i$ (which is unknown to the attacker) and value $c_i$ after the insertion.
-From the formula used to compute the HLL sketch cardinality estimate $c$, we get the equation:
-\begin{eqnarray*}
-   \log_2(c)-\log_2(\alpha 2^{2n} - c2^n + c) &=& c_i
-\end{eqnarray*}
-which enables the adversary to learn the value $c_i$ from $c$.
-(Here $\alpha$ is a constant, and $2^n$ is known since the provider must explicitly tell the customer the expected error of the estimation, hence the number of buckets.) Now $c_i-1$ is (an approximation of) the number of leading zeros in the $l$ rightmost bits of $h(x)$.
-This is precisely the information extracted from $h(x)$ and used by the adversary in his attack in scenario S2, as we detail in Section~\ref{sec:attacks}.
-%the value of the updated estimator of the bucket $x$ was mapped to, and recover some bits of $h(x)$.
-To summarize, we do not consider scenario S1 typical of HLL deployments, and even if such a setup may be possible under certain circumstances, the attacker can infer the information used in our attack in scenario S2.
-
-%\mathilde{Do you think that we should mention the idea behind the attack of \cite{hllvuln} Section 3.2 ?
-As they explained in their emails "The attacker can then test and find elements that do not increase the HLL estimate on the shadow HLL and thus will unlikely increase the cardinality of the target HLL when inserted."}
+We start by highlighting the fact that we do not consider scenario S1 typical of HLL deployments.
+In a networking application, the authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker learns which machine is used to do the monitoring, buys the same, and uses it as the shadow device.
+It is possible that the provider uses a different salt or key in the hash per device.
+This would make any pre-computation on the items using the shadow device irrelevant, and considerably weaken the attack presented in Section 3.2 of~\cite{hllvuln}.
+Hence, we assume for the rest of this work that shadow and targeted devices have the same internals and security parameters.
+If the HLL device receives inputs from several users, the attack from Section 3.2 of~\cite{hllvuln} is the easiest to carry out.
+It works as follow: "after each insertion [in the shadow device], elements that do not increase the cardinality estimate are retained".
+Iterating through the process several times refines the items' filtering before inserting them into the targeted device.
+We explore in this work the setup where inputs are coming only from the adversary, and propose an attack that provides better results than the one we just presented.
 
@@
 The pre-computation and query complexity required can be much less if the HLL sketch is only moderately loaded.
-
 A version of this attack can be carried out to elevate an external adversary to a full-strength insider adversary as per~\cite{cardestprivacy} (instead of our slightly weaker S4 adversary), since it can simply be done repeatedly each time the external adversary wants to know the complete internals of the HLL sketch.
@@
 
-Since the assumptions of S1 appear to us to be unrealistic and since an attacker can escalate from S3 to S4, we focus on scenarios S2 and S4 in what follows.
-We first present an attack in scenario S2, providing much stronger results than the corresponding attack of \cite{hllvuln} (in their equivalent scenario M1).
-We then use the additional information the adversary has access to in scenario S4 to make the attack even more powerful.
+Since an attacker can escalate from S3 to S4, we focus on scenarios S1, S2 and S4 in what follows.
+We first present an attack in scenario S1, providing much stronger results than the corresponding attack of~\cite{hllvuln} (in their equivalent scenario M2).
+We then use the additional information the adversary has access to in scenarios S2 and S4 to make the attack even more powerful.
 
 \subsection{Manipulating the HLL Cardinality Estimate}
+\noindent\paragraph{Scenario S1} The goal of the attacker is to infer some information about targeted items' hash values, precisely their bucket mappings, without knowing neither $h$ nor the salt in use (if any).
+With this knowledge, the adversary is able to retain items that are mapped to a subset of $b$ buckets out of the $2^n$, and minimize the resulting estimated cardinality.
+To achieve this, the attacker can take advantage of the Hit Counting algorithm initial phase to recover the number of non-empty buckets (whose register is not 0) after inserting few items.
+Indeed, in the small cardinality range, the cardinality estimator E* is computed as E*$=2^n$ln$(2^n/V)$ with $V$ the number of empty buckets.
+For typical HLL parameters, e.g.\ $n = 8$, the exact value of $V$, hence $b$, can be recovered from E*.
+Moreover, when inserting an item in the small cardinality range, the adversary will see a increase in the cardinality estimate if and only if $V$ is modified.
+This means that the number of empty buckets decreased by 1, therefore that the freshly inserted item is mapped to a bucket that was not hit before.
+
+Thus, if the adversary blindly picks the first $b$ buckets to be hit as its targeted buckets, he can then discard all items whose insertion causes the cardinality estimator to increase from $b+k$ to $b+k+1$ for $k \in \mathbb{N}$.
+In other words, when $b$ buckets are already hit and the insertion of an item causes the cardinality estimator to increase, the attacker can discard this item as he knows that a new bucket (other than one of the $b$ buckets to be filled) switched to a non-empty state.
+The adversary can keep inserting items and discarding some while in the small cardinality range.
+
+As soon as the cardinality estimator exceeds $\frac{5}{2}\times 2^n$, the adversary can assume that it corresponds to the raw HLL cardinality estimate with high probability.
+In this case, no more useful information can be inferred about the bucket mappings, so the attacker has to reset the sketch to the empty state and start again with the non-discarded items.
+The adversary must pay attention not to shuffle the non-discarded items in between iterations, as the $b$ buckets to be filled are blindly determined by the first inserted items, and must stay the same throughout the attack.
+After several iterations, the adversary discards all items mapped to unwanted buckets, such that $b$ buckets are filled and the others left untouched.
+
+The value of $b$ has to be picked smartly to keep the final cardinality estimate low, as the registers must hold values potentially very high.
+$b$ is going to be chosen such that even in the worst case scenario where all register values $M[j]$ hold the the maximal value of $l+1$, the Hit Counting algorithm is going to be used to produce the final estimation.
+So $b$ must satisfy:
+\begin{align}
+E:=\alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1} & \le \frac{5}{2}m \nonumber\\
+\alpha_mm^2\cdot((m-b)\cdot2^{-0}+b\cdot2^{-(l+1)})^{-1} & \le \frac{5}{2}m \nonumber\\
+b & \le m\cdot(1-\frac{2\alpha_m}{5})\cdot(1-2^{-(l+1)})^{-1} \label{eqn:1} 
+\end{align}
+with $m=2^n$.
+For example with $n = 8$ and $l = 24$, $b$ can take any value up to 184.
+The initial set of items to be filtered can be built by varying the flexible fields of the items $x$ the adversary is trying to insert.
+%so that the adversary can ``hide'' $2^{t-n}$ elements while keeping the HLL cardinality estimate to 1.
 
 \noindent\paragraph{Scenario S2} The adversary knows $h$ so it can adjust fields in the item $x$ it is trying to insert and keep only those for which the $\ell$ rightmost bits of $h(x)$ start with a one bit.
-In this case, the number of leading zero bits computed by the estimator of the destination bucket is always zero, so the estimator of this bucket is unlikely to be updated and the averaged approximate cardinality is likely to be left unchanged.
-For example, in the scenario of \cite{portscanhll}, the adversary could change the source port of its IP packet until it is satisfied with its hash value.
+In this case, the number of leading zero bits computed by the estimator of the destination bucket is always zero, so the estimator of this bucket is unlikely to be updated and the final averaged approximate cardinality is likely to be left unchanged.
+For example, in the scenario of~\cite{portscanhll}, the adversary could change the source port of its IP packet until it is satisfied with its hash value.
 Given sufficient flexibility in the input $x$, the adversary can insert many items into the HLL sketch without increasing the cardinality markedly, by repeatedly sampling from the input domain.
 
+When attacking an empty sketch, the adversary can keep this strategy and already mount a powerful attack, but can do better by following the approach presented under S1.
+Filtering items is easier in this scenario as the adversary can infer the bucket mapping of an item $x$ directly from the first $n$ bits of $h(x)$.
+A trivial approach for the adversary is to retain items mapped to buckets 1 to $b$.
+We argue why targeting $b$ buckets is a better approach in the impact analysis presented in Section~\ref{sec:impact}.
+
 \noindent\paragraph{Scenario S4} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values per bucket in the HLL sketch).
-So, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
-In other words, if $h(x)$ is mapped to the $i$-th bucket with estimated cardinality $2^{c_i + 1}$ (because a value with $c_i$ leading zeroes was previously observed for this bucket), then, instead of the strict restriction of having a one bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a one bit in any of the first $c_i+1$ positions in this substring.
+Therefore, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
+In other words, if $h(x)$ is mapped to the $i$-th bucket with estimated cardinality $2^{M[i]}=2^{c_i + 1}$ (because a value with $c_i$ leading zeroes was previously observed for this bucket), then, instead of the strict restriction of having a one bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a one bit in any of the first $c_i+1$ positions in this substring.
 In case the adversary hits an empty bucket (for which the estimator is 0), it must skip that bucket completely.
 Under these conditions, the estimators in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.
-A special case arises when the HLL is completely empty at the beginning of the attack; here the adversary is forced to increase the estimate in at least one bucket, but with sampling this can obviously be done in such a way as to increase the HLL estimate to $2$ in a single bucket, leaving all the other buckets untouched.
 
-\subsection{Impact}
-We now evaluate the impact of the above attacks on the HLL cardinality estimate.
+A special case arises when the HLL is completely empty at the beginning of the attack.
+Here the attacker is forced to increase the estimate in at least one bucket.
+Once again, the adversary can choose to fill $b$ buckets following the strategy described in S2, leaving all the other buckets untouched.
+
+\subsection{Impact}\label{sec:impact}
+In this section we evaluate the impact of the above attacks on the HLL cardinality estimate.
 We assume that $h$ is a ``good" hash function, so that each of the bits of $h(x)$ can be regarded as being independent and uniformly random.
+We assume that the adversary has $t$ bits of flexibility in its choice of inputs.
+
+\subsubsection{Non-Empty Sketch}\label{sec:nonempty}
+
+We start by analysing our results when the attack is ran against a sketch that already received inputs from other users, first under scenario S2 then S4.
 
 In scenario S2, the adversary needs to craft two distinct candidates for $x$ on average to obtain a one bit at the first position in the $\ell$ rightmost bits of $h(x)$, thus successfully performing the attack for a single insertion.
-Assuming the adversary has $t$ bits of flexibility in its choice of inputs, the adversary can expect to ``hide''  $2^{t-1}$ distinct items without increasing the HLL cardinality estimate at all, under the assumption that all buckets in the HLL sketch already hold a value $c_i \ge 1$.
+The adversary can expect to ``hide''  $2^{t-1}$ distinct items without increasing the HLL cardinality estimate at all, under the assumption that all buckets in the HLL sketch already hold a value $c_i \ge 1$.
 On the other hand, if some buckets are empty (but the adversary does not know which ones in the blind setting of S2), then we may expect a moderate increase in the HLL cardinality estimator.
-In the worst case, when all buckets are initially empty, the cardinality estimate will increase from 0 to at most $\alpha2^{n+1}$ (regardless of the number of items the adversary inserts).
 Notice that in all cases, provided $t$ is large enough, the adversary can keep the cardinality estimate linear in the total number of buckets ($2^n$) while the expected value of the true cardinality, $2^{t-1}$, can be made as large as the adversary pleases.
@@
 \sum_{i=1}^{2^n}\frac{1}{2^n} \cdot (1-\frac{1}{2^{c_i+1}}) & = & 1-\frac{1}{2}\cdot (\frac{1}{2^n}\cdot \sum_{i=1}^{2^n} {(2^{c_i})}^{-1}) \\
-& = &1-(2H_c)^{-1} \\
+& = &1-(2H_c)^{-1} 
 \end{eqnarray*}
 where $H_c$ is the harmonic mean of the counts $2^{c_1}, 2^{c_2}, \ldots, 2^{c_{2^n}}$.
-
-%$P(h(x)$ does NOT update bucket $i| h(x)$ mapped to bucket $i)=P(h(x)$ has LESS than $c_i$ leading 0s$)=1-P(h(x)$ has strictly MORE than $c_i$ leading 0s$)=1-\frac{1}{2^{c_i+1}}$\\
-%Across all buckets, the adversary does not update any estimate with probability $P=
-%\sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i$ AND $h(x)$ does NOT update bucket $i) = \sum_{i=1}^{2^n}P(h(x)$ mapped to bucket $i)P(h(x)$ does NOT update bucket $i| h(x)$ mapped to bucket $i) = \sum_{i=1}^{2^n}\frac{1}{2^n} \times (1-\frac{1}{2^{c_i+1}}) = 1-\frac{1}{2}\times(\frac{1}{2^n}\times \sum_{i=1}^{2^n} {(2^{c_i})}^{-1}) = 1-\frac{1}{2}\times\frac{1}{H_c} = 1-(2H_c)^{-1}$, with $H_c$ the harmonic mean of the numbers $2^{c_1}, 2^{c_2}, ..., 2^{c_{2^n}}$.
 So the adversary needs to craft on average $(1-(2H_c)^{-1})^{-1}$ distinct candidates for $x$ in order to insert a single candidate, and, given $t$ bits of flexibility in its choice of inputs, can expect to ``hide" up to $(1-(2H_c)^{-1}) \cdot 2^t$ distinct items without increasing the HLL cardinality estimate at all.
-
-%\textit{NOTE: does it take into account the special case of empty bucket ?
-counters are initialised to $c_i=-oo$ (or -1 in the code???) thus $P(h(x)$ has LESS than $-oo$ leading $0s) = 0$ but it is hard to add it to the probabilities while staying clear.}\\
-%2$^{len(h(x))}\times \frac{2H_c-1}{H_c}$ items into the sketch.\\
 
@@
 The modified analysis is quite straightforward.
-Suppose $r \le 2^n-1$ out of $2^n$ buckets have a count of zero.
-Then the cost of crafting an input $x$ becomes $(1-(2H_c)^{-1})^{-1} \cdot (1- \frac{r}{2^n})^{-1}$ trials, while the total number of items the adversary can expect to insert without increasing the HLL cardinality estimate becomes $(1-(2H_c)^{-1}) \cdot (1- \frac{r}{2^n})\cdot 2^t$.
+Suppose $V \le 2^n-1$ out of $2^n$ buckets have a count of zero.
+Then the cost of crafting an input $x$ becomes $(1-(2H_c)^{-1})^{-1} \cdot (1- \frac{V}{2^n})^{-1}$ trials, while the total number of items the adversary can expect to insert without increasing the HLL cardinality estimate becomes $(1-(2H_c)^{-1}) \cdot (1- \frac{V}{2^n})\cdot 2^t$.
 Of course, the adversary can relax its attack and allow some of the empty buckets to be hit.
-This increases the insertion rate at the cost of a small increase in the final cardinality estimate.
+This would increase the insertion rate at the cost of a small increase in the final cardinality estimate.
 
-We assumed so far that there are already values coming from other honest users' data in the HLL.
-We now remove this assumption and instead assume that the HLL sketch is empty at the beginning of our attacks.
-The blind adversary in scenario S2 has no way to detect that it is attacking an empty HLL and cannot adapt its strategy, so all buckets that are hit will update their estimator once to $2^{0+1}=2$ items, leading to a worst-case final cardinality estimate of $\alpha2^{n+1}$ regardless of the number of items the adversary inserts.
-However, the adversary in S4 is able to detect that all buckets are empty and can adapt its attack accordingly.
-The adversary can now target and fill a single bucket only.
-This requires fixing $n+1$ bits of $h(x)$ for each input $x$, allowing the adversary to ``hide'' $2^{t-n-1}$ elements while keeping the HLL cardinality estimate to 1.
-Other trade-offs between work and the cardinality estimate are of course possible in scenario S4.
+\textbf{Conclusion:} In this setting, we see that an adversary who has $t$ bits of flexibility in its inputs can expect to insert exponentially (in $t$) many input values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the cases where the HLL sketch has empty buckets in S2).
+The result is largely independent of the HLL parameters, but does depend on the extent to which the HLL sketch is already filled by other users, via a factor $1-(2H_c)^{-1}$ and the number of empty buckets $V$, via a factor $1- \frac{V}{2^n}$.
+Other trade-offs between work and the cardinality estimate are possible.
 
-In both scenarios S2 and S4, we see that an adversary who has $t$ bits of flexibility in its inputs can expect to insert exponentially (in $t$) many inputs values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the cases where the HLL sketch has empty buckets in S2 or where the HLL sketch is empty to begin with in S2 and S4).
-The result is largely independent of the HLL parameters, but does depend on the extent to which the HLL sketch is already filled, via a factor $1-(2H_c)^{-1}$ and the number of empty buckets $r$, via a factor $1- \frac{r}{2^n}$.
+\subsubsection{Empty Sketch}
 
-\subsection{Experimental Results}
+We now consider the case where the HLL sketch is initially empty.
+In that case, the output of the attacks is the same for all scenarios.
+Targeting $b$ buckets allows the adversary to insert a fraction $\frac{b}{2^n}$ of the items, and thus to retain $\frac{b}{2^n}2^t=2^{t-n+\text{log}_2(b)}$ elements.
+After inserting the retained items, the cardinality of the sketch is estimated to $2^n$ln$(2^n/(2^n-b))$, approximating to $b$ for relatively small values of $b$.
+
+When following the initial strategy of S2, we saw in paragraph~\ref{sec:nonempty} that the adversary retains $2^{t-1}$ items.
+The insertion of those items will hit every bucket of the sketch with high probability, and update every register once to 1.
+The cardinality of the sketch is then estimated to $\alpha2^{2n}(\sum_{j=1}^{2^n}2^{-1})^{-1}=\alpha2^{n+1}$ and, as $V=0$, no small range cardinality correction is applied.
+On the other hand, when targeting and filling half the buckets ($b=2^{n-1}$), the adversary can insert as many items but lower the estimated cardinality to $2^n$ln$(2^n/(2^n-2^{n-1}))=2^n\text{ln}(2)$ instead of $\alpha{2^{n+1}}$ that we previously had, at no extra computational cost.
+
+\textbf{Conclusion:} When confronted to an empty sketch, the adversary is forced to concede a slight cardinally estimation increase in order to be able to insert some elements.
+For a fixed $b$, an adversary who has $t$ bits of flexibility in its inputs is then able to insert exponentially (in $t$) many input values into the HLL sketch while keeping the cardinality estimate constant.
+Any value below the upper-bound specified in eq. (4.1) can be chosen for $b$ based on the setup requirements. For example if the reset operation is costly, the attacker can adjust $b$ accordingly.
+
+\section{Experimental Results}\label{sec:exp}
 Many different implementations of HLL are available online.
 Among the repositories collecting more than 350 stars on GitHub, we can cite~\cite{clahll} and~\cite{datasketch}.
-HLL is also featured in many frameworks, including Redis \cite{redis}, Google's BigQuery~\cite{bigquery}, Facebook's Airlift~\cite{airlift} and several products of Apache such as Spark~\cite{spahll} and Druid~\cite{druhll}.
+HLL is also featured in many frameworks, including Redis~\cite{redis}, Google's BigQuery~\cite{bigquery}, Facebook's Airlift~\cite{airlift} and several products of Apache such as Spark~\cite{spahll} and Druid~\cite{druhll}.
 It is interesting to point out that all these implementations use a constant number of buckets and a fixed hash function.
@@
 For simplicity, our items are 4-character long ASCII strings, giving us up to $7,311,616= 2^{22.8}$ distinct items that can be inserted (hence we have $t=22.8$ in our attacks).
-We choose $h$ to be the 32-bit Murmur3 hash function from \cite{murmur3code} (note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
-% \kenny{Why do we get to choose $h$?
-Isn't it fixed by the specific implementation we are attacking?} \mathilde{The library has a place holder for a 32 bit hash function but lets the user choose it.
-This was one of my concerns and I was wondering if it would make sense to attack \cite{datasketch} instead (hardcoded $h$, but when we discussed it, you said that due to Kerckhoff's principle, this hash function should be public so it does not matter that much.}
+We choose $h$ to be the 32-bit Murmur3 hash function from~\cite{murmur3code} (note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
 We create an HLL sketch using $2^n = 256$ buckets and initialise it with 1,000 random items representing honest users' data.
 
-We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $7,311,616$ possible inputs while trying to keep the cardinality estimate as low as possible, either in the S2 or the S4 scenario.
-As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mounted our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose what to insert from a set of 250,000 random items that are given to it.
-Finally, for completeness, we also run the attacks against HLL sketches that are initially empty.
+We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $7,311,616$ possible inputs while trying to keep the cardinality estimate as low as possible, either in the S1, S2 or S4 scenario.
+As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose what to insert from a set of 250,000 random items that are given to it.
+Finally, for completeness, we also run the attacks against HLL sketches filled with adversarial inputs only, thus that are initially empty.
+We vary the value of $b$ in the attack under the S1 scenario to demonstrate possible trade-offs.
+For simplicity and to keep the attack as striking as possible, we set $b$ to 1 in the attacks on empty sketch under S2 and S4.
 All attacks were run 30 times and the outcomes averaged.
@@
 
-Secondly, when attacking an empty HLL sketch in S2, the final cardinality estimate is always exactly $367=\alpha2^{n+1}$ with $\alpha = 0.72$ as specified by equation (29) in~\cite{hll}.
-For S4, the final cardinality estimate is always exactly 1 (this comes from rounding up the estimated cardinality of $\alpha = 0.72$ that arises from our attack targeting a single bucket).
-Note that in S4 with an initially empty HLL sketch, we insert (on average) 14,220 items into the sketch from our candidate set of $7,311,616$ items; this is again consistent with our theoretical predictions of $2^{t-n-1} = 2^{22.8-9} = 2^{13.8} = 14,263$.
-Of course, we could insert many more items into an initially empty HLL sketch in S4 by relaxing the single bucket restriction.
-This would come at the price of allowing the cardinality estimate to increase beyond 1.
+Secondly, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly $2^n$ln$(2^n/(2^n-b))$ after inserting a fraction $\frac{b}{2^n}$ of the items (this comes from the Hit Counting algorithm result that arises from our attack targeting $b$ buckets).
 
 We can also compare our results from Table~\ref{table:tab2} with results from the attack presented in Section 5.1 of~\cite{hllvuln}.
-In that paper, the authors target the Redis~\cite{redis} open-source implementation of HLL and perform the attack against an empty sketch.
-Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780''~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality.
-The authors present this attack as being in their M2 (our S1) scenario, without providing a detailed explanation.
-By looking at the Redis open-source code, we were able to obtain full details of the implementation that any attacker could obtain, and thus argue that based on this knowledge, the adversarial scenario is actually M1 (S2 in this work).
-This makes the comparison between the results of~\cite{hllvuln} and our S2 results appropriate.
-The S2 adversary following our strategy is able to add 125,000 distinct items while keeping the cardinality estimate at 367.
-Compared to~\cite{hllvuln}, we can insert twice as many items and reduce the cardinality estimate by a factor of 340 instead of 5.
+In that paper, the authors target the Redis~\cite{redis} open-source implementation of HLL and perform the attack against an empty sketch filled with adversarial inputs only.
+Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780''~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality after 4 iterations (resets).
+The S1 adversary following our strategy is able to add 977 and 97,433 distinct items while keeping the cardinality estimate respectively at 1 and 126.
+Compared to~\cite{hllvuln}, we can reduce the cardinality estimate by a factor of almost 1,000 instead of 5 at the cost of more resets.
 
@@
 \caption{Attack results, averaged over 30 iterations.}
-\begin{tabular}{| m{8.5em} | m{4em} | m{4em} | m{4em} | m{4em} |}
+     \makebox[\linewidth]{
+\begin{tabular}{| m{9.8em} | m{4em} | m{4.2em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
     \hline
-    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
-    \textbf{Initial data?} & yes & no & yes & no \\ \hline
-    \textbf{Original Card. Est.} & 1,000 & 0 & 1,005 & 0 \\ \hline
-    \textbf{\# Items added} & 3,655,744 & 3,655,740 & 6,090,003 & 14,220 \\ \hline
-    \textbf{Final Card. Est.} & 1'011 & 367 & 1,005 & 1 \\ \hline
-\end{tabular}
+    \textbf{Scenario} &  \multicolumn{2}{c|}{S1} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
+    \textbf{\# Targeted buckets} & 1 & 100 & - & 1 & - & 1 \\ \hline
+    \textbf{Initial data?} & no & no & yes & no & yes & no \\ \hline
+    \textbf{Original Card. Est.} & 0 & 0 & 1,000 & 0 & 1,005 & 0 \\ \hline
+    \textbf{\# Items added} & 28,566 & 2,856,100 & 3,655,744 & 28,566 & 6,090,003 & 28,566 \\ \hline
+    \textbf{Final Card. Est.} & 1 & 126 & 1'011 & 1 & 1,005 & 1 \\ \hline
+\end{tabular}}
 \label{table:tab1}
@@
 \caption{Attack results, averaged over 30 iterations, in the setting of \cite{hllvuln}.}
-\begin{tabular}{| m{8.5em} | m{4em} | m{4em} | m{4em} | m{4em} |}
+     \makebox[\linewidth]{
+\begin{tabular}{| m{9.8em} | m{4em} | m{4.2em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
     \hline
-    \textbf{Scenario} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
-    \textbf{Initial data?} & yes & no & yes & no \\ \hline
-    \textbf{Original Card. Est.} & 1,004 & 0 & 998 & 0 \\ \hline
-    \textbf{\# Items added} & 124,940 & 125,085 & 204,669 & 501 \\ \hline
-    \textbf{Final Card. Est.} & 1,058 & 367 & 998 & 1 \\ \hline
-\end{tabular}
+    \textbf{Scenario} & \multicolumn{2}{c|}{S1} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
+    \textbf{\# Targeted buckets} & 1 & 100 & - & 1 & - & 1 \\ \hline
+    \textbf{Initial data?} & no & no & yes & no & yes & no \\ \hline
+    \textbf{Original Card. Est.} & 0 & 0 & 1,004 & 0 & 998 & 0 \\ \hline
+    \textbf{\# Items added} & 977 & 97,433 & 124,940 & 979 & 204,669 & 988 \\ \hline
+    \textbf{Final Card. Est.} & 1 & 126 & 1,058 & 1 & 998 & 1 \\ \hline
+\end{tabular}}
 \label{table:tab2}
@@
 For example, would the imposed ID setting (like Facebook) solve the issue but still be convenient for the various HLL applications?
-Also left for future work is the task of extending our attacks to HLL variants such as redis alternative using maximum likelihood~\cite{newhll}, HLL++~\cite{hllpratice} and sliding HLL~\cite{slidinghll}.
+Also left for future work is the task of extending our attacks to HLL variants such as sliding HLL~\cite{slidinghll}, HLL++~\cite{hllpratice} and Redis alternative using the maximum likelihood approach from~\cite{newhll}.
 

### 2020-09-22 docs/HLL_exp_bad/main.tex
working through the paper

@@
 \title{HyperLogLog: Exponentially Bad in Adversarial Settings}
-\author{Mathilde Raynal, Kenneth G. Paterson\\
+\author{Kenneth G. Paterson and Mathilde Raynal\\
 \vspace{3mm}
 Department of Computer Science, ETH Zurich\\
-{\tt mraynal@student.ethz.ch, kenny.paterson@inf.ethz.ch}}
+{\tt kenny.paterson@inf.ethz.ch, mraynal@student.ethz.ch}}
 
@@
 \section{Introduction}
-Naive approaches to computing the cardinality of a big data set, such as sorting the elements or simply maintaining the set of unique elements seen, are impractical at scale.
+Naive approaches to computing the cardinality of large data sets, such as sorting the elements or simply maintaining the set of unique elements seen, are impractical at scale and in streaming settings.
 The HyperLogLog algorithm (HLL henceforth)~\cite{hll} is today the most widely used cardinality estimator.
@@
 Very recently, Reviriego and Ting~\cite{hllvuln} initiated work in this direction, showing that in specific attack settings, the cardinality estimate made by HLL could be modified through input selection.
-In this paper, we present a complete analysis of HLL in adversarial setting, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
-In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of $O(2^t)$ where $t$ is the number of bits of flexibility in the input).
+In this paper, we present a complete analysis of HLL in the adversarial setting, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
+In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of $O(2^t)$ where $t$ is the number of bits of flexibility in the adversary's input).
 
 The rest of the paper is organised as follows.
-Section~\ref{sec:overview} gives an overview of the HyperLogLog algorithm.
-A brief summary of the related works can be found in Section \ref{sec:related}.
+Section~\ref{sec:overview} gives an overview of HLL.
+A brief summary of  related work can be found in Section \ref{sec:related}.
 Section~\ref{sec:attacks} presents our different adversarial models.
@@
 Each estimate is stored in its own register, that we call a bucket.
-The collection of buckets is referred to as the HLL \emph{sketch}.
-To map a value $x$ to one of $2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$.
+We refer to the collection of buckets and their contents as the HLL \emph{sketch}.
+To map a value $x$ to one of $m=2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$.
 The output length of $h(\cdot)$ is thus $n+\ell$, fixed to 32 in~\cite{hll}.
-Estimators from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ is empirically computed to correct a systematic multiplicative bias.
+estimates from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ is empirically computed to correct a systematic multiplicative bias.
 
-Relatively small and big cardinalities are handled separately to avoid eventual distortions, as shown in Fig.~\ref{fig:hll}.
-When the raw cardinality estimate is below $\frac{5}{2}\times 2^n$, the Hit Counting algorithm of Whang et al.~\cite{hitcounting} is used to produce the final cardinality estimator, looking at the number of empty buckets (whose register is 0) instead of the number of leading zeros in the stream.
+Relatively small and big cardinalities are handled separately to correct for systematic distortions, as shown in Fig.~\ref{fig:hll}.
+When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang et al.~\cite{hitcounting} is used to produce the final cardinality estimate, looking at the number of empty buckets (whose register is 0) instead of the number of leading zeros in the stream.
 A similar correction is applied for large cardinalities in order to take into account the likelihood of hashing collisions.
-Some works building on top of~\cite{hll} replace the Hit counting algorithm by other schemes and use a hash function with a larger output domain to reduce the chances of collision and thus get rid of the need for large cardinality range correction.
+As a result, HLL is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
+
+Some works building on~\cite{hll} replace the Hit Counting algorithm by other schemes and use a hash function with a larger output domain to reduce the chances of collision and thus get rid of the need for large cardinality range correction.
 For example, \cite{hllpratice} uses a 64-bit hash function and Linear Counting along with bias correction in the small cardinality range.
-In this report we target the original paper's implementation reported in Fig.~\ref{fig:hll} and leave for future work the analysis of its variations.
-
-As result, HLL is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
+In this report we target the HLL formulation in the original paper~\cite{hll} as represented in Fig.~\ref{fig:hll}, and leave for future work the analysis of its variants.
 
@@
 \begin{flushleft}
-\textit{Let $h : D \rightarrow \{0, 1\}^{32}$ hash data from $\mathcal{D}$ to binary $32$–bit words.\\
+\textit{Let $h : \mathcal{D} \rightarrow \{0, 1\}^{32}$ hash data from some domain $\mathcal{D}$ to binary $32$-bit strings.\\
 Let $\rho(s)$ be the position of the leftmost $1$-bit of $s$: e.g., $\rho(1...) = 1, \rho(0001...) = 4, \rho(0^K) = K + 1.$\\}
@@
 \vspace{0.2em}
-\textbf{compute} $E := \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$; \hfill \textit{\{the “raw” HyperLogLog estimate\}}\\
+\textbf{compute} $E := \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$; \hfill \textit{\{the raw HyperLogLog estimate\}}\\
 \vspace{0.2em}
@@
     \hspace{2em}  \textbf{let} $V$ be the number of registers equal to $0$;\\
-    \hspace{2em}  \textbf{if} $V \neq 0$ \textbf{then set} $E^*:= m $ln$(m/V)$ \textbf{else set} $E^*:= E$; \hfill \textit{\{small range correction\}}\\
+    \hspace{2em}  \textbf{if} $V \neq 0$ \textbf{then set} $E^*:=  m\cdot \ln(m/V)$ \textbf{else set} $E^*:= E$; \hfill \textit{\{small range correction\}}\\
 \textbf{if} $E \le \frac{1}{30}2^{32}$ \textbf{then}\\
@@
 
-Regardless of the sensitivity of the above examples, none of the relevant papers are clear about their threat model, especially for the HLL component.
+Regardless of the sensitivity of the above examples, none of the above-cited papers are clear about their threat model, especially for the HLL component.
 They are all potentially vulnerable to the attacks we present below.
@@
 
-The work of~\cite{cardestprivacy} studies the privacy properties of HLL in two different attack scenarios, referred to as \emph{insider} and \emph{external}.
-The attack target of~\cite{cardestprivacy} is different from ours, being concerned with loss of privacy, whereas we care about cardinality estimation.
+The work of Desfontaines \emph{et al.}~\cite{cardestprivacy} studies the privacy properties of HLL in two different attack scenarios, referred to as \emph{insider} and \emph{external}.
+The attack target of~\cite{cardestprivacy} is different from ours, being concerned with loss of privacy, whereas we care about manipulating cardinality estimation.
 It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
 In Section~\ref{sec:attacks} we will discuss the two attack scenarios from~\cite{cardestprivacy} in more detail.
-In particular, we will show that an attacker who can insert items into the HLL and who can access the HLL cardinality estimate (via typical APIs in HLL implementations) as permitted in the external scenario can actually realise the same attacks as the insider attacker can, at least insofar as cardinality estimate manipulation is concerned.
+In particular, we will show that an adversary who can insert items into the HLL and who can access the HLL cardinality estimate (via typical APIs in HLL implementations) as permitted in the external scenario of~\cite{cardestprivacy} can actually realise the same attacks as the insider adversary can, at least insofar as cardinality estimate manipulation is concerned.
 
 In~\cite{hllvuln}, Reviriego and Ting exploit a vulnerability of HLL to manipulate the cardinality estimate, leading to a five-fold reduction in the estimate compared to the true cardinality.
-Their attack model can be considered to be quite artificial, since it gives the attacker the ability to test whether inserting an item \emph{would} increase the cardinality estimate, without actually having to insert the item.
-After some clarifications obtained from personal correspondence with the authors, we finally agree that such setup might be possible under certain circumstances.
-See further discussion in Section~\ref{sec:attacks}.
+We consider the attack model used in~\cite{hllvuln} to be quite artificial, since it gives the adversary the ability to test whether inserting an item \emph{would} increase the cardinality estimate, without actually having to insert the item.
+After some clarifications obtained through personal correspondence with the authors of~\cite{hllvuln}, we understand that such a setup might be possible under certain circumstances, for example where the adversary has access to an identical ``shadow'' device; see further discussion in Section~\ref{sec:attacks}.
 
-Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted sketch, either as replacement (but at the cost of losing mergeability) or in addition to an unsalted sketch (inducing a memory overhead).
+Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted sketch, either as a replacement for (but at the cost of losing mergeability) or in addition to (inducing a memory overhead) an unsalted sketch.
 %This might not be a convenient solution for some applications, they thus encourage further research.
+
+Clayton \emph{et al.}~\cite{CCS:ClaPatShrPS19} provide a provable-security treatment of probabilistic data structures in adversarial environments more generally.
+They use their formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures, but not HLL.
+It is an interesting open problem to use their framework to develop a formal understanding of how HLL can be made secure in the adversarial setting.
 
@@
 
-\subsection{Adversarial Setting}\label{sec:set}
+\subsection{Adversarial Scenarios}\label{sec:set}
 We model four distinct adversarial scenarios based on the adversary's knowledge and capabilities.
-To simulate real-life settings, we consider both cases when the sketch is shared and receives inputs from other honest users but also when inputs are under adversarial control only.
-It is safe to assume that the adversary knows the setting being used. 
-Those two settings differ mainly in the fact that the adversary is attacking a sketch that already contains other users' data in the former option, but is empty in the latter one.
+To simulate real-life settings, we consider both cases when the sketch is shared and receives inputs from other honest users but also when inputs are under the sole control of the adversary.
+It is reasonable to assume that the adversary knows which case it is in. 
+These two settings differ mainly in the fact that the adversary is attacking a sketch that may already contain other users' data in the former option, but is empty in the latter.
 
-Among the scenarios, the \textit{insider} setup of~\cite{cardestprivacy} that we present as S4 makes the strongest assumptions about the adversarial capabilities since it assumes that the attacker has a perfect view on the sketch.
+Among the scenarios, the \textit{insider} setup of~\cite{cardestprivacy} that we present as S4 makes the strongest assumptions about the adversarial capabilities since it assumes that the adversary has a perfect view of the sketch (at some point in time).
 
@@
 \item[S1:] The adversary does not know the details of the target HLL implementation but can access a copy of the HLL instance as a user.
-He can use the API to insert elements into this shadow sketch, get its cardinality estimate, and reset it to its native empty state.
-The attacker does not know the state of the targeted HLL sketch, meaning that he does not have the list of elements previously inserted into it by other users.
-\item[S2:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use, but is otherwise ``blind''.
+It can use the API to insert elements into this shadow sketch, get its cardinality estimate, and reset it to its native empty state.
+The adversary does not know the state of the targeted HLL sketch, meaning that it does not have the list of elements previously inserted into it by other users.
+\item[S2:] The adversary has access to the details of the HLL implementation, i.e.\ it knows the number of buckets $m=2^n$ and the hash function $h$ in use, but is otherwise ``blind''.
 In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack.
 This kind of attack scenario is appropriate in the context of the port scanning attack of~\cite{portscanhll}.
-\item[S3:] The adversary has access to the details of the HLL implementation, i.e., the adversary knows the number of buckets $2^n$ and the hash function $h$ in use.
+\item[S3:] The adversary has access to the details of the HLL implementation, i.e.\ it knows the number of buckets $m=2^n$ and the hash function $h$ in use.
 Additionally, it can access and interact with the sketch via an API provided by the sketch owner, allowing it to insert items into the HLL sketch and ask for the HLL's cardinality estimate at any point in time.
+However, it does not know the individual bucket contents.
 \item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch.
 It then causes chosen inputs to be inserted into the HLL sketch.
-An S4 adversary can compute a lower bound on the HLL cardinality estimate based on its snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
+Note that, using these capabilities, an S4 adversary can compute a lower bound on the HLL cardinality estimate based on its initial snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
 \end{itemize}
@@
 
-We start by highlighting the fact that we do not consider scenario S1 typical of HLL deployments.
-In a networking application, the authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the attacker learns which machine is used to do the monitoring, buys the same, and uses it as the shadow device.
+\subsection{Discussion of Adversarial Scenarios}\label{sec:scenarios-discussion}
+
+We do not consider scenario S1 typical of HLL deployments.
+In a networking application, the authors of~\cite{hllvuln} argue that the assumptions of S1 could be fulfilled if the adversary learns which machine is used to do the monitoring, buys the same, and uses it as the shadow device.
 It is possible that the provider uses a different salt or key in the hash per device.
 This would make any pre-computation on the items using the shadow device irrelevant, and considerably weaken the attack presented in Section 3.2 of~\cite{hllvuln}.
-Hence, we assume for the rest of this work that shadow and targeted devices have the same internals and security parameters.
-If the HLL device receives inputs from several users, the attack from Section 3.2 of~\cite{hllvuln} is the easiest to carry out.
-It works as follow: "after each insertion [in the shadow device], elements that do not increase the cardinality estimate are retained".
-Iterating through the process several times refines the items' filtering before inserting them into the targeted device.
-We explore in this work the setup where inputs are coming only from the adversary, and propose an attack that provides better results than the one we just presented.
+Hence we assume for the rest of this work that, in scenario S1, shadow and targeted devices have the same internals and security parameters.
+If the HLL device receives inputs from several users, the attack from Section 3.2 of~\cite{hllvuln} is then the easiest to carry out.
+It is described in~\cite{hllvuln} as follows: "after each insertion [in the shadow device], elements that do not increase the cardinality estimate are retained".
+The idea is to iterate through the process of trial insertion several times to refine the list of items, before inserting what remains into the targeted device.
+We explore in this work scenario S1 in the special case where inputs come only from the adversary, and propose an attack that provides better results than the one presented in~\cite{hllvuln}.
 
@@
 Indeed, an adversary can easily recover the contents of each bucket in the sketch simply by analysing which inputs increase the cardinality estimate.
-To find the estimate held in a targeted bucket, the attacker gradually inserts items $x$ with an incrementing number of leading zeros in the $\ell$ rightmost bits of $h(x)$ whilst holding the $n$ bits determining the bucket index constant, until the cardinality estimate of the sketch increases.
-Once an increase occurs, the attacker knows that the maximum number of leading zeros observed in that bucket was just updated to a value that can be recovered by looking at the last inserted item.
-This assumes the attacker has sufficient flexibility in its choice of inputs $x$ so that the relevant bit conditions on $h(x)$ can be forced.
-The attacker can determine all bucket contents with, in the worst case, $O(2^{n + \ell})$ pre-computation and $O(\ell \cdot 2^{n})$ insert and cardinality estimate queries, and needs about $n + \ell$ bits of flexibility assuming $h$ behaves like a random function.
-The pre-computation and query complexity required can be much less if the HLL sketch is only moderately loaded.
-A version of this attack can be carried out to elevate an external adversary to a full-strength insider adversary as per~\cite{cardestprivacy} (instead of our slightly weaker S4 adversary), since it can simply be done repeatedly each time the external adversary wants to know the complete internals of the HLL sketch.
+To find the estimate held in a targeted bucket, the adversary  inserts a sequence of items $x$ with an incrementing number of leading zeros in the $\ell$ rightmost bits of $h(x)$ whilst holding the $n$ bits determining the bucket index constant, until the cardinality estimate of the sketch increases.
+Once an increase occurs, the adversary knows that the maximum number of leading zeros observed in that bucket was just updated to a value that can be recovered by looking at the last inserted item.
+This assumes the adversary has sufficient flexibility in its choice of inputs $x$ so that the relevant bit conditions on $h(x)$ can be forced.
+The adversary can determine all bucket contents with, in the worst case, $O(2^{n + \ell})$ pre-computation and $O(\ell \cdot 2^{n})$ insert and cardinality estimate queries, and needs about $n + \ell$ bits of flexibility assuming $h$ behaves like a random function.
+The pre-computation and query complexity required can be much less if the HLL sketch is only moderately loaded, since then most buckets will store values corresponding to inputs with few leading zero bits.
+A version of this attack can be carried out to elevate an external adversary to a full-strength insider adversary as per~\cite{cardestprivacy} (instead of our slightly weaker S4 adversary), since it can simply be done each time the external adversary wants to obtain a snapshot of the internals of the HLL sketch.
 Obvious optimisations can be carried out to reduce the query complexity each time the attack is run (since the bucket contents can only increase over time, some queries are redundant; moreover, the pre-computation is only needed once and is moderate given typical HLL parameters, e.g.\ $n+\ell = 32$).
 
-Since an attacker can escalate from S3 to S4, we focus on scenarios S1, S2 and S4 in what follows.
+Since an adversary can quite efficiently escalate from S3 to S4, we focus on scenarios S1, S2 and S4 in what follows.
 We first present an attack in scenario S1, providing much stronger results than the corresponding attack of~\cite{hllvuln} (in their equivalent scenario M2).
@@
 \subsection{Manipulating the HLL Cardinality Estimate}
-\noindent\paragraph{Scenario S1} The goal of the attacker is to infer some information about targeted items' hash values, precisely their bucket mappings, without knowing neither $h$ nor the salt in use (if any).
-With this knowledge, the adversary is able to retain items that are mapped to a subset of $b$ buckets out of the $2^n$, and minimize the resulting estimated cardinality.
-To achieve this, the attacker can take advantage of the Hit Counting algorithm initial phase to recover the number of non-empty buckets (whose register is not 0) after inserting few items.
-Indeed, in the small cardinality range, the cardinality estimator E* is computed as E*$=2^n$ln$(2^n/V)$ with $V$ the number of empty buckets.
-For typical HLL parameters, e.g.\ $n = 8$, the exact value of $V$, hence $b$, can be recovered from E*.
-Moreover, when inserting an item in the small cardinality range, the adversary will see a increase in the cardinality estimate if and only if $V$ is modified.
-This means that the number of empty buckets decreased by 1, therefore that the freshly inserted item is mapped to a bucket that was not hit before.
 
-Thus, if the adversary blindly picks the first $b$ buckets to be hit as its targeted buckets, he can then discard all items whose insertion causes the cardinality estimator to increase from $b+k$ to $b+k+1$ for $k \in \mathbb{N}$.
-In other words, when $b$ buckets are already hit and the insertion of an item causes the cardinality estimator to increase, the attacker can discard this item as he knows that a new bucket (other than one of the $b$ buckets to be filled) switched to a non-empty state.
-The adversary can keep inserting items and discarding some while in the small cardinality range.
+Although we consider scenario S1 to be quite artificial, we will commence our analysis there, since the techniques we present are more broadly applicable in S2 and S4.
 
-As soon as the cardinality estimator exceeds $\frac{5}{2}\times 2^n$, the adversary can assume that it corresponds to the raw HLL cardinality estimate with high probability.
-In this case, no more useful information can be inferred about the bucket mappings, so the attacker has to reset the sketch to the empty state and start again with the non-discarded items.
-The adversary must pay attention not to shuffle the non-discarded items in between iterations, as the $b$ buckets to be filled are blindly determined by the first inserted items, and must stay the same throughout the attack.
-After several iterations, the adversary discards all items mapped to unwanted buckets, such that $b$ buckets are filled and the others left untouched.
+\noindent\paragraph{Scenario S1} The goal of our adversary is to infer some information about targeted items' hash values, precisely their bucket mappings, knowing neither $h$ nor the salt in use (if any).
+Using this information, the adversary is able to retain items that are mapped to a subset of $b$ buckets out of the $2^n$, and minimize the resulting estimated cardinality.
+We assume the adversary either has flexibility in choosing the items to insert, or is presented with a static list of items at the start of its attack.
+We also assume that the shadow HLL sketch and the target HLL sketch are both initially empty.
 
-The value of $b$ has to be picked smartly to keep the final cardinality estimate low, as the registers must hold values potentially very high.
-$b$ is going to be chosen such that even in the worst case scenario where all register values $M[j]$ hold the the maximal value of $l+1$, the Hit Counting algorithm is going to be used to produce the final estimation.
-So $b$ must satisfy:
-\begin{align}
-E:=\alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1} & \le \frac{5}{2}m \nonumber\\
-\alpha_mm^2\cdot((m-b)\cdot2^{-0}+b\cdot2^{-(l+1)})^{-1} & \le \frac{5}{2}m \nonumber\\
-b & \le m\cdot(1-\frac{2\alpha_m}{5})\cdot(1-2^{-(l+1)})^{-1} \label{eqn:1} 
-\end{align}
-with $m=2^n$.
-For example with $n = 8$ and $l = 24$, $b$ can take any value up to 184.
-The initial set of items to be filtered can be built by varying the flexible fields of the items $x$ the adversary is trying to insert.
+Our adversary takes advantage of the use of HLL's small cardinality Hit Counting correction to recover the number of non-empty buckets (whose register does not contain value 0) after inserting a few items into the initially empty shadow HLL sketch.
+Indeed, in the small cardinality range, the cardinality estimate $E^*$ is computed as $E^*= m \cdot \ln (m/V)$ with $V$ the number of empty buckets.
+For typical HLL parameters, e.g.\ $n = 8, m =256$, the exact value of $V$, hence $b$, can be recovered from $E^*$.
+Moreover, when inserting an item in the small cardinality range, the adversary will see an increase in the cardinality estimate for the shadow HLL sketch if and only if $V$ is modified.
+This in turn means that the number of empty buckets in the shadow HLL sketch decreased by 1, and therefore that the freshly inserted item was mapped to a bucket that was not hit before.
+
+The adversary blindly picks the first $b$ buckets to be hit by simply inserting items into the shadow HLL until the cardinality estimate indicates that $b$ buckets have been touched.
+It puts these items in a final target list ${\cal L}$.
+It continues to insert items, discarding all items whose insertion causes the cardinality estimate for the shadow HLL sketch to increase from $b+k$ to $b+k+1$ for $k \in \mathbb{N}$ (here we allow some slack in the total number of buckets that have been hit in order to reduce the number of resets of the shadow device that are required in the attack).
+In other words, when $b+k$ buckets have already been hit and the insertion of an item causes the cardinality estimate to increase further, the adversary can discard this item as it definitely knows that a new bucket (that is not one of the $b$ targeted buckets) switched to a non-empty state.
+On the other hand, it adds to the list ${\cal L}$ any item that does \emph{not} cause the cardinality estimate to increase.
+Such items may have hit one of the initial $b$ buckets, or one of the later buckets; at this stage the adversary cannot be certain.
+The adversary can keep inserting items and discarding those which cause the cardinality estimate to increase while in the small cardinality range.
+
+As soon as the cardinality estimate exceeds $\frac{5}{2}m$, the adversary can assume that the shadow HLL sketch's output corresponds to the raw HLL cardinality estimate with high probability.
+In this case, no more useful information can be inferred about the bucket mappings in the shadow HLL sketch, so the adversary resets the shadow HLL sketch to the empty state and starts again with the list ${\cal L}$ of non-discarded items obtained so far.
+Reinserting these items will (eventually) determine whether they hit the original $b$ buckets or later buckets, and hence whether they should be retained in ${\cal L}$ or discarded.
+Once ${\cal L}$ is reprocessed (and assuming the cardinality estimate does not exceed $\frac{5}{2}m$), the adversary can start to insert fresh items, as previously described.
+The adversary must be careful not to shuffle the non-discarded items in between iterations, as the $b$ buckets to be filled are blindly determined by the first inserted items, and must stay the same throughout the attack.
+
+After sufficiently many iterations of this procedure, the adversary will have created a list of items ${\cal L}$ such that, when inserted into the \emph{target} HLL, create a sketch in which $b$ target buckets are filled and the others left empty.
+With an appropriate choice of $b$ (as discussed immediately below), the final cardinality estimate for the target HLL sketch will be $E^*= m \cdot \ln (m/V)$ where $V = m-b$ is the number of empty buckets.
+This can be rewritten as $E^*= m \cdot \ln (1+b/(m-b))$.
+
+The value of $b$ has to be picked carefully in order to keep the shadow and target HLL cardinality estimates low, as the registers in both sketches may hold values that (potentially) become very high as more items are inserted.
+The trick is select $b$ so that, even in the worst case scenario where all register values $M[j]$ hold the maximal value of $\ell+1$, the Hit Counting algorithm will still be used to produce the cardinality estimates.
+Noting that the raw estimate $E = \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$ increases with increasing $M[j]$, and that we have $b$ buckets each contributing, in the worst case, $2^{-(\ell+1)}$ to the sum (and hence $m-b$ buckets which contribute $2^{0}=1$ to the sum) we see that for the Hit Counting algorithm still to be used, we require:
+\[
+\alpha_mm^2\cdot\left((m-b)+b\cdot2^{-(\ell+1)}\right)^{-1} \le \frac{5}{2}m.
+\]
+Solving for $b$, we obtain:
+\[
+b \le b_{\max} := \left\lfloor m\cdot\left(1-\frac{2\alpha_m}{5}\right)\cdot\left(1-2^{-(\ell+1)}\right)^{-1} \right\rfloor. \label{eqn:1} 
+\]
+
+For example with $n = 8$ (so $m=256$) and $\ell = 24$, we find that $b_{\max}=184$.
+
+A surprising implication of this analysis is that, for the HLL algorithm as described in Figure~\ref{fig:hll}, provided $b \leq b_{\max}$, no matter which fixed set of $b$ buckets are hit, and no matter how many leading zero bits the items have, the final cardinality estimate will not exceed $m \cdot \ln (1+b/(m-b))$.
+We will exploit this in the sequel in scenarios S2 and S4.
+For now, we note that setting, say, $b=128$ with $n = 8$ and $\ell = 24$ would result in a final cardinality estimate of $256 \ln 2$ in the above attack in Scenario S1, irrespective of how long the list ${\cal L}$ of inserted items is.
+Moreover, setting $b=128$ would mean that half of all items available to the adversary could be inserted into the target HLL.
+
+%The initial set of items to be filtered can be built by varying the flexible fields of the items $x$ the adversary is trying to insert.
 %so that the adversary can ``hide'' $2^{t-n}$ elements while keeping the HLL cardinality estimate to 1.
@@
 \noindent\paragraph{Scenario S2} The adversary knows $h$ so it can adjust fields in the item $x$ it is trying to insert and keep only those for which the $\ell$ rightmost bits of $h(x)$ start with a one bit.
-In this case, the number of leading zero bits computed by the estimator of the destination bucket is always zero, so the estimator of this bucket is unlikely to be updated and the final averaged approximate cardinality is likely to be left unchanged.
+In this case, the number of leading zero bits computed by the estimate of the destination bucket is always zero, so the estimate of this bucket is unlikely to be updated and the final averaged approximate cardinality is likely to be left unchanged.
 For example, in the scenario of~\cite{portscanhll}, the adversary could change the source port of its IP packet until it is satisfied with its hash value.
@@
 In other words, if $h(x)$ is mapped to the $i$-th bucket with estimated cardinality $2^{M[i]}=2^{c_i + 1}$ (because a value with $c_i$ leading zeroes was previously observed for this bucket), then, instead of the strict restriction of having a one bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a one bit in any of the first $c_i+1$ positions in this substring.
-In case the adversary hits an empty bucket (for which the estimator is 0), it must skip that bucket completely.
-Under these conditions, the estimators in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.
+In case the adversary hits an empty bucket (for which the estimate is 0), it must skip that bucket completely.
+Under these conditions, the estimates in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.
 
 A special case arises when the HLL is completely empty at the beginning of the attack.
-Here the attacker is forced to increase the estimate in at least one bucket.
+Here the adversary is forced to increase the estimate in at least one bucket.
 Once again, the adversary can choose to fill $b$ buckets following the strategy described in S2, leaving all the other buckets untouched.
@@
 The adversary can expect to ``hide''  $2^{t-1}$ distinct items without increasing the HLL cardinality estimate at all, under the assumption that all buckets in the HLL sketch already hold a value $c_i \ge 1$.
-On the other hand, if some buckets are empty (but the adversary does not know which ones in the blind setting of S2), then we may expect a moderate increase in the HLL cardinality estimator.
+On the other hand, if some buckets are empty (but the adversary does not know which ones in the blind setting of S2), then we may expect a moderate increase in the HLL cardinality estimate.
 Notice that in all cases, provided $t$ is large enough, the adversary can keep the cardinality estimate linear in the total number of buckets ($2^n$) while the expected value of the true cardinality, $2^{t-1}$, can be made as large as the adversary pleases.
@@
 Assume $h(x)$ is mapped to bucket $i$.
-The probability that $h(x)$ does not increase the estimator is the probability that the  rightmost $\ell$ bits of $h(x)$ have $c_i$ or less leading zeroes; this probability is equal to $1-\frac{1}{2^{c_i+1}}$ assuming $h$ behaves like a random function.
-Averaging over all $2^n$ buckets (and using the fact that the bucket choice is uniformly random since the bits of $h$ are independent and uniformly random), the probability that input $x$ does not increase the estimator is given by:
+The probability that $h(x)$ does not increase the estimate is the probability that the  rightmost $\ell$ bits of $h(x)$ have $c_i$ or less leading zeroes; this probability is equal to $1-\frac{1}{2^{c_i+1}}$ assuming $h$ behaves like a random function.
+Averaging over all $2^n$ buckets (and using the fact that the bucket choice is uniformly random since the bits of $h$ are independent and uniformly random), the probability that input $x$ does not increase the estimate is given by:
 \begin{eqnarray*}
@@
 For a fixed $b$, an adversary who has $t$ bits of flexibility in its inputs is then able to insert exponentially (in $t$) many input values into the HLL sketch while keeping the cardinality estimate constant.
-Any value below the upper-bound specified in eq. (4.1) can be chosen for $b$ based on the setup requirements. For example if the reset operation is costly, the attacker can adjust $b$ accordingly.
+Any value below the upper-bound specified in eq. (4.1) can be chosen for $b$ based on the setup requirements. For example if the reset operation is costly, the adversary can adjust $b$ accordingly.
 

### 2020-09-25 docs/HLL_exp_bad/main.tex
rewriting

@@
 \documentclass[11pt]{article}
-\usepackage{graphicx, amsmath, amsthm, amssymb, setspace, array, url, hyperref, xspace}
+\usepackage{graphicx, amsmath, amsthm, amssymb, setspace, array, url, hyperref, xspace, todonotes, pifont}
 \usepackage{a4wide}
@@
 \usepackage[T1]{fontenc}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+
+\newtheorem{lemma}{Lemma}
 
@@
 Computing the count of distinct elements in massive data sets is a common task but naive approaches are memory-expensive.
-The HyperLogLog (HLL) algorithm (Flajolet \emph{et al.}, 2007) estimates a data stream's cardinality while using significantly less memory than a naive approach at the cost of some accuracy.
+The HyperLogLog (HLL) algorithm (Flajolet \emph{et al.}, 2007) estimates a data set's cardinality while using significantly less memory than a naive approach at the cost of some accuracy.
 This trade-off makes the HLL algorithm very attractive for a wide range of applications such as database management and network monitoring, where an exact count may not be needed.
@@
 In this paper, we present a complete analysis of HLL in the adversarial setting, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
-In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of $O(2^t)$ where $t$ is the number of bits of flexibility in the adversary's input).
+In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of ${\cal O}(2^t)$ where $t$ is the number of bits of flexibility in the adversary's input).
 
@@
 Section~\ref{sec:attacks} presents our different adversarial models.
-There, we also give attacks and evaluate their impacts under the different adversarial models.
+In Section~\ref{sec:manip}, we describe attacks and evaluate their impacts under the different adversarial models.
 The experimental results are discussed in Section~\ref{sec:exp} and Section~\ref{sec:conclusions} provides our conclusions and some ideas for future work.
@@
 The output length of $h(\cdot)$ is thus $n+\ell$, fixed to 32 in~\cite{hll}.
-estimates from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ is empirically computed to correct a systematic multiplicative bias.
+Estimates from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ is empirically computed to correct a systematic multiplicative bias.
 
 Relatively small and big cardinalities are handled separately to correct for systematic distortions, as shown in Fig.~\ref{fig:hll}.
-When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang et al.~\cite{hitcounting} is used to produce the final cardinality estimate, looking at the number of empty buckets (whose register is 0) instead of the number of leading zeros in the stream.
+When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate, looking at the number of empty buckets (whose register is 0) instead of the number of leading zeros in the stream.
 A similar correction is applied for large cardinalities in order to take into account the likelihood of hashing collisions.
@@
 
-The work of Desfontaines \emph{et al.}~\cite{cardestprivacy} studies the privacy properties of HLL in two different attack scenarios, referred to as \emph{insider} and \emph{external}.
+Desfontaines \emph{et al.}~\cite{cardestprivacy} studied the privacy properties of HLL in two different attack scenarios, referred to as \emph{insider} and \emph{external}.
 The attack target of~\cite{cardestprivacy} is different from ours, being concerned with loss of privacy, whereas we care about manipulating cardinality estimation.
@@
 \begin{itemize}
-\item[S1:] The adversary does not know the details of the target HLL implementation but can access a copy of the HLL instance as a user.
+\item[S1:] The adversary does not know the details of the target HLL implementation but can access a shadow copy of the HLL instance via its API.
 It can use the API to insert elements into this shadow sketch, get its cardinality estimate, and reset it to its native empty state.
-The adversary does not know the state of the targeted HLL sketch, meaning that it does not have the list of elements previously inserted into it by other users.
-\item[S2:] The adversary has access to the details of the HLL implementation, i.e.\ it knows the number of buckets $m=2^n$ and the hash function $h$ in use, but is otherwise ``blind''.
+The adversary does not know the state of the targeted HLL sketch, meaning in particular that it does not have the list of elements previously inserted into it by other users.
+It can insert items into the target HLL sketch, but is otherwise ``blind''.
+\item[S2:] The adversary has access to the details of the HLL implementation, i.e.\ it knows the number of buckets $m=2^n$ and the hash function $h$ in use.
+It can insert items into the HLL sketch, but is otherwise blind.
 In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack.
@@
 \item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch.
-It then causes chosen inputs to be inserted into the HLL sketch.
+It can access and interact with the sketch via an API provided by the sketch owner, as per scenario S3.
 Note that, using these capabilities, an S4 adversary can compute a lower bound on the HLL cardinality estimate based on its initial snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
@@
 S4 is a slightly weaker version of the \textit{insider} attack scenario in~\cite{cardestprivacy}, the difference being that an insider adversary in~\cite{cardestprivacy} has continuous access to the HLL sketch internals including all bucket values, while our adversary in S4 only has access to a snapshot of the HLL sketch internals at the start of its attack.
+
+The adversarial capabilities in each of scenarios S1 to S4 are summarised in Table~\ref{tab:scenarios}.
+
+\begin{table}[h]
+\centering
+\caption{Adversarial capabilities in scenarios S1, S2, S3 and S4.}
+\medskip
+     \makebox[\linewidth]{
+\begin{tabular}{| m{11em} | m{2em} | m{2em} | m{2em} | m{2em} |}
+    \hline
+    {Scenario} 							&  {S1} & {S2} & {S3} & {S4} \\ \hline
+   	{API for shadow device} 		& \cmark & \xmark & \xmark & \xmark \\ \hline
+    {Insertion of  items} 				& \cmark & \cmark & \cmark & \cmark \\ \hline
+    {$h$ and $m$ known}			& \xmark & \cmark & \cmark & \cmark \\ \hline
+    {Get cardinality estimate} 		& \xmark & \xmark & \cmark & \cmark \\ \hline
+    {Get snapshot} 						& \xmark & \xmark & \xmark & \cmark \\ \hline
+\end{tabular}}
+\label{tab:scenarios}
+\end{table}
 
@@
 Since an adversary can quite efficiently escalate from S3 to S4, we focus on scenarios S1, S2 and S4 in what follows.
-We first present an attack in scenario S1, providing much stronger results than the corresponding attack of~\cite{hllvuln} (in their equivalent scenario M2).
+
+%We first present an attack in scenario S1, providing much stronger results than the corresponding attack of~\cite{hllvuln} (in their equivalent scenario M2).
 We then use the additional information the adversary has access to in scenarios S2 and S4 to make the attack even more powerful.
 
-\subsection{Manipulating the HLL Cardinality Estimate}
+\section{Manipulating HLL Cardinality Estimates}\label{sec:manip}
 
-Although we consider scenario S1 to be quite artificial, we will commence our analysis there, since the techniques we present are more broadly applicable in S2 and S4.
+We divide our analysis into two cases: when the HLL sketch is initially empty (and the adversary knows this), and when it is not.
+For each case, we consider adversaries operating in scenarios S1, S2 and S4.
 
-\noindent\paragraph{Scenario S1} The goal of our adversary is to infer some information about targeted items' hash values, precisely their bucket mappings, knowing neither $h$ nor the salt in use (if any).
-Using this information, the adversary is able to retain items that are mapped to a subset of $b$ buckets out of the $2^n$, and minimize the resulting estimated cardinality.
-We assume the adversary either has flexibility in choosing the items to insert, or is presented with a static list of items at the start of its attack.
-We also assume that the shadow HLL sketch and the target HLL sketch are both initially empty.
+\subsection{Attacking an Initially Empty HLL Sketch}\label{sec:empty}
 
-Our adversary takes advantage of the use of HLL's small cardinality Hit Counting correction to recover the number of non-empty buckets (whose register does not contain value 0) after inserting a few items into the initially empty shadow HLL sketch.
-Indeed, in the small cardinality range, the cardinality estimate $E^*$ is computed as $E^*= m \cdot \ln (m/V)$ with $V$ the number of empty buckets.
-For typical HLL parameters, e.g.\ $n = 8, m =256$, the exact value of $V$, hence $b$, can be recovered from $E^*$.
-Moreover, when inserting an item in the small cardinality range, the adversary will see an increase in the cardinality estimate for the shadow HLL sketch if and only if $V$ is modified.
-This in turn means that the number of empty buckets in the shadow HLL sketch decreased by 1, and therefore that the freshly inserted item was mapped to a bucket that was not hit before.
+We begin with a preliminary lemma, applying to an adversary operating against an initially empty sketch (in any of S1, S2, S4).
+As usual, we have $m=2^n$ buckets, while $n+\ell$ is the output length of the employed hash function $h$.
 
-The adversary blindly picks the first $b$ buckets to be hit by simply inserting items into the shadow HLL until the cardinality estimate indicates that $b$ buckets have been touched.
+\begin{lemma}\label{lem:bound}
+Suppose an adversary inserts elements $x$ from some set ${\cal X}$ into an initially empty HLL sketch.
+Suppose the adversary can guarantee that in its attack, only a subset of size $B$ out of $2^m$ buckets are hit, where $B \leq B_{\max}$ with
+\[
+B_{\max} := \left\lfloor m\cdot\left(1-\frac{2\alpha_m}{5}\right)\cdot\left(1-2^{-(\ell+1)}\right)^{-1} \right\rfloor. \label{eqn:1} 
+\]
+Then, at the end of the attack, the HLL cardinality estimate is at most $m \cdot \ln (m/V)$ where $V = m-B$.
+\end{lemma}
+
+\begin{proof}
+Recall that the Hit Counting algorithm is used in the HLL algorithm as described in Figure~\ref{fig:hll} when the raw 
+estimate $E = \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$ does not exceed $\frac{5}{2}m$.
+Here $M[j]$ is the counter stored in bucket $j$.
+Note that $E$ increases with increasing $M[j]$ and we have $B$ buckets each contributing, in the worst case, $2^{-(\ell+1)}$ to the sum (and hence $m-B$ buckets which contribute $2^{0}=1$ to the sum).
+Hence, we see that for the Hit Counting algorithm to be used, we require:
+\[
+\alpha_mm^2\cdot\left((m-B)+B\cdot2^{-(\ell+1)}\right)^{-1} \le \frac{5}{2}m.
+\]
+Solving for $B$, we obtain:
+\[
+B \le B_{\max} := \left\lfloor m\cdot\left(1-\frac{2\alpha_m}{5}\right)\cdot\left(1-2^{-(\ell+1)}\right)^{-1} \right\rfloor.
+\]
+When the Hit Counting algorithm is used, the HLL cardinality estimate is set as $m \cdot \ln (m/V)$ where $V = m-B$ is the number of non-empty buckets. This completes the proof.
+\end{proof}
+
+For example with $n = 8$ (so $m=256$) and $\ell = 24$, we have $\alpha_m = 0.72$ and we find that $B_{\max}=184$.
+More generally, since for large $m$ we have $\alpha_m \approx 0.72$, and the term $(1-2^{-(\ell+1)})^{-1}$ is always close to but greater than 1, we see that $B_{\max} \approx  0.71\cdot m$.
+Note also that $m \cdot \ln (m/V)$ can be rewritten as $m \cdot \ln (1+\frac{B}{m-B})$.
+
+A surprising implication of the above analysis is that, for the HLL algorithm as described in Figure~\ref{fig:hll}, provided $B \leq B_{\max} \approx  0.71\cdot m$, no matter which fixed set of $B$ buckets are hit in an attack, and no matter how many leading zero bits the items' hashes have, the final cardinality estimate will not exceed $m \cdot \ln (1+\frac{B}{m-B})$.
+Setting, say, $B = m/2$, so half of the buckets can be hit in an attack, we would obtain a bound of $m \cdot \ln 2$ on the cardinality estimate.
+
+\subsubsection{Initially Empty HLL Sketch, Scenarios S2 and S4:}
+
+Armed with the above lemma, we now consider scenarios S2 and S4.
+We begin by noting that, when dealing with an initially empty sketch, S2 and S4 are actually equivalent, since the only additional information that an S4 adversary has is a snapshot of the HLL sketch, and this information is effectively given to the S2 adversary by virtue of the attack setting.
+So we consider S2 and S4 together.
+
+Now our proposed attack is very simple.
+Suppose the adversary has some set ${\cal X}$ of $2^t$ items from which to pick its inputs.
+The adversary selects a value of $B \leq B_{\max}$, and defines its target buckets to be those numbered $1,\ldots,B$.
+Since the adversary knows $h$, it can tell whether, for each $x \in {\cal X}$, insertion of $x$ into the HLL sketch would hit one of the target buckets or not; it inserts those that do.
+In expectation, assuming $h$ behaves like a random function, the adversary will successfully insert a fraction $B/m$ of the items, totalling $(B/m)\cdot 2^t = 2^{t-n+\log_2(B)}$ items.
+Meanwhile the final cardinality estimate will remain below $m \cdot \ln (1+\frac{B}{m-B})$.
+
+For example, setting $B = m/2$ (which, for practical parameters, satisfies $B \leq B_{\max}$), we can expect to insert $2^{t-1}$ items while keeping the cardinality estimate below $m \ln 2$.
+
+As a second example, setting $B = 1$, we can expect to insert $2^{t-n}$ items with a final cardinality estimate equal to $m \cdot \ln (1+\frac{1}{m-1})$, which is rounded to $1$ by the HLL algorithm (for $m$ of practical interest, e.g.\ $m=256$).
+Thus we keep the HLL cardinality estimate to a constant (1!) whilst being able to insert exponentially many items, provided there is sufficient flexibility in the choice of input set ${\cal X}$.
+
+\subsubsection{Initially Empty HLL Sketch, Scenario S1:}
+
+Here the attack is somewhat more complex.
+We begin by explaining the core idea, then a simplistic version of our attack, then a more complex but performant version, and finally, we give a heuristic performance analysis of the attack.
+
+Recall that in scenario S1, the adversary has access to a shadow HLL sketch which behaves identically to the target HLL sketch, and which can be reset at will.
+We will seek to minimise the number of resets performed in our attack, since we assume these are costly.
+The adversary can also read out the HLL estimate of the shadow device at will.
+
+\paragraph{Core Idea:}
+We assume the adversary again has (or is given) some set ${\cal X}$ of $2^t$ items from which to pick its inputs.
+The core idea of our attack is to use the first $B \leq B_{\max}$ items $x_1,\ldots, x_B$ from ${\cal X}$ to define a set of target buckets, and then to insert into the shadow HLL sketch further items $x \in {\cal X}$, checking for a cardinality estimate increase on each insertion.
+We initialise a list ${\cal L}$ containing $x_1,\ldots, x_B$.
+We add those items that do not increase the cardinality estimate on insertion into the shadow HLL sketch to ${\cal L}$ and reject the rest.
+We reset the shadow device when the cardinality estimate has increased too much, reinserting the first $B$ items from ${\cal X}$ again to re-establish a ``baseline'' of target buckets for the shadow HLL sketch.
+A key observation is that if in the attack we keep the number of ``active'' buckets to values $B^\prime$ that are always below $B_{\max}$, then the Hit Counting algorithm is used to compute the cardinality estimate; given its output $m \cdot \ln (m/V)$ (or, more exactly, a rounded version of this value) we can infer the exact value of $V$, and hence $B^\prime$.
+This enables precise detection of cardinality increases, provided we perform a reset before the Hit Counting algorithm stops being used (this is ensured by resetting when the estimate indicates that $B_{\max}$ buckets have been hit).
+After all of ${\cal X}$ has been processed as described, we take the constructed list ${\cal L}$ and reprocess it in the same way.
+We must ensure that the same set of initial items $x_1,\ldots, x_B$ is inserted into the shadow HLL sketch before any other items, so as to fix the same set of target buckets.
+The order of insertion can be otherwise randomised.
+% (and our analysis will take advantage of the fact that it can be).
+By repeated reprocessing, we will gradually remove items that do not hit the $B$ target buckets (defined by $x_1,\ldots, x_B$).
+Eventually, we obtain a final list ${\cal L}^*$ that we insert into the target HLL sketch.
+
+By construction and from Lemma~\ref{lem:bound}, we are guaranteed that the cardinality estimate for the target HLL sketch will be at most $m \cdot \ln (m/V)$ where $V = m-B$.
+Assuming $h$ behaves like a random function, it is clear that the expected size of the final list ${\cal L}^*$ is $\frac{B}{m}\cdot |{\cal X}|$.
+Thus the attack's final performance is the same as that in scenarios S2 and S4 above (but more processing and resets are required).
+
+\paragraph{Simplistic Attack:}
+The simplistic version of our attack uses the above idea, inserting items $x$ into the shadow HLL sketch but resetting it as soon as the cardinality estimate increases from $B$ to $B+1$ (indicating that a non-target bucket has been hit by item $x$).
+Acting in this way precisely determines whether each inserted item hits the target buckets or not, so has no ``false positives'' which need to be eliminated through reprocessing of the list ${\cal L}$.
+This means that the attack can be carried out in a streaming fashion in a single pass over ${\cal X}$.
+Assuming $h$ behaves like a random function, we must perform a reset with probability $\frac{m-B}{m} = 1 - \frac{B}{m}$ per item inserted, and hence in total an expected number of $(1 - \frac{B}{m})\cdot 2^t$ resets.
+
+\paragraph{Complex Attack:}
+The more complex version of the attack only resets the shadow HLL sketch when its cardinality estimate reaches a value indicating that $B_{\max}$ buckets have been hit (recall this number can be determined precisely because of the use of the Hit Counting algorithm for cardinality estimation).
+Hence, less resets should be required.
+However, when the number of active buckets has reached some value $B^\prime \leq B_{\max}$ this means that an item $x$ may not increase the cardinality estimate, yet still not hit one of the $B$ target buckets (instead it hits one of the other $B^\prime - B$ active buckets).
+Such an $x$ acts as a ``false positive" and needs to be eliminated through reprocessing of the list ${\cal L}$.
+
+\paragraph{Analysis of the Complex Attack:}
+%Suppose that after each reset, the order of the items in ${\cal X}$ (or ${\cal L}$) that have not yet been processed on the current pass is randomised.
+(Here we exclude the special items  $x_1,\ldots, x_B$ from the reordering.) Then
+On each pass, the probability that an item $x$ that should \emph{not} be inserted (i.e.\ $x$ is a true negative) does \emph{not} increase the cardinality estimate (i.e.\ $x$ is a false positive on this pass) is given by $(B^\prime - B)/m$, where $B^\prime$ is the current number of active buckets at the point when we try to insert $x$.
+(Here we ignore any inconvenient dependence issues across the passes.) This is bounded by $q := (B_{\max} - B)/m$ in the attack, since $B^\prime$ does not exceed $B_{\max}$.
+The probability that such an $x$ is correctly rejected is at least $1-q$.
+Hence the expected number of trial insertions done on $x$ (across different passes of the attack) before $x$ is correctly rejected is $1/(1-q) = m/(m - B_{\max} + B)$.
+
+Now we expect to reject in total $(1 - \frac{B}{m})\cdot 2^t$ items, and each one needs an expected number of trial insertions that is bounded by $m/(m - B_{\max} + B)$.
+(In fact, we can model the number of trial insertions needed as an independent geometric distribution with parameter $q$ for each such $x$.) Meanwhile, each of the $\frac{B}{m}\cdot 2^t$ items in the final insertion list ${\cal L}^*$ for the target HLL sketch is trial inserted into the shadow HLL sketch on every pass.
+
+All the passes in which some item $x$ is a false positive do not lead to cardinality estimate increases; only the final pass in which $x$ is identified as a true negative does so.
+Hence each rejected item only increases the number of active buckets once during the entire attack.
+Since we allow the number of active buckets to start at $B$ and reach $B_{\max}$ before resetting, the expected total number of resets (and passes) needed is given by:
+\[
+N_R:= \frac{1 - \frac{B}{m}}{B_{\max} - B}\cdot 2^t.
+\]
+Hence the expectation of the total number of trial insertions $N_I$ is bounded as:
+\[
+N_I \leq \frac{m}{ m-B_{\max} + B} \cdot \left(1 - \frac{B}{m} \right)\cdot 2^t +   N_R \cdot \frac{B}{m}\cdot 2^t.
+\]
+
+As a concrete illustration, we take $B = m/2$ and $B_{\max} = 0.71\cdot m$, to obtain $N_R = (2.38/m) \cdot 2^t$ and $N_I \leq 0.63 \cdot 2^t + 1.19\cdot 2^{2t}/m$.
+Here the dominant term in $N_I$ is ${\cal O}(2^{2t})$, i.e.\ the total insertion cost grows as the square of the size of the starting set ${\cal X}$.
+The result of the attack is that, as in S2 and S4, an expected $2^{t-1}$ items can be inserted into the target HLL sketch while keeping its cardinality estimate below $m \ln 2$.
+
+The insertion cost of the attack can be reduced by more smartly ordering the items $x$ to be trial inserted on each pass.
+Note that, in a given pass, when $B^\prime$ is still close to $B$, an item $x$ that does not increase the number of active buckets is more likely to have hit one of the target buckets than is the case when $B^\prime$ is large; such an $x$ can be tried earlier in the sequence of trials in the next pass and its exact status determined sooner.
+Any $x$ that can be determined to be a true positive (because its insertion into the shadow HLL sketch keeps the number of active buckets at exactly $B$) can be placed immediately in ${\cal L}^*$ and not reprocessed any further.
+We leave the analysis of this optimisation to future work.
+
+%The adversary blindly picks the first $b$ buckets to be hit by simply inserting items into the shadow HLL until the cardinality estimate indicates that $b$ buckets have been touched.
 It puts these items in a final target list ${\cal L}$.
@@
 
-As soon as the cardinality estimate exceeds $\frac{5}{2}m$, the adversary can assume that the shadow HLL sketch's output corresponds to the raw HLL cardinality estimate with high probability.
+%As soon as the cardinality estimate exceeds $\frac{5}{2}m$, the adversary can assume that the shadow HLL sketch's output corresponds to the raw HLL cardinality estimate with high probability.
 In this case, no more useful information can be inferred about the bucket mappings in the shadow HLL sketch, so the adversary resets the shadow HLL sketch to the empty state and starts again with the list ${\cal L}$ of non-discarded items obtained so far.
@@
 
-After sufficiently many iterations of this procedure, the adversary will have created a list of items ${\cal L}$ such that, when inserted into the \emph{target} HLL, create a sketch in which $b$ target buckets are filled and the others left empty.
+%After sufficiently many iterations of this procedure, the adversary will have created a list of items ${\cal L}$ such that, when inserted into the \emph{target} HLL, create a sketch in which $b$ target buckets are filled and the others left empty.
 With an appropriate choice of $b$ (as discussed immediately below), the final cardinality estimate for the target HLL sketch will be $E^*= m \cdot \ln (m/V)$ where $V = m-b$ is the number of empty buckets.
@@
 
-The value of $b$ has to be picked carefully in order to keep the shadow and target HLL cardinality estimates low, as the registers in both sketches may hold values that (potentially) become very high as more items are inserted.
+%The value of $b$ has to be picked carefully in order to keep the shadow and target HLL cardinality estimates low, as the registers in both sketches may hold values that (potentially) become very high as more items are inserted.
 The trick is select $b$ so that, even in the worst case scenario where all register values $M[j]$ hold the maximal value of $\ell+1$, the Hit Counting algorithm will still be used to produce the cardinality estimates.
 Noting that the raw estimate $E = \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$ increases with increasing $M[j]$, and that we have $b$ buckets each contributing, in the worst case, $2^{-(\ell+1)}$ to the sum (and hence $m-b$ buckets which contribute $2^{0}=1$ to the sum) we see that for the Hit Counting algorithm still to be used, we require:
-\[
-\alpha_mm^2\cdot\left((m-b)+b\cdot2^{-(\ell+1)}\right)^{-1} \le \frac{5}{2}m.
-\]
-Solving for $b$, we obtain:
-\[
-b \le b_{\max} := \left\lfloor m\cdot\left(1-\frac{2\alpha_m}{5}\right)\cdot\left(1-2^{-(\ell+1)}\right)^{-1} \right\rfloor. \label{eqn:1} 
-\]
-
-For example with $n = 8$ (so $m=256$) and $\ell = 24$, we find that $b_{\max}=184$.
-
-A surprising implication of this analysis is that, for the HLL algorithm as described in Figure~\ref{fig:hll}, provided $b \leq b_{\max}$, no matter which fixed set of $b$ buckets are hit, and no matter how many leading zero bits the items have, the final cardinality estimate will not exceed $m \cdot \ln (1+b/(m-b))$.
+%\[
+%\alpha_mm^2\cdot\left((m-b)+b\cdot2^{-(\ell+1)}\right)^{-1} \le \frac{5}{2}m.
+%\]
+%Solving for $b$, we obtain:
+%\[
+%b \le b_{\max} := \left\lfloor m\cdot\left(1-\frac{2\alpha_m}{5}\right)\cdot\left(1-2^{-(\ell+1)}\right)^{-1} \right\rfloor. \label{eqn:1} 
+%\]
+%
+%For example with $n = 8$ (so $m=256$) and $\ell = 24$, we find that $b_{\max}=184$.
+%
+%A surprising implication of this analysis is that, for the HLL algorithm as described in Figure~\ref{fig:hll}, provided $b \leq b_{\max}$, no matter which fixed set of $b$ buckets are hit, and no matter how many leading zero bits the items have, the final cardinality estimate will not exceed $m \cdot \ln (1+b/(m-b))$.
 We will exploit this in the sequel in scenarios S2 and S4.
@@
 Moreover, setting $b=128$ would mean that half of all items available to the adversary could be inserted into the target HLL.
+%
+%%The initial set of items to be filtered can be built by varying the flexible fields of the items $x$ the adversary is trying to insert.
+%%so that the adversary can ``hide'' $2^{t-n}$ elements while keeping the HLL cardinality estimate to 1.
+%
 
-%The initial set of items to be filtered can be built by varying the flexible fields of the items $x$ the adversary is trying to insert.
-%so that the adversary can ``hide'' $2^{t-n}$ elements while keeping the HLL cardinality estimate to 1.
-
-\noindent\paragraph{Scenario S2} The adversary knows $h$ so it can adjust fields in the item $x$ it is trying to insert and keep only those for which the $\ell$ rightmost bits of $h(x)$ start with a one bit.
+%\subsubsection{Scenario S2} The adversary knows $h$ so it can adjust fields in the item $x$ it is trying to insert and keep only those for which the $\ell$ rightmost bits of $h(x)$ start with a one bit.
 In this case, the number of leading zero bits computed by the estimate of the destination bucket is always zero, so the estimate of this bucket is unlikely to be updated and the final averaged approximate cardinality is likely to be left unchanged.
-For example, in the scenario of~\cite{portscanhll}, the adversary could change the source port of its IP packet until it is satisfied with its hash value.
+%For example, in the scenario of~\cite{portscanhll}, the adversary could change the source port of its IP packet until it is satisfied with its hash value.
 Given sufficient flexibility in the input $x$, the adversary can insert many items into the HLL sketch without increasing the cardinality markedly, by repeatedly sampling from the input domain.
 
-When attacking an empty sketch, the adversary can keep this strategy and already mount a powerful attack, but can do better by following the approach presented under S1.
+%When attacking an empty sketch, the adversary can keep this strategy and already mount a powerful attack, but can do better by following the approach presented under S1.
 Filtering items is easier in this scenario as the adversary can infer the bucket mapping of an item $x$ directly from the first $n$ bits of $h(x)$.
@@
 
-\noindent\paragraph{Scenario S4} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values per bucket in the HLL sketch).
+%\subsubsection{Scenario S4} The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values per bucket in the HLL sketch).
 Therefore, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
@@
 
-A special case arises when the HLL is completely empty at the beginning of the attack.
+%A special case arises when the HLL is completely empty at the beginning of the attack.
 Here the adversary is forced to increase the estimate in at least one bucket.
 Once again, the adversary can choose to fill $b$ buckets following the strategy described in S2, leaving all the other buckets untouched.
-
-\subsection{Impact}\label{sec:impact}
-In this section we evaluate the impact of the above attacks on the HLL cardinality estimate.
+%
+%\subsection{Impact}\label{sec:impact}
+%In this section we evaluate the impact of the above attacks on the HLL cardinality estimate.
 We assume that $h$ is a ``good" hash function, so that each of the bits of $h(x)$ can be regarded as being independent and uniformly random.
@@
 
-\subsubsection{Non-Empty Sketch}\label{sec:nonempty}
+\subsection{Attacking an Initially Non-Empty HLL Sketch}\label{sec:nonempty}
 
-We start by analysing our results when the attack is ran against a sketch that already received inputs from other users, first under scenario S2 then S4.
+Now we turn to the case where the HLL sketch under attack is not initially empty, but instead contains some honestly inserted items.
+As usual, the adversary's goal is to insert as many items as possible whilst increasing the cardinality estimate as little as possible.
 
-In scenario S2, the adversary needs to craft two distinct candidates for $x$ on average to obtain a one bit at the first position in the $\ell$ rightmost bits of $h(x)$, thus successfully performing the attack for a single insertion.
-The adversary can expect to ``hide''  $2^{t-1}$ distinct items without increasing the HLL cardinality estimate at all, under the assumption that all buckets in the HLL sketch already hold a value $c_i \ge 1$.
-On the other hand, if some buckets are empty (but the adversary does not know which ones in the blind setting of S2), then we may expect a moderate increase in the HLL cardinality estimate.
-Notice that in all cases, provided $t$ is large enough, the adversary can keep the cardinality estimate linear in the total number of buckets ($2^n$) while the expected value of the true cardinality, $2^{t-1}$, can be made as large as the adversary pleases.
+We focus here on scenarios S2 and S4, and only sketch an attack below for S1.
 
-In scenario S4, we compute the average work required of the adversary to find an input $x$ such that $h(x)$ meets the attack's requirements.
-Let us assume for the moment that $c_i \ge 1$ for all buckets $i$.
-Assume $h(x)$ is mapped to bucket $i$.
-The probability that $h(x)$ does not increase the estimate is the probability that the  rightmost $\ell$ bits of $h(x)$ have $c_i$ or less leading zeroes; this probability is equal to $1-\frac{1}{2^{c_i+1}}$ assuming $h$ behaves like a random function.
-Averaging over all $2^n$ buckets (and using the fact that the bucket choice is uniformly random since the bits of $h$ are independent and uniformly random), the probability that input $x$ does not increase the estimate is given by:
+We assume that the target HLL sketch in S2 and S4 has already had sufficiently many items inserted that every one of the $m=2^n$ buckets has counter value $M[j]$ with $M[j] \ge 1$.
+This is highly likely to be the case (with probability at least $1-1/m^2$) as long as at least $m\ln m$ items have already been inserted into the target HLL sketch.
+This follows from the standard analysis of the coupon collector problem under our usual assumption that $h$ behaves as a random function.
+In S4, the adversary can use its snapshot to detect if this is the case or not, and behave differently if not (details for this case follow below).
+
+With $m \ln m$ honest insertions, it is also highly likely that the small range correction (Hit Counting algorithm) will not be used, and the target HLL's cardinality estimate will be computed using the raw estimate:
+\[
+E = \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}.
+\]
+
+We assume the adversary has (or is given) some set of items ${\cal X}$ that it wishes to insert; as usual, we write $|{\cal X}| = 2^t$.
+
+\subsubsection{Initially Non-empty HLL Sketch, Scenario S2:} 
+The adversary knows $h$ so it selects items $x$ from ${\cal X}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit.
+In this case, since the relevant bucket $j$ already has $M[j] \ge 1$, the bucket's counter is not updated, and the cardinality estimate of the HLL sketch remains unchanged.
+
+The adversary then expects to be able to insert half of the items from ${\cal X}$ into the HLL sketch, without increasing the HLL cardinality estimate at all.
+
+If some buckets are actually empty (but the adversary does not know which ones in the blind setting of S2), then we may expect a moderate increase in the HLL cardinality estimate in the above attack.
+If the adversary suspects this is the case, then it can first insert $m$ carefully selected items into the target HLL sketch to ensure that $M[j] \ge 1$ for every bucket.
+This should increase its cardinality estimate by about $m$ (since HLL is a good cardinality estimator!).
+Then it can mount the attack, and be sure that it does not increase the cardinality estimate any further.
+
+\subsubsection{Initially Non-empty HLL Sketch, Scenario S4:} 
+The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch).
+Therefore, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
+In other words, if $h(x)$ is mapped to the $j$-th bucket containing $M[j] \ge 1$, then, instead of the strict requirement of having a ``1'' bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a ``1'' bit in any of the first $M[j]$ positions in this substring.
+Under these conditions, the estimates in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.
+
+We now compute the expected number of elements $x$ from ${\cal X}$ such that $h(x)$ meets the attack's requirements.
+We assume that $M[j] \ge 1$ for all buckets $j$.
+Assume $h(x)$ is mapped to bucket $j$.
+Inserting $x$ does not increase the cardinality estimate if the rightmost $\ell$ bits of $h(x)$ have $M[j]-1$ or fewer leading zeroes; this probability is equal to $1-\frac{1}{2^{M[j]}}$ assuming $h$ behaves like a random function.
+Averaging over all $m$ buckets (and using the fact that the bucket choice is uniformly random since the bits of $h$ are independent and uniformly random), the probability that input $x$ meets the attack's requirements is given by:
 \begin{eqnarray*}
-\sum_{i=1}^{2^n}\frac{1}{2^n} \cdot (1-\frac{1}{2^{c_i+1}}) & = & 1-\frac{1}{2}\cdot (\frac{1}{2^n}\cdot \sum_{i=1}^{2^n} {(2^{c_i})}^{-1}) \\
-& = &1-(2H_c)^{-1} 
+\sum_{j=1}^{m}\frac{1}{m} \cdot (1-\frac{1}{2^{M[j]}}) & = & 1- \frac{1}{m}\cdot \sum_{j=1}^{m} {(2^{M[j]})}^{-1} \\
+& = &1 - H_M^{-1} 
 \end{eqnarray*}
-where $H_c$ is the harmonic mean of the counts $2^{c_1}, 2^{c_2}, \ldots, 2^{c_{2^n}}$.
-So the adversary needs to craft on average $(1-(2H_c)^{-1})^{-1}$ distinct candidates for $x$ in order to insert a single candidate, and, given $t$ bits of flexibility in its choice of inputs, can expect to ``hide" up to $(1-(2H_c)^{-1}) \cdot 2^t$ distinct items without increasing the HLL cardinality estimate at all.
+where $H_M$ is the harmonic mean of the counts $2^{M[1]}, 2^{M[2]}, \ldots, 2^{M[m]}$.
+So, given ${\cal X}$ containing $2^t$ items, the adversary can expect to insert $(1 - H_M)^{-1} \cdot 2^t$ items from ${\cal X}$ without increasing the HLL cardinality estimate at all.
 
-Note that the above analysis for scenario S4 made the assumption that every bucket count $c_i$ was already at least 1.
-When some bucket counts are 0, the above analysis needs to be modified in order to avoid touching those buckets altogether.
-The modified analysis is quite straightforward.
-Suppose $V \le 2^n-1$ out of $2^n$ buckets have a count of zero.
-Then the cost of crafting an input $x$ becomes $(1-(2H_c)^{-1})^{-1} \cdot (1- \frac{V}{2^n})^{-1}$ trials, while the total number of items the adversary can expect to insert without increasing the HLL cardinality estimate becomes $(1-(2H_c)^{-1}) \cdot (1- \frac{V}{2^n})\cdot 2^t$.
-Of course, the adversary can relax its attack and allow some of the empty buckets to be hit.
-This would increase the insertion rate at the cost of a small increase in the final cardinality estimate.
+If some buckets are actually empty (and in S4 the adversary knows which ones using its snapshot), then the above analysis can be reused in order to avoid touching those buckets altogether when inserting elements from ${\cal X}$.
+The above analysis does not change at all because the term $\frac{1}{m} \cdot (1-\frac{1}{2^{M[j]}})$ already signs a probability of 0 to inserting $x$ in bucket $j$ if $M[j] = 0$.
+Hence we still expect to insert $(1 - H_M)^{-1} \cdot 2^t$ items from ${\cal X}$ without increasing the HLL cardinality estimate at all.
+Of course, the adversary can relax its attack and allow some of the empty buckets to be hit as well.
+This would increase the insertion rate at the cost of an increase in the final cardinality estimate.
 
-\textbf{Conclusion:} In this setting, we see that an adversary who has $t$ bits of flexibility in its inputs can expect to insert exponentially (in $t$) many input values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the cases where the HLL sketch has empty buckets in S2).
-The result is largely independent of the HLL parameters, but does depend on the extent to which the HLL sketch is already filled by other users, via a factor $1-(2H_c)^{-1}$ and the number of empty buckets $V$, via a factor $1- \frac{V}{2^n}$.
-Other trade-offs between work and the cardinality estimate are possible.
+In this setting, where the target HLL sketch is initially non-empty, we see that an adversary who has sufficient flexibility in its inputs (as represented by the size $2^t$ of the set ${\cal X}$) can expect to insert exponentially many (in $t$) values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the case where the HLL sketch has some empty buckets in S2).
+The result is largely independent of the HLL parameters, but does depend on the extent to which the HLL sketch is already filled by other users, via a factor $1-(H_M)^{-1}$: the more full the HLL sketch is, the easier it is to insert new items without them increasing the cardinality estimate.
 
-\subsubsection{Empty Sketch}
+\subsubsection{Initially Non-empty HLL Sketch, Scenario S1:}
 
-We now consider the case where the HLL sketch is initially empty.
-In that case, the output of the attacks is the same for all scenarios.
-Targeting $b$ buckets allows the adversary to insert a fraction $\frac{b}{2^n}$ of the items, and thus to retain $\frac{b}{2^n}2^t=2^{t-n+\text{log}_2(b)}$ elements.
-After inserting the retained items, the cardinality of the sketch is estimated to $2^n$ln$(2^n/(2^n-b))$, approximating to $b$ for relatively small values of $b$.
+Finally, we briefly consider scenario S1 with an initially non-empty target HLL sketch.
 
-When following the initial strategy of S2, we saw in paragraph~\ref{sec:nonempty} that the adversary retains $2^{t-1}$ items.
-The insertion of those items will hit every bucket of the sketch with high probability, and update every register once to 1.
-The cardinality of the sketch is then estimated to $\alpha2^{2n}(\sum_{j=1}^{2^n}2^{-1})^{-1}=\alpha2^{n+1}$ and, as $V=0$, no small range cardinality correction is applied.
-On the other hand, when targeting and filling half the buckets ($b=2^{n-1}$), the adversary can insert as many items but lower the estimated cardinality to $2^n$ln$(2^n/(2^n-2^{n-1}))=2^n\text{ln}(2)$ instead of $\alpha{2^{n+1}}$ that we previously had, at no extra computational cost.
+We first make an empty shadow HLL sketch via a reset.
+Consider an adversary that simply inserts $m \ln m$ random items $x$ from ${\cal X}$ into the shadow and target HLL sketches.
+This increases the target HLL sketch's cardinality estimate by about $m \ln m$ since the HLL algorithm is a good estimator.
+It also sets the shadow HLL sketch's estimate to about $m \ln m$, by the same argument.
+At the same time, with probability $1 - 1/m^2$, it ensures that every bucket in the shadow HLL sketch becomes active (i.e.\ has counter values $M[j]$ with $M[j] \ge 1$ for all $j$).
+Now the adversary can carry out a version of our previous attack in scenario S1 (where the target HLL sketch is initially empty).
+By repeatedly processing lists of inputs, it can find a large subset ${\cal L}^*$ of ${\cal X}$ which does not cause the cardinality estimate of the shadow HLL sketch to increase beyond $m \ln m$.
+This assumes that single updates to the vector of counter values $(M[j])_{j=1}^m$ in the shadow HLL sketch can be detected through changes to the HLL cardinality estimate.
+This seems plausible but requires further experimental investigation.
+Inserting all the items from ${\cal L}^*$ into the target HLL sketch then leads to an increase in its cardinality estimate of at most $m \ln m$.
+It is not hard to see that, since every bucket in the shadow HLL sketch is active, we can expect ${\cal L}^*$ to contain at least half of the items of ${\cal X}$.
 
-\textbf{Conclusion:} When confronted to an empty sketch, the adversary is forced to concede a slight cardinally estimation increase in order to be able to insert some elements.
-For a fixed $b$, an adversary who has $t$ bits of flexibility in its inputs is then able to insert exponentially (in $t$) many input values into the HLL sketch while keeping the cardinality estimate constant.
-Any value below the upper-bound specified in eq. (4.1) can be chosen for $b$ based on the setup requirements. For example if the reset operation is costly, the adversary can adjust $b$ accordingly.
+%An alternative approach is to simply run the scenario S1 attack as if the target HLL sketch were empty (i.e.\ as described in Section~\ref{sec:empty} above), take the resulting list ${\cal L}^*$, and insert it into the target HLL sketch, hoping that the cardinality does not increase too much.
+%We report on this approach experimentally in Section~\ref{sec:exp}.
+It turns out to work well in practice.
+%% Nah, that doesn't work -- the S1 attack crucially depends on staying in the Hit Count range, and you go out of that as soon as you try to ``merge'' with a non-empty target.
+
+We omit further analysis of this scenario.
 
@@
 Many different implementations of HLL are available online.
-Among the repositories collecting more than 350 stars on GitHub, we can cite~\cite{clahll} and~\cite{datasketch}.
+Among the repositories collecting more than 350 stars on GitHub are ~\cite{clahll,datasketch}.
 HLL is also featured in many frameworks, including Redis~\cite{redis}, Google's BigQuery~\cite{bigquery}, Facebook's Airlift~\cite{airlift} and several products of Apache such as Spark~\cite{spahll} and Druid~\cite{druhll}.
@@
 
-As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}.
-The attack code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\textit{\url{https://github.com/PizzaWhisperer/HLLVuln}}}.
-For simplicity, our items are 4-character long ASCII strings, giving us up to $7,311,616= 2^{22.8}$ distinct items that can be inserted (hence we have $t=22.8$ in our attacks).
+As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}, since it is faithful to the original description of HLL given in~\cite{hll}.\footnote{Our attack code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\url{https://github.com/PizzaWhisperer/HLLVuln}}.}.
+For simplicity, our items are 4-character long ASCII strings, giving us a set ${\cal X}$ of $7,311,616= 2^{22.8}$ distinct items that can be inserted (hence we have $t=22.8$ in our attacks).
 We choose $h$ to be the 32-bit Murmur3 hash function from~\cite{murmur3code} (note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
-We create an HLL sketch using $2^n = 256$ buckets and initialise it with 1,000 random items representing honest users' data.
+We create an HLL sketch using $n=8$ and hence $m=256$ buckets and, for attacks in the setting of an initially non-empty target HLL sketch, initialise it with 1,000 random items representing honest users' data.
 
-We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $7,311,616$ possible inputs while trying to keep the cardinality estimate as low as possible, either in the S1, S2 or S4 scenario.
+We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $2^{22.8}$ possible inputs while trying to keep the cardinality estimate as low as possible, in each of the scenarios S1, S2 or S4.
+
 As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose what to insert from a set of 250,000 random items that are given to it.
 Finally, for completeness, we also run the attacks against HLL sketches filled with adversarial inputs only, thus that are initially empty.
-We vary the value of $b$ in the attack under the S1 scenario to demonstrate possible trade-offs.
-For simplicity and to keep the attack as striking as possible, we set $b$ to 1 in the attacks on empty sketch under S2 and S4.
-All attacks were run 30 times and the outcomes averaged.
+We vary the value of $B$ in the attack for scenario S1 to demonstrate some of the available trade-offs.
+For simplicity and to keep the attack as striking as possible, we set $B$ to 1 in the attacks on an initially empty HLL sketch in scenarios S2 and S4.
+All attacks were run 30 times; we report averge outcomes.
 Our results are reported in Tables~\ref{table:tab1} and~\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary, and the cardinality estimate after adding the adversary's items.
 
-These empirical results confirm our impact analysis on several points.
+These empirical results confirm our analysis on several points.
 
-First, when attacking a sketch already containing some items, we can see that the adversary is on average able to hide about half of the items in our starting set under S2 (as predicted by our theoretical analysis), and about 83\% under S4.
+First, when attacking a sketch already containing 1000 random items, we can see that the adversary is on average able to insert about half of the items in our starting set in scenario S2, and about 83\% in scenario S4.
+These results align well with our theoretical analysis.
+
+%In scenario S1 we run the attack designed for an initially empty target HLL sketch. It performs well in practice, for example, increasing the HLL cardinality estimate from 100 to 126????
 %( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$)
-Here $0.83 \approx (1-\frac{1}{2\cdot H_c})$ with $H_c\approx\frac{1000}{256}=3.9$, so the results are also consistent with our theoretical analysis.
+%Here $0.83 \approx (1-\frac{1}{2\cdot H_c})$ with $H_c\approx\frac{1000}{256}=3.9$, so the results are also consistent with our theoretical analysis.
+%% That analysis was now updated -- does it still match?
 
-Secondly, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly $2^n$ln$(2^n/(2^n-b))$ after inserting a fraction $\frac{b}{2^n}$ of the items (this comes from the Hit Counting algorithm result that arises from our attack targeting $b$ buckets).
+Secondly, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly $m \cdot \ln (m/(m-B))$ after inserting a fraction $\frac{B}{m}$ of the items, as predicted by our analysis.
 
@@
 Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780''~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality after 4 iterations (resets).
-The S1 adversary following our strategy is able to add 977 and 97,433 distinct items while keeping the cardinality estimate respectively at 1 and 126.
-Compared to~\cite{hllvuln}, we can reduce the cardinality estimate by a factor of almost 1,000 instead of 5 at the cost of more resets.
+The S1 adversary following our strategy is able to add 977 and 97,433 distinct items while holding the cardinality estimate at 1 and 126, respectively.
+Compared to~\cite{hllvuln}, we can reduce the cardinality estimate by a factor of almost 1,000 instead of 5.
+This comes at the cost of performing more resets compared to~\cite{hllvuln}.
+
+\todo{Check why in the first table in S1 we end up with a card.\ est.\ of 126 instead of $B=100$. The S1 attack description says we should stay at $B=100$ so something is off!}
 
@@
 \caption{Attack results, averaged over 30 iterations.}
+\medskip
      \makebox[\linewidth]{
-\begin{tabular}{| m{9.8em} | m{4em} | m{4.2em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
+\begin{tabular}{| m{11em} | m{4em} | m{4.2em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
     \hline
     \textbf{Scenario} &  \multicolumn{2}{c|}{S1} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
-    \textbf{\# Targeted buckets} & 1 & 100 & - & 1 & - & 1 \\ \hline
-    \textbf{Initial data?} & no & no & yes & no & yes & no \\ \hline
+    \textbf{\# Targeted buckets $B$} & 1 & 100 & n/a & 1 & n/a & 1 \\ \hline
+    \textbf{Initially empty?} & yes & yes & no & yes & no & yes \\ \hline
     \textbf{Original Card. Est.} & 0 & 0 & 1,000 & 0 & 1,005 & 0 \\ \hline
     \textbf{\# Items added} & 28,566 & 2,856,100 & 3,655,744 & 28,566 & 6,090,003 & 28,566 \\ \hline
-    \textbf{Final Card. Est.} & 1 & 126 & 1'011 & 1 & 1,005 & 1 \\ \hline
+    \textbf{Final Card. Est.} & 1 & 126 & 1,011 & 1 & 1,005 & 1 \\ \hline
 \end{tabular}}
@@
 \centering
-\caption{Attack results, averaged over 30 iterations, in the setting of \cite{hllvuln}.}
+\caption{Attack results, averaged over 30 iterations, in the setting of~\cite{hllvuln}.}
+\medskip
      \makebox[\linewidth]{
-\begin{tabular}{| m{9.8em} | m{4em} | m{4.2em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
+\begin{tabular}{| m{11em} | m{4em} | m{4.2em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
     \hline
     \textbf{Scenario} & \multicolumn{2}{c|}{S1} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
-    \textbf{\# Targeted buckets} & 1 & 100 & - & 1 & - & 1 \\ \hline
-    \textbf{Initial data?} & no & no & yes & no & yes & no \\ \hline
+    \textbf{\# Targeted buckets $B$ } & 1 & 100 & n/a & 1 & n/a & 1 \\ \hline
+    \textbf{Initially empty?} & yes & yes & no & yes & no & yes \\ \hline
     \textbf{Original Card. Est.} & 0 & 0 & 1,004 & 0 & 998 & 0 \\ \hline
@@
 
-A formal analysis of using salting as a countermeasure to our attacks remains open.
 %Future work could consist in finding mitigations that do not rely on salt.
 For example, would the imposed ID setting (like Facebook) solve the issue but still be convenient for the various HLL applications?
-Also left for future work is the task of extending our attacks to HLL variants such as sliding HLL~\cite{slidinghll}, HLL++~\cite{hllpratice} and Redis alternative using the maximum likelihood approach from~\cite{newhll}.
+Left to future work is the task of extending our attacks to HLL variants such as sliding HLL~\cite{slidinghll}, HLL++~\cite{hllpratice} and the version of HLL used in Redis, which relies on the maximum likelihood approach from~\cite{newhll} to perform corrections on cardinality estimates instead of the Hit Counting algorithm.
+
+It would be interesting to find and analyse a strategy for more cleverly ordering the candidates $x$ in our attack in scenario S1 to reduce the total insertion cost of the attack, whilst also keeping the number of resets low.
+
+A formal analysis of using salting or secretly keyed hash functions as a countermeasure to our attacks remains open.
+This could be done by extending the framework of~\cite{CCS:ClaPatShrPS19}.
+An interesting cryptographic challenge here is to  improve the security of HLL sketches against adversarial inputs whilst retaining some form of the uncoordinated mergeability that is enjoyed by the original HLL algorithm.
 

### 2020-09-25 docs/HLL_exp_bad/main.tex
tweakery

@@
 This prompts an examination of the performance of the HLL algorithm in the face of adversarial inputs.
-In this cautionary note, we show that in such a setting, the HLL algorithm's estimate of cardinality can be exponentially bad: when an adversary has access to the internals of the HLL algorithm and has some flexibility in choosing what inputs will be recorded, it can manipulate the cardinality estimate to be exponentially smaller than the true cardinality.
+We show that in such a setting, the HLL algorithm's estimate of cardinality can be exponentially bad: when an adversary has access to the internals of the HLL algorithm and has some flexibility in choosing what inputs will be recorded, it can manipulate the cardinality estimate to be exponentially smaller than the true cardinality.
 \end{abstract}
@@
     \hspace{2em} \textbf{set} $x := h(v);$\\
-    \hspace{2em} \textbf{set} $j = 1 + <x_1x_2...x_{n}>_2$; \hfill \textit{\{the binary address determined by the first $n$ bits of $x$\}}\\
+    \hspace{2em} \textbf{set} $j = 1 + <x_1x_2...x_{n}>_2$; \hfill \textit{\{the bucket index determined by the first $n$ bits of $x$\}}\\
     \hspace{2em} \textbf{set} $w := x_{n+1}x_{n+2}...$;\\
@@
 Here the attack is somewhat more complex.
-We begin by explaining the core idea, then a simplistic version of our attack, then a more complex but performant version, and finally, we give a heuristic performance analysis of the attack.
+We begin by explaining the core idea, then a simple version of our attack, then a more complex but performant version, and finally, we give a heuristic performance analysis of the attack.
 
@@
 
-\paragraph{Simplistic Attack:}
+\paragraph{Simple Attack:}
 The simplistic version of our attack uses the above idea, inserting items $x$ into the shadow HLL sketch but resetting it as soon as the cardinality estimate increases from $B$ to $B+1$ (indicating that a non-target bucket has been hit by item $x$).

### 2020-09-29 docs/HLL_exp_bad/main.tex
change M to X
change v to x

@@
 \vspace{0.5em}
-\textbf{Algorithm} HyperLogLog (\textbf{input} $\mathcal{M}$ : multiset of items from domain $\mathcal{D}$).\\
+\textbf{Algorithm} HyperLogLog (\textbf{input} ${\cal X}$ : multiset of items from domain $\mathcal{D}$).\\
 \textbf{initialize} a collection of $m$ registers, $M[1], ..., M[m]$, to $0$;\\
 \vspace{0.5em}
-\textbf{for} $v \in \mathcal{M}$ \textbf{do}\\
-    \hspace{2em} \textbf{set} $x := h(v);$\\
-    \hspace{2em} \textbf{set} $j = 1 + <x_1x_2...x_{n}>_2$; \hfill \textit{\{the bucket index determined by the first $n$ bits of $x$\}}\\
-    \hspace{2em} \textbf{set} $w := x_{n+1}x_{n+2}...$;\\
+\textbf{for} $x \in \cal X$ \textbf{do}\\
+    \hspace{2em} \textbf{set} $y := h(x);$\\
+    \hspace{2em} \textbf{set} $j = 1 + <y_1y_2...y_{n}>_2$; \hfill \textit{\{the bucket index determined by the first $n$ bits of $y$\}}\\
+    \hspace{2em} \textbf{set} $w := y_{n+1}y_{n+2}...$;\\
    \hspace{2em} \textbf{set} $M[j] :=$ max$(M[j], \rho(w))$;\\

### 2020-10-04 docs/HLL_exp_bad/main.tex
prep for adding redis and formal security

@@
 
+\todo{fix up the intro to refer to new sections, bring in more of the details of related work and HLL applications here.}
+
 \section{Overview of HLL}\label{sec:overview}
@@
 The output length of $h(\cdot)$ is thus $n+\ell$, fixed to 32 in~\cite{hll}.
-Estimates from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha$, where $\alpha$ is empirically computed to correct a systematic multiplicative bias.
+Estimates from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha_m$, where $\alpha_m$ is empirically computed to correct a systematic multiplicative bias.
 
@@
 Some works building on~\cite{hll} replace the Hit Counting algorithm by other schemes and use a hash function with a larger output domain to reduce the chances of collision and thus get rid of the need for large cardinality range correction.
-For example, \cite{hllpratice} uses a 64-bit hash function and Linear Counting along with bias correction in the small cardinality range.
-In this report we target the HLL formulation in the original paper~\cite{hll} as represented in Fig.~\ref{fig:hll}, and leave for future work the analysis of its variants.
+For example, \cite{hllpractice} uses a 64-bit hash function and Linear Counting along with bias correction in the small cardinality range, while~\cite{hllnew} proposes new estimators based on a more precise statistical analysis of the counting problem.
+In this report we mostly target the HLL formulation in the original paper~\cite{hll} as represented in Fig.~\ref{fig:hll}.
+We also study~\cite{hllnew} because of its recent deployment in Redis.
+We leave for future work the analysis of further HLL variants.
 
@@
 \textit{Let $h : \mathcal{D} \rightarrow \{0, 1\}^{32}$ hash data from some domain $\mathcal{D}$ to binary $32$-bit strings.\\
-Let $\rho(s)$ be the position of the leftmost $1$-bit of $s$: e.g., $\rho(1...) = 1, \rho(0001...) = 4, \rho(0^K) = K + 1.$\\}
+For $s$ of bit-length $\ell$, let $\rho(s)$ denote the position of the leftmost $1$-bit of $s$: e.g., $\rho(1...) = 1, \rho(0001...) = 4, \rho(0^\ell) = \ell + 1.$\\}
 \textbf{define} $m = 2^n$ with $n \in [4..16].$\\
@@
     \hspace{2em} \textbf{set} $y := h(x);$\\
-    \hspace{2em} \textbf{set} $j = 1 + <y_1y_2...y_{n}>_2$; \hfill \textit{\{the bucket index determined by the first $n$ bits of $y$\}}\\
-    \hspace{2em} \textbf{set} $w := y_{n+1}y_{n+2}...$;\\
+    \hspace{2em} \textbf{set} $j = 1 + <y_1y_2...y_{n}>_2$; \hfill \textit{\{bucket index determined by the first $n$ bits of $y$\}}\\
+    \hspace{2em} \textbf{set} $w := y_{n+1}y_{n+2} \ldots y_{n+\ell+1}$;\hfill \textit{\{bit string determined by the last $\ell$ bits of $y$\}}\\
    \hspace{2em} \textbf{set} $M[j] :=$ max$(M[j], \rho(w))$;\\
@@
 Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted sketch, either as a replacement for (but at the cost of losing mergeability) or in addition to (inducing a memory overhead) an unsalted sketch.
+We show in Section~\ref{sec:formal} that when a salt is used appropriately (i.e.\ as the secret key for a pseudo-random function that is used in place of hash function $h$), security against adversarial inputs can be achieved even in our strongest adversarial setting.
 %This might not be a convenient solution for some applications, they thus encourage further research.
@@
 They use their formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures, but not HLL.
-It is an interesting open problem to use their framework to develop a formal understanding of how HLL can be made secure in the adversarial setting.
+%It is an interesting open problem to use their framework to develop a formal understanding of how HLL can be made secure in the adversarial setting.
 
@@
 The adversary's goal is to reduce the estimate as much as possible, since this would cause more damage when  considering the above-mentioned applications of HLL.
-It is easy to modify our attack strategies for an adversary who instead wishes to artificially inflate the HLL estimate given a limited input capability.
+It is easy to modify our attack strategies for an adversary who instead wishes to artificially inflate the HLL cardinality estimate given a limited input capability.
 We also analyze the impact on the HLL cardinality estimate and the computational effort required.
@@
 For example it can manipulate IPID and other header fields in the context of IP packets.
-This enables the adversary to control the output of hash function $h$ and flexibly choose which items will be inserted into the HLL sketch.
+We model this by assuming the adversary has access to a large set ${\cal X}$ of potential inputs from which it can select items to insert into the HLL.
+Our adversary's objective, then, is to insert as many items from ${\cal X}$ as possible whilst keeping the HLL cardinality estimate as low as possible.
+
+% This enables the adversary to control the output of hash function $h$ and flexibly choose which items will be inserted into the HLL sketch.
 
@@
 Indeed, an adversary can easily recover the contents of each bucket in the sketch simply by analysing which inputs increase the cardinality estimate.
-To find the estimate held in a targeted bucket, the adversary  inserts a sequence of items $x$ with an incrementing number of leading zeros in the $\ell$ rightmost bits of $h(x)$ whilst holding the $n$ bits determining the bucket index constant, until the cardinality estimate of the sketch increases.
+To find the estimate held in a targeted bucket, the adversary  inserts a sequence of items $x$ with an incrementing number of leading zeros in the $\ell$ rightmost bits of $h(x)$ whilst holding the $n$ bits determining the bucket index constant, until the cardinality estimate of the sketch increases.\todo{Really?
+Can the adversary detect a change to an individual bucket through a cardinality estimate change?
+It's not really clear -- depends on an analysis of the raw HLL estimate using the harmonic mean which we haven't done!
+We may need to heavily revise all this text.}
  Once an increase occurs, the adversary knows that the maximum number of leading zeros observed in that bucket was just updated to a value that can be recovered by looking at the last inserted item.
@@
 
-We focus here on scenarios S2 and S4, and only sketch an attack below for S1.
+We focus here on scenarios S2 and S4, and make brief remarks on S1.
 
@@
 
-We first make an empty shadow HLL sketch via a reset.
+\todo{Fix this -- it's not clear the attack works because it's not clear we really can detect cardinality changes in the way desired (and we've not demonstrated that we can).}
+
+%We first make an empty shadow HLL sketch via a reset.
 Consider an adversary that simply inserts $m \ln m$ random items $x$ from ${\cal X}$ into the shadow and target HLL sketches.
@@
 
-We omit further analysis of this scenario.
+%We omit further analysis of this scenario.
+
+\section{Analysis of the Redis Implementation}\label{sec:redis}
+
+We present a brief analysis of the HLL implementation used in Redis, since Redis is a very popular key-value store implementation and the version of HLL used there differs significantly from that described in Figure~\ref{fig:hll}.
+
+%and the version of HLL used in Redis, which relies on the maximum likelihood approach from~\cite{newhll} to perform corrections on cardinality estimates instead of the Hit Counting algorithm.
 
@@
 It is interesting to point out that all these implementations use a constant number of buckets and a fixed hash function.
-Thus they are vulnerable to our attacks in scenarios S2 and S4.
+Thus they are potentially vulnerable to our attacks in scenarios S2 and S4.
 
-As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}, since it is faithful to the original description of HLL given in~\cite{hll}.\footnote{Our attack code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\url{https://github.com/PizzaWhisperer/HLLVuln}}.}.
+As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}, since it is faithful to the original description of HLL given in~\cite{hll}.\footnote{Our attack code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\url{https://github.com/PizzaWhisperer/HLLVuln}}.}
+\todo{We need to make sure the repo is publicly accessible but anonymised for submission, if we are going to include it.}
 For simplicity, our items are 4-character long ASCII strings, giving us a set ${\cal X}$ of $7,311,616= 2^{22.8}$ distinct items that can be inserted (hence we have $t=22.8$ in our attacks).
@@
 Secondly, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly $m \cdot \ln (m/(m-B))$ after inserting a fraction $\frac{B}{m}$ of the items, as predicted by our analysis.
+For example, in scenario S1 with $B=100$, the final cardinality estimate is 126 (in agreement with the formula) while with $B=1$ in scenarios S2, S4, the final cardinality estimate is just 1 (again, as predicted by the formula).
 
-We can also compare our results from Table~\ref{table:tab2} with results from the attack presented in Section 5.1 of~\cite{hllvuln}.
+We can also compare our results from Table~\ref{table:tab2} with results from the attack presented in~\cite[Section 5.1]{hllvuln}.
 In that paper, the authors target the Redis~\cite{redis} open-source implementation of HLL and perform the attack against an empty sketch filled with adversarial inputs only.
 Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780''~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality after 4 iterations (resets).
-The S1 adversary following our strategy is able to add 977 and 97,433 distinct items while holding the cardinality estimate at 1 and 126, respectively.
+It should also be noted that the Redis implementation analysed in~\cite{hllvuln} does not conform precisely to the original HLL algorithm as presented in Figure~\ref{fig:hll}.
+As discussed in Section~\ref{sec:redis}, the Redis implementation has changed its HLL algorithm several times in recent releases, but the authors of~\cite{hllvuln} do not specify which version was used in their experiments.
+This makes it difficult to offer a precise comparison with their results.
+Nevertheless, the S1 adversary following our strategy is able to add 977 and 97,433 distinct items while holding the cardinality estimate at 1 and 126, respectively.
 Compared to~\cite{hllvuln}, we can reduce the cardinality estimate by a factor of almost 1,000 instead of 5.
 This comes at the cost of performing more resets compared to~\cite{hllvuln}.
-
-\todo{Check why in the first table in S1 we end up with a card.\ est.\ of 126 instead of $B=100$. The S1 attack description says we should stay at $B=100$ so something is off!}
 
@@
 
+\section{Formal Security Analysis}\label{sec:formal}
+
+\todo{S1: no dice, even with keys everything is broken if shadow device is identical. But that also means identical key.}
+
+\todo{S2-S4: we work in S4, and show that simply using a secretly keyed PRF in palce of the hash function reduces the adversary to the random case which is already used to analyse average case performance of sketches.
+Truncated version of AES would do the job perfectly, and is very fast with hardware instructions.
+PRPness does not matter due to truncation and max number of queries.}
+
+\todo{Discussion: note again that no-one currently does this in practice; also point out that you want different keys in different sketches for security, but this sacrifices the merge capability of HLL.}
+
+\todo{Point out that in the ostensibly weakest attack scenario S1, the shadow device is actually very powerful! It's the rest capability that gives this power and that is missing }
+
 \section{Conclusion}\label{sec:conclusions}
@@
 Its simple structure makes it easy to code and use, as shown by the growing number of available open-source implementations.
-Nonetheless, malicious users can manipulate the HLL cardinality estimate and thence break the security properties of systems relying on HLL.
+Nonetheless, we have shown that malicious users can manipulate the HLL cardinality estimate and thence break the security properties of systems relying on HLL.
 Our attacks are simple but powerful and should raise awareness of the limitations of HLL.
 Our analysis may assist software developers in understanding the risks they run when using HLL in adversarial settings.
+We have also provided a simple method for securing HLL in adversarial settings: simply replace the internal hash function by a keyed PRF, and one obtains full protection even in scenario S4, our strongest attack scenario.
 
@@
 For example, would the imposed ID setting (like Facebook) solve the issue but still be convenient for the various HLL applications?
-Left to future work is the task of extending our attacks to HLL variants such as sliding HLL~\cite{slidinghll}, HLL++~\cite{hllpratice} and the version of HLL used in Redis, which relies on the maximum likelihood approach from~\cite{newhll} to perform corrections on cardinality estimates instead of the Hit Counting algorithm.
+Left to future work is the task of extending our attacks to HLL variants such as sliding HLL~\cite{slidinghll} and HLL++~\cite{hllpractice}.
 
 It would be interesting to find and analyse a strategy for more cleverly ordering the candidates $x$ in our attack in scenario S1 to reduce the total insertion cost of the attack, whilst also keeping the number of resets low.
+We also lack an attack in scenario S1 when the target HLL filter is initially non-empty.
 
-A formal analysis of using salting or secretly keyed hash functions as a countermeasure to our attacks remains open.
-This could be done by extending the framework of~\cite{CCS:ClaPatShrPS19}.
-An interesting cryptographic challenge here is to  improve the security of HLL sketches against adversarial inputs whilst retaining some form of the uncoordinated mergeability that is enjoyed by the original HLL algorithm.
+We have given a formal analysis of using secretly keyed hash functions as a countermeasure to our attacks.
+The reliance on secret keys means sacrificing the mergeability property of HLL sketches in distributed environments (where the keys in the different HLLs are independently generated).
+It is an interesting cryptographic challenge to find methods for securing HLL sketches against adversarial inputs whilst retaining some form of mergeability.
+This may also require the introduction of entirely new data structures that trade increased storage for improved security.
 

### 2020-10-04 docs/HLL_exp_bad/main.tex
adding redis analysis

@@
 
-\subsubsection{Initially Non-empty HLL Sketch, Scenario S2:} 
+\subsubsection{Initially Non-empty HLL Sketch, Scenario S2:}\label{sec:nonemptyHLLS2}
 The adversary knows $h$ so it selects items $x$ from ${\cal X}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit.
@@
 
-\subsubsection{Initially Non-empty HLL Sketch, Scenario S4:} 
+\subsubsection{Initially Non-empty HLL Sketch, Scenario S4:}\label{sec:nonemptyHLLS4}
 The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch).
@@
 where $H_M$ is the harmonic mean of the counts $2^{M[1]}, 2^{M[2]}, \ldots, 2^{M[m]}$.
-So, given ${\cal X}$ containing $2^t$ items, the adversary can expect to insert $(1 - H_M)^{-1} \cdot 2^t$ items from ${\cal X}$ without increasing the HLL cardinality estimate at all.
+So, given ${\cal X}$ containing $2^t$ items, the adversary can expect to insert $(1 - H_M^{-1}) \cdot 2^t$ items from ${\cal X}$ without increasing the HLL cardinality estimate at all.
 
@@
 The above analysis does not change at all because the term $\frac{1}{m} \cdot (1-\frac{1}{2^{M[j]}})$ already signs a probability of 0 to inserting $x$ in bucket $j$ if $M[j] = 0$.
-Hence we still expect to insert $(1 - H_M)^{-1} \cdot 2^t$ items from ${\cal X}$ without increasing the HLL cardinality estimate at all.
+Hence we still expect to insert $(1 - H_M^{-1}) \cdot 2^t$ items from ${\cal X}$ without increasing the HLL cardinality estimate at all.
 Of course, the adversary can relax its attack and allow some of the empty buckets to be hit as well.
@@
 In this setting, where the target HLL sketch is initially non-empty, we see that an adversary who has sufficient flexibility in its inputs (as represented by the size $2^t$ of the set ${\cal X}$) can expect to insert exponentially many (in $t$) values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the case where the HLL sketch has some empty buckets in S2).
-The result is largely independent of the HLL parameters, but does depend on the extent to which the HLL sketch is already filled by other users, via a factor $1-(H_M)^{-1}$: the more full the HLL sketch is, the easier it is to insert new items without them increasing the cardinality estimate.
+The result is largely independent of the HLL parameters, but does depend on the extent to which the HLL sketch is already filled by other users, via a factor $1-H_M^{-1}$: the more full the HLL sketch is, the easier it is to insert new items without them increasing the cardinality estimate.
 
@@
 
-\section{Analysis of the Redis Implementation}\label{sec:redis}
+\section{Analysis of the Redis HLL Implementation}\label{sec:redis}
 
-We present a brief analysis of the HLL implementation used in Redis, since Redis is a very popular key-value store implementation and the version of HLL used there differs significantly from that described in Figure~\ref{fig:hll}.
+We present a brief analysis of the HLL implementation used in Redis, since Redis is a very popular distributed, in-memory key-value store implementation and the version of HLL used there differs significantly from that described in Figure~\ref{fig:hll}.
 
-%and the version of HLL used in Redis, which relies on the maximum likelihood approach from~\cite{newhll} to perform corrections on cardinality estimates instead of the Hit Counting algorithm.
+HLL was first added to Redis in release 2.8.9 in April 2014.
+Redis 4.0 released in April 2017 introduced HLL Beta, which improves on the raw HLL cardinality estimator by using polynomial interpolation to reduce systematic errors.
+Redis 5.0 released in October 2018 replaces this with a new cardinality estimator based closely on the work of Ertl~\cite{hllnew}.
+This method is still used in the current release (Redis 6.0.8) and so is the object of our study here.
+
+Ertl~\cite{hllnew} proposed using the following cardinality estimator in place of the raw estimator in Figure~\ref{fig:hll}:
+\begin{equation}\label{eqn:ertl}
+\hat{\lambda}  = \frac{\alpha_\infty m^2}{m\sigma(C_0/m) + \sum_{k=1}^{\ell}C_k 2^{-k} + m \tau(1-C_{\ell+1}/m)2^{-\ell}} \, .
+\end{equation}
+Here, in the denominator, $C_k$ denotes the number of buckets containing value $k$ in the HLL filter, so $C_0$ is in particular the number of empty buckets; $\sigma(\cdot)$ and $\tau(\cdot)$ are specific functions defined as:
+\[
+\sigma(x) :=  x + \sum_{k=1}^{\infty} x^{2^k} 2^{k-1}
+\]
+and
+\[
+\tau(x) :=  \frac{1}{2} \left( -x + \sum_{k=1}^{\infty} x^{2^{-k}} 2^{-k} \right).
+\]
+Finally,  $\alpha_\infty  = 0.72134752\ldots$ is a constant. (Note that we have translated from the original notation in~\cite{hllnew} to ours.)
+
+Redis 5.0 onwards uses exactly the formula in equation~(\ref{eqn:ertl}) to form its cardinality estimate.
+It uses $n=14$, $m=16,384$ and $\ell = 50$ (so $n+\ell = 64$ and a 64-bit hash function is used).
+
+Note that when evaluating (\ref{eqn:ertl}), functions $\sigma$ and $\tau$ are only called on inputs in the range $[0,1]$.
+Moreover, in this range, we have $\sigma(x) \ge x$.
+Also it is highly likely that $C_{\ell+1} = 0$ (since it is very unlikely that a 50-bit all-0 string will result from the application of the Redis HLL hash function, and in any case in scenarios S2 and S4, we can always arrange for this to be avoided).
+When $C_{\ell+1} = 0$, $\tau$ is evaluated at 1, where its value is zero.
+Combining this information, we see that, provided $C_{\ell+1} = 0$, we have:
+\[
+m\sigma(C_0/m) + \sum_{k=1}^{\ell}C_k 2^{-k} + m \tau(1-C_{\ell+1}/m)2^{-\ell} \ge m\sigma(C_0/m) \ge C_0
+\]
+and so
+\[
+\hat{\lambda}  \leq \frac{\alpha_\infty m^2}{C_0}
+\]
+where, recall, $C_0$ is the number of empty buckets.
+
+Attacks are now straightforward.
+Suppose we can arrange that $C_0$, the number of empty buckets, stays above ${\epsilon}m$, a linear function of the total number of buckets, and that $C_{\ell+1} = 0$.
+Then we obtain $\hat{\lambda}  \leq \alpha_\infty \epsilon^{-1} m$, a linear function of $m$.
+For example, we can set $\epsilon = 1/2$ to obtain $\hat{\lambda}  \leq 2\alpha_\infty m$, where recall that $\alpha_\infty$ is about $0.72$, hence $\hat{\lambda}  \leq 1.44m$.
+
+The requirement on $C_0 \geq {\epsilon}m$ is easily met in scenarios S2 and S4 with an initially empty HLL filter, through the usual trick of sampling inputs $x \in {\cal X}$ so that only a fraction $1-\epsilon$ of the buckets are ever hit.
+We can expect to insert a fraction $1-\epsilon-\epsilon2^{-\ell}$ of the elements of  ${\cal X}$ whilst keeping the cardinality estimate $\hat{\lambda}$ below  $\alpha_\infty \epsilon^{-1} m$.
+(The term $\epsilon2^{-\ell}$ arises from the requirement to keep $C_{\ell+1} =0$ in the attack; recall that $\ell=50$ in Redis, so the effect of this term is very small.)
+
+In scenarios S2 and S4 with an initially non-empty HLL filter, we assume as for our earlier analysis of the standard HLL algorithm that all buckets have already been hit (this will be the case with high probability as soon as $m \log m$ items have been honestly inserted).
+Hence $C_0 = 0$ and $\sigma$ is evaluated at 0 in (\ref{eqn:ertl}), where its value is 0.
+This means that the estimate $\hat{\lambda}$ is computed using a denominator of the form:
+\[
+\sum_{k=1}^{\ell}C_k 2^{-k} + m \tau(1-C_{\ell+1}/m)2^{-\ell}.
+\]
+Now we can proceed exactly as in Sections~\ref{sec:nonemptyHLLS2} and~~\ref{sec:nonemptyHLLS4}, inserting values $x$ whose hashes have rightmost $\ell$ bits beginning with a ``1'' bit (and with a more relaxed condition in S4).
+These insertions do not alter the values of the $C_k$, and hence $\hat{\lambda}$ is unaffected.
+In S2, we expect to insert half the items in ${\cal X}$ without increasing the cardinality estimate $\hat{\lambda}$ at all.
+In S4, we can insert even more items.
+The analysis can be extended in S4 to the case where some of the buckets are still empty: in S4 we can avoid hitting these empty buckets and again the cardinality estimate does not increase at all.
+
+This completes our analysis of Redis.
+The key takeaway is that an adversary can expect to insert a constant fraction of ${\cal X}$ into a non-empty Redis HLL filter whilst not increasing its cardinality estimate at all (and whilst keeping the estimated cardinality linear in $m$ in the case where the filter is initially empty).
+
+\todo{What about S1 here?
+Seems harder because we may not get accurate HLL estimates due to rounding again.
+We need to show that decreasing $C_0$ by 1 is noticeable in Ertl's formula even with rounding.}
 
@@
 It is interesting to point out that all these implementations use a constant number of buckets and a fixed hash function.
-Thus they are potentially vulnerable to our attacks in scenarios S2 and S4.
+Thus they are potentially vulnerable to our attacks.
 

### 2020-10-04 docs/HLL_exp_bad/main.tex
some more notes and tweaks

@@
 \subsubsection{Initially Non-empty HLL Sketch, Scenario S4:}\label{sec:nonemptyHLLS4}
-The adversary can adopt the same strategy as for S2, but make the attack more efficient and reduce the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch).
-Therefore, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
+The adversary can adopt the same strategy as for S2, but make relax the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch).
+Specifically, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
 In other words, if $h(x)$ is mapped to the $j$-th bucket containing $M[j] \ge 1$, then, instead of the strict requirement of having a ``1'' bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a ``1'' bit in any of the first $M[j]$ positions in this substring.
@@
 
-\todo{Fix this -- it's not clear the attack works because it's not clear we really can detect cardinality changes in the way desired (and we've not demonstrated that we can).}
+\todo{Fix this -- it's not clear the attack works because it's not clear we really can detect cardinality changes in the way desired (and we've not demonstrated that we can).
+Explain why it does not work -- to stop reviewers finding wrong solutions.}
 
@@
 This method is still used in the current release (Redis 6.0.8) and so is the object of our study here.
+
+\todo{include some urls for code and release notes.}
 

### 2020-10-12 docs/HLL_exp_bad/main.tex
work on abstract, intro, section 3

@@
 This trade-off makes the HLL algorithm very attractive for a wide range of applications such as database management and network monitoring, where an exact count may not be needed.
-Recently, the HLL algorithm has started to be proposed for use in scenarios where the inputs may be adversarially generated, for example detection of network scanning attacks.
+Recently, the HLL algorithm has started to be proposed for use in scenarios where the inputs may be adversarially generated, for example counting social network users or detection of network scanning attacks.
 This prompts an examination of the performance of the HLL algorithm in the face of adversarial inputs.
 We show that in such a setting, the HLL algorithm's estimate of cardinality can be exponentially bad: when an adversary has access to the internals of the HLL algorithm and has some flexibility in choosing what inputs will be recorded, it can manipulate the cardinality estimate to be exponentially smaller than the true cardinality.
+We study both the original HLL algorithm and a more modern version of it (Ertl, 2017) that is used in Redis.
+We present experimental results confirming our theoretical analysis.
+Finally, we consider attack prevention: we show to modify HLL in a simple way that provably prevents cardinality estimate manipulation attacks.
 \end{abstract}
@@
 It is increasingly used in settings where adversaries may have  incentives to manipulate the estimate made by the algorithm.
-For example,~\cite{portscanhll} proposes its use for detecting network scanning attacks, while Facebook engineers have reported~\cite{fbhll} that they use HLL for estimating the number of active users.
-These applications demand a careful evaluation of the performance of HLL under adversarial input.
+For example:
+\begin{itemize}
+    \item Cardinality estimation provides a significantly faster way for Facebook to compute statistics on their users.
+They use HLL to find out how many distinct people visited their website in the past week~\cite{fbhll}.
+    \item HLL has been proposed as a solution to detecting Denial-of-Service attacks~\cite{portscanhll} (this paper actually uses a variation on HLL called \emph{sliding HLL}~\cite{slidinghll} which estimates the cardinality over a moving time window).
+Specifically, HLL is used to count how many different ports are used as destination port over all packets received, as an attempt to identify when a port scan attack is underway.
+    \item HLL has been implemented in network soft-switches to approximate the number of distinct packets in traffic flows and overcome the switch resource limits~\cite{flexswitch}.
+The number of unique flows traversing a switch can then be used to provide a better congestion control protocol.
+\end{itemize}
 
-Very recently, Reviriego and Ting~\cite{hllvuln} initiated work in this direction, showing that in specific attack settings, the cardinality estimate made by HLL could be modified through input selection.
-In this paper, we present a complete analysis of HLL in the adversarial setting, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
+Regardless of the sensitivity of the above examples, none of the above-cited sources are clear about their threat model, especially for the HLL component.
+They are all potentially vulnerable to the attacks we present in this paper.
+
+Very recently, Reviriego and Ting~\cite{hllvuln} initiated work on the performance of HLL in adversarial settings.
+They showed that in very specific attack settings, the cardinality estimate made by HLL could be modified through input selection.
+In this paper, we present a complete analysis of HLL in adversarial settings, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
 In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of ${\cal O}(2^t)$ where $t$ is the number of bits of flexibility in the adversary's input).
+We demonstrate this theoretically and experimentally for the ``classic'' version of HLL.
+We also study a specific ``modern'' version of HLL which adopts recommendation from~\cite{hllnew} and show how variants of our attacks apply there too.
+This modern version of HLL is interesting because it is the one used in Redis, a popular key-value database technology.
+
+To complement our attack results, we provide a formal security analysis of HLL.
+We show that in the ``shadow device'' attack scenario of~\cite{hllvuln}, no security is possible even if the HLL is secretly keyed.
+On the other hand, we also show that in the most powerful attack setting introduced here, where the attacker has full access to the internals of the HLL (except for a secret key!), it is possible to provably secure HLL against cardinality manipulation by the simple expedient of replacing its internal hash function by a keyed, variable-input-length pseudo-random function (VIL-PRF).
+We describe a suitable instantiation of the required VIL-PRF using off-the-shelf cryptographic components.
 
 The rest of the paper is organised as follows.
+Section~\ref{sec:related} immediately below discusses related work.
 Section~\ref{sec:overview} gives an overview of HLL.
-A brief summary of  related work can be found in Section \ref{sec:related}.
 Section~\ref{sec:attacks} presents our different adversarial models.
-In Section~\ref{sec:manip}, we describe attacks and evaluate their impacts under the different adversarial models.
-The experimental results are discussed in Section~\ref{sec:exp} and Section~\ref{sec:conclusions} provides our conclusions and some ideas for future work.
+In Section~\ref{sec:manip}, we describe attacks on HLL and evaluate their impacts under the different adversarial models.
+Section~\ref{sec:redis} considers the security of the more modern version of HLL due to Ertl~\cite{hllnew} that is used in Redis.
+Our experimental results are discussed in Section~\ref{sec:exp} while Section~\ref{sec:formal} provides our formal security analysis of HLL.
+Section~\ref{sec:conclusions} provides our conclusions and some ideas for future work.
 
-\todo{fix up the intro to refer to new sections, bring in more of the details of related work and HLL applications here.}
+\subsection{Related Work}\label{sec:related}
+
+Desfontaines \emph{et al.}~\cite{cardestprivacy} studied the privacy properties of HLL in two different attack scenarios, referred to as \emph{insider} and \emph{external}.
+The attack target of~\cite{cardestprivacy} is different from ours, being concerned with loss of privacy, whereas we care about manipulating cardinality estimation.
+It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
+In Section~\ref{sec:attacks} we will discuss the two attack scenarios from~\cite{cardestprivacy} in more detail.
+%In particular, we will show that an adversary who can insert items into the HLL and who can access the HLL cardinality estimate (via typical APIs in HLL implementations) as permitted in the external scenario of~\cite{cardestprivacy} can actually realise the same attacks as the insider adversary can, at least insofar as cardinality estimate manipulation is concerned.
+
+In~\cite{hllvuln}, Reviriego and Ting exploit a vulnerability of HLL to manipulate the cardinality estimate, leading to a five-fold reduction in the estimate compared to the true cardinality.
+We consider the attack model used in~\cite{hllvuln} to be quite artificial, since it gives the adversary the ability to test whether inserting an item \emph{would} increase the cardinality estimate, without actually having to insert the item.
+After some clarifications obtained through personal correspondence with the authors of~\cite{hllvuln}, we understand that such a setup might be possible under certain circumstances, for example where the adversary has access to an identical ``shadow'' device; see further discussion in Section~\ref{sec:attacks}.
+For completeness, we include analyses of this attack scenario, showing that we can drastically improve on the attacks in~\cite{hllvuln} (e.g.\ we can reduce the HLL cardinality estimate by a factor of almost 1,000 instead of the factor of 5 achieved in~\cite{hllvuln}).
+
+Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted sketch, either as a replacement for (but at the cost of losing mergeability) or in addition to (inducing a memory overhead) an unsalted sketch.
+We show in Section~\ref{sec:formal} that when a salt is used appropriately (i.e.\ as the secret key for a pseudo-random function that is used in place of hash function $h$), security against adversarial inputs can be achieved even in our strongest adversarial setting.
+%This might not be a convenient solution for some applications, they thus encourage further research.
+
+Clayton \emph{et al.}~\cite{CCS:ClaPatShrPS19} provide a provable-security treatment of probabilistic data structures in adversarial environments more generally.
+They use their formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures, but not HLL.
+%It is an interesting open problem to use their framework to develop a formal understanding of how HLL can be made secure in the adversarial setting.
 
@@
 A similar correction is applied for large cardinalities in order to take into account the likelihood of hashing collisions.
-As a result, HLL is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
+In addition, the cardinality estimate is usually rounded to a whole number before being output.
+
+HLL as described in Figure~\ref{fig:hll} is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
 
@@
 Such a simple set union operation allows easy parallelisation of operations among multiple machines independently, provided they use the same hash function and the same number of buckets.
+Clearly this procedure cannot be carried out if the different HLL sketches are differently seeded or keyed.
 
-\begin{figure}[ht]
+\begin{figure}[tb!]
 \centering
@@
 \textbf{if} $E \le \frac{1}{30}2^{32}$ \textbf{then}\\
-    \hspace{2em} \textbf{set} $E^*:= E$; \hfill \textit{\{intermediate range—no correction\}}\\
+    \hspace{2em} \textbf{set} $E^*:= E$; \hfill \textit{\{intermediate range --- no correction\}}\\
 \textbf{if} $E > \frac{1}{30}2^{32}$ \textbf{then}\\
@@
 \end{figure}
-
-\subsection{Applications}
-For completeness and motivation, we provide examples of concrete uses of HLL:
-\begin{itemize}
-    \item Cardinality estimation provides a significantly faster way for Facebook to compute statistics on their users.
-They use HLL to find out how many distinct people visited their website in the past week~\cite{fbhll}.
-    \item HLL has been proposed as a solution to detecting Denial-of-Service attacks~\cite{portscanhll} (this paper actually uses a variation on HLL called \emph{sliding HLL}~\cite{slidinghll} which estimates the cardinality over a moving time window).
-Specifically, HLL is used to count how many different ports are used as destination port over all packets received, as an attempt to identify when a port scan attack is underway.
-    \item HLL has been implemented in network soft-switches to approximate the number of distinct packets in traffic flows and overcome the switch resource limits~\cite{flexswitch}.
-The number of unique flows traversing a switch can then be used to provide a better congestion control protocol.
-\end{itemize}
-
-Regardless of the sensitivity of the above examples, none of the above-cited papers are clear about their threat model, especially for the HLL component.
-They are all potentially vulnerable to the attacks we present below.
-
-\section{Related Work}\label{sec:related}
-
-Desfontaines \emph{et al.}~\cite{cardestprivacy} studied the privacy properties of HLL in two different attack scenarios, referred to as \emph{insider} and \emph{external}.
-The attack target of~\cite{cardestprivacy} is different from ours, being concerned with loss of privacy, whereas we care about manipulating cardinality estimation.
-It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
-In Section~\ref{sec:attacks} we will discuss the two attack scenarios from~\cite{cardestprivacy} in more detail.
-In particular, we will show that an adversary who can insert items into the HLL and who can access the HLL cardinality estimate (via typical APIs in HLL implementations) as permitted in the external scenario of~\cite{cardestprivacy} can actually realise the same attacks as the insider adversary can, at least insofar as cardinality estimate manipulation is concerned.
-
-In~\cite{hllvuln}, Reviriego and Ting exploit a vulnerability of HLL to manipulate the cardinality estimate, leading to a five-fold reduction in the estimate compared to the true cardinality.
-We consider the attack model used in~\cite{hllvuln} to be quite artificial, since it gives the adversary the ability to test whether inserting an item \emph{would} increase the cardinality estimate, without actually having to insert the item.
-After some clarifications obtained through personal correspondence with the authors of~\cite{hllvuln}, we understand that such a setup might be possible under certain circumstances, for example where the adversary has access to an identical ``shadow'' device; see further discussion in Section~\ref{sec:attacks}.
-
-Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted sketch, either as a replacement for (but at the cost of losing mergeability) or in addition to (inducing a memory overhead) an unsalted sketch.
-We show in Section~\ref{sec:formal} that when a salt is used appropriately (i.e.\ as the secret key for a pseudo-random function that is used in place of hash function $h$), security against adversarial inputs can be achieved even in our strongest adversarial setting.
-%This might not be a convenient solution for some applications, they thus encourage further research.
-
-Clayton \emph{et al.}~\cite{CCS:ClaPatShrPS19} provide a provable-security treatment of probabilistic data structures in adversarial environments more generally.
-They use their formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures, but not HLL.
-%It is an interesting open problem to use their framework to develop a formal understanding of how HLL can be made secure in the adversarial setting.
 
@@
 
+Scenario S1 is rather distinctive in that it is the only one in which the adversary has access via an API to a shadow device (and no other capabilities).
 We do not consider scenario S1 typical of HLL deployments.
@@
 This would make any pre-computation on the items using the shadow device irrelevant, and considerably weaken the attack presented in Section 3.2 of~\cite{hllvuln}.
-Hence we assume for the rest of this work that, in scenario S1, shadow and targeted devices have the same internals and security parameters.
+Hence we assume for the rest of this work that, in scenario S1, shadow and targeted devices have the same internals and security parameters (if there are any).
 If the HLL device receives inputs from several users, the attack from Section 3.2 of~\cite{hllvuln} is then the easiest to carry out.
@@
 The idea is to iterate through the process of trial insertion several times to refine the list of items, before inserting what remains into the targeted device.
-We explore in this work scenario S1 in the special case where inputs come only from the adversary, and propose an attack that provides better results than the one presented in~\cite{hllvuln}.
+We explore in this work scenario S1 in the special case where inputs come only from the adversary, and propose an attack that provides significantly better results than the one presented in~\cite{hllvuln}.
 
-Furthermore, we show that although~\cite{cardestprivacy} presents S3 (external adversary) and the stronger version of our S4 (insider adversary) as two separate scenarios, an S3 adversary can easily infer the knowledge given to our S4 adversary in its snapshot.
+Scenarios S2, S3 and S4 gradually increase the power of the adversary.
+
+At first sight it might appear that an S3 adversary can easily infer the knowledge given to our S4 adversary in its snapshot.
+Indeed, an S3 adversary could try to recover the contents of each bucket in the sketch simply by analysing which inputs lead to increases in the cardinality estimate: to find the value held in a targeted bucket, the adversary would insert a sequence of items $x$ with an incrementing number of leading zeros in the $\ell$ rightmost bits of $h(x)$ whilst holding the $n$ bits determining the bucket index constant, until the cardinality estimate of the sketch increases.
+This assumes the adversary has sufficient flexibility in its input that the conditions on $h(x)$ can be forced.
+
+However, this procedure does not work as simply as just described because of numerical issues that arise in the computation of the cardinality estimate.
+Specifically, increasing the value held in some bucket from $M[j]$ to $M[j]+1$ \emph{decreases} the denominator in the raw cardinality estimate $\sum_{j=1}^m2^{-M[j]}$ by an amount $2^{-M[j]-1}$.
+This does lead to a change in the raw cardinality estimate $E = \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$.
+However, in practice, the raw cardinality estimate is rounded before being presented to the adversary, so this change may not actually be detectable.
+We have done experiments which shows this to be the case for parameters arising in practice.
+
+We do not know of attacks which are able to use the extra information available to an S3 adversary over an S2 adversary (namely, the ability to obtain cardinality estimates).
+Moreover, as we will see below, our attacks in the S2 setting are already quite powerful.
+For this reason, we will focus in the remainder of the paper on scenarios S1, S2 and S4.
+
+%Furthermore, we show that although~\cite{cardestprivacy} presents S3 (external adversary) and the stronger version of our S4 (insider adversary) as two separate scenarios, an S3 adversary can easily infer the knowledge given to our S4 adversary in its snapshot.
 Indeed, an adversary can easily recover the contents of each bucket in the sketch simply by analysing which inputs increase the cardinality estimate.
@@
 We may need to heavily revise all this text.}
- Once an increase occurs, the adversary knows that the maximum number of leading zeros observed in that bucket was just updated to a value that can be recovered by looking at the last inserted item.
+% Once an increase occurs, the adversary knows that the maximum number of leading zeros observed in that bucket was just updated to a value that can be recovered by looking at the last inserted item.
 This assumes the adversary has sufficient flexibility in its choice of inputs $x$ so that the relevant bit conditions on $h(x)$ can be forced.
@@
 
-Since an adversary can quite efficiently escalate from S3 to S4, we focus on scenarios S1, S2 and S4 in what follows.
+%Since an adversary can quite efficiently escalate from S3 to S4, we focus on scenarios S1, S2 and S4 in what follows.
 

### 2020-10-12 docs/HLL_exp_bad/main.tex
more rewriting

@@
 Relatively small and big cardinalities are handled separately to correct for systematic distortions, as shown in Fig.~\ref{fig:hll}.
-When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate, looking at the number of empty buckets (whose register is 0) instead of the number of leading zeros in the stream.
+When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate, looking at the number of empty buckets (whose register value is 0) instead of the number of leading zeros in the stream.
 A similar correction is applied for large cardinalities in order to take into account the likelihood of hashing collisions.
@@
 Finally, we briefly consider scenario S1 with an initially non-empty target HLL sketch.
-
-\todo{Fix this -- it's not clear the attack works because it's not clear we really can detect cardinality changes in the way desired (and we've not demonstrated that we can).
-Explain why it does not work -- to stop reviewers finding wrong solutions.}
+Since the target sketch has an unknown status for the adversary, our previous strategy for S1 of keeping many buckets empty does not work.
+One can attempt to use the shadow device to identify items that would result in small entries $M[j]$ (for unknown indices $j$) since such items are unlikely to increase the cardinality estimate of the target device by much.
+This requires first filling the shadow device with sufficiently many values to escape from the ``small range correction'' phase.
+Then one could enter random values $x$ into the shadow device and check for cardinality increases.
+However this runs into the rounding issue described in Section~\ref{sec:scenarios-discussion}.
+Still, one could retain those items which do lead to a detectable change in cardinality and reject the others (and also estimate the value of $M[j]$ for the unknown bucket $j$ from the magnitude of the change, retaining only those for which $M[j]$ is small).
+But here the lack of a cardinality change could result either from rounding issues or because insertion of $x$ into the shadow device really did lead to no increase in cardinality.
+This indicates that  ``good'' items would be rejected unnecessarily.
+This could be addressed by using multiple resets, filling with random items, and then multiple insertions per item to try to estimate the true value of $M[j]$ for each item.
+The details of such an attack require further analysis which we do not pursue here, since we consider S1 to be the least realistic attack scenario.
 
@@
 Redis 5.0 released in October 2018 replaces this with a new cardinality estimator based closely on the work of Ertl~\cite{hllnew}.
-This method is still used in the current release (Redis 6.0.8) and so is the object of our study here.
-
-\todo{include some urls for code and release notes.}
+This method is still used in the current release (Redis 6.0.8) and so is the object of our study here.\footnote{For details of Redis releases, see \url{https://github.com/redis/redis/blob/6.0/00-RELEASENOTES} and other files obtained by replacing ``6.0'' in this URL with the desired version number.}
 
@@
 
-The requirement on $C_0 \geq {\epsilon}m$ is easily met in scenarios S2 and S4 with an initially empty HLL filter, through the usual trick of sampling inputs $x \in {\cal X}$ so that only a fraction $1-\epsilon$ of the buckets are ever hit.
+The requirement on $C_0 \geq {\epsilon}m$ is easily met in scenarios S2 and S4 with an initially empty HLL filter, through the trick of sampling inputs $x \in {\cal X}$ so that only a fraction $1-\epsilon$ of the buckets are ever hit.
 We can expect to insert a fraction $1-\epsilon-\epsilon2^{-\ell}$ of the elements of  ${\cal X}$ whilst keeping the cardinality estimate $\hat{\lambda}$ below  $\alpha_\infty \epsilon^{-1} m$.
@@
 This completes our analysis of Redis.
-The key takeaway is that an adversary can expect to insert a constant fraction of ${\cal X}$ into a non-empty Redis HLL filter whilst not increasing its cardinality estimate at all (and whilst keeping the estimated cardinality linear in $m$ in the case where the filter is initially empty).
+The key takeaway is that, in scenarios S2 and S4, an adversary can expect to insert a constant fraction of ${\cal X}$ into a non-empty Redis HLL filter whilst not increasing its cardinality estimate at all (and whilst keeping the estimated cardinality linear in $m$ in the case where the filter is initially empty).
 
-\todo{What about S1 here?
+%\todo{What about S1 here?
 Seems harder because we may not get accurate HLL estimates due to rounding again.
@@
 
-As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}, since it is faithful to the original description of HLL given in~\cite{hll}.\footnote{Our attack code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\url{https://github.com/PizzaWhisperer/HLLVuln}}.}
-\todo{We need to make sure the repo is publicly accessible but anonymised for submission, if we are going to include it.}
+As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}, since it is faithful to the original description of HLL given in~\cite{hll}.\footnote{We will make our attack code available on acceptance.}
+%at \href{https://github.com/PizzaWhisperer/HLLVuln}{\url{https://github.com/PizzaWhisperer/HLLVuln}}.} 
+%\todo{We need to make sure the repo is publicly accessible but anonymised for submission, if we are going to include it.}
 For simplicity, our items are 4-character long ASCII strings, giving us a set ${\cal X}$ of $7,311,616= 2^{22.8}$ distinct items that can be inserted (hence we have $t=22.8$ in our attacks).
@@
 For simplicity and to keep the attack as striking as possible, we set $B$ to 1 in the attacks on an initially empty HLL sketch in scenarios S2 and S4.
-All attacks were run 30 times; we report averge outcomes.
+All attacks were run 30 times; we report average outcomes.
 Our results are reported in Tables~\ref{table:tab1} and~\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary, and the cardinality estimate after adding the adversary's items.
@@
 
-Secondly, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly $m \cdot \ln (m/(m-B))$ after inserting a fraction $\frac{B}{m}$ of the items, as predicted by our analysis.
+Secondly, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly a rounded version of the Hit Counting estimator $m \cdot \ln (m/(m-B))$ after inserting a fraction $\frac{B}{m}$ of the items, as predicted by our analysis.
 For example, in scenario S1 with $B=100$, the final cardinality estimate is 126 (in agreement with the formula) while with $B=1$ in scenarios S2, S4, the final cardinality estimate is just 1 (again, as predicted by the formula).
@@
 We can also compare our results from Table~\ref{table:tab2} with results from the attack presented in~\cite[Section 5.1]{hllvuln}.
-In that paper, the authors target the Redis~\cite{redis} open-source implementation of HLL and perform the attack against an empty sketch filled with adversarial inputs only.
+In that paper, the authors target the Redis implementation of HLL and perform the attack against an empty sketch filled with adversarial inputs only.
 Given a set of 250,000 items, the adversary could choose ``74,390 distinct items and obtain an HLL estimate of only 15,780''~\cite{hllvuln}, achieving a five-fold reduction from the true cardinality after 4 iterations (resets).
-It should also be noted that the Redis implementation analysed in~\cite{hllvuln} does not conform precisely to the original HLL algorithm as presented in Figure~\ref{fig:hll}.
-As discussed in Section~\ref{sec:redis}, the Redis implementation has changed its HLL algorithm several times in recent releases, but the authors of~\cite{hllvuln} do not specify which version was used in their experiments.
+As explained in Section~\ref{sec:redis}, the Redis implementation analysed in~\cite{hllvuln} does not conform precisely to the original HLL algorithm as presented in Figure~\ref{fig:hll}.
+Also as discussed in Section~\ref{sec:redis}, the Redis implementation has changed its HLL algorithm several times in recent releases, but the authors of~\cite{hllvuln} do not specify which version was used in their experiments.
 This makes it difficult to offer a precise comparison with their results.
@@
     \textbf{\# Targeted buckets $B$} & 1 & 100 & n/a & 1 & n/a & 1 \\ \hline
-    \textbf{Initially empty?} & yes & yes & no & yes & no & yes \\ \hline
+    \textbf{HLL initially empty?} & yes & yes & no & yes & no & yes \\ \hline
     \textbf{Original Card. Est.} & 0 & 0 & 1,000 & 0 & 1,005 & 0 \\ \hline
@@
     \textbf{\# Targeted buckets $B$ } & 1 & 100 & n/a & 1 & n/a & 1 \\ \hline
-    \textbf{Initially empty?} & yes & yes & no & yes & no & yes \\ \hline
+    \textbf{HLL initially empty?} & yes & yes & no & yes & no & yes \\ \hline
     \textbf{Original Card. Est.} & 0 & 0 & 1,004 & 0 & 998 & 0 \\ \hline
@@
 
-It would be interesting to find and analyse a strategy for more cleverly ordering the candidates $x$ in our attack in scenario S1 to reduce the total insertion cost of the attack, whilst also keeping the number of resets low.
-We also lack an attack in scenario S1 when the target HLL filter is initially non-empty.
+It would be interesting to find and analyse a strategy for more cleverly ordering the candidates in our attack in scenario S1 to reduce the total insertion cost of the attack, whilst also keeping the number of resets low.
+%We also lack an attack in scenario S1 when the target HLL filter is initially non-empty.
 
@@
 It is an interesting cryptographic challenge to find methods for securing HLL sketches against adversarial inputs whilst retaining some form of mergeability.
-This may also require the introduction of entirely new data structures that trade increased storage for improved security.
+This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security.
 

### 2020-10-12 docs/HLL_exp_bad/main.tex
first complete draft

@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
 
@@
     \hspace{2em} \textbf{set} $j = 1 + <y_1y_2...y_{n}>_2$; \hfill \textit{\{bucket index determined by the first $n$ bits of $y$\}}\\
-    \hspace{2em} \textbf{set} $w := y_{n+1}y_{n+2} \ldots y_{n+\ell+1}$;\hfill \textit{\{bit string determined by the last $\ell$ bits of $y$\}}\\
+    \hspace{2em} \textbf{set} $w := y_{n+1}y_{n+2} \ldots y_{n+\ell}$;\hfill \textit{\{bit string determined by the last $\ell$ bits of $y$\}}\\
    \hspace{2em} \textbf{set} $M[j] :=$ max$(M[j], \rho(w))$;\\
@@
 S4 is a slightly weaker version of the \textit{insider} attack scenario in~\cite{cardestprivacy}, the difference being that an insider adversary in~\cite{cardestprivacy} has continuous access to the HLL sketch internals including all bucket values, while our adversary in S4 only has access to a snapshot of the HLL sketch internals at the start of its attack.
+We consider a weaker version of S4 than in~\cite{cardestprivacy} because it is sufficient for our attacks; when considering security proofs in Section~\ref{sec:formal}, we will switch back to the stronger version of S4.
 
@@
 
-\begin{table}[h]
+\begin{table}[tb!]
 \centering
@@
 
-\todo{S1: no dice, even with keys everything is broken if shadow device is identical. But that also means identical key.}
+In this section we present a formal security analysis of HLL, focussing on whether we can augment the original HLL design with cryptographic mechanisms and thereby provably prevent attacks.
 
-\todo{S2-S4: we work in S4, and show that simply using a secretly keyed PRF in palce of the hash function reduces the adversary to the random case which is already used to analyse average case performance of sketches.
-Truncated version of AES would do the job perfectly, and is very fast with hardware instructions.
-PRPness does not matter due to truncation and max number of queries.}
+\subsection{Scenario S1}
+We begin by disposing of scenario S1.
+In that scenario, the attacker has access to insertion and reset capabilities, and to cardinality estimates, for a shadow HLL that is identical in every respect to the target HLL.
+The attacks we presented previously are oblivious to any internals of the implementation (other than that they follow the HLL description in Figure~\ref{fig:hll}).
+This means that, say, replacing the hash function $h$ with a keyed function or any other cryptographic primitive will be futile in preventing attacks.
+This illustrates the powerful nature of the resettable shadow device (despite our attacks in that setting being more complex and requiring more effort than in scenarios S2 and S4).
+In short, we cannot hope to gain any security in scenario S1 by adding cryptography.
 
-\todo{Discussion: note again that no-one currently does this in practice; also point out that you want different keys in different sketches for security, but this sacrifices the merge capability of HLL.}
+\subsection{Formal Security Analysis for S2, S3, S4}
 
-\todo{Point out that in the ostensibly weakest attack scenario S1, the shadow device is actually very powerful! It's the rest capability that gives this power and that is missing }
+For scenarios S2, S3 and S4, the picture is much more positive.
+In the remainder of this section, we introduce a formal attack model covering all three scenarios.
+This is done by allowing an adversary ${\cal A}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios.
+In the security model, ${\cal A}$ interacts in one of two worlds.
+In the real world, ${\cal A}$ interacts with a real HLL, operating as in Figure~\ref{fig:hll}.
+In the ideal world, ${\cal A}$ interacts with an ideal HLL in which the operation $y: = h(x)$ is replaced by random sampling; that is, by setting $y \leftarrow_{\$} \{0,1\}^{n+\ell}$.
+
+We then show that, if $h$ is replaced by a keyed Pseudo Random Function (PRF) $F$ whose key is not available to the adversary, then any adversary ${\cal A}$ that can distinguish between the real and random worlds with oracles as in S4 can be used to build an adversary ${\cal B}$ which breaks the PRF security of $F$.
+Under the assumption that $F$ is a good PRF, no such efficient adversary ${\cal B}$ can exist, and we may conclude that we can safely switch from the real world to the ideal world without incurring any significant change to the behaviour of ${\cal A}$.
+
+However, it is clear that any adversary in the ideal world is reduced to operating in the setting used in the average case analysis of HLL (wherein every insertion query results in a random value $y$ being used to update the HLL).
+Thus, ${\cal A}$ can have no advantage when it comes to influencing the HLL cardinality estimate.
+It follows that HLL, when instantiated with a keyed PRF, is secure against attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
+
+We will conclude by describing how the necessary keyed PRF can be instantiated.
+
+\subsection{Formal Security Model for HLL}
+
+We consider adversary ${\cal A}$ to be an algorithm running against a challenger ${\cal C}$ in a security game.
+The game begins with ${\cal C}$ selecting a bit $b$ uniformly at random.
+Adversary ${\cal A}$ interacts with up to four oracles which are handled by ${\cal C}$; different combinations of oracles define scenarios S2, S3, S4.
+After making some queries, ${\cal A}$ returns a bit $b'$ estimating whether it is interacting with a real HLL instantiation ($b'=0$) or an ideal one ($b'=1$).
+We define ${\cal A}$'s S$i$ Real-or-Ideal advantage against HLL to be:
+\[
+\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny HLL}}({\cal A}) := \lvert 2\Pr[b' = b] - 1 \rvert \, .
+\]
+%Using a standard transformation, this probability can be rewritten as $\lvert \Pr[b' = 1| b = 1] - \Pr[b'=0|b=1] \rvert$. 
+The four oracles that  ${\cal A}$ can interact with are as follows:
+\begin{itemize}
+\item \texttt{init}: this oracle performs initialisation of the HLL (and selection of the PRF key when $h$ is replaced by a PRF $f$); this oracle is available in all of S2, S3, S4, but can only be called once as ${\cal A}$'s first oracle query.
+\item \texttt{insert($\cdot$)}: on input $x \in {\cal X}$, the oracle inserts value $x$ into the HLL.
+This oracle is available in all of S2, S3, S4.
+Without loss of generality (since the HLL and variants of it we consider here have a deterministic insertion procedure), we assume all queries $x$ made to \texttt{insert($\cdot$)} are distinct.
+We let $q$ denote the number of insertion queries made by ${\cal A}$ during the course of its attack.
+\item \texttt{read}: this oracle returns the HLL cardinality estimate. We allow multiple queries to this oracle and make it available to ${\cal A}$ in S3 and S4, but not S2.
+\item \texttt{corrupt}: this oracle returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents.
+This oracle is available only in S4.
+We allow the adversary to query \texttt{corrupt} as often as it wishes during its attack.
+Our description of S4 in Section~\ref{sec:attacks} limited the adversary to at most one such query; we allow multiple queries here because doing so does not increase the difficulty of achieving security.
+\end{itemize}
+
+We note that in S4 (and only in S4) the adversary has access to all four oracles.
+This makes it the strongest model, and hereafter we only target security in this model.
+We count queries $q$ to \texttt{insert($\cdot$)} because this will enable us to quantify the quality of our security reduction to PRF security below.
+
+\subsection{Security Proof for HLL Using a PRF}
+
+We now replace the hash function $h$ acting on ${\cal X}$ with a keyed function $F: \{0,1\}^k \times {\cal X} \rightarrow  \{0,1\}^{n+\ell}$.
+Here the first input to $F$ is a $k$-bit string, the key, denoted $K$.
+We assume that \texttt{init} in the HLL security model now initialises key $K$ by selecting it uniformly at random from $\{0,1\}^k$.
+
+We define the PRF advantage of an adversary ${\cal B}$ against $F$ outputting a bit $d'$ as being its advantage in distinguishing whether it is interacting with $F$ for a uniformly random key $K$ ($d=0$) or interacting with a truly random function ($d=1$).
+Thus:
+\[
+\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B}) := \lvert 2\Pr[d' = d] - 1 \rvert\,.
+\]
+Here, we assume without loss of generality that all of ${\cal B}$'s queries are distinct. See~\cite[Section 4.4.1]{BonehShoup} for details.
+
+Now we are ready to state:
+
+\begin{theorem}\label{thm:security-of-HLL-with-PRF}
+Suppose ${\cal A}$ is an S4 adversary against HLL instantiated with a keyed function $F: \{0,1\}^k \times {\cal X} \rightarrow  \{0,1\}^{n+\ell}$.
+Then we can construct from ${\cal A}$ a PRF adversary ${\cal B}$ against $F$ such that:
+\[
+\mbox{Adv}^{\mbox{\tiny S4-RoI}}_{\mbox{\tiny HLL}}({\cal A}) = \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B}) \,.
+\]
+Moreover, if ${\cal A}$ makes $q$ queries to its \texttt{insert($\cdot$)} oracle, then ${\cal B}$ makes $q$ queries to its real or random oracle.
+Finally, the running time of ${\cal B}$ is (roughly) the same as that of ${\cal A}$.
+\end{theorem}
+
+The interpretation of the above theorem is that if $F$ is a good PRF, then $\mbox{Adv}^{\tiny\mbox{PRF}}_{F}({\cal B})$ should be small for any adversary ${\cal B}$ running in a reasonable amount of time.
+Hence we can conclude that the advantage of any S4 adversary against HLL using PRF $F$ is also small.
+
+\begin{proof}
+PRF adversary ${\cal B}$ against $F$ is built as follows.
+Let $d$ denote the hidden bit determining whether ${\cal B}$ receives real responses of the form $F(K,x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${\cal B}$ runs S4 adversary ${\cal A}$ and responds to each of its queries as follows:
+\begin{itemize}
+\item \texttt{init}: here ${\cal B}$ initialises an HLL filter in accordance with Figure~\ref{fig:hll}. Thus it initialises an array $M[j]$ in which each entry is set to 0. 
+\item \texttt{insert($\cdot$)}: on input $x \in {\cal X}$,  ${\cal B}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${\cal B}$ uses $y$ to update the HLL bucket $j$ generated via $j=1 + <y_1y_2...y_{n}>_2$ an in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
+\item \texttt{read}: ${\cal B}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate in accordance with Figure~\ref{fig:hll}. 
+\item \texttt{corrupt}: this oracle returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents. 
+\end{itemize}
+When ${\cal A}$ outputs a bit $b'$, ${\cal B}$ outputs its own bit $d'=b'$.
+Note that when $d=0$, ${\cal B}$ plays against the real $F$ with a random key $K$ and provides a simulation for ${\cal A}$ that perfectly instantiates the real world.
+On the other hand, when $d=1$, ${\cal B}$ plays against a truly random function and provides a simulation for ${\cal A}$ that perfectly instantiates the ideal world.
+The equality of advantages in the statement of the theorem now follows immediately.
+The remainder of the proof follows on counting queries and noting that the HLL algorithm for computing cardinality is efficient, so ${\cal B}$'s running time when handling \texttt{read} queries is not increased significantly compared to that of ${\cal A}$.
+\end{proof}
+
+Note that the above proof still goes through if the classical HLL algorithm is replaced with the improved algorithm of~\cite{hllnew} that is used in Redis.
+The only changes to the proof are to use the Redis approach for computing responses to \texttt{read} queries in place of the classical HLL approach.
+
+Finally, we conclude our analysis by reiterating the key point: we have shown that no efficient S4 adversary ${\cal A}$ can tell that it is playing against HLL implemented using truly random values $y$ instead of with values obtained from a PRF, except with a small advantage.
+This means that ${\cal A}$ is effectively reduced to attacking HLL in the exact same setting used in the average case analysis of HLL from~\cite{hll}, wherein it is proved that the HLL cardinality estimate endures relative errors of the form $1.04/\sqrt{m}$.
+
+\subsection{Instantiating the PRF}
+
+The above construction requires a PRF $F: \{0,1\}^k \times {\cal X} \rightarrow  \{0,1\}^{n+\ell}$.
+Here, the set ${\cal X}$ represents the space of possible inputs to the HLL which can be modelled by a set of bit-strings of some maximum length.
+This mean that $F$ needs to be Variable Input Length PRF (VIL-PRF).
+
+Making $F$ competitive in speed with existing implementations of HLL, which use non-cryptographic hash functions such as Murmur3, is challenging, and in practice we must sacrifice some speed for security.
+
+A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {\cal X} \rightarrow  {\cal D}$ onto some domain ${\cal D}$ with a PRF $G: \{0,1\}^k \times {\cal D} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
+It is easy to show that the result is a PRF of the required input-output signature (see for example~\cite[Section 8.2]{BonehShoup}).
+As a concrete instantiation, one could take $H$ to be SHA-256 with outputs truncated to 128 bits and $G$ to be AES-128 truncated to ${n+\ell}$ bits (since ${n+\ell} \leq 64 < 128$ in practice, this truncation is always possible).
+
+This construction would be relatively fast on commodity platforms due to widespread hardware support for SHA-256 and AES operations.
+The exact slowdown relative to, say, Murmur3, depends on platform specifics, but can be expected to be in the range 2-fold to 5-fold.
+
+An alternative is to simply use HMAC-SHA-256 with an appropriately truncated output. HMAC is a PRF under reasonable assumptions on the hash function used in its construction~\cite{Bellare15}.
+
+More complex but faster constructions using randomised or nonce-based MACs constructed from universal hashing could be envisaged.
+However, these do not seem to be directly applicable here since we need the function $F$ to be both stateless and deterministic so as to cope with the repeated inputs one would encounter during the normal operation of HLL.
+
+It is an important open problem to close the performance gap that exists between fast but non-cryptographic hashing used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security in the face of adversarial inputs.
+
+We also note that using keyed functions sacrifices the mergeability property that HLL enjoys, if different keys $K$ are used in different HLL filters.
+The reason is that one can no longer emulate running a single filter by taking the maximum over all buckets with each index $j$ in the different filters, since now the bucket contents are created using different keys in a PRF and hence bear no relation to one another.
+
+%\todo{Discussion: note again that no-one currently does this in practice; also point out that you want different keys in different sketches for security, but this sacrifices the merge capability of HLL.}
 
@@
 For example, would the imposed ID setting (like Facebook) solve the issue but still be convenient for the various HLL applications?
-Left to future work is the task of extending our attacks to HLL variants such as sliding HLL~\cite{slidinghll} and HLL++~\cite{hllpractice}.
+Left to future work is the task of extending our attacks and defences to HLL variants such as sliding HLL~\cite{slidinghll} and HLL++~\cite{hllpractice}.
 
 It would be interesting to find and analyse a strategy for more cleverly ordering the candidates in our attack in scenario S1 to reduce the total insertion cost of the attack, whilst also keeping the number of resets low.
-%We also lack an attack in scenario S1 when the target HLL filter is initially non-empty.
+Finding an attack in scenario S1 when the target HLL filter is initially non-empty is also open.
 
-We have given a formal analysis of using secretly keyed hash functions as a countermeasure to our attacks.
+We have given a formal analysis of using a secretly-keyed PRF as a countermeasure to our attacks.
 The reliance on secret keys means sacrificing the mergeability property of HLL sketches in distributed environments (where the keys in the different HLLs are independently generated).

### 2020-10-12 docs/HLL_exp_bad/main.tex
typo

@@
 
-The interpretation of the above theorem is that if $F$ is a good PRF, then $\mbox{Adv}^{\tiny\mbox{PRF}}_{F}({\cal B})$ should be small for any adversary ${\cal B}$ running in a reasonable amount of time.
+The interpretation of the above theorem is that if $F$ is a good PRF, then $\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B})$ should be small for any adversary ${\cal B}$ running in a reasonable amount of time.
 Hence we can conclude that the advantage of any S4 adversary against HLL using PRF $F$ is also small.

### 2020-10-13 docs/HLL_exp_bad/main.tex
small changes to formal analysis

@@
 Many different implementations of HLL are available online.
-Among the repositories collecting more than 350 stars on GitHub are ~\cite{clahll,datasketch}.
+Among the repositories collecting more than 350 stars on GitHub are~\cite{clahll,datasketch}.
 HLL is also featured in many frameworks, including Redis~\cite{redis}, Google's BigQuery~\cite{bigquery}, Facebook's Airlift~\cite{airlift} and several products of Apache such as Spark~\cite{spahll} and Druid~\cite{druhll}.
@@
 
-First, when attacking a sketch already containing 1000 random items, we can see that the adversary is on average able to insert about half of the items in our starting set in scenario S2, and about 83\% in scenario S4.
+First, when attacking a sketch already containing 1000 random items, we can see that the adversary is on average able to insert about half of the $2^{22.8}$ items in our starting set in scenario S2, and about 83\% in scenario S4.
+The cardinality estimate is, on average, increased slightly in S2 (from 1,000 to 1,011) and remains at the same starting value (1,005 on average) in S4.
+In both cases, we inserted more than 3 million items whilst increasing the cardinality estimate by about $0.1\%$!
 These results align well with our theoretical analysis.
@@
 Also as discussed in Section~\ref{sec:redis}, the Redis implementation has changed its HLL algorithm several times in recent releases, but the authors of~\cite{hllvuln} do not specify which version was used in their experiments.
-This makes it difficult to offer a precise comparison with their results.
-Nevertheless, the S1 adversary following our strategy is able to add 977 and 97,433 distinct items while holding the cardinality estimate at 1 and 126, respectively.
+This makes it difficult to make a precise comparison with their results.
+Nevertheless, the S1 adversary following our strategy against the classical HLL algorithm is able to add 977 and 97,433 distinct items while holding the cardinality estimate at 1 and 126, respectively.
 Compared to~\cite{hllvuln}, we can reduce the cardinality estimate by a factor of almost 1,000 instead of 5.
@@
 The attacks we presented previously are oblivious to any internals of the implementation (other than that they follow the HLL description in Figure~\ref{fig:hll}).
-This means that, say, replacing the hash function $h$ with a keyed function or any other cryptographic primitive will be futile in preventing attacks.
-This illustrates the powerful nature of the resettable shadow device (despite our attacks in that setting being more complex and requiring more effort than in scenarios S2 and S4).
-In short, we cannot hope to gain any security in scenario S1 by adding cryptography.
+This means that, say, replacing the hash function $h$ with a keyed function or any other cryptographic primitive would be futile in preventing attacks in scenario S1.
+Interestingly, our attacks in scenario S1 are more complex and require more effort than in scenarios S2 and S4.
+The fact that we cannot prove security in S1 with improved cryptography is not in contradiction to this, but rather an illustration of the power conferred by giving the adversary access to an exact and resettable replica of the target device in the keyed setting.
+
+In summary, we cannot hope to gain any security in scenario S1 by adding more cryptography.
 
@@
 
+We note that our formal security analysis could be conducted using the framework recently laid out in~\cite{CCS:ClaPatShrPS19}. However, we prefer to give a compact and self-contained treatment here.
+
 \subsection{Formal Security Model for HLL}
@@
 After making some queries, ${\cal A}$ returns a bit $b'$ estimating whether it is interacting with a real HLL instantiation ($b'=0$) or an ideal one ($b'=1$).
-We define ${\cal A}$'s S$i$ Real-or-Ideal advantage against HLL to be:
+For $i=1, 2, 3$, we define ${\cal A}$'s S$i$-Real-or-Ideal advantage against HLL to be:
 \[
@@
 \begin{itemize}
-\item \texttt{init}: this oracle performs initialisation of the HLL (and selection of the PRF key when $h$ is replaced by a PRF $f$); this oracle is available in all of S2, S3, S4, but can only be called once as ${\cal A}$'s first oracle query.
+\item \texttt{init}: this oracle performs initialisation of the HLL (and selection of the PRF key when $h$ is replaced by a PRF $F$); this oracle is available in all of S2, S3, S4, but can only be called once as ${\cal A}$'s first oracle query.
 \item \texttt{insert($\cdot$)}: on input $x \in {\cal X}$, the oracle inserts value $x$ into the HLL.
 This oracle is available in all of S2, S3, S4.
-Without loss of generality (since the HLL and variants of it we consider here have a deterministic insertion procedure), we assume all queries $x$ made to \texttt{insert($\cdot$)} are distinct.
+Without loss of generality (since HLL and variants of it we consider here have a deterministic insertion procedure), we assume all queries $x$ made by ${\cal A}$ to \texttt{insert($\cdot$)} are distinct.
 We let $q$ denote the number of insertion queries made by ${\cal A}$ during the course of its attack.
 \item \texttt{read}: this oracle returns the HLL cardinality estimate. We allow multiple queries to this oracle and make it available to ${\cal A}$ in S3 and S4, but not S2.
-\item \texttt{corrupt}: this oracle returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents.
+\item \texttt{corrupt}: this oracle returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents (but not any secret key $K$).
 This oracle is available only in S4.
@@
 
-We now replace the hash function $h$ acting on ${\cal X}$ with a keyed function $F: \{0,1\}^k \times {\cal X} \rightarrow  \{0,1\}^{n+\ell}$.
+We now replace the hash function $h$ acting on ${\cal X}$ from the HLL algorithm with a keyed function $F: \{0,1\}^k \times {\cal X} \rightarrow  \{0,1\}^{n+\ell}$.
 Here the first input to $F$ is a $k$-bit string, the key, denoted $K$.
@@
 The interpretation of the above theorem is that if $F$ is a good PRF, then $\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B})$ should be small for any adversary ${\cal B}$ running in a reasonable amount of time.
-Hence we can conclude that the advantage of any S4 adversary against HLL using PRF $F$ is also small.
+Hence we can conclude that the advantage of any S4 adversary  ${\cal A}$ against HLL using PRF $F$ is also small.
 
@@
 \item \texttt{init}: here ${\cal B}$ initialises an HLL filter in accordance with Figure~\ref{fig:hll}. Thus it initialises an array $M[j]$ in which each entry is set to 0. 
-\item \texttt{insert($\cdot$)}: on input $x \in {\cal X}$,  ${\cal B}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${\cal B}$ uses $y$ to update the HLL bucket $j$ generated via $j=1 + <y_1y_2...y_{n}>_2$ an in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
+\item \texttt{insert($\cdot$)}: on input $x \in {\cal X}$,  ${\cal B}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${\cal B}$ uses $y$ to update the HLL bucket $j$ generated via $j=1 + <y_1y_2...y_{n}>_2$ as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
 \item \texttt{read}: ${\cal B}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate in accordance with Figure~\ref{fig:hll}. 
-\item \texttt{corrupt}: this oracle returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents. 
+\item \texttt{corrupt}: ${\cal B}$ returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents. 
 \end{itemize}
@@
 The equality of advantages in the statement of the theorem now follows immediately.
-The remainder of the proof follows on counting queries and noting that the HLL algorithm for computing cardinality is efficient, so ${\cal B}$'s running time when handling \texttt{read} queries is not increased significantly compared to that of ${\cal A}$.
+The remainder of the proof follows on counting queries and noting that the HLL algorithm for computing cardinality is efficient, so ${\cal B}$'s running time due to handling \texttt{read} queries is not increased significantly compared to that of ${\cal A}$.
 \end{proof}
@@
 This means that ${\cal A}$ is effectively reduced to attacking HLL in the exact same setting used in the average case analysis of HLL from~\cite{hll}, wherein it is proved that the HLL cardinality estimate endures relative errors of the form $1.04/\sqrt{m}$.
+It does not matter that ${\cal A}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries.
+For each \texttt{insert($\cdot$)} query, a fresh value $x$ leads to a new random $y$ being generated and used to update a bucket, as in the average case analysis of HLL.
 

### 2020-10-13 docs/HLL_exp_bad/main.tex
more tweakery

@@
 For each \texttt{insert($\cdot$)} query, a fresh value $x$ leads to a new random $y$ being generated and used to update a bucket, as in the average case analysis of HLL.
+The \texttt{read} and \texttt{corrupt} queries do leak information about $F$'s outputs to the adversary, but such outputs are indistinguishable from random under the assumption that $F$ is a PRF.
 
@@
 
-This construction would be relatively fast on commodity platforms due to widespread hardware support for SHA-256 and AES operations.
-The exact slowdown relative to, say, Murmur3, depends on platform specifics, but can be expected to be in the range 2-fold to 5-fold.
+An alternative is to simply use HMAC-SHA-256 with an appropriately truncated output. HMAC is a PRF under reasonable assumptions on the hash function used in its construction~\cite{Bellare15}.
 
-An alternative is to simply use HMAC-SHA-256 with an appropriately truncated output. HMAC is a PRF under reasonable assumptions on the hash function used in its construction~\cite{Bellare15}.
+These constructions would be relatively fast on commodity platforms due to widespread hardware support for SHA-256 and AES operations.
+It would be a simple exercise to add the above PRF proposals to an existing HLL implementation and benchmark them against, say, Murmur3.
+The exact slowdown depends on platform specifics, but based on microbenchmarks can be expected to be in the range 2-fold to 5-fold compared to Murmur3.
 
@@
 
-It is an important open problem to close the performance gap that exists between fast but non-cryptographic hashing used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security in the face of adversarial inputs.
-
-We also note that using keyed functions sacrifices the mergeability property that HLL enjoys, if different keys $K$ are used in different HLL filters.
+We note that using keyed functions sacrifices the mergeability property that HLL enjoys, if different keys $K$ are used in different HLL filters.
 The reason is that one can no longer emulate running a single filter by taking the maximum over all buckets with each index $j$ in the different filters, since now the bucket contents are created using different keys in a PRF and hence bear no relation to one another.
@@
 This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security.
+It is also an important open problem to close the performance gap that exists between fast but non-cryptographic hashing used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security in the face of adversarial inputs.
 

### 2020-10-13 docs/HLL_exp_bad/main.tex
prepping for usenix submission

@@
-\documentclass[11pt]{article}
+\newif\ifeprint
+%\eprinttrue
+\eprintfalse
+
+\ifeprint
+\documentclass{article}
+\usepackage{geometry}
+\usepackage[dvipsnames]{xcolor}
+\else
+\documentclass[letterpaper,twocolumn,10pt]{article}
+\usepackage[dvipsnames]{xcolor}
+\usepackage{usenix-2020-09}
+\fi
+
+%\documentclass[11pt]{article}
 \usepackage{graphicx, amsmath, amsthm, amssymb, setspace, array, url, hyperref, xspace, todonotes, pifont}
@@
 
-\begin{figure}[tb!]
+\begin{figure*}[tb!]
 \centering
@@
 \label{fig:hll}
-\end{figure}
+\end{figure*}
 
@@
 When $C_{\ell+1} = 0$, $\tau$ is evaluated at 1, where its value is zero.
-Combining this information, we see that, provided $C_{\ell+1} = 0$, we have:
+Combining this information, we see that, provided $C_{\ell+1} = 0$, and letting $D$ denote the denominator in equation~(\ref{eqn:ertl}), we have:
 \[
-m\sigma(C_0/m) + \sum_{k=1}^{\ell}C_k 2^{-k} + m \tau(1-C_{\ell+1}/m)2^{-\ell} \ge m\sigma(C_0/m) \ge C_0
+D \ge m\sigma(C_0/m) \ge C_0
 \]
@@
 
-\begin{table}[tb!]
-\centering
+\begin{table*}[tb!]
+\small\centering
 \caption{Attack results, averaged over 30 iterations.}
@@
 \label{table:tab1}
-\end{table}
+\end{table*}
 
-\begin{table}[tb!]
-\centering
+\begin{table*}[tb!]
+\small\centering
 \caption{Attack results, averaged over 30 iterations, in the setting of~\cite{hllvuln}.}
@@
 \label{table:tab2}
-\end{table}
+\end{table*}
 
@@
 Thus, ${\cal A}$ can have no advantage when it comes to influencing the HLL cardinality estimate.
-It follows that HLL, when instantiated with a keyed PRF, is secure against attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
-
-We will conclude by describing how the necessary keyed PRF can be instantiated.
+It follows that HLL, when instantiated with PRF $F$, is secure against attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
+We will conclude by discussing how $F$ can be instantiated.
 
@@
 \item \texttt{init}: this oracle performs initialisation of the HLL (and selection of the PRF key when $h$ is replaced by a PRF $F$); this oracle is available in all of S2, S3, S4, but can only be called once as ${\cal A}$'s first oracle query.
-\item \texttt{insert($\cdot$)}: on input $x \in {\cal X}$, the oracle inserts value $x$ into the HLL.
+\item \texttt{insert($\cdot$)}: on input $x \in {\cal D}$, the oracle inserts value $x$ into the HLL.
 This oracle is available in all of S2, S3, S4.
@@
 
-We now replace the hash function $h$ acting on ${\cal X}$ from the HLL algorithm with a keyed function $F: \{0,1\}^k \times {\cal X} \rightarrow  \{0,1\}^{n+\ell}$.
+We now replace the hash function $h$ acting on ${\cal D}$ from the HLL algorithm with a keyed function $F: \{0,1\}^k \times {\cal D} \rightarrow  \{0,1\}^{n+\ell}$.
 Here the first input to $F$ is a $k$-bit string, the key, denoted $K$.
@@
 \begin{theorem}\label{thm:security-of-HLL-with-PRF}
-Suppose ${\cal A}$ is an S4 adversary against HLL instantiated with a keyed function $F: \{0,1\}^k \times {\cal X} \rightarrow  \{0,1\}^{n+\ell}$.
+Suppose ${\cal A}$ is an S4 adversary against HLL instantiated with a keyed function $F: \{0,1\}^k \times {\cal D} \rightarrow  \{0,1\}^{n+\ell}$.
 Then we can construct from ${\cal A}$ a PRF adversary ${\cal B}$ against $F$ such that:
@@
 \item \texttt{init}: here ${\cal B}$ initialises an HLL filter in accordance with Figure~\ref{fig:hll}. Thus it initialises an array $M[j]$ in which each entry is set to 0. 
-\item \texttt{insert($\cdot$)}: on input $x \in {\cal X}$,  ${\cal B}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${\cal B}$ uses $y$ to update the HLL bucket $j$ generated via $j=1 + <y_1y_2...y_{n}>_2$ as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
+\item \texttt{insert($\cdot$)}: on input $x \in {\cal D}$,  ${\cal B}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${\cal B}$ uses $y$ to update the HLL bucket $j$ generated via $j=1 + <y_1y_2...y_{n}>_2$ as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
 \item \texttt{read}: ${\cal B}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate in accordance with Figure~\ref{fig:hll}. 
@@
 
-The above construction requires a PRF $F: \{0,1\}^k \times {\cal X} \rightarrow  \{0,1\}^{n+\ell}$.
-Here, the set ${\cal X}$ represents the space of possible inputs to the HLL which can be modelled by a set of bit-strings of some maximum length.
+The above construction requires a PRF $F: \{0,1\}^k \times {\cal D} \rightarrow  \{0,1\}^{n+\ell}$.
+Here, the set ${\cal D}$ represents the space of possible inputs to the HLL which can be modelled by a set of bit-strings of some maximum length.
 This mean that $F$ needs to be Variable Input Length PRF (VIL-PRF).
@@
 
-A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {\cal X} \rightarrow  {\cal D}$ onto some domain ${\cal D}$ with a PRF $G: \{0,1\}^k \times {\cal D} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
+A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {\cal D} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
 It is easy to show that the result is a PRF of the required input-output signature (see for example~\cite[Section 8.2]{BonehShoup}).

### 2020-10-13 docs/HLL_exp_bad/main.tex
prepping for usenix submission
cutting to  13 pages

@@
 \title{HyperLogLog: Exponentially Bad in Adversarial Settings}
-\author{Kenneth G. Paterson and Mathilde Raynal\\
-\vspace{3mm}
-Department of Computer Science, ETH Zurich\\
-{\tt kenny.paterson@inf.ethz.ch, mraynal@student.ethz.ch}}
+%\author{Kenneth G. Paterson and Mathilde Raynal\\
+%\vspace{3mm}
+%Department of Computer Science, ETH Zurich\\
+%{\tt kenny.paterson@inf.ethz.ch, mraynal@student.ethz.ch}}
 
@@
 \maketitle
+%\IEEEtitleabstractindextext{%
 
-%\IEEEtitleabstractindextext{%
 \begin{abstract}
@@
 We show that in the ``shadow device'' attack scenario of~\cite{hllvuln}, no security is possible even if the HLL is secretly keyed.
-On the other hand, we also show that in the most powerful attack setting introduced here, where the attacker has full access to the internals of the HLL (except for a secret key!), it is possible to provably secure HLL against cardinality manipulation by the simple expedient of replacing its internal hash function by a keyed, variable-input-length pseudo-random function (VIL-PRF).
+On the other hand, we also show that in the most powerful attack setting introduced here, where the attacker has full access to the internals of the HLL algorithm (except for a secret key!), it is possible to provably secure HLL against cardinality manipulation by the simple expedient of replacing its internal hash function by a keyed, variable-input-length pseudo-random function (VIL-PRF).
 We describe a suitable instantiation of the required VIL-PRF using off-the-shelf cryptographic components.
@@
 
-Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted sketch, either as a replacement for (but at the cost of losing mergeability) or in addition to (inducing a memory overhead) an unsalted sketch.
+Both~\cite{cardestprivacy,hllvuln} propose mitigations using a salted HLL sketch, either as a replacement for (but at the cost of losing mergeability) or in addition to (inducing a memory overhead) an unsalted HLL.
 We show in Section~\ref{sec:formal} that when a salt is used appropriately (i.e.\ as the secret key for a pseudo-random function that is used in place of hash function $h$), security against adversarial inputs can be achieved even in our strongest adversarial setting.
@@
 It can access and interact with the sketch via an API provided by the sketch owner, as per scenario S3.
-Note that, using these capabilities, an S4 adversary can compute a lower bound on the HLL cardinality estimate based on its initial snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
+%Note that, using these capabilities, an S4 adversary can compute a lower bound on the HLL cardinality estimate based on its initial snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
 \end{itemize}
@@
 
-\begin{table}[h]
+\begin{table}[tb!]
+\small
 \centering
@@
 
-Scenario S1 is rather distinctive in that it is the only one in which the adversary has access via an API to a shadow device (and no other capabilities).
+Scenario S1 is distinctive in that it is the only one in which the adversary has access via an API to a shadow device (and no other capabilities).
 We do not consider scenario S1 typical of HLL deployments.
@@
 Hence we assume for the rest of this work that, in scenario S1, shadow and targeted devices have the same internals and security parameters (if there are any).
-If the HLL device receives inputs from several users, the attack from Section 3.2 of~\cite{hllvuln} is then the easiest to carry out.
+%If the HLL device receives inputs from several users, the attack from Section 3.2 of~\cite{hllvuln} is then the easiest to carry out.
 It is described in~\cite{hllvuln} as follows: "after each insertion [in the shadow device], elements that do not increase the cardinality estimate are retained".
@@
 This does lead to a change in the raw cardinality estimate $E = \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$.
-However, in practice, the raw cardinality estimate is rounded before being presented to the adversary, so this change may not actually be detectable.
+However, in practice, the raw cardinality estimate is rounded before being presented to the adversary, so this change may not actually be detectable, especially when $M[j]$ becomes large.
 We have done experiments which shows this to be the case for parameters arising in practice.
@@
 As a second example, setting $B = 1$, we can expect to insert $2^{t-n}$ items with a final cardinality estimate equal to $m \cdot \ln (1+\frac{1}{m-1})$, which is rounded to $1$ by the HLL algorithm (for $m$ of practical interest, e.g.\ $m=256$).
-Thus we keep the HLL cardinality estimate to a constant (1!) whilst being able to insert exponentially many items, provided there is sufficient flexibility in the choice of input set ${\cal X}$.
+Thus we keep the HLL cardinality estimate to a constant (1!) whilst being able to insert exponentially many items, provided there is sufficient flexibility in the input set ${\cal X}$.
 
@@
 Now we expect to reject in total $(1 - \frac{B}{m})\cdot 2^t$ items, and each one needs an expected number of trial insertions that is bounded by $m/(m - B_{\max} + B)$. 
-(In fact, we can model the number of trial insertions needed as an independent geometric distribution with parameter $q$ for each such $x$.) Meanwhile, each of the $\frac{B}{m}\cdot 2^t$ items in the final insertion list ${\cal L}^*$ for the target HLL sketch is trial inserted into the shadow HLL sketch on every pass.
+%(In fact, we can model the number of trial insertions needed as an independent geometric distribution with parameter $q$ for each such $x$.) 
+Meanwhile, each of the $\frac{B}{m}\cdot 2^t$ items in the final insertion list ${\cal L}^*$ for the target HLL sketch is trial inserted into the shadow HLL sketch on every pass.
 
@@
 
-Now we turn to the case where the HLL sketch under attack is not initially empty, but instead contains some honestly inserted items.
+Now we turn to the case where the HLL under attack is not initially empty, but instead contains some honestly inserted items.
 As usual, the adversary's goal is to insert as many items as possible whilst increasing the cardinality estimate as little as possible.
+We focus on scenarios S2 and S4, and make brief remarks on S1.
 
-We focus here on scenarios S2 and S4, and make brief remarks on S1.
-
-We assume that the target HLL sketch in S2 and S4 has already had sufficiently many items inserted that every one of the $m=2^n$ buckets has counter value $M[j]$ with $M[j] \ge 1$.
-This is highly likely to be the case (with probability at least $1-1/m^2$) as long as at least $m\ln m$ items have already been inserted into the target HLL sketch.
+We assume that the target HLL in S2 and S4 has already had sufficiently many items inserted that every one of the $m=2^n$ buckets has counter value $M[j]$ with $M[j] \ge 1$.
+This is highly likely to be the case (with probability at least $1-1/m^2$) as long as at least $m\ln m$ items have already been inserted into the target HLL.
 This follows from the standard analysis of the coupon collector problem under our usual assumption that $h$ behaves as a random function.
@@
 
-With $m \ln m$ honest insertions, it is also highly likely that the small range correction (Hit Counting algorithm) will not be used, and the target HLL's cardinality estimate will be computed using the raw estimate:
-\[
-E = \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}.
-\]
+With $m \ln m$ honest insertions, it is also highly likely that the small range correction (Hit Counting algorithm) will not be used, and the target HLL's cardinality estimate will be computed using the raw estimate.
+%\[
+%E = \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}.
+%\]
 
-We assume the adversary has (or is given) some set of items ${\cal X}$ that it wishes to insert; as usual, we write $|{\cal X}| = 2^t$.
+%We assume the adversary has (or is given) some set of items ${\cal X}$ that it wishes to insert; as usual, we write $|{\cal X}| = 2^t$.
 
@@
 The adversary knows $h$ so it selects items $x$ from ${\cal X}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit.
-In this case, since the relevant bucket $j$ already has $M[j] \ge 1$, the bucket's counter is not updated, and the cardinality estimate of the HLL sketch remains unchanged.
+In this case, since the relevant bucket $j$ already has $M[j] \ge 1$, the bucket's counter is not updated, and the cardinality estimate of the HLL remains unchanged.
 
-The adversary then expects to be able to insert half of the items from ${\cal X}$ into the HLL sketch, without increasing the HLL cardinality estimate at all.
+The adversary then expects to be able to insert half of the items from ${\cal X}$ into the HLL, without increasing the HLL cardinality estimate at all.
 
@@
 Specifically, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
-In other words, if $h(x)$ is mapped to the $j$-th bucket containing $M[j] \ge 1$, then, instead of the strict requirement of having a ``1'' bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a ``1'' bit in any of the first $M[j]$ positions in this substring.
-Under these conditions, the estimates in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.
+In other words, if $h(x)$ is mapped to the $j$-th bucket containing $M[j] \ge 0$, then, instead of the strict requirement of having a ``1'' bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a ``1'' bit in any of the first $M[j]$ positions in this substring.
+(A value $M[j]=0$ then never permits insertions, meaning an empty bucket stays empty.) Under these conditions, the estimates in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.
 
-We now compute the expected number of elements $x$ from ${\cal X}$ such that $h(x)$ meets the attack's requirements.
-We assume that $M[j] \ge 1$ for all buckets $j$.
+We now compute the expected number of elements $x$ from ${\cal X}$ such that $h(x)$ meets the attack's requirements.  %We assume that $M[j] \ge 0$ for all buckets $j$. 
 Assume $h(x)$ is mapped to bucket $j$.
@@
 
-If some buckets are actually empty (and in S4 the adversary knows which ones using its snapshot), then the above analysis can be reused in order to avoid touching those buckets altogether when inserting elements from ${\cal X}$.
+%If some buckets are actually empty (and in S4 the adversary knows which ones using its snapshot), then the above analysis can be reused in order to avoid touching those buckets altogether when inserting elements from ${\cal X}$.
 The above analysis does not change at all because the term $\frac{1}{m} \cdot (1-\frac{1}{2^{M[j]}})$ already signs a probability of 0 to inserting $x$ in bucket $j$ if $M[j] = 0$.
@@
 In this setting, where the target HLL sketch is initially non-empty, we see that an adversary who has sufficient flexibility in its inputs (as represented by the size $2^t$ of the set ${\cal X}$) can expect to insert exponentially many (in $t$) values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the case where the HLL sketch has some empty buckets in S2).
-The result is largely independent of the HLL parameters, but does depend on the extent to which the HLL sketch is already filled by other users, via a factor $1-H_M^{-1}$: the more full the HLL sketch is, the easier it is to insert new items without them increasing the cardinality estimate.
+The result is largely independent of the HLL parameters.
+The more full is the HLL sketch, the larger is $1-H_M^{-1}$ and the easier it is to insert new items without them increasing the cardinality estimate.
 
@@
 
-Finally, we briefly consider scenario S1 with an initially non-empty target HLL sketch.
+We briefly sketch an attack in this setting in Appendix~\ref{sec:non-empty-HLL-sketch-S1}.
+
+%Finally, we briefly consider scenario S1 with an initially non-empty target HLL sketch.
 Since the target sketch has an unknown status for the adversary, our previous strategy for S1 of keeping many buckets empty does not work.
@@
 \end{equation}
-Here, in the denominator, $C_k$ denotes the number of buckets containing value $k$ in the HLL filter, so $C_0$ is in particular the number of empty buckets; $\sigma(\cdot)$ and $\tau(\cdot)$ are specific functions defined as:
+Here, in the denominator, $C_k$ denotes the number of buckets containing value $k$ in the HLL sketch, so $C_0$ is in particular the number of empty buckets; $\sigma(\cdot)$ and $\tau(\cdot)$ are specific functions defined as:
 \[
@@
 Moreover, in this range, we have $\sigma(x) \ge x$.
-Also it is highly likely that $C_{\ell+1} = 0$ (since it is very unlikely that a 50-bit all-0 string will result from the application of the Redis HLL hash function, and in any case in scenarios S2 and S4, we can always arrange for this to be avoided).
+It is also highly likely that $C_{\ell+1} = 0$ (since it is very unlikely that a 50-bit all-0 string will result from the application of the Redis HLL hash function, and in scenarios S2 and S4, we can always arrange for this to be avoided).
 When $C_{\ell+1} = 0$, $\tau$ is evaluated at 1, where its value is zero.
@@
 
-The requirement on $C_0 \geq {\epsilon}m$ is easily met in scenarios S2 and S4 with an initially empty HLL filter, through the trick of sampling inputs $x \in {\cal X}$ so that only a fraction $1-\epsilon$ of the buckets are ever hit.
+The requirement on $C_0 \geq {\epsilon}m$ is easily met in scenarios S2 and S4 with an initially empty HLL sketch, through the trick of sampling inputs $x \in {\cal X}$ so that only a fraction $1-\epsilon$ of the buckets are ever hit.
 We can expect to insert a fraction $1-\epsilon-\epsilon2^{-\ell}$ of the elements of  ${\cal X}$ whilst keeping the cardinality estimate $\hat{\lambda}$ below  $\alpha_\infty \epsilon^{-1} m$.
@@
 
-In scenarios S2 and S4 with an initially non-empty HLL filter, we assume as for our earlier analysis of the standard HLL algorithm that all buckets have already been hit (this will be the case with high probability as soon as $m \log m$ items have been honestly inserted).
+In scenarios S2 and S4 with an initially non-empty HLL sketch, we assume as for our earlier analysis of the standard HLL algorithm that all buckets have already been hit (this will be the case with high probability as soon as $m \log m$ items have been honestly inserted).
 Hence $C_0 = 0$ and $\sigma$ is evaluated at 0 in (\ref{eqn:ertl}), where its value is 0.
@@
 In S2, we expect to insert half the items in ${\cal X}$ without increasing the cardinality estimate $\hat{\lambda}$ at all.
-In S4, we can insert even more items.
-The analysis can be extended in S4 to the case where some of the buckets are still empty: in S4 we can avoid hitting these empty buckets and again the cardinality estimate does not increase at all.
+In S4, the analysis is the same as in Section~\label{sec:nonemptyHLLS4}.
 
 This completes our analysis of Redis.
-The key takeaway is that, in scenarios S2 and S4, an adversary can expect to insert a constant fraction of ${\cal X}$ into a non-empty Redis HLL filter whilst not increasing its cardinality estimate at all (and whilst keeping the estimated cardinality linear in $m$ in the case where the filter is initially empty).
+The key takeaway is that, in scenarios S2 and S4, an adversary can expect to insert a constant fraction of ${\cal X}$ into a non-empty Redis HLL sketch whilst not increasing its cardinality estimate at all (and whilst keeping the estimated cardinality linear in $m$ in the case where the sketch is initially empty).
 
@@
 %\todo{We need to make sure the repo is publicly accessible but anonymised for submission, if we are going to include it.}
-For simplicity, our items are 4-character long ASCII strings, giving us a set ${\cal X}$ of $7,311,616= 2^{22.8}$ distinct items that can be inserted (hence we have $t=22.8$ in our attacks).
-We choose $h$ to be the 32-bit Murmur3 hash function from~\cite{murmur3code} (note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
+For simplicity, we take ${\cal X}$ to be the set of 4-character long ASCII strings, with $\lvert {\cal X} \rvert = 7,311,616= 2^{22.8}$.
+We choose $h$ to be the 32-bit Murmur3 hash function from~\cite{murmur3code}.
+%(note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
 We create an HLL sketch using $n=8$ and hence $m=256$ buckets and, for attacks in the setting of an initially non-empty target HLL sketch, initialise it with 1,000 random items representing honest users' data.
@@
 These results align well with our theoretical analysis.
-
 %In scenario S1 we run the attack designed for an initially empty target HLL sketch. It performs well in practice, for example, increasing the HLL cardinality estimate from 100 to 126????
@@
 %% That analysis was now updated -- does it still match?
-
-Secondly, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly a rounded version of the Hit Counting estimator $m \cdot \ln (m/(m-B))$ after inserting a fraction $\frac{B}{m}$ of the items, as predicted by our analysis.
+Second, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly a rounded version of the Hit Counting estimator $m \cdot \ln (m/(m-B))$ after inserting a fraction $\frac{B}{m}$ of the items, as predicted by our analysis.
 For example, in scenario S1 with $B=100$, the final cardinality estimate is 126 (in agreement with the formula) while with $B=1$ in scenarios S2, S4, the final cardinality estimate is just 1 (again, as predicted by the formula).
@@
 
-\subsection{Security Proof for HLL Using a PRF}
+\subsection{Security Proof for HLL-PRF}
 
@@
 \]
-Here, we assume without loss of generality that all of ${\cal B}$'s queries are distinct. See~\cite[Section 4.4.1]{BonehShoup} for details.
+Here, we assume without loss of generality that all of ${\cal B}$'s queries are distinct. See~\cite[Section 4.4.1]{BonehShoup} for details of the PRF security definition.
 
@@
 
-The interpretation of the above theorem is that if $F$ is a good PRF, then $\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B})$ should be small for any adversary ${\cal B}$ running in a reasonable amount of time.
-Hence we can conclude that the advantage of any S4 adversary  ${\cal A}$ against HLL using PRF $F$ is also small.
-
-\begin{proof}
-PRF adversary ${\cal B}$ against $F$ is built as follows.
-Let $d$ denote the hidden bit determining whether ${\cal B}$ receives real responses of the form $F(K,x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${\cal B}$ runs S4 adversary ${\cal A}$ and responds to each of its queries as follows:
-\begin{itemize}
-\item \texttt{init}: here ${\cal B}$ initialises an HLL filter in accordance with Figure~\ref{fig:hll}. Thus it initialises an array $M[j]$ in which each entry is set to 0. 
-\item \texttt{insert($\cdot$)}: on input $x \in {\cal D}$,  ${\cal B}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${\cal B}$ uses $y$ to update the HLL bucket $j$ generated via $j=1 + <y_1y_2...y_{n}>_2$ as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
-\item \texttt{read}: ${\cal B}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate in accordance with Figure~\ref{fig:hll}. 
-\item \texttt{corrupt}: ${\cal B}$ returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents. 
-\end{itemize}
-When ${\cal A}$ outputs a bit $b'$, ${\cal B}$ outputs its own bit $d'=b'$.
-Note that when $d=0$, ${\cal B}$ plays against the real $F$ with a random key $K$ and provides a simulation for ${\cal A}$ that perfectly instantiates the real world.
-On the other hand, when $d=1$, ${\cal B}$ plays against a truly random function and provides a simulation for ${\cal A}$ that perfectly instantiates the ideal world.
-The equality of advantages in the statement of the theorem now follows immediately.
-The remainder of the proof follows on counting queries and noting that the HLL algorithm for computing cardinality is efficient, so ${\cal B}$'s running time due to handling \texttt{read} queries is not increased significantly compared to that of ${\cal A}$.
-\end{proof}
-
-Note that the above proof still goes through if the classical HLL algorithm is replaced with the improved algorithm of~\cite{hllnew} that is used in Redis.
+Informally, if $F$ is a good PRF, then $\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B})$ should be small for any adversary ${\cal B}$ running in a reasonable amount of time; hence the advantage of any S4 adversary  ${\cal A}$ against HLL using PRF $F$ is also small.
+The proof can be found in Appendix~\ref{sec:proof}.
+Our proof still applies for the improved algorithm of~\cite{hllnew} that is used in Redis.
 The only changes to the proof are to use the Redis approach for computing responses to \texttt{read} queries in place of the classical HLL approach.
 
-Finally, we conclude our analysis by reiterating the key point: we have shown that no efficient S4 adversary ${\cal A}$ can tell that it is playing against HLL implemented using truly random values $y$ instead of with values obtained from a PRF, except with a small advantage.
+We conclude by reiterating our central point: we have shown that no efficient S4 adversary ${\cal A}$ can tell that it is playing against HLL implemented using truly random values $y$ instead of with values obtained from a PRF, except with a small advantage.
 This means that ${\cal A}$ is effectively reduced to attacking HLL in the exact same setting used in the average case analysis of HLL from~\cite{hll}, wherein it is proved that the HLL cardinality estimate endures relative errors of the form $1.04/\sqrt{m}$.
-It does not matter that ${\cal A}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries.
+%It does not matter that ${\cal A}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries.
 For each \texttt{insert($\cdot$)} query, a fresh value $x$ leads to a new random $y$ being generated and used to update a bucket, as in the average case analysis of HLL.
-The \texttt{read} and \texttt{corrupt} queries do leak information about $F$'s outputs to the adversary, but such outputs are indistinguishable from random under the assumption that $F$ is a PRF.
+%The \texttt{read} and \texttt{corrupt} queries do leak information about $F$'s outputs to the adversary, but such outputs are indistinguishable from random under the assumption that $F$ is a PRF.
 
@@
 
-Making $F$ competitive in speed with existing implementations of HLL, which use non-cryptographic hash functions such as Murmur3, is challenging, and in practice we must sacrifice some speed for security.
+%Making $F$ competitive in speed with existing implementations of HLL, which use non-cryptographic hash functions such as Murmur3, is challenging, and in practice we must sacrifice some speed for security.
 
 A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {\cal D} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
-It is easy to show that the result is a PRF of the required input-output signature (see for example~\cite[Section 8.2]{BonehShoup}).
+The result is a PRF with the required input-output signature~\cite[Section 8.2]{BonehShoup}.
 As a concrete instantiation, one could take $H$ to be SHA-256 with outputs truncated to 128 bits and $G$ to be AES-128 truncated to ${n+\ell}$ bits (since ${n+\ell} \leq 64 < 128$ in practice, this truncation is always possible).
@@
 The exact slowdown depends on platform specifics, but based on microbenchmarks can be expected to be in the range 2-fold to 5-fold compared to Murmur3.
+\todo{Put a hard number here.}
 
-More complex but faster constructions using randomised or nonce-based MACs constructed from universal hashing could be envisaged.
+%More complex but faster constructions using randomised or nonce-based MACs constructed from universal hashing could be envisaged.
 However, these do not seem to be directly applicable here since we need the function $F$ to be both stateless and deterministic so as to cope with the repeated inputs one would encounter during the normal operation of HLL.
 
-We note that using keyed functions sacrifices the mergeability property that HLL enjoys, if different keys $K$ are used in different HLL filters.
-The reason is that one can no longer emulate running a single filter by taking the maximum over all buckets with each index $j$ in the different filters, since now the bucket contents are created using different keys in a PRF and hence bear no relation to one another.
+We note that using keyed functions sacrifices the mergeability property that HLL enjoys, if different keys $K$ are used in different HLL sketches.
+The reason is that one can no longer emulate running a single HLL sketch by taking the maximum over all buckets with each index $j$ in the different sketches, since now the bucket contents are created using different keys in a PRF and hence bear no relation to one another.
 
@@
 Left to future work is the task of extending our attacks and defences to HLL variants such as sliding HLL~\cite{slidinghll} and HLL++~\cite{hllpractice}.
+It would be interesting to find and analyse a strategy for more cleverly ordering the candidates in our attack in scenario S1 to reduce the total insertion cost of the attack, whilst also keeping the number of resets low.
+Finding an attack in scenario S1 when the target HLL sketch is initially non-empty is also open.
 
-It would be interesting to find and analyse a strategy for more cleverly ordering the candidates in our attack in scenario S1 to reduce the total insertion cost of the attack, whilst also keeping the number of resets low.
-Finding an attack in scenario S1 when the target HLL filter is initially non-empty is also open.
-
-We have given a formal analysis of using a secretly-keyed PRF as a countermeasure to our attacks.
-The reliance on secret keys means sacrificing the mergeability property of HLL sketches in distributed environments (where the keys in the different HLLs are independently generated).
-It is an interesting cryptographic challenge to find methods for securing HLL sketches against adversarial inputs whilst retaining some form of mergeability.
+We have given a formal analysis of using a PRF as a countermeasure to our attacks.
+The use of a keyed function sacrifices HLL mergeability when the keys in the different HLLs are independently generated.
+It is an open problem to find methods for securing HLL sketches against adversarial inputs whilst retaining some form of mergeability.
 This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security.
-It is also an important open problem to close the performance gap that exists between fast but non-cryptographic hashing used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security in the face of adversarial inputs.
+It is also an open problem to close the performance gap that exists between fast but non-cryptographic hashes used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security.
 
@@
 
+\appendix
+
+\section{Initially Non-empty HLL Sketch, Scenario S1}\label{sec:non-empty-HLL-sketch-S1}
+
+We briefly sketch an attack in the setting of an S1 adversary attacking an initially non-empty HLL sketch.
+An alternative attack with moderate performance in this setting is presented in~\cite{hllvuln}.
+
+Since the target sketch has an unknown status for the adversary, our previous strategy for S1 of keeping many buckets empty no longer works.
+One can attempt to use the shadow device to identify items that would result in small entries $M[j]$ (for unknown indices $j$) since such items are unlikely to increase the cardinality estimate of the target device by much.
+This can be done by first filling the shadow device with sufficiently many values to escape from the ``small range correction'' phase.
+Then one could enter target values $x$ into the shadow device and check for cardinality increases.
+However this runs into the rounding issue described in Section~\ref{sec:scenarios-discussion}.
+Still, one could retain those items which do lead to a detectable change in cardinality and reject the others (and also estimate the value of $M[j]$ for the unknown bucket $j$ from the magnitude of the change, retaining only those for which $M[j]$ is small).
+But here the lack of a cardinality change could result either from rounding issues or because insertion of target value $x$ into the shadow device really did lead to no increase in cardinality.
+This indicates that  ``good'' items would be rejected unnecessarily.
+This could be addressed by using multiple resets of the shadow HLL sketch, filling it with random items after each reset, and performing multiple insertions per target item $x$ to try to estimate the true value of $M[j]$ for each item $x$ (since after at least some of the resets followed by random insertions, we could expect $M[j]$ to increase when $x$ is inserted, and if $M[j]$ is small enough relative to the cardinality estimate, the change could be detected).
+The details of such an attack require further analysis which we do not pursue here, since we consider S1 to be the least realistic attack scenario.
+
+\section{Proof of Theorem~\ref{thm:security-of-HLL-with-PRF}}\label{sec:proof}
+
+\begin{proof}
+PRF adversary ${\cal B}$ against $F$ is built as follows.
+Let $d$ denote the hidden bit determining whether ${\cal B}$ receives real responses of the form $F(K,x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${\cal B}$ runs S4 adversary ${\cal A}$ and responds to each of its queries as follows:
+\begin{itemize}
+\item \texttt{init}: here ${\cal B}$ initialises an HLL sketch in accordance with Figure~\ref{fig:hll}. Thus it initialises an array $M[j]$ in which each entry is set to 0. 
+\item \texttt{insert($\cdot$)}: on input $x \in {\cal D}$,  ${\cal B}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${\cal B}$ uses $y$ to update the HLL bucket $j$ generated via $j=1 + <y_1y_2...y_{n}>_2$ as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
+\item \texttt{read}: ${\cal B}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate in accordance with Figure~\ref{fig:hll}. 
+\item \texttt{corrupt}: ${\cal B}$ returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents. 
+\end{itemize}
+When ${\cal A}$ outputs a bit $b'$, ${\cal B}$ outputs its own bit $d'=b'$.
+Note that when $d=0$, ${\cal B}$ plays against the real $F$ with a random key $K$ and provides a simulation for ${\cal A}$ that perfectly instantiates the real world.
+On the other hand, when $d=1$, ${\cal B}$ plays against a truly random function and provides a simulation for ${\cal A}$ that perfectly instantiates the ideal world.
+The equality of advantages in the statement of the theorem now follows immediately.
+The remainder of the proof follows on counting queries and noting that the HLL algorithm for computing cardinality is efficient, so ${\cal B}$'s running time due to handling \texttt{read} queries is not increased significantly compared to that of ${\cal A}$.
+\end{proof}
+
 \end{document}

### 2020-10-13 docs/HLL_exp_bad/main.tex
cutting to  13 pages

@@
 This can be done by first filling the shadow device with sufficiently many values to escape from the ``small range correction'' phase.
-Then one could enter target values $x$ into the shadow device and check for cardinality increases.
+Then one could enter candidate values $x$ into the shadow device and check for cardinality increases.
 However this runs into the rounding issue described in Section~\ref{sec:scenarios-discussion}.
 Still, one could retain those items which do lead to a detectable change in cardinality and reject the others (and also estimate the value of $M[j]$ for the unknown bucket $j$ from the magnitude of the change, retaining only those for which $M[j]$ is small).
-But here the lack of a cardinality change could result either from rounding issues or because insertion of target value $x$ into the shadow device really did lead to no increase in cardinality.
+But here the lack of a cardinality change could result either from rounding issues or because insertion of candidate value $x$ into the shadow device really did lead to no increase in cardinality.
 This indicates that  ``good'' items would be rejected unnecessarily.
-This could be addressed by using multiple resets of the shadow HLL sketch, filling it with random items after each reset, and performing multiple insertions per target item $x$ to try to estimate the true value of $M[j]$ for each item $x$ (since after at least some of the resets followed by random insertions, we could expect $M[j]$ to increase when $x$ is inserted, and if $M[j]$ is small enough relative to the cardinality estimate, the change could be detected).
+This could be addressed by using multiple resets of the shadow HLL sketch, filling it with random items after each reset, and performing multiple insertions per candidate item $x$ to try to estimate the true value of $M[j]$ for each item $x$ (since after at least some of the resets followed by random insertions, we could expect $M[j]$ to increase when $x$ is inserted, and if $M[j]$ is small enough relative to the cardinality estimate, the change could be detected).
 The details of such an attack require further analysis which we do not pursue here, since we consider S1 to be the least realistic attack scenario.

### 2020-10-14 docs/HLL_exp_bad/main.tex
bringing in SipHash and updating related work

@@
 On the other hand, we also show that in the most powerful attack setting introduced here, where the attacker has full access to the internals of the HLL algorithm (except for a secret key!), it is possible to provably secure HLL against cardinality manipulation by the simple expedient of replacing its internal hash function by a keyed, variable-input-length pseudo-random function (VIL-PRF).
-We describe a suitable instantiation of the required VIL-PRF using off-the-shelf cryptographic components.
+We discuss and benchmark a suitable choice for the required VIL-PRF.
 
@@
 Desfontaines \emph{et al.}~\cite{cardestprivacy} studied the privacy properties of HLL in two different attack scenarios, referred to as \emph{insider} and \emph{external}.
-The attack target of~\cite{cardestprivacy} is different from ours, being concerned with loss of privacy, whereas we care about manipulating cardinality estimation.
-It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
-In Section~\ref{sec:attacks} we will discuss the two attack scenarios from~\cite{cardestprivacy} in more detail.
+The attack target of~\cite{cardestprivacy} is different from ours, being focussed on privacy breaches, whereas we care about manipulating cardinality estimation.
+%It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
+We discuss the two attack scenarios from~\cite{cardestprivacy} in more detail in Section~\ref{sec:attacks}.
+
 %In particular, we will show that an adversary who can insert items into the HLL and who can access the HLL cardinality estimate (via typical APIs in HLL implementations) as permitted in the external scenario of~\cite{cardestprivacy} can actually realise the same attacks as the insider adversary can, at least insofar as cardinality estimate manipulation is concerned.
@@
 We consider the attack model used in~\cite{hllvuln} to be quite artificial, since it gives the adversary the ability to test whether inserting an item \emph{would} increase the cardinality estimate, without actually having to insert the item.
-After some clarifications obtained through personal correspondence with the authors of~\cite{hllvuln}, we understand that such a setup might be possible under certain circumstances, for example where the adversary has access to an identical ``shadow'' device; see further discussion in Section~\ref{sec:attacks}.
+Such a setup might be possible under certain circumstances, for example where the adversary has access to an identical ``shadow'' device; see further discussion in Section~\ref{sec:attacks}.
 For completeness, we include analyses of this attack scenario, showing that we can drastically improve on the attacks in~\cite{hllvuln} (e.g.\ we can reduce the HLL cardinality estimate by a factor of almost 1,000 instead of the factor of 5 achieved in~\cite{hllvuln}).
@@
 
-Clayton \emph{et al.}~\cite{CCS:ClaPatShrPS19} provide a provable-security treatment of probabilistic data structures in adversarial environments more generally.
-They use their formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures, but not HLL.
+Our work relates to the broader study of the performance of probabilistic data structures in adversarial environments.
+Clayton \emph{et al.}~\cite{CCS:ClaPatShrPS19} recently provided a provable-security treatment of this topic.
+They used their formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures, but not HLL.
+Prior work also focussed on Bloom filters~\cite{NaorY15,GerbetKL15} or flooding of hash tables~\cite{LiptonN93,CrosbyW03}.
+References~\cite{GerbetKL15,CCS:ClaPatShrPS19} provide comprehensive summaries of prior work in this area.
+
 %It is an interesting open problem to use their framework to develop a formal understanding of how HLL can be made secure in the adversarial setting.
@@
 \begin{itemize}
-\item[S1:] The adversary does not know the details of the target HLL implementation but can access a shadow copy of the HLL instance via its API.
-It can use the API to insert elements into this shadow sketch, get its cardinality estimate, and reset it to its native empty state.
+\item[S1:] The adversary does not know the details of the target HLL sketch but can access a shadow copy of the HLL sketch via its API.
+It can use the API to insert elements into this shadow HLL sketch, get its cardinality estimate, and reset it to its native empty state.
 The adversary does not know the state of the targeted HLL sketch, meaning in particular that it does not have the list of elements previously inserted into it by other users.
@@
 For simplicity, we take ${\cal X}$ to be the set of 4-character long ASCII strings, with $\lvert {\cal X} \rvert = 7,311,616= 2^{22.8}$.
-We choose $h$ to be the 32-bit Murmur3 hash function from~\cite{murmur3code}.
+We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}.
 %(note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
@@
 
-\subsection{Security Proof for HLL-PRF}
+\subsection{Security Analysis of HLL with a PRF}
 
@@
 \[
-\mbox{Adv}^{\mbox{\tiny S4-RoI}}_{\mbox{\tiny HLL}}({\cal A}) = \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B}) \,.
+ \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B}) = \mbox{Adv}^{\mbox{\tiny S4-RoI}}_{\mbox{\tiny HLL}}({\cal A}) \,.
 \]
@@
 
+We want $F$ to be competitive with non-cryptographic hashing on small input sizes, as typical in current HLL implementations and applications (e.g.\ MurmurHash3 applied to an IP header).
+This makes off-the-shelf approaches like HMAC uncompetitive, since even for the shortest inputs, HMAC would require 4 evaluations of the underlying hash compression function.
+
+SipHash~\cite{AumassonB12} is designed specifically for this use case.
+It is competitive in speed with non-cryptographic hash functions commonly used in HLL implementations such as MurmurHash3.
+It has also withstood cryptanalysis well~\cite{DobraunigMS14,XinLSL19}.
+It has 128-bit outputs that can be truncated to the needs of typical HLL implementations (32 and 64 bits).%and is well supported in libraries such as OpenSSL.
+
+\todo{Add a sentence on SipHash-2-4 performance here. Microbenchmark on a short input suffices.}
+
 %Making $F$ competitive in speed with existing implementations of HLL, which use non-cryptographic hash functions such as Murmur3, is challenging, and in practice we must sacrifice some speed for security.
 
-A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {\cal D} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
+%A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {\cal D} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
 The result is a PRF with the required input-output signature~\cite[Section 8.2]{BonehShoup}.
@@
 
-An alternative is to simply use HMAC-SHA-256 with an appropriately truncated output. HMAC is a PRF under reasonable assumptions on the hash function used in its construction~\cite{Bellare15}.
+%An alternative is to simply use HMAC-SHA-256 with an appropriately truncated output. HMAC is a PRF under reasonable assumptions on the hash function used in its construction~\cite{Bellare15}.
 
-These constructions would be relatively fast on commodity platforms due to widespread hardware support for SHA-256 and AES operations.
+%These constructions would be relatively fast on commodity platforms due to widespread hardware support for SHA-256 and AES operations.
 It would be a simple exercise to add the above PRF proposals to an existing HLL implementation and benchmark them against, say, Murmur3.
@@
 This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security.
-It is also an open problem to close the performance gap that exists between fast but non-cryptographic hashes used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security.
+%It is also an open problem to close the performance gap that exists between fast but non-cryptographic hashes used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security.
 

### 2020-10-14 docs/HLL_exp_bad/main.tex
tweaks

@@
 To complement our attack results, we provide a formal security analysis of HLL.
-We show that in the ``shadow device'' attack scenario of~\cite{hllvuln}, no security is possible even if the HLL is secretly keyed.
+We show that in the ``shadow device'' attack scenario of~\cite{hllvuln}, no security is possible even if HLL is secretly keyed.
 On the other hand, we also show that in the most powerful attack setting introduced here, where the attacker has full access to the internals of the HLL algorithm (except for a secret key!), it is possible to provably secure HLL against cardinality manipulation by the simple expedient of replacing its internal hash function by a keyed, variable-input-length pseudo-random function (VIL-PRF).
@@
 For example it can manipulate IPID and other header fields in the context of IP packets.
-We model this by assuming the adversary has access to a large set ${\cal X}$ of potential inputs from which it can select items to insert into the HLL.
+We model this by assuming the adversary has access to a large set ${\cal X}$ of potential inputs from which it can select items to insert into the HLL sketch.
 Our adversary's objective, then, is to insert as many items from ${\cal X}$ as possible whilst keeping the HLL cardinality estimate as low as possible.
@@
 However, it does not know the individual bucket contents.
-\item[S4:] The adversary has direct access to the HLL in use and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch.
+\item[S4:] The adversary has direct access to the HLL sketch and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch.
 It can access and interact with the sketch via an API provided by the sketch owner, as per scenario S3.
@@
 
-Now we turn to the case where the HLL under attack is not initially empty, but instead contains some honestly inserted items.
+Now we turn to the case where the HLL sketch under attack is not initially empty, but instead contains some honestly inserted items.
 As usual, the adversary's goal is to insert as many items as possible whilst increasing the cardinality estimate as little as possible.
@@
 The adversary knows $h$ so it selects items $x$ from ${\cal X}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit.
-In this case, since the relevant bucket $j$ already has $M[j] \ge 1$, the bucket's counter is not updated, and the cardinality estimate of the HLL remains unchanged.
+In this case, since the relevant bucket $j$ already has $M[j] \ge 1$, the bucket's counter is not updated, and the cardinality estimate of the HLL sketch remains unchanged.
 
-The adversary then expects to be able to insert half of the items from ${\cal X}$ into the HLL, without increasing the HLL cardinality estimate at all.
+The adversary then expects to be able to insert half of the items from ${\cal X}$ into the HLL sketch, without increasing the HLL cardinality estimate at all.
 
@@
 
-However, it is clear that any adversary in the ideal world is reduced to operating in the setting used in the average case analysis of HLL (wherein every insertion query results in a random value $y$ being used to update the HLL).
+However, it is clear that any adversary in the ideal world is reduced to operating in the setting used in the average case analysis of HLL (wherein every insertion query results in a random value $y$ being used to update the HLL sketch).
 Thus, ${\cal A}$ can have no advantage when it comes to influencing the HLL cardinality estimate.
@@
 \begin{itemize}
-\item \texttt{init}: this oracle performs initialisation of the HLL (and selection of the PRF key when $h$ is replaced by a PRF $F$); this oracle is available in all of S2, S3, S4, but can only be called once as ${\cal A}$'s first oracle query.
-\item \texttt{insert($\cdot$)}: on input $x \in {\cal D}$, the oracle inserts value $x$ into the HLL.
+\item \texttt{init}: this oracle performs initialisation of the HLL sketch (and selection of the PRF key when $h$ is replaced by a PRF $F$); this oracle is available in all of S2, S3, S4, but can only be called once as ${\cal A}$'s first oracle query.
+\item \texttt{insert($\cdot$)}: on input $x \in {\cal D}$, the oracle inserts value $x$ into the HLL sketch.
 This oracle is available in all of S2, S3, S4.
@@
 This means that ${\cal A}$ is effectively reduced to attacking HLL in the exact same setting used in the average case analysis of HLL from~\cite{hll}, wherein it is proved that the HLL cardinality estimate endures relative errors of the form $1.04/\sqrt{m}$.
-%It does not matter that ${\cal A}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries.
+It does not matter that ${\cal A}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries.
 For each \texttt{insert($\cdot$)} query, a fresh value $x$ leads to a new random $y$ being generated and used to update a bucket, as in the average case analysis of HLL.
-%The \texttt{read} and \texttt{corrupt} queries do leak information about $F$'s outputs to the adversary, but such outputs are indistinguishable from random under the assumption that $F$ is a PRF.
+The \texttt{read} and \texttt{corrupt} queries do leak information about $F$'s outputs to the adversary, but such outputs are indistinguishable from random under the assumption that $F$ is a PRF.
 
@@
 The above construction requires a PRF $F: \{0,1\}^k \times {\cal D} \rightarrow  \{0,1\}^{n+\ell}$.
-Here, the set ${\cal D}$ represents the space of possible inputs to the HLL which can be modelled by a set of bit-strings of some maximum length.
+Here, the set ${\cal D}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
 This mean that $F$ needs to be Variable Input Length PRF (VIL-PRF).
@@
 
-\todo{Add a sentence on SipHash-2-4 performance here. Microbenchmark on a short input suffices.}
+For example, our microbenchmarks created on an Intel i7 CPU show that SipHash-2-4 is as fast as the 128-bit version of MurmurHash3 and only slightly slower than the original 32-bit version.
 

### 2020-10-15 docs/HLL_exp_bad/main.tex
adding perf numbers

@@
 It has also withstood cryptanalysis well~\cite{DobraunigMS14,XinLSL19}.
-It has 128-bit outputs that can be truncated to the needs of typical HLL implementations (32 and 64 bits).%and is well supported in libraries such as OpenSSL.
+It has 128-bit outputs that can be truncated to the needs of typical HLL implementations (32 and 64 bits).
+%and is well supported in libraries such as OpenSSL.
+Our microbenchmarks created on an Intel i7 CPU show that SipHash-2-4 on 90-byte inputs requires 102ns per evaluation, while the 128-bit version of MurmurHash3 takes 60ns.
+Meanwhile, HMAC-SHA256 requires 705ns.
+We obtained similar performance ratios for other input sizes.
+These together show that SipHash is quite competitive with MurmurHash3.
 
-For example, our microbenchmarks created on an Intel i7 CPU show that SipHash-2-4 is as fast as the 128-bit version of MurmurHash3 and only slightly slower than the original 32-bit version.
+%is as fast as the 128-bit version of MurmurHash3 and only slightly slower than the original 32-bit version.
+
+%length: 720, time h_mac 705ns, time h_murmur 60ns and time h_sip 102ns
 

### 2020-10-21 docs/HLL_exp_bad/main.tex
fixing an unfortunate typo
Updates from Overleaf

@@
 Desfontaines \emph{et al.}~\cite{cardestprivacy} studied the privacy properties of HLL in two different attack scenarios, referred to as \emph{insider} and \emph{external}.
-The attack target of~\cite{cardestprivacy} is different from ours, being focussed on privacy breaches, whereas we care about manipulating cardinality estimation.
+The attack target of~\cite{cardestprivacy} is different from ours, being focused on privacy breaches, whereas we care about manipulating cardinality estimation.
 %It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
@@
 They used their formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures, but not HLL.
-Prior work also focussed on Bloom filters~\cite{NaorY15,GerbetKL15} or flooding of hash tables~\cite{LiptonN93,CrosbyW03}.
+Prior work also focused on Bloom filters~\cite{NaorY15,GerbetKL15} or flooding of hash tables~\cite{LiptonN93,CrosbyW03}.
 References~\cite{GerbetKL15,CCS:ClaPatShrPS19} provide comprehensive summaries of prior work in this area.

### 2020-10-21 docs/HLL_exp_bad/main.tex
making a non-anon version

--- 2020-10-15 docs/HLL_exp_bad/main.tex
@@
 \title{HyperLogLog: Exponentially Bad in Adversarial Settings}
-%\author{Kenneth G. Paterson and Mathilde Raynal\\
-%\vspace{3mm}
-%Department of Computer Science, ETH Zurich\\
+\author{{\rm Kenneth G. Paterson}\\ {\rm ETH Zurich} \and {\rm Mathilde Raynal\thanks{Work done while an exchange student at ETH Zurich.}}\\ {\rm EPFL}}
 %{\tt kenny.paterson@inf.ethz.ch, mraynal@student.ethz.ch}}
@@
 Suppose an adversary inserts elements $x$ from some set ${\cal X}$ into an initially empty HLL sketch.
-Suppose the adversary can guarantee that in its attack, only a subset of size $B$ out of $2^m$ buckets are hit, where $B \leq B_{\max}$ with
+Suppose the adversary can guarantee that in its attack, only a subset of size $B$ out of $m$ buckets are hit, where $B \leq B_{\max}$ with
 \[

### 2021-01-13 docs/HLL_exp_bad/main.tex
fixing typos in reviews

@@
 The more complex version of the attack only resets the shadow HLL sketch when its cardinality estimate reaches a value indicating that $B_{\max}$ buckets have been hit (recall this number can be determined precisely because of the use of the Hit Counting algorithm for cardinality estimation).
-Hence, less resets should be required.
+Hence, fewer resets should be required.
 However, when the number of active buckets has reached some value $B^\prime \leq B_{\max}$ this means that an item $x$ may not increase the cardinality estimate, yet still not hit one of the $B$ target buckets (instead it hits one of the other $B^\prime - B$ active buckets).
@@
 In S2, we expect to insert half the items in ${\cal X}$ without increasing the cardinality estimate $\hat{\lambda}$ at all.
-In S4, the analysis is the same as in Section~\label{sec:nonemptyHLLS4}.
+In S4, the analysis is the same as in Section~\ref{sec:nonemptyHLLS4}.
 
@@
 Here, the set ${\cal D}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
-This mean that $F$ needs to be Variable Input Length PRF (VIL-PRF).
+This means that $F$ needs to be Variable Input Length PRF (VIL-PRF).
 

### 2021-01-13 docs/HLL_exp_bad/rebuttal.txt
adding rebuttal

Rebuttal text

We thank the reviewers for their detailed comments. We will fix all typos and reduce (!) the margins. We address key comments from each reviewer in turn.

Reviewer A:

We agree that HLL is obviously unsound in S2, S3, S4. Yet we seem to be amongst the first to point this out. Every system we looked at uses HLL in an unkeyed fashion. Examples include Google BigQuery, PrestoDB and Redis. We think usage of HLL is becoming widespread. It certainly has been proposed for use in adversarial settings without any mention of keying [2,3,5]. Moreover, our attacks in S2, S3, S4 go beyond the trivially obvious attacks by exploiting HLL’s use of “small range correction” to get more powerful attacks (we will add explicit mention of this and comparison to the paper). So this paper is more than a blogpost/bug report.

Resetting a shadow device (which could be a network switch) might involve rebooting it. This would be costly in time, so the attacker wants to minimise the number of resets (as envisaged by Reviriego and Ting [6]).

We do give estimates (including the constants) for both the number of insertions N_I and resets N_R in both the “simple” and “complex” attacks in S1. We only used the “O()” notation to highlight the main term in a complicated expression for the “complex” attack. Of course, we can make the comparison more explicit, but we politely ask the reviewer to please carefully reread our text, page 7, where everything is laid out (even using displayed equations for the complex case).

It is common in distributed environments to share keys across devices. Think of TLS private keys in a distributed server setting. Your review itself makes this comparison later (in its last paragraph). We do not try to prove security in S1. The fact that our security model has “independent key generation” says nothing about whether that key ends up in a single HLL instance or in multiple instances (but our analysis for S2-S4 only applies to the single instance case).

Lemma 1: yes, 2^m should be m = 2^n, this is an unfortunate typo.

We will add the number of resets to Table 2. Unfortunately, Reviriego and Ting [6] are not precise about which implementation of HLL they used. See the last paragraph of Section 6, where we discuss this in detail.

Yes, various network switches could be equipped with same key. Of course, getting the key from one device then compromises the security of all the others. So there are scenarios where being able to merge under different keys would be nice.

Reviewer B:

In the port-scanning application, the attacker could exploit flexibility in other fields in the IP header, e.g. the 16-bit IPID field. Whether this would work would depend on what IP header fields exactly are fed into the HLL hash. Alternatively, with our S2 attack, the attacker could scan roughly half the ports whilst keeping the cardinality estimate below mlog2 = 0.693m.

We will investigate whether something weaker than a PRF would suffice - indeed a pairwise independent hash (with an key unknown to the attacker) may be enough - thanks!

We believe there is value in the generality of our analysis (simply put, provided there is enough flexibility in the input, an attacker can make the HLL severely underestimate the input cardinality). But we will make the connection between our analysis and its effect on specific security mechanisms mentioned in the introduction clearer.

Reviewer C:

We will add a simplified description of HLL in Section 2 by expanding the description of the “single bucket” version. In Figure 1, “m” does indeed occur multiple times, but has the same meaning everywhere and always m = 2^n with n in [4,16].

Reviewer D:

We tried to present our attacks in a generic way - the attacker has a set X of strings from which he can select and wants to insert as much of X as possible into the HLL - rather than tying them to a specific setting. We can work to make the connections back to specific settings clearer. The introduction cites 3 documented scenarios - estimation of Facebook user statistics, DoS attack detection, network flow estimation. In all 3 cases, our attacks would lead to undercounting, with various real-world consequences, e.g. missing DoS attacks or wrongly optimised routing protocols.

We will improve the description of HLL itself - see response to Reviewer C.

The attacks all work identically if SHA is used in place of a non-cryptographic hash function. All that is needed is that the hash maps the inputs to roughly uniform outputs. The adversary does not need to “force” exact outputs and we did not use any specific weakness in Murmur in the attacks (only its “strength” in providing roughly uniform outputs).

The attacks no longer work if a PRF is used (which is what we assume the referee means by “MAC”). Indeed, we prove security in this case for S2-S4, provided the adversary does not know the PRF key. If it does know the key, the attacks work exactly as before.

Reviewer E:

The attacks are not sensitive at all to the form of the inputs - all that is needed is that the hash function h, acting on those inputs, should have approximately uniform outputs. So 4-character ACSII is just as good as, say, IP packet headers, for experimental evaluation.

We agree that having a target HLL with no honest inputs is idealised. Section 4.2 provides attacks in the setting where there are honest inputs. For example, in S2, assuming the HLL is at least lightly loaded, an attacker can insert about half of all items without increasing the cardinality estimate at all, and can do even better in S4. We can add experiments that confirm this analysis.

<end>
967 words.

TODO:

fix margin size
Missing reference in Section 5: "In S4, the analysis is the same as in Section ."
Probable typo in Section 6: "only, thus that are initially empty." Probably should be "thus they are".
Better HLL description (by adding a simplified version?)

REVIEWS QUESTIONS:

* Wouldn't a more reasonable set of adversarial models be "insertion of items + m known + (maybe) get cardinality estimate + (maybe) get  snapshot"? 
   * If we assume that the knowledge of the adv is limited to m and can only insert items, then because of the Hit Counting phase, no attack can be done on crafting/picking items. The only thing that comes to my mind is a probabilistic attack where the adversary limits itself to x(m) items hoping to fall under a certain cardinality estimate.
   * When adding cardinality estimate, we obtain an elevated S1 where an adversary could filter items not based on the effect of the insertion on the shadow device but on the target itself. 
   * When adding snapshot, one can compute the cardinality estimate by itself and partially recreate the map of the (even keyed) h: x -> {0, 1}^n so can end up in S4 with some computational work.

* Are there really deployments of HLL in adversarial settings that don't key $h$?
   * The original implementation does not mention any key, and many implementations do not allow to input a key. If the user is not aware of that and use those libraries, then yes they would deploy something key-less.
   * https://github.com/svpcom/hyperloglog; https://prestodb.io/docs/current/functions/hyperloglog.html; https://github.com/prestodb/presto/blob/6ff3052e9df1a04613157746e7e779fd54c0c2a1/presto-tests/src/main/java/com/facebook/presto/tests/CreateHll.java
   * Add a link for Redis - same, has no key: https://github.com/redis/redis/blob/efaf09ee4b6437c69c467acdb0c62a510207e993/src/hyperloglog.c#L466 Seed is hardcoded ? Could you please check because I am not sure at all.
   * Yep, it’s a hardcoded seed.

* Why would a shadow device have the same _cryptographic keys_ as a real device?
   * Quite common to have multiple webserver machines or TLS terminators (for load balancing) sharing certificate private keys
   * Also inherited from Riviregio and Ting

* Why would resets of a shadow device the attacker controls be  costly? Constant comparison.
   * Resets could be costly depending on the application. For the facebook application, a reset occurs only every day so you don’t want to wait that long. It might also be the case that you pay at each reset (if you have a network monitoring device) so the less resets = the less bills.
   * Constants comparison would be nice indeed…
   * Number of resets should be reported in table 2

* And the results of applying their method to the same implementation of HLL you used as well?
   *

* Analysis of whether a weaker hash assumption could be used.  (Or analysis of further variations of HLL, if they are needed to make a concrete attack example.)
* Analysis about whether something less than a PRF would work, analogous to pairwise-independent hashing for hash tables.
* Unify the use of $m$
   * it is and represents the number of registers.

* SHA instead of Murmur?
   * Murmur is designed to be fast, yet remain fairly unique to avoid collisions. Going for SHA-256 would increase the insertion time on the server side, which is not desired when handling big streams.
   * 25 cycles/byte for SHA 256 vs 1 cycle/byte for Murmur3

* A keyed hash is described but if it is computationally difficult to produce a value that leads to a certain element being inserted into the sketch how does this affect the attack?
   * If it is computationally difficult, it means that the attack takes more time.

* While there is no secret, it may still be difficult for the adversary to insert values of their choice.
   * Insertion should be one API call or one user action (log in to Facebook for example). It does not make much sense that it should be difficult.
   * Values are found via brute force, so it cannot be difficult?

* Can you describe the difference between use of a non-cryptographic hash like Murmur3, a cryptographic hash and a MAC?
   * They all have different goals. Non cryptographic hash aims for speed while the crypto alternatives aims for security. MAC on the other hand aim for unforgeability.
   * This is why in the context of HLL applications, non-cryptographic hashes are often picked as they are extremely fast and “unique enough” as we expect collisions to happen anyway for big tables.

* Empirical results: could do again with another set of inputs (the set of 4-character ASCII strings, or a set of 250K random strings)
   * Totally

* Realistic to consider a scenario where attack a target with no honest inputs already in the sketch?
   * Yes, example of network monitoring

* Map scenarios to real life examples. For ex, ZMap or nmap that successfully evades detection against a defender using Redis for log analysis

Original reviews (A, B, C, D, E with scores 1, 5, 4, 2, 5 resp.)

Review #186A
===========================================================================

Review recommendation
---------------------
1. Reject

Reviewer expertise
------------------
3. Knowledgeable

Overall merit
-------------
1. Bottom 50% of submitted papers

Writing quality
---------------
4. Well-written

Paper summary
-------------
This paper describes a family of attacks against the well-known
cardinality estimator HyperLogLog (HLL) that allow an adversary who
knows the hash function used in the HLL deployment to insert a large
number of items into the sketch, while having HLL report only a small
value.

Strengths
---------
+ The paper is quite well written; thanks for that!  It appears to use
  the wrong margin size, though?

+ The formal security modelling in Section 7 is a good exercise to
  actually do, though the answer it quickly comes to (key the hash
  function to get a PRF) is indeed the answer that one would immediately
  expect.

+ SipHash is a good choice for the PRF.

Weaknesses
----------
- The assumption in S2, S3, S4 that the adversary knows $h$ makes HLL
  obviously unsound, since HLL fundamentally relies on incoming items
  being put into random buckets (and, though you do not use this fact in
  this paper, having geometrically distributed numbers of 0s starting at
  bit n).  I don't see that observing this is a research contribution.
  Similarly in S1, the adversary doesn't have $h$, but _does_ have a
  black box that implements $h$ and pretty easily returns the result.
  Wouldn't a more reasonable set of adversarial models be "insertion of
  items + m known + (maybe) get cardinality estimate + (maybe) get
  snapshot"?  Are there really deployments of HLL in adversarial
  settings that don't key $h$?  (You indicate in Section 6 that there
  are a number of implementations with a fixed $h$, so if they're indeed
  used in adversarial situations, that's for sure a bug report, but not
  a research contribution.)

- The optimizations in the "complex attack" paragraphs of Section 4.1.2
  are based on the assumption that "We will seek to minimise the number
  of resets performed in our attack, since we assume these are costly."
  (p.6) Why would resets of a shadow device the attacker controls be
  costly?  What is your cost model here?  Even with your optimizations,
  you come to a total number of insertions that you cite as
  "$O(2^{2t})$" (p.7).  But the "simple attack" you presented previously
  also has $O(2^{2t})$ insertions, so if you're claiming a difference,
  the constants are what matter, and you don't directly compare the
  constants in your two attacks in this section.

- Why would a shadow device have the same _cryptographic keys_ as a real
  device?  What kind of deployment scenario are you envisioning here?
  In your security model, you even explicitly say that the key is chosen
  at runtime during the init call (for settings S2, S3, and S4), so why
  should setting S1 be different?

Comments for author
-------------------
In Lemma 1, you say "$B$ out of $2^m$ buckets".  But there are only $m$
buckets, right?

At the end of Section 6, you say that your method has a higher reduction
to the cardinality estimate than that of Reviriego and Ting [6], but
at the cost of more resets.  Shouldn't the number of resets then be
reported in Table 2?  And the results of applying their method to the
same implementation of HLL you used as well?

In Section 7.5, "This mean that" -> "This means that"

Also in Section 7.5, you correctly point out that instantiating HLL with
keyed hashed with different keys make them no longer mergeable.  Would
it be difficult to arrange for the various, say, network switches run by
the same entity (that want to merge their counts) use the same keys?
Having multiple webserver machines or TLS terminators (for load
balancing) sharing certificate private keys is quite common; what is
different about this deployment scenario?

* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * *

Review #186B
===========================================================================

Review recommendation
---------------------
5. Accept

Reviewer expertise
------------------
3. Knowledgeable

Overall merit
-------------
3. Top 25% but not top 10% of submitted papers

Writing quality
---------------
4. Well-written

Paper summary
-------------
This paper presents several attacks on the HyperLogLog unique item sketch algorithm.  This algorithm is used to estimate unique item counts in several widely-used database and networking products.  Previous work had shown that an attacker can cause the algorithm to underestimate the number of unique items seen by a factor of five; this work shows that an attacker that knows the parameters of the algorithm (hash function, number of buckets, number of bits) can essentially insert an unbounded number of items into a stream without appreciably increasing the count, at the cost of rejecting about half of the items sampled.  Several variants of the attack are investigated theoretically and empirically, and a simple countermeasure - using a PRF instead of a hash function - is shown to provably eliminate attacks, although this countermeasure does rule out the useful property of being able to merge distributed HLL sketches.

Strengths
---------
+ Well-written paper

+ Establishes a useful adversary model for investigating security of HLL schemes

+ Exponentially improves the performance of previous attacks

+ Includes both empirical and theoretical evaluation of attacks

+ Gives a simple and easy-to-analyze countermeasure

Weaknesses
----------
- The implications of the attacks are a little unclear: e.g. one example is evading port-scan detection, but if I can only scan some ports, doesn't that defeat the purpose?

- The attacks and defense are both a little obvious: basically, insert only items that won't advance the counter in a bucket, and use a keyed hash.  It would be nice to see some analysis about whether something less than a PRF would work, analogous to pairwise-independent hashing for hash tables.

Comments for author
-------------------
This is a quite nice paper, and I really don't have much to say beyond the bullet points above.  Basically, the only things that I would like to see improved in the paper is a better connection from the actual attacks back to the motivating examples - e.g., show us how some specific security mechanism can fail due to this attack - and maybe some analysis of whether a weaker hash assumption could be used.  (Or analysis of further variations of HLL, if they are needed to make a concrete attack example.)

* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * *

Review #186C
===========================================================================

Review recommendation
---------------------
4. Minor revision

Reviewer expertise
------------------
1. No familiarity

Overall merit
-------------
4. Top 10% but not top 5% of submitted papers

Writing quality
---------------
4. Well-written

Paper summary
-------------
The paper shows that the HyperLogLog algorithm, which is used as a cardinality estimation algorithm for very large data sets, as described in the literature is vulnerable to adversarial attacks. As a countermeasure, the authors propose to use a secret hashing function (a keyed PRF) instead of the public hashing function, and prove the security of this construction.

Strengths
---------
- Novel attack scenarios against an important Big Data algorithm
- Provably secure solution, based on a convincing observation
- Both formal and empirical security analysis

Weaknesses
----------
- HLL algorithm description a bit sketchy
- Formal model needs refinement

Comments for author
-------------------
This paper presents a very interesting security problem, and a convincing solution. The presentation however could benefit from a number of clarifications:
- A section 2.1a 'Simplified HLL' which describes a simplified version of the HLL algorithm would help in understanding the full algorithm. It may also be used as an abstraction for your proofs in 7.4 and the appendix, since you do not go into the details of HLL there.
- In Fig. 1, the variable m is used for different purposes.
(1) $m=2^n$ Perhaps it would be clearer to use the notation $m_n$ since there seem to be 13 different values of m.
(2) $m$ registers
(3) Which value $m$ is meant in the computation of $E$?

Requested Changes
-----------------
Clarifications to the questions above.

* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * *

Review #186D
===========================================================================

Review recommendation
---------------------
2. Reject and resubmit

Reviewer expertise
------------------
1. No familiarity

Overall merit
-------------
2. Top 50% but not top 25% of submitted papers

Writing quality
---------------
2. Needs improvement

Paper summary
-------------
The authors present an attack on they HLL algorithm used to efficiently and probabilistically estimate the cardinality of a large set of values by sampling values from the set.  They show that by manipulating the sample set, they can cause HLL to under-estimate the cardinality.

Strengths
---------
The attack is simple and would likely work.

Weaknesses
----------
Once the mechanism behind HLL is described, the attack is mostly straight forward.  However, it is also fairly easy to stop as a keyed hash makes it difficult for the adversary to mount the attack.  The writing is fairly abstract and it is hard to see how practical this attack is in practice.

Comments for author
-------------------
It would be good to start with a solid motivating scenario that describes end-to-end how an adversary would use this exploit to cause a large amount of damage or harm to the victim.  I also feel the writing could be improved.  The description of HLL itself, as well as the attack, were difficult to follow at times, and the different attack scenarios felt contrived and it was difficult to map them to real-life scenarios.

I was also curious as to how the attack works if a cryptographic (but non-keyed) hash like SHA were used instead of Murmur.  A keyed hash is described but if it is computationally difficult to produce a value that leads to a certain element being inserted into the sketch how does this affect the attack?  While there is no secret, it may still be difficult for the adversary to insert values of their choice.

Requested Changes
-----------------
Please improve the descriptions of HLL and your attack.

Questions for authors' response
-------------------------------
Can you describe the difference between use of a non-cryptographic hash like Murmur3, a cryptographic hash and a MAC?

* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * *

Review #186E
===========================================================================

Review recommendation
---------------------
5. Accept

Reviewer expertise
------------------
2. Some familiarity

Overall merit
-------------
5. Top 5% of submitted papers!

Writing quality
---------------
4. Well-written

Paper summary
-------------
HyperLogLog (HLL) is an algorithm for efficiently estimating the cardinality of sets. HLL is useful for, among many other applications, detecting anomalies in data sets such as network logs, application event logs, and so on. If an attacker could cause the algorithm to significantly underestimate the cardinality, important information about the set (such as data anomalies) would go undetected.

The paper presents, proves, and analyzes a practical attack that induces significant underestimation, improving on previous work.

Strengths
---------
Solid theoretical foundation combined with solid empirical results on a real, meaningful target (Redis with modern HLL).

Results are significant — attack induces a large underestimate in cardinality.

Weaknesses
----------
The empirical results described in Section 6, while promising, are not necessarily something a real-world attacker could meaningfully use (the set of 4-character ASCII strings, or a set of 250K random strings). Similarly, a real-world attacker is unlikely to be lucky enough to get to attack a target with no honest inputs already in the sketch. Those model scenarios might or might not faithfully model a real attack like suggested in section 3 (e.g. a port scanner that evades detection).

Comments for author
-------------------
It entirely depends on your audience, but as an industry schlub I find pseudocode notation easier to understand than mathematical notation. I understand the good reasons for using math notation. I'll be excited to see the source code when you can release it!

I would love to see future work that operationalizes the attack. For example, a modification to ZMap or nmap that successfully evades detection against a defender using Redis for log analysis would be an excellent paper to present at USENIX WOOT or another similar conference.

Requested Changes
-----------------
No significant changes.

Missing reference in Section 5: "In S4, the analysis is the same as in Section ."

Probable typo in Section 6: "only, thus that are initially empty." Probably should be "thus they are".

### 2021-04-22 docs/HLL_exp_bad/main.tex
prep for CCS

@@
-\newif\ifeprint
-%\eprinttrue
-\eprintfalse
+\newif\ifacmversion % turn on ACM style (can be short or full)
+\acmversiontrue
 
-\ifeprint
-\documentclass{article}
-\usepackage{geometry}
-\usepackage[dvipsnames]{xcolor}
+\ifacmversion
+
+\documentclass[sigconf, anonymous, dvipsnames, draft]{acmart} %dvipsnames for old/new macros
+
+\fancyhf{} % Remove fancy page headers 
+\fancyhead[C]{Anonymous submission \#9999 to ACM CCS 2021} % TODO: replace 9999 with your paper number
+\fancyfoot[C]{\thepage}
+
+\setcopyright{none} % No copyright notice required for submissions
+\acmConference[Anonymous Submission to ACM CCS 2021]{ACM Conference on Computer and Communications Security}{Due 6 May 2021}{Seoul, South Korea}
+\acmYear{2021}
+
+\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions
+
+%%\ccsPaper{9999} % TODO: replace with your paper number once obtained
+
 \else
-\documentclass[letterpaper,twocolumn,10pt]{article}
-\usepackage[dvipsnames]{xcolor}
-\usepackage{usenix-2020-09}
+\documentclass[%
+	dvipsnames,	%% xcolor
+	draft,
+]{article}
 \fi
 
-%\documentclass[11pt]{article}
 \usepackage{graphicx, amsmath, amsthm, amssymb, setspace, array, url, hyperref, xspace, todonotes, pifont}
-\usepackage{a4wide}
+%\usepackage{a4wide}
 \usepackage[utf8]{inputenc}
@@
 
-\newtheorem{lemma}{Lemma}
-\newtheorem{theorem}{Theorem}
+%\newtheorem{lemma}{Lemma}
+%\newtheorem{theorem}{Theorem}
 
 \title{HyperLogLog: Exponentially Bad in Adversarial Settings}
-\author{{\rm Kenneth G. Paterson}\\ {\rm ETH Zurich} \and {\rm Mathilde Raynal\thanks{Work done while an exchange student at ETH Zurich.}}\\ {\rm EPFL}}
+%\author{{\rm Kenneth G. Paterson}\\ {\rm ETH Zurich} \and {\rm Mathilde Raynal\thanks{Work done while an exchange student at ETH Zurich.}}\\ {\rm EPFL}}
 %{\tt kenny.paterson@inf.ethz.ch, mraynal@student.ethz.ch}}
@@
 This trade-off makes the HLL algorithm very attractive for a wide range of applications such as database management and network monitoring, where an exact count may not be needed.
+The HLL algorithm and variants of it are implemented in systems such as Redis and Google Big Query.
 Recently, the HLL algorithm has started to be proposed for use in scenarios where the inputs may be adversarially generated, for example counting social network users or detection of network scanning attacks.
@@
 \begin{itemize}
+\item Google and counting stuff.\todo{add more detail here -- their blogs, their paper, the use in Google Big Query.} 
     \item Cardinality estimation provides a significantly faster way for Facebook to compute statistics on their users.
@@
 We discuss and benchmark a suitable choice for the required VIL-PRF.
+\todo{make the model a contribution...
+``To achieve this...
+introduce simulation-based security model for cardinality estimators'', etc}
+
+\todo{Simplicity of attacks as a strength; only complexity comes in S1 attack which we don't consider that interesting!}
+\todo{No e2e attack on a specific system; rather paper is written in general form since the attacks apply broadly.}
 
@@
 Section~\ref{sec:redis} considers the security of the more modern version of HLL due to Ertl~\cite{hllnew} that is used in Redis.
-Our experimental results are discussed in Section~\ref{sec:exp} while Section~\ref{sec:formal} provides our formal security analysis of HLL.
+Our experimental results are discussed and related back to specific use cases in Section~\ref{sec:exp}.
+Section~\ref{sec:formal} provides our formal security analysis of HLL.
 Section~\ref{sec:conclusions} provides our conclusions and some ideas for future work.
@@
 We discuss the two attack scenarios from~\cite{cardestprivacy} in more detail in Section~\ref{sec:attacks}.
+\todo{update: we will get privacy for free from our sim based security model if sim does not use input $x$.}
 
@@
 We show in Section~\ref{sec:formal} that when a salt is used appropriately (i.e.\ as the secret key for a pseudo-random function that is used in place of hash function $h$), security against adversarial inputs can be achieved even in our strongest adversarial setting.
+\todo{Add a bit more detail here.}
 %This might not be a convenient solution for some applications, they thus encourage further research.
@@
 HLL~\cite{hll} is a streaming algorithm based on the key observation that, for a stream of randomly distributed values represented as bit-strings of some fixed length, if we observe a value with the maximum of $k$ leading zero-bits, then the cardinality of the stream (i.e.\ the number of distinct values it contains) is likely to be on the order of $2^{k+1}$.
-In practice, to ensure a uniform distribution, the stream values are first processed by a hash function $h$ before the leading bits are inspected.
+To ensure a uniform distribution, the stream values are first processed by a hash function $h$ before the leading bits are inspected.
 So to estimate the cardinality of a stream, we need only to store a representation of $k$, the length of the largest observed string of leading zero-bits for numbers in the stream.
@@
 When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate, looking at the number of empty buckets (whose register value is 0) instead of the number of leading zeros in the stream.
-A similar correction is applied for large cardinalities in order to take into account the likelihood of hashing collisions.
+A similar correction is applied for large cardinalities in order to take into account the likelihood of hash collisions.
 In addition, the cardinality estimate is usually rounded to a whole number before being output.
@@
 For example, \cite{hllpractice} uses a 64-bit hash function and Linear Counting along with bias correction in the small cardinality range, while~\cite{hllnew} proposes new estimators based on a more precise statistical analysis of the counting problem.
-In this report we mostly target the HLL formulation in the original paper~\cite{hll} as represented in Fig.~\ref{fig:hll}.
+In this report we mostly target the HLL formulation in the original paper~\cite{hll} as represented in Fig.~\ref{fig:hll} and as widely used in practice.
 We also study~\cite{hllnew} because of its recent deployment in Redis.
@@
 S4 is a slightly weaker version of the \textit{insider} attack scenario in~\cite{cardestprivacy}, the difference being that an insider adversary in~\cite{cardestprivacy} has continuous access to the HLL sketch internals including all bucket values, while our adversary in S4 only has access to a snapshot of the HLL sketch internals at the start of its attack.
-We consider a weaker version of S4 than in~\cite{cardestprivacy} because it is sufficient for our attacks; when considering security proofs in Section~\ref{sec:formal}, we will switch back to the stronger version of S4.
+We consider a weaker version of S4 than in~\cite{cardestprivacy} because it is sufficient for our attacks; when considering formal security analysis in Section~\ref{sec:formal}, we will switch back to the stronger version of S4.
 
@@
 
+\todo{What did referee say about other attack scenarios we should consider?}
+
 %Furthermore, we show that although~\cite{cardestprivacy} presents S3 (external adversary) and the stronger version of our S4 (insider adversary) as two separate scenarios, an S3 adversary can easily infer the knowledge given to our S4 adversary in its snapshot.
@@
 For each case, we consider adversaries operating in scenarios S1, S2 and S4.
+
+\todo{Provide a brief sketch here of the ``obvious attack'', and then say that we target improvements over this attack.}
+\todo{remove $t$ parameter or use consistently.}
 
@@
 Here the attack is somewhat more complex.
-We begin by explaining the core idea, then a simple version of our attack, then a more complex but performant version, and finally, we give a heuristic performance analysis of the attack.
+We begin by explaining the core idea, then a simple version of our attack, then a more complex but performant version, and finally, we give a heuristic performance analysis of the more complex attack.
 
@@
 
-The insertion cost of the attack can be reduced by more smartly ordering the items $x$ to be trial inserted on each pass.
+The insertion cost of the attack can be reduced by more cleverly ordering the items $x$ to be trial inserted on each pass.
 Note that, in a given pass, when $B^\prime$ is still close to $B$, an item $x$ that does not increase the number of active buckets is more likely to have hit one of the target buckets than is the case when $B^\prime$ is large; such an $x$ can be tried earlier in the sequence of trials in the next pass and its exact status determined sooner.
@@
 This is highly likely to be the case (with probability at least $1-1/m^2$) as long as at least $m\ln m$ items have already been inserted into the target HLL.
-This follows from the standard analysis of the coupon collector problem under our usual assumption that $h$ behaves as a random function.
+This follows from the standard analysis of the coupon collector problem\todo{add a reference for this} under our usual assumption that $h$ behaves as a random function.
 In S4, the adversary can use its snapshot to detect if this is the case or not, and behave differently if not (details for this case follow below).
@@
 
-We present a brief analysis of the HLL implementation used in Redis, since Redis is a very popular distributed, in-memory key-value store implementation and the version of HLL used there differs significantly from that described in Figure~\ref{fig:hll}.
+We present a brief analysis of the HLL implementation used in Redis, since Redis is a popular distributed, in-memory key-value store implementation and the version of HLL used there differs significantly from that described in Figure~\ref{fig:hll}.
 
@@
 Redis 5.0 released in October 2018 replaces this with a new cardinality estimator based closely on the work of Ertl~\cite{hllnew}.
-This method is still used in the current release (Redis 6.0.8) and so is the object of our study here.\footnote{For details of Redis releases, see \url{https://github.com/redis/redis/blob/6.0/00-RELEASENOTES} and other files obtained by replacing ``6.0'' in this URL with the desired version number.}
+This method is still used in the current release (Redis 6.2.2) and so is the object of our study here.\footnote{For details of Redis releases, see \url{https://github.com/redis/redis/blob/6.2/00-RELEASENOTES} and other files obtained by replacing ``6.2'' in this URL with the desired version number.}
 
@@
 HLL is also featured in many frameworks, including Redis~\cite{redis}, Google's BigQuery~\cite{bigquery}, Facebook's Airlift~\cite{airlift} and several products of Apache such as Spark~\cite{spahll} and Druid~\cite{druhll}.
-It is interesting to point out that all these implementations use a constant number of buckets and a fixed hash function.
-Thus they are potentially vulnerable to our attacks.
+All these implementations use a constant number of buckets and a fixed hash function.
+Thus they are  vulnerable to our attacks.
+\todo{maybe switch to one of these, or somehow be more clear about how the attacks transfer.}
 
@@
 For simplicity, we take ${\cal X}$ to be the set of 4-character long ASCII strings, with $\lvert {\cal X} \rvert = 7,311,616= 2^{22.8}$.
-We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}.
+\todo{Explain there's nothing special about these inputs; anything would do; processed by hash function anyway.}We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}.
 %(note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
@@
 This comes at the cost of performing more resets compared to~\cite{hllvuln}.
+\todo{explain why counting resets matter; also that we don't care too much about this scenario.}
 
@@
 
+\todo{Add a couple of paragraphs to relate all this back to the initial attack scenarios, e.g.
+packet flows.
+May need to look at internals of some of those papers to see where flex in inputs will come from.
+}
+
 \section{Formal Security Analysis}\label{sec:formal}
@@
 
+\todo{update! may need to formally define algs of an HLL as well as oracles. Since Sim will need to keep a shadow copy for itself.}
+
 We then show that, if $h$ is replaced by a keyed Pseudo Random Function (PRF) $F$ whose key is not available to the adversary, then any adversary ${\cal A}$ that can distinguish between the real and random worlds with oracles as in S4 can be used to build an adversary ${\cal B}$ which breaks the PRF security of $F$.
@@
 
+\todo{explain the above better with simulators + conf for free.}
+
 We note that our formal security analysis could be conducted using the framework recently laid out in~\cite{CCS:ClaPatShrPS19}. However, we prefer to give a compact and self-contained treatment here.
 
+\todo{explain why a sim-based approach is better, more general, does not require a specific winning condition.}
+
+\todo{Why UHFs don't work in general -- key recovery is possible in S4; partial info about processed values and therefore keys  leaks from Card Est in S2, though harder to make an attack from this. }
+
 \subsection{Formal Security Model for HLL}
+
+\todo{Upgrade to use simulator S as a typical simulation-based security notion should; reference to Lindell's ``How to simulate it'' paper?
+Order of quantifiers; input to simulator: nothing, not even number of queries, assuming adversary never repeats an insert query.}
 
@@
 The \texttt{read} and \texttt{corrupt} queries do leak information about $F$'s outputs to the adversary, but such outputs are indistinguishable from random under the assumption that $F$ is a PRF.
+
+\todo{revisit the above in view of introduction of simulator, plus the confidentiality aspect for free. NB corruption query does not give PRF key.}
 

### 2021-04-23 docs/HLL_exp_bad/main.tex
work on paper

@@
 
-\maketitle
-%\IEEEtitleabstractindextext{%
-
 \begin{abstract}
@@
 \end{abstract}
-%}
 
-%\IEEEdisplaynontitleabstractindextext
+\maketitle
 
@@
 In this paper, we present a complete analysis of HLL in adversarial settings, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
-In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of ${\cal O}(2^t)$ where $t$ is the number of bits of flexibility in the adversary's input).
+In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of ${\mathcal{O}}(2^t)$ where $t$ is the number of bits of flexibility in the adversary's input).
 We demonstrate this theoretically and experimentally for the ``classic'' version of HLL.
@@
 Desfontaines \emph{et al.}~\cite{cardestprivacy} studied the privacy properties of HLL in two different attack scenarios, referred to as \emph{insider} and \emph{external}.
-The attack target of~\cite{cardestprivacy} is different from ours, being focused on privacy breaches, whereas we care about manipulating cardinality estimation.
-%It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
+The attack target of~\cite{cardestprivacy} is different from ours, being focused on privacy breaches.
+It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
+In contrast, we are concerned with an attacker attempting to manipulate the cardinality estimation process itself (an issue not considered in~\cite{cardestprivacy}).
 We discuss the two attack scenarios from~\cite{cardestprivacy} in more detail in Section~\ref{sec:attacks}. 
-\todo{update: we will get privacy for free from our sim based security model if sim does not use input $x$.}
+%\todo{update: we will get privacy for free from our sim based security model if sim does not use input $x$.}
 
@@
 For completeness, we include analyses of this attack scenario, showing that we can drastically improve on the attacks in~\cite{hllvuln} (e.g.\ we can reduce the HLL cardinality estimate by a factor of almost 1,000 instead of the factor of 5 achieved in~\cite{hllvuln}).
+However, we stress that we consider this scenario to be the least realistic (and therefore the least interesting).
 
@@
 We show in Section~\ref{sec:formal} that when a salt is used appropriately (i.e.\ as the secret key for a pseudo-random function that is used in place of hash function $h$), security against adversarial inputs can be achieved even in our strongest adversarial setting.
-\todo{Add a bit more detail here.}
-%This might not be a convenient solution for some applications, they thus encourage further research.
+In addition, our simulation-based security definition, in which the simulator in the ideal world has no access to the inserted values, automatically guarantees confidentiality of the values inserted into the HLL sketch, even in the strongest attack model considered in~\cite{cardestprivacy}.
+So, while our main objective is to protect against adversarial inputs, we also get privacy for free when using a keyed PRF in place of the HLL hash function.
 
@@
 Clayton \emph{et al.}~\cite{CCS:ClaPatShrPS19} recently provided a provable-security treatment of this topic.
-They used their formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures, but not HLL.
+They used a game-based formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures.
+They did not study HLL.
+Our use of a simulation-based security definition in place of the game-based ones used in~\cite{CCS:ClaPatShrPS19} is another point of distinction; see further discussion in Section~\ref{sec:formal}.
+
 Prior work also focused on Bloom filters~\cite{NaorY15,GerbetKL15} or flooding of hash tables~\cite{LiptonN93,CrosbyW03}.
@@
 
-%It is an interesting open problem to use their framework to develop a formal understanding of how HLL can be made secure in the adversarial setting.
-
 \section{Overview of HLL}\label{sec:overview}
 \subsection{The Algorithm}
-HLL~\cite{hll} is a streaming algorithm based on the key observation that, for a stream of randomly distributed values represented as bit-strings of some fixed length, if we observe a value with the maximum of $k$ leading zero-bits, then the cardinality of the stream (i.e.\ the number of distinct values it contains) is likely to be on the order of $2^{k+1}$.
-To ensure a uniform distribution, the stream values are first processed by a hash function $h$ before the leading bits are inspected.
-So to estimate the cardinality of a stream, we need only to store a representation of $k$, the length of the largest observed string of leading zero-bits for numbers in the stream.
-This provides a very compact mechanism for estimating the stream cardinality -- if the true cardinality is $N$, then just $\log\log(N)$ bits are needed to store the estimate.
+HLL~\cite{hll} is based on the key observation that, for a stream of items represented as bit-strings of some fixed length, if all items have at most $k$ leading zero-bits (and one or more items actually have this many leading zero-bits), then the cardinality of the stream (i.e.\ the number of distinct values it contains) is likely to be on the order of $2^{k+1}$.
+To ensure a uniform distribution on the items' representations as bit-strings, the items are processed by a hash function $h$ before the leading bits are inspected.
 
-This simple approach suffers from large variance in the estimation.
-In order to improve the estimate, we can use many estimators in parallel instead of one, and average the results.
-Each estimate is stored in its own register, that we call a bucket.
+So to roughly estimate the cardinality of a stream, we need only to store a representation of $k$, the length of the largest observed string of leading zero-bits for items in the stream.
+This provides a very compact mechanism for estimating a stream's cardinality -- if the true cardinality is $N$, then just $\mathcal{O}(\log\log(N))$ bits of storage are needed to compute the cardinality estimate.
+
+This simple approach to cardinality estimation suffers from large variance in the estimation.
+In order to improve the variance, HLL uses many estimators in parallel instead of one and averages the results.
+Each estimate is stored in its own register that we call a bucket.
 We refer to the collection of buckets and their contents as the HLL \emph{sketch}.
 To map a value $x$ to one of $m=2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$.
-The output length of $h(\cdot)$ is thus $n+\ell$, fixed to 32 in~\cite{hll}.
+The output length of $h(\cdot)$ is thus $n+\ell$, fixed to 32 in the original HLL paper~\cite{hll}.
 Estimates from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha_m$, where $\alpha_m$ is empirically computed to correct a systematic multiplicative bias.
 
-Relatively small and big cardinalities are handled separately to correct for systematic distortions, as shown in Fig.~\ref{fig:hll}.
-When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate, looking at the number of empty buckets (whose register value is 0) instead of the number of leading zeros in the stream.
+In the original HLL proposal~\cite{hll}, relatively small and large cardinalities are handled separately.
+This is to correct for systematic errors.
+When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate.
+This estimator uses the number of empty buckets (whose register value is 0) instead of the number of leading zeros in the stream.
 A similar correction is applied for large cardinalities in order to take into account the likelihood of hash collisions.
 In addition, the cardinality estimate is usually rounded to a whole number before being output.
-
-HLL as described in Figure~\ref{fig:hll} is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
+The original HLL algorithm from~\cite{hll} is shown in Fig.~\ref{fig:hll}.
+This algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
 
 Some works building on~\cite{hll} replace the Hit Counting algorithm by other schemes and use a hash function with a larger output domain to reduce the chances of collision and thus get rid of the need for large cardinality range correction.
-For example, \cite{hllpractice} uses a 64-bit hash function and Linear Counting along with bias correction in the small cardinality range, while~\cite{hllnew} proposes new estimators based on a more precise statistical analysis of the counting problem.
-In this report we mostly target the HLL formulation in the original paper~\cite{hll} as represented in Fig.~\ref{fig:hll} and as widely used in practice.
+For example, Google engineers proposed using a 64-bit hash function and Linear Counting along with bias correction in the small cardinality range~\cite{hllpractice},\footnote{See also the Google blogpost describing the implementation in Google Big Query at~\url{https://cloud.google.com/blog/products/gcp/counting-uniques-faster-in-bigquery-with-hyperloglog}.} while Redis has adopted a new estimator based on a more precise statistical analysis of the counting problem provided by~\cite{hllnew}.
+
+In this paper we mostly focus on the HLL formulation in the original paper~\cite{hll} as represented in Fig.~\ref{fig:hll} and as widely used in practice.
 We also study~\cite{hllnew} because of its recent deployment in Redis.
 We leave for future work the analysis of further HLL variants.
+\todo{revise this if we add Google BQ.}
 
@@
 \vspace{0.5em}
-\textbf{Algorithm} HyperLogLog (\textbf{input} ${\cal X}$ : multiset of items from domain $\mathcal{D}$).\\
+\textbf{Algorithm} HyperLogLog (\textbf{input} ${\mathcal{X}}$ : multiset of items from domain $\mathcal{D}$).\\
 \textbf{initialize} a collection of $m$ registers, $M[1], ..., M[m]$, to $0$;\\
 \vspace{0.5em}
-\textbf{for} $x \in \cal X$ \textbf{do}\\
+\textbf{for} $x \in {\mathcal{X}}$ \textbf{do}\\
     \hspace{2em} \textbf{set} $y := h(x);$\\
@@
 This section presents different adversarial settings and corresponding attacks that manipulate the HLL cardinality estimate.
-The adversary's goal is to reduce the estimate as much as possible, since this would cause more damage when  considering the above-mentioned applications of HLL.
+The adversary's goal is to reduce the estimate as much as possible, since this would cause more damage when  considering the previously mentioned applications of HLL.
 It is easy to modify our attack strategies for an adversary who instead wishes to artificially inflate the HLL cardinality estimate given a limited input capability.
@@
 
-Throughout, we assume the adversary has the ability to vary the contents of free fields in the input strings which will be inserted into the stream of values.
-For example it can manipulate IPID and other header fields in the context of IP packets.
-We model this by assuming the adversary has access to a large set ${\cal X}$ of potential inputs from which it can select items to insert into the HLL sketch.
-Our adversary's objective, then, is to insert as many items from ${\cal X}$ as possible whilst keeping the HLL cardinality estimate as low as possible.
+\subsection{Attacker's Objective, Abstractly}
+Throughout, we assume the adversary has the ability to vary the contents of free fields in the strings which will be inserted into the stream of values.
+For example it might manipulate IPID and other header fields in the context of IP packets.
+We model this in a general way by assuming the adversary has access to a large set ${\mathcal{X}}$ of potential inputs from which it can select items to insert into the HLL sketch.
+Our adversary's objective, then, is to insert as many items from ${\mathcal{X}}$ as possible whilst keeping the HLL cardinality estimate as low as possible.
 
@@
 This would make any pre-computation on the items using the shadow device irrelevant, and considerably weaken the attack presented in Section 3.2 of~\cite{hllvuln}.
-Hence we assume for the rest of this work that, in scenario S1, shadow and targeted devices have the same internals and security parameters (if there are any).
+Hence we assume for the rest of this work that, in scenario S1, shadow and targeted devices have the same internals and security parameters (if there are any).\footnote{This scenario is roughly analogous to the common situation where a fixed private key is stored in multiple devices acting as TLS termination points for a single web site.}
 %If the HLL device receives inputs from several users, the attack from Section 3.2 of~\cite{hllvuln} is then the easiest to carry out.
@@
 
-\todo{What did referee say about other attack scenarios we should consider?}
+%\todo{What did referee say about other attack scenarios we should consider?}
 
@@
 For each case, we consider adversaries operating in scenarios S1, S2 and S4.
+We operate in the situation where the adversary has a set $\mathcal{X}$ of potential inputs which it can choose to insert into the HLL sketch, and where it wishes to insert as many of them as it can whilst keeping the HLL cardinality estimate as low as possible.
 
-\todo{Provide a brief sketch here of the ``obvious attack'', and then say that we target improvements over this attack.}
-\todo{remove $t$ parameter or use consistently.}
+\subsection{A Simple Attack}\label{sec:simple}
+
+In scenarios S2 and S4 with an initially empty HLL sketch, there is a very simple attack based on simply evaluating $h(x)$ for each $x \in \mathcal{X}$, looking at the bits of $h(x)$ to see which bucket $j$ would be affected if $x$ was inserted, and then deciding whether to insert or not based on some rule.
+
+For example, the rule could be to insert $x$ if and only if doing so does not increase the entry $M[j]$ stored in the relevant bucket beyond some pre-defined threshold value $v$.
+A simple computation based on the HLL raw cardinality estimator function in Fig.~\ref{fig:hll} shows that this rule would allow the insertion of a fraction of  $1-2^{-v}$ of the set $\mathcal{X}$ whilst keeping the raw cardinality estimate below $\alpha_m m 2^v$.
+This attack assumes $h$ behaves like a random function and exploits the geometric distribution of the position of the leading ``1'' bit in $w$, where $w = y_{n+1}y_{n+2}\ldots y_{n+1}$ and $y=h(x)$ (as in the HLL algorithm).
+
+This simple attack is already very effective.
+For example, setting $v=1$ in the above rule, the attacker can insert half of all elements of $\mathcal{X}$ whilst keeping the cardinality estimate below $2\alpha_m m$ (where, recall, $m$ is the fixed number of buckets in the HLL sketch).
+
+In the remainder of this section, we target attacks that are more effective than this simple attack.
 
@@
 \begin{lemma}\label{lem:bound}
-Suppose an adversary inserts elements $x$ from some set ${\cal X}$ into an initially empty HLL sketch.
+Suppose an adversary inserts elements $x$ from some set ${\mathcal{X}}$ into an initially empty HLL sketch.
 Suppose the adversary can guarantee that in its attack, only a subset of size $B$ out of $m$ buckets are hit, where $B \leq B_{\max}$ with
@@
 Now our proposed attack is very simple.
-Suppose the adversary has some set ${\cal X}$ of $2^t$ items from which to pick its inputs.
+Suppose the adversary has some set ${\mathcal{X}}$ of items from which to pick its inputs.
 The adversary selects a value of $B \leq B_{\max}$, and defines its target buckets to be those numbered $1,\ldots,B$.
-Since the adversary knows $h$, it can tell whether, for each $x \in {\cal X}$, insertion of $x$ into the HLL sketch would hit one of the target buckets or not; it inserts those that do.
-In expectation, assuming $h$ behaves like a random function, the adversary will successfully insert a fraction $B/m$ of the items, totalling $(B/m)\cdot 2^t = 2^{t-n+\log_2(B)}$ items.
+Since the adversary knows $h$, it can tell whether, for each $x \in {\mathcal{X}}$, insertion of $x$ into the HLL sketch would hit one of the target buckets or not; it inserts those that do.
+In expectation, assuming $h$ behaves like a random function, the adversary will successfully insert a fraction $B/m$ of the items, totalling $(B/m)\cdot |{\mathcal{X}}|$ items.
 Meanwhile the final cardinality estimate will remain below $m \cdot \ln (1+\frac{B}{m-B})$.
 
-For example, setting $B = m/2$ (which, for practical parameters, satisfies $B \leq B_{\max}$), we can expect to insert $2^{t-1}$ items while keeping the cardinality estimate below $m \ln 2$.
+For example, setting $B = m/2$ (which, for practical parameters, satisfies $B \leq B_{\max}$), we can expect to insert half of all the items in $\mathcal{X}$ while keeping the cardinality estimate below $(\ln 2)m \approx 0.693m$.
+This is superior to the simple attack in Section~\ref{sec:simple} (which, recall, could insert half of $\mathcal{X}$ while keeping the cardinality estimate below $2\alpha_m m \approx 1.44m$).
 
-As a second example, setting $B = 1$, we can expect to insert $2^{t-n}$ items with a final cardinality estimate equal to $m \cdot \ln (1+\frac{1}{m-1})$, which is rounded to $1$ by the HLL algorithm (for $m$ of practical interest, e.g.\ $m=256$).
-Thus we keep the HLL cardinality estimate to a constant (1!) whilst being able to insert exponentially many items, provided there is sufficient flexibility in the input set ${\cal X}$.
+As a second example, setting $B = 1$, we can expect to insert $|{\mathcal{X}}|/2^m$ items with a final cardinality estimate equal to $m \cdot \ln (1+\frac{1}{m-1})$, which is rounded to $1$ by the HLL algorithm (for $m$ of practical interest, e.g.\ $m=256$).
+Thus we keep the HLL cardinality estimate to a constant (1!) whilst being able to insert a constant fraction of the items from the input set ${\mathcal{X}}$.
 
@@
 \paragraph{Core Idea:}
-We assume the adversary again has (or is given) some set ${\cal X}$ of $2^t$ items from which to pick its inputs.
-The core idea of our attack is to use the first $B \leq B_{\max}$ items $x_1,\ldots, x_B$ from ${\cal X}$ to define a set of target buckets, and then to insert into the shadow HLL sketch further items $x \in {\cal X}$, checking for a cardinality estimate increase on each insertion.
-We initialise a list ${\cal L}$ containing $x_1,\ldots, x_B$.
-We add those items that do not increase the cardinality estimate on insertion into the shadow HLL sketch to ${\cal L}$ and reject the rest.
-We reset the shadow device when the cardinality estimate has increased too much, reinserting the first $B$ items from ${\cal X}$ again to re-establish a ``baseline'' of target buckets for the shadow HLL sketch.
+We assume the adversary again has (or is given) some set ${\mathcal{X}}$ of items from which to pick its inputs.
+The core idea of our attack is to use the first $B \leq B_{\max}$ items $x_1,\ldots, x_B$ from ${\mathcal{X}}$ to define a set of target buckets, and then to insert into the shadow HLL sketch further items $x \in {\mathcal{X}}$, checking for a cardinality estimate increase on each insertion.
+We initialise a list ${\mathcal{L}}$ containing $x_1,\ldots, x_B$.
+We add those items that do not increase the cardinality estimate on insertion into the shadow HLL sketch to ${\mathcal{L}}$ and reject the rest.
+We reset the shadow device when the cardinality estimate has increased too much, reinserting the first $B$ items from ${\mathcal{X}}$ again to re-establish a ``baseline'' of target buckets for the shadow HLL sketch.
 A key observation is that if in the attack we keep the number of ``active'' buckets to values $B^\prime$ that are always below $B_{\max}$, then the Hit Counting algorithm is used to compute the cardinality estimate; given its output $m \cdot \ln (m/V)$ (or, more exactly, a rounded version of this value) we can infer the exact value of $V$, and hence $B^\prime$.
 This enables precise detection of cardinality increases, provided we perform a reset before the Hit Counting algorithm stops being used (this is ensured by resetting when the estimate indicates that $B_{\max}$ buckets have been hit).
-After all of ${\cal X}$ has been processed as described, we take the constructed list ${\cal L}$ and reprocess it in the same way.
+After all of ${\mathcal{X}}$ has been processed as described, we take the constructed list ${\mathcal{L}}$ and reprocess it in the same way.
 We must ensure that the same set of initial items $x_1,\ldots, x_B$ is inserted into the shadow HLL sketch before any other items, so as to fix the same set of target buckets.
@@
 By repeated reprocessing, we will gradually remove items that do not hit the $B$ target buckets (defined by $x_1,\ldots, x_B$).
-Eventually, we obtain a final list ${\cal L}^*$ that we insert into the target HLL sketch.
+Eventually, we obtain a final list ${\mathcal{L}}^*$ that we insert into the target HLL sketch.
 
 By construction and from Lemma~\ref{lem:bound}, we are guaranteed that the cardinality estimate for the target HLL sketch will be at most $m \cdot \ln (m/V)$ where $V = m-B$.
-Assuming $h$ behaves like a random function, it is clear that the expected size of the final list ${\cal L}^*$ is $\frac{B}{m}\cdot |{\cal X}|$.
-Thus the attack's final performance is the same as that in scenarios S2 and S4 above (but more processing and resets are required).
+Assuming $h$ behaves like a random function, it is clear that the expected size of the final list ${\mathcal{L}}^*$ is $\frac{B}{m}\cdot |{\mathcal{X}}|$.
+Thus the attack's final performance is the same as that in scenarios S2 and S4 above.
+However, more processing and resets are required (and the number of resets should be considered as something to be minimised since resetting the shadow device may take a certain amount of time).
 
@@
 The simplistic version of our attack uses the above idea, inserting items $x$ into the shadow HLL sketch but resetting it as soon as the cardinality estimate increases from $B$ to $B+1$ (indicating that a non-target bucket has been hit by item $x$).
-Acting in this way precisely determines whether each inserted item hits the target buckets or not, so has no ``false positives'' which need to be eliminated through reprocessing of the list ${\cal L}$.
-This means that the attack can be carried out in a streaming fashion in a single pass over ${\cal X}$.
-Assuming $h$ behaves like a random function, we must perform a reset with probability $\frac{m-B}{m} = 1 - \frac{B}{m}$ per item inserted, and hence in total an expected number of $(1 - \frac{B}{m})\cdot 2^t$ resets.
+Acting in this way precisely determines whether each inserted item hits the target buckets or not, so has no ``false positives'' which need to be eliminated through reprocessing of the list ${\mathcal{L}}$.
+This means that the attack can be carried out in a streaming fashion in a single pass over ${\mathcal{X}}$.
+Assuming $h$ behaves like a random function, we must perform a reset with probability $\frac{m-B}{m} = 1 - \frac{B}{m}$ per item inserted, and hence in total an expected number of $(1 - \frac{B}{m})\cdot | \mathcal{X} |$ resets.
 
@@
 However, when the number of active buckets has reached some value $B^\prime \leq B_{\max}$ this means that an item $x$ may not increase the cardinality estimate, yet still not hit one of the $B$ target buckets (instead it hits one of the other $B^\prime - B$ active buckets).
-Such an $x$ acts as a ``false positive" and needs to be eliminated through reprocessing of the list ${\cal L}$.
+Such an $x$ acts as a ``false positive" and needs to be eliminated through reprocessing of the list ${\mathcal{L}}$.
 
 \paragraph{Analysis of the Complex Attack:}
-%Suppose that after each reset, the order of the items in ${\cal X}$ (or ${\cal L}$) that have not yet been processed on the current pass is randomised.
+%Suppose that after each reset, the order of the items in ${\mathcal{X}}$ (or ${\mathcal{L}}$) that have not yet been processed on the current pass is randomised.
 (Here we exclude the special items  $x_1,\ldots, x_B$ from the reordering.) Then
@@
 
-Now we expect to reject in total $(1 - \frac{B}{m})\cdot 2^t$ items, and each one needs an expected number of trial insertions that is bounded by $m/(m - B_{\max} + B)$. 
+Now we expect to reject in total $(1 - \frac{B}{m})\cdot | \mathcal{X} |$ items, and each one needs an expected number of trial insertions that is bounded by $m/(m - B_{\max} + B)$. 
 %(In fact, we can model the number of trial insertions needed as an independent geometric distribution with parameter $q$ for each such $x$.) 
-Meanwhile, each of the $\frac{B}{m}\cdot 2^t$ items in the final insertion list ${\cal L}^*$ for the target HLL sketch is trial inserted into the shadow HLL sketch on every pass.
+Meanwhile, each of the $\frac{B}{m}\cdot | \mathcal{X} |$ items in the final insertion list ${\mathcal{L}}^*$ for the target HLL sketch is trial inserted into the shadow HLL sketch on every pass.
 
@@
 \[
-N_R:= \frac{1 - \frac{B}{m}}{B_{\max} - B}\cdot 2^t.
+N_R:= \frac{1 - \frac{B}{m}}{B_{\max} - B}\cdot | \mathcal{X} |.
 \]
@@
 \[
-N_I \leq \frac{m}{ m-B_{\max} + B} \cdot \left(1 - \frac{B}{m} \right)\cdot 2^t +   N_R \cdot \frac{B}{m}\cdot 2^t.
+N_I \leq \frac{m}{ m-B_{\max} + B} \cdot \left(1 - \frac{B}{m} \right)\cdot | \mathcal{X} | +   N_R \cdot \frac{B}{m}\cdot | \mathcal{X} |.
 \]
 
-As a concrete illustration, we take $B = m/2$ and $B_{\max} = 0.71\cdot m$, to obtain $N_R = (2.38/m) \cdot 2^t$ and $N_I \leq 0.63 \cdot 2^t + 1.19\cdot 2^{2t}/m$.
-Here the dominant term in $N_I$ is ${\cal O}(2^{2t})$, i.e.\ the total insertion cost grows as the square of the size of the starting set ${\cal X}$.
-The result of the attack is that, as in S2 and S4, an expected $2^{t-1}$ items can be inserted into the target HLL sketch while keeping its cardinality estimate below $m \ln 2$.
+As a concrete illustration, we take $B = m/2$ and $B_{\max} = 0.71\cdot m$, to obtain $N_R = (2.38/m) \cdot | \mathcal{X} |$ and $N_I \leq 0.63 \cdot | \mathcal{X} | + 1.19\cdot | \mathcal{X} |^2/m$.
+Here the dominant term in $N_I$ is ${\mathcal{O}}(| \mathcal{X} |^2)$, i.e.\ the total insertion cost of this attack in S1 grows as the square of the size of the starting set ${\mathcal{X}}$ (instead of linearly as in S2 and S4).
+The end result of the attack is that, as in S2 and S4, an expected $| \mathcal{X} |/2$ items can be inserted into the target HLL sketch while keeping its cardinality estimate below $m \ln 2$.
 
@@
 Note that, in a given pass, when $B^\prime$ is still close to $B$, an item $x$ that does not increase the number of active buckets is more likely to have hit one of the target buckets than is the case when $B^\prime$ is large; such an $x$ can be tried earlier in the sequence of trials in the next pass and its exact status determined sooner.
-Any $x$ that can be determined to be a true positive (because its insertion into the shadow HLL sketch keeps the number of active buckets at exactly $B$) can be placed immediately in ${\cal L}^*$ and not reprocessed any further.
+Any $x$ that can be determined to be a true positive (because its insertion into the shadow HLL sketch keeps the number of active buckets at exactly $B$) can be placed immediately in ${\mathcal{L}}^*$ and not reprocessed any further.
 We leave the analysis of this optimisation to future work.
@@
 %The adversary blindly picks the first $b$ buckets to be hit by simply inserting items into the shadow HLL until the cardinality estimate indicates that $b$ buckets have been touched.
-It puts these items in a final target list ${\cal L}$.
+It puts these items in a final target list ${\mathcal{L}}$.
 It continues to insert items, discarding all items whose insertion causes the cardinality estimate for the shadow HLL sketch to increase from $b+k$ to $b+k+1$ for $k \in \mathbb{N}$ (here we allow some slack in the total number of buckets that have been hit in order to reduce the number of resets of the shadow device that are required in the attack).
 In other words, when $b+k$ buckets have already been hit and the insertion of an item causes the cardinality estimate to increase further, the adversary can discard this item as it definitely knows that a new bucket (that is not one of the $b$ targeted buckets) switched to a non-empty state.
-On the other hand, it adds to the list ${\cal L}$ any item that does \emph{not} cause the cardinality estimate to increase.
+On the other hand, it adds to the list ${\mathcal{L}}$ any item that does \emph{not} cause the cardinality estimate to increase.
 Such items may have hit one of the initial $b$ buckets, or one of the later buckets; at this stage the adversary cannot be certain.
@@
 %As soon as the cardinality estimate exceeds $\frac{5}{2}m$, the adversary can assume that the shadow HLL sketch's output corresponds to the raw HLL cardinality estimate with high probability.
-In this case, no more useful information can be inferred about the bucket mappings in the shadow HLL sketch, so the adversary resets the shadow HLL sketch to the empty state and starts again with the list ${\cal L}$ of non-discarded items obtained so far.
-Reinserting these items will (eventually) determine whether they hit the original $b$ buckets or later buckets, and hence whether they should be retained in ${\cal L}$ or discarded.
-Once ${\cal L}$ is reprocessed (and assuming the cardinality estimate does not exceed $\frac{5}{2}m$), the adversary can start to insert fresh items, as previously described.
+In this case, no more useful information can be inferred about the bucket mappings in the shadow HLL sketch, so the adversary resets the shadow HLL sketch to the empty state and starts again with the list ${\mathcal{L}}$ of non-discarded items obtained so far.
+Reinserting these items will (eventually) determine whether they hit the original $b$ buckets or later buckets, and hence whether they should be retained in ${\mathcal{L}}$ or discarded.
+Once ${\mathcal{L}}$ is reprocessed (and assuming the cardinality estimate does not exceed $\frac{5}{2}m$), the adversary can start to insert fresh items, as previously described.
 The adversary must be careful not to shuffle the non-discarded items in between iterations, as the $b$ buckets to be filled are blindly determined by the first inserted items, and must stay the same throughout the attack.
 
-%After sufficiently many iterations of this procedure, the adversary will have created a list of items ${\cal L}$ such that, when inserted into the \emph{target} HLL, create a sketch in which $b$ target buckets are filled and the others left empty.
+%After sufficiently many iterations of this procedure, the adversary will have created a list of items ${\mathcal{L}}$ such that, when inserted into the \emph{target} HLL, create a sketch in which $b$ target buckets are filled and the others left empty.
 With an appropriate choice of $b$ (as discussed immediately below), the final cardinality estimate for the target HLL sketch will be $E^*= m \cdot \ln (m/V)$ where $V = m-b$ is the number of empty buckets.
@@
 We will exploit this in the sequel in scenarios S2 and S4.
-For now, we note that setting, say, $b=128$ with $n = 8$ and $\ell = 24$ would result in a final cardinality estimate of $256 \ln 2$ in the above attack in Scenario S1, irrespective of how long the list ${\cal L}$ of inserted items is.
+For now, we note that setting, say, $b=128$ with $n = 8$ and $\ell = 24$ would result in a final cardinality estimate of $256 \ln 2$ in the above attack in Scenario S1, irrespective of how long the list ${\mathcal{L}}$ of inserted items is.
 Moreover, setting $b=128$ would mean that half of all items available to the adversary could be inserted into the target HLL.
@@
 This is highly likely to be the case (with probability at least $1-1/m^2$) as long as at least $m\ln m$ items have already been inserted into the target HLL.
-This follows from the standard analysis of the coupon collector problem\todo{add a reference for this} under our usual assumption that $h$ behaves as a random function.
+This follows from the standard analysis of the coupon collector problem~\cite[Section 3.6]{MR95} under our usual assumption that $h$ behaves as a random function.
 In S4, the adversary can use its snapshot to detect if this is the case or not, and behave differently if not (details for this case follow below).
@@
 
-%We assume the adversary has (or is given) some set of items ${\cal X}$ that it wishes to insert; as usual, we write $|{\cal X}| = 2^t$.
+%We assume the adversary has (or is given) some set of items ${\mathcal{X}}$ that it wishes to insert; as usual, we write $|{\mathcal{X}}| = 2^t$.
 
 \subsubsection{Initially Non-empty HLL Sketch, Scenario S2:}\label{sec:nonemptyHLLS2}
-The adversary knows $h$ so it selects items $x$ from ${\cal X}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit.
+The adversary knows $h$ so it selects items $x$ from ${\mathcal{X}}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit.
 In this case, since the relevant bucket $j$ already has $M[j] \ge 1$, the bucket's counter is not updated, and the cardinality estimate of the HLL sketch remains unchanged.
 
-The adversary then expects to be able to insert half of the items from ${\cal X}$ into the HLL sketch, without increasing the HLL cardinality estimate at all.
+The adversary then expects to be able to insert half of the items from ${\mathcal{X}}$ into the HLL sketch, without increasing the HLL cardinality estimate at all.
 
@@
 \subsubsection{Initially Non-empty HLL Sketch, Scenario S4:}\label{sec:nonemptyHLLS4}
-The adversary can adopt the same strategy as for S2, but make relax the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch).
+The adversary can adopt the same strategy as for S2, but relax the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch).
 Specifically, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
@@
 
-We now compute the expected number of elements $x$ from ${\cal X}$ such that $h(x)$ meets the attack's requirements.  %We assume that $M[j] \ge 0$ for all buckets $j$. 
+We now compute the expected number of elements $x$ from ${\mathcal{X}}$ such that $h(x)$ meets the requirements of the attack.  %We assume that $M[j] \ge 0$ for all buckets $j$. 
 Assume $h(x)$ is mapped to bucket $j$.
 Inserting $x$ does not increase the cardinality estimate if the rightmost $\ell$ bits of $h(x)$ have $M[j]-1$ or fewer leading zeroes; this probability is equal to $1-\frac{1}{2^{M[j]}}$ assuming $h$ behaves like a random function.
-Averaging over all $m$ buckets (and using the fact that the bucket choice is uniformly random since the bits of $h$ are independent and uniformly random), the probability that input $x$ meets the attack's requirements is given by:
+Averaging over all $m$ buckets (and using the fact that the bucket choice is uniformly random since the bits of $h$ are independent and uniformly random), the probability that input $x$ meets the requirements of the attack is given by:
 \begin{eqnarray*}
@@
 where $H_M$ is the harmonic mean of the counts $2^{M[1]}, 2^{M[2]}, \ldots, 2^{M[m]}$.
-So, given ${\cal X}$ containing $2^t$ items, the adversary can expect to insert $(1 - H_M^{-1}) \cdot 2^t$ items from ${\cal X}$ without increasing the HLL cardinality estimate at all.
+So the adversary can expect to insert a fraction $(1 - H_M^{-1})$ of the items from ${\mathcal{X}}$ without increasing the HLL cardinality estimate at all.
 
-%If some buckets are actually empty (and in S4 the adversary knows which ones using its snapshot), then the above analysis can be reused in order to avoid touching those buckets altogether when inserting elements from ${\cal X}$.
+%If some buckets are actually empty (and in S4 the adversary knows which ones using its snapshot), then the above analysis can be reused in order to avoid touching those buckets altogether when inserting elements from ${\mathcal{X}}$.
 The above analysis does not change at all because the term $\frac{1}{m} \cdot (1-\frac{1}{2^{M[j]}})$ already signs a probability of 0 to inserting $x$ in bucket $j$ if $M[j] = 0$.
-Hence we still expect to insert $(1 - H_M^{-1}) \cdot 2^t$ items from ${\cal X}$ without increasing the HLL cardinality estimate at all.
+Hence we still expect to insert $(1 - H_M^{-1}) \cdot 2^t$ items from ${\mathcal{X}}$ without increasing the HLL cardinality estimate at all.
 Of course, the adversary can relax its attack and allow some of the empty buckets to be hit as well.
@@
 
-In this setting, where the target HLL sketch is initially non-empty, we see that an adversary who has sufficient flexibility in its inputs (as represented by the size $2^t$ of the set ${\cal X}$) can expect to insert exponentially many (in $t$) values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the case where the HLL sketch has some empty buckets in S2).
+In this setting, where the target HLL sketch is initially non-empty, we see that an adversary who has sufficient flexibility in its inputs (as represented by the size of the set ${\mathcal{X}}$) can expect to insert many values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the case where the HLL sketch has some empty buckets in S2).
 The result is largely independent of the HLL parameters.
-The more full is the HLL sketch, the larger is $1-H_M^{-1}$ and the easier it is to insert new items without them increasing the cardinality estimate.
+The more full is the HLL sketch, the larger is $1-H_M^{-1}$ and the easier it is to insert new items without increasing the cardinality estimate.
 
@@
 
-We briefly sketch an attack in this setting in Appendix~\ref{sec:non-empty-HLL-sketch-S1}.
+We briefly sketch an attack in the setting of an S1 adversary attacking an initially non-empty HLL sketch.
+An alternative attack with moderate performance in this setting is presented in~\cite{hllvuln}.
+
+Since the target sketch has an unknown status for the adversary, our previous strategy for S1 of keeping many buckets empty no longer works.
+One can attempt to use the shadow device to identify items that would result in small entries $M[j]$ (for unknown indices $j$) since such items are unlikely to increase the cardinality estimate of the target device by much.
+This can be done by first filling the shadow device with sufficiently many values to escape from the ``small range correction'' phase.
+Then one could enter candidate values $x$ into the shadow device and check for cardinality increases.
+However this runs into the rounding issue described in Section~\ref{sec:scenarios-discussion}.
+Still, one could retain those items which do lead to a detectable change in cardinality and reject the others (and also estimate the value of $M[j]$ for the unknown bucket $j$ from the magnitude of the change, retaining only those for which $M[j]$ is small).
+But here the lack of a cardinality change could result either from rounding issues or because insertion of candidate value $x$ into the shadow device really did lead to no increase in cardinality.
+This indicates that  ``good'' items would be rejected unnecessarily.
+This could be addressed by using multiple resets of the shadow HLL sketch, filling it with random items after each reset, and performing multiple insertions per candidate item $x$ to try to estimate the true value of $M[j]$ for each item $x$ (since after at least some of the resets followed by random insertions, we could expect $M[j]$ to increase when $x$ is inserted, and if $M[j]$ is small enough relative to the cardinality estimate, the change could be detected).
+The details of such an attack require further analysis which we do not pursue here, since we consider S1 to be the least realistic attack scenario.
 
@@
 %We first make an empty shadow HLL sketch via a reset.
-Consider an adversary that simply inserts $m \ln m$ random items $x$ from ${\cal X}$ into the shadow and target HLL sketches.
+Consider an adversary that simply inserts $m \ln m$ random items $x$ from ${\mathcal{X}}$ into the shadow and target HLL sketches.
 This increases the target HLL sketch's cardinality estimate by about $m \ln m$ since the HLL algorithm is a good estimator.
@@
 Now the adversary can carry out a version of our previous attack in scenario S1 (where the target HLL sketch is initially empty).
-By repeatedly processing lists of inputs, it can find a large subset ${\cal L}^*$ of ${\cal X}$ which does not cause the cardinality estimate of the shadow HLL sketch to increase beyond $m \ln m$.
+By repeatedly processing lists of inputs, it can find a large subset ${\mathcal{L}}^*$ of ${\mathcal{X}}$ which does not cause the cardinality estimate of the shadow HLL sketch to increase beyond $m \ln m$.
 This assumes that single updates to the vector of counter values $(M[j])_{j=1}^m$ in the shadow HLL sketch can be detected through changes to the HLL cardinality estimate.
 This seems plausible but requires further experimental investigation.
-Inserting all the items from ${\cal L}^*$ into the target HLL sketch then leads to an increase in its cardinality estimate of at most $m \ln m$.
-It is not hard to see that, since every bucket in the shadow HLL sketch is active, we can expect ${\cal L}^*$ to contain at least half of the items of ${\cal X}$.
+Inserting all the items from ${\mathcal{L}}^*$ into the target HLL sketch then leads to an increase in its cardinality estimate of at most $m \ln m$.
+It is not hard to see that, since every bucket in the shadow HLL sketch is active, we can expect ${\mathcal{L}}^*$ to contain at least half of the items of ${\mathcal{X}}$.
 
-%An alternative approach is to simply run the scenario S1 attack as if the target HLL sketch were empty (i.e.\ as described in Section~\ref{sec:empty} above), take the resulting list ${\cal L}^*$, and insert it into the target HLL sketch, hoping that the cardinality does not increase too much.
+%An alternative approach is to simply run the scenario S1 attack as if the target HLL sketch were empty (i.e.\ as described in Section~\ref{sec:empty} above), take the resulting list ${\mathcal{L}}^*$, and insert it into the target HLL sketch, hoping that the cardinality does not increase too much.
 %We report on this approach experimentally in Section~\ref{sec:exp}.
@@
 Redis 5.0 onwards uses exactly the formula in equation~(\ref{eqn:ertl}) to form its cardinality estimate.
-It uses $n=14$, $m=16,384$ and $\ell = 50$ (so $n+\ell = 64$ and a 64-bit hash function is used).
+It uses $n=14$, $m=16,384$ and $\ell = 50$ (so $n+\ell = 64$ and a 64-bit unkeyed hash function is used).
 
@@
 
-Attacks are now straightforward.
+Attacks against the Redis implementation are now straightforward.
 Suppose we can arrange that $C_0$, the number of empty buckets, stays above ${\epsilon}m$, a linear function of the total number of buckets, and that $C_{\ell+1} = 0$.
@@
 
-The requirement on $C_0 \geq {\epsilon}m$ is easily met in scenarios S2 and S4 with an initially empty HLL sketch, through the trick of sampling inputs $x \in {\cal X}$ so that only a fraction $1-\epsilon$ of the buckets are ever hit.
-We can expect to insert a fraction $1-\epsilon-\epsilon2^{-\ell}$ of the elements of  ${\cal X}$ whilst keeping the cardinality estimate $\hat{\lambda}$ below  $\alpha_\infty \epsilon^{-1} m$.
+The requirement on $C_0 \geq {\epsilon}m$ is easily met in scenarios S2 and S4 with an initially empty HLL sketch, through the trick of sampling inputs $x \in {\mathcal{X}}$ so that only a fraction $1-\epsilon$ of the buckets are ever hit.
+We can expect to insert a fraction $1-\epsilon-\epsilon2^{-\ell}$ of the elements of  ${\mathcal{X}}$ whilst keeping the cardinality estimate $\hat{\lambda}$ below  $\alpha_\infty \epsilon^{-1} m$.
 (The term $\epsilon2^{-\ell}$ arises from the requirement to keep $C_{\ell+1} =0$ in the attack; recall that $\ell=50$ in Redis, so the effect of this term is very small.)
@@
 These insertions do not alter the values of the $C_k$, and hence $\hat{\lambda}$ is unaffected.
-In S2, we expect to insert half the items in ${\cal X}$ without increasing the cardinality estimate $\hat{\lambda}$ at all.
+In S2, we expect to insert half the items in ${\mathcal{X}}$ without increasing the cardinality estimate $\hat{\lambda}$ at all.
 In S4, the analysis is the same as in Section~\ref{sec:nonemptyHLLS4}.
@@
 This completes our analysis of Redis.
-The key takeaway is that, in scenarios S2 and S4, an adversary can expect to insert a constant fraction of ${\cal X}$ into a non-empty Redis HLL sketch whilst not increasing its cardinality estimate at all (and whilst keeping the estimated cardinality linear in $m$ in the case where the sketch is initially empty).
+The key takeaway is that, in scenarios S2 and S4, an adversary can expect to insert a constant fraction of ${\mathcal{X}}$ into a non-empty Redis HLL sketch whilst not increasing its cardinality estimate at all (and whilst keeping the estimated cardinality linear in $m$ in the case where the sketch is initially empty).
 
@@
 Thus they are  vulnerable to our attacks.
-\todo{maybe switch to one of these, or somehow be more clear about how the attacks transfer.}
+%\todo[inline]{maybe switch to one of these, or somehow be more clear about how the attacks transfer.}
 
@@
 %\todo{We need to make sure the repo is publicly accessible but anonymised for submission, if we are going to include it.}
-For simplicity, we take ${\cal X}$ to be the set of 4-character long ASCII strings, with $\lvert {\cal X} \rvert = 7,311,616= 2^{22.8}$.
-\todo{Explain there's nothing special about these inputs; anything would do; processed by hash function anyway.}We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}.
+For simplicity, we take ${\mathcal{X}}$ to be the set of 4-character long ASCII strings, with $\lvert {\mathcal{X}} \rvert = 7,311,616= 2^{22.8}$.
+We stress that this choice of inputs is merely illustrative and the results do not in any way depend on choosing ASCII strings.
+We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}.
 %(note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
@@
 For simplicity and to keep the attack as striking as possible, we set $B$ to 1 in the attacks on an initially empty HLL sketch in scenarios S2 and S4.
-All attacks were run 30 times; we report average outcomes.
+All attacks were run 30 times and we report average outcomes.
 Our results are reported in Tables~\ref{table:tab1} and~\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary, and the cardinality estimate after adding the adversary's items.
@@
 Also as discussed in Section~\ref{sec:redis}, the Redis implementation has changed its HLL algorithm several times in recent releases, but the authors of~\cite{hllvuln} do not specify which version was used in their experiments.
-This makes it difficult to make a precise comparison with their results.
+This makes it difficult to make a precise comparison with the results of~\cite{hllvuln}.
 Nevertheless, the S1 adversary following our strategy against the classical HLL algorithm is able to add 977 and 97,433 distinct items while holding the cardinality estimate at 1 and 126, respectively.
@@
 This comes at the cost of performing more resets compared to~\cite{hllvuln}.
-\todo{explain why counting resets matter; also that we don't care too much about this scenario.}
+We recall that in practice each reset may cause a delay (for example, it may involve rebooting a network device); we also recall that we consider S1 to be a somewhat artificial attack scenario.
 
@@
 In the remainder of this section, we introduce a formal attack model covering all three scenarios.
-This is done by allowing an adversary ${\cal A}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios.
-In the security model, ${\cal A}$ interacts in one of two worlds.
-In the real world, ${\cal A}$ interacts with a real HLL, operating as in Figure~\ref{fig:hll}.
-In the ideal world, ${\cal A}$ interacts with an ideal HLL in which the operation $y: = h(x)$ is replaced by random sampling; that is, by setting $y \leftarrow_{\$} \{0,1\}^{n+\ell}$.
+This is done by allowing an adversary ${\mathcal{A}}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios.
+In the security model, ${\mathcal{A}}$ interacts in one of two worlds.
+In the real world, ${\mathcal{A}}$ interacts with a real HLL, operating as in Figure~\ref{fig:hll}.
+In the ideal world, ${\mathcal{A}}$ interacts with an ideal HLL in which the operation $y: = h(x)$ is replaced by random sampling; that is, by setting $y \leftarrow_{\$} \{0,1\}^{n+\ell}$.
 
@@
 
-We then show that, if $h$ is replaced by a keyed Pseudo Random Function (PRF) $F$ whose key is not available to the adversary, then any adversary ${\cal A}$ that can distinguish between the real and random worlds with oracles as in S4 can be used to build an adversary ${\cal B}$ which breaks the PRF security of $F$.
-Under the assumption that $F$ is a good PRF, no such efficient adversary ${\cal B}$ can exist, and we may conclude that we can safely switch from the real world to the ideal world without incurring any significant change to the behaviour of ${\cal A}$.
+We then show that, if $h$ is replaced by a keyed Pseudo Random Function (PRF) $F$ whose key is not available to the adversary, then any adversary ${\mathcal{A}}$ that can distinguish between the real and random worlds with oracles as in S4 can be used to build an adversary ${{\mathcal{B}}}$ which breaks the PRF security of $F$.
+Under the assumption that $F$ is a good PRF, no such efficient adversary ${{\mathcal{B}}}$ can exist, and we may conclude that we can safely switch from the real world to the ideal world without incurring any significant change to the behaviour of ${\mathcal{A}}$.
 
 However, it is clear that any adversary in the ideal world is reduced to operating in the setting used in the average case analysis of HLL (wherein every insertion query results in a random value $y$ being used to update the HLL sketch).
-Thus, ${\cal A}$ can have no advantage when it comes to influencing the HLL cardinality estimate.
+Thus, ${\mathcal{A}}$ can have no advantage when it comes to influencing the HLL cardinality estimate.
 It follows that HLL, when instantiated with PRF $F$, is secure against attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
@@
 
-We consider adversary ${\cal A}$ to be an algorithm running against a challenger ${\cal C}$ in a security game.
-The game begins with ${\cal C}$ selecting a bit $b$ uniformly at random.
-Adversary ${\cal A}$ interacts with up to four oracles which are handled by ${\cal C}$; different combinations of oracles define scenarios S2, S3, S4.
-After making some queries, ${\cal A}$ returns a bit $b'$ estimating whether it is interacting with a real HLL instantiation ($b'=0$) or an ideal one ($b'=1$).
-For $i=1, 2, 3$, we define ${\cal A}$'s S$i$-Real-or-Ideal advantage against HLL to be:
+We consider adversary ${\mathcal{A}}$ to be an algorithm running against a challenger ${{\mathcal{C}}}$ in a security game.
+The game begins with ${{\mathcal{C}}}$ selecting a bit $b$ uniformly at random.
+Adversary ${\mathcal{A}}$ interacts with up to four oracles which are handled by ${{\mathcal{C}}}$; different combinations of oracles define scenarios S2, S3, S4.
+After making some queries, ${\mathcal{A}}$ returns a bit $b'$ estimating whether it is interacting with a real HLL instantiation ($b'=0$) or an ideal one ($b'=1$).
+For $i=1, 2, 3$, we define ${\mathcal{A}}$'s S$i$-Real-or-Ideal advantage against HLL to be:
 \[
-\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny HLL}}({\cal A}) := \lvert 2\Pr[b' = b] - 1 \rvert \, .
+\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny HLL}}({\mathcal{A}}) := \lvert 2\Pr[b' = b] - 1 \rvert \, .
 \]
 %Using a standard transformation, this probability can be rewritten as $\lvert \Pr[b' = 1| b = 1] - \Pr[b'=0|b=1] \rvert$. 
-The four oracles that  ${\cal A}$ can interact with are as follows:
+The four oracles that  ${\mathcal{A}}$ can interact with are as follows:
 \begin{itemize}
-\item \texttt{init}: this oracle performs initialisation of the HLL sketch (and selection of the PRF key when $h$ is replaced by a PRF $F$); this oracle is available in all of S2, S3, S4, but can only be called once as ${\cal A}$'s first oracle query.
-\item \texttt{insert($\cdot$)}: on input $x \in {\cal D}$, the oracle inserts value $x$ into the HLL sketch.
+\item \texttt{init}: this oracle performs initialisation of the HLL sketch (and selection of the PRF key when $h$ is replaced by a PRF $F$); this oracle is available in all of S2, S3, S4, but can only be called once as ${\mathcal{A}}$'s first oracle query.
+\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$, the oracle inserts value $x$ into the HLL sketch.
 This oracle is available in all of S2, S3, S4.
-Without loss of generality (since HLL and variants of it we consider here have a deterministic insertion procedure), we assume all queries $x$ made by ${\cal A}$ to \texttt{insert($\cdot$)} are distinct.
-We let $q$ denote the number of insertion queries made by ${\cal A}$ during the course of its attack.
-\item \texttt{read}: this oracle returns the HLL cardinality estimate. We allow multiple queries to this oracle and make it available to ${\cal A}$ in S3 and S4, but not S2.
+Without loss of generality (since HLL and variants of it we consider here have a deterministic insertion procedure), we assume all queries $x$ made by ${\mathcal{A}}$ to \texttt{insert($\cdot$)} are distinct.
+We let $q$ denote the number of insertion queries made by ${\mathcal{A}}$ during the course of its attack.
+\item \texttt{read}: this oracle returns the HLL cardinality estimate. We allow multiple queries to this oracle and make it available to ${\mathcal{A}}$ in S3 and S4, but not S2.
 \item \texttt{corrupt}: this oracle returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents (but not any secret key $K$).
@@
 
-We now replace the hash function $h$ acting on ${\cal D}$ from the HLL algorithm with a keyed function $F: \{0,1\}^k \times {\cal D} \rightarrow  \{0,1\}^{n+\ell}$.
+We now replace the hash function $h$ acting on ${{\mathcal{D}}}$ from the HLL algorithm with a keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
 Here the first input to $F$ is a $k$-bit string, the key, denoted $K$.
@@
 
-We define the PRF advantage of an adversary ${\cal B}$ against $F$ outputting a bit $d'$ as being its advantage in distinguishing whether it is interacting with $F$ for a uniformly random key $K$ ($d=0$) or interacting with a truly random function ($d=1$).
+We define the PRF advantage of an adversary ${{\mathcal{B}}}$ against $F$ outputting a bit $d'$ as being its advantage in distinguishing whether it is interacting with $F$ for a uniformly random key $K$ ($d=0$) or interacting with a truly random function ($d=1$).
 Thus:
 \[
-\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B}) := \lvert 2\Pr[d' = d] - 1 \rvert\,.
+\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}}) := \lvert 2\Pr[d' = d] - 1 \rvert\,.
 \]
-Here, we assume without loss of generality that all of ${\cal B}$'s queries are distinct. See~\cite[Section 4.4.1]{BonehShoup} for details of the PRF security definition.
+Here, we assume without loss of generality that all of ${{\mathcal{B}}}$'s queries are distinct. See~\cite[Section 4.4.1]{BonehShoup} for details of the PRF security definition.
 
@@
 \begin{theorem}\label{thm:security-of-HLL-with-PRF}
-Suppose ${\cal A}$ is an S4 adversary against HLL instantiated with a keyed function $F: \{0,1\}^k \times {\cal D} \rightarrow  \{0,1\}^{n+\ell}$.
-Then we can construct from ${\cal A}$ a PRF adversary ${\cal B}$ against $F$ such that:
+Suppose ${\mathcal{A}}$ is an S4 adversary against HLL instantiated with a keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
+Then we can construct from ${\mathcal{A}}$ a PRF adversary ${{\mathcal{B}}}$ against $F$ such that:
 \[
- \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B}) = \mbox{Adv}^{\mbox{\tiny S4-RoI}}_{\mbox{\tiny HLL}}({\cal A}) \,.
+ \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}}) = \mbox{Adv}^{\mbox{\tiny S4-RoI}}_{\mbox{\tiny HLL}}({\mathcal{A}}) \,.
 \]
-Moreover, if ${\cal A}$ makes $q$ queries to its \texttt{insert($\cdot$)} oracle, then ${\cal B}$ makes $q$ queries to its real or random oracle.
-Finally, the running time of ${\cal B}$ is (roughly) the same as that of ${\cal A}$.
+Moreover, if ${\mathcal{A}}$ makes $q$ queries to its \texttt{insert($\cdot$)} oracle, then ${{\mathcal{B}}}$ makes $q$ queries to its real or random oracle.
+Finally, the running time of ${{\mathcal{B}}}$ is (roughly) the same as that of ${\mathcal{A}}$.
 \end{theorem}
 
-Informally, if $F$ is a good PRF, then $\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B})$ should be small for any adversary ${\cal B}$ running in a reasonable amount of time; hence the advantage of any S4 adversary  ${\cal A}$ against HLL using PRF $F$ is also small.
-The proof can be found in Appendix~\ref{sec:proof}.
-Our proof still applies for the improved algorithm of~\cite{hllnew} that is used in Redis.
+Informally, if $F$ is a good PRF, then $\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}})$ should be small for any adversary ${{\mathcal{B}}}$ running in a reasonable amount of time; hence the advantage of any S4 adversary  ${\mathcal{A}}$ against HLL using PRF $F$ is also small.
+Note that our proof still applies for the improved algorithm of~\cite{hllnew} that is used in Redis.
 The only changes to the proof are to use the Redis approach for computing responses to \texttt{read} queries in place of the classical HLL approach.
 
-We conclude by reiterating our central point: we have shown that no efficient S4 adversary ${\cal A}$ can tell that it is playing against HLL implemented using truly random values $y$ instead of with values obtained from a PRF, except with a small advantage.
-This means that ${\cal A}$ is effectively reduced to attacking HLL in the exact same setting used in the average case analysis of HLL from~\cite{hll}, wherein it is proved that the HLL cardinality estimate endures relative errors of the form $1.04/\sqrt{m}$.
-It does not matter that ${\cal A}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries.
+\begin{proof}
+PRF adversary ${{\mathcal{B}}}$ against $F$ is built as follows.
+Let $d$ denote the hidden bit determining whether ${{\mathcal{B}}}$ receives real responses of the form $F(K,x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${{\mathcal{B}}}$ runs S4 adversary ${\mathcal{A}}$ and responds to each of its queries as follows:
+\begin{itemize}
+\item \texttt{init}: here ${{\mathcal{B}}}$ initialises an HLL sketch in accordance with Figure~\ref{fig:hll}. Thus it initialises an array $M[j]$ in which each entry is set to 0. 
+\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$,  ${{\mathcal{B}}}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${{\mathcal{B}}}$ uses $y$ to update the HLL bucket $j$ generated via $j=1 + <y_1y_2...y_{n}>_2$ as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
+\item \texttt{read}: ${{\mathcal{B}}}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate in accordance with Figure~\ref{fig:hll}. 
+\item \texttt{corrupt}: ${{\mathcal{B}}}$ returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents. 
+\end{itemize}
+When ${\mathcal{A}}$ outputs a bit $b'$, ${{\mathcal{B}}}$ outputs its own bit $d'=b'$.
+Note that when $d=0$, ${{\mathcal{B}}}$ plays against the real $F$ with a random key $K$ and provides a simulation for ${\mathcal{A}}$ that perfectly instantiates the real world.
+On the other hand, when $d=1$, ${{\mathcal{B}}}$ plays against a truly random function and provides a simulation for ${\mathcal{A}}$ that perfectly instantiates the ideal world.
+The equality of advantages in the statement of the theorem now follows immediately.
+The remainder of the proof follows on counting queries and noting that the HLL algorithm for computing cardinality is efficient, so ${{\mathcal{B}}}$'s running time due to handling \texttt{read} queries is not increased significantly compared to that of ${\mathcal{A}}$.
+\end{proof}
+
+We conclude by reiterating our central point: we have shown that no efficient S4 adversary ${\mathcal{A}}$ can tell that it is playing against HLL implemented using truly random values $y$ instead of with values obtained from a PRF, except with a small advantage.
+This means that ${\mathcal{A}}$ is effectively reduced to attacking HLL in the exact same setting used in the average case analysis of HLL from~\cite{hll}, wherein it is proved that the HLL cardinality estimate endures relative errors of the form $1.04/\sqrt{m}$.
+It does not matter that ${\mathcal{A}}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries.
 For each \texttt{insert($\cdot$)} query, a fresh value $x$ leads to a new random $y$ being generated and used to update a bucket, as in the average case analysis of HLL.
@@
 
-The above construction requires a PRF $F: \{0,1\}^k \times {\cal D} \rightarrow  \{0,1\}^{n+\ell}$.
-Here, the set ${\cal D}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
+The above construction requires a PRF $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
+Here, the set ${{\mathcal{D}}}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
 This means that $F$ needs to be Variable Input Length PRF (VIL-PRF).
@@
 
-%A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {\cal D} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
+%A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {{\mathcal{D}}} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
 The result is a PRF with the required input-output signature~\cite[Section 8.2]{BonehShoup}.
@@
 
-\appendix
-
-\section{Initially Non-empty HLL Sketch, Scenario S1}\label{sec:non-empty-HLL-sketch-S1}
-
-We briefly sketch an attack in the setting of an S1 adversary attacking an initially non-empty HLL sketch.
-An alternative attack with moderate performance in this setting is presented in~\cite{hllvuln}.
-
-Since the target sketch has an unknown status for the adversary, our previous strategy for S1 of keeping many buckets empty no longer works.
-One can attempt to use the shadow device to identify items that would result in small entries $M[j]$ (for unknown indices $j$) since such items are unlikely to increase the cardinality estimate of the target device by much.
-This can be done by first filling the shadow device with sufficiently many values to escape from the ``small range correction'' phase.
-Then one could enter candidate values $x$ into the shadow device and check for cardinality increases.
-However this runs into the rounding issue described in Section~\ref{sec:scenarios-discussion}.
-Still, one could retain those items which do lead to a detectable change in cardinality and reject the others (and also estimate the value of $M[j]$ for the unknown bucket $j$ from the magnitude of the change, retaining only those for which $M[j]$ is small).
-But here the lack of a cardinality change could result either from rounding issues or because insertion of candidate value $x$ into the shadow device really did lead to no increase in cardinality.
-This indicates that  ``good'' items would be rejected unnecessarily.
-This could be addressed by using multiple resets of the shadow HLL sketch, filling it with random items after each reset, and performing multiple insertions per candidate item $x$ to try to estimate the true value of $M[j]$ for each item $x$ (since after at least some of the resets followed by random insertions, we could expect $M[j]$ to increase when $x$ is inserted, and if $M[j]$ is small enough relative to the cardinality estimate, the change could be detected).
-The details of such an attack require further analysis which we do not pursue here, since we consider S1 to be the least realistic attack scenario.
-
-\section{Proof of Theorem~\ref{thm:security-of-HLL-with-PRF}}\label{sec:proof}
-
-\begin{proof}
-PRF adversary ${\cal B}$ against $F$ is built as follows.
-Let $d$ denote the hidden bit determining whether ${\cal B}$ receives real responses of the form $F(K,x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${\cal B}$ runs S4 adversary ${\cal A}$ and responds to each of its queries as follows:
-\begin{itemize}
-\item \texttt{init}: here ${\cal B}$ initialises an HLL sketch in accordance with Figure~\ref{fig:hll}. Thus it initialises an array $M[j]$ in which each entry is set to 0. 
-\item \texttt{insert($\cdot$)}: on input $x \in {\cal D}$,  ${\cal B}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${\cal B}$ uses $y$ to update the HLL bucket $j$ generated via $j=1 + <y_1y_2...y_{n}>_2$ as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
-\item \texttt{read}: ${\cal B}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate in accordance with Figure~\ref{fig:hll}. 
-\item \texttt{corrupt}: ${\cal B}$ returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents. 
-\end{itemize}
-When ${\cal A}$ outputs a bit $b'$, ${\cal B}$ outputs its own bit $d'=b'$.
-Note that when $d=0$, ${\cal B}$ plays against the real $F$ with a random key $K$ and provides a simulation for ${\cal A}$ that perfectly instantiates the real world.
-On the other hand, when $d=1$, ${\cal B}$ plays against a truly random function and provides a simulation for ${\cal A}$ that perfectly instantiates the ideal world.
-The equality of advantages in the statement of the theorem now follows immediately.
-The remainder of the proof follows on counting queries and noting that the HLL algorithm for computing cardinality is efficient, so ${\cal B}$'s running time due to handling \texttt{read} queries is not increased significantly compared to that of ${\cal A}$.
-\end{proof}
-
 \end{document}

### 2021-04-23 docs/HLL_exp_bad/main.tex
work on paper

@@
 
-Very recently, Reviriego and Ting~\cite{hllvuln} initiated work on the performance of HLL in adversarial settings.
+Recently, Reviriego and Ting~\cite{hllvuln} initiated work on the performance of HLL in adversarial settings.
 They showed that in very specific attack settings, the cardinality estimate made by HLL could be modified through input selection.
@@
 We discuss and benchmark a suitable choice for the required VIL-PRF.
-\todo{make the model a contribution...
-``To achieve this...
-introduce simulation-based security model for cardinality estimators'', etc}
-
-\todo{Simplicity of attacks as a strength; only complexity comes in S1 attack which we don't consider that interesting!}
-\todo{No e2e attack on a specific system; rather paper is written in general form since the attacks apply broadly.}
+
+\todo[inline]{make the model a contribution... ``To achieve this... introduce simulation-based security model for cardinality estimators'', etc}
+\todo[inline]{Simplicity of attacks as a strength; only complexity comes in S1 attack which we don't consider that interesting!}
+\todo[inline]{No e2e attack on a specific system; rather paper is written in general form since the attacks apply broadly.}
+\todo[inline]{Highlight the analysis of OSDI paper~\cite{flexswitch} as an example of the potential of the attacks.}
 
@@
 
-\todo{Add a couple of paragraphs to relate all this back to the initial attack scenarios, e.g.
-packet flows.
-May need to look at internals of some of those papers to see where flex in inputs will come from.
-}
+\todo[inline]{Results from Redis or GBQ if we get them.}
+
+\subsection{Relating the Experimental Results to Applications}\label{sec:exp-apps}
+
+So far, we have treated attacks in an abstract way, working with an off-the-shelf implementation of HLL.
+We now illustrate what effect such attacks might have on applications relying on HLL for cardinality estimation.
+In an effort to be as specific a possible, we analyse in detail a recent paper~\cite{flexswitch} from the networking systems community.
+
+The authors of~\cite{flexswitch} propose to use HLL for approximating traffic flow statistics in datacenter network switches.
+In the main example considered in~\cite[Section 3.1]{flexswitch}, each flow is identified by a 4-tuple consisting of source IP address, destination IP address, source port, and destination port.
+Various parameters for the HLL sketch to be used are then explored in~\cite{flexswitch}, trading memory consumed for accuracy of estimation of the number of concurrent flows.
+The paper also proposes to use the low range correction method of Fig.~\ref{fig:hll}.
+At no point does the discussion in~\cite{flexswitch} consider adversarial input (it does mention worst-case performance but only in passing, and no indication is given of how to handle this).
+It is clear that, by modifying the 16-bit source port (for example), an adversary could fool the switch into undercounting the flows it is responsible for producing.
+Depending on how the flow count is used, this could lead to undetected DoS conditions, mis-provisioning of network capacity, or under-billing of the adversary for its traffic.
+
+More concretely,~\cite[Section 4.2]{flexswitch} proposes using HLL for detecting TCP port scans by counting observed destination port numbers in TCP packets with set SYN flags and reporting a scan event if the estimate exceeds a
+set threshold.
+Using our attack techniques, an attacker could complete a partial port scan of roughly half of all $2^{16}$ ports whilst keeping the HLL estimate below $0.693m$ where $m$ is the number of buckets used in HLL.
+While~\cite{flexswitch} is not specific about the value of $m$ actually used, the authors' focus on using low memory indicates that $m$ could be $1024$.\footnote{The paper mentions using 1 byte counters in the buckets and states that using more than 1KB of memory ``only improves average error rates marginally''; we infer a choice of $m=1024$ from this.} It is not stated in~\cite{flexswitch} what threshold needs to be crossed to trigger an alarm, but we consider it plausible that it would be above $0.693 \cdot 1024 = 710$.
+
+Similarly,~\cite[Section 4.2]{flexswitch} proposes using HLL for detecting NTP amplification attacks by counting distinct source IP addresses arriving on UDP port 123 over a small time window.
+No further details are given, but it is clear that an attacker with even limited flexibility in spoofing source IP addresses (e.g.\ within a private range used in the datacenter) could easily evade detection whilst delivering a large number of NTP requests.
+
+In a more advanced application,~\cite[Section 4.3]{flexswitch} proposes to use HLL as a component in a novel version of the RCP congestion control algorithm called RCP-Approx.
+The idea of RCP-Approx is to replace exact counting with approximate counting to save on consumed memory in switches.
+Again HLL is used for estimating numbers of traffic flows.
+Obviously, our attack techniques could be used (in scenario S2) to cause RCP-Approx to underestimate the number of traffic flows, leading to non-detection of congestion and so enabling DoS attacks to go undetected.
+
+Similar remarks apply to the treatment of CONGA in~\cite[Section 4.5]{flexswitch}.
+CONGA is a congestion-aware load balancing system that splits TCP flows into flowlets (collections of IP packets) and allocates them to paths based on congestion feedback from remote switches.
+In the enhancements to CONGA proposed in~\cite{flexswitch}, HLL is used in the ``flow statistics'' building block for estimating flowlet statistics in switches; these estimates are sent back to a monitoring system and used to determine the load balancing configuration.
+This overall system can be abused by an S2 adversary, to cause flowlet numbers to be underestimated (or overestimated) and the load balancing to perform sub-optimally.
+For example, an adversary could cause all the switches to \emph{over}-estimate the number of flowlets they are handling and ask the load balancing system to send less traffic through them.
+This could result in a degradation in the total amount of traffic the load balancing system can handle (depending on what fail-safe mode it has).
+Or an attacker could cause all the switches to \emph{under}-estimate the number of flowlets they are handling, causing undetected congestion in the switches.
+
+\todo[inline]{This ends up being pretty harsh on this NSDI paper, but hey.}
 

### 2021-04-24 docs/HLL_exp_bad/main.tex
new subsection to make impact on some system more clear
mods to proof

@@
 We show in Section~\ref{sec:formal} that when a salt is used appropriately (i.e.\ as the secret key for a pseudo-random function that is used in place of hash function $h$), security against adversarial inputs can be achieved even in our strongest adversarial setting.
-In addition, our simulation-based security definition, in which the simulator in the ideal world has no access to the inserted values, automatically guarantees confidentiality of the values inserted into the HLL sketch, even in the strongest attack model considered in~\cite{cardestprivacy}.
+%In addition, our simulation-based security definition, in which the simulator in the ideal world has no access to the inserted values, automatically guarantees confidentiality of the values inserted into the HLL sketch, even in the strongest attack model considered in~\cite{cardestprivacy}.
 So, while our main objective is to protect against adversarial inputs, we also get privacy for free when using a keyed PRF in place of the HLL hash function.
@@
 
-\subsection{Relating the Experimental Results to Applications}\label{sec:exp-apps}
+\subsection{Relating the Attacks to Applications}\label{sec:exp-apps}
 
@@
 We now illustrate what effect such attacks might have on applications relying on HLL for cardinality estimation.
-In an effort to be as specific a possible, we analyse in detail a recent paper~\cite{flexswitch} from the networking systems community.
-
-The authors of~\cite{flexswitch} propose to use HLL for approximating traffic flow statistics in datacenter network switches.
+In an effort to be more concrete, we analyse in detail a recent paper~\cite{flexswitch} from the networking systems community.
+
+The authors of~\cite{flexswitch} propose to use HLL for recording traffic flow statistics in datacenter network switches.
+The idea is to use HLL so as to reduce the amount of memory needed in the switches.
 In the main example considered in~\cite[Section 3.1]{flexswitch}, each flow is identified by a 4-tuple consisting of source IP address, destination IP address, source port, and destination port.
@@
 At no point does the discussion in~\cite{flexswitch} consider adversarial input (it does mention worst-case performance but only in passing, and no indication is given of how to handle this).
-It is clear that, by modifying the 16-bit source port (for example), an adversary could fool the switch into undercounting the flows it is responsible for producing.
+It is clear that, by modifying the 16-bit source port (for example), an adversary could fool the switch into undercounting the flows the adversary is responsible for producing.
 Depending on how the flow count is used, this could lead to undetected DoS conditions, mis-provisioning of network capacity, or under-billing of the adversary for its traffic.
 
-More concretely,~\cite[Section 4.2]{flexswitch} proposes using HLL for detecting TCP port scans by counting observed destination port numbers in TCP packets with set SYN flags and reporting a scan event if the estimate exceeds a
+Going further, \cite[Section 4.2]{flexswitch} proposes using HLL for detecting TCP port scans by counting observed destination port numbers in TCP packets with set SYN flags and reporting a scan event if the estimate exceeds a
 set threshold.
 Using our attack techniques, an attacker could complete a partial port scan of roughly half of all $2^{16}$ ports whilst keeping the HLL estimate below $0.693m$ where $m$ is the number of buckets used in HLL.
-While~\cite{flexswitch} is not specific about the value of $m$ actually used, the authors' focus on using low memory indicates that $m$ could be $1024$.\footnote{The paper mentions using 1 byte counters in the buckets and states that using more than 1KB of memory ``only improves average error rates marginally''; we infer a choice of $m=1024$ from this.} It is not stated in~\cite{flexswitch} what threshold needs to be crossed to trigger an alarm, but we consider it plausible that it would be above $0.693 \cdot 1024 = 710$.
+While~\cite{flexswitch} is not specific about the value of $m$ actually used, the authors' focus on reducing memory usage indicates that $m$ would not be larger than $1024$.\footnote{The paper mentions using 1 byte counters in the buckets and states that using more than 1KB of memory ``only improves average error rates marginally''; we infer a maximum value of $m=1024$ from this.} It is not stated in~\cite{flexswitch} what threshold needs to be crossed to trigger an alarm, but we consider it plausible that it would be above $0.693 \cdot 1024 = 710$.
 
 Similarly,~\cite[Section 4.2]{flexswitch} proposes using HLL for detecting NTP amplification attacks by counting distinct source IP addresses arriving on UDP port 123 over a small time window.
-No further details are given, but it is clear that an attacker with even limited flexibility in spoofing source IP addresses (e.g.\ within a private range used in the datacenter) could easily evade detection whilst delivering a large number of NTP requests.
-
-In a more advanced application,~\cite[Section 4.3]{flexswitch} proposes to use HLL as a component in a novel version of the RCP congestion control algorithm called RCP-Approx.
-The idea of RCP-Approx is to replace exact counting with approximate counting to save on consumed memory in switches.
+No further details are given, but it is clear that an attacker with even limited flexibility in spoofing source IP addresses (e.g.\ within a private range used in the datacenter) could easily evade detection whilst delivering a large number of NTP requests, thus re-enabling the original attack.
+
+In a more advanced application,~\cite[Section 4.3]{flexswitch} proposes to use HLL as a component in a novel, low-memory version of the RCP congestion control algorithm called RCP-Approx.
 Again HLL is used for estimating numbers of traffic flows.
@@
 
-In this section we present a formal security analysis of HLL, focussing on whether we can augment the original HLL design with cryptographic mechanisms and thereby provably prevent attacks.
+In this section we present a formal security analysis of HLL, focussing on how to augment the original HLL design with cryptographic mechanisms and thereby provably prevent attacks.
 
@@
 
-\subsection{Formal Security Analysis for S2, S3, S4}
+\subsection{Intuition for Formal Security Analysis for S2, S3, S4}
 
 For scenarios S2, S3 and S4, the picture is much more positive.
-In the remainder of this section, we introduce a formal attack model covering all three scenarios.
-This is done by allowing an adversary ${\mathcal{A}}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios.
+In the remainder of this section, we introduce a formal security model covering all three scenarios.
+We work with a simulation-based framework.
+This is in contrast to the game-based approach introduced for probabilistic data structures in~\cite{CCS:ClaPatShrPS19}.
+We discuss the strengths of our alternative approach in more detail below.
+
+We first begin by explaining the high-level idea of our approach.
+We allow an adversary ${\mathcal{A}}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios S2, S3 and S4: it may have an \texttt{insert($\cdot$)} oracle allowing it to add values to the HLL, a \texttt{read} oracle that returns the HLL cardinality estimate, and a \texttt{corrupt} that returns all the HLL bucket contents (but not any secret key $K$); in S4 it has all three oracles.
+In addition, ${\mathcal{A}}$ always has access to an \texttt{init($\cdot$)} oracle by which it receives the HLL parameters $n$, $\ell$, $\alpha_n$ needed to initialise the HLL sketch.
+
 In the security model, ${\mathcal{A}}$ interacts in one of two worlds.
-In the real world, ${\mathcal{A}}$ interacts with a real HLL, operating as in Figure~\ref{fig:hll}.
-In the ideal world, ${\mathcal{A}}$ interacts with an ideal HLL in which the operation $y: = h(x)$ is replaced by random sampling; that is, by setting $y \leftarrow_{\$} \{0,1\}^{n+\ell}$.
-
-\todo{update! may need to formally define algs of an HLL as well as oracles. Since Sim will need to keep a shadow copy for itself.}
-
-We then show that, if $h$ is replaced by a keyed Pseudo Random Function (PRF) $F$ whose key is not available to the adversary, then any adversary ${\mathcal{A}}$ that can distinguish between the real and random worlds with oracles as in S4 can be used to build an adversary ${{\mathcal{B}}}$ which breaks the PRF security of $F$.
-Under the assumption that $F$ is a good PRF, no such efficient adversary ${{\mathcal{B}}}$ can exist, and we may conclude that we can safely switch from the real world to the ideal world without incurring any significant change to the behaviour of ${\mathcal{A}}$.
-
-However, it is clear that any adversary in the ideal world is reduced to operating in the setting used in the average case analysis of HLL (wherein every insertion query results in a random value $y$ being used to update the HLL sketch).
-Thus, ${\mathcal{A}}$ can have no advantage when it comes to influencing the HLL cardinality estimate.
+In the real world, ${\mathcal{A}}$ interacts with a real HLL, operating as in Figure~\ref{fig:hll} but with the hash function $h$ replaced by a keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$ whose key is not available to the adversary.
+In the ideal world, ${\mathcal{A}}$ interacts with a simulator ${\mathcal{S}}$ which ignores the input for ${\mathcal{A}}$'s \texttt{insert($\cdot$)} queries but nevertheless provides a ``good'' simulation of the oracle responses to ${\mathcal{A}}$.
+Importantly, ${\mathcal{S}}$ will handle all of ${\mathcal{A}}$'s requests to insert an item $x$ using \emph{random sampling}, that is, by setting $y \leftarrow_{\$} \{0,1\}^{n+\ell}$ instead of $y: = F_K(x)$.
+
+We will show that when $h$ is replaced by the keyed function $F$	 then any adversary ${\mathcal{A}}$ that can distinguish between the real and ideal worlds can be used to build an adversary ${{\mathcal{B}}}$ which breaks the PRF security of $F$.
+Under the assumption that $F$ is a good PRF, no such efficient adversary ${{\mathcal{B}}}$ can exist, and we may conclude that (up to PRF security) ${\mathcal{A}}$ cannot tell that it is operating in the ideal world instead of the real world.
+
+But in the ideal world, ${\mathcal{A}}$ interacts with a simulator that uses random sampling for the $y$ values.
+This effectively reduces ${\mathcal{A}}$ to operating in the setting where the average case analysis of HLL applies, as per~\cite{hll}.
+Here, ${\mathcal{A}}$ has no advantage when it comes to influencing the HLL cardinality estimate.
 It follows that HLL, when instantiated with PRF $F$, is secure against attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
-We will conclude by discussing how $F$ can be instantiated.
-
-\todo{explain the above better with simulators + conf for free.}
-
-We note that our formal security analysis could be conducted using the framework recently laid out in~\cite{CCS:ClaPatShrPS19}. However, we prefer to give a compact and self-contained treatment here.
-
-\todo{explain why a sim-based approach is better, more general, does not require a specific winning condition.}
-
-\todo{Why UHFs don't work in general -- key recovery is possible in S4; partial info about processed values and therefore keys  leaks from Card Est in S2, though harder to make an attack from this. }
+%Recall also that the simulator ${\mathcal{S}}$ ignores the values $x$ being inserted by ${\mathcal{A}}$.
+This enables us to get certain confidentiality properties for free when HLL is implemented using a PRF in place of $h$.
+For example, we can argue that the bucket contents (obtained via a \texttt{corrupt} query) reveal nothing to ${\mathcal{A}}$ about what values $x$ were inserted into the HLL.
+
+We will conclude our analysis by discussing how $F$ can be instantiated.
 
@@
 
-\todo{Upgrade to use simulator S as a typical simulation-based security notion should; reference to Lindell's ``How to simulate it'' paper?
-Order of quantifiers; input to simulator: nothing, not even number of queries, assuming adversary never repeats an insert query.}
-
-We consider adversary ${\mathcal{A}}$ to be an algorithm running against a challenger ${{\mathcal{C}}}$ in a security game.
-The game begins with ${{\mathcal{C}}}$ selecting a bit $b$ uniformly at random.
-Adversary ${\mathcal{A}}$ interacts with up to four oracles which are handled by ${{\mathcal{C}}}$; different combinations of oracles define scenarios S2, S3, S4.
+We consider adversary ${\mathcal{A}}$ to be an algorithm running in a security game.
+The game begins with the selection of a bit $b$ uniformly at random.
+Adversary ${\mathcal{A}}$ interacts with up to four oracles which are handled by the game; different combinations of oracles define scenarios S2, S3, S4.
 After making some queries, ${\mathcal{A}}$ returns a bit $b'$ estimating whether it is interacting with a real HLL instantiation ($b'=0$) or an ideal one ($b'=1$).
@@
 %Using a standard transformation, this probability can be rewritten as $\lvert \Pr[b' = 1| b = 1] - \Pr[b'=0|b=1] \rvert$. 
-The four oracles that  ${\mathcal{A}}$ can interact with are as follows:
+The four oracles that  ${\mathcal{A}}$ can interact with behave as follows in the real world:
 \begin{itemize}
-\item \texttt{init}: this oracle performs initialisation of the HLL sketch (and selection of the PRF key when $h$ is replaced by a PRF $F$); this oracle is available in all of S2, S3, S4, but can only be called once as ${\mathcal{A}}$'s first oracle query.
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$, the oracle inserts value $x$ into the HLL sketch.
+\item \texttt{init($\cdot$)}: this oracle receives as input the HLL parameters ($n$, $\ell$, $\alpha_n$) and performs initialisation of the HLL sketch.
+It also selects a key $K \leftarrow_{\$} \{0,1\}^k$ for keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
+This oracle is available in all of S2, S3, S4, but can only be called once as ${\mathcal{A}}$'s first oracle query.
+\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$, the oracle sets $y = F_K(x)$ and follows the procedure in Figure~\ref{fig:hll} to insert value $x$ into the HLL sketch (that is, it uses the first $m$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
 This oracle is available in all of S2, S3, S4.
@@
 We let $q$ denote the number of insertion queries made by ${\mathcal{A}}$ during the course of its attack.
-\item \texttt{read}: this oracle returns the HLL cardinality estimate. We allow multiple queries to this oracle and make it available to ${\mathcal{A}}$ in S3 and S4, but not S2.
+\item \texttt{read}: this oracle returns the HLL cardinality estimate as per Figure~\ref{fig:hll}.
+We allow multiple queries to this oracle and make it available to ${\mathcal{A}}$ in S3 and S4, but not S2.
 \item \texttt{corrupt}: this oracle returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents (but not any secret key $K$).
@@
 
-\subsection{Security Analysis of HLL with a PRF}
-
-We now replace the hash function $h$ acting on ${{\mathcal{D}}}$ from the HLL algorithm with a keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
-Here the first input to $F$ is a $k$-bit string, the key, denoted $K$.
-We assume that \texttt{init} in the HLL security model now initialises key $K$ by selecting it uniformly at random from $\{0,1\}^k$.
+In many simulation-based security proofs, the simulator ${\mathcal{S}}$ depends on ${\mathcal{A}}$, and security is quantified via a statement of the form ``For all adversaries ${\mathcal{A}}$, there exists a simulator ${\mathcal{S}}$ such that $\ldots$'' with ${\mathcal{S}}$ being described in the proof.
+In our analysis we are able to provide a universal simulator, enabling us to make a stronger statement of the form ``There exists a simulator ${\mathcal{S}}$ such that, for all adversaries ${\mathcal{A}}\ldots$''.
+We describe this simulator now, in terms of how it handles  ${\mathcal{A}}$'s oracle queries in the ideal world:
+
+\begin{itemize}
+\item \texttt{init($\cdot$)}: on input $n$, $\ell$, $\alpha_m$, ${\mathcal{S}}$ initialises an HLL sketch as in Figure~\ref{fig:hll}; that is, it creates an array of $m=2^n$ buckets $M[j]$ and initialises them all to 0.
+\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$,  ${\mathcal{S}}$ ignores $x$, sets $y \leftarrow_{\$} \{0,1\}^{n+\ell}$, and proceeds to update the buckets using this value of $y$ as in the original HLL algorithm in Figure~\ref{fig:hll}.
+\item \texttt{read}:  ${\mathcal{S}}$ returns the HLL cardinality estimate computed as per Figure~\ref{fig:hll}.
+\item \texttt{corrupt}: ${\mathcal{S}}$ returns the current vector $(M[j])_{j=1}^{m}$ of HLL bucket contents that it has stored.
+\end{itemize}
 
@@
 
-Now we are ready to state:
+Now we are ready to state our main result:
 
@@
 
-Informally, if $F$ is a good PRF, then $\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}})$ should be small for any adversary ${{\mathcal{B}}}$ running in a reasonable amount of time; hence the advantage of any S4 adversary  ${\mathcal{A}}$ against HLL using PRF $F$ is also small.
+%Informally, if $F$ is a good PRF, then $\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}})$ should be small for any adversary ${{\mathcal{B}}}$ running in a reasonable amount of time; hence the advantage of any S4 adversary  ${\mathcal{A}}$ against HLL using PRF $F$ is also small.
 Note that our proof still applies for the improved algorithm of~\cite{hllnew} that is used in Redis.
@@
 PRF adversary ${{\mathcal{B}}}$ against $F$ is built as follows.
-Let $d$ denote the hidden bit determining whether ${{\mathcal{B}}}$ receives real responses of the form $F(K,x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${{\mathcal{B}}}$ runs S4 adversary ${\mathcal{A}}$ and responds to each of its queries as follows:
+Let $d$ denote the hidden bit determining whether ${{\mathcal{B}}}$ receives real responses of the form $F_K(x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${{\mathcal{B}}}$ runs S4 adversary ${\mathcal{A}}$ and responds to each of its queries as follows:
 \begin{itemize}
@@
 Note that when $d=0$, ${{\mathcal{B}}}$ plays against the real $F$ with a random key $K$ and provides a simulation for ${\mathcal{A}}$ that perfectly instantiates the real world.
-On the other hand, when $d=1$, ${{\mathcal{B}}}$ plays against a truly random function and provides a simulation for ${\mathcal{A}}$ that perfectly instantiates the ideal world.
+On the other hand, when $d=1$, ${{\mathcal{B}}}$ plays against a truly random function and provides a simulation for ${\mathcal{A}}$ that perfectly instantiates the ideal world with simulator $\mathcal{S}$ (to see this, note that sampling the outputs $y$ of a truly random function at distinct inputs $x$ is identical to simply sampling $y$ uniformly at random).
 The equality of advantages in the statement of the theorem now follows immediately.
@@
 
-We conclude by reiterating our central point: we have shown that no efficient S4 adversary ${\mathcal{A}}$ can tell that it is playing against HLL implemented using truly random values $y$ instead of with values obtained from a PRF, except with a small advantage.
+We reiterate our central point: the theorem shows that no efficient S4 adversary ${\mathcal{A}}$ can tell that it is playing against our simulator ${\mathcal{S}}$ using truly random values $y$ instead of with values obtained from a PRF, except with a small advantage.
 This means that ${\mathcal{A}}$ is effectively reduced to attacking HLL in the exact same setting used in the average case analysis of HLL from~\cite{hll}, wherein it is proved that the HLL cardinality estimate endures relative errors of the form $1.04/\sqrt{m}$.
-It does not matter that ${\mathcal{A}}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries.
+From this we can conclude that no efficient S4 adversary ${\mathcal{A}}$ can manipulate the HLL cardinality estimate to its benefit.
+
+\subsection{On Confidentiality}
+We also recall that our ${\mathcal{S}}$ ignores ${\mathcal{A}}$'s inputs $x$ in handling \texttt{insert($\cdot$)} queries.
+In traditional simulation-based security notions, this would imply that the ``leakage'' given to the simulator does not include the input values $x$, and one would then conclude by analogy to, say, semantic security for encryption that the HLL implemented using a PRF offers confidentiality for the inputs $x$.
+This is not quite correct here, since ${\mathcal{A}}$ chooses all the inputs $x$ and so ``knows'' them already.
+To address confidentiality, one would need instead to consider $x$ that are drawn from some distribution on $\mathcal{D}$ or consider two-stage adversaries that ``forget'' the $x$ values chosen in its first stage by not passing state to its second stage.
+
+Even here, however, an HLL can leak information about values $x$ that were previously inserted.
+Consider the following attack from~\cite{cardestprivacy}: to test whether a value $x$ was previously inserted into the HLL, make the sequence of queries: \texttt{read}, \texttt{insert($x$)}, \texttt{read}.
+If the two cardinality estimates are different, one can conclude definitively that $x$ was \emph{not} previously inserted into the HLL; if they are the same, one can estimate (with some error) that $x$ was previously inserted.
+This attack still works even if HLL is implemented using a PRF and is unavoidable in view of the oracle queries (and in real settings, the API) available to the adversary.
+So such privacy leaks are an inherent feature of HLL if a suitably rich API is provided.
+Note that this attack does not contradict our main result, which concerns prevention of cardinality manipulation rather than confidentiality.
+
+More positively, what we can say is that our simulation-based security definition does imply confidentiality of inputs in the  case of a model where an honest party makes a sequence of insertions and where ${\mathcal{A}}$ is then limited to \texttt{read} and \texttt{corrupt} queries.
+\todo[inline]{explain why.}
+
+We leave the further exploration of how to adapt our simulation-based security definitions to capture confidentiality notions to future work.
+
+%It does not matter that ${\mathcal{A}}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries.
 For each \texttt{insert($\cdot$)} query, a fresh value $x$ leads to a new random $y$ being generated and used to update a bucket, as in the average case analysis of HLL.
@@
 
-\todo{revisit the above in view of introduction of simulator, plus the confidentiality aspect for free. NB corruption query does not give PRF key.}
+\todo{explain why a sim-based approach is better, more general, does not require a specific winning condition.}
+
+\todo{Why UHFs don't work in general so why we need a PRF -- key recovery is possible in S4; partial info about processed values and therefore keys  leaks from Card Est in S2, though harder to make an attack from this.
+}
 
@@
 
-The above construction requires a PRF $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
+Our HLL construction requires a PRF $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
 Here, the set ${{\mathcal{D}}}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
@@
 This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security.
+
+\todo{Sim based of indepednent interest, arguably more powerful than game-based notions with fixed winning predicates.}
 %It is also an open problem to close the performance gap that exists between fast but non-cryptographic hashes used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security.

### 2021-04-26 docs/HLL_exp_bad/main.tex
clean up on proof and confidentiality for free

@@
 \acmversiontrue
+
+\newif\ifanon % turn on ACM style (can be short or full)
+\anontrue
 
@@
 In the remainder of this section, we introduce a formal security model covering all three scenarios.
-We work with a simulation-based framework.
-This is in contrast to the game-based approach introduced for probabilistic data structures in~\cite{CCS:ClaPatShrPS19}.
-We discuss the strengths of our alternative approach in more detail below.
+We work with a simulation-based framework, in contrast to the game-based approach introduced for probabilistic data structures in~\cite{CCS:ClaPatShrPS19}.
+%We discuss the strengths of our alternative approach in more detail below.
 
@@
 %Using a standard transformation, this probability can be rewritten as $\lvert \Pr[b' = 1| b = 1] - \Pr[b'=0|b=1] \rvert$.
+
 The four oracles that  ${\mathcal{A}}$ can interact with behave as follows in the real world:
@@
 
-In many simulation-based security proofs, the simulator ${\mathcal{S}}$ depends on ${\mathcal{A}}$, and security is quantified via a statement of the form ``For all adversaries ${\mathcal{A}}$, there exists a simulator ${\mathcal{S}}$ such that $\ldots$'' with ${\mathcal{S}}$ being described in the proof.
+In many simulation-based security proofs, the simulator ${\mathcal{S}}$ depends on ${\mathcal{A}}$, and security is quantified via a statement of the form ``For all adversaries ${\mathcal{A}}$, there exists a simulator ${\mathcal{S}}$ such that $\ldots$'' with ${\mathcal{S}}$ being described in the proof of a specific scheme.
 In our analysis we are able to provide a universal simulator, enabling us to make a stronger statement of the form ``There exists a simulator ${\mathcal{S}}$ such that, for all adversaries ${\mathcal{A}}\ldots$''.
@@
 \begin{theorem}\label{thm:security-of-HLL-with-PRF}
-Suppose ${\mathcal{A}}$ is an S4 adversary against HLL instantiated with a keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
+Let simulator $\mathcal{S}$ be as described above.
+Suppose ${\mathcal{A}}$ is any S4 adversary against HLL instantiated with a keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
+Suppose ${\mathcal{A}}$ either plays in the real world or in the ideal world with $\mathcal{S}$.
 Then we can construct from ${\mathcal{A}}$ a PRF adversary ${{\mathcal{B}}}$ against $F$ such that:
@@
 
+Note that our security analysis using a simulation-based definition does not require us to specify a winning condition for the adversary.
+Rather, we infer from inspecting the simulator what the adversary can (or cannot) achieve in the ideal world, and show that no efficient adversary can distinguish the real and ideal worlds.
+This stands in contrast to the game-based approach of~\cite{CCS:ClaPatShrPS19} which needs a complex definition to say when the adversary has successfully manipulated a probabilistic data structure.
+This reveals a strength of the simulation-based approach in more complex settings like ours.
+
 \subsection{On Confidentiality}
-We also recall that our ${\mathcal{S}}$ ignores ${\mathcal{A}}$'s inputs $x$ in handling \texttt{insert($\cdot$)} queries.
-In traditional simulation-based security notions, this would imply that the ``leakage'' given to the simulator does not include the input values $x$, and one would then conclude by analogy to, say, semantic security for encryption that the HLL implemented using a PRF offers confidentiality for the inputs $x$.
-This is not quite correct here, since ${\mathcal{A}}$ chooses all the inputs $x$ and so ``knows'' them already.
-To address confidentiality, one would need instead to consider $x$ that are drawn from some distribution on $\mathcal{D}$ or consider two-stage adversaries that ``forget'' the $x$ values chosen in its first stage by not passing state to its second stage.
-
-Even here, however, an HLL can leak information about values $x$ that were previously inserted.
-Consider the following attack from~\cite{cardestprivacy}: to test whether a value $x$ was previously inserted into the HLL, make the sequence of queries: \texttt{read}, \texttt{insert($x$)}, \texttt{read}.
-If the two cardinality estimates are different, one can conclude definitively that $x$ was \emph{not} previously inserted into the HLL; if they are the same, one can estimate (with some error) that $x$ was previously inserted.
-This attack still works even if HLL is implemented using a PRF and is unavoidable in view of the oracle queries (and in real settings, the API) available to the adversary.
+
+An HLL can leak information about values $x$ that were previously inserted.
+Consider the following attack from~\cite{cardestprivacy}: to test whether a value $x$ was previously inserted into the HLL by an honest user (say), the adversary makes the sequence of queries: \texttt{read}, \texttt{insert($x$)}, \texttt{read}.
+If the two cardinality estimates are different, the adversary can conclude definitively that $x$ was \emph{not} previously inserted into the HLL; if they are the same, it can estimate (with some error) that $x$ was previously inserted.\footnote{Errors can arise in two ways when $x$ was not previously inserted: either inserting $x$ changes a bucket value, but the change is masked by rounding when the cardinality estimate is computed, or inserting $x$ does lead to a change in a bucket value because $\rho(w)$ computed from $x$ is too small.} This attack still works even if HLL is implemented using a PRF and is unavoidable in view of the oracle queries (and in real settings, the API) available to the adversary.
 So such privacy leaks are an inherent feature of HLL if a suitably rich API is provided.
@@
 
-More positively, what we can say is that our simulation-based security definition does imply confidentiality of inputs in the  case of a model where an honest party makes a sequence of insertions and where ${\mathcal{A}}$ is then limited to \texttt{read} and \texttt{corrupt} queries.
-\todo[inline]{explain why.}
+Still, simulation-based security notions are often used to model confidentiality properties, so it is tempting to think that our security definition might say something about confidentiality of the HLL inputs $x$.
+Our existing model does not, since ${\mathcal{A}}$ chooses all the inputs $x$ and so ``knows'' them already.
+However, it is easily modified to provide a confidentiality notion.
+For example, we can consider the insertion of honestly generated values $x$ that are drawn from some probability distribution $p$ on $\mathcal{D}$ in a first phase, with ${\mathcal{A}}$ then having access to only \texttt{read} and \texttt{corrupt} oracles in a second phase (we have to remove the \texttt{insert($\cdot$)} oracle in view of the general attack from~\cite{cardestprivacy}).
+
+In this case, the simulator ${\mathcal{S}}$ in the ideal world can proceed without ever seeing any inputs but instead knowing only the distribution $p$ and the number of distinct samples drawn from it in the first phase.
+It is not hard to modify the proof of Theorem~\ref{thm:security-of-HLL-with-PRF} to show that no adversary can distinguish real and ideal worlds in this case (up to PRF security).
+Now, since ${\mathcal{S}}$ has no access to inputs $x$, we can conclude that the HLL implemented using a PRF ensures confidentiality for the inputs $x$.
+Thus, the cardinality estimates and even the bucket values themselves do not leak information about the honestly inserted values $x$ (beyond their number).
 
@@
 
-%It does not matter that ${\mathcal{A}}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries.
-For each \texttt{insert($\cdot$)} query, a fresh value $x$ leads to a new random $y$ being generated and used to update a bucket, as in the average case analysis of HLL.
-The \texttt{read} and \texttt{corrupt} queries do leak information about $F$'s outputs to the adversary, but such outputs are indistinguishable from random under the assumption that $F$ is a PRF.
-
-\todo{explain why a sim-based approach is better, more general, does not require a specific winning condition.}
-
-\todo{Why UHFs don't work in general so why we need a PRF -- key recovery is possible in S4; partial info about processed values and therefore keys  leaks from Card Est in S2, though harder to make an attack from this.
-}
+\subsection{Is a PRF Required?}
+
+It is natural to wonder whether a PRF is really required in place of hash function $h$ in order to make HLL secure, or whether a weaker cryptographic primitive would suffice.
+For example, perhaps using a keyed universal hash function (UHF) $h_K(\cdot)$ would be sufficient.
+
+In general, UHFs do not preserve confidentiality of their keys: the key may be recovered from a small number of UHF outputs.
+Consider, for example, the UHF built from polynomial hashing over a finite field $\mathcal{F}$.
+Here, inputs $x$ are encoded as polynomials  $p_x \in \mathcal{F}(X)$, the key is a uniformly random field element, and the UHF output $h_K(x)$  is equal to $p_x(K)$, obtained by polynomial evaluation at $K$.
+It is well-known that an adversary who knows the input $x$ can determine the key $K$ by finding roots of polynomials over $\mathcal{F}$, using Berlekamp's algorithm~\cite{Berlekamp}.
+If the adversary can choose $x$, then it can make the polynomials $p_x$ have low degree, including of degree 1, making the attack easier to perform.
+
+Now consider using such a UHF in place of $h$ in attack scenario S4.
+For concreteness, we assume $\mathcal{F}$ equals $\textsf{GF}(2^{128})$, so our UHF has 128-bit keys and 128-bit outputs.
+The output is then truncated to provide the 32 bits needed by HLL (or, say, the 64 bits consumed by the Redis variant of HLL).
+In S4 an adversary $\mathcal{A}$ can perform a series of pairs of \texttt{insert($x_i$)} and \texttt{corrupt} queries on chosen values $x_1, x_2, \ldots$.
+With high probability, each  \texttt{insert} query touches a fresh bucket in the HLL sketch.
+Then each \texttt{corrupt} query reveals to $\mathcal{A}$ which bucket was touched by the preceding \texttt{insert} query.
+In turn, this leaks partial information about the corresponding UHF output $p_{x_i}(K)$.
+
+Going further, the S4 adversary has complete control over the values $x_i$, so it can choose them so that the polynomials $p_{x_i}$ are of degree 1 over $\textsf{GF}(2^{128})$.
+Then $K$ can be determined by writing everything bit-wise and solving a linear system.
+The idea is that, when $x_i$ has 128 bits, we can write $h_K(x_i) = p_{x_i}(K) = x_i \cdot K$ where the multiplication is over $\textsf{GF}(2^{128})$ (and we turn bit-strings into field elements in a standard way).
+Then each bit of $h_K(x_i)$ is a linear combination of the bits of $K$, with the coefficients being determined by the known bits of $x_i$ and the degree 128 irreducible polynomial $f$ over $\textsf{GF}(2)$ that determines how field multiplication works at the bit level.
+
+The above attack exploits specific properties of the chosen UHF, but demonstrates that UHFs are not a suitable tool for protecting HLL in general.
+It is still possible that UHFs could be safely used in weaker attack scenarios, e.g.\ S2.
+We leave a detailed study to future work.
 
@@
 
+\ifanon
+\else
+\section*{Acknowledgements}
+The work of Paterson is supported in part by a gift from VMware.
+
+The authors thank Anupama Unnikrishnan for useful discussions on simulation-based definitions.
+\fi
+
 \bibliographystyle{ieeetr}

### 2021-04-27 docs/HLL_exp_bad/main.tex
better attack on UHF

@@
 
+\subsection{Paper Organisation}
 The rest of the paper is organised as follows.
@@
 \section{Overview of HLL}\label{sec:overview}
-\subsection{The Algorithm}
-HLL~\cite{hll} is based on the key observation that, for a stream of items represented as bit-strings of some fixed length, if all items have at most $k$ leading zero-bits (and one or more items actually have this many leading zero-bits), then the cardinality of the stream (i.e.\ the number of distinct values it contains) is likely to be on the order of $2^{k+1}$.
-To ensure a uniform distribution on the items' representations as bit-strings, the items are processed by a hash function $h$ before the leading bits are inspected.
-
-So to roughly estimate the cardinality of a stream, we need only to store a representation of $k$, the length of the largest observed string of leading zero-bits for items in the stream.
-This provides a very compact mechanism for estimating a stream's cardinality -- if the true cardinality is $N$, then just $\mathcal{O}(\log\log(N))$ bits of storage are needed to compute the cardinality estimate.
-
-This simple approach to cardinality estimation suffers from large variance in the estimation.
-In order to improve the variance, HLL uses many estimators in parallel instead of one and averages the results.
-Each estimate is stored in its own register that we call a bucket.
-We refer to the collection of buckets and their contents as the HLL \emph{sketch}.
-To map a value $x$ to one of $m=2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$.
-The output length of $h(\cdot)$ is thus $n+\ell$, fixed to 32 in the original HLL paper~\cite{hll}.
-Estimates from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha_m$, where $\alpha_m$ is empirically computed to correct a systematic multiplicative bias.
-
-In the original HLL proposal~\cite{hll}, relatively small and large cardinalities are handled separately.
-This is to correct for systematic errors.
-When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate.
-This estimator uses the number of empty buckets (whose register value is 0) instead of the number of leading zeros in the stream.
-A similar correction is applied for large cardinalities in order to take into account the likelihood of hash collisions.
-In addition, the cardinality estimate is usually rounded to a whole number before being output.
-The original HLL algorithm from~\cite{hll} is shown in Fig.~\ref{fig:hll}.
-This algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
-
-Some works building on~\cite{hll} replace the Hit Counting algorithm by other schemes and use a hash function with a larger output domain to reduce the chances of collision and thus get rid of the need for large cardinality range correction.
-For example, Google engineers proposed using a 64-bit hash function and Linear Counting along with bias correction in the small cardinality range~\cite{hllpractice},\footnote{See also the Google blogpost describing the implementation in Google Big Query at~\url{https://cloud.google.com/blog/products/gcp/counting-uniques-faster-in-bigquery-with-hyperloglog}.} while Redis has adopted a new estimator based on a more precise statistical analysis of the counting problem provided by~\cite{hllnew}.
-
-In this paper we mostly focus on the HLL formulation in the original paper~\cite{hll} as represented in Fig.~\ref{fig:hll} and as widely used in practice.
-We also study~\cite{hllnew} because of its recent deployment in Redis.
-We leave for future work the analysis of further HLL variants.
-\todo{revise this if we add Google BQ.}
-
-An interesting property of HLL is that it supports merging in a lossless way.
-To combine two buckets at index $i$ in two different HLL instances, one can take the maximum of the two bucket entries and assign that to the matching bucket $i$ in the merged HLL sketch.
-Such a simple set union operation allows easy parallelisation of operations among multiple machines independently, provided they use the same hash function and the same number of buckets.
-Clearly this procedure cannot be carried out if the different HLL sketches are differently seeded or keyed.
 
@@
 
+HLL~\cite{hll} is based on the key observation that, for a stream of items represented as bit-strings of some fixed length, if all items have at most $k$ leading zero-bits (and one or more items actually have this many leading zero-bits), then the cardinality of the stream (i.e.\ the number of distinct values it contains) is likely to be on the order of $2^{k+1}$.
+To ensure a uniform distribution on the items' representations as bit-strings, the items are processed by a hash function $h$ before the leading bits are inspected.
+
+So to roughly estimate the cardinality of a stream, we need only to store a representation of $k$, the length of the largest observed string of leading zero-bits for items in the stream.
+This provides a very compact mechanism for estimating a stream's cardinality -- if the true cardinality is $N$, then just $\mathcal{O}(\log\log(N))$ bits of storage are needed to compute the cardinality estimate.
+
+This simple approach suffers from large variance.
+In order to reduce the variance, HLL uses many estimators in parallel instead of one and averages the results.
+Each estimate is stored in its own register that we call a bucket.
+We refer to the collection of buckets and their contents as the HLL \emph{sketch}.
+To map a value $x$ to one of $m=2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$.
+The output length of $h(\cdot)$ is thus $n+\ell$.
+This sum is fixed to 32 in the original HLL paper~\cite{hll}.
+Estimates from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha_m$, where $\alpha_m$ is empirically computed to correct a systematic multiplicative bias.
+
+In the original HLL proposal~\cite{hll}, relatively small and large cardinalities are handled separately to correct for systematic errors.
+When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate.
+This estimator uses the number of empty buckets (whose register value is 0) instead of the number of leading zeros in the stream.
+A similar correction is applied for large cardinalities in order to take into account the likelihood of hash collisions.
+In addition, the cardinality estimate is usually rounded to a whole number before being output.
+The original HLL algorithm from~\cite{hll} is shown in Fig.~\ref{fig:hll}.
+This algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 4\%, using only 1.5 kB of memory.
+Crucially, the proof of this~\cite[Theorem 1]{hll} makes the assumption that the inputs are drawn independently and uniformly at random from the input domain.\footnote{Technically, the results of~\cite{hll} are proven for a distribution obtained by making arbitrary replications and permutations after sampling in this way, to simulate input taken from an \emph{ideal multiset}.
+But they also hold without replication and permutation because HLL's internal state is invariant under repeated inputs.}
+
+Some works building on~\cite{hll} replace the Hit Counting algorithm by other schemes and use a hash function with a larger output domain to reduce the chances of collision and thus get rid of the need for large cardinality range correction.
+For example, Google engineers proposed using a 64-bit hash function and Linear Counting along with bias correction in the small cardinality range~\cite{hllpractice},\footnote{See also the Google blogpost describing the implementation in Google Big Query at~\url{https://cloud.google.com/blog/products/gcp/counting-uniques-faster-in-bigquery-with-hyperloglog}.} while Redis has adopted a new estimator based on a more precise statistical analysis of the counting problem provided by~\cite{hllnew}.
+
+In this paper we mostly focus on the HLL formulation in the original paper~\cite{hll} as represented in Fig.~\ref{fig:hll} and as widely used in practice.
+We also study~\cite{hllnew} because of its recent deployment in Redis.
+We leave for future work the analysis of further HLL variants.
+\todo{revise this if we add Google BQ.}
+
+An interesting property of HLL is that it supports merging in a lossless way.
+To combine two buckets at index $i$ in two different HLL instances, one can take the maximum of the two bucket entries and assign that to the matching bucket $i$ in the merged HLL sketch.
+Such a simple set union operation allows easy parallelisation of operations among multiple machines independently, provided they use the same hash function and the same number of buckets.
+Clearly this procedure cannot be carried out if the different HLL sketches are differently seeded or keyed.
+
 \section{HLL in Adversarial Settings}\label{sec:attacks}
 
-This section presents different adversarial settings and corresponding attacks that manipulate the HLL cardinality estimate.
-The adversary's goal is to reduce the estimate as much as possible, since this would cause more damage when  considering the previously mentioned applications of HLL.
+This section presents the different adversarial settings that we consider.
+In all settings, our adversary's aim is to manipulate the HLL cardinality estimate.
+We focus on reducing the estimate as much as possible, since this would cause more damage when  considering the previously mentioned applications of HLL.
 It is easy to modify our attack strategies for an adversary who instead wishes to artificially inflate the HLL cardinality estimate given a limited input capability.
-We also analyze the impact on the HLL cardinality estimate and the computational effort required.
+%We also analyze the impact on the HLL cardinality estimate and the computational effort required.
 
@@
 \paragraph{Simple Attack:}
-The simplistic version of our attack uses the above idea, inserting items $x$ into the shadow HLL sketch but resetting it as soon as the cardinality estimate increases from $B$ to $B+1$ (indicating that a non-target bucket has been hit by item $x$).
+The simple version of our attack uses the above idea, inserting items $x$ into the shadow HLL sketch but resetting it as soon as the cardinality estimate increases from $B$ to $B+1$ (indicating that a non-target bucket has been hit by item $x$).
 Acting in this way precisely determines whether each inserted item hits the target buckets or not, so has no ``false positives'' which need to be eliminated through reprocessing of the list ${\mathcal{L}}$.
@@
 \end{eqnarray*}
-where $H_M$ is the harmonic mean of the counts $2^{M[1]}, 2^{M[2]}, \ldots, 2^{M[m]}$.
+where $H_M$ is the harmonic mean of $2^{M[1]}, 2^{M[2]}, \ldots, 2^{M[m]}$.
 So the adversary can expect to insert a fraction $(1 - H_M^{-1})$ of the items from ${\mathcal{X}}$ without increasing the HLL cardinality estimate at all.
@@
 We now illustrate what effect such attacks might have on applications relying on HLL for cardinality estimation.
-In an effort to be more concrete, we analyse in detail a recent paper~\cite{flexswitch} from the networking systems community.
+Specifically, we analyse in detail a recent paper~\cite{flexswitch} from the networking systems community.
 
@@
 It is natural to wonder whether a PRF is really required in place of hash function $h$ in order to make HLL secure, or whether a weaker cryptographic primitive would suffice.
-For example, perhaps using a keyed universal hash function (UHF) $h_K(\cdot)$ would be sufficient.
-
-In general, UHFs do not preserve confidentiality of their keys: the key may be recovered from a small number of UHF outputs.
+For example, since it has (roughly) uniform outputs, perhaps using a universal hash function (UHF) $h_K(\cdot)$ would be sufficient.
+
+In general the key of a UHF may be recovered from a small number of its outputs.
 Consider, for example, the UHF built from polynomial hashing over a finite field $\mathcal{F}$.
-Here, inputs $x$ are encoded as polynomials  $p_x \in \mathcal{F}(X)$, the key is a uniformly random field element, and the UHF output $h_K(x)$  is equal to $p_x(K)$, obtained by polynomial evaluation at $K$.
-It is well-known that an adversary who knows the input $x$ can determine the key $K$ by finding roots of polynomials over $\mathcal{F}$, using Berlekamp's algorithm~\cite{Berlekamp}.
-If the adversary can choose $x$, then it can make the polynomials $p_x$ have low degree, including of degree 1, making the attack easier to perform.
+Here, inputs $x$ are encoded as polynomials $p_x \in \mathcal{F}(Z)$, the key is a uniformly random field element, and the UHF output $h_K(x)$  is equal to $p_x(K)$, obtained by polynomial evaluation at $K$.
+The specific translation from $x$ to $p_x$ involves encoding $x$ as a vector $(x_1,x_1,\ldots,x_v) \in \mathcal{F}^v$ for some $v$ and then setting:
+\[
+p_x(Z) = Z^v + x_1Z^{v-1} + \cdots + x_{v-1}Z + x_v.
+\]
+An adversary who knows the input $x$, and therefore the vector $(x_1,x_2,\ldots,x_v)$ can determine the key $K$ by finding roots of polynomials over $\mathcal{F}$, using Berlekamp's algorithm~\cite{Berlekamp}.
+If the adversary can choose $x$, then it can make the polynomials $p_x$ have low degree, making the attack easier to perform.
 
 Now consider using such a UHF in place of $h$ in attack scenario S4.
-For concreteness, we assume $\mathcal{F}$ equals $\textsf{GF}(2^{128})$, so our UHF has 128-bit keys and 128-bit outputs.
+For concreteness, we assume $\mathcal{F} = \textsf{GF}(2^{128})$, so our UHF has 128-bit keys and 128-bit outputs.
 The output is then truncated to provide the 32 bits needed by HLL (or, say, the 64 bits consumed by the Redis variant of HLL).
-In S4 an adversary $\mathcal{A}$ can perform a series of pairs of \texttt{insert($x_i$)} and \texttt{corrupt} queries on chosen values $x_1, x_2, \ldots$.
-With high probability, each  \texttt{insert} query touches a fresh bucket in the HLL sketch.
-Then each \texttt{corrupt} query reveals to $\mathcal{A}$ which bucket was touched by the preceding \texttt{insert} query.
-In turn, this leaks partial information about the corresponding UHF output $p_{x_i}(K)$.
-
-Going further, the S4 adversary has complete control over the values $x_i$, so it can choose them so that the polynomials $p_{x_i}$ are of degree 1 over $\textsf{GF}(2^{128})$.
+The encoding step, taking bit-string $x$ to vector $(x_1,x_2,\ldots,x_v)$, simply splits $x$ into 128-bit strings $x_i$, each $x_i$ representing a field element.
+The attack we present below is specific to this UHF, but can be extended to other cases.
+
+An S4 adversary $\mathcal{A}$ can perform a series of pairs of \texttt{insert($x^i$)} and \texttt{corrupt} queries on chosen values $x^0, x^1, \ldots, x^{t-1}$.
+With high probability (assuming $t$ is not too large), each \texttt{insert} query hits a fresh bucket in the HLL sketch.
+Then comparing the results of the \texttt{corrupt} queries either side of the \texttt{insert($x^i$)} query reveals to $\mathcal{A}$ which bucket was hit by that  \texttt{insert} query.
+In turn, this leaks partial information about the corresponding UHF output $p_{x^i}(K)$, specifically, its $n$ least significant bits (since those bits are used to generate the bucket index for insertion).
+We now show how to exploit this partial information on $p_{x^i}(K)$ to recover $K$.
+
+%For example when $x^i$ is short enough that $v=1$, then $p_{x^i}(K) = K + x^i$ and the bucket index leaks $n$ bits of $K$ immediately.
+
+By choosing $x^i$ to have 256 bits, we have $v=2$ and $p_{x^i}(K) = K^2 + x^i_1K+x^i_2$; then further selecting $x^i$ such that  $x^i_2$ is a constant value (e.g.\ $0 \in \textsf{GF}(2^{128})$) and considering consecutive pairs of $x_i$ values, we obtain equations of the form:
+\[
+p_{x^{2j+1}}(K) \oplus p_{x^{2j}}(K) = (x^{2j+1}_1 \oplus x^{2j}_1)K
+\]
+where $\oplus$ denotes addition over $\textsf{GF}(2^{128})$.
+Here, the $n$ least significant bits of the left-hand side leak through the indices of the buckets hit for each index.
+By selecting the $x^i_1$ uniformly at random and analysing the equations of the above form that result, we obtain a non-singular linear system over $\textsf{GF}(2)$ in the unknown bits of $K$.
+The number $t$ of $x^i$ needed is roughly $2\cdot128/n$ since we get $n$ independent linear equations per pair $(x^{2j},x^{2j+1})$ and there are 128 unknown bits in $K$.
+The required value of $t$ is small for typical HLL parameters (e.g.
+$n=16$ would yield $t=32$).
+
+%Going further, the S4 adversary has complete control over the values $x_i$, so it can choose them so that the polynomials $p_{x_i}$ are of degree 1 over $\textsf{GF}(2^{128})$.
 Then $K$ can be determined by writing everything bit-wise and solving a linear system.
@@
 
-The above attack exploits specific properties of the chosen UHF, but demonstrates that UHFs are not a suitable tool for protecting HLL in general.
+The above attack exploits specific properties of the chosen UHF, but suffices to demonstrate that UHFs are not a suitable tool for protecting HLL in general.
 It is still possible that UHFs could be safely used in weaker attack scenarios, e.g.\ S2.

### 2021-04-29 docs/HLL_exp_bad/main.tex
redoing model and proof

@@
 \begin{abstract}
-Computing the count of distinct elements in massive data sets is a common task but naive approaches are memory-expensive.
-The HyperLogLog (HLL) algorithm (Flajolet \emph{et al.}, 2007) estimates a data set's cardinality while using significantly less memory than a naive approach at the cost of some accuracy.
+Computing the count of distinct elements in large data sets is a common task but naive approaches are memory-expensive.
+The HyperLogLog (HLL) algorithm (Flajolet \emph{et al.}, 2007) estimates a data set's cardinality while using significantly less memory than a naive approach, at the cost of some accuracy.
 This trade-off makes the HLL algorithm very attractive for a wide range of applications such as database management and network monitoring, where an exact count may not be needed.
@@
 \section{Introduction}
+\todo{more here introducing big data and why we want statistical estimators on them; data streams; cannot store, or large data sets, want summary statistics.
+Such things are supported by GBQ, Apache Spark, Redis.
+Cardinality as a key metric; HLL as the algorithm of choice.
+''analytics engines for big data processing.''}
 Naive approaches to computing the cardinality of large data sets, such as sorting the elements or simply maintaining the set of unique elements seen, are impractical at scale and in streaming settings.
@@
 Recently, Reviriego and Ting~\cite{hllvuln} initiated work on the performance of HLL in adversarial settings.
-They showed that in very specific attack settings, the cardinality estimate made by HLL could be modified through input selection.
+They showed that in very specific attack settings, the cardinality estimate made by HLL can be modified through input selection.
 In this paper, we present a complete analysis of HLL in adversarial settings, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
@@
 The attack target of~\cite{cardestprivacy} is different from ours, being focused on privacy breaches.
-It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
+%It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
 In contrast, we are concerned with an attacker attempting to manipulate the cardinality estimation process itself (an issue not considered in~\cite{cardestprivacy}).
@@
 This algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 4\%, using only 1.5 kB of memory.
-Crucially, the proof of this~\cite[Theorem 1]{hll} makes the assumption that the inputs are drawn independently and uniformly at random from the input domain.\footnote{Technically, the results of~\cite{hll} are proven for a distribution obtained by making arbitrary replications and permutations after sampling in this way, to simulate input taken from an \emph{ideal multiset}.
+Crucially, the proof of this~\cite[Theorem 1]{hll} makes the assumption that the inputs are drawn independently and uniformly at random from the input domain.\footnote{Technically, the results of~\cite{hll} are proven for a distribution obtained by making arbitrary replications and permutations after sampling independently and uniformly, to simulate input taken from an \emph{ideal multiset}.
 But they also hold without replication and permutation because HLL's internal state is invariant under repeated inputs.}
@@
 We first begin by explaining the high-level idea of our approach.
-We allow an adversary ${\mathcal{A}}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios S2, S3 and S4: it may have an \texttt{insert($\cdot$)} oracle allowing it to add values to the HLL, a \texttt{read} oracle that returns the HLL cardinality estimate, and a \texttt{corrupt} that returns all the HLL bucket contents (but not any secret key $K$); in S4 it has all three oracles.
-In addition, ${\mathcal{A}}$ always has access to an \texttt{init($\cdot$)} oracle by which it receives the HLL parameters $n$, $\ell$, $\alpha_n$ needed to initialise the HLL sketch.
-
-In the security model, ${\mathcal{A}}$ interacts in one of two worlds.
-In the real world, ${\mathcal{A}}$ interacts with a real HLL, operating as in Figure~\ref{fig:hll} but with the hash function $h$ replaced by a keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$ whose key is not available to the adversary.
-In the ideal world, ${\mathcal{A}}$ interacts with a simulator ${\mathcal{S}}$ which ignores the input for ${\mathcal{A}}$'s \texttt{insert($\cdot$)} queries but nevertheless provides a ``good'' simulation of the oracle responses to ${\mathcal{A}}$.
-Importantly, ${\mathcal{S}}$ will handle all of ${\mathcal{A}}$'s requests to insert an item $x$ using \emph{random sampling}, that is, by setting $y \leftarrow_{\$} \{0,1\}^{n+\ell}$ instead of $y: = F_K(x)$.
-
-We will show that when $h$ is replaced by the keyed function $F$	 then any adversary ${\mathcal{A}}$ that can distinguish between the real and ideal worlds can be used to build an adversary ${{\mathcal{B}}}$ which breaks the PRF security of $F$.
-Under the assumption that $F$ is a good PRF, no such efficient adversary ${{\mathcal{B}}}$ can exist, and we may conclude that (up to PRF security) ${\mathcal{A}}$ cannot tell that it is operating in the ideal world instead of the real world.
-
-But in the ideal world, ${\mathcal{A}}$ interacts with a simulator that uses random sampling for the $y$ values.
-This effectively reduces ${\mathcal{A}}$ to operating in the setting where the average case analysis of HLL applies, as per~\cite{hll}.
-Here, ${\mathcal{A}}$ has no advantage when it comes to influencing the HLL cardinality estimate.
-It follows that HLL, when instantiated with PRF $F$, is secure against attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
-%Recall also that the simulator ${\mathcal{S}}$ ignores the values $x$ being inserted by ${\mathcal{A}}$.
-This enables us to get certain confidentiality properties for free when HLL is implemented using a PRF in place of $h$.
-For example, we can argue that the bucket contents (obtained via a \texttt{corrupt} query) reveal nothing to ${\mathcal{A}}$ about what values $x$ were inserted into the HLL.
-
-We will conclude our analysis by discussing how $F$ can be instantiated.
+We allow an adversary ${\mathcal{A}}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios S2, S3 and S4: it may have an \texttt{insert($\cdot$)} oracle allowing it to add values to the HLL sketch, a \texttt{read} oracle that returns the HLL cardinality estimate, and a \texttt{corrupt} that returns all the HLL bucket contents (but not any secret key $K$); in S4 it has all three oracles.
+In addition, in all 3 scenarios, ${\mathcal{A}}$ has access to an \texttt{init($\cdot$)} oracle by which it sets the HLL parameters $n$, $\ell$, $\alpha_n$ used to initialise the HLL sketch.
+We assume ${\mathcal{A}}$ always makes exactly one \texttt{init($\cdot$)} query and that it is its first oracle call.
+
+In the security model, the game selects at random one of two worlds by sampling a bit $b$.
+In the real world ($b=0$), ${\mathcal{A}}$ interacts with the scheme HLL-$F$.
+This is the same as HLL operating as in Figure~\ref{fig:hll} but with the hash function $h$ replaced by a keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$ whose key $K$ is sampled uniformly at random at the beginning of the game and not available to the adversary; ${\mathcal{A}}$ has access to the relevant selection of oracles and completes its execution with some output $\textsf{out}_{\mathcal{A}}$ (a bit-string).
+
+In the ideal world ($b=1$), ${\mathcal{A}}$ is replaced by a simulator ${\mathcal{S}}$ with blackbox oracle access to ${\mathcal{A}}$ (that is, ${\mathcal{S}}$ can run ${\mathcal{A}}$ in a blackbox manner, receiving its oracle queries and providing responses to ${\mathcal{A}}$ as it sees fit).
+In particular, ${\mathcal{S}}$ receives parameters $n$, $\ell$, $\alpha_n$ from ${\mathcal{A}}$'s \texttt{init($\cdot$)} call.
+Finally, ${\mathcal{S}}$ receives $\textsf{out}_{\mathcal{A}}$ and completes its execution by outputting its own output $\textsf{out}_{\mathcal{S}}$.
+
+Importantly, we will specify a construction of ${\mathcal{S}}$ that handles any of ${\mathcal{A}}$'s \texttt{insert($x$)} queries using \emph{random sampling}, that is, by setting $y \leftarrow_{\$} \{0,1\}^{n+\ell}$ instead of $y: = F_K(x)$ and then executing the rest of the insertion process as in Figure~\ref{fig:hll}.
+In essence, our ${\mathcal{S}}$ will build its own ``shadow'' HLL in which all insertion queries are handled by random sampling.
+${\mathcal{S}}$ then uses the shadow HLL for handling ${\mathcal{A}}$'s \texttt{read} and \texttt{corrupt} queries.
+Finally, we will specify ${\mathcal{S}}$ so that it outputs whatever ${\mathcal{A}}$ outputs.
+
+We will prove that  ${\mathcal{S}}$ is such that, given any adversary ${\mathcal{A}}$, any distinguisher ${\mathcal{D}}$ that can distinguish between the outputs $\textsf{out}_{\mathcal{A}}$ in the real world and $\textsf{out}_{\mathcal{S}}$ in the ideal world can be used to build an adversary ${{\mathcal{B}}}$ which breaks the PRF security of $F$.
+Under the assumption that $F$ is a good PRF, no such efficient distinguisher ${{\mathcal{D}}}$ can exist, and we may conclude that (up to PRF security) ${\mathcal{D}}$ cannot tell that it is interacting with the ideal world instead of the real world.
+
+Why does this proof imply that ${\mathcal{A}}$ cannot manipulate the cardinality estimate?
+To see this, note that (amongst many other possibilities) ${\mathcal{A}}$'s output could be the vector of bucket values $M[j]$ obtained from a \texttt{corrupt} query, enabling a specific ${\mathcal{D}}$ to compute the cardinality estimate for itself and to attempt to use that estimate as a distinguisher between the two worlds.
+Yet, by the preceding argument, the simulator in the ideal world is able to produce an output that ${\mathcal{D}}$ cannot distinguish from that of ${\mathcal{A}}$ in the real world.
+In particular, our simulator will output whatever ${\mathcal{A}}$ outputs, so if ${\mathcal{A}}$ outputs the result of a \texttt{corrupt} query in the ideal world, then so will ${\mathcal{S}}$.
+But in the ideal world, ${\mathcal{S}}$ uses random sampling for the $y$ values, so there we are operating in the setting where the average case analysis of HLL applies, as per~\cite{hll}.
+So in the ideal world this specific ${\mathcal{D}}$ would end up computing an accurate average case cardinality estimate.
+Since ${\mathcal{D}}$ tries to distinguish the two worlds by computing the cardinality estimate and we already showed that no efficient distinguisher ${{\mathcal{D}}}$ can exist, it must follow that the cardinality estimates produced in the real and ideal worlds are indistinguishable (up to PRF security).
+It then follows that ${\mathcal{A}}$ has no advantage when it comes to manipulating the HLL cardinality estimate.
+It follows that HLL, when instantiated with PRF $F$, is secure against cardinality manipulation attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
+
+%We will conclude our analysis by discussing how $F$ can be instantiated.
 
@@
 
-We consider adversary ${\mathcal{A}}$ to be an algorithm running in a security game.
-The game begins with the selection of a bit $b$ uniformly at random.
-Adversary ${\mathcal{A}}$ interacts with up to four oracles which are handled by the game; different combinations of oracles define scenarios S2, S3, S4.
-After making some queries, ${\mathcal{A}}$ returns a bit $b'$ estimating whether it is interacting with a real HLL instantiation ($b'=0$) or an ideal one ($b'=1$).
-For $i=1, 2, 3$, we define ${\mathcal{A}}$'s S$i$-Real-or-Ideal advantage against HLL to be:
-\[
-\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny HLL}}({\mathcal{A}}) := \lvert 2\Pr[b' = b] - 1 \rvert \, .
+We consider distinguisher ${\mathcal{D}}$ to be an algorithm running in a security game.
+The game begins with the selection of a bit $b$ uniformly at random and then runs either ${\mathcal{A}}$ (the real world, $b=0$) or ${\mathcal{S}}$  (the ideal world, $b=0$).
+Adversary ${\mathcal{A}}$ interacts with up to four oracles which are handled by the game using the algorithms of HLL-$F$; different combinations of oracles define scenarios S2, S3, S4.
+After making some queries, ${\mathcal{A}}$ returns its output $\textsf{out}_{\mathcal{A}}$ to ${\mathcal{D}}$.
+Simulator ${\mathcal{S}}$ runs ${\mathcal{A}}$ in a blackbox manner (as specified below) and receives ${\mathcal{A}}$'s output.
+It then provides its own output $\textsf{out}_{\mathcal{S}}$ to ${\mathcal{D}}$.
+Finally, ${\mathcal{D}}$ outputs a bit $b'$ estimating whether it is interacting with the real world ($b'=0$) or the ideal world ($b'=1$).
+For $i=2, 3, 4$, we define ${\mathcal{D}}$'s S$i$-Real-or-Ideal advantage against HLL-$F$ with adversary ${\mathcal{A}}$ and simulator ${\mathcal{S}}$ to be:
+\[
+\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{D}}) := \lvert 2\Pr[b' = b] - 1 \rvert \, .
 \]
 %Using a standard transformation, this probability can be rewritten as $\lvert \Pr[b' = 1| b = 1] - \Pr[b'=0|b=1] \rvert$.
+
+We say that the scheme HLL-$F$ is S$i$-simulation-secure with parameters $(q,t_{\mathcal{S}},t_{\mathcal{A}},t_{\mathcal{D}},\epsilon)$ if there exists a simulator $\mathcal{S}$ running in time at most $t_{\mathcal{S}}$ such that, for all adversaries ${\mathcal{A}}$ running in time at most $t_{\mathcal{A}}$ and making at most $q$ \texttt{insert($\cdot$)} queries and all distinguishers ${\mathcal{D}}$ running in time at most $t_{\mathcal{D}}$, we have
+$\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{D}}) \leq \epsilon$.
 
@@
 It also selects a key $K \leftarrow_{\$} \{0,1\}^k$ for keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
-This oracle is available in all of S2, S3, S4, but can only be called once as ${\mathcal{A}}$'s first oracle query.
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$, the oracle sets $y = F_K(x)$ and follows the procedure in Figure~\ref{fig:hll} to insert value $x$ into the HLL sketch (that is, it uses the first $m$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
+This oracle has no output.
+This oracle is available in all of S2, S3, S4.
+We assume it is only called once as ${\mathcal{A}}$'s first oracle query.
+\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$, the oracle implements HLL-$F$ by setting $y = F_K(x)$ and following the procedure in Figure~\ref{fig:hll} to insert value $x$ into the HLL sketch (that is, it uses the first $m$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
+This oracle has no output.
 This oracle is available in all of S2, S3, S4.
 Without loss of generality (since HLL and variants of it we consider here have a deterministic insertion procedure), we assume all queries $x$ made by ${\mathcal{A}}$ to \texttt{insert($\cdot$)} are distinct.
-We let $q$ denote the number of insertion queries made by ${\mathcal{A}}$ during the course of its attack.
-\item \texttt{read}: this oracle returns the HLL cardinality estimate as per Figure~\ref{fig:hll}.
+\item \texttt{read}: this oracle has no input and returns the HLL cardinality estimate as per Figure~\ref{fig:hll}.
 We allow multiple queries to this oracle and make it available to ${\mathcal{A}}$ in S3 and S4, but not S2.
-\item \texttt{corrupt}: this oracle returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents (but not any secret key $K$).
+\item \texttt{corrupt}: this oracle has no input returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents (but not any secret key $K$).
 This oracle is available only in S4.
 We allow the adversary to query \texttt{corrupt} as often as it wishes during its attack.
-Our description of S4 in Section~\ref{sec:attacks} limited the adversary to at most one such query; we allow multiple queries here because doing so does not increase the difficulty of achieving security.
+Our description of S4 in Section~\ref{sec:attacks} limited the adversary to at most one such query; here we allow multiple queries because doing so does not increase the difficulty of achieving security.
 \end{itemize}
 
-We note that in S4 (and only in S4) the adversary has access to all four oracles.
-This makes it the strongest model, and hereafter we only target security in this model.
-We count queries $q$ to \texttt{insert($\cdot$)} because this will enable us to quantify the quality of our security reduction to PRF security below.
-
-In many simulation-based security proofs, the simulator ${\mathcal{S}}$ depends on ${\mathcal{A}}$, and security is quantified via a statement of the form ``For all adversaries ${\mathcal{A}}$, there exists a simulator ${\mathcal{S}}$ such that $\ldots$'' with ${\mathcal{S}}$ being described in the proof of a specific scheme.
-In our analysis we are able to provide a universal simulator, enabling us to make a stronger statement of the form ``There exists a simulator ${\mathcal{S}}$ such that, for all adversaries ${\mathcal{A}}\ldots$''.
-We describe this simulator now, in terms of how it handles  ${\mathcal{A}}$'s oracle queries in the ideal world:
-
+We note that in S4 (and only in S4) the adversary has access to all four oracles. This makes it the strongest model, and hereafter we only target security in this model.
+
+We next describe our simulator ${\mathcal{S}}$. Recall that ${\mathcal{S}}$ has blackbox access to ${\mathcal{A}}$, so we describe how it handles ${\mathcal{A}}$'s queries: 
 \begin{itemize}
@@
 \end{itemize}
+Simulator ${\mathcal{S}}$ operates by running ${\mathcal{A}}$ and handling all of its queries as above.
+At the end of ${\mathcal{A}}$'s execution, ${\mathcal{S}}$ takes ${\mathcal{A}}$'s output and forwards it to ${\mathcal{D}}$ as its own output.
 
@@
 
-Now we are ready to state our main result:
+Now we are ready to formally state and prove our main result:
+
+\todo{pickup from here}
 
@@
 \[
- \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}}) = \mbox{Adv}^{\mbox{\tiny S4-RoI}}_{\mbox{\tiny HLL}}({\mathcal{A}}) \,.
+ \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}}) = \mbox{Adv}^{\mbox{\tiny S4-RoI}}_{\mbox{\tiny HLL}}({\mathcal{D}}) \,.
 \]
@@
 
-\subsection{Is a PRF Required?}
+\subsection{Is Universal Hashing Sufficient?}
 

### 2021-04-30 docs/HLL_exp_bad/main.tex
fixing up proof

@@
 
-\todo[inline]{This ends up being pretty harsh on this NSDI paper, but hey.}
+\todo[inline]{This ends up being pretty harsh on this NSDI paper, but hey. We may shorten or move some details to an appendix.}
 
@@
 
-\subsection{Intuition for Formal Security Analysis for S2, S3, S4}
+\subsection{Intuition for Formal Security Analysis}
 
@@
 In particular, ${\mathcal{S}}$ receives parameters $n$, $\ell$, $\alpha_n$ from ${\mathcal{A}}$'s \texttt{init($\cdot$)} call.
-Finally, ${\mathcal{S}}$ receives $\textsf{out}_{\mathcal{A}}$ and completes its execution by outputting its own output $\textsf{out}_{\mathcal{S}}$.
-
-Importantly, we will specify a construction of ${\mathcal{S}}$ that handles any of ${\mathcal{A}}$'s \texttt{insert($x$)} queries using \emph{random sampling}, that is, by setting $y \leftarrow_{\$} \{0,1\}^{n+\ell}$ instead of $y: = F_K(x)$ and then executing the rest of the insertion process as in Figure~\ref{fig:hll}.
+Finally, ${\mathcal{S}}$ receives $\textsf{out}_{\mathcal{A}}$ and completes its execution by producing its own output $\textsf{out}_{\mathcal{S}}$.
+
+Importantly, we will specify an ${\mathcal{S}}$ that handles all of ${\mathcal{A}}$'s \texttt{insert($x$)} queries using \emph{random sampling}, that is, by setting $y \leftarrow_{\$} \{0,1\}^{n+\ell}$ instead of $y: = F_K(x)$ and then executing the rest of the insertion process as in Figure~\ref{fig:hll}.
 In essence, our ${\mathcal{S}}$ will build its own ``shadow'' HLL in which all insertion queries are handled by random sampling.
@@
 
-We will prove that  ${\mathcal{S}}$ is such that, given any adversary ${\mathcal{A}}$, any distinguisher ${\mathcal{D}}$ that can distinguish between the outputs $\textsf{out}_{\mathcal{A}}$ in the real world and $\textsf{out}_{\mathcal{S}}$ in the ideal world can be used to build an adversary ${{\mathcal{B}}}$ which breaks the PRF security of $F$.
-Under the assumption that $F$ is a good PRF, no such efficient distinguisher ${{\mathcal{D}}}$ can exist, and we may conclude that (up to PRF security) ${\mathcal{D}}$ cannot tell that it is interacting with the ideal world instead of the real world.
+We will prove that  ${\mathcal{S}}$ is such that, given any adversary ${\mathcal{A}}$, then any distinguisher $\mathcal{E}$ that can distinguish between the $\textsf{out}_{\mathcal{A}}$ in the real world and $\textsf{out}_{\mathcal{S}}$ in the ideal world can be used to build an adversary ${{\mathcal{B}}}$ which breaks the PRF security of $F$.
+Under the assumption that $F$ is a good PRF, no such efficient distinguisher $\mathcal{E}$ can exist, and we may conclude that (up to PRF security) ${\mathcal{E}}$ cannot tell that it is interacting with the ideal world instead of the real world.
 
 Why does this proof imply that ${\mathcal{A}}$ cannot manipulate the cardinality estimate?
-To see this, note that (amongst many other possibilities) ${\mathcal{A}}$'s output could be the vector of bucket values $M[j]$ obtained from a \texttt{corrupt} query, enabling a specific ${\mathcal{D}}$ to compute the cardinality estimate for itself and to attempt to use that estimate as a distinguisher between the two worlds.
-Yet, by the preceding argument, the simulator in the ideal world is able to produce an output that ${\mathcal{D}}$ cannot distinguish from that of ${\mathcal{A}}$ in the real world.
+To see this, note that (amongst many other possibilities) ${\mathcal{A}}$'s output could be the vector of bucket values $M[j]$ obtained from a \texttt{corrupt} query, enabling a specific ${\mathcal{E}}$ to compute the cardinality estimate for itself and to attempt to use that estimate as a distinguisher between the two worlds.
+Yet, by the preceding argument, the simulator in the ideal world is able to produce an output that ${\mathcal{E}}$ cannot distinguish from that of ${\mathcal{A}}$ in the real world.
 In particular, our simulator will output whatever ${\mathcal{A}}$ outputs, so if ${\mathcal{A}}$ outputs the result of a \texttt{corrupt} query in the ideal world, then so will ${\mathcal{S}}$.
-But in the ideal world, ${\mathcal{S}}$ uses random sampling for the $y$ values, so there we are operating in the setting where the average case analysis of HLL applies, as per~\cite{hll}.
-So in the ideal world this specific ${\mathcal{D}}$ would end up computing an accurate average case cardinality estimate.
-Since ${\mathcal{D}}$ tries to distinguish the two worlds by computing the cardinality estimate and we already showed that no efficient distinguisher ${{\mathcal{D}}}$ can exist, it must follow that the cardinality estimates produced in the real and ideal worlds are indistinguishable (up to PRF security).
-It then follows that ${\mathcal{A}}$ has no advantage when it comes to manipulating the HLL cardinality estimate.
-It follows that HLL, when instantiated with PRF $F$, is secure against cardinality manipulation attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
+But in the ideal world, ${\mathcal{S}}$ uses random sampling for the $y$ values, so there ${\mathcal{S}}$ (and ${\mathcal{A}}$ that it runs) are operating in the setting where the average case analysis of HLL applies, as per~\cite{hll}.
+So in the ideal world this specific ${\mathcal{E}}$ would end up computing an accurate average case cardinality estimate.
+Since ${\mathcal{E}}$ tries to distinguish the two worlds by computing the cardinality estimate and we already showed that no efficient distinguisher ${{\mathcal{E}}}$ can exist, it must follow that the cardinality estimates produced in the real and ideal worlds are indistinguishable (up to PRF security).
+It then follows that ${\mathcal{A}}$ has no advantage when it comes to manipulating the HLL-$F$ cardinality estimate.
+So HLL, when instantiated with PRF $F$, is secure against cardinality manipulation attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
 
@@
 
-We consider distinguisher ${\mathcal{D}}$ to be an algorithm running in a security game.
-The game begins with the selection of a bit $b$ uniformly at random and then runs either ${\mathcal{A}}$ (the real world, $b=0$) or ${\mathcal{S}}$  (the ideal world, $b=0$).
+We now formalise the above intuition.
+We consider distinguisher ${\mathcal{E}}$ to be an algorithm running in a security game.
+The game begins with the selection of a bit $b$ uniformly at random and then runs either ${\mathcal{A}}$ (the real world, $b=0$) or ${\mathcal{S}}$ (the ideal world, $b=1$).
 Adversary ${\mathcal{A}}$ interacts with up to four oracles which are handled by the game using the algorithms of HLL-$F$; different combinations of oracles define scenarios S2, S3, S4.
-After making some queries, ${\mathcal{A}}$ returns its output $\textsf{out}_{\mathcal{A}}$ to ${\mathcal{D}}$.
+After making some queries, ${\mathcal{A}}$ returns its output $\textsf{out}_{\mathcal{A}}$ to ${\mathcal{E}}$.
 Simulator ${\mathcal{S}}$ runs ${\mathcal{A}}$ in a blackbox manner (as specified below) and receives ${\mathcal{A}}$'s output.
-It then provides its own output $\textsf{out}_{\mathcal{S}}$ to ${\mathcal{D}}$.
-Finally, ${\mathcal{D}}$ outputs a bit $b'$ estimating whether it is interacting with the real world ($b'=0$) or the ideal world ($b'=1$).
-For $i=2, 3, 4$, we define ${\mathcal{D}}$'s S$i$-Real-or-Ideal advantage against HLL-$F$ with adversary ${\mathcal{A}}$ and simulator ${\mathcal{S}}$ to be:
-\[
-\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{D}}) := \lvert 2\Pr[b' = b] - 1 \rvert \, .
+It then provides its own output $\textsf{out}_{\mathcal{S}}$ to ${\mathcal{E}}$.
+Finally, ${\mathcal{E}}$ outputs a bit $b'$ estimating whether it is interacting with the real world ($b'=0$) or the ideal world ($b'=1$).
+For $i=2, 3, 4$, we define ${\mathcal{E}}$'s S$i$-Real-or-Ideal advantage against HLL-$F$ with adversary ${\mathcal{A}}$ and simulator ${\mathcal{S}}$ to be:
+\[
+\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{E}}) := \lvert 2\Pr[b' = b] - 1 \rvert \, .
 \]
@@
 
-We say that the scheme HLL-$F$ is S$i$-simulation-secure with parameters $(q,t_{\mathcal{S}},t_{\mathcal{A}},t_{\mathcal{D}},\epsilon)$ if there exists a simulator $\mathcal{S}$ running in time at most $t_{\mathcal{S}}$ such that, for all adversaries ${\mathcal{A}}$ running in time at most $t_{\mathcal{A}}$ and making at most $q$ \texttt{insert($\cdot$)} queries and all distinguishers ${\mathcal{D}}$ running in time at most $t_{\mathcal{D}}$, we have
-$\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{D}}) \leq \epsilon$.
+We say that the scheme HLL-$F$ is S$i$-simulation-secure with parameters $(q,t_{\mathcal{S}},t_{\mathcal{A}},t_{\mathcal{E}},\epsilon)$ if there exists a simulator $\mathcal{S}$ running in time at most $t_{\mathcal{S}}$ such that, for all adversaries ${\mathcal{A}}$ running in time at most $t_{\mathcal{A}}$ and making at most $q$ \texttt{insert($\cdot$)} queries and all distinguishers ${\mathcal{E}}$ running in time at most $t_{\mathcal{E}}$, we have
+$\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{E}}) \leq \epsilon$.
+We quantify over \texttt{insert($\cdot$)} queries (via $q$) and not other types of query because only \texttt{insert($\cdot$)} queries will appear in our concrete security analysis and we do not want to clutter the notation.
 
@@
 \begin{itemize}
-\item \texttt{init($\cdot$)}: this oracle receives as input the HLL parameters ($n$, $\ell$, $\alpha_n$) and performs initialisation of the HLL sketch.
-It also selects a key $K \leftarrow_{\$} \{0,1\}^k$ for keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
+\item \texttt{init($\cdot$)}: this oracle receives as input the HLL-$F$ parameters ($n$, $\ell$, $\alpha_n$) and performs initialisation of the HLL-$F$ sketch, including generating the array of buckets $M[j]$ and initialising all the entries to $0$, and selecting a key $K \leftarrow_{\$} \{0,1\}^k$ for keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
 This oracle has no output.
-This oracle is available in all of S2, S3, S4.
+It is available in all of S2, S3, S4.
 We assume it is only called once as ${\mathcal{A}}$'s first oracle query.
@@
 This oracle has no output.
-This oracle is available in all of S2, S3, S4.
+It is available in all of S2, S3, S4.
 Without loss of generality (since HLL and variants of it we consider here have a deterministic insertion procedure), we assume all queries $x$ made by ${\mathcal{A}}$ to \texttt{insert($\cdot$)} are distinct.
-\item \texttt{read}: this oracle has no input and returns the HLL cardinality estimate as per Figure~\ref{fig:hll}.
+\item \texttt{read}: this oracle has no input and returns the HLL-$F$ cardinality estimate as per Figure~\ref{fig:hll}, i.e.\ it computes a raw cardinality estimate via $E := \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$ and then performs corrections for small and large values.
 We allow multiple queries to this oracle and make it available to ${\mathcal{A}}$ in S3 and S4, but not S2.
-\item \texttt{corrupt}: this oracle has no input returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents (but not any secret key $K$).
+\item \texttt{corrupt}: this oracle has no input and returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents (but not any secret key $K$).
 This oracle is available only in S4.
@@
 
-We next describe our simulator ${\mathcal{S}}$. Recall that ${\mathcal{S}}$ has blackbox access to ${\mathcal{A}}$, so we describe how it handles ${\mathcal{A}}$'s queries: 
+We next describe the simulator ${\mathcal{S}}$ for the ideal world. Simulator ${\mathcal{S}}$ operates by running ${\mathcal{A}}$ in a blackbox manner and handling all of its queries as follows: 
 \begin{itemize}
@@
 \end{itemize}
-Simulator ${\mathcal{S}}$ operates by running ${\mathcal{A}}$ and handling all of its queries as above.
-At the end of ${\mathcal{A}}$'s execution, ${\mathcal{S}}$ takes ${\mathcal{A}}$'s output and forwards it to ${\mathcal{D}}$ as its own output.
+At the end of ${\mathcal{A}}$'s execution, ${\mathcal{S}}$ takes ${\mathcal{A}}$'s output and forwards it to ${\mathcal{E}}$ as its own output.
+Notice that the running time of  ${\mathcal{S}}$ is (roughly) equal to that of ${\mathcal{A}}$.
 
@@
 Now we are ready to formally state and prove our main result:
-
-\todo{pickup from here}
-
 \begin{theorem}\label{thm:security-of-HLL-with-PRF}
 Let simulator $\mathcal{S}$ be as described above.
-Suppose ${\mathcal{A}}$ is any S4 adversary against HLL instantiated with a keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
-Suppose ${\mathcal{A}}$ either plays in the real world or in the ideal world with $\mathcal{S}$.
-Then we can construct from ${\mathcal{A}}$ a PRF adversary ${{\mathcal{B}}}$ against $F$ such that:
-\[
- \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}}) = \mbox{Adv}^{\mbox{\tiny S4-RoI}}_{\mbox{\tiny HLL}}({\mathcal{D}}) \,.
+For any S4 adversary ${\mathcal{A}}$ against HLL-$F$ and any distinguisher $\mathcal{E}$, we can construct a PRF adversary ${{\mathcal{B}}}$ against $F$ such that:
+\[
+ \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}}) = 
+ \mbox{Adv}^{\mbox{\tiny S$4$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{E}}) \,.
 \]
 Moreover, if ${\mathcal{A}}$ makes $q$ queries to its \texttt{insert($\cdot$)} oracle, then ${{\mathcal{B}}}$ makes $q$ queries to its real or random oracle.
-Finally, the running time of ${{\mathcal{B}}}$ is (roughly) the same as that of ${\mathcal{A}}$.
+Finally, the running time of ${{\mathcal{B}}}$ is (roughly) $t_{\mathcal{A}}+t_{\mathcal{E}}$.
 \end{theorem}
@@
 PRF adversary ${{\mathcal{B}}}$ against $F$ is built as follows.
-Let $d$ denote the hidden bit determining whether ${{\mathcal{B}}}$ receives real responses of the form $F_K(x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${{\mathcal{B}}}$ runs S4 adversary ${\mathcal{A}}$ and responds to each of its queries as follows:
+Let $d$ denote the hidden bit determining whether ${{\mathcal{B}}}$ receives real responses of the form $F_K(x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${{\mathcal{B}}}$ provides a simulation of either the real or random world for distinguisher ${\mathcal{E}}$ as follows.
+${{\mathcal{B}}}$ runs ${{\mathcal{A}}}$ and answers each of ${{\mathcal{A}}}$'s queries as follows:
 \begin{itemize}
-\item \texttt{init}: here ${{\mathcal{B}}}$ initialises an HLL sketch in accordance with Figure~\ref{fig:hll}. Thus it initialises an array $M[j]$ in which each entry is set to 0. 
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$,  ${{\mathcal{B}}}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${{\mathcal{B}}}$ uses $y$ to update the HLL bucket $j$ generated via $j=1 + <y_1y_2...y_{n}>_2$ as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
-\item \texttt{read}: ${{\mathcal{B}}}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate in accordance with Figure~\ref{fig:hll}. 
-\item \texttt{corrupt}: ${{\mathcal{B}}}$ returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents. 
+\item \texttt{init($\cdot$)}: ${{\mathcal{B}}}$ receives parameters $n$, $\ell$, $\alpha_m$ and initialises an array $M[j]$ in which each of the $2^n$ entries is set to 0. 
+\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$,  ${{\mathcal{B}}}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${{\mathcal{B}}}$ uses $y$ to update the HLL sketch as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
+That is, it uses the first $m$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$.
+\item \texttt{read}: ${{\mathcal{B}}}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate as per Figure~\ref{fig:hll}. 
+\item \texttt{corrupt}: ${{\mathcal{B}}}$ returns the vector $(M[j])_{j=1}^{m}$. 
 \end{itemize}
-When ${\mathcal{A}}$ outputs a bit $b'$, ${{\mathcal{B}}}$ outputs its own bit $d'=b'$.
+When ${\mathcal{A}}$ returns its final output $\textsf{out}_{\mathcal{S}}$, ${{\mathcal{B}}}$ forwards this output to ${{\mathcal{E}}}$.
+Finally, ${{\mathcal{E}}}$ outputs a bit $b'$.
+Then ${{\mathcal{B}}}$ outputs its own bit $d'=b'$.
+
 Note that when $d=0$, ${{\mathcal{B}}}$ plays against the real $F$ with a random key $K$ and provides a simulation for ${\mathcal{A}}$ that perfectly instantiates the real world.
-On the other hand, when $d=1$, ${{\mathcal{B}}}$ plays against a truly random function and provides a simulation for ${\mathcal{A}}$ that perfectly instantiates the ideal world with simulator $\mathcal{S}$ (to see this, note that sampling the outputs $y$ of a truly random function at distinct inputs $x$ is identical to simply sampling $y$ uniformly at random).
-The equality of advantages in the statement of the theorem now follows immediately.
-The remainder of the proof follows on counting queries and noting that the HLL algorithm for computing cardinality is efficient, so ${{\mathcal{B}}}$'s running time due to handling \texttt{read} queries is not increased significantly compared to that of ${\mathcal{A}}$.
+On the other hand, when $d=1$, ${{\mathcal{B}}}$ plays against a truly random function and perfectly instantiates our simulator $\mathcal{S}$ in the ideal world.
+To see this, note that sampling the outputs $y$ of a truly random function at distinct inputs $x$ is identical to simply sampling $y$ uniformly at random.\footnote{Our proof relies on the $x$ being distinct, but this is just to simplify book-keeping and is without loss of generality.} The equality of advantages in the statement of the theorem now follows immediately.
+The remainder of the proof follows on counting queries and noting that the HLL sub-algorithms for insertion and computing cardinality are efficient, so ${{\mathcal{B}}}$'s running time due to handling ${\mathcal{A}}$'s queries is not increased significantly compared to that of ${\mathcal{A}}$.
+The final running time for ${\mathcal{B}}$, namely $t_{\mathcal{A}}+t_{\mathcal{E}}$, follows on noting that ${\mathcal{B}}$ first runs ${\mathcal{A}}$ and then ${\mathcal{E}}$.
 \end{proof}
-
-We reiterate our central point: the theorem shows that no efficient S4 adversary ${\mathcal{A}}$ can tell that it is playing against our simulator ${\mathcal{S}}$ using truly random values $y$ instead of with values obtained from a PRF, except with a small advantage.
-This means that ${\mathcal{A}}$ is effectively reduced to attacking HLL in the exact same setting used in the average case analysis of HLL from~\cite{hll}, wherein it is proved that the HLL cardinality estimate endures relative errors of the form $1.04/\sqrt{m}$.
-From this we can conclude that no efficient S4 adversary ${\mathcal{A}}$ can manipulate the HLL cardinality estimate to its benefit.
 
@@
 Rather, we infer from inspecting the simulator what the adversary can (or cannot) achieve in the ideal world, and show that no efficient adversary can distinguish the real and ideal worlds.
-This stands in contrast to the game-based approach of~\cite{CCS:ClaPatShrPS19} which needs a complex definition to say when the adversary has successfully manipulated a probabilistic data structure.
+This is in contrast to the game-based approach of~\cite{CCS:ClaPatShrPS19} which needs a complex definition to say when the adversary has successfully manipulated a probabilistic data structure.
 This reveals a strength of the simulation-based approach in more complex settings like ours.
@@
 If the two cardinality estimates are different, the adversary can conclude definitively that $x$ was \emph{not} previously inserted into the HLL; if they are the same, it can estimate (with some error) that $x$ was previously inserted.\footnote{Errors can arise in two ways when $x$ was not previously inserted: either inserting $x$ changes a bucket value, but the change is masked by rounding when the cardinality estimate is computed, or inserting $x$ does lead to a change in a bucket value because $\rho(w)$ computed from $x$ is too small.} This attack still works even if HLL is implemented using a PRF and is unavoidable in view of the oracle queries (and in real settings, the API) available to the adversary.
-So such privacy leaks are an inherent feature of HLL if a suitably rich API is provided.
+So such privacy leaks are an inherent feature of HLL if a rich enough API is provided.
 Note that this attack does not contradict our main result, which concerns prevention of cardinality manipulation rather than confidentiality.
 
-Still, simulation-based security notions are often used to model confidentiality properties, so it is tempting to think that our security definition might say something about confidentiality of the HLL inputs $x$.
+%Still, simulation-based security notions are often used to model confidentiality properties, so it is tempting to think that our security definition might say something about confidentiality of the HLL inputs $x$.
 Our existing model does not, since ${\mathcal{A}}$ chooses all the inputs $x$ and so ``knows'' them already.
@@
 
-In this case, the simulator ${\mathcal{S}}$ in the ideal world can proceed without ever seeing any inputs but instead knowing only the distribution $p$ and the number of distinct samples drawn from it in the first phase.
+%In this case, the simulator ${\mathcal{S}}$ in the ideal world can proceed without ever seeing any inputs but instead knowing only the distribution $p$ and the number of distinct samples drawn from it in the first phase.
 It is not hard to modify the proof of Theorem~\ref{thm:security-of-HLL-with-PRF} to show that no adversary can distinguish real and ideal worlds in this case (up to PRF security).
@@
 
-We leave the further exploration of how to adapt our simulation-based security definitions to capture confidentiality notions to future work.
+We leave the exploration of the adaptation of our simulation-based security definitions to capture confidentiality notions for more restrictred APIs to future work.
 

### 2021-04-30 docs/HLL_exp_bad/main.tex
work on intro

@@
 \section{Introduction}
-\todo{more here introducing big data and why we want statistical estimators on them; data streams; cannot store, or large data sets, want summary statistics.
-Such things are supported by GBQ, Apache Spark, Redis.
-Cardinality as a key metric; HLL as the algorithm of choice.
-''analytics engines for big data processing.''}
-Naive approaches to computing the cardinality of large data sets, such as sorting the elements or simply maintaining the set of unique elements seen, are impractical at scale and in streaming settings.
-The HyperLogLog algorithm (HLL henceforth)~\cite{hll} is today the most widely used cardinality estimator.
-It is increasingly used in settings where adversaries may have  incentives to manipulate the estimate made by the algorithm.
+Data scientists today make use of systems like Google Big Query, Apache Spark, Presto and Redis to handle large data sets.
+Such systems include analytics engines capable of performing powerful operations on the data.
+Particularly useful to data scientists are summary statistics, with a key statistic being the number of distinct data items in a data set, i.e.\ the data set's cardinality.
+In many use cases, obtaining an approximate cardinality is sufficient.
+Over the last decade, the HyperLogLog (HLL) cardinality estimator~\cite{hll} has emerged as the algorithm of choice for this task, due to its excellent speed, low memory requirements and high proven accuracy.
+It is also suitable for use in applications where it is not possible (or desirable) to store an entire data set, but instead where the data set is presented in a streaming fashion.
+HLL is used by all of the above-mentioned systems; section~\ref{sec:overview} contains a complete specification of HLL.
+
+However, HLL is increasingly used in settings where adversaries may have incentives to manipulate the cardinality estimates made by the algorithm, i.e.\ HLL is being used in adversarial settings.
 For example:
 \begin{itemize}
-\item Google and counting stuff.\todo{add more detail here -- their blogs, their paper, the use in Google Big Query.} 
     \item Cardinality estimation provides a significantly faster way for Facebook to compute statistics on their users.
 They use HLL to find out how many distinct people visited their website in the past week~\cite{fbhll}.
+    \item Google uses a variant of HLL described in~\cite{hllpractice} in its PowerDrill system (first described in~\cite{HallBBGN12}) and possibly in a variety of other internal systems reported in~\cite{hllpractice}.
+For example, cardinality estimation is used to to determine the number of distinct search queries on \url{google.com} over a time period~\cite{hllpractice}.
+HLL is also used for cardinality estimation via an API in Google Big Query.
+    \item HLL has been implemented in network soft-switches to approximate the number of distinct packets in traffic flows and overcome the switch memory limits~\cite{flexswitch}.
+The number of unique flows traversing a switch is then used to  inform a congestion control mechanism.
     \item HLL has been proposed as a solution to detecting Denial-of-Service attacks~\cite{portscanhll} (this paper actually uses a variation on HLL called \emph{sliding HLL}~\cite{slidinghll} which estimates the cardinality over a moving time window).
-Specifically, HLL is used to count how many different ports are used as destination port over all packets received, as an attempt to identify when a port scan attack is underway.
-    \item HLL has been implemented in network soft-switches to approximate the number of distinct packets in traffic flows and overcome the switch resource limits~\cite{flexswitch}.
-The number of unique flows traversing a switch can then be used to provide a better congestion control protocol.
+Specifically, HLL is used to count how many different destination ports are seen over all packets received, in an attempt to identify when a port scan attack is underway.
 \end{itemize}
 
-Regardless of the sensitivity of the above examples, none of the above-cited sources are clear about their threat model, especially for the HLL component.
-They are all potentially vulnerable to the attacks we present in this paper.
+None of the above-cited examples discuss any kind of threat model where an adversary may try to manipulate the HLL cardinality estimate to its advantage, for example, through input selection.
+This is despite them all operating in potentially adversarial settings.
+We infer that the developers of these systems may not be fully aware of the risks they are taking in relying on HLL.
 
@@
 We demonstrate this theoretically and experimentally for the ``classic'' version of HLL.
-We also study a specific ``modern'' version of HLL which adopts recommendation from~\cite{hllnew} and show how variants of our attacks apply there too.
+We also study a specific ``modern'' version of HLL which adopts recommendations from~\cite{hllnew} and show how variants of our attacks apply there too.
 This modern version of HLL is interesting because it is the one used in Redis, a popular key-value database technology.
+
+We stress that none of our attacks are particularly deep in cryptanalytic terms.
+Yet HLL is being increasingly widely deployed, and never with protections against our attacks (insofar as we are aware).
+So despite their technical simplicity, we argue that our attacks are important in establishing the insecurity of HLL in adversarial setting.
+We present our attacks in an abstract way rather than tying them to any specific setting in order to illustrate their generality.
+We do give a more detailed exploration of their impact on the data-centre network monitoring system proposed in~\cite{flexswitch}, see Section~\ref{sec:exp}.
 
@@
 We show that in the ``shadow device'' attack scenario of~\cite{hllvuln}, no security is possible even if HLL is secretly keyed.
-On the other hand, we also show that in the most powerful attack setting introduced here, where the attacker has full access to the internals of the HLL algorithm (except for a secret key!), it is possible to provably secure HLL against cardinality manipulation by the simple expedient of replacing its internal hash function by a keyed, variable-input-length pseudo-random function (VIL-PRF).
+On the other hand, we also show that in the most powerful attack setting introduced here, where the attacker has full access to the internals of the HLL algorithm (except for a secret key!), it is possible to provably secure HLL against cardinality manipulation by the simple expedient of replacing its internal hash function $h$ by a keyed, variable-input-length pseudo-random function (VIL-PRF) $F$.
 We discuss and benchmark a suitable choice for the required VIL-PRF.
-
-\todo[inline]{make the model a contribution... ``To achieve this... introduce simulation-based security model for cardinality estimators'', etc}
-\todo[inline]{Simplicity of attacks as a strength; only complexity comes in S1 attack which we don't consider that interesting!}
-\todo[inline]{No e2e attack on a specific system; rather paper is written in general form since the attacks apply broadly.}
-\todo[inline]{Highlight the analysis of OSDI paper~\cite{flexswitch} as an example of the potential of the attacks.}
+To support our formal analysis, we introduce a novel simulation-based security model for HLL.
+We believe our model and approach is of independent interest and will be useful beyond the analysis of cardinality manipulation attacks, for example when studying privacy properties of HLL and other probabilistic data structures.
 
@@
 Section~\ref{sec:redis} considers the security of the more modern version of HLL due to Ertl~\cite{hllnew} that is used in Redis.
-Our experimental results are discussed and related back to specific use cases in Section~\ref{sec:exp}.
+Our experimental results are discussed and related back to a specific application setting in Section~\ref{sec:exp}.
 Section~\ref{sec:formal} provides our formal security analysis of HLL.
-Section~\ref{sec:conclusions} provides our conclusions and some ideas for future work.
+Section~\ref{sec:conclusions} provides our conclusions and ideas for future work.
 
@@
 
-So to roughly estimate the cardinality of a stream, we need only to store a representation of $k$, the length of the largest observed string of leading zero-bits for items in the stream.
+So to roughly estimate the cardinality of a stream, we need only to store a representation of $k$, the length of the largest observed string of leading zero-bits for hashes of items in the stream.
 This provides a very compact mechanism for estimating a stream's cardinality -- if the true cardinality is $N$, then just $\mathcal{O}(\log\log(N))$ bits of storage are needed to compute the cardinality estimate.
@@
 Estimates from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha_m$, where $\alpha_m$ is empirically computed to correct a systematic multiplicative bias.
+See Figure~\ref{fig:hll}.
 
@@
 In addition, the cardinality estimate is usually rounded to a whole number before being output.
-The original HLL algorithm from~\cite{hll} is shown in Fig.~\ref{fig:hll}.
+The original HLL algorithm from~\cite{hll} is shown in Figure~\ref{fig:hll}.
 This algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 4\%, using only 1.5 kB of memory.
@@
 \subsection{Attacker's Objective, Abstractly}
-Throughout, we assume the adversary has the ability to vary the contents of free fields in the strings which will be inserted into the stream of values.
+Throughout, we assume the adversary has the ability to vary the contents of free fields in the strings which will be inserted into the stream of items.
 For example it might manipulate IPID and other header fields in the context of IP packets.
@@
 
-In this setting, where the target HLL sketch is initially non-empty, we see that an adversary who has sufficient flexibility in its inputs (as represented by the size of the set ${\mathcal{X}}$) can expect to insert many values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the case where the HLL sketch has some empty buckets in S2).
+In this setting, where the target HLL sketch is initially non-empty, we see that an adversary who has sufficient flexibility in its inputs (as represented by the size of the set ${\mathcal{X}}$) can expect to insert many items into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the case where the HLL sketch has some empty buckets in S2).
 The result is largely independent of the HLL parameters.
@@
 One can attempt to use the shadow device to identify items that would result in small entries $M[j]$ (for unknown indices $j$) since such items are unlikely to increase the cardinality estimate of the target device by much.
-This can be done by first filling the shadow device with sufficiently many values to escape from the ``small range correction'' phase.
-Then one could enter candidate values $x$ into the shadow device and check for cardinality increases.
+This can be done by first filling the shadow device with sufficiently many items to escape from the ``small range correction'' phase.
+Then one could enter candidate items $x$ into the shadow device and check for cardinality increases.
 However this runs into the rounding issue described in Section~\ref{sec:scenarios-discussion}.
 Still, one could retain those items which do lead to a detectable change in cardinality and reject the others (and also estimate the value of $M[j]$ for the unknown bucket $j$ from the magnitude of the change, retaining only those for which $M[j]$ is small).
-But here the lack of a cardinality change could result either from rounding issues or because insertion of candidate value $x$ into the shadow device really did lead to no increase in cardinality.
+But here the lack of a cardinality change could result either from rounding issues or because insertion of candidate item $x$ into the shadow device really did lead to no increase in cardinality.
 This indicates that  ``good'' items would be rejected unnecessarily.
@@
 One can attempt to use the shadow device to identify items that would result in small entries $M[j]$ (for unknown indices $j$) since such items are unlikely to increase the cardinality estimate of the target device by much.
-This requires first filling the shadow device with sufficiently many values to escape from the ``small range correction'' phase.
-Then one could enter random values $x$ into the shadow device and check for cardinality increases.
+This requires first filling the shadow device with sufficiently many items to escape from the ``small range correction'' phase.
+Then one could enter random items $x$ into the shadow device and check for cardinality increases.
 However this runs into the rounding issue described in Section~\ref{sec:scenarios-discussion}.
@@
 \]
-Now we can proceed exactly as in Sections~\ref{sec:nonemptyHLLS2} and~~\ref{sec:nonemptyHLLS4}, inserting values $x$ whose hashes have rightmost $\ell$ bits beginning with a ``1'' bit (and with a more relaxed condition in S4).
+Now we can proceed exactly as in Sections~\ref{sec:nonemptyHLLS2} and~~\ref{sec:nonemptyHLLS4}, inserting items $x$ whose hashes have rightmost $\ell$ bits beginning with a ``1'' bit (and with a more relaxed condition in S4).
 These insertions do not alter the values of the $C_k$, and hence $\hat{\lambda}$ is unaffected.
@@
 We first begin by explaining the high-level idea of our approach.
-We allow an adversary ${\mathcal{A}}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios S2, S3 and S4: it may have an \texttt{insert($\cdot$)} oracle allowing it to add values to the HLL sketch, a \texttt{read} oracle that returns the HLL cardinality estimate, and a \texttt{corrupt} that returns all the HLL bucket contents (but not any secret key $K$); in S4 it has all three oracles.
+We allow an adversary ${\mathcal{A}}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios S2, S3 and S4: it may have an \texttt{insert($\cdot$)} oracle allowing it to add items to the HLL sketch, a \texttt{read} oracle that returns the HLL cardinality estimate, and a \texttt{corrupt} that returns all the HLL bucket contents (but not any secret key $K$); in S4 it has all three oracles.
 In addition, in all 3 scenarios, ${\mathcal{A}}$ has access to an \texttt{init($\cdot$)} oracle by which it sets the HLL parameters $n$, $\ell$, $\alpha_n$ used to initialise the HLL sketch.
@@
 We assume it is only called once as ${\mathcal{A}}$'s first oracle query.
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$, the oracle implements HLL-$F$ by setting $y = F_K(x)$ and following the procedure in Figure~\ref{fig:hll} to insert value $x$ into the HLL sketch (that is, it uses the first $m$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
+\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$, the oracle implements HLL-$F$ by setting $y = F_K(x)$ and following the procedure in Figure~\ref{fig:hll} to insert $x$ into the HLL sketch (that is, it uses the first $m$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
 This oracle has no output.
@@
 
-An HLL can leak information about values $x$ that were previously inserted.
-Consider the following attack from~\cite{cardestprivacy}: to test whether a value $x$ was previously inserted into the HLL by an honest user (say), the adversary makes the sequence of queries: \texttt{read}, \texttt{insert($x$)}, \texttt{read}.
+An HLL can leak information about items $x$ that were previously inserted.
+Consider the following attack from~\cite{cardestprivacy}: to test whether an item $x$ was previously inserted into the HLL by an honest user (say), the adversary makes the sequence of queries: \texttt{read}, \texttt{insert($x$)}, \texttt{read}.
 If the two cardinality estimates are different, the adversary can conclude definitively that $x$ was \emph{not} previously inserted into the HLL; if they are the same, it can estimate (with some error) that $x$ was previously inserted.\footnote{Errors can arise in two ways when $x$ was not previously inserted: either inserting $x$ changes a bucket value, but the change is masked by rounding when the cardinality estimate is computed, or inserting $x$ does lead to a change in a bucket value because $\rho(w)$ computed from $x$ is too small.} This attack still works even if HLL is implemented using a PRF and is unavoidable in view of the oracle queries (and in real settings, the API) available to the adversary.
@@
 However, it is easily modified to provide a confidentiality notion.
-For example, we can consider the insertion of honestly generated values $x$ that are drawn from some probability distribution $p$ on $\mathcal{D}$ in a first phase, with ${\mathcal{A}}$ then having access to only \texttt{read} and \texttt{corrupt} oracles in a second phase (we have to remove the \texttt{insert($\cdot$)} oracle in view of the general attack from~\cite{cardestprivacy}).
+For example, we can consider the insertion of honestly generated items $x$ that are drawn from some probability distribution $p$ on $\mathcal{D}$ in a first phase, with ${\mathcal{A}}$ then having access to only \texttt{read} and \texttt{corrupt} oracles in a second phase (we have to remove the \texttt{insert($\cdot$)} oracle in view of the general attack from~\cite{cardestprivacy}).
 
@@
 
-An S4 adversary $\mathcal{A}$ can perform a series of pairs of \texttt{insert($x^i$)} and \texttt{corrupt} queries on chosen values $x^0, x^1, \ldots, x^{t-1}$.
+An S4 adversary $\mathcal{A}$ can perform a series of pairs of \texttt{insert($x^i$)} and \texttt{corrupt} queries on chosen items $x^0, x^1, \ldots, x^{t-1}$.
 With high probability (assuming $t$ is not too large), each \texttt{insert} query hits a fresh bucket in the HLL sketch.
@@
 
-By choosing $x^i$ to have 256 bits, we have $v=2$ and $p_{x^i}(K) = K^2 + x^i_1K+x^i_2$; then further selecting $x^i$ such that  $x^i_2$ is a constant value (e.g.\ $0 \in \textsf{GF}(2^{128})$) and considering consecutive pairs of $x_i$ values, we obtain equations of the form:
+By choosing $x^i$ to have 256 bits, we have $v=2$ and $p_{x^i}(K) = K^2 + x^i_1K+x^i_2$; then further selecting $x^i$ such that  $x^i_2$ is a constant value (e.g.\ $0 \in \textsf{GF}(2^{128})$) and considering consecutive pairs of $x_i$ items, we obtain equations of the form:
 \[
@@
 
-%Going further, the S4 adversary has complete control over the values $x_i$, so it can choose them so that the polynomials $p_{x_i}$ are of degree 1 over $\textsf{GF}(2^{128})$.
+%Going further, the S4 adversary has complete control over the items $x_i$, so it can choose them so that the polynomials $p_{x_i}$ are of degree 1 over $\textsf{GF}(2^{128})$.
 Then $K$ can be determined by writing everything bit-wise and solving a linear system.
@@
 
-\todo{Sim based of indepednent interest, arguably more powerful than game-based notions with fixed winning predicates.}
+%\todo{Sim based of indepednent interest, arguably more powerful than game-based notions with fixed winning predicates.}
 %It is also an open problem to close the performance gap that exists between fast but non-cryptographic hashes used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security.

### 2021-04-30 docs/HLL_exp_bad/main.tex
work on intro

@@
 On the other hand, when $d=1$, ${{\mathcal{B}}}$ plays against a truly random function and perfectly instantiates our simulator $\mathcal{S}$ in the ideal world.
-To see this, note that sampling the outputs $y$ of a truly random function at distinct inputs $x$ is identical to simply sampling $y$ uniformly at random.\footnote{Our proof relies on the $x$ being distinct, but this is just to simplify book-keeping and is without loss of generality.} The equality of advantages in the statement of the theorem now follows immediately.
+To see this, note that sampling the outputs $y$ of a truly random function at distinct inputs $x$ is identical to simply sampling $y$ uniformly at random.\footnote{Our proof relies on the $x$ being distinct, but this is to simplify book-keeping and is without loss of generality.} The equality of advantages in the statement of the theorem now follows immediately.
 The remainder of the proof follows on counting queries and noting that the HLL sub-algorithms for insertion and computing cardinality are efficient, so ${{\mathcal{B}}}$'s running time due to handling ${\mathcal{A}}$'s queries is not increased significantly compared to that of ${\mathcal{A}}$.

### 2021-04-30 docs/HLL_exp_bad/main.tex
Updates from Overleaf

--- 2021-04-30 docs/HLL_exp_bad/main.tex
@@
 \acmversiontrue
-
-\newif\ifanon % turn on ACM style (can be short or full)
-\anontrue
 
@@
 \begin{document}
+%\IEEEtitleabstractindextext{%
 
 \begin{abstract}
-Computing the count of distinct elements in large data sets is a common task but naive approaches are memory-expensive.
-The HyperLogLog (HLL) algorithm (Flajolet \emph{et al.}, 2007) estimates a data set's cardinality while using significantly less memory than a naive approach, at the cost of some accuracy.
+Computing the count of distinct elements in massive data sets is a common task but naive approaches are memory-expensive.
+The HyperLogLog (HLL) algorithm (Flajolet \emph{et al.}, 2007) estimates a data set's cardinality while using significantly less memory than a naive approach at the cost of some accuracy.
 This trade-off makes the HLL algorithm very attractive for a wide range of applications such as database management and network monitoring, where an exact count may not be needed.
@@
 \end{abstract}
+%}
 
 \maketitle
+%\IEEEdisplaynontitleabstractindextext
 
 \section{Introduction}
-\todo{more here introducing big data and why we want statistical estimators on them; data streams; cannot store, or large data sets, want summary statistics.
-Such things are supported by GBQ, Apache Spark, Redis.
-Cardinality as a key metric; HLL as the algorithm of choice.
-''analytics engines for big data processing.''}
 Naive approaches to computing the cardinality of large data sets, such as sorting the elements or simply maintaining the set of unique elements seen, are impractical at scale and in streaming settings.
@@
 
-Recently, Reviriego and Ting~\cite{hllvuln} initiated work on the performance of HLL in adversarial settings.
-They showed that in very specific attack settings, the cardinality estimate made by HLL can be modified through input selection.
+Very recently, Reviriego and Ting~\cite{hllvuln} initiated work on the performance of HLL in adversarial settings.
+They showed that in very specific attack settings, the cardinality estimate made by HLL could be modified through input selection.
 In this paper, we present a complete analysis of HLL in adversarial settings, considering more realistic attack scenarios and giving more powerful attacks than~\cite{hllvuln}.
-In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of ${\mathcal{O}}(2^t)$ where $t$ is the number of bits of flexibility in the adversary's input).
+In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of ${\cal O}(2^t)$ where $t$ is the number of bits of flexibility in the adversary's input).
 We demonstrate this theoretically and experimentally for the ``classic'' version of HLL.
@@
 We discuss and benchmark a suitable choice for the required VIL-PRF.
-
-\todo[inline]{make the model a contribution... ``To achieve this... introduce simulation-based security model for cardinality estimators'', etc}
-\todo[inline]{Simplicity of attacks as a strength; only complexity comes in S1 attack which we don't consider that interesting!}
-\todo[inline]{No e2e attack on a specific system; rather paper is written in general form since the attacks apply broadly.}
-\todo[inline]{Highlight the analysis of OSDI paper~\cite{flexswitch} as an example of the potential of the attacks.}
-
-\subsection{Paper Organisation}
+\todo{make the model a contribution...
+``To achieve this...
+introduce simulation-based security model for cardinality estimators'', etc}
+
+\todo{Simplicity of attacks as a strength; only complexity comes in S1 attack which we don't consider that interesting!}
+\todo{No e2e attack on a specific system; rather paper is written in general form since the attacks apply broadly.}
+
 The rest of the paper is organised as follows.
@@
 Desfontaines \emph{et al.}~\cite{cardestprivacy} studied the privacy properties of HLL in two different attack scenarios, referred to as \emph{insider} and \emph{external}.
-The attack target of~\cite{cardestprivacy} is different from ours, being focused on privacy breaches.
+The attack target of~\cite{cardestprivacy} is different from ours, being focused on privacy breaches, whereas we care about manipulating cardinality estimation.
 %It is argued in~\cite{cardestprivacy} that HLL can be used by organisations to store and process location data, which is inherently sensitive, so losing privacy would leak users' locations and potentially cause great harm.
-In contrast, we are concerned with an attacker attempting to manipulate the cardinality estimation process itself (an issue not considered in~\cite{cardestprivacy}).
 We discuss the two attack scenarios from~\cite{cardestprivacy} in more detail in Section~\ref{sec:attacks}.
-%\todo{update: we will get privacy for free from our sim based security model if sim does not use input $x$.}
+\todo{update: we will get privacy for free from our sim based security model if sim does not use input $x$.}
 
@@
 For completeness, we include analyses of this attack scenario, showing that we can drastically improve on the attacks in~\cite{hllvuln} (e.g.\ we can reduce the HLL cardinality estimate by a factor of almost 1,000 instead of the factor of 5 achieved in~\cite{hllvuln}).
-However, we stress that we consider this scenario to be the least realistic (and therefore the least interesting).
 
@@
 We show in Section~\ref{sec:formal} that when a salt is used appropriately (i.e.\ as the secret key for a pseudo-random function that is used in place of hash function $h$), security against adversarial inputs can be achieved even in our strongest adversarial setting.
-%In addition, our simulation-based security definition, in which the simulator in the ideal world has no access to the inserted values, automatically guarantees confidentiality of the values inserted into the HLL sketch, even in the strongest attack model considered in~\cite{cardestprivacy}.
-So, while our main objective is to protect against adversarial inputs, we also get privacy for free when using a keyed PRF in place of the HLL hash function.
+\todo{Add a bit more detail here.}
+%This might not be a convenient solution for some applications, they thus encourage further research.
 
@@
 Clayton \emph{et al.}~\cite{CCS:ClaPatShrPS19} recently provided a provable-security treatment of this topic.
-They used a game-based formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures.
-They did not study HLL.
-Our use of a simulation-based security definition in place of the game-based ones used in~\cite{CCS:ClaPatShrPS19} is another point of distinction; see further discussion in Section~\ref{sec:formal}.
-
+They used their formalism to analyze Bloom filters, counting (Bloom) filters and count-min sketch data structures, but not HLL.
 Prior work also focused on Bloom filters~\cite{NaorY15,GerbetKL15} or flooding of hash tables~\cite{LiptonN93,CrosbyW03}.
@@
 
+%It is an interesting open problem to use their framework to develop a formal understanding of how HLL can be made secure in the adversarial setting.
+
 \section{Overview of HLL}\label{sec:overview}
+\subsection{The Algorithm}
+HLL~\cite{hll} is a streaming algorithm based on the key observation that, for a stream of randomly distributed values represented as bit-strings of some fixed length, if we observe a value with the maximum of $k$ leading zero-bits, then the cardinality of the stream (i.e.\ the number of distinct values it contains) is likely to be on the order of $2^{k+1}$.
+To ensure a uniform distribution, the stream values are first processed by a hash function $h$ before the leading bits are inspected.
+So to estimate the cardinality of a stream, we need only to store a representation of $k$, the length of the largest observed string of leading zero-bits for numbers in the stream.
+This provides a very compact mechanism for estimating the stream cardinality -- if the true cardinality is $N$, then just $\log\log(N)$ bits are needed to store the estimate.
+
+This simple approach suffers from large variance in the estimation.
+In order to improve the estimate, we can use many estimators in parallel instead of one, and average the results.
+Each estimate is stored in its own register, that we call a bucket.
+We refer to the collection of buckets and their contents as the HLL \emph{sketch}.
+To map a value $x$ to one of $m=2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$.
+The output length of $h(\cdot)$ is thus $n+\ell$, fixed to 32 in~\cite{hll}.
+Estimates from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha_m$, where $\alpha_m$ is empirically computed to correct a systematic multiplicative bias.
+
+Relatively small and big cardinalities are handled separately to correct for systematic distortions, as shown in Fig.~\ref{fig:hll}.
+When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate, looking at the number of empty buckets (whose register value is 0) instead of the number of leading zeros in the stream.
+A similar correction is applied for large cardinalities in order to take into account the likelihood of hash collisions.
+In addition, the cardinality estimate is usually rounded to a whole number before being output.
+
+HLL as described in Figure~\ref{fig:hll} is able to estimate cardinalities greater than $10^9$ with a typical standard error of 2\%, using only 1.5 kB of memory~\cite{hll}.
+
+Some works building on~\cite{hll} replace the Hit Counting algorithm by other schemes and use a hash function with a larger output domain to reduce the chances of collision and thus get rid of the need for large cardinality range correction.
+For example, \cite{hllpractice} uses a 64-bit hash function and Linear Counting along with bias correction in the small cardinality range, while~\cite{hllnew} proposes new estimators based on a more precise statistical analysis of the counting problem.
+In this report we mostly target the HLL formulation in the original paper~\cite{hll} as represented in Fig.~\ref{fig:hll} and as widely used in practice.
+We also study~\cite{hllnew} because of its recent deployment in Redis.
+We leave for future work the analysis of further HLL variants.
+
+An interesting property of HLL is that it supports merging in a lossless way.
+To combine two buckets at index $i$ in two different HLL instances, one can take the maximum of the two bucket entries and assign that to the matching bucket $i$ in the merged HLL sketch.
+Such a simple set union operation allows easy parallelisation of operations among multiple machines independently, provided they use the same hash function and the same number of buckets.
+Clearly this procedure cannot be carried out if the different HLL sketches are differently seeded or keyed.
 
@@
 \vspace{0.5em}
-\textbf{Algorithm} HyperLogLog (\textbf{input} ${\mathcal{X}}$ : multiset of items from domain $\mathcal{D}$).\\
+\textbf{Algorithm} HyperLogLog (\textbf{input} ${\cal X}$ : multiset of items from domain $\mathcal{D}$).\\
 \textbf{initialize} a collection of $m$ registers, $M[1], ..., M[m]$, to $0$;\\
 \vspace{0.5em}
-\textbf{for} $x \in {\mathcal{X}}$ \textbf{do}\\
+\textbf{for} $x \in \cal X$ \textbf{do}\\
     \hspace{2em} \textbf{set} $y := h(x);$\\
@@
 
-HLL~\cite{hll} is based on the key observation that, for a stream of items represented as bit-strings of some fixed length, if all items have at most $k$ leading zero-bits (and one or more items actually have this many leading zero-bits), then the cardinality of the stream (i.e.\ the number of distinct values it contains) is likely to be on the order of $2^{k+1}$.
-To ensure a uniform distribution on the items' representations as bit-strings, the items are processed by a hash function $h$ before the leading bits are inspected.
-
-So to roughly estimate the cardinality of a stream, we need only to store a representation of $k$, the length of the largest observed string of leading zero-bits for items in the stream.
-This provides a very compact mechanism for estimating a stream's cardinality -- if the true cardinality is $N$, then just $\mathcal{O}(\log\log(N))$ bits of storage are needed to compute the cardinality estimate.
-
-This simple approach suffers from large variance.
-In order to reduce the variance, HLL uses many estimators in parallel instead of one and averages the results.
-Each estimate is stored in its own register that we call a bucket.
-We refer to the collection of buckets and their contents as the HLL \emph{sketch}.
-To map a value $x$ to one of $m=2^n$ buckets, \cite{loglog} suggests using the first $n$ bits of $h(x)$ as a bucket index, and then computing the longest sequence of leading zero bits on the remaining $\ell$ bits of $h(x)$.
-The output length of $h(\cdot)$ is thus $n+\ell$.
-This sum is fixed to 32 in the original HLL paper~\cite{hll}.
-Estimates from all buckets are averaged using the harmonic mean and scaled by a constant $\alpha_m$, where $\alpha_m$ is empirically computed to correct a systematic multiplicative bias.
-
-In the original HLL proposal~\cite{hll}, relatively small and large cardinalities are handled separately to correct for systematic errors.
-When the raw cardinality estimate is below $\frac{5}{2}m$, the Hit Counting algorithm of Whang \emph{et al.}~\cite{hitcounting} is used to produce the final cardinality estimate.
-This estimator uses the number of empty buckets (whose register value is 0) instead of the number of leading zeros in the stream.
-A similar correction is applied for large cardinalities in order to take into account the likelihood of hash collisions.
-In addition, the cardinality estimate is usually rounded to a whole number before being output.
-The original HLL algorithm from~\cite{hll} is shown in Fig.~\ref{fig:hll}.
-This algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 4\%, using only 1.5 kB of memory.
-Crucially, the proof of this~\cite[Theorem 1]{hll} makes the assumption that the inputs are drawn independently and uniformly at random from the input domain.\footnote{Technically, the results of~\cite{hll} are proven for a distribution obtained by making arbitrary replications and permutations after sampling independently and uniformly, to simulate input taken from an \emph{ideal multiset}.
-But they also hold without replication and permutation because HLL's internal state is invariant under repeated inputs.}
-
-Some works building on~\cite{hll} replace the Hit Counting algorithm by other schemes and use a hash function with a larger output domain to reduce the chances of collision and thus get rid of the need for large cardinality range correction.
-For example, Google engineers proposed using a 64-bit hash function and Linear Counting along with bias correction in the small cardinality range~\cite{hllpractice},\footnote{See also the Google blogpost describing the implementation in Google Big Query at~\url{https://cloud.google.com/blog/products/gcp/counting-uniques-faster-in-bigquery-with-hyperloglog}.} while Redis has adopted a new estimator based on a more precise statistical analysis of the counting problem provided by~\cite{hllnew}.
-
-In this paper we mostly focus on the HLL formulation in the original paper~\cite{hll} as represented in Fig.~\ref{fig:hll} and as widely used in practice.
-We also study~\cite{hllnew} because of its recent deployment in Redis.
-We leave for future work the analysis of further HLL variants.
-\todo{revise this if we add Google BQ.}
-
-An interesting property of HLL is that it supports merging in a lossless way.
-To combine two buckets at index $i$ in two different HLL instances, one can take the maximum of the two bucket entries and assign that to the matching bucket $i$ in the merged HLL sketch.
-Such a simple set union operation allows easy parallelisation of operations among multiple machines independently, provided they use the same hash function and the same number of buckets.
-Clearly this procedure cannot be carried out if the different HLL sketches are differently seeded or keyed.
-
 \section{HLL in Adversarial Settings}\label{sec:attacks}
 
-This section presents the different adversarial settings that we consider.
-In all settings, our adversary's aim is to manipulate the HLL cardinality estimate.
-We focus on reducing the estimate as much as possible, since this would cause more damage when  considering the previously mentioned applications of HLL.
+This section presents different adversarial settings and corresponding attacks that manipulate the HLL cardinality estimate.
+The adversary's goal is to reduce the estimate as much as possible, since this would cause more damage when  considering the above-mentioned applications of HLL.
 It is easy to modify our attack strategies for an adversary who instead wishes to artificially inflate the HLL cardinality estimate given a limited input capability.
-%We also analyze the impact on the HLL cardinality estimate and the computational effort required.
-
-\subsection{Attacker's Objective, Abstractly}
-Throughout, we assume the adversary has the ability to vary the contents of free fields in the strings which will be inserted into the stream of values.
-For example it might manipulate IPID and other header fields in the context of IP packets.
-We model this in a general way by assuming the adversary has access to a large set ${\mathcal{X}}$ of potential inputs from which it can select items to insert into the HLL sketch.
-Our adversary's objective, then, is to insert as many items from ${\mathcal{X}}$ as possible whilst keeping the HLL cardinality estimate as low as possible.
+We also analyze the impact on the HLL cardinality estimate and the computational effort required.
+
+Throughout, we assume the adversary has the ability to vary the contents of free fields in the input strings which will be inserted into the stream of values.
+For example it can manipulate IPID and other header fields in the context of IP packets.
+We model this by assuming the adversary has access to a large set ${\cal X}$ of potential inputs from which it can select items to insert into the HLL sketch.
+Our adversary's objective, then, is to insert as many items from ${\cal X}$ as possible whilst keeping the HLL cardinality estimate as low as possible.
 
@@
 This would make any pre-computation on the items using the shadow device irrelevant, and considerably weaken the attack presented in Section 3.2 of~\cite{hllvuln}.
-Hence we assume for the rest of this work that, in scenario S1, shadow and targeted devices have the same internals and security parameters (if there are any).\footnote{This scenario is roughly analogous to the common situation where a fixed private key is stored in multiple devices acting as TLS termination points for a single web site.}
+Hence we assume for the rest of this work that, in scenario S1, shadow and targeted devices have the same internals and security parameters (if there are any).
 %If the HLL device receives inputs from several users, the attack from Section 3.2 of~\cite{hllvuln} is then the easiest to carry out.
@@
 
-%\todo{What did referee say about other attack scenarios we should consider?}
+\todo{What did referee say about other attack scenarios we should consider?}
 
@@
 For each case, we consider adversaries operating in scenarios S1, S2 and S4.
-We operate in the situation where the adversary has a set $\mathcal{X}$ of potential inputs which it can choose to insert into the HLL sketch, and where it wishes to insert as many of them as it can whilst keeping the HLL cardinality estimate as low as possible.
-
-\subsection{A Simple Attack}\label{sec:simple}
-
-In scenarios S2 and S4 with an initially empty HLL sketch, there is a very simple attack based on simply evaluating $h(x)$ for each $x \in \mathcal{X}$, looking at the bits of $h(x)$ to see which bucket $j$ would be affected if $x$ was inserted, and then deciding whether to insert or not based on some rule.
-
-For example, the rule could be to insert $x$ if and only if doing so does not increase the entry $M[j]$ stored in the relevant bucket beyond some pre-defined threshold value $v$.
-A simple computation based on the HLL raw cardinality estimator function in Fig.~\ref{fig:hll} shows that this rule would allow the insertion of a fraction of  $1-2^{-v}$ of the set $\mathcal{X}$ whilst keeping the raw cardinality estimate below $\alpha_m m 2^v$.
-This attack assumes $h$ behaves like a random function and exploits the geometric distribution of the position of the leading ``1'' bit in $w$, where $w = y_{n+1}y_{n+2}\ldots y_{n+1}$ and $y=h(x)$ (as in the HLL algorithm).
-
-This simple attack is already very effective.
-For example, setting $v=1$ in the above rule, the attacker can insert half of all elements of $\mathcal{X}$ whilst keeping the cardinality estimate below $2\alpha_m m$ (where, recall, $m$ is the fixed number of buckets in the HLL sketch).
-
-In the remainder of this section, we target attacks that are more effective than this simple attack.
+
+\todo{Provide a brief sketch here of the ``obvious attack'', and then say that we target improvements over this attack.}
+\todo{remove $t$ parameter or use consistently.}
 
@@
 \begin{lemma}\label{lem:bound}
-Suppose an adversary inserts elements $x$ from some set ${\mathcal{X}}$ into an initially empty HLL sketch.
+Suppose an adversary inserts elements $x$ from some set ${\cal X}$ into an initially empty HLL sketch.
 Suppose the adversary can guarantee that in its attack, only a subset of size $B$ out of $m$ buckets are hit, where $B \leq B_{\max}$ with
@@
 Now our proposed attack is very simple.
-Suppose the adversary has some set ${\mathcal{X}}$ of items from which to pick its inputs.
+Suppose the adversary has some set ${\cal X}$ of $2^t$ items from which to pick its inputs.
 The adversary selects a value of $B \leq B_{\max}$, and defines its target buckets to be those numbered $1,\ldots,B$.
-Since the adversary knows $h$, it can tell whether, for each $x \in {\mathcal{X}}$, insertion of $x$ into the HLL sketch would hit one of the target buckets or not; it inserts those that do.
-In expectation, assuming $h$ behaves like a random function, the adversary will successfully insert a fraction $B/m$ of the items, totalling $(B/m)\cdot |{\mathcal{X}}|$ items.
+Since the adversary knows $h$, it can tell whether, for each $x \in {\cal X}$, insertion of $x$ into the HLL sketch would hit one of the target buckets or not; it inserts those that do.
+In expectation, assuming $h$ behaves like a random function, the adversary will successfully insert a fraction $B/m$ of the items, totalling $(B/m)\cdot 2^t = 2^{t-n+\log_2(B)}$ items.
 Meanwhile the final cardinality estimate will remain below $m \cdot \ln (1+\frac{B}{m-B})$.
 
-For example, setting $B = m/2$ (which, for practical parameters, satisfies $B \leq B_{\max}$), we can expect to insert half of all the items in $\mathcal{X}$ while keeping the cardinality estimate below $(\ln 2)m \approx 0.693m$.
-This is superior to the simple attack in Section~\ref{sec:simple} (which, recall, could insert half of $\mathcal{X}$ while keeping the cardinality estimate below $2\alpha_m m \approx 1.44m$).
-
-As a second example, setting $B = 1$, we can expect to insert $|{\mathcal{X}}|/2^m$ items with a final cardinality estimate equal to $m \cdot \ln (1+\frac{1}{m-1})$, which is rounded to $1$ by the HLL algorithm (for $m$ of practical interest, e.g.\ $m=256$).
-Thus we keep the HLL cardinality estimate to a constant (1!) whilst being able to insert a constant fraction of the items from the input set ${\mathcal{X}}$.
+For example, setting $B = m/2$ (which, for practical parameters, satisfies $B \leq B_{\max}$), we can expect to insert $2^{t-1}$ items while keeping the cardinality estimate below $m \ln 2$.
+
+As a second example, setting $B = 1$, we can expect to insert $2^{t-n}$ items with a final cardinality estimate equal to $m \cdot \ln (1+\frac{1}{m-1})$, which is rounded to $1$ by the HLL algorithm (for $m$ of practical interest, e.g.\ $m=256$).
+Thus we keep the HLL cardinality estimate to a constant (1!) whilst being able to insert exponentially many items, provided there is sufficient flexibility in the input set ${\cal X}$.
 
@@
 \paragraph{Core Idea:}
-We assume the adversary again has (or is given) some set ${\mathcal{X}}$ of items from which to pick its inputs.
-The core idea of our attack is to use the first $B \leq B_{\max}$ items $x_1,\ldots, x_B$ from ${\mathcal{X}}$ to define a set of target buckets, and then to insert into the shadow HLL sketch further items $x \in {\mathcal{X}}$, checking for a cardinality estimate increase on each insertion.
-We initialise a list ${\mathcal{L}}$ containing $x_1,\ldots, x_B$.
-We add those items that do not increase the cardinality estimate on insertion into the shadow HLL sketch to ${\mathcal{L}}$ and reject the rest.
-We reset the shadow device when the cardinality estimate has increased too much, reinserting the first $B$ items from ${\mathcal{X}}$ again to re-establish a ``baseline'' of target buckets for the shadow HLL sketch.
+We assume the adversary again has (or is given) some set ${\cal X}$ of $2^t$ items from which to pick its inputs.
+The core idea of our attack is to use the first $B \leq B_{\max}$ items $x_1,\ldots, x_B$ from ${\cal X}$ to define a set of target buckets, and then to insert into the shadow HLL sketch further items $x \in {\cal X}$, checking for a cardinality estimate increase on each insertion.
+We initialise a list ${\cal L}$ containing $x_1,\ldots, x_B$.
+We add those items that do not increase the cardinality estimate on insertion into the shadow HLL sketch to ${\cal L}$ and reject the rest.
+We reset the shadow device when the cardinality estimate has increased too much, reinserting the first $B$ items from ${\cal X}$ again to re-establish a ``baseline'' of target buckets for the shadow HLL sketch.
 A key observation is that if in the attack we keep the number of ``active'' buckets to values $B^\prime$ that are always below $B_{\max}$, then the Hit Counting algorithm is used to compute the cardinality estimate; given its output $m \cdot \ln (m/V)$ (or, more exactly, a rounded version of this value) we can infer the exact value of $V$, and hence $B^\prime$.
 This enables precise detection of cardinality increases, provided we perform a reset before the Hit Counting algorithm stops being used (this is ensured by resetting when the estimate indicates that $B_{\max}$ buckets have been hit).
-After all of ${\mathcal{X}}$ has been processed as described, we take the constructed list ${\mathcal{L}}$ and reprocess it in the same way.
+After all of ${\cal X}$ has been processed as described, we take the constructed list ${\cal L}$ and reprocess it in the same way.
 We must ensure that the same set of initial items $x_1,\ldots, x_B$ is inserted into the shadow HLL sketch before any other items, so as to fix the same set of target buckets.
@@
 By repeated reprocessing, we will gradually remove items that do not hit the $B$ target buckets (defined by $x_1,\ldots, x_B$).
-Eventually, we obtain a final list ${\mathcal{L}}^*$ that we insert into the target HLL sketch.
+Eventually, we obtain a final list ${\cal L}^*$ that we insert into the target HLL sketch.
 
 By construction and from Lemma~\ref{lem:bound}, we are guaranteed that the cardinality estimate for the target HLL sketch will be at most $m \cdot \ln (m/V)$ where $V = m-B$.
-Assuming $h$ behaves like a random function, it is clear that the expected size of the final list ${\mathcal{L}}^*$ is $\frac{B}{m}\cdot |{\mathcal{X}}|$.
-Thus the attack's final performance is the same as that in scenarios S2 and S4 above.
-However, more processing and resets are required (and the number of resets should be considered as something to be minimised since resetting the shadow device may take a certain amount of time).
+Assuming $h$ behaves like a random function, it is clear that the expected size of the final list ${\cal L}^*$ is $\frac{B}{m}\cdot |{\cal X}|$.
+Thus the attack's final performance is the same as that in scenarios S2 and S4 above (but more processing and resets are required).
 
 \paragraph{Simple Attack:}
-The simple version of our attack uses the above idea, inserting items $x$ into the shadow HLL sketch but resetting it as soon as the cardinality estimate increases from $B$ to $B+1$ (indicating that a non-target bucket has been hit by item $x$).
-Acting in this way precisely determines whether each inserted item hits the target buckets or not, so has no ``false positives'' which need to be eliminated through reprocessing of the list ${\mathcal{L}}$.
-This means that the attack can be carried out in a streaming fashion in a single pass over ${\mathcal{X}}$.
-Assuming $h$ behaves like a random function, we must perform a reset with probability $\frac{m-B}{m} = 1 - \frac{B}{m}$ per item inserted, and hence in total an expected number of $(1 - \frac{B}{m})\cdot | \mathcal{X} |$ resets.
+The simplistic version of our attack uses the above idea, inserting items $x$ into the shadow HLL sketch but resetting it as soon as the cardinality estimate increases from $B$ to $B+1$ (indicating that a non-target bucket has been hit by item $x$).
+Acting in this way precisely determines whether each inserted item hits the target buckets or not, so has no ``false positives'' which need to be eliminated through reprocessing of the list ${\cal L}$.
+This means that the attack can be carried out in a streaming fashion in a single pass over ${\cal X}$.
+Assuming $h$ behaves like a random function, we must perform a reset with probability $\frac{m-B}{m} = 1 - \frac{B}{m}$ per item inserted, and hence in total an expected number of $(1 - \frac{B}{m})\cdot 2^t$ resets.
 
@@
 However, when the number of active buckets has reached some value $B^\prime \leq B_{\max}$ this means that an item $x$ may not increase the cardinality estimate, yet still not hit one of the $B$ target buckets (instead it hits one of the other $B^\prime - B$ active buckets).
-Such an $x$ acts as a ``false positive" and needs to be eliminated through reprocessing of the list ${\mathcal{L}}$.
+Such an $x$ acts as a ``false positive" and needs to be eliminated through reprocessing of the list ${\cal L}$.
 
 \paragraph{Analysis of the Complex Attack:}
-%Suppose that after each reset, the order of the items in ${\mathcal{X}}$ (or ${\mathcal{L}}$) that have not yet been processed on the current pass is randomised.
+%Suppose that after each reset, the order of the items in ${\cal X}$ (or ${\cal L}$) that have not yet been processed on the current pass is randomised.
 (Here we exclude the special items  $x_1,\ldots, x_B$ from the reordering.) Then
@@
 
-Now we expect to reject in total $(1 - \frac{B}{m})\cdot | \mathcal{X} |$ items, and each one needs an expected number of trial insertions that is bounded by $m/(m - B_{\max} + B)$. 
+Now we expect to reject in total $(1 - \frac{B}{m})\cdot 2^t$ items, and each one needs an expected number of trial insertions that is bounded by $m/(m - B_{\max} + B)$. 
 %(In fact, we can model the number of trial insertions needed as an independent geometric distribution with parameter $q$ for each such $x$.) 
-Meanwhile, each of the $\frac{B}{m}\cdot | \mathcal{X} |$ items in the final insertion list ${\mathcal{L}}^*$ for the target HLL sketch is trial inserted into the shadow HLL sketch on every pass.
+Meanwhile, each of the $\frac{B}{m}\cdot 2^t$ items in the final insertion list ${\cal L}^*$ for the target HLL sketch is trial inserted into the shadow HLL sketch on every pass.
 
@@
 \[
-N_R:= \frac{1 - \frac{B}{m}}{B_{\max} - B}\cdot | \mathcal{X} |.
+N_R:= \frac{1 - \frac{B}{m}}{B_{\max} - B}\cdot 2^t.
 \]
@@
 \[
-N_I \leq \frac{m}{ m-B_{\max} + B} \cdot \left(1 - \frac{B}{m} \right)\cdot | \mathcal{X} | +   N_R \cdot \frac{B}{m}\cdot | \mathcal{X} |.
-\]
-
-As a concrete illustration, we take $B = m/2$ and $B_{\max} = 0.71\cdot m$, to obtain $N_R = (2.38/m) \cdot | \mathcal{X} |$ and $N_I \leq 0.63 \cdot | \mathcal{X} | + 1.19\cdot | \mathcal{X} |^2/m$.
-Here the dominant term in $N_I$ is ${\mathcal{O}}(| \mathcal{X} |^2)$, i.e.\ the total insertion cost of this attack in S1 grows as the square of the size of the starting set ${\mathcal{X}}$ (instead of linearly as in S2 and S4).
-The end result of the attack is that, as in S2 and S4, an expected $| \mathcal{X} |/2$ items can be inserted into the target HLL sketch while keeping its cardinality estimate below $m \ln 2$.
+N_I \leq \frac{m}{ m-B_{\max} + B} \cdot \left(1 - \frac{B}{m} \right)\cdot 2^t +   N_R \cdot \frac{B}{m}\cdot 2^t.
+\]
+
+As a concrete illustration, we take $B = m/2$ and $B_{\max} = 0.71\cdot m$, to obtain $N_R = (2.38/m) \cdot 2^t$ and $N_I \leq 0.63 \cdot 2^t + 1.19\cdot 2^{2t}/m$.
+Here the dominant term in $N_I$ is ${\cal O}(2^{2t})$, i.e.\ the total insertion cost grows as the square of the size of the starting set ${\cal X}$.
+The result of the attack is that, as in S2 and S4, an expected $2^{t-1}$ items can be inserted into the target HLL sketch while keeping its cardinality estimate below $m \ln 2$.
 
@@
 Note that, in a given pass, when $B^\prime$ is still close to $B$, an item $x$ that does not increase the number of active buckets is more likely to have hit one of the target buckets than is the case when $B^\prime$ is large; such an $x$ can be tried earlier in the sequence of trials in the next pass and its exact status determined sooner.
-Any $x$ that can be determined to be a true positive (because its insertion into the shadow HLL sketch keeps the number of active buckets at exactly $B$) can be placed immediately in ${\mathcal{L}}^*$ and not reprocessed any further.
+Any $x$ that can be determined to be a true positive (because its insertion into the shadow HLL sketch keeps the number of active buckets at exactly $B$) can be placed immediately in ${\cal L}^*$ and not reprocessed any further.
 We leave the analysis of this optimisation to future work.
@@
 %The adversary blindly picks the first $b$ buckets to be hit by simply inserting items into the shadow HLL until the cardinality estimate indicates that $b$ buckets have been touched.
-It puts these items in a final target list ${\mathcal{L}}$.
+It puts these items in a final target list ${\cal L}$.
 It continues to insert items, discarding all items whose insertion causes the cardinality estimate for the shadow HLL sketch to increase from $b+k$ to $b+k+1$ for $k \in \mathbb{N}$ (here we allow some slack in the total number of buckets that have been hit in order to reduce the number of resets of the shadow device that are required in the attack).
 In other words, when $b+k$ buckets have already been hit and the insertion of an item causes the cardinality estimate to increase further, the adversary can discard this item as it definitely knows that a new bucket (that is not one of the $b$ targeted buckets) switched to a non-empty state.
-On the other hand, it adds to the list ${\mathcal{L}}$ any item that does \emph{not} cause the cardinality estimate to increase.
+On the other hand, it adds to the list ${\cal L}$ any item that does \emph{not} cause the cardinality estimate to increase.
 Such items may have hit one of the initial $b$ buckets, or one of the later buckets; at this stage the adversary cannot be certain.
@@
 %As soon as the cardinality estimate exceeds $\frac{5}{2}m$, the adversary can assume that the shadow HLL sketch's output corresponds to the raw HLL cardinality estimate with high probability.
-In this case, no more useful information can be inferred about the bucket mappings in the shadow HLL sketch, so the adversary resets the shadow HLL sketch to the empty state and starts again with the list ${\mathcal{L}}$ of non-discarded items obtained so far.
-Reinserting these items will (eventually) determine whether they hit the original $b$ buckets or later buckets, and hence whether they should be retained in ${\mathcal{L}}$ or discarded.
-Once ${\mathcal{L}}$ is reprocessed (and assuming the cardinality estimate does not exceed $\frac{5}{2}m$), the adversary can start to insert fresh items, as previously described.
+In this case, no more useful information can be inferred about the bucket mappings in the shadow HLL sketch, so the adversary resets the shadow HLL sketch to the empty state and starts again with the list ${\cal L}$ of non-discarded items obtained so far.
+Reinserting these items will (eventually) determine whether they hit the original $b$ buckets or later buckets, and hence whether they should be retained in ${\cal L}$ or discarded.
+Once ${\cal L}$ is reprocessed (and assuming the cardinality estimate does not exceed $\frac{5}{2}m$), the adversary can start to insert fresh items, as previously described.
 The adversary must be careful not to shuffle the non-discarded items in between iterations, as the $b$ buckets to be filled are blindly determined by the first inserted items, and must stay the same throughout the attack.
 
-%After sufficiently many iterations of this procedure, the adversary will have created a list of items ${\mathcal{L}}$ such that, when inserted into the \emph{target} HLL, create a sketch in which $b$ target buckets are filled and the others left empty.
+%After sufficiently many iterations of this procedure, the adversary will have created a list of items ${\cal L}$ such that, when inserted into the \emph{target} HLL, create a sketch in which $b$ target buckets are filled and the others left empty.
 With an appropriate choice of $b$ (as discussed immediately below), the final cardinality estimate for the target HLL sketch will be $E^*= m \cdot \ln (m/V)$ where $V = m-b$ is the number of empty buckets.
@@
 We will exploit this in the sequel in scenarios S2 and S4.
-For now, we note that setting, say, $b=128$ with $n = 8$ and $\ell = 24$ would result in a final cardinality estimate of $256 \ln 2$ in the above attack in Scenario S1, irrespective of how long the list ${\mathcal{L}}$ of inserted items is.
+For now, we note that setting, say, $b=128$ with $n = 8$ and $\ell = 24$ would result in a final cardinality estimate of $256 \ln 2$ in the above attack in Scenario S1, irrespective of how long the list ${\cal L}$ of inserted items is.
 Moreover, setting $b=128$ would mean that half of all items available to the adversary could be inserted into the target HLL.
@@
 This is highly likely to be the case (with probability at least $1-1/m^2$) as long as at least $m\ln m$ items have already been inserted into the target HLL.
-This follows from the standard analysis of the coupon collector problem~\cite[Section 3.6]{MR95} under our usual assumption that $h$ behaves as a random function.
+This follows from the standard analysis of the coupon collector problem\todo{add a reference for this} under our usual assumption that $h$ behaves as a random function.
 In S4, the adversary can use its snapshot to detect if this is the case or not, and behave differently if not (details for this case follow below).
@@
 
-%We assume the adversary has (or is given) some set of items ${\mathcal{X}}$ that it wishes to insert; as usual, we write $|{\mathcal{X}}| = 2^t$.
+%We assume the adversary has (or is given) some set of items ${\cal X}$ that it wishes to insert; as usual, we write $|{\cal X}| = 2^t$.
 
 \subsubsection{Initially Non-empty HLL Sketch, Scenario S2:}\label{sec:nonemptyHLLS2}
-The adversary knows $h$ so it selects items $x$ from ${\mathcal{X}}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit.
+The adversary knows $h$ so it selects items $x$ from ${\cal X}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit.
 In this case, since the relevant bucket $j$ already has $M[j] \ge 1$, the bucket's counter is not updated, and the cardinality estimate of the HLL sketch remains unchanged.
 
-The adversary then expects to be able to insert half of the items from ${\mathcal{X}}$ into the HLL sketch, without increasing the HLL cardinality estimate at all.
+The adversary then expects to be able to insert half of the items from ${\cal X}$ into the HLL sketch, without increasing the HLL cardinality estimate at all.
 
@@
 \subsubsection{Initially Non-empty HLL Sketch, Scenario S4:}\label{sec:nonemptyHLLS4}
-The adversary can adopt the same strategy as for S2, but relax the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch).
+The adversary can adopt the same strategy as for S2, but make relax the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch).
 Specifically, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
@@
 
-We now compute the expected number of elements $x$ from ${\mathcal{X}}$ such that $h(x)$ meets the requirements of the attack.  %We assume that $M[j] \ge 0$ for all buckets $j$. 
+We now compute the expected number of elements $x$ from ${\cal X}$ such that $h(x)$ meets the attack's requirements.  %We assume that $M[j] \ge 0$ for all buckets $j$. 
 Assume $h(x)$ is mapped to bucket $j$.
 Inserting $x$ does not increase the cardinality estimate if the rightmost $\ell$ bits of $h(x)$ have $M[j]-1$ or fewer leading zeroes; this probability is equal to $1-\frac{1}{2^{M[j]}}$ assuming $h$ behaves like a random function.
-Averaging over all $m$ buckets (and using the fact that the bucket choice is uniformly random since the bits of $h$ are independent and uniformly random), the probability that input $x$ meets the requirements of the attack is given by:
+Averaging over all $m$ buckets (and using the fact that the bucket choice is uniformly random since the bits of $h$ are independent and uniformly random), the probability that input $x$ meets the attack's requirements is given by:
 \begin{eqnarray*}
@@
 \end{eqnarray*}
-where $H_M$ is the harmonic mean of $2^{M[1]}, 2^{M[2]}, \ldots, 2^{M[m]}$.
-So the adversary can expect to insert a fraction $(1 - H_M^{-1})$ of the items from ${\mathcal{X}}$ without increasing the HLL cardinality estimate at all.
-
-%If some buckets are actually empty (and in S4 the adversary knows which ones using its snapshot), then the above analysis can be reused in order to avoid touching those buckets altogether when inserting elements from ${\mathcal{X}}$.
+where $H_M$ is the harmonic mean of the counts $2^{M[1]}, 2^{M[2]}, \ldots, 2^{M[m]}$.
+So, given ${\cal X}$ containing $2^t$ items, the adversary can expect to insert $(1 - H_M^{-1}) \cdot 2^t$ items from ${\cal X}$ without increasing the HLL cardinality estimate at all.
+
+%If some buckets are actually empty (and in S4 the adversary knows which ones using its snapshot), then the above analysis can be reused in order to avoid touching those buckets altogether when inserting elements from ${\cal X}$.
 The above analysis does not change at all because the term $\frac{1}{m} \cdot (1-\frac{1}{2^{M[j]}})$ already signs a probability of 0 to inserting $x$ in bucket $j$ if $M[j] = 0$.
-Hence we still expect to insert $(1 - H_M^{-1}) \cdot 2^t$ items from ${\mathcal{X}}$ without increasing the HLL cardinality estimate at all.
+Hence we still expect to insert $(1 - H_M^{-1}) \cdot 2^t$ items from ${\cal X}$ without increasing the HLL cardinality estimate at all.
 Of course, the adversary can relax its attack and allow some of the empty buckets to be hit as well.
@@
 
-In this setting, where the target HLL sketch is initially non-empty, we see that an adversary who has sufficient flexibility in its inputs (as represented by the size of the set ${\mathcal{X}}$) can expect to insert many values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the case where the HLL sketch has some empty buckets in S2).
+In this setting, where the target HLL sketch is initially non-empty, we see that an adversary who has sufficient flexibility in its inputs (as represented by the size $2^t$ of the set ${\cal X}$) can expect to insert exponentially many (in $t$) values into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the case where the HLL sketch has some empty buckets in S2).
 The result is largely independent of the HLL parameters.
-The more full is the HLL sketch, the larger is $1-H_M^{-1}$ and the easier it is to insert new items without increasing the cardinality estimate.
+The more full is the HLL sketch, the larger is $1-H_M^{-1}$ and the easier it is to insert new items without them increasing the cardinality estimate.
 
@@
 
-We briefly sketch an attack in the setting of an S1 adversary attacking an initially non-empty HLL sketch.
-An alternative attack with moderate performance in this setting is presented in~\cite{hllvuln}.
-
-Since the target sketch has an unknown status for the adversary, our previous strategy for S1 of keeping many buckets empty no longer works.
-One can attempt to use the shadow device to identify items that would result in small entries $M[j]$ (for unknown indices $j$) since such items are unlikely to increase the cardinality estimate of the target device by much.
-This can be done by first filling the shadow device with sufficiently many values to escape from the ``small range correction'' phase.
-Then one could enter candidate values $x$ into the shadow device and check for cardinality increases.
-However this runs into the rounding issue described in Section~\ref{sec:scenarios-discussion}.
-Still, one could retain those items which do lead to a detectable change in cardinality and reject the others (and also estimate the value of $M[j]$ for the unknown bucket $j$ from the magnitude of the change, retaining only those for which $M[j]$ is small).
-But here the lack of a cardinality change could result either from rounding issues or because insertion of candidate value $x$ into the shadow device really did lead to no increase in cardinality.
-This indicates that  ``good'' items would be rejected unnecessarily.
-This could be addressed by using multiple resets of the shadow HLL sketch, filling it with random items after each reset, and performing multiple insertions per candidate item $x$ to try to estimate the true value of $M[j]$ for each item $x$ (since after at least some of the resets followed by random insertions, we could expect $M[j]$ to increase when $x$ is inserted, and if $M[j]$ is small enough relative to the cardinality estimate, the change could be detected).
-The details of such an attack require further analysis which we do not pursue here, since we consider S1 to be the least realistic attack scenario.
+We briefly sketch an attack in this setting in Appendix~\ref{sec:non-empty-HLL-sketch-S1}.
 
@@
 %We first make an empty shadow HLL sketch via a reset.
-Consider an adversary that simply inserts $m \ln m$ random items $x$ from ${\mathcal{X}}$ into the shadow and target HLL sketches.
+Consider an adversary that simply inserts $m \ln m$ random items $x$ from ${\cal X}$ into the shadow and target HLL sketches.
 This increases the target HLL sketch's cardinality estimate by about $m \ln m$ since the HLL algorithm is a good estimator.
@@
 Now the adversary can carry out a version of our previous attack in scenario S1 (where the target HLL sketch is initially empty).
-By repeatedly processing lists of inputs, it can find a large subset ${\mathcal{L}}^*$ of ${\mathcal{X}}$ which does not cause the cardinality estimate of the shadow HLL sketch to increase beyond $m \ln m$.
+By repeatedly processing lists of inputs, it can find a large subset ${\cal L}^*$ of ${\cal X}$ which does not cause the cardinality estimate of the shadow HLL sketch to increase beyond $m \ln m$.
 This assumes that single updates to the vector of counter values $(M[j])_{j=1}^m$ in the shadow HLL sketch can be detected through changes to the HLL cardinality estimate.
 This seems plausible but requires further experimental investigation.
-Inserting all the items from ${\mathcal{L}}^*$ into the target HLL sketch then leads to an increase in its cardinality estimate of at most $m \ln m$.
-It is not hard to see that, since every bucket in the shadow HLL sketch is active, we can expect ${\mathcal{L}}^*$ to contain at least half of the items of ${\mathcal{X}}$.
-
-%An alternative approach is to simply run the scenario S1 attack as if the target HLL sketch were empty (i.e.\ as described in Section~\ref{sec:empty} above), take the resulting list ${\mathcal{L}}^*$, and insert it into the target HLL sketch, hoping that the cardinality does not increase too much.
+Inserting all the items from ${\cal L}^*$ into the target HLL sketch then leads to an increase in its cardinality estimate of at most $m \ln m$.
+It is not hard to see that, since every bucket in the shadow HLL sketch is active, we can expect ${\cal L}^*$ to contain at least half of the items of ${\cal X}$.
+
+%An alternative approach is to simply run the scenario S1 attack as if the target HLL sketch were empty (i.e.\ as described in Section~\ref{sec:empty} above), take the resulting list ${\cal L}^*$, and insert it into the target HLL sketch, hoping that the cardinality does not increase too much.
 %We report on this approach experimentally in Section~\ref{sec:exp}.
@@
 Redis 5.0 onwards uses exactly the formula in equation~(\ref{eqn:ertl}) to form its cardinality estimate.
-It uses $n=14$, $m=16,384$ and $\ell = 50$ (so $n+\ell = 64$ and a 64-bit unkeyed hash function is used).
+It uses $n=14$, $m=16,384$ and $\ell = 50$ (so $n+\ell = 64$ and a 64-bit hash function is used).
 
@@
 
-Attacks against the Redis implementation are now straightforward.
+Attacks are now straightforward.
 Suppose we can arrange that $C_0$, the number of empty buckets, stays above ${\epsilon}m$, a linear function of the total number of buckets, and that $C_{\ell+1} = 0$.
@@
 
-The requirement on $C_0 \geq {\epsilon}m$ is easily met in scenarios S2 and S4 with an initially empty HLL sketch, through the trick of sampling inputs $x \in {\mathcal{X}}$ so that only a fraction $1-\epsilon$ of the buckets are ever hit.
-We can expect to insert a fraction $1-\epsilon-\epsilon2^{-\ell}$ of the elements of  ${\mathcal{X}}$ whilst keeping the cardinality estimate $\hat{\lambda}$ below  $\alpha_\infty \epsilon^{-1} m$.
+The requirement on $C_0 \geq {\epsilon}m$ is easily met in scenarios S2 and S4 with an initially empty HLL sketch, through the trick of sampling inputs $x \in {\cal X}$ so that only a fraction $1-\epsilon$ of the buckets are ever hit.
+We can expect to insert a fraction $1-\epsilon-\epsilon2^{-\ell}$ of the elements of  ${\cal X}$ whilst keeping the cardinality estimate $\hat{\lambda}$ below  $\alpha_\infty \epsilon^{-1} m$.
 (The term $\epsilon2^{-\ell}$ arises from the requirement to keep $C_{\ell+1} =0$ in the attack; recall that $\ell=50$ in Redis, so the effect of this term is very small.)
@@
 These insertions do not alter the values of the $C_k$, and hence $\hat{\lambda}$ is unaffected.
-In S2, we expect to insert half the items in ${\mathcal{X}}$ without increasing the cardinality estimate $\hat{\lambda}$ at all.
+In S2, we expect to insert half the items in ${\cal X}$ without increasing the cardinality estimate $\hat{\lambda}$ at all.
 In S4, the analysis is the same as in Section~\ref{sec:nonemptyHLLS4}.
@@
 This completes our analysis of Redis.
-The key takeaway is that, in scenarios S2 and S4, an adversary can expect to insert a constant fraction of ${\mathcal{X}}$ into a non-empty Redis HLL sketch whilst not increasing its cardinality estimate at all (and whilst keeping the estimated cardinality linear in $m$ in the case where the sketch is initially empty).
+The key takeaway is that, in scenarios S2 and S4, an adversary can expect to insert a constant fraction of ${\cal X}$ into a non-empty Redis HLL sketch whilst not increasing its cardinality estimate at all (and whilst keeping the estimated cardinality linear in $m$ in the case where the sketch is initially empty).
 
@@
 Thus they are  vulnerable to our attacks.
-%\todo[inline]{maybe switch to one of these, or somehow be more clear about how the attacks transfer.}
+\todo{maybe switch to one of these, or somehow be more clear about how the attacks transfer.}
 
@@
 %\todo{We need to make sure the repo is publicly accessible but anonymised for submission, if we are going to include it.}
-For simplicity, we take ${\mathcal{X}}$ to be the set of 4-character long ASCII strings, with $\lvert {\mathcal{X}} \rvert = 7,311,616= 2^{22.8}$.
-We stress that this choice of inputs is merely illustrative and the results do not in any way depend on choosing ASCII strings.
-We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}.
+For simplicity, we take ${\cal X}$ to be the set of 4-character long ASCII strings, with $\lvert {\cal X} \rvert = 7,311,616= 2^{22.8}$.
+\todo{Explain there's nothing special about these inputs; anything would do; processed by hash function anyway.}We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}.
 %(note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
@@
 For simplicity and to keep the attack as striking as possible, we set $B$ to 1 in the attacks on an initially empty HLL sketch in scenarios S2 and S4.
-All attacks were run 30 times and we report average outcomes.
+All attacks were run 30 times; we report average outcomes.
 Our results are reported in Tables~\ref{table:tab1} and~\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary, and the cardinality estimate after adding the adversary's items.
@@
 Also as discussed in Section~\ref{sec:redis}, the Redis implementation has changed its HLL algorithm several times in recent releases, but the authors of~\cite{hllvuln} do not specify which version was used in their experiments.
-This makes it difficult to make a precise comparison with the results of~\cite{hllvuln}.
+This makes it difficult to make a precise comparison with their results.
 Nevertheless, the S1 adversary following our strategy against the classical HLL algorithm is able to add 977 and 97,433 distinct items while holding the cardinality estimate at 1 and 126, respectively.
@@
 This comes at the cost of performing more resets compared to~\cite{hllvuln}.
-We recall that in practice each reset may cause a delay (for example, it may involve rebooting a network device); we also recall that we consider S1 to be a somewhat artificial attack scenario.
+\todo{explain why counting resets matter; also that we don't care too much about this scenario.}
 
@@
 
-\todo[inline]{Results from Redis or GBQ if we get them.}
-
-\subsection{Relating the Attacks to Applications}\label{sec:exp-apps}
-
-So far, we have treated attacks in an abstract way, working with an off-the-shelf implementation of HLL.
-We now illustrate what effect such attacks might have on applications relying on HLL for cardinality estimation.
-Specifically, we analyse in detail a recent paper~\cite{flexswitch} from the networking systems community.
-
-The authors of~\cite{flexswitch} propose to use HLL for recording traffic flow statistics in datacenter network switches.
-The idea is to use HLL so as to reduce the amount of memory needed in the switches.
-In the main example considered in~\cite[Section 3.1]{flexswitch}, each flow is identified by a 4-tuple consisting of source IP address, destination IP address, source port, and destination port.
-Various parameters for the HLL sketch to be used are then explored in~\cite{flexswitch}, trading memory consumed for accuracy of estimation of the number of concurrent flows.
-The paper also proposes to use the low range correction method of Fig.~\ref{fig:hll}.
-At no point does the discussion in~\cite{flexswitch} consider adversarial input (it does mention worst-case performance but only in passing, and no indication is given of how to handle this).
-It is clear that, by modifying the 16-bit source port (for example), an adversary could fool the switch into undercounting the flows the adversary is responsible for producing.
-Depending on how the flow count is used, this could lead to undetected DoS conditions, mis-provisioning of network capacity, or under-billing of the adversary for its traffic.
-
-Going further, \cite[Section 4.2]{flexswitch} proposes using HLL for detecting TCP port scans by counting observed destination port numbers in TCP packets with set SYN flags and reporting a scan event if the estimate exceeds a
-set threshold.
-Using our attack techniques, an attacker could complete a partial port scan of roughly half of all $2^{16}$ ports whilst keeping the HLL estimate below $0.693m$ where $m$ is the number of buckets used in HLL.
-While~\cite{flexswitch} is not specific about the value of $m$ actually used, the authors' focus on reducing memory usage indicates that $m$ would not be larger than $1024$.\footnote{The paper mentions using 1 byte counters in the buckets and states that using more than 1KB of memory ``only improves average error rates marginally''; we infer a maximum value of $m=1024$ from this.} It is not stated in~\cite{flexswitch} what threshold needs to be crossed to trigger an alarm, but we consider it plausible that it would be above $0.693 \cdot 1024 = 710$.
-
-Similarly,~\cite[Section 4.2]{flexswitch} proposes using HLL for detecting NTP amplification attacks by counting distinct source IP addresses arriving on UDP port 123 over a small time window.
-No further details are given, but it is clear that an attacker with even limited flexibility in spoofing source IP addresses (e.g.\ within a private range used in the datacenter) could easily evade detection whilst delivering a large number of NTP requests, thus re-enabling the original attack.
-
-In a more advanced application,~\cite[Section 4.3]{flexswitch} proposes to use HLL as a component in a novel, low-memory version of the RCP congestion control algorithm called RCP-Approx.
-Again HLL is used for estimating numbers of traffic flows.
-Obviously, our attack techniques could be used (in scenario S2) to cause RCP-Approx to underestimate the number of traffic flows, leading to non-detection of congestion and so enabling DoS attacks to go undetected.
-
-Similar remarks apply to the treatment of CONGA in~\cite[Section 4.5]{flexswitch}.
-CONGA is a congestion-aware load balancing system that splits TCP flows into flowlets (collections of IP packets) and allocates them to paths based on congestion feedback from remote switches.
-In the enhancements to CONGA proposed in~\cite{flexswitch}, HLL is used in the ``flow statistics'' building block for estimating flowlet statistics in switches; these estimates are sent back to a monitoring system and used to determine the load balancing configuration.
-This overall system can be abused by an S2 adversary, to cause flowlet numbers to be underestimated (or overestimated) and the load balancing to perform sub-optimally.
-For example, an adversary could cause all the switches to \emph{over}-estimate the number of flowlets they are handling and ask the load balancing system to send less traffic through them.
-This could result in a degradation in the total amount of traffic the load balancing system can handle (depending on what fail-safe mode it has).
-Or an attacker could cause all the switches to \emph{under}-estimate the number of flowlets they are handling, causing undetected congestion in the switches.
-
-\todo[inline]{This ends up being pretty harsh on this NSDI paper, but hey. We may shorten or move some details to an appendix.}
+\todo{Add a couple of paragraphs to relate all this back to the initial attack scenarios, e.g.
+packet flows.
+May need to look at internals of some of those papers to see where flex in inputs will come from.
+}
 
@@
 
-In this section we present a formal security analysis of HLL, focussing on how to augment the original HLL design with cryptographic mechanisms and thereby provably prevent attacks.
+In this section we present a formal security analysis of HLL, focussing on whether we can augment the original HLL design with cryptographic mechanisms and thereby provably prevent attacks.
 
@@
 
-\subsection{Intuition for Formal Security Analysis}
+\subsection{Formal Security Analysis for S2, S3, S4}
 
 For scenarios S2, S3 and S4, the picture is much more positive.
-In the remainder of this section, we introduce a formal security model covering all three scenarios.
-We work with a simulation-based framework, in contrast to the game-based approach introduced for probabilistic data structures in~\cite{CCS:ClaPatShrPS19}.
-%We discuss the strengths of our alternative approach in more detail below.
-
-We first begin by explaining the high-level idea of our approach.
-We allow an adversary ${\mathcal{A}}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios S2, S3 and S4: it may have an \texttt{insert($\cdot$)} oracle allowing it to add values to the HLL sketch, a \texttt{read} oracle that returns the HLL cardinality estimate, and a \texttt{corrupt} that returns all the HLL bucket contents (but not any secret key $K$); in S4 it has all three oracles.
-In addition, in all 3 scenarios, ${\mathcal{A}}$ has access to an \texttt{init($\cdot$)} oracle by which it sets the HLL parameters $n$, $\ell$, $\alpha_n$ used to initialise the HLL sketch.
-We assume ${\mathcal{A}}$ always makes exactly one \texttt{init($\cdot$)} query and that it is its first oracle call.
-
-In the security model, the game selects at random one of two worlds by sampling a bit $b$.
-In the real world ($b=0$), ${\mathcal{A}}$ interacts with the scheme HLL-$F$.
-This is the same as HLL operating as in Figure~\ref{fig:hll} but with the hash function $h$ replaced by a keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$ whose key $K$ is sampled uniformly at random at the beginning of the game and not available to the adversary; ${\mathcal{A}}$ has access to the relevant selection of oracles and completes its execution with some output $\textsf{out}_{\mathcal{A}}$ (a bit-string).
-
-In the ideal world ($b=1$), ${\mathcal{A}}$ is replaced by a simulator ${\mathcal{S}}$ with blackbox oracle access to ${\mathcal{A}}$ (that is, ${\mathcal{S}}$ can run ${\mathcal{A}}$ in a blackbox manner, receiving its oracle queries and providing responses to ${\mathcal{A}}$ as it sees fit).
-In particular, ${\mathcal{S}}$ receives parameters $n$, $\ell$, $\alpha_n$ from ${\mathcal{A}}$'s \texttt{init($\cdot$)} call.
-Finally, ${\mathcal{S}}$ receives $\textsf{out}_{\mathcal{A}}$ and completes its execution by producing its own output $\textsf{out}_{\mathcal{S}}$.
-
-Importantly, we will specify an ${\mathcal{S}}$ that handles all of ${\mathcal{A}}$'s \texttt{insert($x$)} queries using \emph{random sampling}, that is, by setting $y \leftarrow_{\$} \{0,1\}^{n+\ell}$ instead of $y: = F_K(x)$ and then executing the rest of the insertion process as in Figure~\ref{fig:hll}.
-In essence, our ${\mathcal{S}}$ will build its own ``shadow'' HLL in which all insertion queries are handled by random sampling.
-${\mathcal{S}}$ then uses the shadow HLL for handling ${\mathcal{A}}$'s \texttt{read} and \texttt{corrupt} queries.
-Finally, we will specify ${\mathcal{S}}$ so that it outputs whatever ${\mathcal{A}}$ outputs.
-
-We will prove that  ${\mathcal{S}}$ is such that, given any adversary ${\mathcal{A}}$, then any distinguisher $\mathcal{E}$ that can distinguish between the $\textsf{out}_{\mathcal{A}}$ in the real world and $\textsf{out}_{\mathcal{S}}$ in the ideal world can be used to build an adversary ${{\mathcal{B}}}$ which breaks the PRF security of $F$.
-Under the assumption that $F$ is a good PRF, no such efficient distinguisher $\mathcal{E}$ can exist, and we may conclude that (up to PRF security) ${\mathcal{E}}$ cannot tell that it is interacting with the ideal world instead of the real world.
-
-Why does this proof imply that ${\mathcal{A}}$ cannot manipulate the cardinality estimate?
-To see this, note that (amongst many other possibilities) ${\mathcal{A}}$'s output could be the vector of bucket values $M[j]$ obtained from a \texttt{corrupt} query, enabling a specific ${\mathcal{E}}$ to compute the cardinality estimate for itself and to attempt to use that estimate as a distinguisher between the two worlds.
-Yet, by the preceding argument, the simulator in the ideal world is able to produce an output that ${\mathcal{E}}$ cannot distinguish from that of ${\mathcal{A}}$ in the real world.
-In particular, our simulator will output whatever ${\mathcal{A}}$ outputs, so if ${\mathcal{A}}$ outputs the result of a \texttt{corrupt} query in the ideal world, then so will ${\mathcal{S}}$.
-But in the ideal world, ${\mathcal{S}}$ uses random sampling for the $y$ values, so there ${\mathcal{S}}$ (and ${\mathcal{A}}$ that it runs) are operating in the setting where the average case analysis of HLL applies, as per~\cite{hll}.
-So in the ideal world this specific ${\mathcal{E}}$ would end up computing an accurate average case cardinality estimate.
-Since ${\mathcal{E}}$ tries to distinguish the two worlds by computing the cardinality estimate and we already showed that no efficient distinguisher ${{\mathcal{E}}}$ can exist, it must follow that the cardinality estimates produced in the real and ideal worlds are indistinguishable (up to PRF security).
-It then follows that ${\mathcal{A}}$ has no advantage when it comes to manipulating the HLL-$F$ cardinality estimate.
-So HLL, when instantiated with PRF $F$, is secure against cardinality manipulation attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
-
-%We will conclude our analysis by discussing how $F$ can be instantiated.
+In the remainder of this section, we introduce a formal attack model covering all three scenarios.
+This is done by allowing an adversary ${\cal A}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios.
+In the security model, ${\cal A}$ interacts in one of two worlds.
+In the real world, ${\cal A}$ interacts with a real HLL, operating as in Figure~\ref{fig:hll}.
+In the ideal world, ${\cal A}$ interacts with an ideal HLL in which the operation $y: = h(x)$ is replaced by random sampling; that is, by setting $y \leftarrow_{\$} \{0,1\}^{n+\ell}$.
+
+\todo{update! may need to formally define algs of an HLL as well as oracles. Since Sim will need to keep a shadow copy for itself.}
+
+We then show that, if $h$ is replaced by a keyed Pseudo Random Function (PRF) $F$ whose key is not available to the adversary, then any adversary ${\cal A}$ that can distinguish between the real and random worlds with oracles as in S4 can be used to build an adversary ${\cal B}$ which breaks the PRF security of $F$.
+Under the assumption that $F$ is a good PRF, no such efficient adversary ${\cal B}$ can exist, and we may conclude that we can safely switch from the real world to the ideal world without incurring any significant change to the behaviour of ${\cal A}$.
+
+However, it is clear that any adversary in the ideal world is reduced to operating in the setting used in the average case analysis of HLL (wherein every insertion query results in a random value $y$ being used to update the HLL sketch).
+Thus, ${\cal A}$ can have no advantage when it comes to influencing the HLL cardinality estimate.
+It follows that HLL, when instantiated with PRF $F$, is secure against attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
+We will conclude by discussing how $F$ can be instantiated.
+
+\todo{explain the above better with simulators + conf for free.}
+
+We note that our formal security analysis could be conducted using the framework recently laid out in~\cite{CCS:ClaPatShrPS19}. However, we prefer to give a compact and self-contained treatment here.
+
+\todo{explain why a sim-based approach is better, more general, does not require a specific winning condition.}
+
+\todo{Why UHFs don't work in general -- key recovery is possible in S4; partial info about processed values and therefore keys  leaks from Card Est in S2, though harder to make an attack from this. }
 
@@
 
-We now formalise the above intuition.
-We consider distinguisher ${\mathcal{E}}$ to be an algorithm running in a security game.
-The game begins with the selection of a bit $b$ uniformly at random and then runs either ${\mathcal{A}}$ (the real world, $b=0$) or ${\mathcal{S}}$ (the ideal world, $b=1$).
-Adversary ${\mathcal{A}}$ interacts with up to four oracles which are handled by the game using the algorithms of HLL-$F$; different combinations of oracles define scenarios S2, S3, S4.
-After making some queries, ${\mathcal{A}}$ returns its output $\textsf{out}_{\mathcal{A}}$ to ${\mathcal{E}}$.
-Simulator ${\mathcal{S}}$ runs ${\mathcal{A}}$ in a blackbox manner (as specified below) and receives ${\mathcal{A}}$'s output.
-It then provides its own output $\textsf{out}_{\mathcal{S}}$ to ${\mathcal{E}}$.
-Finally, ${\mathcal{E}}$ outputs a bit $b'$ estimating whether it is interacting with the real world ($b'=0$) or the ideal world ($b'=1$).
-For $i=2, 3, 4$, we define ${\mathcal{E}}$'s S$i$-Real-or-Ideal advantage against HLL-$F$ with adversary ${\mathcal{A}}$ and simulator ${\mathcal{S}}$ to be:
-\[
-\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{E}}) := \lvert 2\Pr[b' = b] - 1 \rvert \, .
+\todo{Upgrade to use simulator S as a typical simulation-based security notion should; reference to Lindell's ``How to simulate it'' paper?
+Order of quantifiers; input to simulator: nothing, not even number of queries, assuming adversary never repeats an insert query.}
+
+We consider adversary ${\cal A}$ to be an algorithm running against a challenger ${\cal C}$ in a security game.
+The game begins with ${\cal C}$ selecting a bit $b$ uniformly at random.
+Adversary ${\cal A}$ interacts with up to four oracles which are handled by ${\cal C}$; different combinations of oracles define scenarios S2, S3, S4.
+After making some queries, ${\cal A}$ returns a bit $b'$ estimating whether it is interacting with a real HLL instantiation ($b'=0$) or an ideal one ($b'=1$).
+For $i=1, 2, 3$, we define ${\cal A}$'s S$i$-Real-or-Ideal advantage against HLL to be:
+\[
+\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny HLL}}({\cal A}) := \lvert 2\Pr[b' = b] - 1 \rvert \, .
 \]
 %Using a standard transformation, this probability can be rewritten as $\lvert \Pr[b' = 1| b = 1] - \Pr[b'=0|b=1] \rvert$. 
-
-We say that the scheme HLL-$F$ is S$i$-simulation-secure with parameters $(q,t_{\mathcal{S}},t_{\mathcal{A}},t_{\mathcal{E}},\epsilon)$ if there exists a simulator $\mathcal{S}$ running in time at most $t_{\mathcal{S}}$ such that, for all adversaries ${\mathcal{A}}$ running in time at most $t_{\mathcal{A}}$ and making at most $q$ \texttt{insert($\cdot$)} queries and all distinguishers ${\mathcal{E}}$ running in time at most $t_{\mathcal{E}}$, we have
-$\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{E}}) \leq \epsilon$.
-We quantify over \texttt{insert($\cdot$)} queries (via $q$) and not other types of query because only \texttt{insert($\cdot$)} queries will appear in our concrete security analysis and we do not want to clutter the notation.
-
-The four oracles that  ${\mathcal{A}}$ can interact with behave as follows in the real world:
+The four oracles that  ${\cal A}$ can interact with are as follows:
 \begin{itemize}
-\item \texttt{init($\cdot$)}: this oracle receives as input the HLL-$F$ parameters ($n$, $\ell$, $\alpha_n$) and performs initialisation of the HLL-$F$ sketch, including generating the array of buckets $M[j]$ and initialising all the entries to $0$, and selecting a key $K \leftarrow_{\$} \{0,1\}^k$ for keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
-This oracle has no output.
-It is available in all of S2, S3, S4.
-We assume it is only called once as ${\mathcal{A}}$'s first oracle query.
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$, the oracle implements HLL-$F$ by setting $y = F_K(x)$ and following the procedure in Figure~\ref{fig:hll} to insert value $x$ into the HLL sketch (that is, it uses the first $m$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
-This oracle has no output.
-It is available in all of S2, S3, S4.
-Without loss of generality (since HLL and variants of it we consider here have a deterministic insertion procedure), we assume all queries $x$ made by ${\mathcal{A}}$ to \texttt{insert($\cdot$)} are distinct.
-\item \texttt{read}: this oracle has no input and returns the HLL-$F$ cardinality estimate as per Figure~\ref{fig:hll}, i.e.\ it computes a raw cardinality estimate via $E := \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$ and then performs corrections for small and large values.
-We allow multiple queries to this oracle and make it available to ${\mathcal{A}}$ in S3 and S4, but not S2.
-\item \texttt{corrupt}: this oracle has no input and returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents (but not any secret key $K$).
+\item \texttt{init}: this oracle performs initialisation of the HLL sketch (and selection of the PRF key when $h$ is replaced by a PRF $F$); this oracle is available in all of S2, S3, S4, but can only be called once as ${\cal A}$'s first oracle query.
+\item \texttt{insert($\cdot$)}: on input $x \in {\cal D}$, the oracle inserts value $x$ into the HLL sketch.
+This oracle is available in all of S2, S3, S4.
+Without loss of generality (since HLL and variants of it we consider here have a deterministic insertion procedure), we assume all queries $x$ made by ${\cal A}$ to \texttt{insert($\cdot$)} are distinct.
+We let $q$ denote the number of insertion queries made by ${\cal A}$ during the course of its attack.
+\item \texttt{read}: this oracle returns the HLL cardinality estimate. We allow multiple queries to this oracle and make it available to ${\cal A}$ in S3 and S4, but not S2.
+\item \texttt{corrupt}: this oracle returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents (but not any secret key $K$).
 This oracle is available only in S4.
 We allow the adversary to query \texttt{corrupt} as often as it wishes during its attack.
-Our description of S4 in Section~\ref{sec:attacks} limited the adversary to at most one such query; here we allow multiple queries because doing so does not increase the difficulty of achieving security.
+Our description of S4 in Section~\ref{sec:attacks} limited the adversary to at most one such query; we allow multiple queries here because doing so does not increase the difficulty of achieving security.
 \end{itemize}
 
-We note that in S4 (and only in S4) the adversary has access to all four oracles. This makes it the strongest model, and hereafter we only target security in this model.
-
-We next describe the simulator ${\mathcal{S}}$ for the ideal world. Simulator ${\mathcal{S}}$ operates by running ${\mathcal{A}}$ in a blackbox manner and handling all of its queries as follows: 
-\begin{itemize}
-\item \texttt{init($\cdot$)}: on input $n$, $\ell$, $\alpha_m$, ${\mathcal{S}}$ initialises an HLL sketch as in Figure~\ref{fig:hll}; that is, it creates an array of $m=2^n$ buckets $M[j]$ and initialises them all to 0.
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$,  ${\mathcal{S}}$ ignores $x$, sets $y \leftarrow_{\$} \{0,1\}^{n+\ell}$, and proceeds to update the buckets using this value of $y$ as in the original HLL algorithm in Figure~\ref{fig:hll}.
-\item \texttt{read}:  ${\mathcal{S}}$ returns the HLL cardinality estimate computed as per Figure~\ref{fig:hll}.
-\item \texttt{corrupt}: ${\mathcal{S}}$ returns the current vector $(M[j])_{j=1}^{m}$ of HLL bucket contents that it has stored.
-\end{itemize}
-At the end of ${\mathcal{A}}$'s execution, ${\mathcal{S}}$ takes ${\mathcal{A}}$'s output and forwards it to ${\mathcal{E}}$ as its own output.
-Notice that the running time of  ${\mathcal{S}}$ is (roughly) equal to that of ${\mathcal{A}}$.
-
-We define the PRF advantage of an adversary ${{\mathcal{B}}}$ against $F$ outputting a bit $d'$ as being its advantage in distinguishing whether it is interacting with $F$ for a uniformly random key $K$ ($d=0$) or interacting with a truly random function ($d=1$).
+We note that in S4 (and only in S4) the adversary has access to all four oracles.
+This makes it the strongest model, and hereafter we only target security in this model.
+We count queries $q$ to \texttt{insert($\cdot$)} because this will enable us to quantify the quality of our security reduction to PRF security below.
+
+\subsection{Security Analysis of HLL with a PRF}
+
+We now replace the hash function $h$ acting on ${\cal D}$ from the HLL algorithm with a keyed function $F: \{0,1\}^k \times {\cal D} \rightarrow  \{0,1\}^{n+\ell}$.
+Here the first input to $F$ is a $k$-bit string, the key, denoted $K$.
+We assume that \texttt{init} in the HLL security model now initialises key $K$ by selecting it uniformly at random from $\{0,1\}^k$.
+
+We define the PRF advantage of an adversary ${\cal B}$ against $F$ outputting a bit $d'$ as being its advantage in distinguishing whether it is interacting with $F$ for a uniformly random key $K$ ($d=0$) or interacting with a truly random function ($d=1$).
 Thus:
 \[
-\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}}) := \lvert 2\Pr[d' = d] - 1 \rvert\,.
-\]
-Here, we assume without loss of generality that all of ${{\mathcal{B}}}$'s queries are distinct. See~\cite[Section 4.4.1]{BonehShoup} for details of the PRF security definition.
-
-Now we are ready to formally state and prove our main result:
+\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B}) := \lvert 2\Pr[d' = d] - 1 \rvert\,.
+\]
+Here, we assume without loss of generality that all of ${\cal B}$'s queries are distinct. See~\cite[Section 4.4.1]{BonehShoup} for details of the PRF security definition.
+
+Now we are ready to state:
+
 \begin{theorem}\label{thm:security-of-HLL-with-PRF}
-Let simulator $\mathcal{S}$ be as described above.
-For any S4 adversary ${\mathcal{A}}$ against HLL-$F$ and any distinguisher $\mathcal{E}$, we can construct a PRF adversary ${{\mathcal{B}}}$ against $F$ such that:
-\[
- \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}}) = 
- \mbox{Adv}^{\mbox{\tiny S$4$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{E}}) \,.
-\]
-Moreover, if ${\mathcal{A}}$ makes $q$ queries to its \texttt{insert($\cdot$)} oracle, then ${{\mathcal{B}}}$ makes $q$ queries to its real or random oracle.
-Finally, the running time of ${{\mathcal{B}}}$ is (roughly) $t_{\mathcal{A}}+t_{\mathcal{E}}$.
+Suppose ${\cal A}$ is an S4 adversary against HLL instantiated with a keyed function $F: \{0,1\}^k \times {\cal D} \rightarrow  \{0,1\}^{n+\ell}$.
+Then we can construct from ${\cal A}$ a PRF adversary ${\cal B}$ against $F$ such that:
+\[
+ \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B}) = \mbox{Adv}^{\mbox{\tiny S4-RoI}}_{\mbox{\tiny HLL}}({\cal A}) \,.
+\]
+Moreover, if ${\cal A}$ makes $q$ queries to its \texttt{insert($\cdot$)} oracle, then ${\cal B}$ makes $q$ queries to its real or random oracle.
+Finally, the running time of ${\cal B}$ is (roughly) the same as that of ${\cal A}$.
 \end{theorem}
 
-%Informally, if $F$ is a good PRF, then $\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}})$ should be small for any adversary ${{\mathcal{B}}}$ running in a reasonable amount of time; hence the advantage of any S4 adversary  ${\mathcal{A}}$ against HLL using PRF $F$ is also small.
-Note that our proof still applies for the improved algorithm of~\cite{hllnew} that is used in Redis.
+Informally, if $F$ is a good PRF, then $\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({\cal B})$ should be small for any adversary ${\cal B}$ running in a reasonable amount of time; hence the advantage of any S4 adversary  ${\cal A}$ against HLL using PRF $F$ is also small.
+The proof can be found in Appendix~\ref{sec:proof}.
+Our proof still applies for the improved algorithm of~\cite{hllnew} that is used in Redis.
 The only changes to the proof are to use the Redis approach for computing responses to \texttt{read} queries in place of the classical HLL approach.
 
-\begin{proof}
-PRF adversary ${{\mathcal{B}}}$ against $F$ is built as follows.
-Let $d$ denote the hidden bit determining whether ${{\mathcal{B}}}$ receives real responses of the form $F_K(x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${{\mathcal{B}}}$ provides a simulation of either the real or random world for distinguisher ${\mathcal{E}}$ as follows.
-${{\mathcal{B}}}$ runs ${{\mathcal{A}}}$ and answers each of ${{\mathcal{A}}}$'s queries as follows:
-\begin{itemize}
-\item \texttt{init($\cdot$)}: ${{\mathcal{B}}}$ receives parameters $n$, $\ell$, $\alpha_m$ and initialises an array $M[j]$ in which each of the $2^n$ entries is set to 0. 
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$,  ${{\mathcal{B}}}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${{\mathcal{B}}}$ uses $y$ to update the HLL sketch as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
-That is, it uses the first $m$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$.
-\item \texttt{read}: ${{\mathcal{B}}}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate as per Figure~\ref{fig:hll}. 
-\item \texttt{corrupt}: ${{\mathcal{B}}}$ returns the vector $(M[j])_{j=1}^{m}$. 
-\end{itemize}
-When ${\mathcal{A}}$ returns its final output $\textsf{out}_{\mathcal{S}}$, ${{\mathcal{B}}}$ forwards this output to ${{\mathcal{E}}}$.
-Finally, ${{\mathcal{E}}}$ outputs a bit $b'$.
-Then ${{\mathcal{B}}}$ outputs its own bit $d'=b'$.
-
-Note that when $d=0$, ${{\mathcal{B}}}$ plays against the real $F$ with a random key $K$ and provides a simulation for ${\mathcal{A}}$ that perfectly instantiates the real world.
-On the other hand, when $d=1$, ${{\mathcal{B}}}$ plays against a truly random function and perfectly instantiates our simulator $\mathcal{S}$ in the ideal world.
-To see this, note that sampling the outputs $y$ of a truly random function at distinct inputs $x$ is identical to simply sampling $y$ uniformly at random.\footnote{Our proof relies on the $x$ being distinct, but this is just to simplify book-keeping and is without loss of generality.} The equality of advantages in the statement of the theorem now follows immediately.
-The remainder of the proof follows on counting queries and noting that the HLL sub-algorithms for insertion and computing cardinality are efficient, so ${{\mathcal{B}}}$'s running time due to handling ${\mathcal{A}}$'s queries is not increased significantly compared to that of ${\mathcal{A}}$.
-The final running time for ${\mathcal{B}}$, namely $t_{\mathcal{A}}+t_{\mathcal{E}}$, follows on noting that ${\mathcal{B}}$ first runs ${\mathcal{A}}$ and then ${\mathcal{E}}$.
-\end{proof}
-
-Note that our security analysis using a simulation-based definition does not require us to specify a winning condition for the adversary.
-Rather, we infer from inspecting the simulator what the adversary can (or cannot) achieve in the ideal world, and show that no efficient adversary can distinguish the real and ideal worlds.
-This is in contrast to the game-based approach of~\cite{CCS:ClaPatShrPS19} which needs a complex definition to say when the adversary has successfully manipulated a probabilistic data structure.
-This reveals a strength of the simulation-based approach in more complex settings like ours.
-
-\subsection{On Confidentiality}
-
-An HLL can leak information about values $x$ that were previously inserted.
-Consider the following attack from~\cite{cardestprivacy}: to test whether a value $x$ was previously inserted into the HLL by an honest user (say), the adversary makes the sequence of queries: \texttt{read}, \texttt{insert($x$)}, \texttt{read}.
-If the two cardinality estimates are different, the adversary can conclude definitively that $x$ was \emph{not} previously inserted into the HLL; if they are the same, it can estimate (with some error) that $x$ was previously inserted.\footnote{Errors can arise in two ways when $x$ was not previously inserted: either inserting $x$ changes a bucket value, but the change is masked by rounding when the cardinality estimate is computed, or inserting $x$ does lead to a change in a bucket value because $\rho(w)$ computed from $x$ is too small.} This attack still works even if HLL is implemented using a PRF and is unavoidable in view of the oracle queries (and in real settings, the API) available to the adversary.
-So such privacy leaks are an inherent feature of HLL if a rich enough API is provided.
-Note that this attack does not contradict our main result, which concerns prevention of cardinality manipulation rather than confidentiality.
-
-%Still, simulation-based security notions are often used to model confidentiality properties, so it is tempting to think that our security definition might say something about confidentiality of the HLL inputs $x$.
-Our existing model does not, since ${\mathcal{A}}$ chooses all the inputs $x$ and so ``knows'' them already.
-However, it is easily modified to provide a confidentiality notion.
-For example, we can consider the insertion of honestly generated values $x$ that are drawn from some probability distribution $p$ on $\mathcal{D}$ in a first phase, with ${\mathcal{A}}$ then having access to only \texttt{read} and \texttt{corrupt} oracles in a second phase (we have to remove the \texttt{insert($\cdot$)} oracle in view of the general attack from~\cite{cardestprivacy}).
-
-%In this case, the simulator ${\mathcal{S}}$ in the ideal world can proceed without ever seeing any inputs but instead knowing only the distribution $p$ and the number of distinct samples drawn from it in the first phase.
-It is not hard to modify the proof of Theorem~\ref{thm:security-of-HLL-with-PRF} to show that no adversary can distinguish real and ideal worlds in this case (up to PRF security).
-Now, since ${\mathcal{S}}$ has no access to inputs $x$, we can conclude that the HLL implemented using a PRF ensures confidentiality for the inputs $x$.
-Thus, the cardinality estimates and even the bucket values themselves do not leak information about the honestly inserted values $x$ (beyond their number).
-
-We leave the exploration of the adaptation of our simulation-based security definitions to capture confidentiality notions for more restrictred APIs to future work.
-
-\subsection{Is Universal Hashing Sufficient?}
-
-It is natural to wonder whether a PRF is really required in place of hash function $h$ in order to make HLL secure, or whether a weaker cryptographic primitive would suffice.
-For example, since it has (roughly) uniform outputs, perhaps using a universal hash function (UHF) $h_K(\cdot)$ would be sufficient.
-
-In general the key of a UHF may be recovered from a small number of its outputs.
-Consider, for example, the UHF built from polynomial hashing over a finite field $\mathcal{F}$.
-Here, inputs $x$ are encoded as polynomials $p_x \in \mathcal{F}(Z)$, the key is a uniformly random field element, and the UHF output $h_K(x)$  is equal to $p_x(K)$, obtained by polynomial evaluation at $K$.
-The specific translation from $x$ to $p_x$ involves encoding $x$ as a vector $(x_1,x_1,\ldots,x_v) \in \mathcal{F}^v$ for some $v$ and then setting:
-\[
-p_x(Z) = Z^v + x_1Z^{v-1} + \cdots + x_{v-1}Z + x_v.
-\]
-An adversary who knows the input $x$, and therefore the vector $(x_1,x_2,\ldots,x_v)$ can determine the key $K$ by finding roots of polynomials over $\mathcal{F}$, using Berlekamp's algorithm~\cite{Berlekamp}.
-If the adversary can choose $x$, then it can make the polynomials $p_x$ have low degree, making the attack easier to perform.
-
-Now consider using such a UHF in place of $h$ in attack scenario S4.
-For concreteness, we assume $\mathcal{F} = \textsf{GF}(2^{128})$, so our UHF has 128-bit keys and 128-bit outputs.
-The output is then truncated to provide the 32 bits needed by HLL (or, say, the 64 bits consumed by the Redis variant of HLL).
-The encoding step, taking bit-string $x$ to vector $(x_1,x_2,\ldots,x_v)$, simply splits $x$ into 128-bit strings $x_i$, each $x_i$ representing a field element.
-The attack we present below is specific to this UHF, but can be extended to other cases.
-
-An S4 adversary $\mathcal{A}$ can perform a series of pairs of \texttt{insert($x^i$)} and \texttt{corrupt} queries on chosen values $x^0, x^1, \ldots, x^{t-1}$.
-With high probability (assuming $t$ is not too large), each \texttt{insert} query hits a fresh bucket in the HLL sketch.
-Then comparing the results of the \texttt{corrupt} queries either side of the \texttt{insert($x^i$)} query reveals to $\mathcal{A}$ which bucket was hit by that  \texttt{insert} query.
-In turn, this leaks partial information about the corresponding UHF output $p_{x^i}(K)$, specifically, its $n$ least significant bits (since those bits are used to generate the bucket index for insertion).
-We now show how to exploit this partial information on $p_{x^i}(K)$ to recover $K$.
-
-%For example when $x^i$ is short enough that $v=1$, then $p_{x^i}(K) = K + x^i$ and the bucket index leaks $n$ bits of $K$ immediately.
-
-By choosing $x^i$ to have 256 bits, we have $v=2$ and $p_{x^i}(K) = K^2 + x^i_1K+x^i_2$; then further selecting $x^i$ such that  $x^i_2$ is a constant value (e.g.\ $0 \in \textsf{GF}(2^{128})$) and considering consecutive pairs of $x_i$ values, we obtain equations of the form:
-\[
-p_{x^{2j+1}}(K) \oplus p_{x^{2j}}(K) = (x^{2j+1}_1 \oplus x^{2j}_1)K
-\]
-where $\oplus$ denotes addition over $\textsf{GF}(2^{128})$.
-Here, the $n$ least significant bits of the left-hand side leak through the indices of the buckets hit for each index.
-By selecting the $x^i_1$ uniformly at random and analysing the equations of the above form that result, we obtain a non-singular linear system over $\textsf{GF}(2)$ in the unknown bits of $K$.
-The number $t$ of $x^i$ needed is roughly $2\cdot128/n$ since we get $n$ independent linear equations per pair $(x^{2j},x^{2j+1})$ and there are 128 unknown bits in $K$.
-The required value of $t$ is small for typical HLL parameters (e.g.
-$n=16$ would yield $t=32$).
-
-%Going further, the S4 adversary has complete control over the values $x_i$, so it can choose them so that the polynomials $p_{x_i}$ are of degree 1 over $\textsf{GF}(2^{128})$.
-Then $K$ can be determined by writing everything bit-wise and solving a linear system.
-The idea is that, when $x_i$ has 128 bits, we can write $h_K(x_i) = p_{x_i}(K) = x_i \cdot K$ where the multiplication is over $\textsf{GF}(2^{128})$ (and we turn bit-strings into field elements in a standard way).
-Then each bit of $h_K(x_i)$ is a linear combination of the bits of $K$, with the coefficients being determined by the known bits of $x_i$ and the degree 128 irreducible polynomial $f$ over $\textsf{GF}(2)$ that determines how field multiplication works at the bit level.
-
-The above attack exploits specific properties of the chosen UHF, but suffices to demonstrate that UHFs are not a suitable tool for protecting HLL in general.
-It is still possible that UHFs could be safely used in weaker attack scenarios, e.g.\ S2.
-We leave a detailed study to future work.
+We conclude by reiterating our central point: we have shown that no efficient S4 adversary ${\cal A}$ can tell that it is playing against HLL implemented using truly random values $y$ instead of with values obtained from a PRF, except with a small advantage.
+This means that ${\cal A}$ is effectively reduced to attacking HLL in the exact same setting used in the average case analysis of HLL from~\cite{hll}, wherein it is proved that the HLL cardinality estimate endures relative errors of the form $1.04/\sqrt{m}$.
+It does not matter that ${\cal A}$ can adaptively choose its inputs $x$ in \texttt{insert($\cdot$)} queries, nor that it can access \texttt{read} and \texttt{corrupt} queries.
+For each \texttt{insert($\cdot$)} query, a fresh value $x$ leads to a new random $y$ being generated and used to update a bucket, as in the average case analysis of HLL.
+The \texttt{read} and \texttt{corrupt} queries do leak information about $F$'s outputs to the adversary, but such outputs are indistinguishable from random under the assumption that $F$ is a PRF.
+
+\todo{revisit the above in view of introduction of simulator, plus the confidentiality aspect for free. NB corruption query does not give PRF key.}
 
@@
 
-Our HLL construction requires a PRF $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
-Here, the set ${{\mathcal{D}}}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
+The above construction requires a PRF $F: \{0,1\}^k \times {\cal D} \rightarrow  \{0,1\}^{n+\ell}$.
+Here, the set ${\cal D}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
 This means that $F$ needs to be Variable Input Length PRF (VIL-PRF).
@@
 
-%A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {{\mathcal{D}}} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
+%A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {\cal D} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
 The result is a PRF with the required input-output signature~\cite[Section 8.2]{BonehShoup}.
@@
 This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security.
-
-\todo{Sim based of indepednent interest, arguably more powerful than game-based notions with fixed winning predicates.}
 %It is also an open problem to close the performance gap that exists between fast but non-cryptographic hashes used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security.
-
-\ifanon
-\else
-\section*{Acknowledgements}
-The work of Paterson is supported in part by a gift from VMware.
-
-The authors thank Anupama Unnikrishnan for useful discussions on simulation-based definitions.
-\fi
 
@@
 
+\appendix
+
+\section{Initially Non-empty HLL Sketch, Scenario S1}\label{sec:non-empty-HLL-sketch-S1}
+
+We briefly sketch an attack in the setting of an S1 adversary attacking an initially non-empty HLL sketch.
+An alternative attack with moderate performance in this setting is presented in~\cite{hllvuln}.
+
+Since the target sketch has an unknown status for the adversary, our previous strategy for S1 of keeping many buckets empty no longer works.
+One can attempt to use the shadow device to identify items that would result in small entries $M[j]$ (for unknown indices $j$) since such items are unlikely to increase the cardinality estimate of the target device by much.
+This can be done by first filling the shadow device with sufficiently many values to escape from the ``small range correction'' phase.
+Then one could enter candidate values $x$ into the shadow device and check for cardinality increases.
+However this runs into the rounding issue described in Section~\ref{sec:scenarios-discussion}.
+Still, one could retain those items which do lead to a detectable change in cardinality and reject the others (and also estimate the value of $M[j]$ for the unknown bucket $j$ from the magnitude of the change, retaining only those for which $M[j]$ is small).
+But here the lack of a cardinality change could result either from rounding issues or because insertion of candidate value $x$ into the shadow device really did lead to no increase in cardinality.
+This indicates that  ``good'' items would be rejected unnecessarily.
+This could be addressed by using multiple resets of the shadow HLL sketch, filling it with random items after each reset, and performing multiple insertions per candidate item $x$ to try to estimate the true value of $M[j]$ for each item $x$ (since after at least some of the resets followed by random insertions, we could expect $M[j]$ to increase when $x$ is inserted, and if $M[j]$ is small enough relative to the cardinality estimate, the change could be detected).
+The details of such an attack require further analysis which we do not pursue here, since we consider S1 to be the least realistic attack scenario.
+
+\section{Proof of Theorem~\ref{thm:security-of-HLL-with-PRF}}\label{sec:proof}
+
+\begin{proof}
+PRF adversary ${\cal B}$ against $F$ is built as follows.
+Let $d$ denote the hidden bit determining whether ${\cal B}$ receives real responses of the form $F(K,x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${\cal B}$ runs S4 adversary ${\cal A}$ and responds to each of its queries as follows:
+\begin{itemize}
+\item \texttt{init}: here ${\cal B}$ initialises an HLL sketch in accordance with Figure~\ref{fig:hll}. Thus it initialises an array $M[j]$ in which each entry is set to 0. 
+\item \texttt{insert($\cdot$)}: on input $x \in {\cal D}$,  ${\cal B}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${\cal B}$ uses $y$ to update the HLL bucket $j$ generated via $j=1 + <y_1y_2...y_{n}>_2$ as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
+\item \texttt{read}: ${\cal B}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate in accordance with Figure~\ref{fig:hll}. 
+\item \texttt{corrupt}: ${\cal B}$ returns the vector $(M[j])_{j=1}^{m}$ of HLL bucket contents. 
+\end{itemize}
+When ${\cal A}$ outputs a bit $b'$, ${\cal B}$ outputs its own bit $d'=b'$.
+Note that when $d=0$, ${\cal B}$ plays against the real $F$ with a random key $K$ and provides a simulation for ${\cal A}$ that perfectly instantiates the real world.
+On the other hand, when $d=1$, ${\cal B}$ plays against a truly random function and provides a simulation for ${\cal A}$ that perfectly instantiates the ideal world.
+The equality of advantages in the statement of the theorem now follows immediately.
+The remainder of the proof follows on counting queries and noting that the HLL algorithm for computing cardinality is efficient, so ${\cal B}$'s running time due to handling \texttt{read} queries is not increased significantly compared to that of ${\cal A}$.
+\end{proof}
+
 \end{document}

### 2021-05-03 docs/HLL_exp_bad/main.tex
typo
tweaks to intro and elsewhere

--- 2021-04-30 docs/HLL_exp_bad/main.tex
@@
 In many use cases, obtaining an approximate cardinality is sufficient.
-Over the last decade, the HyperLogLog (HLL) cardinality estimator~\cite{hll} has emerged as the algorithm of choice for this task, due to its excellent speed, low memory requirements and high proven accuracy.
-It is also suitable for use in applications where it is not possible (or desirable) to store an entire data set, but instead where the data set is presented in a streaming fashion.
-HLL is used by all of the above-mentioned systems; section~\ref{sec:overview} contains a complete specification of HLL.
-
-However, HLL is increasingly used in settings where adversaries may have incentives to manipulate the cardinality estimates made by the algorithm, i.e.\ HLL is being used in adversarial settings.
+Over the last decade, the HyperLogLog (HLL) cardinality estimator~\cite{hll} has emerged as the algorithm of choice for this task, due to its excellent speed, low memory requirements and high proven accuracy.\footnote{Section~\ref{sec:overview} contains a complete specification of HLL.} It is also suitable for use in applications where it is not possible (or desirable) to store an entire data set, but instead where the data set is presented in a streaming fashion.
+HLL is used by all of the above-mentioned systems.
+
+However, HLL is increasingly used in settings where adversaries may have incentives to manipulate the cardinality estimates made by the algorithm, i.e.\ HLL is being used in \emph{adversarial settings}.
 For example:
 \begin{itemize}
-    \item Cardinality estimation provides a significantly faster way for Facebook to compute statistics on their users.
-They use HLL to find out how many distinct people visited their website in the past week~\cite{fbhll}.
+    \item Facebook uses HLL to find out how many distinct people visited their website in the past week~\cite{fbhll}.
     \item Google uses a variant of HLL described in~\cite{hllpractice} in its PowerDrill system (first described in~\cite{HallBBGN12}) and possibly in a variety of other internal systems reported in~\cite{hllpractice}.
 For example, cardinality estimation is used to to determine the number of distinct search queries on \url{google.com} over a time period~\cite{hllpractice}.
-HLL is also used for cardinality estimation via an API in Google Big Query.
+%HLL is also used for cardinality estimation via an API in Google Big Query.
     \item HLL has been implemented in network soft-switches to approximate the number of distinct packets in traffic flows and overcome the switch memory limits~\cite{flexswitch}.
@@
 
-None of the above-cited examples discuss any kind of threat model where an adversary may try to manipulate the HLL cardinality estimate to its advantage, for example, through input selection.
+None of the above-cited examples discuss any kind of threat model where an adversary may try to manipulate the HLL cardinality estimate to its advantage.
 This is despite them all operating in potentially adversarial settings.
@@
 In particular, we show that with only modest knowledge of the HLL internals and a moderate amount of computation, an adversary with sufficient flexibility in the choice of inputs can make the HLL cardinality estimate exponentially smaller than the true cardinality (e.g.\ constant instead of ${\mathcal{O}}(2^t)$ where $t$ is the number of bits of flexibility in the adversary's input).
-We demonstrate this theoretically and experimentally for the ``classic'' version of HLL.
+We demonstrate this theoretically and experimentally for the ``classic'' version of HLL from~\cite{hll}.
 We also study a specific ``modern'' version of HLL which adopts recommendations from~\cite{hllnew} and show how variants of our attacks apply there too.
-This modern version of HLL is interesting because it is the one used in Redis, a popular key-value database technology.
+This modern version of HLL is interesting because it is the one used in Redis.
 
@@
 Yet HLL is being increasingly widely deployed, and never with protections against our attacks (insofar as we are aware).
-So despite their technical simplicity, we argue that our attacks are important in establishing the insecurity of HLL in adversarial setting.
+So despite their technical simplicity, we argue that our attacks are important in establishing the limitations of HLL in the broadest set of use cases.
 We present our attacks in an abstract way rather than tying them to any specific setting in order to illustrate their generality.
@@
 To support our formal analysis, we introduce a novel simulation-based security model for HLL.
-We believe our model and approach is of independent interest and will be useful beyond the analysis of cardinality manipulation attacks, for example when studying privacy properties of HLL and other probabilistic data structures.
+We believe our simulation-based approach will be of independent interest and will be useful beyond the analysis of cardinality manipulation attacks, for example when studying privacy properties of HLL and other probabilistic data structures.
 
@@
 
-\subsection{Related Work}\label{sec:related}
+\section{Related Work}\label{sec:related}
 
@@
 We leave for future work the analysis of further HLL variants.
-\todo{revise this if we add Google BQ.}
+%\todo{revise this if we add Google BQ.}
 
@@
 
-\todo[inline]{This ends up being pretty harsh on this NSDI paper, but hey. We may shorten or move some details to an appendix.}
+%\todo[inline]{This ends up being pretty harsh on this NSDI paper, but hey. We may shorten or move some details to an appendix.}
 
@@
 We will prove that  ${\mathcal{S}}$ is such that, given any adversary ${\mathcal{A}}$, then any distinguisher $\mathcal{E}$ that can distinguish between the $\textsf{out}_{\mathcal{A}}$ in the real world and $\textsf{out}_{\mathcal{S}}$ in the ideal world can be used to build an adversary ${{\mathcal{B}}}$ which breaks the PRF security of $F$.
-Under the assumption that $F$ is a good PRF, no such efficient distinguisher $\mathcal{E}$ can exist, and we may conclude that (up to PRF security) ${\mathcal{E}}$ cannot tell that it is interacting with the ideal world instead of the real world.
+Under the assumption that $F$ is a good PRF, no such efficient distinguisher $\mathcal{E}$ can exist, and we may conclude that (up to PRF security) no efficient distinguisher can tell that it is interacting with the ideal world instead of the real world.
 
 Why does this proof imply that ${\mathcal{A}}$ cannot manipulate the cardinality estimate?
-To see this, note that (amongst many other possibilities) ${\mathcal{A}}$'s output could be the vector of bucket values $M[j]$ obtained from a \texttt{corrupt} query, enabling a specific ${\mathcal{E}}$ to compute the cardinality estimate for itself and to attempt to use that estimate as a distinguisher between the two worlds.
+To see this, note that (amongst many other possibilities) ${\mathcal{A}}$'s output could be the cardinality estimate obtained from a \texttt{read} query, allowing a specific ${\mathcal{E}}$ to try to use that to distinguish between the two worlds.
 Yet, by the preceding argument, the simulator in the ideal world is able to produce an output that ${\mathcal{E}}$ cannot distinguish from that of ${\mathcal{A}}$ in the real world.
-In particular, our simulator will output whatever ${\mathcal{A}}$ outputs, so if ${\mathcal{A}}$ outputs the result of a \texttt{corrupt} query in the ideal world, then so will ${\mathcal{S}}$.
+In particular, our simulator will output whatever ${\mathcal{A}}$ outputs, so if ${\mathcal{A}}$ outputs the result of a \texttt{read} query in the ideal world, then so will ${\mathcal{S}}$.
 But in the ideal world, ${\mathcal{S}}$ uses random sampling for the $y$ values, so there ${\mathcal{S}}$ (and ${\mathcal{A}}$ that it runs) are operating in the setting where the average case analysis of HLL applies, as per~\cite{hll}.
-So in the ideal world this specific ${\mathcal{E}}$ would end up computing an accurate average case cardinality estimate.
-Since ${\mathcal{E}}$ tries to distinguish the two worlds by computing the cardinality estimate and we already showed that no efficient distinguisher ${{\mathcal{E}}}$ can exist, it must follow that the cardinality estimates produced in the real and ideal worlds are indistinguishable (up to PRF security).
+So in the ideal world this specific ${\mathcal{E}}$ would receive an accurate average case cardinality estimate.
+Since ${\mathcal{E}}$ tries to distinguish the two worlds from the cardinality estimate and we already showed that no efficient distinguisher ${{\mathcal{E}}}$ can exist, it must follow that the cardinality estimates produced in the real and ideal worlds are indistinguishable (up to PRF security).
 It then follows that ${\mathcal{A}}$ has no advantage when it comes to manipulating the HLL-$F$ cardinality estimate.
@@
 An HLL can leak information about items $x$ that were previously inserted.
-Consider the following attack from~\cite{cardestprivacy}: to test whether an item $x$ was previously inserted into the HLL by an honest user (say), the adversary makes the sequence of queries: \texttt{read}, \texttt{insert($x$)}, \texttt{read}.
+Consider the following attack from~\cite{cardestprivacy}: to test whether an item $x$ was previously inserted into the HLL by an honest user (say), the adversary makes a sequence of queries: \texttt{read}, \texttt{insert($x$)}, \texttt{read}.
 If the two cardinality estimates are different, the adversary can conclude definitively that $x$ was \emph{not} previously inserted into the HLL; if they are the same, it can estimate (with some error) that $x$ was previously inserted.\footnote{Errors can arise in two ways when $x$ was not previously inserted: either inserting $x$ changes a bucket value, but the change is masked by rounding when the cardinality estimate is computed, or inserting $x$ does lead to a change in a bucket value because $\rho(w)$ computed from $x$ is too small.} This attack still works even if HLL is implemented using a PRF and is unavoidable in view of the oracle queries (and in real settings, the API) available to the adversary.
@@
 
+%\todo[inline]{
 %Still, simulation-based security notions are often used to model confidentiality properties, so it is tempting to think that our security definition might say something about confidentiality of the HLL inputs $x$.
@@
 Thus, the cardinality estimates and even the bucket values themselves do not leak information about the honestly inserted values $x$ (beyond their number).
+}
 

### 2021-05-03 docs/HLL_exp_bad/main.tex
creating appendices

@@
 The idea is to use HLL so as to reduce the amount of memory needed in the switches.
+The resulting switches are called \emph{flexswitches} in~\cite{flexswitch}.
 In the main example considered in~\cite[Section 3.1]{flexswitch}, each flow is identified by a 4-tuple consisting of source IP address, destination IP address, source port, and destination port.
@@
 Depending on how the flow count is used, this could lead to undetected DoS conditions, mis-provisioning of network capacity, or under-billing of the adversary for its traffic.
-
-Going further, \cite[Section 4.2]{flexswitch} proposes using HLL for detecting TCP port scans by counting observed destination port numbers in TCP packets with set SYN flags and reporting a scan event if the estimate exceeds a
-set threshold.
-Using our attack techniques, an attacker could complete a partial port scan of roughly half of all $2^{16}$ ports whilst keeping the HLL estimate below $0.693m$ where $m$ is the number of buckets used in HLL.
-While~\cite{flexswitch} is not specific about the value of $m$ actually used, the authors' focus on reducing memory usage indicates that $m$ would not be larger than $1024$.\footnote{The paper mentions using 1 byte counters in the buckets and states that using more than 1KB of memory ``only improves average error rates marginally''; we infer a maximum value of $m=1024$ from this.} It is not stated in~\cite{flexswitch} what threshold needs to be crossed to trigger an alarm, but we consider it plausible that it would be above $0.693 \cdot 1024 = 710$.
-
-Similarly,~\cite[Section 4.2]{flexswitch} proposes using HLL for detecting NTP amplification attacks by counting distinct source IP addresses arriving on UDP port 123 over a small time window.
-No further details are given, but it is clear that an attacker with even limited flexibility in spoofing source IP addresses (e.g.\ within a private range used in the datacenter) could easily evade detection whilst delivering a large number of NTP requests, thus re-enabling the original attack.
-
-In a more advanced application,~\cite[Section 4.3]{flexswitch} proposes to use HLL as a component in a novel, low-memory version of the RCP congestion control algorithm called RCP-Approx.
-Again HLL is used for estimating numbers of traffic flows.
-Obviously, our attack techniques could be used (in scenario S2) to cause RCP-Approx to underestimate the number of traffic flows, leading to non-detection of congestion and so enabling DoS attacks to go undetected.
-
-Similar remarks apply to the treatment of CONGA in~\cite[Section 4.5]{flexswitch}.
-CONGA is a congestion-aware load balancing system that splits TCP flows into flowlets (collections of IP packets) and allocates them to paths based on congestion feedback from remote switches.
-In the enhancements to CONGA proposed in~\cite{flexswitch}, HLL is used in the ``flow statistics'' building block for estimating flowlet statistics in switches; these estimates are sent back to a monitoring system and used to determine the load balancing configuration.
-This overall system can be abused by an S2 adversary, to cause flowlet numbers to be underestimated (or overestimated) and the load balancing to perform sub-optimally.
-For example, an adversary could cause all the switches to \emph{over}-estimate the number of flowlets they are handling and ask the load balancing system to send less traffic through them.
-This could result in a degradation in the total amount of traffic the load balancing system can handle (depending on what fail-safe mode it has).
-Or an attacker could cause all the switches to \emph{under}-estimate the number of flowlets they are handling, causing undetected congestion in the switches.
+In Appendix~\ref{sec:flexswitch} we provide further analysis of~\cite{flexswitch}, showing that each proposed use of HLL can be subverted by an S2 adversary.
 
@@
 
-Now we are ready to formally state and prove our main result:
+Now we are ready to formally state our main result. The proof can be found in Appendix~\ref{sec:proof}.
 \begin{theorem}\label{thm:security-of-HLL-with-PRF}
@@
 
-\begin{proof}
+Note that our security analysis using a simulation-based definition does not require us to specify a winning condition for the adversary.
+Rather, we infer from inspecting the simulator what the adversary can (or cannot) achieve in the ideal world, and show that no efficient adversary can distinguish the real and ideal worlds.
+This is in contrast to the game-based approach of~\cite{CCS:ClaPatShrPS19} which needs a complex definition to say when the adversary has successfully manipulated a probabilistic data structure.
+This reveals a strength of the simulation-based approach in more complex settings like ours.
+
+\subsection{On Confidentiality}
+
+An HLL can leak information about items $x$ that were previously inserted.
+Consider the following attack from~\cite{cardestprivacy}: to test whether an item $x$ was previously inserted into the HLL by an honest user (say), the adversary makes a sequence of queries: \texttt{read}, \texttt{insert($x$)}, \texttt{read}.
+If the two cardinality estimates are different, the adversary can conclude definitively that $x$ was \emph{not} previously inserted into the HLL; if they are the same, it can estimate (with some error) that $x$ was previously inserted.\footnote{Errors can arise in two ways when $x$ was not previously inserted: either inserting $x$ changes a bucket value, but the change is masked by rounding when the cardinality estimate is computed, or inserting $x$ does lead to a change in a bucket value because $\rho(w)$ computed from $x$ is too small.} This attack still works even if HLL is implemented using a PRF and is unavoidable in view of the oracle queries (and in real settings, the API) available to the adversary.
+So such privacy leaks are an inherent feature of HLL if a rich enough API is provided.
+Note that this attack does not contradict our main result, which concerns prevention of cardinality manipulation rather than confidentiality.
+
+%\todo[inline]{
+%Still, simulation-based security notions are often used to model confidentiality properties, so it is tempting to think that our security definition might say something about confidentiality of the HLL inputs $x$.
+Our existing model does not, since ${\mathcal{A}}$ chooses all the inputs $x$ and so ``knows'' them already.
+However, it is easily modified to provide a confidentiality notion.
+For example, we can consider the insertion of honestly generated items $x$ that are drawn from some probability distribution $p$ on $\mathcal{D}$ in a first phase, with ${\mathcal{A}}$ then having access to only \texttt{read} and \texttt{corrupt} oracles in a second phase (we have to remove the \texttt{insert($\cdot$)} oracle in view of the general attack from~\cite{cardestprivacy}).
+
+%In this case, the simulator ${\mathcal{S}}$ in the ideal world can proceed without ever seeing any inputs but instead knowing only the distribution $p$ and the number of distinct samples drawn from it in the first phase.
+It is not hard to modify the proof of Theorem~\ref{thm:security-of-HLL-with-PRF} to show that no adversary can distinguish real and ideal worlds in this case (up to PRF security).
+Now, since ${\mathcal{S}}$ has no access to inputs $x$, we can conclude that the HLL implemented using a PRF ensures confidentiality for the inputs $x$.
+Thus, the cardinality estimates and even the bucket values themselves do not leak information about the honestly inserted values $x$ (beyond their number).
+}
+
+We leave the exploration of the adaptation of our simulation-based security definitions to capture confidentiality notions for more restrictred APIs to future work.
+
+\subsection{Is Universal Hashing Sufficient?}
+
+It is natural to wonder whether a PRF is really required in place of hash function $h$ in order to make HLL secure, or whether a weaker cryptographic primitive would suffice.
+For example, since it has (roughly) uniform outputs, perhaps using a universal hash function (UHF) $h_K(\cdot)$ would be sufficient.
+In Appendix~\ref{app:UHF} we show that, in general, UHFs do not suffice by providing a key recovery attack against HLL when $h$ is replaced by a specific UHF based on polynomial evaluation over a finite field.
+
+\subsection{Instantiating the PRF}
+
+Our HLL construction requires a PRF $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
+Here, the set ${{\mathcal{D}}}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
+This means that $F$ needs to be Variable Input Length PRF (VIL-PRF).
+
+We want $F$ to be competitive with non-cryptographic hashing on small input sizes, as typical in current HLL implementations and applications (e.g.\ MurmurHash3 applied to an IP header).
+This makes off-the-shelf approaches like HMAC uncompetitive, since even for the shortest inputs, HMAC would require 4 evaluations of the underlying hash compression function.
+
+SipHash~\cite{AumassonB12} is designed specifically for this use case.
+It is competitive in speed with non-cryptographic hash functions commonly used in HLL implementations such as MurmurHash3.
+It has also withstood cryptanalysis well~\cite{DobraunigMS14,XinLSL19}.
+It has 128-bit outputs that can be truncated to the needs of typical HLL implementations (32 and 64 bits).
+%and is well supported in libraries such as OpenSSL.
+Our microbenchmarks created on an Intel i7 CPU show that SipHash-2-4 on 90-byte inputs requires 102ns per evaluation, while the 128-bit version of MurmurHash3 takes 60ns.
+Meanwhile, HMAC-SHA256 requires 705ns.
+We obtained similar performance ratios for other input sizes.
+These together show that SipHash is quite competitive with MurmurHash3.
+
+%is as fast as the 128-bit version of MurmurHash3 and only slightly slower than the original 32-bit version.
+
+%length: 720, time h_mac 705ns, time h_murmur 60ns and time h_sip 102ns
+
+%Making $F$ competitive in speed with existing implementations of HLL, which use non-cryptographic hash functions such as Murmur3, is challenging, and in practice we must sacrifice some speed for security.
+
+%A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {{\mathcal{D}}} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
+The result is a PRF with the required input-output signature~\cite[Section 8.2]{BonehShoup}.
+As a concrete instantiation, one could take $H$ to be SHA-256 with outputs truncated to 128 bits and $G$ to be AES-128 truncated to ${n+\ell}$ bits (since ${n+\ell} \leq 64 < 128$ in practice, this truncation is always possible).
+
+%An alternative is to simply use HMAC-SHA-256 with an appropriately truncated output. HMAC is a PRF under reasonable assumptions on the hash function used in its construction~\cite{Bellare15}.
+
+%These constructions would be relatively fast on commodity platforms due to widespread hardware support for SHA-256 and AES operations.
+It would be a simple exercise to add the above PRF proposals to an existing HLL implementation and benchmark them against, say, Murmur3.
+The exact slowdown depends on platform specifics, but based on microbenchmarks can be expected to be in the range 2-fold to 5-fold compared to Murmur3.
+\todo{Put a hard number here.}
+
+%More complex but faster constructions using randomised or nonce-based MACs constructed from universal hashing could be envisaged.
+However, these do not seem to be directly applicable here since we need the function $F$ to be both stateless and deterministic so as to cope with the repeated inputs one would encounter during the normal operation of HLL.
+
+We note that using keyed functions sacrifices the mergeability property that HLL enjoys, if different keys $K$ are used in different HLL sketches.
+The reason is that one can no longer emulate running a single HLL sketch by taking the maximum over all buckets with each index $j$ in the different sketches, since now the bucket contents are created using different keys in a PRF and hence bear no relation to one another.
+
+%\todo{Discussion: note again that no-one currently does this in practice; also point out that you want different keys in different sketches for security, but this sacrifices the merge capability of HLL.}
+
+\section{Conclusion}\label{sec:conclusions}
+The HyperLogLog algorithm is an elegant and efficient solution to the problem of estimating the cardinality of large sets.
+Its simple structure makes it easy to code and use, as shown by the growing number of available open-source implementations.
+Nonetheless, we have shown that malicious users can manipulate the HLL cardinality estimate and thence break the security properties of systems relying on HLL.
+Our attacks are simple but powerful and should raise awareness of the limitations of HLL.
+Our analysis may assist software developers in understanding the risks they run when using HLL in adversarial settings.
+We have also provided a simple method for securing HLL in adversarial settings: simply replace the internal hash function by a keyed PRF, and one obtains full protection even in scenario S4, our strongest attack scenario.
+
+%Future work could consist in finding mitigations that do not rely on salt.
+For example, would the imposed ID setting (like Facebook) solve the issue but still be convenient for the various HLL applications?
+Left to future work is the task of extending our attacks and defences to HLL variants such as sliding HLL~\cite{slidinghll} and HLL++~\cite{hllpractice}.
+It would be interesting to find and analyse a strategy for more cleverly ordering the candidates in our attack in scenario S1 to reduce the total insertion cost of the attack, whilst also keeping the number of resets low.
+Finding an attack in scenario S1 when the target HLL sketch is initially non-empty is also open.
+
+We have given a formal analysis of using a PRF as a countermeasure to our attacks.
+The use of a keyed function sacrifices HLL mergeability when the keys in the different HLLs are independently generated.
+It is an open problem to find methods for securing HLL sketches against adversarial inputs whilst retaining some form of mergeability.
+This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security.
+
+%\todo{Sim based of indepednent interest, arguably more powerful than game-based notions with fixed winning predicates.}
+%It is also an open problem to close the performance gap that exists between fast but non-cryptographic hashes used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security.
+
+\ifanon
+\else
+\section*{Acknowledgements}
+The work of Paterson is supported in part by a gift from VMware.
+
+The authors thank Anupama Unnikrishnan for useful discussions on simulation-based definitions.
+\fi
+
+\bibliographystyle{ieeetr}
+\bibliography{ref.bib}
+
+\appendix
+
+\section{Further analysis of Flexswitches}\label{sec:flexswitch}
+
+Extending our analysis of~\cite{flexswitch} from the main body, \cite[Section 4.2]{flexswitch} proposes using HLL for detecting TCP port scans by counting observed destination port numbers in TCP packets with set SYN flags and reporting a scan event if the estimate exceeds a
+set threshold.
+Using our attack techniques (specifically, in scenario S2), an attacker could complete a partial port scan of roughly half of all $2^{16}$ ports whilst keeping the HLL estimate below $0.693m$ where $m$ is the number of buckets used in HLL.
+While~\cite{flexswitch} is not specific about the value of $m$ actually used, the authors' focus on reducing memory usage indicates that $m$ would not be larger than $1024$.\footnote{The paper mentions using 1 byte counters in the buckets and states that using more than 1KB of memory ``only improves average error rates marginally''; we infer a maximum value of $m=1024$ from this.} It is not stated in~\cite{flexswitch} what threshold needs to be crossed to trigger an alarm, but we consider it plausible that it would be above $0.693 \cdot 1024 = 710$.
+
+Similarly,~\cite[Section 4.2]{flexswitch} proposes using HLL for detecting NTP amplification attacks by counting distinct source IP addresses arriving on UDP port 123 over a small time window.
+No further details are given, but it is clear that an S2 attacker with even limited flexibility in spoofing source IP addresses (e.g.\ within a private range used in the datacenter) could easily evade detection whilst delivering a large number of NTP requests, thus re-enabling the original attack.
+
+In a more advanced application,~\cite[Section 4.3]{flexswitch} proposes to use HLL as a component in a novel, low-memory version of the RCP congestion control algorithm called RCP-Approx.
+Again HLL is used for estimating numbers of traffic flows.
+Obviously, our attack techniques could be used (in scenario S2) to cause RCP-Approx to underestimate the number of traffic flows, leading to non-detection of congestion and so enabling DoS attacks to go undetected.
+
+Similar remarks apply to the treatment of CONGA in~\cite[Section 4.5]{flexswitch}.
+CONGA is a congestion-aware load balancing system that splits TCP flows into flowlets (collections of IP packets) and allocates them to paths based on congestion feedback from remote switches.
+In the enhancements to CONGA proposed in~\cite{flexswitch}, HLL is used in the ``flow statistics'' building block for estimating flowlet statistics in switches; these estimates are sent back to a monitoring system and used to determine the load balancing configuration.
+This overall system can be abused by an S2 adversary, to cause flowlet numbers to be underestimated (or overestimated) and the load balancing to perform sub-optimally.
+For example, an adversary could cause all the switches to \emph{over}-estimate the number of flowlets they are handling and ask the load balancing system to send less traffic through them.
+This could result in a degradation in the total amount of traffic the load balancing system can handle (depending on what fail-safe mode it has).
+Or an attacker could cause all the switches to \emph{under}-estimate the number of flowlets they are handling, causing undetected congestion in the switches.
+
+\section{Proof of Theorem 8.1}\label{sec:proof} %ridiculous, but had to hard-code this reference
+
 PRF adversary ${{\mathcal{B}}}$ against $F$ is built as follows.
@@
 The final running time for ${\mathcal{B}}$, namely $t_{\mathcal{A}}+t_{\mathcal{E}}$, follows on noting that ${\mathcal{B}}$ first runs ${\mathcal{A}}$ and then ${\mathcal{E}}$.
-\end{proof}
-
-Note that our security analysis using a simulation-based definition does not require us to specify a winning condition for the adversary.
-Rather, we infer from inspecting the simulator what the adversary can (or cannot) achieve in the ideal world, and show that no efficient adversary can distinguish the real and ideal worlds.
-This is in contrast to the game-based approach of~\cite{CCS:ClaPatShrPS19} which needs a complex definition to say when the adversary has successfully manipulated a probabilistic data structure.
-This reveals a strength of the simulation-based approach in more complex settings like ours.
-
-\subsection{On Confidentiality}
-
-An HLL can leak information about items $x$ that were previously inserted.
-Consider the following attack from~\cite{cardestprivacy}: to test whether an item $x$ was previously inserted into the HLL by an honest user (say), the adversary makes a sequence of queries: \texttt{read}, \texttt{insert($x$)}, \texttt{read}.
-If the two cardinality estimates are different, the adversary can conclude definitively that $x$ was \emph{not} previously inserted into the HLL; if they are the same, it can estimate (with some error) that $x$ was previously inserted.\footnote{Errors can arise in two ways when $x$ was not previously inserted: either inserting $x$ changes a bucket value, but the change is masked by rounding when the cardinality estimate is computed, or inserting $x$ does lead to a change in a bucket value because $\rho(w)$ computed from $x$ is too small.} This attack still works even if HLL is implemented using a PRF and is unavoidable in view of the oracle queries (and in real settings, the API) available to the adversary.
-So such privacy leaks are an inherent feature of HLL if a rich enough API is provided.
-Note that this attack does not contradict our main result, which concerns prevention of cardinality manipulation rather than confidentiality.
-
-%\todo[inline]{
-%Still, simulation-based security notions are often used to model confidentiality properties, so it is tempting to think that our security definition might say something about confidentiality of the HLL inputs $x$.
-Our existing model does not, since ${\mathcal{A}}$ chooses all the inputs $x$ and so ``knows'' them already.
-However, it is easily modified to provide a confidentiality notion.
-For example, we can consider the insertion of honestly generated items $x$ that are drawn from some probability distribution $p$ on $\mathcal{D}$ in a first phase, with ${\mathcal{A}}$ then having access to only \texttt{read} and \texttt{corrupt} oracles in a second phase (we have to remove the \texttt{insert($\cdot$)} oracle in view of the general attack from~\cite{cardestprivacy}).
-
-%In this case, the simulator ${\mathcal{S}}$ in the ideal world can proceed without ever seeing any inputs but instead knowing only the distribution $p$ and the number of distinct samples drawn from it in the first phase.
-It is not hard to modify the proof of Theorem~\ref{thm:security-of-HLL-with-PRF} to show that no adversary can distinguish real and ideal worlds in this case (up to PRF security).
-Now, since ${\mathcal{S}}$ has no access to inputs $x$, we can conclude that the HLL implemented using a PRF ensures confidentiality for the inputs $x$.
-Thus, the cardinality estimates and even the bucket values themselves do not leak information about the honestly inserted values $x$ (beyond their number).
-}
-
-We leave the exploration of the adaptation of our simulation-based security definitions to capture confidentiality notions for more restrictred APIs to future work.
-
-\subsection{Is Universal Hashing Sufficient?}
-
-It is natural to wonder whether a PRF is really required in place of hash function $h$ in order to make HLL secure, or whether a weaker cryptographic primitive would suffice.
-For example, since it has (roughly) uniform outputs, perhaps using a universal hash function (UHF) $h_K(\cdot)$ would be sufficient.
+
+\section{Breaking HLL-UHF}\label{sec:UHF}
 
@@
 
-\subsection{Instantiating the PRF}
-
-Our HLL construction requires a PRF $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
-Here, the set ${{\mathcal{D}}}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
-This means that $F$ needs to be Variable Input Length PRF (VIL-PRF).
-
-We want $F$ to be competitive with non-cryptographic hashing on small input sizes, as typical in current HLL implementations and applications (e.g.\ MurmurHash3 applied to an IP header).
-This makes off-the-shelf approaches like HMAC uncompetitive, since even for the shortest inputs, HMAC would require 4 evaluations of the underlying hash compression function.
-
-SipHash~\cite{AumassonB12} is designed specifically for this use case.
-It is competitive in speed with non-cryptographic hash functions commonly used in HLL implementations such as MurmurHash3.
-It has also withstood cryptanalysis well~\cite{DobraunigMS14,XinLSL19}.
-It has 128-bit outputs that can be truncated to the needs of typical HLL implementations (32 and 64 bits).
-%and is well supported in libraries such as OpenSSL.
-Our microbenchmarks created on an Intel i7 CPU show that SipHash-2-4 on 90-byte inputs requires 102ns per evaluation, while the 128-bit version of MurmurHash3 takes 60ns.
-Meanwhile, HMAC-SHA256 requires 705ns.
-We obtained similar performance ratios for other input sizes.
-These together show that SipHash is quite competitive with MurmurHash3.
-
-%is as fast as the 128-bit version of MurmurHash3 and only slightly slower than the original 32-bit version.
-
-%length: 720, time h_mac 705ns, time h_murmur 60ns and time h_sip 102ns
-
-%Making $F$ competitive in speed with existing implementations of HLL, which use non-cryptographic hash functions such as Murmur3, is challenging, and in practice we must sacrifice some speed for security.
-
-%A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {{\mathcal{D}}} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
-The result is a PRF with the required input-output signature~\cite[Section 8.2]{BonehShoup}.
-As a concrete instantiation, one could take $H$ to be SHA-256 with outputs truncated to 128 bits and $G$ to be AES-128 truncated to ${n+\ell}$ bits (since ${n+\ell} \leq 64 < 128$ in practice, this truncation is always possible).
-
-%An alternative is to simply use HMAC-SHA-256 with an appropriately truncated output. HMAC is a PRF under reasonable assumptions on the hash function used in its construction~\cite{Bellare15}.
-
-%These constructions would be relatively fast on commodity platforms due to widespread hardware support for SHA-256 and AES operations.
-It would be a simple exercise to add the above PRF proposals to an existing HLL implementation and benchmark them against, say, Murmur3.
-The exact slowdown depends on platform specifics, but based on microbenchmarks can be expected to be in the range 2-fold to 5-fold compared to Murmur3.
-\todo{Put a hard number here.}
-
-%More complex but faster constructions using randomised or nonce-based MACs constructed from universal hashing could be envisaged.
-However, these do not seem to be directly applicable here since we need the function $F$ to be both stateless and deterministic so as to cope with the repeated inputs one would encounter during the normal operation of HLL.
-
-We note that using keyed functions sacrifices the mergeability property that HLL enjoys, if different keys $K$ are used in different HLL sketches.
-The reason is that one can no longer emulate running a single HLL sketch by taking the maximum over all buckets with each index $j$ in the different sketches, since now the bucket contents are created using different keys in a PRF and hence bear no relation to one another.
-
-%\todo{Discussion: note again that no-one currently does this in practice; also point out that you want different keys in different sketches for security, but this sacrifices the merge capability of HLL.}
-
-\section{Conclusion}\label{sec:conclusions}
-The HyperLogLog algorithm is an elegant and efficient solution to the problem of estimating the cardinality of large sets.
-Its simple structure makes it easy to code and use, as shown by the growing number of available open-source implementations.
-Nonetheless, we have shown that malicious users can manipulate the HLL cardinality estimate and thence break the security properties of systems relying on HLL.
-Our attacks are simple but powerful and should raise awareness of the limitations of HLL.
-Our analysis may assist software developers in understanding the risks they run when using HLL in adversarial settings.
-We have also provided a simple method for securing HLL in adversarial settings: simply replace the internal hash function by a keyed PRF, and one obtains full protection even in scenario S4, our strongest attack scenario.
-
-%Future work could consist in finding mitigations that do not rely on salt.
-For example, would the imposed ID setting (like Facebook) solve the issue but still be convenient for the various HLL applications?
-Left to future work is the task of extending our attacks and defences to HLL variants such as sliding HLL~\cite{slidinghll} and HLL++~\cite{hllpractice}.
-It would be interesting to find and analyse a strategy for more cleverly ordering the candidates in our attack in scenario S1 to reduce the total insertion cost of the attack, whilst also keeping the number of resets low.
-Finding an attack in scenario S1 when the target HLL sketch is initially non-empty is also open.
-
-We have given a formal analysis of using a PRF as a countermeasure to our attacks.
-The use of a keyed function sacrifices HLL mergeability when the keys in the different HLLs are independently generated.
-It is an open problem to find methods for securing HLL sketches against adversarial inputs whilst retaining some form of mergeability.
-This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security.
-
-%\todo{Sim based of indepednent interest, arguably more powerful than game-based notions with fixed winning predicates.}
-%It is also an open problem to close the performance gap that exists between fast but non-cryptographic hashes used in HLL implementations today and the slower but cryptographically strong VIL-PRF that is needed to obtain security.
-
-\ifanon
-\else
-\section*{Acknowledgements}
-The work of Paterson is supported in part by a gift from VMware.
-
-The authors thank Anupama Unnikrishnan for useful discussions on simulation-based definitions.
-\fi
-
-\bibliographystyle{ieeetr}
-\bibliography{ref.bib}
-
 \end{document}

### 2021-05-03 docs/HLL_exp_bad/main.tex
tweaks to expt section

@@
 We stress that this choice of inputs is merely illustrative and the results do not in any way depend on choosing ASCII strings.
-We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}.
+We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}; again the attacks do not depend on the specifics of what fixed hash function is used.
 %(note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
@@
 
-We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $2^{22.8}$ possible inputs while trying to keep the cardinality estimate as low as possible, in each of the scenarios S1, S2 or S4.
+We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $2^{22.8}$ possible inputs while trying to keep the cardinality estimate as low as possible, in each of the scenarios S1, S2 and S4.
 

### 2021-05-03 docs/HLL_exp_bad/main.tex
removing draft flag  and adding submission number

@@
 
-\documentclass[sigconf, anonymous, dvipsnames, draft]{acmart} %dvipsnames for old/new macros
+\documentclass[sigconf, anonymous, dvipsnames]{acmart} %dvipsnames for old/new macros
 
 \fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Anonymous submission \#9999 to ACM CCS 2021} % TODO: replace 9999 with your paper number
+\fancyhead[C]{Anonymous submission \#147 to ACM CCS 2021} % TODO: replace 9999 with your paper number
 \fancyfoot[C]{\thepage}

### 2021-05-05 docs/HLL_exp_bad/main.tex
updates after feedback from Mia and Anu

@@
 We present experimental results confirming our theoretical analysis.
-Finally, we consider attack prevention: we show to modify HLL in a simple way that provably prevents cardinality estimate manipulation attacks.
+Finally, we consider attack prevention: we show how to modify HLL in a simple way that provably prevents cardinality estimate manipulation attacks.
 \end{abstract}
@@
 In this section we present a formal security analysis of HLL, focussing on how to augment the original HLL design with cryptographic mechanisms and thereby provably prevent attacks.
+We focus on the case of an initially empty HLL sketch, giving the adversary full control over which items are inserted into it.
 
@@
 We allow an adversary ${\mathcal{A}}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios S2, S3 and S4: it may have an \texttt{insert($\cdot$)} oracle allowing it to add items to the HLL sketch, a \texttt{read} oracle that returns the HLL cardinality estimate, and a \texttt{corrupt} that returns all the HLL bucket contents (but not any secret key $K$); in S4 it has all three oracles.
-In addition, in all 3 scenarios, ${\mathcal{A}}$ has access to an \texttt{init($\cdot$)} oracle by which it sets the HLL parameters $n$, $\ell$, $\alpha_n$ used to initialise the HLL sketch.
+In addition, in all 3 scenarios, ${\mathcal{A}}$ has access to an \texttt{init($\cdot$)} oracle by which it sets the HLL parameters $n$, $m=2^n$, $\ell$, $\alpha_m$ used to initialise the HLL sketch.
 We assume ${\mathcal{A}}$ always makes exactly one \texttt{init($\cdot$)} query and that it is its first oracle call.
@@
 In the ideal world ($b=1$), ${\mathcal{A}}$ is replaced by a simulator ${\mathcal{S}}$ with blackbox oracle access to ${\mathcal{A}}$ (that is, ${\mathcal{S}}$ can run ${\mathcal{A}}$ in a blackbox manner, receiving its oracle queries and providing responses to ${\mathcal{A}}$ as it sees fit).
-In particular, ${\mathcal{S}}$ receives parameters $n$, $\ell$, $\alpha_n$ from ${\mathcal{A}}$'s \texttt{init($\cdot$)} call.
+In particular, ${\mathcal{S}}$ receives parameters $n$, $\ell$, $\alpha_m$ from ${\mathcal{A}}$'s \texttt{init($\cdot$)} call.
 Finally, ${\mathcal{S}}$ receives $\textsf{out}_{\mathcal{A}}$ and completes its execution by producing its own output $\textsf{out}_{\mathcal{S}}$.
@@
 \begin{itemize}
-\item \texttt{init($\cdot$)}: this oracle receives as input the HLL-$F$ parameters ($n$, $\ell$, $\alpha_n$) and performs initialisation of the HLL-$F$ sketch, including generating the array of buckets $M[j]$ and initialising all the entries to $0$, and selecting a key $K \leftarrow_{\$} \{0,1\}^k$ for keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
+\item \texttt{init($\cdot$)}: this oracle receives as input the HLL-$F$ parameters ($n$, $\ell$, $\alpha_m$) and performs initialisation of the HLL-$F$ sketch, including generating the array of $m=2^n$ buckets $M[j]$ and initialising all the entries to $0$, and selecting a key $K \leftarrow_{\$} \{0,1\}^k$ for keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
 This oracle has no output.
@@
 We assume it is only called once as ${\mathcal{A}}$'s first oracle query.
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$, the oracle implements HLL-$F$ by setting $y = F_K(x)$ and following the procedure in Figure~\ref{fig:hll} to insert $x$ into the HLL sketch (that is, it uses the first $m$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
+\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$, the oracle implements HLL-$F$ by setting $y = F_K(x)$ and following the procedure in Figure~\ref{fig:hll} to insert $x$ into the HLL sketch (that is, it uses the first $n$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
 This oracle has no output.
 It is available in all of S2, S3, S4.
-Without loss of generality (since HLL and variants of it we consider here have a deterministic insertion procedure), we assume all queries $x$ made by ${\mathcal{A}}$ to \texttt{insert($\cdot$)} are distinct.
+Since HLL and variants of it we consider here have the property that reinserting an item $x$ does not change the internal state of the HLL sketch (nor the cardinality estimate computed from it) we assume without loss of generality that all queries $x$ made by ${\mathcal{A}}$ to \texttt{insert($\cdot$)} are distinct.
 \item \texttt{read}: this oracle has no input and returns the HLL-$F$ cardinality estimate as per Figure~\ref{fig:hll}, i.e.\ it computes a raw cardinality estimate via $E := \alpha_mm^2\cdot(\sum_{j=1}^m2^{-M[j]})^{-1}$ and then performs corrections for small and large values.
@@
 
-We leave the exploration of the adaptation of our simulation-based security definitions to capture confidentiality notions for more restrictred APIs to future work.
+We leave the exploration of the adaptation of our simulation-based security definitions to capture confidentiality notions for more restricted APIs to future work.
 
@@
 For example, since it has (roughly) uniform outputs, perhaps using a universal hash function (UHF) $h_K(\cdot)$ would be sufficient.
-In Appendix~\ref{app:UHF} we show that, in general, UHFs do not suffice by providing a key recovery attack against HLL when $h$ is replaced by a specific UHF based on polynomial evaluation over a finite field.
+In Appendix~\ref{sec:UHF} we show that, in general, UHFs do not suffice by providing a key recovery attack against HLL when $h$ is replaced by a specific UHF based on polynomial evaluation over a finite field.
 
@@
 PRF adversary ${{\mathcal{B}}}$ against $F$ is built as follows.
-Let $d$ denote the hidden bit determining whether ${{\mathcal{B}}}$ receives real responses of the form $F_K(x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${{\mathcal{B}}}$ provides a simulation of either the real or random world for distinguisher ${\mathcal{E}}$ as follows.
+Let $d$ denote the hidden bit determining whether ${{\mathcal{B}}}$ receives real responses of the form $F_K(x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${{\mathcal{B}}}$ provides a simulation of either the real or ideal world for distinguisher ${\mathcal{E}}$ as follows.
 ${{\mathcal{B}}}$ runs ${{\mathcal{A}}}$ and answers each of ${{\mathcal{A}}}$'s queries as follows:
 \begin{itemize}
-\item \texttt{init($\cdot$)}: ${{\mathcal{B}}}$ receives parameters $n$, $\ell$, $\alpha_m$ and initialises an array $M[j]$ in which each of the $2^n$ entries is set to 0. 
+\item \texttt{init($\cdot$)}: ${{\mathcal{B}}}$ receives parameters $n$, $\ell$, $\alpha_m$, sets $m=2^n$ and initialises an array $M[j]$ of size $m$ in which each of the entries is set to 0. 
 \item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$,  ${{\mathcal{B}}}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${{\mathcal{B}}}$ uses $y$ to update the HLL sketch as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
-That is, it uses the first $m$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$.
+That is, it uses the first $n$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$.
 \item \texttt{read}: ${{\mathcal{B}}}$ uses its knowledge of the buckets' contents to compute and return the HLL cardinality estimate as per Figure~\ref{fig:hll}. 
@@
 \end{itemize}
-When ${\mathcal{A}}$ returns its final output $\textsf{out}_{\mathcal{S}}$, ${{\mathcal{B}}}$ forwards this output to ${{\mathcal{E}}}$.
+When ${\mathcal{A}}$ returns its final output $\textsf{out}_{\mathcal{A}}$, ${{\mathcal{B}}}$ forwards this output to ${{\mathcal{E}}}$.
 Finally, ${{\mathcal{E}}}$ outputs a bit $b'$.

### 2021-05-05 docs/HLL_exp_bad/main.tex
updates after feedback from Mia and Anu

@@
 We now illustrate what effect such attacks might have on applications relying on HLL for cardinality estimation.
-Specifically, we analyse in detail a recent paper~\cite{flexswitch} from the networking systems community.
+Specifically, we analyse a recent paper~\cite{flexswitch} from the networking systems community.
 
@@
 Depending on how the flow count is used, this could lead to undetected DoS conditions, mis-provisioning of network capacity, or under-billing of the adversary for its traffic.
-In Appendix~\ref{sec:flexswitch} we provide further analysis of~\cite{flexswitch}, showing that each proposed use of HLL can be subverted by an S2 adversary.
+In Appendix~\ref{sec:flexswitch} we provide further analysis showing that each use of HLL proposed in~\cite{flexswitch} can be subverted by an S2 adversary.
 

### 2021-05-05 docs/HLL_exp_bad/main.tex
removing last todo to make something that can be submitted
adding info on sim-based defns

@@
 This algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 4\%, using only 1.5 kB of memory.
-Crucially, the proof of this~\cite[Theorem 1]{hll} makes the assumption that the inputs are drawn independently and uniformly at random from the input domain.\footnote{Technically, the results of~\cite{hll} are proven for a distribution obtained by making arbitrary replications and permutations after sampling independently and uniformly, to simulate input taken from an \emph{ideal multiset}.
+Crucially, the proof of this~\cite[Theorem 1]{hll} makes the assumption that the inputs are drawn independently and uniformly at random from the input domain -- loosely, it is an average case analysis.\footnote{Technically, the results of~\cite{hll} are proven for a distribution obtained by making arbitrary replications and permutations after sampling independently and uniformly, to simulate input taken from an \emph{ideal multiset}.
 But they also hold without replication and permutation because HLL's internal state is invariant under repeated inputs.}
@@
 
-\todo[inline]{Results from Redis or GBQ if we get them.}
+%\todo[inline]{Results from Redis or GBQ if we get them.}
 
@@
 
+Note also that, for our definition to meaningfully capture security against cardinality manipulation, it is important to restrict  ${\mathcal{S}}$ to doing random sampling when handling \texttt{insert($\cdot$)} queries.
+Otherwise, our definition could be vacuously satisfied (for example ${\mathcal{S}}$ could just sample the PRF key $K$ for itself and provide a perfect real-world view).
+This is somewhat different to common simulation-based security definitions, where ${\mathcal{S}}$'s input might be restricted (e.g.\ in a confidentiality definition, ${\mathcal{S}}$ might receive only $|m|$ in place of $m$) but its behaviour is not.
+Crucially, in our setting, security arises from inspecting the details of ${\mathcal{S}}$ and observing how it instantiates HLL in the average case.
+
 \subsection{On Confidentiality}

### 2021-05-07 docs/HLL_exp_bad/main.tex
Sec7: add redis attack; change the input set; add sentence for inputs form,

@@
 \centering
-\caption{Adversarial capabilities in scenarios S1, S2, S3 and S4.}
+\caption{Adversarial capabilities in scenarios S1, S2, S3, and S4.}
 \medskip
@@
 
-As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}, since it is faithful to the original description of HLL given in~\cite{hll}.\footnote{We will make our attack code available on acceptance.}
+As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}, since it is faithful to the original description of HLL given in~\cite{hll}, and against the HLL Beta implementation used in the popular Redis framework~\cite{redis}.\footnote{We will make our attack code available on acceptance.}
 %at \href{https://github.com/PizzaWhisperer/HLLVuln}{\url{https://github.com/PizzaWhisperer/HLLVuln}}.} 
 %\todo{We need to make sure the repo is publicly accessible but anonymised for submission, if we are going to include it.}
-For simplicity, we take ${\mathcal{X}}$ to be the set of 4-character long ASCII strings, with $\lvert {\mathcal{X}} \rvert = 7,311,616= 2^{22.8}$.
-We stress that this choice of inputs is merely illustrative and the results do not in any way depend on choosing ASCII strings.
+For simplicity, we take ${\mathcal{X}}$ to be the set of integers ranging from 0 to 1,048,576, with $\lvert {\mathcal{X}} \rvert = 2^{20}$.
+We stress that this choice of inputs is merely illustrative, as the attacks are not sensitive at all to the form of the inputs - all that is needed is that the hash function $h$, acting on those inputs, should have approximately uniform outputs.
+So 20-bit integers are just as good as, say, IP packet headers, for experimental evaluation.
+
+\subsection{Attacking ~\cite{clahll}}
 We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}; again the attacks do not depend on the specifics of what fixed hash function is used.
@@
 
-We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $2^{22.8}$ possible inputs while trying to keep the cardinality estimate as low as possible, in each of the scenarios S1, S2 and S4.
+We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $2^{20}$ possible inputs while trying to keep the cardinality estimate as low as possible, in each of the scenarios S1, S2 and S4.
 
@@
 
-First, when attacking a sketch already containing 1000 random items, we can see that the adversary is on average able to insert about half of the $2^{22.8}$ items in our starting set in scenario S2, and about 83\% in scenario S4.
+First, when attacking a sketch already containing 1000 random items, we can see that the adversary is on average able to insert about half of the $2^{20}$ items in our starting set in scenario S2, and about 83\% in scenario S4.
 The cardinality estimate is, on average, increased slightly in S2 (from 1,000 to 1,011) and remains at the same starting value (1,005 on average) in S4.
-In both cases, we inserted more than 3 million items whilst increasing the cardinality estimate by about $0.1\%$!
+In both cases, we inserted more than half a million items whilst increasing the cardinality estimate by about $0.1\%$!
 These results align well with our theoretical analysis.
@@
     \textbf{Original Card. Est.} & 0 & 0 & 1,000 & 0 & 1,005 & 0 \\ \hline
-    \textbf{\# Items added} & 28,566 & 2,856,100 & 3,655,744 & 28,566 & 6,090,003 & 28,566 \\ \hline
+    \textbf{\# Items added} & 4,096 & 409,595 & 524,288 & 28,566 & 870,320 & 409,595 \\ \hline
     \textbf{Final Card. Est.} & 1 & 126 & 1,011 & 1 & 1,005 & 1 \\ \hline
@@
 
+\subsection{Attacking Redis implementation}
+For completeness, we attack the HLL Beta implementation used in the widespread Redis framework~\cite{redis}, which we recall is different than the original HLL algorithm given in~\cite{hll}.
+The hash function $h$ is fixed as the 64-bit MurmurHash3 function. The seed used internally is hardcoded and can be extracted by inspection of the openly-available Redis codebase.
+Furthermore, the number of buckets is constant and set to 16,384. Those informations allow to gain the knowledge of an S2 adversary.
+As a classic Redis application separates the server, owning the sketch, from the client, interacting with the sketch, we consider that the scenarios S3 or S4, while possible, are not the typical Redis use case.
+Thus, we only consider a limited adversary, and implement our attacks under the scenario S2 as described in Section~\ref{sec:redis}.
+For simplicity, we set $\epsilon$ to $\frac{1}{2}$ in the attack on an initially empty HLL sketch.
+The results are averaged over 30 iterations and reported in Table~\ref{table:tab3}.
+Similarly to the section above, our empirical results confirm our theoretical analysis.
+
+\begin{table*}[tb!]
+\small\centering
+\caption{Attack results, averaged over 30 iterations.}
+\medskip
+     \makebox[\linewidth]{
+\begin{tabular}{| m{11em} | m{4em} | m{4.2em} |}
+    \hline
+    \textbf{Scenario} &  \multicolumn{2}{c|}{S2} \\ \hline
+    \textbf{$\epsilon$} & n/a & 0.5 \\ \hline
+    \textbf{HLL initially empty?} & no & yes \\ \hline
+    \textbf{Original Card. Est.} & 70,140 & 0 \\ \hline
+    \textbf{\# Items added} & 524,455 & 524,275 \\ \hline
+    \textbf{Final Card. Est.} & 73,249 & 13,188 \\ \hline
+\end{tabular}}
+\label{table:tab3}
+\end{table*}
 %\todo[inline]{Results from Redis or GBQ if we get them.}

### 2021-05-07 redis/hyperloglog.c
Added redis attack
When replacing those in the redis codebase of 06/05/2021, implements attack.

@@ -1,0 +1,1753 @@
+/* hyperloglog.c - Redis HyperLogLog probabilistic cardinality approximation.
+ * This file implements the algorithm and the exported Redis commands.
+ *
+ * Copyright (c) 2014, Salvatore Sanfilippo <antirez at gmail dot com>
+ * All rights reserved.
+ *
+ * Redistribution and use in source and binary forms, with or without
+ * modification, are permitted provided that the following conditions are met:
+ *
+ *   * Redistributions of source code must retain the above copyright notice,
+ *     this list of conditions and the following disclaimer.
+ *   * Redistributions in binary form must reproduce the above copyright
+ *     notice, this list of conditions and the following disclaimer in the
+ *     documentation and/or other materials provided with the distribution.
+ *   * Neither the name of Redis nor the names of its contributors may be used
+ *     to endorse or promote products derived from this software without
+ *     specific prior written permission.
+ *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+ * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+ * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
+ * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
+ * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
+ * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
+ * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
+ * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
+ * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
+ * POSSIBILITY OF SUCH DAMAGE.
+ */
+
+#include "server.h"
+
+#include <stdint.h>
+#include <math.h>
+
[…]

### 2021-05-07 redis/server.h

@@ -1,0 +1,2750 @@
+/*
+ * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
+ * All rights reserved.
+ *
+ * Redistribution and use in source and binary forms, with or without
+ * modification, are permitted provided that the following conditions are met:
+ *
+ *   * Redistributions of source code must retain the above copyright notice,
+ *     this list of conditions and the following disclaimer.
+ *   * Redistributions in binary form must reproduce the above copyright
+ *     notice, this list of conditions and the following disclaimer in the
+ *     documentation and/or other materials provided with the distribution.
+ *   * Neither the name of Redis nor the names of its contributors may be used
+ *     to endorse or promote products derived from this software without
+ *     specific prior written permission.
+ *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+ * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+ * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
+ * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
+ * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
+ * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
+ * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
+ * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
+ * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
+ * POSSIBILITY OF SUCH DAMAGE.
+ */
+
+#ifndef __REDIS_H
+#define __REDIS_H
+
+#include "fmacros.h"
+#include "config.h"
+#include "solarisfixes.h"
+#include "rio.h"
+#include "atomicvar.h"
+
+#include <stdio.h>
[…]

### 2021-08-27 docs/HLL_exp_bad/main.tex
update after CCS reviews and smoother integration of Redis results

@@
 \newif\ifanon % turn on ACM style (can be short or full)
-\anontrue
+\anonfalse
 
@@
 
-\documentclass[sigconf, anonymous, dvipsnames]{acmart} %dvipsnames for old/new macros
+%\documentclass[sigconf, anonymous, dvipsnames]{acmart} %dvipsnames for old/new macros
+\documentclass[sigconf, dvipsnames]{acmart} %dvipsnames for old/new macros
 
 \fancyhf{} % Remove fancy page headers 
-\fancyhead[C]{Anonymous submission \#147 to ACM CCS 2021} % TODO: replace 9999 with your paper number
+%\fancyhead[C]{Anonymous submission \#147 to ACM CCS 2021} % TODO: replace 9999 with your paper number
 \fancyfoot[C]{\thepage}
@@
 \setcopyright{none} % No copyright notice required for submissions
-\acmConference[Anonymous Submission to ACM CCS 2021]{ACM Conference on Computer and Communications Security}{Due 6 May 2021}{Seoul, South Korea}
-\acmYear{2021}
+%\acmConference[Anonymous Submission to ACM CCS 2021]{ACM Conference on Computer and Communications Security}{Due 6 May 2021}{Seoul, South Korea}
+%\acmYear{2021}
 
@@
 \title{HyperLogLog: Exponentially Bad in Adversarial Settings}
-%\author{{\rm Kenneth G. Paterson}\\ {\rm ETH Zurich} \and {\rm Mathilde Raynal\thanks{Work done while an exchange student at ETH Zurich.}}\\ {\rm EPFL}}
-%{\tt kenny.paterson@inf.ethz.ch, mraynal@student.ethz.ch}}
+
+\author{Kenneth G. Paterson}
+%\orcid{...}
+\affiliation{ETH Zurich}
+\email{kenny.paterson@inf.ethz.ch}
+
+\author{Mathilde Raynal}
+%\orcid{...}
+\affiliation{EPFL}
+\email{mraynal@student.ethz.ch}
 
@@
 
+The S2 adversary can insert items and knows the hash function $h$ and number of buckets $m$, but is otherwise blind.
+Such an adversary is quite realistic.
+For example, both the Redis and Presto implementations of HLL use a fixed hash function and parameters that can be read off from source code.\footnote{See \url{https://github.com/redis/redis/blob/unstable/src/hyperloglog.c\#L396}, \url{
+https://github.com/redis/redis/blob/unstable/src/hyperloglog.c\#L466} for Redis and 
+\url{https://github.com/prestodb/presto/blob/2ad67dcf000be86ebc5ff7732bbb9994c8e324a8/presto-main/src/main/java/com/facebook/presto/type/khyperloglog/KHyperLogLog.java\#L20}
+for Presto.}
+
+The S3 and S4 adversaries have stronger capabilities.
+The access to cardinality estimates in S3 would typically be provided via an API, while the additional access to the bucket contents in S4 may require insider capabilities.
 At first sight it might appear that an S3 adversary can easily infer the knowledge given to our S4 adversary in its snapshot.
@@
 However, when the number of active buckets has reached some value $B^\prime \leq B_{\max}$ this means that an item $x$ may not increase the cardinality estimate, yet still not hit one of the $B$ target buckets (instead it hits one of the other $B^\prime - B$ active buckets).
-Such an $x$ acts as a ``false positive" and needs to be eliminated through reprocessing of the list ${\mathcal{L}}$.
+Such an $x$ acts as a ``false positive" and needs to be eliminated through reprocessing of the list ${\mathcal{L}}$ (as previously described).
 
@@
 
-We present a brief analysis of the HLL implementation used in Redis, since Redis is a popular distributed, in-memory key-value store implementation and the version of HLL used there differs significantly from that described in Figure~\ref{fig:hll}.
+We present an analysis of the HLL implementation used in Redis, since Redis is a popular distributed, in-memory key-value store implementation and the version of HLL used there differs significantly from that described in Figure~\ref{fig:hll}.
 
@@
 Redis 5.0 released in October 2018 replaces this with a new cardinality estimator based closely on the work of Ertl~\cite{hllnew}.
-This method is still used in the current release (Redis 6.2.2) and so is the object of our study here.\footnote{For details of Redis releases, see \url{https://github.com/redis/redis/blob/6.2/00-RELEASENOTES} and other files obtained by replacing ``6.2'' in this URL with the desired version number.}
+This method is still used in the current release (Redis 6.2.5) and so is the object of our study here.\footnote{For details of Redis releases, see \url{https://github.com/redis/redis/blob/6.2/00-RELEASENOTES} and other files obtained by replacing ``6.2'' in this URL with the desired version number.}
 
@@
 
-As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}, since it is faithful to the original description of HLL given in~\cite{hll}, and against the HLL Beta implementation used in the popular Redis framework~\cite{redis}.\footnote{We will make our attack code available on acceptance.}
+As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}, since it is faithful to the original description of HLL given in~\cite{hll}, and against the HLL implementation used in Redis~\cite{redis}.\footnote{Our attack code is available on request.}
 %at \href{https://github.com/PizzaWhisperer/HLLVuln}{\url{https://github.com/PizzaWhisperer/HLLVuln}}.} 
 %\todo{We need to make sure the repo is publicly accessible but anonymised for submission, if we are going to include it.}
-For simplicity, we take ${\mathcal{X}}$ to be the set of integers ranging from 0 to 1,048,576, with $\lvert {\mathcal{X}} \rvert = 2^{20}$.
-We stress that this choice of inputs is merely illustrative, as the attacks are not sensitive at all to the form of the inputs - all that is needed is that the hash function $h$, acting on those inputs, should have approximately uniform outputs.
+For simplicity, we take ${\mathcal{X}}$ to be the set of integers ranging from 0 to $2^{20}-1$, so $\lvert {\mathcal{X}} \rvert = 2^{20}$.
+We stress that this choice of inputs is merely illustrative, as the attacks are not sensitive at all to the form of the inputs -- all that is needed is that the hash function $h$, acting on those inputs, should have approximately uniform outputs.
 So 20-bit integers are just as good as, say, IP packet headers, for experimental evaluation.
 
-\subsection{Attacking ~\cite{clahll}}
+\subsection{Results for a ``Classic'' HLL Implementation}
+
+We first focus on a classic HLL implementation as per the original HLL paper~\cite{hll}, namely the implementation of~\cite{clahll}.
 We choose $h$ to be the 32-bit MurmurHash3 function from~\cite{murmur3code}; again the attacks do not depend on the specifics of what fixed hash function is used.
@@
 \small\centering
-\caption{Attack results, averaged over 30 iterations.}
+\caption{Attack results for a classic HLL implementation, averaged over 30 iterations.}
 \medskip
@@
 \small\centering
-\caption{Attack results, averaged over 30 iterations, in the setting of~\cite{hllvuln}.}
+\caption{Attack results for a classic HLL implementation, averaged over 30 iterations, in the setting of~\cite{hllvuln}.}
 \medskip
@@
 
-\subsection{Attacking Redis implementation}
-For completeness, we attack the HLL Beta implementation used in the widespread Redis framework~\cite{redis}, which we recall is different than the original HLL algorithm given in~\cite{hll}.
+\subsection{Results for Redis}
+We also implemented our attacks for the HLL implementation used in Redis as described in Section~\ref{sec:redis}.
 The hash function $h$ is fixed as the 64-bit MurmurHash3 function. The seed used internally is hardcoded and can be extracted by inspection of the openly-available Redis codebase.
-Furthermore, the number of buckets is constant and set to 16,384. Those informations allow to gain the knowledge of an S2 adversary.
-As a classic Redis application separates the server, owning the sketch, from the client, interacting with the sketch, we consider that the scenarios S3 or S4, while possible, are not the typical Redis use case.
-Thus, we only consider a limited adversary, and implement our attacks under the scenario S2 as described in Section~\ref{sec:redis}.
-For simplicity, we set $\epsilon$ to $\frac{1}{2}$ in the attack on an initially empty HLL sketch.
+Furthermore, the number of buckets is constant and set to 16,384. Knowledge of theese parameters places us in the setting of an S2 adversary.
+Since a classic Redis application separates the server owning the sketch from the client interacting with the sketch, we consider that the scenarios S3 or S4, while possible, are not typical Redis use case.
+Thus, we only consider scenario S2 for Redis.
+For simplicity, we set attack parameter $\epsilon$ to $\frac{1}{2}$ in the attack on an initially empty HLL sketch.
 The results are averaged over 30 iterations and reported in Table~\ref{table:tab3}.
-Similarly to the section above, our empirical results confirm our theoretical analysis.
+Similarly to the case for classic HLL, our empirical results match with our theoretical analysis.
 
@@
 \small\centering
-\caption{Attack results, averaged over 30 iterations.}
+\caption{Attack results for HLL in Redis, averaged over 30 iterations.}
 \medskip
@@
 This is somewhat different to common simulation-based security definitions, where ${\mathcal{S}}$'s input might be restricted (e.g.\ in a confidentiality definition, ${\mathcal{S}}$ might receive only $|m|$ in place of $m$) but its behaviour is not.
-Crucially, in our setting, security arises from inspecting the details of ${\mathcal{S}}$ and observing how it instantiates HLL in the average case.
+Crucially, in our setting, security arises from inspecting the details of ${\mathcal{S}}$ and observing that it instantiates HLL in the average case.
 
@@
 Our analysis may assist software developers in understanding the risks they run when using HLL in adversarial settings.
-We have also provided a simple method for securing HLL in adversarial settings: simply replace the internal hash function by a keyed PRF, and one obtains full protection even in scenario S4, our strongest attack scenario.
+We have also provided a method for securing HLL in adversarial settings: simply replace the internal hash function by a keyed PRF, and one obtains full protection even in scenario S4, our strongest attack scenario.
+%However, this solution requires the introduction of a cryptographic key, and all the attendant key management issues this brings.
+So while the solution is simple on paper, in practice, it is less so.
 
@@
 
-The authors thank Anupama Unnikrishnan for useful discussions on simulation-based definitions.
+The authors thank Anupama Unnikrishnan for useful discussions on simulation-based definitions and Tom Shrimpton for suggesting to look at GoogleBigQuery.
 \fi

### 2021-08-28 docs/HLL_exp_bad/main.tex
tweaks to UHF attack

@@
 
-By choosing $x^i$ to have 256 bits, we have $v=2$ and $p_{x^i}(K) = K^2 + x^i_1K+x^i_2$; then further selecting $x^i$ such that  $x^i_2$ is a constant value (e.g.\ $0 \in \textsf{GF}(2^{128})$) and considering consecutive pairs of $x_i$ items, we obtain equations of the form:
+By choosing $x^i$ to have 256 bits, we have $v=2$ and $p_{x^i}(K) = K^2 + x^i_1K+x^i_2$; then further selecting $x^i$ such that  $x^i_2$ is a constant value (e.g.\ $0 \in \textsf{GF}(2^{128})$) and considering consecutive pairs of $x^i$ items, we obtain equations of the form:
 \[
@@
 Here, the $n$ least significant bits of the left-hand side leak through the indices of the buckets hit for each index.
-By selecting the $x^i_1$ uniformly at random and analysing the equations of the above form that result, we obtain a non-singular linear system over $\textsf{GF}(2)$ in the unknown bits of $K$.
-The number $t$ of $x^i$ needed is roughly $2\cdot128/n$ since we get $n$ independent linear equations per pair $(x^{2j},x^{2j+1})$ and there are 128 unknown bits in $K$.
+By selecting the $x^i_1$ uniformly at random and analysing the equations of the above form that result, we obtain a linear system over $\textsf{GF}(2)$ in the unknown bits of $K$.
+The number $t$ of $x^i$ needed to recover $K$ in its entirety is roughly $2\cdot128/n$ since we get $n$ independent linear equations per pair $(x^{2j},x^{2j+1})$ and there are 128 unknown bits in $K$.
 The required value of $t$ is small for typical HLL parameters (e.g.

### 2021-08-28 docs/HLL_exp_bad/main.tex
switching to IEEE foramt, adding flags for preprint and anonymity

@@
-\newif\ifacmversion % turn on ACM style (can be short or full)
-\acmversiontrue
-
-\newif\ifanon % turn on ACM style (can be short or full)
+\newif\ifIEEEversion % turn on IEEE style 
+\IEEEversionfalse
+\newif\ifanon % needed for acks, suppression of authors, etc
 \anonfalse
 
-\ifacmversion
-
-%\documentclass[sigconf, anonymous, dvipsnames]{acmart} %dvipsnames for old/new macros
-\documentclass[sigconf, dvipsnames]{acmart} %dvipsnames for old/new macros
-
-\fancyhf{} % Remove fancy page headers 
-%\fancyhead[C]{Anonymous submission \#147 to ACM CCS 2021} % TODO: replace 9999 with your paper number
-\fancyfoot[C]{\thepage}
-
-\setcopyright{none} % No copyright notice required for submissions
-%\acmConference[Anonymous Submission to ACM CCS 2021]{ACM Conference on Computer and Communications Security}{Due 6 May 2021}{Seoul, South Korea}
-%\acmYear{2021}
-
-\settopmatter{printacmref=false, printccs=true, printfolios=true} % We want page numbers on submissions
-
-%%\ccsPaper{9999} % TODO: replace with your paper number once obtained
-
-\else
-\documentclass[%
-	dvipsnames,	%% xcolor
-	draft,
-]{article}
-\fi
-
-\usepackage{graphicx, amsmath, amsthm, amssymb, setspace, array, url, hyperref, xspace, todonotes, pifont}
+\ifIEEEversion
+%% what follows is for the IEEEtran format required for IEEE Euro S&P
+
+\documentclass[compsoc,conference,a4paper,10pt,times]{IEEEtran}
+
+%% our own choice of packages 
+\usepackage{graphicx, amsmath, amsthm, amssymb, setspace, array, url, xspace, todonotes, pifont} %hyperref is supplied by IEEEtran below
 %\usepackage{a4wide}
@@
 
-%\newtheorem{lemma}{Lemma}
-%\newtheorem{theorem}{Theorem}
+%% packages and whatnot inherited from IEEEtran example file
+
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{bmpsize}
+\usepackage{xcolor}
+\usepackage{lipsum}
+\usepackage[colorlinks=true,urlcolor=black]{hyperref}
+%% commented out this nonsense
+%\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+%   T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+%% need these for IEEEtran - not defined in cls file
+\newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
+
+\begin{document}
 
@@
 
-\author{Kenneth G. Paterson}
-%\orcid{...}
-\affiliation{ETH Zurich}
-\email{kenny.paterson@inf.ethz.ch}
-
-\author{Mathilde Raynal}
-%\orcid{...}
-\affiliation{EPFL}
-\email{mraynal@student.ethz.ch}
+\ifanon %% suppress authors in anonymous version
+\author{}
+\else
+\author{\IEEEauthorblockN{Kenneth G. Paterson}
+\IEEEauthorblockA{\textit{ETH Zurich} \\
+kenny.paterson@inf.ethz.ch}
+\and
+\IEEEauthorblockN{Mathilde Raynal}
+\IEEEauthorblockA{\textit{EPFL} \\
+mraynal@student.ethz.ch}
+}
+\fi %end anon processing of authors
+
+\else  %% what follows is a vanilla format for posting on eprint
+
+\documentclass[a4paper,10pt]{article}
+
+%% our own choice of packages 
+\usepackage{graphicx, amsmath, amsthm, amssymb, setspace, array, url, xspace, todonotes, pifont, hyperref} 
+\usepackage{a4wide}
+%\usepackage{authblk}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+\newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
 
 \begin{document}
+
+\title{HyperLogLog: Exponentially Bad in Adversarial Settings}
+\ifanon %% suppress authors in anonymous version
+\author{}
+\else
+\author{Kenneth G. Paterson\\ETH Zurich\\\texttt{kenny.paterson@inf.ethz.ch} \and Mathilde Raynal\\EPFL\\\texttt{mraynal@student.ethz.ch}}
+\fi % end anon processing of authors 
+\fi % end of vanilla format
+
+\maketitle
 
@@
 
-\maketitle
+%\begin{IEEEkeywords}
+%yadda yadda
+%\end{IEEEkeywords}
 

### 2021-08-28 docs/HLL_exp_bad/main.tex
reordering formal security  analysis section to improve flow

@@
 \newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
 
@@
 Yet, by the preceding argument, the simulator in the ideal world is able to produce an output that ${\mathcal{E}}$ cannot distinguish from that of ${\mathcal{A}}$ in the real world.
-In particular, our simulator will output whatever ${\mathcal{A}}$ outputs, so if ${\mathcal{A}}$ outputs the result of a \texttt{read} query in the ideal world, then so will ${\mathcal{S}}$.
-But in the ideal world, ${\mathcal{S}}$ uses random sampling for the $y$ values, so there ${\mathcal{S}}$ (and ${\mathcal{A}}$ that it runs) are operating in the setting where the average case analysis of HLL applies, as per~\cite{hll}.
+In particular, our simulator will output whatever ${\mathcal{A}}$ outputs, so if ${\mathcal{A}}$'s output $\textsf{out}_{\mathcal{A}}$ is the result of a \texttt{read} query in the ideal world, then so will ${\mathcal{S}}$'s.
+But in the ideal world, ${\mathcal{S}}$ uses random sampling for the $y$ values, so in that world ${\mathcal{S}}$ (and ${\mathcal{A}}$ that it runs) are operating in the setting where the average case analysis of HLL applies, as per~\cite{hll}.
 So in the ideal world this specific ${\mathcal{E}}$ would receive an accurate average case cardinality estimate.
@@
 It then follows that ${\mathcal{A}}$ has no advantage when it comes to manipulating the HLL-$F$ cardinality estimate.
-So HLL, when instantiated with PRF $F$, is secure against cardinality manipulation attacks in S4 (and therefore in the weaker attacks models S2 and S3 as well).
+So HLL, when instantiated with PRF $F$, is secure against cardinality manipulation attacks.
+% in S4 (and therefore in the weaker attacks models S2 and S3 as well).
 
@@
 %Using a standard transformation, this probability can be rewritten as $\lvert \Pr[b' = 1| b = 1] - \Pr[b'=0|b=1] \rvert$.
-
-We say that the scheme HLL-$F$ is S$i$-simulation-secure with parameters $(q,t_{\mathcal{S}},t_{\mathcal{A}},t_{\mathcal{E}},\epsilon)$ if there exists a simulator $\mathcal{S}$ running in time at most $t_{\mathcal{S}}$ such that, for all adversaries ${\mathcal{A}}$ running in time at most $t_{\mathcal{A}}$ and making at most $q$ \texttt{insert($\cdot$)} queries and all distinguishers ${\mathcal{E}}$ running in time at most $t_{\mathcal{E}}$, we have
-$\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{E}}) \leq \epsilon$.
-We quantify over \texttt{insert($\cdot$)} queries (via $q$) and not other types of query because only \texttt{insert($\cdot$)} queries will appear in our concrete security analysis and we do not want to clutter the notation.
 
@@
 
-We next describe the simulator ${\mathcal{S}}$ for the ideal world. Simulator ${\mathcal{S}}$ operates by running ${\mathcal{A}}$ in a blackbox manner and handling all of its queries as follows: 
+\begin{definition}
+We say that the scheme HLL-$F$ is S$i$-simulation-secure with parameters $(q,t_{\mathcal{S}},t_{\mathcal{A}},t_{\mathcal{E}},\epsilon)$ if there exists a simulator $\mathcal{S}$ running in time at most $t_{\mathcal{S}}$ such that, for all adversaries ${\mathcal{A}}$ running in time at most $t_{\mathcal{A}}$ and making at most $q$ \texttt{insert($\cdot$)} queries and all distinguishers ${\mathcal{E}}$ running in time at most $t_{\mathcal{E}}$, we have
+$\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{E}}) \leq \epsilon$. 
+\end{definition}
+
+In the above definition, we quantify over \texttt{insert($\cdot$)} queries (via $q$) and not other types of query because only \texttt{insert($\cdot$)} queries will appear in our concrete security analysis and we do not want to clutter the notation.
+
+Note that our security analysis using a simulation-based definition does not require us to specify a winning condition for adversary $\mathcal{A}$.
+Rather, we infer from inspecting the simulator what the adversary can (or cannot) achieve in the ideal world, and show that no efficient algorithm $\mathcal{E}$ can distinguish the real and ideal worlds.
+This is in contrast to the game-based approach of~\cite{CCS:ClaPatShrPS19} which needs a complex definition to say when the adversary has successfully manipulated a probabilistic data structure.
+This reveals a strength of the simulation-based approach in more complex settings like ours.
+
+We next describe the simulator ${\mathcal{S}}$ that we will use in our further analysis: ${\mathcal{S}}$ operates by running ${\mathcal{A}}$ in a blackbox manner and handling all of its queries as follows:
 \begin{itemize}
@@
 
+For our definition to meaningfully capture security against cardinality manipulation, it is important to restrict ${\mathcal{S}}$ to doing random sampling when handling \texttt{insert($\cdot$)} queries.
+Otherwise, our definition could be vacuously satisfied (for example ${\mathcal{S}}$ could just sample the PRF key $K$ for itself and provide a perfect real-world view).
+This is somewhat different to common simulation-based security definitions, where ${\mathcal{S}}$'s input might be restricted (e.g.\ in a confidentiality definition, ${\mathcal{S}}$ might receive only $|m|$ in place of $m$) but its behaviour is not.
+Crucially, in our setting, security arises from inspecting the details of ${\mathcal{S}}$ and observing that it instantiates HLL in the average case.
+
+We require one further standard definition.
+\begin{definition}
 We define the PRF advantage of an adversary ${{\mathcal{B}}}$ against $F$ outputting a bit $d'$ as being its advantage in distinguishing whether it is interacting with $F$ for a uniformly random key $K$ ($d=0$) or interacting with a truly random function ($d=1$).
@@
 \]
+\end{definition}
 Here, we assume without loss of generality that all of ${{\mathcal{B}}}$'s queries are distinct. See~\cite[Section 4.4.1]{BonehShoup} for details of the PRF security definition.
@@
 The only changes to the proof are to use the Redis approach for computing responses to \texttt{read} queries in place of the classical HLL approach.
-
-Note that our security analysis using a simulation-based definition does not require us to specify a winning condition for the adversary.
-Rather, we infer from inspecting the simulator what the adversary can (or cannot) achieve in the ideal world, and show that no efficient adversary can distinguish the real and ideal worlds.
-This is in contrast to the game-based approach of~\cite{CCS:ClaPatShrPS19} which needs a complex definition to say when the adversary has successfully manipulated a probabilistic data structure.
-This reveals a strength of the simulation-based approach in more complex settings like ours.
-
-Note also that, for our definition to meaningfully capture security against cardinality manipulation, it is important to restrict  ${\mathcal{S}}$ to doing random sampling when handling \texttt{insert($\cdot$)} queries.
-Otherwise, our definition could be vacuously satisfied (for example ${\mathcal{S}}$ could just sample the PRF key $K$ for itself and provide a perfect real-world view).
-This is somewhat different to common simulation-based security definitions, where ${\mathcal{S}}$'s input might be restricted (e.g.\ in a confidentiality definition, ${\mathcal{S}}$ might receive only $|m|$ in place of $m$) but its behaviour is not.
-Crucially, in our setting, security arises from inspecting the details of ${\mathcal{S}}$ and observing that it instantiates HLL in the average case.
 

### 2021-08-28 docs/HLL_exp_bad/main.tex
more work on formal analysis to improve presentation

@@
 \newif\ifIEEEversion % turn on IEEE style 
-\IEEEversionfalse
+\IEEEversiontrue
 \newif\ifanon % needed for acks, suppression of authors, etc
-\anonfalse
+\anontrue
 
@@
 
-\subsection{Intuition for Formal Security Analysis}
+\subsection{Intuition for Formal Security Analysis}\label{sec:informal}
 
@@
 
+\vspace{0.1cm}
+\noindent\textbf{A specific simulator ${\mathcal{S}}$.}
 We next describe the simulator ${\mathcal{S}}$ that we will use in our further analysis: ${\mathcal{S}}$ operates by running ${\mathcal{A}}$ in a blackbox manner and handling all of its queries as follows:
@@
 \begin{definition}
+Let $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
 We define the PRF advantage of an adversary ${{\mathcal{B}}}$ against $F$ outputting a bit $d'$ as being its advantage in distinguishing whether it is interacting with $F$ for a uniformly random key $K$ ($d=0$) or interacting with a truly random function ($d=1$).
-Thus:
+We write:
 \[
@@
 \end{theorem}
+
+This result tightly relates the security of HLL-$F$ against cardinality manipulation to that of the PRF security of the function $F$ used in its construction (for the reasons explained in Section~\ref{sec:informal}).
 

### 2021-09-07 docs/HLL_exp_bad/main.tex
update email address
updating email address and fixing a typo or two

@@
 \newif\ifIEEEversion % turn on IEEE style 
-\IEEEversiontrue
+\IEEEversionfalse
 \newif\ifanon % needed for acks, suppression of authors, etc
-\anontrue
+\anonfalse
 
@@
 \IEEEauthorblockA{\textit{EPFL} \\
-mraynal@student.ethz.ch}
+mathilde.raynal@epfl.ch}
 }
@@
 \else
-\author{Kenneth G. Paterson\\ETH Zurich\\\texttt{kenny.paterson@inf.ethz.ch} \and Mathilde Raynal\\EPFL\\\texttt{mraynal@student.ethz.ch}}
+\author{Kenneth G. Paterson\\ETH Zurich\\\texttt{kenny.paterson@inf.ethz.ch} \and Mathilde Raynal\\EPFL\\\texttt{mathilde.raynal@epfl.ch}}
 \fi % end anon processing of authors 
@@
 The hash function $h$ is fixed as the 64-bit MurmurHash3 function. The seed used internally is hardcoded and can be extracted by inspection of the openly-available Redis codebase.
-Furthermore, the number of buckets is constant and set to 16,384. Knowledge of theese parameters places us in the setting of an S2 adversary.
-Since a classic Redis application separates the server owning the sketch from the client interacting with the sketch, we consider that the scenarios S3 or S4, while possible, are not typical Redis use case.
+Furthermore, the number of buckets is constant and set to 16,384. Knowledge of these parameters places us in the setting of an S2 adversary.
+Since a classic Redis application separates the server holding the sketch from the client interacting with the sketch, we consider that attack scenarios S3 or S4, while possible, are not typical for Redis.
 Thus, we only consider scenario S2 for Redis.

### 2021-09-07 docs/HLL_exp_bad/main.tex
rejigging security analysis once again

@@
 \begin{flushleft}
-\textit{Let $h : \mathcal{D} \rightarrow \{0, 1\}^{32}$ hash data from some domain $\mathcal{D}$ to binary $32$-bit strings.\\
+\textit{Let $h : \mathcal{Y} \rightarrow \{0, 1\}^{32}$ hash data from some domain $\mathcal{Y}$ to binary $32$-bit strings.\\
 For $s$ of bit-length $\ell$, let $\rho(s)$ denote the position of the leftmost $1$-bit of $s$: e.g., $\rho(1...) = 1, \rho(0001...) = 4, \rho(0^\ell) = \ell + 1.$\\}
@@
 \vspace{0.5em}
-\textbf{Algorithm} HyperLogLog (\textbf{input} ${\mathcal{X}}$ : multiset of items from domain $\mathcal{D}$).\\
+\textbf{Algorithm} HyperLogLog (\textbf{input} ${\mathcal{X}}$ : multiset of items from domain $\mathcal{Y}$).\\
 \textbf{initialize} a collection of $m$ registers, $M[1], ..., M[m]$, to $0$;\\
@@
 This algorithm is able to estimate cardinalities greater than $10^9$ with a typical standard error of 4\%, using only 1.5 kB of memory.
-Crucially, the proof of this~\cite[Theorem 1]{hll} makes the assumption that the inputs are drawn independently and uniformly at random from the input domain -- loosely, it is an average case analysis.\footnote{Technically, the results of~\cite{hll} are proven for a distribution obtained by making arbitrary replications and permutations after sampling independently and uniformly, to simulate input taken from an \emph{ideal multiset}.
+Crucially, the proof of this~\cite[Theorem 1]{hll} makes the assumption that the inputs are drawn independently and uniformly at random from the input domain -- loosely put, it is an average case analysis.\footnote{Technically, the results of~\cite{hll} are proven for a distribution obtained by making arbitrary replications and permutations after sampling independently and uniformly, to simulate input taken from an \emph{ideal multiset}.
 But they also hold without replication and permutation because HLL's internal state is invariant under repeated inputs.}
@@
 We first begin by explaining the high-level idea of our approach.
-We allow an adversary ${\mathcal{A}}$ to interact with an HLL via various oracles that capture its capabilities in each of the three scenarios S2, S3 and S4: it may have an \texttt{insert($\cdot$)} oracle allowing it to add items to the HLL sketch, a \texttt{read} oracle that returns the HLL cardinality estimate, and a \texttt{corrupt} that returns all the HLL bucket contents (but not any secret key $K$); in S4 it has all three oracles.
+We allow an adversary ${\mathcal{A}}$ to interact with an HLL sketch via various oracles that capture its capabilities in each of the three scenarios S2, S3 and S4: it may have an \texttt{insert($\cdot$)} oracle allowing it to add items to the HLL sketch, a \texttt{read} oracle that returns the HLL cardinality estimate, and a \texttt{corrupt} that returns all the HLL sketch's bucket contents (but not any secret key $K$); in S4 it has all three oracles.
 In addition, in all 3 scenarios, ${\mathcal{A}}$ has access to an \texttt{init($\cdot$)} oracle by which it sets the HLL parameters $n$, $m=2^n$, $\ell$, $\alpha_m$ used to initialise the HLL sketch.
@@
 In the real world ($b=0$), ${\mathcal{A}}$ interacts with the scheme HLL-$F$.
-This is the same as HLL operating as in Figure~\ref{fig:hll} but with the hash function $h$ replaced by a keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$ whose key $K$ is sampled uniformly at random at the beginning of the game and not available to the adversary; ${\mathcal{A}}$ has access to the relevant selection of oracles and completes its execution with some output $\textsf{out}_{\mathcal{A}}$ (a bit-string).
+This is the same as HLL operating as in Figure~\ref{fig:hll} but where the hash function $h$ is replaced by a keyed function $F: \{0,1\}^k \times {{\mathcal{Y}}} \rightarrow  \{0,1\}^{n+\ell}$ whose key $K$ is sampled uniformly at random at the beginning of the game and not available to the adversary; ${\mathcal{A}}$ has access to the relevant selection of oracles and completes its execution with some output $\textsf{out}_{\mathcal{A}}$ (a bit-string).
 
@@
 
-We will prove that  ${\mathcal{S}}$ is such that, given any adversary ${\mathcal{A}}$, then any distinguisher $\mathcal{E}$ that can distinguish between the $\textsf{out}_{\mathcal{A}}$ in the real world and $\textsf{out}_{\mathcal{S}}$ in the ideal world can be used to build an adversary ${{\mathcal{B}}}$ which breaks the PRF security of $F$.
-Under the assumption that $F$ is a good PRF, no such efficient distinguisher $\mathcal{E}$ can exist, and we may conclude that (up to PRF security) no efficient distinguisher can tell that it is interacting with the ideal world instead of the real world.
+We will prove that  ${\mathcal{S}}$ is such that, given any adversary ${\mathcal{A}}$, then any distinguisher $\mathcal{D}$ that can distinguish between the $\textsf{out}_{\mathcal{A}}$ in the real world and $\textsf{out}_{\mathcal{S}}$ in the ideal world can be used to build an adversary ${{\mathcal{B}}}$ which breaks the PRF security of $F$.
+Under the assumption that $F$ is a good PRF, no such efficient distinguisher $\mathcal{D}$ can exist, and we may conclude that (up to PRF security) no efficient distinguisher can tell that it is interacting with the ideal world instead of the real world.
 
 Why does this proof imply that ${\mathcal{A}}$ cannot manipulate the cardinality estimate?
-To see this, note that (amongst many other possibilities) ${\mathcal{A}}$'s output could be the cardinality estimate obtained from a \texttt{read} query, allowing a specific ${\mathcal{E}}$ to try to use that to distinguish between the two worlds.
-Yet, by the preceding argument, the simulator in the ideal world is able to produce an output that ${\mathcal{E}}$ cannot distinguish from that of ${\mathcal{A}}$ in the real world.
+To see this, note that (amongst many other possibilities) ${\mathcal{A}}$'s output could be the cardinality estimate obtained from a \texttt{read} query, allowing a specific ${\mathcal{D}}$ to try to use that to distinguish between the two worlds.
+Yet, by the preceding argument, the simulator in the ideal world is able to produce an output that ${\mathcal{D}}$ cannot distinguish from that of ${\mathcal{A}}$ in the real world.
 In particular, our simulator will output whatever ${\mathcal{A}}$ outputs, so if ${\mathcal{A}}$'s output $\textsf{out}_{\mathcal{A}}$ is the result of a \texttt{read} query in the ideal world, then so will ${\mathcal{S}}$'s.
 But in the ideal world, ${\mathcal{S}}$ uses random sampling for the $y$ values, so in that world ${\mathcal{S}}$ (and ${\mathcal{A}}$ that it runs) are operating in the setting where the average case analysis of HLL applies, as per~\cite{hll}.
-So in the ideal world this specific ${\mathcal{E}}$ would receive an accurate average case cardinality estimate.
-Since ${\mathcal{E}}$ tries to distinguish the two worlds from the cardinality estimate and we already showed that no efficient distinguisher ${{\mathcal{E}}}$ can exist, it must follow that the cardinality estimates produced in the real and ideal worlds are indistinguishable (up to PRF security).
+So in the ideal world this specific ${\mathcal{D}}$ would receive an accurate average case cardinality estimate.
+Since ${\mathcal{D}}$ tries to distinguish the two worlds from the cardinality estimate and we already showed that no efficient distinguisher ${{\mathcal{D}}}$ can exist, it must follow that the cardinality estimates produced in the real and ideal worlds are indistinguishable (up to PRF security).
 It then follows that ${\mathcal{A}}$ has no advantage when it comes to manipulating the HLL-$F$ cardinality estimate.
@@
 We now formalise the above intuition.
-We consider distinguisher ${\mathcal{E}}$ to be an algorithm running in a security game.
-The game begins with the selection of a bit $b$ uniformly at random and then runs either ${\mathcal{A}}$ (the real world, $b=0$) or ${\mathcal{S}}$ (the ideal world, $b=1$).
-Adversary ${\mathcal{A}}$ interacts with up to four oracles which are handled by the game using the algorithms of HLL-$F$; different combinations of oracles define scenarios S2, S3, S4.
-After making some queries, ${\mathcal{A}}$ returns its output $\textsf{out}_{\mathcal{A}}$ to ${\mathcal{E}}$.
+Our definitions are specific to HLL-$F$ as the object of analysis, but are easily generalised to any cardinality estimator with the same syntax as HLL-$F$ (formally defined by a triple of algorithms (\texttt{Init}, \texttt{Insert}, \texttt{Read})).
+
+\vspace{0.2cm}
+\noindent\textbf{The distinguisher ${\mathcal{D}}$ and its advantage.} We consider distinguisher ${\mathcal{D}}$ to be an algorithm running in a security game.
+The game begins with the selection of a bit $b$ uniformly at random and then runs either adversary ${\mathcal{A}}$ (the real world, $b=0$) or simulator ${\mathcal{S}}$ (the ideal world, $b=1$).
+Adversary ${\mathcal{A}}$ interacts with up to four oracles which are handled by the game using the algorithms of HLL-$F$.
+These are defined below; different combinations of oracles define scenarios S2, S3, S4.
+After making some queries, ${\mathcal{A}}$ returns its output $\textsf{out}_{\mathcal{A}}$ to ${\mathcal{D}}$.
 Simulator ${\mathcal{S}}$ runs ${\mathcal{A}}$ in a blackbox manner (as specified below) and receives ${\mathcal{A}}$'s output.
-It then provides its own output $\textsf{out}_{\mathcal{S}}$ to ${\mathcal{E}}$.
-Finally, ${\mathcal{E}}$ outputs a bit $b'$ estimating whether it is interacting with the real world ($b'=0$) or the ideal world ($b'=1$).
-For $i=2, 3, 4$, we define ${\mathcal{E}}$'s S$i$-Real-or-Ideal advantage against HLL-$F$ with adversary ${\mathcal{A}}$ and simulator ${\mathcal{S}}$ to be:
-\[
-\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{E}}) := \lvert 2\Pr[b' = b] - 1 \rvert \, .
+It then provides its own output $\textsf{out}_{\mathcal{S}}$ to ${\mathcal{D}}$.
+Finally, ${\mathcal{D}}$ outputs a bit $b'$ estimating whether it is interacting with the real world ($b'=0$) or the ideal world ($b'=1$).
+For $i=2, 3, 4$, we define ${\mathcal{D}}$'s S$i$-Real-or-Ideal advantage against HLL-$F$ with adversary ${\mathcal{A}}$ and simulator ${\mathcal{S}}$ to be:
+\[
+\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$,${\mathcal{A}}$,${\mathcal{S}}$}}}({\mathcal{D}}) := \lvert 2\Pr[b' = b] - 1 \rvert \, .
 \]
@@
 
+\vspace{0.2cm}
+\noindent\textbf{The adversary ${\mathcal{A}}$ and its oracles.}
 The four oracles that  ${\mathcal{A}}$ can interact with behave as follows in the real world:
 \begin{itemize}
-\item \texttt{init($\cdot$)}: this oracle receives as input the HLL-$F$ parameters ($n$, $\ell$, $\alpha_m$) and performs initialisation of the HLL-$F$ sketch, including generating the array of $m=2^n$ buckets $M[j]$ and initialising all the entries to $0$, and selecting a key $K \leftarrow_{\$} \{0,1\}^k$ for keyed function $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
+\item \texttt{init($\cdot$)}: this oracle receives as input the HLL-$F$ parameters ($n$, $\ell$, $\alpha_m$) and performs initialisation of the HLL-$F$ sketch, including generating the array of $m=2^n$ buckets $M[j]$ and initialising all the entries to $0$, and selecting a key $K \leftarrow_{\$} \{0,1\}^k$ for keyed function $F: \{0,1\}^k \times {{\mathcal{Y}}} \rightarrow  \{0,1\}^{n+\ell}$.
 This oracle has no output.
@@
 We assume it is only called once as ${\mathcal{A}}$'s first oracle query.
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$, the oracle implements HLL-$F$ by setting $y = F_K(x)$ and following the procedure in Figure~\ref{fig:hll} to insert $x$ into the HLL sketch (that is, it uses the first $n$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
+\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{Y}}}$, the oracle implements HLL-$F$ by setting $y = F_K(x)$ and following the procedure in Figure~\ref{fig:hll} to insert $x$ into the HLL sketch (that is, it uses the first $n$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
 This oracle has no output.
@@
 
-\begin{definition}
-We say that the scheme HLL-$F$ is S$i$-simulation-secure with parameters $(q,t_{\mathcal{S}},t_{\mathcal{A}},t_{\mathcal{E}},\epsilon)$ if there exists a simulator $\mathcal{S}$ running in time at most $t_{\mathcal{S}}$ such that, for all adversaries ${\mathcal{A}}$ running in time at most $t_{\mathcal{A}}$ and making at most $q$ \texttt{insert($\cdot$)} queries and all distinguishers ${\mathcal{E}}$ running in time at most $t_{\mathcal{E}}$, we have
-$\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{E}}) \leq \epsilon$. 
-\end{definition}
-
-In the above definition, we quantify over \texttt{insert($\cdot$)} queries (via $q$) and not other types of query because only \texttt{insert($\cdot$)} queries will appear in our concrete security analysis and we do not want to clutter the notation.
-
-Note that our security analysis using a simulation-based definition does not require us to specify a winning condition for adversary $\mathcal{A}$.
-Rather, we infer from inspecting the simulator what the adversary can (or cannot) achieve in the ideal world, and show that no efficient algorithm $\mathcal{E}$ can distinguish the real and ideal worlds.
-This is in contrast to the game-based approach of~\cite{CCS:ClaPatShrPS19} which needs a complex definition to say when the adversary has successfully manipulated a probabilistic data structure.
-This reveals a strength of the simulation-based approach in more complex settings like ours.
-
-\vspace{0.1cm}
-\noindent\textbf{A specific simulator ${\mathcal{S}}$.}
-We next describe the simulator ${\mathcal{S}}$ that we will use in our further analysis: ${\mathcal{S}}$ operates by running ${\mathcal{A}}$ in a blackbox manner and handling all of its queries as follows:
+\vspace{0.2cm}
+\noindent\textbf{A specific simulator ${\mathcal{S}}$.} In order to obtain a meaningful security definition that prevents cardinality manipulation, we need to restrict ${\mathcal{S}}$ to doing random sampling when handling \texttt{insert($\cdot$)} queries coming from ${\mathcal{A}}$.
+Otherwise, our definition could be vacuously satisfied (for example ${\mathcal{S}}$ could just sample the PRF key $K$ for itself and provide a perfect real-world view).
+This is somewhat different to typical simulation-based security definitions, where ${\mathcal{S}}$'s behaviour is not restricted but its input might be (e.g.\ in a confidentiality definition, ${\mathcal{S}}$ might receive only $|m|$ in place of plaintext $m$).
+Crucially, in our setting, security arises from inspecting the details of ${\mathcal{S}}$ and observing that, due to its use of random sampling, it instantiates HLL in the average case.
+
+In full detail, our specific ${\mathcal{S}}$ operates by running ${\mathcal{A}}$ in a blackbox manner and handling all of its queries as follows:
 \begin{itemize}
 \item \texttt{init($\cdot$)}: on input $n$, $\ell$, $\alpha_m$, ${\mathcal{S}}$ initialises an HLL sketch as in Figure~\ref{fig:hll}; that is, it creates an array of $m=2^n$ buckets $M[j]$ and initialises them all to 0.
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$,  ${\mathcal{S}}$ ignores $x$, sets $y \leftarrow_{\$} \{0,1\}^{n+\ell}$, and proceeds to update the buckets using this value of $y$ as in the original HLL algorithm in Figure~\ref{fig:hll}.
+\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{Y}}}$,  ${\mathcal{S}}$ ignores $x$, sets $y \leftarrow_{\$} \{0,1\}^{n+\ell}$, and proceeds to update the buckets using this value of $y$ as in the original HLL algorithm in Figure~\ref{fig:hll}.
 \item \texttt{read}:  ${\mathcal{S}}$ returns the HLL cardinality estimate computed as per Figure~\ref{fig:hll}.
@@
 \end{itemize}
-At the end of ${\mathcal{A}}$'s execution, ${\mathcal{S}}$ takes ${\mathcal{A}}$'s output and forwards it to ${\mathcal{E}}$ as its own output.
-Notice that the running time of  ${\mathcal{S}}$ is (roughly) equal to that of ${\mathcal{A}}$.
-
-For our definition to meaningfully capture security against cardinality manipulation, it is important to restrict ${\mathcal{S}}$ to doing random sampling when handling \texttt{insert($\cdot$)} queries.
-Otherwise, our definition could be vacuously satisfied (for example ${\mathcal{S}}$ could just sample the PRF key $K$ for itself and provide a perfect real-world view).
-This is somewhat different to common simulation-based security definitions, where ${\mathcal{S}}$'s input might be restricted (e.g.\ in a confidentiality definition, ${\mathcal{S}}$ might receive only $|m|$ in place of $m$) but its behaviour is not.
-Crucially, in our setting, security arises from inspecting the details of ${\mathcal{S}}$ and observing that it instantiates HLL in the average case.
-
-We require one further standard definition.
+At the end of ${\mathcal{A}}$'s execution, ${\mathcal{S}}$ takes ${\mathcal{A}}$'s output $\textsf{out}_{\mathcal{A}}$ and forwards it to ${\mathcal{D}}$ as its own output $\textsf{out}_{\mathcal{S}}$.
+Notice that the running time $t_{\mathcal{S}}$ of ${\mathcal{S}}$ is (roughly) equal to that of ${\mathcal{A}}$.
+
+\vspace{0.2cm}
+\noindent\textbf{Formal security definition and main result.} Now we are ready to give:
 \begin{definition}
-Let $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
+The scheme HLL-$F$ is S$i$-simulation-secure with parameters $(q,t_{\mathcal{S}},t_{\mathcal{A}},t_{\mathcal{D}},\epsilon)$ if, for the specific simulator $\mathcal{S}$ given above, for all adversaries ${\mathcal{A}}$ running in time at most $t_{\mathcal{A}}$ and making at most $q$ \texttt{insert($\cdot$)} queries, and for all distinguishers ${\mathcal{D}}$ running in time at most $t_{\mathcal{D}}$, we have
+$\mbox{Adv}^{\mbox{\tiny S$i$-RoI}}_{\mbox{\tiny{HLL-$F$,${\mathcal{A}}$,${\mathcal{S}}$}}}({\mathcal{D}}) \leq \epsilon$. 
+\end{definition}
+
+In the above definition, we quantify over \texttt{insert($\cdot$)} queries (via $q$) and not other types of query because only \texttt{insert($\cdot$)} queries will appear in our concrete security analysis and we do not want to clutter the notation.
+
+Note that our security analysis using a simulation-based definition does not require us to specify a winning condition for adversary $\mathcal{A}$.
+Rather, we infer from inspecting the simulator what the adversary can (or cannot) achieve in the ideal world, and show that no efficient algorithm $\mathcal{D}$ can distinguish the real and ideal worlds.
+This is in contrast to the game-based approach of~\cite{CCS:ClaPatShrPS19} which needs a complex definition to say when the adversary has successfully manipulated a probabilistic data structure.
+This reveals a strength of the simulation-based approach.
+
+Note also that our definition is given with respect to a \emph{specific} simulator, but could be generalised to allow any simulator that performs random sampling when handling \texttt{insert($\cdot$)} queries.
+
+Before stating our main result, we require one further standard definition:
+\begin{definition}
+Let $F: \{0,1\}^k \times {{\mathcal{Y}}} \rightarrow  \{0,1\}^{n+\ell}$.
 We define the PRF advantage of an adversary ${{\mathcal{B}}}$ against $F$ outputting a bit $d'$ as being its advantage in distinguishing whether it is interacting with $F$ for a uniformly random key $K$ ($d=0$) or interacting with a truly random function ($d=1$).
@@
 Let simulator $\mathcal{S}$ be as described above.
-For any S4 adversary ${\mathcal{A}}$ against HLL-$F$ and any distinguisher $\mathcal{E}$, we can construct a PRF adversary ${{\mathcal{B}}}$ against $F$ such that:
+For any S4 adversary ${\mathcal{A}}$ against HLL-$F$ and any distinguisher $\mathcal{D}$, we can construct a PRF adversary ${{\mathcal{B}}}$ against $F$ such that:
 \[
  \mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}}) = 
- \mbox{Adv}^{\mbox{\tiny S$4$-RoI}}_{\mbox{\tiny{HLL-$F$, ${\mathcal{A}}$, ${\mathcal{S}}$}}}({\mathcal{E}}) \,.
+ \mbox{Adv}^{\mbox{\tiny S$4$-RoI}}_{\mbox{\tiny{HLL-$F$,${\mathcal{A}}$,${\mathcal{S}}$}}}({\mathcal{D}}) \,.
 \]
 Moreover, if ${\mathcal{A}}$ makes $q$ queries to its \texttt{insert($\cdot$)} oracle, then ${{\mathcal{B}}}$ makes $q$ queries to its real or random oracle.
-Finally, the running time of ${{\mathcal{B}}}$ is (roughly) $t_{\mathcal{A}}+t_{\mathcal{E}}$.
+Finally, the running time of ${{\mathcal{B}}}$ is (roughly) $t_{\mathcal{A}}+t_{\mathcal{D}}$.
 \end{theorem}
@@
 %Informally, if $F$ is a good PRF, then $\mbox{Adv}^{\mbox{\tiny PRF}}_{F}({{\mathcal{B}}})$ should be small for any adversary ${{\mathcal{B}}}$ running in a reasonable amount of time; hence the advantage of any S4 adversary  ${\mathcal{A}}$ against HLL using PRF $F$ is also small.
-Note that our proof still applies for the improved algorithm of~\cite{hllnew} that is used in Redis.
-The only changes to the proof are to use the Redis approach for computing responses to \texttt{read} queries in place of the classical HLL approach.
+%Note that our proof would still apply for the algorithm of~\cite{hllnew} that is used in Redis if it used a PRF in place of a hash function.
+The only change needed is to use the Redis approach for computing responses to \texttt{read} queries in place of the classical HLL approach.
 
@@
 
-An HLL can leak information about items $x$ that were previously inserted.
+An HLL sketch can leak information about items $x$ that were previously inserted.
 Consider the following attack from~\cite{cardestprivacy}: to test whether an item $x$ was previously inserted into the HLL by an honest user (say), the adversary makes a sequence of queries: \texttt{read}, \texttt{insert($x$)}, \texttt{read}.
@@
 However, it is easily modified to provide a confidentiality notion.
-For example, we can consider the insertion of honestly generated items $x$ that are drawn from some probability distribution $p$ on $\mathcal{D}$ in a first phase, with ${\mathcal{A}}$ then having access to only \texttt{read} and \texttt{corrupt} oracles in a second phase (we have to remove the \texttt{insert($\cdot$)} oracle in view of the general attack from~\cite{cardestprivacy}).
+For example, we can consider the insertion of honestly generated items $x$ that are drawn from some probability distribution $p$ on $\mathcal{Y}$ in a first phase, with ${\mathcal{A}}$ then having access to only \texttt{read} and \texttt{corrupt} oracles in a second phase (we have to remove the \texttt{insert($\cdot$)} oracle in view of the general attack from~\cite{cardestprivacy}).
 
@@
 
-Our HLL construction requires a PRF $F: \{0,1\}^k \times {{\mathcal{D}}} \rightarrow  \{0,1\}^{n+\ell}$.
-Here, the set ${{\mathcal{D}}}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
+Our HLL construction requires a PRF $F: \{0,1\}^k \times {{\mathcal{Y}}} \rightarrow  \{0,1\}^{n+\ell}$.
+Here, the set ${{\mathcal{Y}}}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
 This means that $F$ needs to be Variable Input Length PRF (VIL-PRF).
@@
 
-%A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {{\mathcal{D}}} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
+%A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {{\mathcal{Y}}} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
 The result is a PRF with the required input-output signature~\cite[Section 8.2]{BonehShoup}.
@@
 PRF adversary ${{\mathcal{B}}}$ against $F$ is built as follows.
-Let $d$ denote the hidden bit determining whether ${{\mathcal{B}}}$ receives real responses of the form $F_K(x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${{\mathcal{B}}}$ provides a simulation of either the real or ideal world for distinguisher ${\mathcal{E}}$ as follows.
+Let $d$ denote the hidden bit determining whether ${{\mathcal{B}}}$ receives real responses of the form $F_K(x)$ where $K$ is a uniformly random key ($d=0$) or responses from a truly random function ($d=1$); ${{\mathcal{B}}}$ provides a simulation of either the real or ideal world for distinguisher ${\mathcal{D}}$ as follows.
 ${{\mathcal{B}}}$ runs ${{\mathcal{A}}}$ and answers each of ${{\mathcal{A}}}$'s queries as follows:
@@
 \item \texttt{init($\cdot$)}: ${{\mathcal{B}}}$ receives parameters $n$, $\ell$, $\alpha_m$, sets $m=2^n$ and initialises an array $M[j]$ of size $m$ in which each of the entries is set to 0. 
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{D}}}$,  ${{\mathcal{B}}}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${{\mathcal{B}}}$ uses $y$ to update the HLL sketch as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
+\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{Y}}}$,  ${{\mathcal{B}}}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${{\mathcal{B}}}$ uses $y$ to update the HLL sketch as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
 That is, it uses the first $n$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$.
@@
 \end{itemize}
-When ${\mathcal{A}}$ returns its final output $\textsf{out}_{\mathcal{A}}$, ${{\mathcal{B}}}$ forwards this output to ${{\mathcal{E}}}$.
-Finally, ${{\mathcal{E}}}$ outputs a bit $b'$.
+When ${\mathcal{A}}$ returns its final output $\textsf{out}_{\mathcal{A}}$, ${{\mathcal{B}}}$ forwards this output to ${{\mathcal{D}}}$.
+Finally, ${{\mathcal{D}}}$ outputs a bit $b'$.
 Then ${{\mathcal{B}}}$ outputs its own bit $d'=b'$.
@@
 The remainder of the proof follows on counting queries and noting that the HLL sub-algorithms for insertion and computing cardinality are efficient, so ${{\mathcal{B}}}$'s running time due to handling ${\mathcal{A}}$'s queries is not increased significantly compared to that of ${\mathcal{A}}$.
-The final running time for ${\mathcal{B}}$, namely $t_{\mathcal{A}}+t_{\mathcal{E}}$, follows on noting that ${\mathcal{B}}$ first runs ${\mathcal{A}}$ and then ${\mathcal{E}}$.
+The final running time for ${\mathcal{B}}$, namely $t_{\mathcal{A}}+t_{\mathcal{D}}$, follows on noting that ${\mathcal{B}}$ first runs ${\mathcal{A}}$ and then ${\mathcal{D}}$.
 

### 2021-09-07 docs/HLL_exp_bad/main.tex
sorting out domain with a hat on a D and a distinguisher D

@@
 \begin{flushleft}
-\textit{Let $h : \mathcal{Y} \rightarrow \{0, 1\}^{32}$ hash data from some domain $\mathcal{Y}$ to binary $32$-bit strings.\\
+\textit{Let $h : \hat{\mathcal{D}} \rightarrow \{0, 1\}^{32}$ hash data from some domain $\hat{\mathcal{D}}$ to binary $32$-bit strings.\\
 For $s$ of bit-length $\ell$, let $\rho(s)$ denote the position of the leftmost $1$-bit of $s$: e.g., $\rho(1...) = 1, \rho(0001...) = 4, \rho(0^\ell) = \ell + 1.$\\}
@@
 \vspace{0.5em}
-\textbf{Algorithm} HyperLogLog (\textbf{input} ${\mathcal{X}}$ : multiset of items from domain $\mathcal{Y}$).\\
+\textbf{Algorithm} HyperLogLog (\textbf{input} ${\mathcal{X}}$ : multiset of items from domain $\hat{\mathcal{D}}$).\\
 \textbf{initialize} a collection of $m$ registers, $M[1], ..., M[m]$, to $0$;\\
@@
 In the real world ($b=0$), ${\mathcal{A}}$ interacts with the scheme HLL-$F$.
-This is the same as HLL operating as in Figure~\ref{fig:hll} but where the hash function $h$ is replaced by a keyed function $F: \{0,1\}^k \times {{\mathcal{Y}}} \rightarrow  \{0,1\}^{n+\ell}$ whose key $K$ is sampled uniformly at random at the beginning of the game and not available to the adversary; ${\mathcal{A}}$ has access to the relevant selection of oracles and completes its execution with some output $\textsf{out}_{\mathcal{A}}$ (a bit-string).
+This is the same as HLL operating as in Figure~\ref{fig:hll} but where the hash function $h$ is replaced by a keyed function $F: \{0,1\}^k \times {{\hat{\mathcal{D}}}} \rightarrow  \{0,1\}^{n+\ell}$ whose key $K$ is sampled uniformly at random at the beginning of the game and not available to the adversary; ${\mathcal{A}}$ has access to the relevant selection of oracles and completes its execution with some output $\textsf{out}_{\mathcal{A}}$ (a bit-string).
 
@@
 \begin{itemize}
-\item \texttt{init($\cdot$)}: this oracle receives as input the HLL-$F$ parameters ($n$, $\ell$, $\alpha_m$) and performs initialisation of the HLL-$F$ sketch, including generating the array of $m=2^n$ buckets $M[j]$ and initialising all the entries to $0$, and selecting a key $K \leftarrow_{\$} \{0,1\}^k$ for keyed function $F: \{0,1\}^k \times {{\mathcal{Y}}} \rightarrow  \{0,1\}^{n+\ell}$.
+\item \texttt{init($\cdot$)}: this oracle receives as input the HLL-$F$ parameters ($n$, $\ell$, $\alpha_m$) and performs initialisation of the HLL-$F$ sketch, including generating the array of $m=2^n$ buckets $M[j]$ and initialising all the entries to $0$, and selecting a key $K \leftarrow_{\$} \{0,1\}^k$ for keyed function $F: \{0,1\}^k \times {{\hat{\mathcal{D}}}} \rightarrow  \{0,1\}^{n+\ell}$.
 This oracle has no output.
@@
 We assume it is only called once as ${\mathcal{A}}$'s first oracle query.
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{Y}}}$, the oracle implements HLL-$F$ by setting $y = F_K(x)$ and following the procedure in Figure~\ref{fig:hll} to insert $x$ into the HLL sketch (that is, it uses the first $n$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
+\item \texttt{insert($\cdot$)}: on input $x \in {{\hat{\mathcal{D}}}}$, the oracle implements HLL-$F$ by setting $y = F_K(x)$ and following the procedure in Figure~\ref{fig:hll} to insert $x$ into the HLL sketch (that is, it uses the first $n$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$).
 This oracle has no output.
@@
 \item \texttt{init($\cdot$)}: on input $n$, $\ell$, $\alpha_m$, ${\mathcal{S}}$ initialises an HLL sketch as in Figure~\ref{fig:hll}; that is, it creates an array of $m=2^n$ buckets $M[j]$ and initialises them all to 0.
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{Y}}}$,  ${\mathcal{S}}$ ignores $x$, sets $y \leftarrow_{\$} \{0,1\}^{n+\ell}$, and proceeds to update the buckets using this value of $y$ as in the original HLL algorithm in Figure~\ref{fig:hll}.
+\item \texttt{insert($\cdot$)}: on input $x \in {{\hat{\mathcal{D}}}}$,  ${\mathcal{S}}$ ignores $x$, sets $y \leftarrow_{\$} \{0,1\}^{n+\ell}$, and proceeds to update the buckets using this value of $y$ as in the original HLL algorithm in Figure~\ref{fig:hll}.
 \item \texttt{read}:  ${\mathcal{S}}$ returns the HLL cardinality estimate computed as per Figure~\ref{fig:hll}.
@@
 \begin{definition}
-Let $F: \{0,1\}^k \times {{\mathcal{Y}}} \rightarrow  \{0,1\}^{n+\ell}$.
+Let $F: \{0,1\}^k \times {{\hat{\mathcal{D}}}} \rightarrow  \{0,1\}^{n+\ell}$.
 We define the PRF advantage of an adversary ${{\mathcal{B}}}$ against $F$ outputting a bit $d'$ as being its advantage in distinguishing whether it is interacting with $F$ for a uniformly random key $K$ ($d=0$) or interacting with a truly random function ($d=1$).
@@
 However, it is easily modified to provide a confidentiality notion.
-For example, we can consider the insertion of honestly generated items $x$ that are drawn from some probability distribution $p$ on $\mathcal{Y}$ in a first phase, with ${\mathcal{A}}$ then having access to only \texttt{read} and \texttt{corrupt} oracles in a second phase (we have to remove the \texttt{insert($\cdot$)} oracle in view of the general attack from~\cite{cardestprivacy}).
+For example, we can consider the insertion of honestly generated items $x$ that are drawn from some probability distribution $p$ on $\hat{\mathcal{D}}$ in a first phase, with ${\mathcal{A}}$ then having access to only \texttt{read} and \texttt{corrupt} oracles in a second phase (we have to remove the \texttt{insert($\cdot$)} oracle in view of the general attack from~\cite{cardestprivacy}).
 
@@
 
-Our HLL construction requires a PRF $F: \{0,1\}^k \times {{\mathcal{Y}}} \rightarrow  \{0,1\}^{n+\ell}$.
-Here, the set ${{\mathcal{Y}}}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
+Our HLL construction requires a PRF $F: \{0,1\}^k \times {{\hat{\mathcal{D}}}} \rightarrow  \{0,1\}^{n+\ell}$.
+Here, the set ${{\hat{\mathcal{D}}}}$ represents the space of possible inputs to HLL which can be modelled by a set of bit-strings of some maximum length.
 This means that $F$ needs to be Variable Input Length PRF (VIL-PRF).
@@
 
-%A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {{\mathcal{Y}}} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
+%A particularly simple construction for $F$ involves composing a collision-resistant hash function $H: {{\hat{\mathcal{D}}}} \rightarrow  {\cal E}$ onto some domain ${\cal E}$ with a PRF $G: \{0,1\}^k \times {\cal E} \rightarrow \{0,1\}^{n+\ell}$, viz: $F(K,x) = G(K,H(x))$.
 The result is a PRF with the required input-output signature~\cite[Section 8.2]{BonehShoup}.
@@
 
-The authors thank Anupama Unnikrishnan for useful discussions on simulation-based definitions and Tom Shrimpton for suggesting to look at GoogleBigQuery.
+The authors thank Dennis Hofheinz and Anupama Unnikrishnan for useful discussions on simulation-based definitions and Tom Shrimpton for suggesting to look at GoogleBigQuery.
 \fi
@@
 \item \texttt{init($\cdot$)}: ${{\mathcal{B}}}$ receives parameters $n$, $\ell$, $\alpha_m$, sets $m=2^n$ and initialises an array $M[j]$ of size $m$ in which each of the entries is set to 0. 
-\item \texttt{insert($\cdot$)}: on input $x \in {{\mathcal{Y}}}$,  ${{\mathcal{B}}}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${{\mathcal{B}}}$ uses $y$ to update the HLL sketch as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
+\item \texttt{insert($\cdot$)}: on input $x \in {{\hat{\mathcal{D}}}}$,  ${{\mathcal{B}}}$ makes a query on $x$ to its real-or-random oracle and receives value $y$; ${{\mathcal{B}}}$ uses $y$ to update the HLL sketch as in the regular HLL algorithm (shown in Figure~\ref{fig:hll}).
 That is, it uses the first $n$ bits of $y$ to determine a bucket index $j$ and the next $\ell$ bits of $y$ to determine a bit-string $w$; then $\rho(w)$ is used to update the bucket at index $j$ via $M[j]:= \max(M[j],\rho(w))$.

### 2021-09-07 docs/HLL_exp_bad/main.tex
final changes to proof

@@
 We now formalise the above intuition.
-Our definitions are specific to HLL-$F$ as the object of analysis, but are easily generalised to any cardinality estimator with the same syntax as HLL-$F$ (formally defined by a triple of algorithms (\texttt{Init}, \texttt{Insert}, \texttt{Read})).
+Our definitions are specific to HLL-$F$ as the object of analysis, but are easily generalised to any cardinality estimator with the same syntax as HLL-$F$ (that can be formally defined by a triple of algorithms (\texttt{Init}, \texttt{Insert}, \texttt{Read})).
 
@@
 In the above definition, we quantify over \texttt{insert($\cdot$)} queries (via $q$) and not other types of query because only \texttt{insert($\cdot$)} queries will appear in our concrete security analysis and we do not want to clutter the notation.
-
-Note that our security analysis using a simulation-based definition does not require us to specify a winning condition for adversary $\mathcal{A}$.
+Note also that our definition is given with respect to a \emph{specific} simulator, but could be generalised to allow any simulator that performs random sampling when handling \texttt{insert($\cdot$)} queries.
+
+Our simulation-based security definition does not require us to specify a winning condition for adversary $\mathcal{A}$.
 Rather, we infer from inspecting the simulator what the adversary can (or cannot) achieve in the ideal world, and show that no efficient algorithm $\mathcal{D}$ can distinguish the real and ideal worlds.
@@
 This reveals a strength of the simulation-based approach.
-
-Note also that our definition is given with respect to a \emph{specific} simulator, but could be generalised to allow any simulator that performs random sampling when handling \texttt{insert($\cdot$)} queries.
 

### 2021-09-18 docs/HLL_exp_bad/main.tex
updating related work to include new papers

@@
     \item Google uses a variant of HLL described in~\cite{hllpractice} in its PowerDrill system (first described in~\cite{HallBBGN12}) and possibly in a variety of other internal systems reported in~\cite{hllpractice}.
-For example, cardinality estimation is used to to determine the number of distinct search queries on \url{google.com} over a time period~\cite{hllpractice}.
+For example, cardinality estimation is used to determine the number of distinct search queries on \url{google.com} over a time period~\cite{hllpractice}.
 %HLL is also used for cardinality estimation via an API in Google Big Query.
@@
 We discuss the two attack scenarios from~\cite{cardestprivacy} in more detail in Section~\ref{sec:attacks}.
-%\todo{update: we will get privacy for free from our sim based security model if sim does not use input $x$.}
-
-%In particular, we will show that an adversary who can insert items into the HLL and who can access the HLL cardinality estimate (via typical APIs in HLL implementations) as permitted in the external scenario of~\cite{cardestprivacy} can actually realise the same attacks as the insider adversary can, at least insofar as cardinality estimate manipulation is concerned.
 
@@
 Our use of a simulation-based security definition in place of the game-based ones used in~\cite{CCS:ClaPatShrPS19} is another point of distinction; see further discussion in Section~\ref{sec:formal}.
-
 Prior work also focused on Bloom filters~\cite{NaorY15,GerbetKL15} or flooding of hash tables~\cite{LiptonN93,CrosbyW03}.
-References~\cite{GerbetKL15,CCS:ClaPatShrPS19} provide comprehensive summaries of prior work in this area.
+%References~\cite{GerbetKL15,CCS:ClaPatShrPS19} provide summaries of prior work in this area.
+Recent papers~\cite{Ben-EliezerJWY20,DBLP:journals/corr/abs-2011-07471,HassidimKMMS20,KaplanMNS21} have studied broad classes streaming algorithms under adversarial input from a theoretical viewpoint.
+Our applied viewpoint complements this work, focussing on the real-world usage of a specific cardinality estimator.
+%The authors of~\cite{Ben-EliezerJWY20} give a compiler that shows how to generically transform a non-robust streaming algorithm into a robust one in various scenarios.
+For the particular case of cardinality estimation, this results in solutions that are much less efficient than ours using a PRF; the security model used in this line of work is also weaker than our strongest model.
 
@@
 
-The authors thank Dennis Hofheinz and Anupama Unnikrishnan for useful discussions on simulation-based definitions and Tom Shrimpton for suggesting to look at GoogleBigQuery.
+The authors thank Dennis Hofheinz and Anupama Unnikrishnan for useful discussions on simulation-based definitions, Tom Shrimpton for suggesting to look at GoogleBigQuery, and Eylon Yogev for drawing our attention to related work on streaming algorithms with adversarial inputs.
 \fi

### 2021-12-17 docs/HLL_exp_bad/main.tex
fix appendix sections

@@
 
-\section{Further analysis of Flexswitches}\label{sec:flexswitch}
+\subsection{Further analysis of Flexswitches}\label{sec:flexswitch}
 
@@
 
-\section{Proof of Theorem 8.1}\label{sec:proof} %ridiculous, but had to hard-code this reference
+\subsection{Proof of Theorem 8.1}\label{sec:proof} %ridiculous, but had to hard-code this reference
 
@@
 
-\section{Breaking HLL-UHF}\label{sec:UHF}
+\subsection{Breaking HLL-UHF}\label{sec:UHF}
 

### 2022-01-24 docs/HLL_exp_bad/main.tex
first set of small changes, referenced in red in reviews in google doc

@@
 We stress that none of our attacks are particularly deep in cryptanalytic terms.
+Fundamentally, they arise from implementations of HLL using a fixed hash function $h$ internally.
 Yet HLL is being increasingly widely deployed, and never with protections against our attacks (insofar as we are aware).
@@
 Section~\ref{sec:overview} gives an overview of HLL.
-Section~\ref{sec:attacks} presents our different adversarial models.
+Section~\ref{sec:attacks} informally presents our different adversarial models.
 In Section~\ref{sec:manip}, we describe attacks on HLL and evaluate their impacts under the different adversarial models.
@@
 Our experimental results are discussed and related back to a specific application setting in Section~\ref{sec:exp}.
-Section~\ref{sec:formal} provides our formal security analysis of HLL.
+Section~\ref{sec:formal} provides our formal security model and security proof for HLL using a VIL-PRF.
 Section~\ref{sec:conclusions} provides our conclusions and ideas for future work.
@@
 We show in Section~\ref{sec:formal} that when a salt is used appropriately (i.e.\ as the secret key for a pseudo-random function that is used in place of hash function $h$), security against adversarial inputs can be achieved even in our strongest adversarial setting.
+However, this comes at the cost of losing mergeability.
 %In addition, our simulation-based security definition, in which the simulator in the ideal world has no access to the inserted values, automatically guarantees confidentiality of the values inserted into the HLL sketch, even in the strongest attack model considered in~\cite{cardestprivacy}.
@@
 Recent papers~\cite{Ben-EliezerJWY20,DBLP:journals/corr/abs-2011-07471,HassidimKMMS20,KaplanMNS21} have studied broad classes streaming algorithms under adversarial input from a theoretical viewpoint.
-Our applied viewpoint complements this work, focussing on the real-world usage of a specific cardinality estimator.
+Our applied perspective complements this work, focussing on the real-world usage of a specific cardinality estimator.
 %The authors of~\cite{Ben-EliezerJWY20} give a compiler that shows how to generically transform a non-robust streaming algorithm into a robust one in various scenarios.
@@
 In particular, an S2 adversary does not have access to the values of the HLL cardinality estimate at any point in its attack.
-This kind of attack scenario is appropriate in the context of the port scanning attack of~\cite{portscanhll}.
+This kind of attack scenario is appropriate in the context of the port scanning attack of~\cite{portscanhll}, where the adversary has no API to the target HLL sketch but can cause entries to be inserted into it by a network monitor.
 \item[S3:] The adversary has access to the details of the HLL implementation, i.e.\ it knows the number of buckets $m=2^n$ and the hash function $h$ in use.
@@
 However, it does not know the individual bucket contents.
+This scenario corresponds to a client-server setting with the adversary playing the role of the client.
 \item[S4:] The adversary has direct access to the HLL sketch and all its internals, i.e., the number of buckets, the hash function $h$, and also the contents of each bucket in the HLL sketch \emph{at one specific point in time}, i.e.\ it is given a \emph{snapshot} of the sketch.
 It can access and interact with the sketch via an API provided by the sketch owner, as per scenario S3.
+This scenario models settings where the service hosting the HLL sketch has been compromised by the adversary (and where the adversary continues to act as a client, as in S3).
 %Note that, using these capabilities, an S4 adversary can compute a lower bound on the HLL cardinality estimate based on its initial snapshot and whatever it causes to be inserted into the HLL sketch during its attack (this is irrespective of whether its attack proceeds with input from other users being added to the HLL sketch in parallel with its attack or not; if it knows that there is no other input, then it can compute the \emph{exact} value of the HLL cardinality estimate at every point in its attack instead of a lower bound).
@@
 However, in practice, the raw cardinality estimate is rounded before being presented to the adversary, so this change may not actually be detectable, especially when $M[j]$ becomes large.
-We have done experiments which shows this to be the case for parameters arising in practice.
+We have done experiments which show this to be the case for parameters arising in practice.
 
@@
 A simple computation based on the HLL raw cardinality estimator function in Fig.~\ref{fig:hll} shows that this rule would allow the insertion of a fraction of  $1-2^{-v}$ of the set $\mathcal{X}$ whilst keeping the raw cardinality estimate below $\alpha_m m 2^v$.
-This attack assumes $h$ behaves like a random function and exploits the geometric distribution of the position of the leading ``1'' bit in $w$, where $w = y_{n+1}y_{n+2}\ldots y_{n+1}$ and $y=h(x)$ (as in the HLL algorithm).
+This attack assumes $h$ behaves like a random function and exploits the geometric distribution of the position of the leading ``1'' bit in $w$, where $w = y_{n+1}y_{n+2}\ldots y_{n+\ell}$ and $y=h(x)$ (as in the HLL algorithm).
 
@@
 
-\subsubsection{Initially Empty HLL Sketch, Scenarios S2 and S4:}
+\subsubsection{Initially Empty HLL Sketch, Scenarios S2 and S4}
 
@@
 
-As a second example, setting $B = 1$, we can expect to insert $|{\mathcal{X}}|/2^m$ items with a final cardinality estimate equal to $m \cdot \ln (1+\frac{1}{m-1})$, which is rounded to $1$ by the HLL algorithm (for $m$ of practical interest, e.g.\ $m=256$).
-Thus we keep the HLL cardinality estimate to a constant (1!) whilst being able to insert a constant fraction of the items from the input set ${\mathcal{X}}$.
-
-\subsubsection{Initially Empty HLL Sketch, Scenario S1:}
+As a second example, setting $B = 1$, we can expect to insert $|{\mathcal{X}}|/m$ items with a final cardinality estimate equal to $m \cdot \ln (1+\frac{1}{m-1})$, which is rounded to $1$ by the HLL algorithm (for $m$ of practical interest, e.g.\ $m=256$).
+Thus we keep the HLL cardinality estimate to a constant (1!) whilst being able to insert a constant fraction $1/m$ of the items from the input set ${\mathcal{X}}$.
+
+\subsubsection{Initially Empty HLL Sketch, Scenario S1}
 
@@
 
-\paragraph{Core Idea:}
+\paragraph{\textbf{Core Idea}}
 We assume the adversary again has (or is given) some set ${\mathcal{X}}$ of items from which to pick its inputs.
@@
 We initialise a list ${\mathcal{L}}$ containing $x_1,\ldots, x_B$.
-We add those items that do not increase the cardinality estimate on insertion into the shadow HLL sketch to ${\mathcal{L}}$ and reject the rest.
+We add those items $x \in {\mathcal{X}}$ that do not increase the cardinality estimate on insertion into the shadow HLL sketch to ${\mathcal{L}}$ and reject the rest.
 We reset the shadow device when the cardinality estimate has increased too much, reinserting the first $B$ items from ${\mathcal{X}}$ again to re-establish a ``baseline'' of target buckets for the shadow HLL sketch.
@@
 
-\paragraph{Simple Attack:}
+\paragraph{\textbf{Simple Attack}}
 The simple version of our attack uses the above idea, inserting items $x$ into the shadow HLL sketch but resetting it as soon as the cardinality estimate increases from $B$ to $B+1$ (indicating that a non-target bucket has been hit by item $x$).
@@
 
-\paragraph{Complex Attack:}
+\paragraph{\textbf{Complex Attack}}
 The more complex version of the attack only resets the shadow HLL sketch when its cardinality estimate reaches a value indicating that $B_{\max}$ buckets have been hit (recall this number can be determined precisely because of the use of the Hit Counting algorithm for cardinality estimation).
@@
 
-\paragraph{Analysis of the Complex Attack:}
+\paragraph{\textbf{Analysis of the Complex Attack}}
 %Suppose that after each reset, the order of the items in ${\mathcal{X}}$ (or ${\mathcal{L}}$) that have not yet been processed on the current pass is randomised.
@@
 
-\subsubsection{Initially Non-empty HLL Sketch, Scenario S2:}\label{sec:nonemptyHLLS2}
+\subsubsection{Initially Non-empty HLL Sketch, Scenario S2}\label{sec:nonemptyHLLS2}
 The adversary knows $h$ so it selects items $x$ from ${\mathcal{X}}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit.
@@
 
-\subsubsection{Initially Non-empty HLL Sketch, Scenario S4:}\label{sec:nonemptyHLLS4}
+\subsubsection{Initially Non-empty HLL Sketch, Scenario S4}\label{sec:nonemptyHLLS4}
 The adversary can adopt the same strategy as for S2, but relax the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch).
@@
 
-\subsubsection{Initially Non-empty HLL Sketch, Scenario S1:}
+\subsubsection{Initially Non-empty HLL Sketch, Scenario S1}
 
@@
 For example, since it has (roughly) uniform outputs, perhaps using a universal hash function (UHF) $h_K(\cdot)$ would be sufficient.
-In Appendix~\ref{sec:UHF} we show that, in general, UHFs do not suffice by providing a key recovery attack against HLL when $h$ is replaced by a specific UHF based on polynomial evaluation over a finite field.
+In Appendix~\ref{sec:UHF} we provide a key recovery attack against HLL when $h$ is replaced by a \emph{specific} UHF based on polynomial evaluation over a finite field.
+This is enough to show that UHFs do not suffice in general.
 
@@
 
-We want $F$ to be competitive with non-cryptographic hashing on small input sizes, as typical in current HLL implementations and applications (e.g.\ MurmurHash3 applied to an IP header).
+From inspection of Figure~\ref{fig:hll}, it is evident that hashing is the dominant cost when inserting an item into HLL.
+So we want $F$ to be competitive with non-cryptographic hashing, especially on small input sizes (as typical in current HLL implementations and applications, e.g.\ MurmurHash3 applied to an IP header).
 This makes off-the-shelf approaches like HMAC uncompetitive, since even for the shortest inputs, HMAC would require 4 evaluations of the underlying hash compression function.
@@
 This may also require the introduction of entirely new data structures that trade increased storage for mergeability and improved security.
+It is also an open problem to extend our simulation-based approach to formal analysis to other data structures.
 
@@
 \bibliographystyle{ieeetr}
-\bibliography{ref.bib}
+\bibliography{./ref.bib}
 

### 2022-01-24 docs/HLL_exp_bad/main.tex
more changes

@@
 
-As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}, since it is faithful to the original description of HLL given in~\cite{hll}, and against the HLL implementation used in Redis~\cite{redis}.\footnote{Our attack code is available on request.}
+As a proof-of-concept, we implemented our attacks against the HLL implementation from~\cite{clahll}, since it is faithful to the original description of HLL given in~\cite{hll}, and against the HLL implementation used in Redis~\cite{redis}.\footnote{All our attack code is available at \href{https://github.com/PizzaWhisperer/HLLVuln}{\url{https://github.com/PizzaWhisperer/HLLVuln}}.}
 %at \href{https://github.com/PizzaWhisperer/HLLVuln}{\url{https://github.com/PizzaWhisperer/HLLVuln}}.} 

### 2022-01-26 attacks/attack_hll.go
working on code

@@ -1,0 +1,290 @@
+package main
+
+import (
+	"fmt"
+	"hash"
+  "math/rand"
+	"math"
+	"time"
+	"encoding/binary"
+
+	"./clarkduvall/hyperloglog"
+	"./spaolacci/murmur3"
+)
+
+var n = uint8(8)
+var m = 1 << n
+var maxValue = int(math.Pow(2, 20))-1
+var iterations = 50
+
+
+func main() {
+
+	fmt.Printf("Results for %d iterations:\n", iterations)
+
+  //S1 scenario
+var originalEst, finalEst, nItermInserted int
+	//originalEst, finalEst, nItermInserted, nResets = runAttack("S1", 0, false, 1)
+	//fmt.Printf("Results in the S1 scenario: %d, %d, %d, %d\n", originalEst, finalEst, nItermInserted, nResets)
+
+  //originalEst, finalEst, nItermInserted, nResets = runAttack("S1", 0, false, 100)
+  //fmt.Printf("Results in the S1 scenario with B = 100: %d, %d, %d, %d\n", originalEst, finalEst, nItermInserted, nResets)
+
+  //S2 scenario
+
+  originalEst, finalEst, nItermInserted, _ = runAttack("S2", 0, false, 1)
+  fmt.Printf("Results in the S2 scenario: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+
+  originalEst, finalEst, nItermInserted, _ = runAttack("S2", 500, false, 1)
+  fmt.Printf("Results in the S2 scenario with 500 initial items: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+
+  originalEst, finalEst, nItermInserted, _ = runAttack("S2", 1000, false, 1)
+  fmt.Printf("Results in the S2 scenario with 1000 initial items: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+
+  originalEst, finalEst, nItermInserted, _ = runAttack("S2", 2000, false, 1)
+  fmt.Printf("Results in the S2 scenario with 2000 initial items: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+
+  //S4 scenario
+
+  originalEst, finalEst, nItermInserted, _ = runAttack("S4", 0, false, 1)
+  fmt.Printf("Results in the S4 scenario with B = 1: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+
+  originalEst, finalEst, nItermInserted, _ = runAttack("S4", 1000, false, 1)
[…]

### 2022-01-26 docs/HLL_exp_bad/main.tex
updates to Section 5.2 on attacks when HLL non-empty

@@
 
-We do not know of attacks which are able to use the extra information available to an S3 adversary over an S2 adversary (namely, the ability to obtain cardinality estimates).
+We only know of one attack setting where the extra power available to an S3 adversary over an S2 adversary (namely, the ability to obtain cardinality estimates) is useful -- see Section~\ref{sec:nonemptyHLLS2}.
 Moreover, as we will see below, our attacks in the S2 setting are already quite powerful.
@@
 
-We assume that the target HLL in S2 and S4 has already had sufficiently many items inserted that every one of the $m=2^n$ buckets has counter value $M[j]$ with $M[j] \ge 1$.
-This is highly likely to be the case (with probability at least $1-1/m^2$) as long as at least $m\ln m$ items have already been inserted into the target HLL.
-This follows from the standard analysis of the coupon collector problem~\cite[Section 3.6]{MR95} under our usual assumption that $h$ behaves as a random function.
-In S4, the adversary can use its snapshot to detect if this is the case or not, and behave differently if not (details for this case follow below).
-
-With $m \ln m$ honest insertions, it is also highly likely that the small range correction (Hit Counting algorithm) will not be used, and the target HLL's cardinality estimate will be computed using the raw estimate.
+\subsubsection{Initially Non-empty HLL Sketch, Scenario S2}\label{sec:nonemptyHLLS2}
+
+We assume that at least $2^{t-1}\cdot m\ln m$ items have already been honestly inserted into the target HLL, for some integer $t$ with $1 \leq t \leq \ell$ which will be a parameter of our attack.
+We assume the S2 adversary somehow knows that this condition holds (e.g.\ through knowledge of when the HLL was initialised and the average rate at which items are inserted into it).
+Notice that an S3 adversary can use its ability to get cardinality estimates to obtain this knowledge; this is one situation where an S3 adversary can make good use of the extra information available to it as compared to an S2 adversary.
+In S4, the adversary can use its access to a snapshot to mount a more powerful attack; see Section~\ref{sec:nonemptyHLLS4}.
+
+Under the condition that at least $2^{t-1}\cdot m\ln m$ items have already been honestly inserted, it is highly likely that every one of the $m=2^n$ buckets has counter value $M[j]$ with $M[j] \ge t$.
+This follows from an extension of the standard analysis of the coupon collector problem~\cite[Section 3.6]{MR95} under our usual assumption that $h$ behaves as a random function.
+To see this, note that when an item $x$ is inserted into a specific bucket $j$, the condition $M[j] \ge t$ holds with probability $2^{-t+1}$; we then need about $m\ln m$ such events to occur to ensure the desired condition holds in every bucket.
+
+With $2^{t-1}m \ln m$ honest insertions, it is also highly likely that the small range correction (Hit Counting algorithm) will not be used, and the target HLL's cardinality estimate will be computed using the raw estimate.
 %\[
@@
 
-\subsubsection{Initially Non-empty HLL Sketch, Scenario S2}\label{sec:nonemptyHLLS2}
-The adversary knows $h$ so it selects items $x$ from ${\mathcal{X}}$ for which the $\ell$ rightmost bits of $h(x)$ start with a ``1'' bit.
-In this case, since the relevant bucket $j$ already has $M[j] \ge 1$, the bucket's counter is not updated, and the cardinality estimate of the HLL sketch remains unchanged.
-
-The adversary then expects to be able to insert half of the items from ${\mathcal{X}}$ into the HLL sketch, without increasing the HLL cardinality estimate at all.
-
-If some buckets are actually empty (but the adversary does not know which ones in the blind setting of S2), then we may expect a moderate increase in the HLL cardinality estimate in the above attack.
-If the adversary suspects this is the case, then it can first insert $m$ carefully selected items into the target HLL sketch to ensure that $M[j] \ge 1$ for every bucket.
-This should increase its cardinality estimate by about $m$ (since HLL is a good cardinality estimator!).
+The S2 adversary knows $h$ so it selects items $x$ from ${\mathcal{X}}$ for which the $\ell$ rightmost bits of $h(x)$ start with up to $t-1$ ``0'' bits.
+In this case, since the relevant bucket $j$ already has $M[j] \ge t$, the bucket's counter is not updated, and the cardinality estimate of the HLL sketch remains unchanged.
+
+%% half of hash values begin 1...; 1/4 begin 01 (t=2), etc....
+
+The S2 adversary then expects to be able to insert a fraction $1-1/2^t$ of the items from ${\mathcal{X}}$ into the HLL sketch, without increasing the HLL cardinality estimate at all.
+Notice how the fraction increases towards 1 as parameter $t$ increases: the more items are originally in the sketch, the more items the adversary expects to be able to insert without penalty.
+
+If some buckets are actually empty or contain values less than $t$ (but the adversary does not know which ones in the setting of S2), then we may expect an increase in the HLL cardinality estimate in the above attack.
+If the adversary suspects this is the case, then it can first insert $m$ carefully selected items into the target HLL sketch to ensure that $M[j] \ge t$ for every bucket.
+This should increase the HLL cardinality estimate to about $\alpha_m 2^t m$.
 Then it can mount the attack, and be sure that it does not increase the cardinality estimate any further.
@@
 The adversary can adopt the same strategy as for S2, but relax the sampling requirements with the additional information available in this scenario (namely, the count values $M[j]$ per bucket in the HLL sketch).
-Specifically, it can allow some leading zero bits in the $\ell$ rightmost bits of $h(x)$ so long as this does not increase the current bucket estimate.
-In other words, if $h(x)$ is mapped to the $j$-th bucket containing $M[j] \ge 0$, then, instead of the strict requirement of having a ``1'' bit in the first position in the $\ell$ rightmost bits of $h(x)$, the adversary is satisfied when there is a ``1'' bit in any of the first $M[j]$ positions in this substring.
+Specifically, if $h(x)$ is mapped to the $j$-th bucket containing $M[j] \ge 0$, then the adversary is satisfied when there is a ``1'' bit in any of the first $M[j]$ positions in this substring.
 (A value $M[j]=0$ then never permits insertions, meaning an empty bucket stays empty.) Under these conditions, the estimates in the buckets are never updated, and there is no effect on the cardinality estimate made by HLL.
@@
 where $H_M$ is the harmonic mean of $2^{M[1]}, 2^{M[2]}, \ldots, 2^{M[m]}$.
-So the adversary can expect to insert a fraction $(1 - H_M^{-1})$ of the items from ${\mathcal{X}}$ without increasing the HLL cardinality estimate at all.
+So the adversary can expect to insert a fraction $1 - H_M^{-1}$ of the items from ${\mathcal{X}}$ without increasing the HLL cardinality estimate at all.
 
@@
 
-In this setting, where the target HLL sketch is initially non-empty, we see that an adversary who has sufficient flexibility in its inputs (as represented by the size of the set ${\mathcal{X}}$) can expect to insert many items into the HLL sketch without increasing the cardinality estimate (or whilst increasing the cardinality estimate slightly in the case where the HLL sketch has some empty buckets in S2).
+In this setting, where the target HLL sketch is initially non-empty, we see that an S4 adversary who has sufficient flexibility in its inputs (as represented by the size of the set ${\mathcal{X}}$) can expect to insert many items into the HLL sketch without increasing the cardinality estimate.
 The result is largely independent of the HLL parameters.
-The more full is the HLL sketch, the larger is $1-H_M^{-1}$ and the easier it is to insert new items without increasing the cardinality estimate.
+The more full is the HLL sketch, the larger is $1 - H_M^{-1}$ and the easier it is to insert new items without increasing the cardinality estimate.
 

### 2022-01-26 docs/HLL_exp_bad/main.tex
fixing appendix numbering

@@
 
-\appendix
-
-\subsection{Further analysis of Flexswitches}\label{sec:flexswitch}
+\appendices
+
+\section{Further analysis of Flexswitches}\label{sec:flexswitch}
 
@@
 
-\subsection{Proof of Theorem 8.1}\label{sec:proof} %ridiculous, but had to hard-code this reference
+\section{Proof of Theorem 8.1}\label{sec:proof} %ridiculous, but had to hard-code this reference
 
@@
 
-\subsection{Breaking HLL-UHF}\label{sec:UHF}
+\section{Breaking HLL-UHF}\label{sec:UHF}
 

### 2022-01-26 docs/HLL_exp_bad/main.tex
adding harms section

@@
 We believe our simulation-based approach will be of independent interest and will be useful beyond the analysis of cardinality manipulation attacks, for example when studying privacy properties of HLL and other probabilistic data structures.
+
+\subsection{Potential Harms and Mitigations}\label{sec-harms}
+
+Our attacks are presented in a general way and evaluated either abstractly or in the context of a specific system described in the scientific literature~\cite{flexswitch}.
+We also propose an easily deployable countermeasure.
+Given this, we think the potential for harm to systems and individuals relying on HLL today is small.
+On the other hand, we hope our results will spur practitioners to consider the possibility of adversarial inputs in their specific applications and design accordingly.
+For the specific case of Redis, a real system that we analyse in detail, we have contacted the Redis core team and shared our results, following their advertised security policy.\footnote{\url{https://github.com/redis/redis/security/policy}.}
 

### 2022-01-27 attacks/attack_hll.go
final attack code

@@ -1,12 +1,12 @@
 package main
 
 import (
+	"encoding/binary"
 	"fmt"
 	"hash"
-  "math/rand"
 	"math"
+	"math/rand"
 	"time"
-	"encoding/binary"
 
 	"./clarkduvall/hyperloglog"
 	"./spaolacci/murmur3"
@@ -14,76 +14,75 @@ import (
 
 var n = uint8(8)
 var m = 1 << n
-var maxValue = int(math.Pow(2, 20))-1
+var maxValue = int(math.Pow(2, 20)) - 1
 var iterations = 50
-
 
 func main() {
 
-	fmt.Printf("Results for %d iterations:\n", iterations)
+	fmt.Printf("(original cardinality estimation, final cardinality estimation, number of items inserted by adversary) for %d iterations:\n", iterations)
 
-  //S1 scenario
-var originalEst, finalEst, nItermInserted int
-	//originalEst, finalEst, nItermInserted, nResets = runAttack("S1", 0, false, 1)
-	//fmt.Printf("Results in the S1 scenario: %d, %d, %d, %d\n", originalEst, finalEst, nItermInserted, nResets)
+	//S1 scenario
+	var originalEst, finalEst, nItermInserted, nResets int
 
-  //originalEst, finalEst, nItermInserted, nResets = runAttack("S1", 0, false, 100)
-  //fmt.Printf("Results in the S1 scenario with B = 100: %d, %d, %d, %d\n", originalEst, finalEst, nItermInserted, nResets)
+	originalEst, finalEst, nItermInserted, nResets = runAttack("S1", 0, false, 1)
+	fmt.Printf("Results in the S1 scenario: %d, %d, %d, resets: %d\n", originalEst, finalEst, nItermInserted, nResets)
 
-  //S2 scenario
+	originalEst, finalEst, nItermInserted, nResets = runAttack("S1", 0, false, uint32(m/2))
+	fmt.Printf("Results in the S1 scenario with B = %d: %d, %d, %d, resets: %d\n", uint32(m/2), originalEst, finalEst, nItermInserted, nResets)
 
-  originalEst, finalEst, nItermInserted, _ = runAttack("S2", 0, false, 1)
-  fmt.Printf("Results in the S2 scenario: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+	//S2 scenario
 
-  originalEst, finalEst, nItermInserted, _ = runAttack("S2", 500, false, 1)
[…]

### 2022-01-27 docs/HLL_exp_bad/main.tex
adding results

@@
 \small\centering
-\caption{Attack results for a classic HLL implementation, averaged over 30 iterations.}
+\caption{Attack results for a classic HLL implementation, averaged over 50 iterations.}
 \medskip
      \makebox[\linewidth]{
-\begin{tabular}{| m{11em} | m{4em} | m{4.2em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
+\begin{tabular}{| m{11em} | m{3.4em} | m{4.8em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
     \hline
     \textbf{Scenario} &  \multicolumn{2}{c|}{S1} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
-    \textbf{\# Targeted buckets $B$} & 1 & 100 & n/a & 1 & n/a & 1 \\ \hline
-    \textbf{HLL initially empty?} & yes & yes & no & yes & no & yes \\ \hline
-    \textbf{Original Card. Est.} & 0 & 0 & 1,000 & 0 & 1,005 & 0 \\ \hline
-    \textbf{\# Items added} & 4,096 & 409,595 & 524,288 & 28,566 & 870,320 & 409,595 \\ \hline
-    \textbf{Final Card. Est.} & 1 & 126 & 1,011 & 1 & 1,005 & 1 \\ \hline
+    %\textbf{HLL initially empty?} & yes & yes & yes & no & yes & no \\ \hline
+    \textbf{Original Card. Est.} & 0 & 0 &  0 & 1,001 & 0 & 1,007  \\ \hline
+    \textbf{\# Targeted buckets $B$} & 1 & 128 (= m/2) & 1 & - & 1 & - \\ \hline
+      \textbf{\# Items added} & 4,132 & 523,791  & 4,106 & 524,851 & 4,106 & 857,150 \\ \hline
+     \textbf{Final Card. Est.} & 1 & 177 & 1 & 1,025  & 1 &  1,007 \\ \hline
+    \textbf{\# Resets} & 4,718 & 6,973 & - & - & - & - \\ \hline
+
 \end{tabular}}
@@
 \small\centering
-\caption{Attack results for a classic HLL implementation, averaged over 30 iterations, in the setting of~\cite{hllvuln}.}
+\caption{Attack results for a classic HLL implementation, averaged over 50 iterations, in the setting of~\cite{hllvuln}.}
 \medskip
      \makebox[\linewidth]{
-\begin{tabular}{| m{11em} | m{4em} | m{4.2em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
+\begin{tabular}{| m{11em} | m{3.4em} | m{4.8em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
     \hline
     \textbf{Scenario} & \multicolumn{2}{c|}{S1} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
-    \textbf{\# Targeted buckets $B$ } & 1 & 100 & n/a & 1 & n/a & 1 \\ \hline
-    \textbf{HLL initially empty?} & yes & yes & no & yes & no & yes \\ \hline
-    \textbf{Original Card. Est.} & 0 & 0 & 1,004 & 0 & 998 & 0 \\ \hline
-    \textbf{\# Items added} & 977 & 97,433 & 124,940 & 979 & 204,669 & 988 \\ \hline
-    \textbf{Final Card. Est.} & 1 & 126 & 1,058 & 1 & 998 & 1 \\ \hline
+    %\textbf{HLL initially empty?} & yes & yes & yes & no & yes & no \\ \hline
+      \textbf{Original Card. Est.} & 0 & 0 & 0 & 1,001 & 0 & 1,003 \\ \hline
+       \textbf{\# Targeted buckets $B$} & 1 & 128 (= m/2) & 1 & - & 1 & - \\ \hline
+     \textbf{\# Items added} & 977 & 125,091 & 980 & 125,153 & 1,003 & 204,127 \\ \hline
+    \textbf{Final Card. Est.} & 1 & 177 & 1 & 1,058  & 1 & 1,003 \\ \hline
+    \textbf{\# Resets} & 1,148 & 1,655 & - & - & - & - \\ \hline
+
 \end{tabular}}

### 2022-01-28 attacks/attack_hll.go
clean code Redis

@@ -244,6 +244,41 @@ func AttackS2(hll *hyperloglog.HyperLogLog, m int, hllHash hash.Hash32, RT20 boo
 	return inserted
 }
 
+//Additional attack strategy in the S2 scenario when the adversary has information on the preload
+func AttackS2Preload(hll *hyperloglog.HyperLogLog, m int, hllHash hash.Hash32, RT20 bool, B uint32, T uint8) int {
+	inserted := 0
+	items := rand.Perm(maxValue)
+	if RT20 {
+		items = items[:250000]
+	}
+
+	// 0 < T < 24, in "normal" attack, T == 1
+	mask := GenMask(T)
+
+	emptyBool := hll.Count() == 0
+
+	for _, i := range items {
+
+		_, err := hllHash.Write(itob(i))
+		if err != nil {
+			fmt.Printf("Hash error: err %v\n", err)
+			continue
+		}
+		result := hllHash.Sum32()
+		bucket := (result & uint32(((1<<8)-1)<<24)) >> 24
+
+		if (emptyBool && bucket < B) || (!emptyBool && (result&mask != 0)) {
+			//There is a leading 1, or in case of empty sketch, to targeted bucket.
+			hll.Add(hllHash)
+			inserted++
+		}
+
+		hllHash.Reset()
+	}
+
+	return inserted
+}
+
 //Attack in the S4 scenario as described in the paper
 func AttackS4(hll *hyperloglog.HyperLogLog, m int, hllHash hash.Hash32, RT20 bool, B uint32) int {
 	inserted := 0

### 2022-01-28 attacks/redis/help.h

@@ -1,1448 +1,0 @@
-/* Automatically generated by ./utils/generate-command-help.rb, do not edit. */
-
-#ifndef __REDIS_HELP_H
-#define __REDIS_HELP_H
-
-static char *commandGroups[] = {
-    "generic",
-    "string",
-    "list",
-    "set",
-    "sorted_set",
-    "hash",
-    "pubsub",
-    "transactions",
-    "connection",
-    "server",
-    "scripting",
-    "hyperloglog",
-    "cluster",
-    "geo",
-    "stream"
-};
-
-struct commandHelp {
-  char *name;
-  char *params;
-  char *summary;
-  int group;
-  char *since;
-} commandHelp[] = {
-    { "ACL CAT",
-    "[categoryname]",
-    "List the ACL categories or the commands inside a category",
-    9,
-    "6.0.0" },
-    { "ACL DELUSER",
-    "username [username ...]",
-    "Remove the specified ACL users and the associated rules",
-    9,
-    "6.0.0" },
-    { "ACL GENPASS",
-    "[bits]",
-    "Generate a pseudorandom secure password to use for ACL users",
-    9,
-    "6.0.0" },
-    { "ACL GETUSER",
-    "username",
-    "Get the rules for a specific ACL user",
-    9,
-    "6.0.0" },
-    { "ACL HELP",
-    "-",
-    "Show helpful text about the different subcommands",
-    9,
-    "6.0.0" },
-    { "ACL LIST",
-    "-",
-    "List the current ACL rules in ACL config file format",
-    9,
-    "6.0.0" },
-    { "ACL LOAD",
-    "-",
-    "Reload the ACLs from the configured ACL file",
-    9,
-    "6.0.0" },
-    { "ACL LOG",
-    "[count or RESET]",
-    "List latest events denied because of ACLs in place",
-    9,
-    "6.0.0" },
-    { "ACL SAVE",
-    "-",
-    "Save the current ACL rules in the configured ACL file",
-    9,
-    "6.0.0" },
-    { "ACL SETUSER",
-    "username [rule [rule ...]]",
-    "Modify or create the rules for a specific ACL user",
-    9,
-    "6.0.0" },
-    { "ACL USERS",
-    "-",
-    "List the username of all the configured ACL rules",
-    9,
-    "6.0.0" },
-    { "ACL WHOAMI",
-    "-",
[…]

### 2022-01-28 attacks/redis/hyperloglog.c

@@ -1,1753 +1,0 @@
-/* hyperloglog.c - Redis HyperLogLog probabilistic cardinality approximation.
- * This file implements the algorithm and the exported Redis commands.
- *
- * Copyright (c) 2014, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-
-#include <stdint.h>
-#include <math.h>
-
-/* The Redis HyperLogLog implementation is based on the following ideas:
- *
[…]

### 2022-01-28 attacks/redis/server.h

@@ -1,2750 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#ifndef __REDIS_H
-#define __REDIS_H
-
-#include "fmacros.h"
-#include "config.h"
-#include "solarisfixes.h"
-#include "rio.h"
-#include "atomicvar.h"
-
-#include <stdio.h>
-#include <stdlib.h>
[…]

### 2022-01-28 attacks/redis/src/aof.c

@@ -1,0 +1,2541 @@
+/*
+ * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
+ * All rights reserved.
+ *
+ * Redistribution and use in source and binary forms, with or without
+ * modification, are permitted provided that the following conditions are met:
+ *
+ *   * Redistributions of source code must retain the above copyright notice,
+ *     this list of conditions and the following disclaimer.
+ *   * Redistributions in binary form must reproduce the above copyright
+ *     notice, this list of conditions and the following disclaimer in the
+ *     documentation and/or other materials provided with the distribution.
+ *   * Neither the name of Redis nor the names of its contributors may be used
+ *     to endorse or promote products derived from this software without
+ *     specific prior written permission.
+ *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+ * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+ * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
+ * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
+ * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
+ * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
+ * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
+ * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
+ * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
+ * POSSIBILITY OF SUCH DAMAGE.
+ */
+
+#include "server.h"
+#include "bio.h"
+#include "rio.h"
+#include "functions.h"
+
+#include <signal.h>
+#include <fcntl.h>
+#include <sys/stat.h>
+#include <sys/types.h>
[…]

### 2022-01-28 attacks/redis/src/cluster.h

@@ -1,0 +1,369 @@
+#ifndef __CLUSTER_H
+#define __CLUSTER_H
+
+/*-----------------------------------------------------------------------------
+ * Redis cluster data structures, defines, exported API.
+ *----------------------------------------------------------------------------*/
+
+#define CLUSTER_SLOTS 16384
+#define CLUSTER_OK 0            /* Everything looks ok */
+#define CLUSTER_FAIL 1          /* The cluster can't work */
+#define CLUSTER_NAMELEN 40      /* sha1 hex length */
+#define CLUSTER_PORT_INCR 10000 /* Cluster port = baseport + PORT_INCR */
+
+/* The following defines are amount of time, sometimes expressed as
+ * multiplicators of the node timeout value (when ending with MULT). */
+#define CLUSTER_FAIL_REPORT_VALIDITY_MULT 2 /* Fail report validity. */
+#define CLUSTER_FAIL_UNDO_TIME_MULT 2 /* Undo fail if master is back. */
+#define CLUSTER_MF_TIMEOUT 5000 /* Milliseconds to do a manual failover. */
+#define CLUSTER_MF_PAUSE_MULT 2 /* Master pause manual failover mult. */
+#define CLUSTER_SLAVE_MIGRATION_DELAY 5000 /* Delay for slave migration. */
+
+/* Redirection errors returned by getNodeByQuery(). */
+#define CLUSTER_REDIR_NONE 0          /* Node can serve the request. */
+#define CLUSTER_REDIR_CROSS_SLOT 1    /* -CROSSSLOT request. */
+#define CLUSTER_REDIR_UNSTABLE 2      /* -TRYAGAIN redirection required */
+#define CLUSTER_REDIR_ASK 3           /* -ASK redirection required. */
+#define CLUSTER_REDIR_MOVED 4         /* -MOVED redirection required. */
+#define CLUSTER_REDIR_DOWN_STATE 5    /* -CLUSTERDOWN, global state. */
+#define CLUSTER_REDIR_DOWN_UNBOUND 6  /* -CLUSTERDOWN, unbound slot. */
+#define CLUSTER_REDIR_DOWN_RO_STATE 7 /* -CLUSTERDOWN, allow reads. */
+
+struct clusterNode;
+
+/* clusterLink encapsulates everything needed to talk with a remote node. */
+typedef struct clusterLink {
+    mstime_t ctime;             /* Link creation time */
+    connection *conn;           /* Connection to remote node */
[…]

### 2022-01-28 attacks/redis/src/db.c

@@ -1,0 +1,2260 @@
+/*
+ * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
+ * All rights reserved.
+ *
+ * Redistribution and use in source and binary forms, with or without
+ * modification, are permitted provided that the following conditions are met:
+ *
+ *   * Redistributions of source code must retain the above copyright notice,
+ *     this list of conditions and the following disclaimer.
+ *   * Redistributions in binary form must reproduce the above copyright
+ *     notice, this list of conditions and the following disclaimer in the
+ *     documentation and/or other materials provided with the distribution.
+ *   * Neither the name of Redis nor the names of its contributors may be used
+ *     to endorse or promote products derived from this software without
+ *     specific prior written permission.
+ *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+ * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+ * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
+ * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
+ * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
+ * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
+ * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
+ * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
+ * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
+ * POSSIBILITY OF SUCH DAMAGE.
+ */
+
+#include "server.h"
+#include "cluster.h"
+#include "atomicvar.h"
+#include "latency.h"
+#include "script.h"
+#include "functions.h"
+
+#include <signal.h>
+#include <ctype.h>
+
+/*-----------------------------------------------------------------------------
+ * C-level DB API
[…]

### 2022-01-28 attacks/redis/src/help.h

@@ -1,0 +1,1443 @@
+/* Automatically generated by ./utils/generate-command-help.rb, do not edit. */
+
+#ifndef __REDIS_HELP_H
+#define __REDIS_HELP_H
+
+static char *commandGroups[] = {
+    "generic",
+    "string",
+    "list",
+    "set",
+    "sorted_set",
+    "hash",
+    "pubsub",
+    "transactions",
+    "connection",
+    "server",
+    "scripting",
+    "hyperloglog",
+    "cluster",
+    "geo",
+    "stream"
+};
+
+struct commandHelp {
+  char *name;
+  char *params;
+  char *summary;
+  int group;
+  char *since;
+} commandHelp[] = {
+    { "ACL CAT",
+    "[categoryname]",
+    "List the ACL categories or the commands inside a category",
+    9,
+    "6.0.0" },
+    { "ACL DELUSER",
+    "username [username ...]",
+    "Remove the specified ACL users and the associated rules",
+    9,
+    "6.0.0" },
+    { "ACL GENPASS",
+    "[bits]",
+    "Generate a pseudorandom secure password to use for ACL users",
+    9,
+    "6.0.0" },
+    { "ACL GETUSER",
+    "username",
+    "Get the rules for a specific ACL user",
+    9,
+    "6.0.0" },
+    { "ACL HELP",
+    "",
+    "Show helpful text about the different subcommands",
+    9,
+    "6.0.0" },
+    { "ACL LIST",
+    "",
+    "List the current ACL rules in ACL config file format",
+    9,
+    "6.0.0" },
+    { "ACL LOAD",
+    "",
+    "Reload the ACLs from the configured ACL file",
+    9,
+    "6.0.0" },
+    { "ACL LOG",
+    "[count or RESET]",
+    "List latest events denied because of ACLs in place",
+    9,
+    "6.0.0" },
+    { "ACL SAVE",
+    "",
+    "Save the current ACL rules in the configured ACL file",
+    9,
+    "6.0.0" },
+    { "ACL SETUSER",
+    "username [rule [rule ...]]",
+    "Modify or create the rules for a specific ACL user",
+    9,
+    "6.0.0" },
+    { "ACL USERS",
+    "",
+    "List the username of all the configured ACL rules",
+    9,
+    "6.0.0" },
+    { "ACL WHOAMI",
+    "",
+    "Return the name of the user associated to the current connection",
[…]

### 2022-01-28 attacks/redis/src/hyperloglog.c

@@ -1,0 +1,1693 @@
+/* hyperloglog.c - Redis HyperLogLog probabilistic cardinality approximation.
+ * This file implements the algorithm and the exported Redis commands.
+ *
+ * Copyright (c) 2014, Salvatore Sanfilippo <antirez at gmail dot com>
+ * All rights reserved.
+ *
+ * Redistribution and use in source and binary forms, with or without
+ * modification, are permitted provided that the following conditions are met:
+ *
+ *   * Redistributions of source code must retain the above copyright notice,
+ *     this list of conditions and the following disclaimer.
+ *   * Redistributions in binary form must reproduce the above copyright
+ *     notice, this list of conditions and the following disclaimer in the
+ *     documentation and/or other materials provided with the distribution.
+ *   * Neither the name of Redis nor the names of its contributors may be used
+ *     to endorse or promote products derived from this software without
+ *     specific prior written permission.
+ *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+ * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+ * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
+ * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
+ * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
+ * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
+ * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
+ * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
+ * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
+ * POSSIBILITY OF SUCH DAMAGE.
+ */
+
+#include "server.h"
+
+#include <stdint.h>
+#include <math.h>
+
[…]

### 2022-01-28 attacks/redis/src/release.h

@@ -1,0 +1,3 @@
+#define REDIS_GIT_SHA1 "b36391c5"
+#define REDIS_GIT_DIRTY "   12469"
+#define REDIS_BUILD_ID "Mathildes-MacBook-Air.local-1643356764"

### 2022-01-28 attacks/redis/src/server.h

@@ -1,0 +1,3432 @@
+/*
+ * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
+ * All rights reserved.
+ *
+ * Redistribution and use in source and binary forms, with or without
+ * modification, are permitted provided that the following conditions are met:
+ *
+ *   * Redistributions of source code must retain the above copyright notice,
+ *     this list of conditions and the following disclaimer.
+ *   * Redistributions in binary form must reproduce the above copyright
+ *     notice, this list of conditions and the following disclaimer in the
+ *     documentation and/or other materials provided with the distribution.
+ *   * Neither the name of Redis nor the names of its contributors may be used
+ *     to endorse or promote products derived from this software without
+ *     specific prior written permission.
+ *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+ * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+ * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
+ * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
+ * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
+ * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
+ * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
+ * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
+ * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
+ * POSSIBILITY OF SUCH DAMAGE.
+ */
+
+#ifndef __REDIS_H
+#define __REDIS_H
+
+#include "fmacros.h"
+#include "config.h"
+#include "solarisfixes.h"
+#include "rio.h"
+#include "atomicvar.h"
+
+#include <stdio.h>
+#include <stdlib.h>
[…]

### 2022-01-28 attacks/redis/utils/generate-command-code.py

@@ -1,0 +1,425 @@
+#!/usr/bin/env python3
+
+import os
+import glob
+import json
+
+ARG_TYPES = {
+    "string": "ARG_TYPE_STRING",
+    "integer": "ARG_TYPE_INTEGER",
+    "double": "ARG_TYPE_DOUBLE",
+    "key": "ARG_TYPE_KEY",
+    "pattern": "ARG_TYPE_PATTERN",
+    "unix-time": "ARG_TYPE_UNIX_TIME",
+    "pure-token": "ARG_TYPE_PURE_TOKEN",
+    "oneof": "ARG_TYPE_ONEOF",
+    "block": "ARG_TYPE_BLOCK",
+}
+
+GROUPS = {
+    "generic": "COMMAND_GROUP_GENERIC",
+    "string": "COMMAND_GROUP_STRING",
+    "list": "COMMAND_GROUP_LIST",
+    "set": "COMMAND_GROUP_SET",
+    "sorted_set": "COMMAND_GROUP_SORTED_SET",
+    "hash": "COMMAND_GROUP_HASH",
+    "pubsub": "COMMAND_GROUP_PUBSUB",
+    "transactions": "COMMAND_GROUP_TRANSACTIONS",
+    "connection": "COMMAND_GROUP_CONNECTION",
+    "server": "COMMAND_GROUP_SERVER",
+    "scripting": "COMMAND_GROUP_SCRIPTING",
+    "hyperloglog": "COMMAND_GROUP_HYPERLOGLOG",
+    "cluster": "COMMAND_GROUP_CLUSTER",
+    "sentinel": "COMMAND_GROUP_SENTINEL",
+    "geo": "COMMAND_GROUP_GEO",
+    "stream": "COMMAND_GROUP_STREAM",
+    "bitmap": "COMMAND_GROUP_BITMAP",
+}
+
+RESP2_TYPES = {
+    "simple-string": "RESP2_SIMPLE_STRING",
+    "error": "RESP2_ERROR",
+    "integer": "RESP2_INTEGER",
+    "bulk-string": "RESP2_BULK_STRING",
+    "null-bulk-string": "RESP2_NULL_BULK_STRING",
+    "array": "RESP2_ARRAY",
+    "null-array": "RESP2_NULL_ARRAY",
+}
+
+RESP3_TYPES = {
+    "simple-string": "RESP3_SIMPLE_STRING",
+    "error": "RESP3_ERROR",
[…]

### 2022-01-28 docs/HLL_exp_bad/main.tex
update paper

@@
 
-As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from the universe of possible strings, the adversary has to choose what to insert from a set of 250,000 random items that are given to it.
+As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from ${\mathcal{X}}$, the adversary has to choose what to insert from a subset of 250,000 random items that are given to it.
+
 Finally, for completeness, we also run the attacks against HLL sketches filled with adversarial inputs only, thus that are initially empty.
@@
 For simplicity and to keep the attack as striking as possible, we set $B$ to 1 in the attacks on an initially empty HLL sketch in scenarios S2 and S4.
-All attacks were run 30 times and we report average outcomes.
+All attacks were run 50 times and we report average outcomes.
 Our results are reported in Tables~\ref{table:tab1} and~\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary, and the cardinality estimate after adding the adversary's items.
@@
 First, when attacking a sketch already containing 1000 random items, we can see that the adversary is on average able to insert about half of the $2^{20}$ items in our starting set in scenario S2, and about 83\% in scenario S4.
-The cardinality estimate is, on average, increased slightly in S2 (from 1,000 to 1,011) and remains at the same starting value (1,005 on average) in S4.
-In both cases, we inserted more than half a million items whilst increasing the cardinality estimate by about $0.1\%$!
+The cardinality estimate is, on average, increased slightly in S2 (from 1,001 to 1,025) and remains at the same starting value (1,007 on average) in S4.
+In both cases, we inserted more than half a million items whilst increasing the cardinality estimate by about $0.2\%$!
 These results align well with our theoretical analysis.
@@
 Second, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly a rounded version of the Hit Counting estimator $m \cdot \ln (m/(m-B))$ after inserting a fraction $\frac{B}{m}$ of the items, as predicted by our analysis.
-For example, in scenario S1 with $B=100$, the final cardinality estimate is 126 (in agreement with the formula) while with $B=1$ in scenarios S2, S4, the final cardinality estimate is just 1 (again, as predicted by the formula).
+For example, in scenario S1 with $B=128$, the final cardinality estimate is 177 (in agreement with the formula) while with $B=1$ in scenarios S2, S4, the final cardinality estimate is just 1 (again, as predicted by the formula).
 
@@
 This makes it difficult to make a precise comparison with the results of~\cite{hllvuln}.
-Nevertheless, the S1 adversary following our strategy against the classical HLL algorithm is able to add 977 and 97,433 distinct items while holding the cardinality estimate at 1 and 126, respectively.
+Nevertheless, the S1 adversary following our strategy against the classical HLL algorithm is able to add 977 and 125,091 distinct items while holding the cardinality estimate at 1 and 177, respectively.
 Compared to~\cite{hllvuln}, we can reduce the cardinality estimate by a factor of almost 1,000 instead of 5.
@@
 For simplicity, we set attack parameter $\epsilon$ to $\frac{1}{2}$ in the attack on an initially empty HLL sketch.
-The results are averaged over 30 iterations and reported in Table~\ref{table:tab3}.
+The results are averaged over 50 iterations and reported in Table~\ref{table:tab3}.
 Similarly to the case for classic HLL, our empirical results match with our theoretical analysis.
@@
 \small\centering
-\caption{Attack results for HLL in Redis, averaged over 30 iterations.}
+\caption{Attack results for HLL in Redis, averaged over 50 iterations.}
 \medskip
@@
     \textbf{Scenario} &  \multicolumn{2}{c|}{S2} \\ \hline
-    \textbf{$\epsilon$} & n/a & 0.5 \\ \hline
-    \textbf{HLL initially empty?} & no & yes \\ \hline
-    \textbf{Original Card. Est.} & 70,140 & 0 \\ \hline
-    \textbf{\# Items added} & 524,455 & 524,275 \\ \hline
-    \textbf{Final Card. Est.} & 73,249 & 13,188 \\ \hline
+    %\textbf{HLL initially empty?} & no & yes \\ \hline
+    \textbf{Original Card. Est.} & 0 & 139,967 \\ \hline
+     \textbf{$\epsilon$} &  0.5 & - \\ \hline
+    \textbf{\# Items added} & 524,210 & 524,454\\ \hline
+    \textbf{Final Card. Est.} & 13,185 & 139,967 \\ \hline
 \end{tabular}}

### 2022-01-31 docs/HLL_exp_bad/main.tex
redo tables

@@
 
+\begin{table*}[htp]
+\caption{Attacks on empty sketches for a classic HLL implementation, averaged over 50 iterations.}
+\begin{center}
+\begin{tabular}{| l | >{\raggedleft\arraybackslash}p{5em} | >{\raggedleft\arraybackslash}p{5em} | >{\raggedleft\arraybackslash}p{5em} | >{\raggedleft\arraybackslash}p{5em} |}
+	\hline
+	\textbf{Scenario} &  \multicolumn{2}{c|}{S1} & S2 & S4  \\ \hline
+	\textbf{Original Card. Est.} & 0 & 0 &  0  & 0   \\ \hline
+	\textbf{\# Targeted buckets $B$} & 1 & 128 (= m/2) & 1 &  1 \\ \hline
+	\textbf{\# Items added} & 4,132 & 523,791  & 4,106 & 4,106 \\ \hline
+	\textbf{Final Card. Est.} & 1 & 177 & 1  & 1 \\ \hline
+	\textbf{\# Resets} & 4,718 & 6,973 & - & - \\ \hline
+\end{tabular}
+\end{center}
+\label{table:tab1}
+\end{table*}%
+
+%\begin{table*}[tb!]
+%\small\centering
+%\caption{Original Table: Attacks on empty sketches for a classic HLL implementation, averaged over 50 iterations.}
+%\medskip
+%     \makebox[\linewidth]{
+%\begin{tabular}{| m{11em} | m{3.4em} | m{4.8em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
+%    \hline
+%    \textbf{Scenario} &  \multicolumn{2}{c|}{S1} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
+%    %\textbf{HLL initially empty?} & yes & yes & yes & no & yes & no \\ \hline
+%    \textbf{Original Card. Est.} & 0 & 0 &  0 & 1,001 & 0 & 1,007  \\ \hline
+%    \textbf{\# Targeted buckets $B$} & 1 & 128 (= m/2) & 1 & - & 1 & - \\ \hline
+%      \textbf{\# Items added} & 4,132 & 523,791  & 4,106 & 524,851 & 4,106 & 857,150 \\ \hline
+%     \textbf{Final Card. Est.} & 1 & 177 & 1 & 1,025  & 1 &  1,007 \\ \hline
+%    \textbf{\# Resets} & 4,718 & 6,973 & - & - & - & - \\ \hline
+%
+%\end{tabular}}
+%\label{table:tab1bis}
+%\end{table*}
+
 \begin{table*}[tb!]
-\small\centering
-\caption{Attack results for a classic HLL implementation, averaged over 50 iterations.}
-\medskip
-     \makebox[\linewidth]{
-\begin{tabular}{| m{11em} | m{3.4em} | m{4.8em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
+\caption{Comparison of attack results in the S1 adversarial scenario, in the setting of~\cite{hllvuln}.}
+\begin{center}
+\begin{tabular}{| l | >{\raggedleft\arraybackslash}p{6.8em} | >{\raggedleft\arraybackslash}p{6.8em} | >{\raggedleft\arraybackslash}p{6.8em} |}
     \hline
-    \textbf{Scenario} &  \multicolumn{2}{c|}{S1} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
-    %\textbf{HLL initially empty?} & yes & yes & yes & no & yes & no \\ \hline
-    \textbf{Original Card. Est.} & 0 & 0 &  0 & 1,001 & 0 & 1,007  \\ \hline
-    \textbf{\# Targeted buckets $B$} & 1 & 128 (= m/2) & 1 & - & 1 & - \\ \hline
-      \textbf{\# Items added} & 4,132 & 523,791  & 4,106 & 524,851 & 4,106 & 857,150 \\ \hline
-     \textbf{Final Card. Est.} & 1 & 177 & 1 & 1,025  & 1 &  1,007 \\ \hline
-    \textbf{\# Resets} & 4,718 & 6,973 & - & - & - & - \\ \hline
-
-\end{tabular}}
-\label{table:tab1}
-\end{table*}
-
-\begin{table*}[tb!]
-\small\centering
-\caption{Attack results for a classic HLL implementation, averaged over 50 iterations, in the setting of~\cite{hllvuln}.}
-\medskip
-     \makebox[\linewidth]{
-\begin{tabular}{| m{11em} | m{3.4em} | m{4.8em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
-    \hline
-    \textbf{Scenario} & \multicolumn{2}{c|}{S1} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
-    %\textbf{HLL initially empty?} & yes & yes & yes & no & yes & no \\ \hline
-      \textbf{Original Card. Est.} & 0 & 0 & 0 & 1,001 & 0 & 1,003 \\ \hline
-       \textbf{\# Targeted buckets $B$} & 1 & 128 (= m/2) & 1 & - & 1 & - \\ \hline
-     \textbf{\# Items added} & 977 & 125,091 & 980 & 125,153 & 1,003 & 204,127 \\ \hline
-    \textbf{Final Card. Est.} & 1 & 177 & 1 & 1,058  & 1 & 1,003 \\ \hline
-    \textbf{\# Resets} & 1,148 & 1,655 & - & - & - & - \\ \hline
-
-\end{tabular}}
+    \textbf{Attack} & \multicolumn{2}{c|}{Ours} & ~\cite{hllvuln} \\ \hline
+      \textbf{Original Card. Est.} & 0 & 0 & 0  \\ \hline
+       \textbf{\# Targeted buckets $B$} & 1 & 128 (= m/2) & - \\ \hline
+     \textbf{\# Items added} & 977 & 125,091 & 74,390 \\ \hline
+    \textbf{Final Card. Est.} & 1 & 177 &  15,780 \\ \hline
+    \textbf{\# Resets} & 1,148 & 1,655 & 3 \\ \hline
+
+\end{tabular}
+\end{center}
 \label{table:tab2}
 \end{table*}
+
+\begin{figure*}[htp]
+    \centering
+    \begin{minipage}{0.45\textwidth}
+        \centering
+        \caption{Attack results as function of t in the S2 scenario.}
+        \includegraphics[width=0.9\textwidth]{figures/graph_S2.pdf}
+    \end{minipage}\hfill
+    \begin{minipage}{0.45\textwidth}
+        \centering
+        \caption{Attack results as function of t in the S4 scenario.}
+        \includegraphics[width=0.9\textwidth]{figures/graph_S4.pdf} % second figure itself
+     \end{minipage}
+    \label{fig:graphs}
+\end{figure*}
+
+%\begin{table*}[tb!]
+%\small\centering
+%\caption{Original Table: Attack results for a classic HLL implementation, averaged over 50 iterations, in the setting of~\cite{hllvuln}.}
+%\medskip
+%     \makebox[\linewidth]{
+%\begin{tabular}{| m{11em} | m{3.4em} | m{4.8em} | m{4.1em} | m{4.1em} | m{4.1em} | m{4em} |}
+%    \hline
+%    \textbf{Scenario} & \multicolumn{2}{c|}{S1} & \multicolumn{2}{c|}{S2} & \multicolumn{2}{c|}{S4} \\ \hline
+%    %\textbf{HLL initially empty?} & yes & yes & yes & no & yes & no \\ \hline
+%      \textbf{Original Card. Est.} & 0 & 0 & 0 & 1,001 & 0 & 1,003 \\ \hline
+%       \textbf{\# Targeted buckets $B$} & 1 & 128 (= m/2) & 1 & - & 1 & - \\ \hline
+%     \textbf{\# Items added} & 977 & 125,091 & 980 & 125,153 & 1,003 & 204,127 \\ \hline
+%    \textbf{Final Card. Est.} & 1 & 177 & 1 & 1,058  & 1 & 1,003 \\ \hline
+%    \textbf{\# Resets} & 1,148 & 1,655 & - & - & - & - \\ \hline
+%
+%\end{tabular}}
+%\label{table:tab2bis}
+%\end{table*}
 
@@
 \begin{table*}[tb!]
-\small\centering
+\begin{center}
 \caption{Attack results for HLL in Redis, averaged over 50 iterations.}
-\medskip
-     \makebox[\linewidth]{
 \begin{tabular}{| m{11em} | m{4em} | m{4.2em} |}
@@
     \textbf{Final Card. Est.} & 13,185 & 139,967 \\ \hline
-\end{tabular}}
+\end{tabular}
+\end{center}
 \label{table:tab3}

### 2022-01-31 attacks/attack_hll.go

@@ -22,7 +22,8 @@ func main() {
 	fmt.Printf("(original cardinality estimation, final cardinality estimation, number of items inserted by adversary) for %d iterations:\n", iterations)
 
 	//S1 scenario
-	var originalEst, finalEst, nItermInserted, nResets int
+	var originalEst, finalEst, nItermInserted int
+	/**
 
 	originalEst, finalEst, nItermInserted, nResets = runAttack("S1", 0, false, 1)
 	fmt.Printf("Results in the S1 scenario: %d, %d, %d, resets: %d\n", originalEst, finalEst, nItermInserted, nResets)
@@ -51,6 +52,27 @@ func main() {
 
 	originalEst, finalEst, nItermInserted, _ = runAttack("S4", 1000, false, 1)
 	fmt.Printf("Results in the S4 scenario with 1000 initial items: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+	**/
+
+	originalEst, finalEst, nItermInserted, _ = runAttack("S2", 1500, false, 1)
+	fmt.Printf("Results in the S2 scenario with 1500 initial items: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+
+	originalEst, finalEst, nItermInserted, _ = runAttack("S2", 10000, false, 1)
+	fmt.Printf("Results in the S2 scenario with 10000 initial items: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+
+	originalEst, finalEst, nItermInserted, _ = runAttack("S2", 15000, false, 1)
+	fmt.Printf("Results in the S2 scenario with 15000 initial items: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+
+	originalEst, finalEst, nItermInserted, _ = runAttack("S4", 1500, false, 1)
+	fmt.Printf("Results in the S4 scenario with 1500 initial items: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+
+	originalEst, finalEst, nItermInserted, _ = runAttack("S4", 10000, false, 1)
+	fmt.Printf("Results in the S4 scenario with 10000 initial items: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+
+	originalEst, finalEst, nItermInserted, _ = runAttack("S4", 15000, false, 1)
[…]

### 2022-02-05 docs/HLL_exp_bad/main.tex
align c title
adding a cdot

@@
 
-With $2^{t-1}m \ln m$ honest insertions, it is also highly likely that the small range correction (Hit Counting algorithm) will not be used, and the target HLL's cardinality estimate will be computed using the raw estimate.
+With $2^{t-1}\cdot m \ln m$ honest insertions, it is also highly likely that the small range correction (Hit Counting algorithm) will not be used, and the target HLL's cardinality estimate will be computed using the raw estimate.
 %\[
@@
 	\hline
-	\textbf{Scenario} &  \multicolumn{2}{c|}{S1} & S2 & S4  \\ \hline
+	\textbf{Scenario} &  \multicolumn{2}{c|}{S1} & \multicolumn{1}{c|}{S2} & \multicolumn{1}{c|}{S4}  \\ \hline
 	\textbf{Original Card. Est.} & 0 & 0 &  0  & 0   \\ \hline
@@
     \hline
-    \textbf{Attack} & \multicolumn{2}{c|}{Ours} & ~\cite{hllvuln} \\ \hline
+    \textbf{Attack} & \multicolumn{2}{c|}{Ours} &\multicolumn{1}{c|}{~\cite{hllvuln}} \\ \hline
       \textbf{Original Card. Est.} & 0 & 0 & 0  \\ \hline

### 2022-02-05 attacks/attack_hll.go
new graphs

@@ -23,30 +23,23 @@ var empty = 0
 
 func main() {
 
-	fmt.Printf("(original cardinality estimation, final cardinality estimation, number of items inserted by adversary) for %d iterations:\n", iterations)
+	fmt.Printf("For %d iterations:\n", iterations)
 
 	//S1 scenario
-	var originalEst, finalEst, nItermInserted, expected, nResets int
+
+	fmt.Printf("S1 Scenario: (original, final, number of items inserted, number of resets)\n")
+	var originalEst, finalEst, nItermInserted, nResets, expected int
 
 	originalEst, finalEst, nItermInserted, nResets = runAttack("S1", empty, false, defaultB, defaultT)
-	fmt.Printf("Results in the S1 scenario: %d, %d, %d, resets: %d\n", originalEst, finalEst, nItermInserted, nResets)
+	fmt.Printf("Empty sketch, B = 1: %d, %d, %d, resets: %d\n", originalEst, finalEst, nItermInserted, nResets)
 
 	originalEst, finalEst, nItermInserted, nResets = runAttack("S1", empty, false, uint32(m/2), defaultT)
-	fmt.Printf("Results in the S1 scenario with B = %d: %d, %d, %d, resets: %d\n", uint32(m/2), originalEst, finalEst, nItermInserted, nResets)
+	fmt.Printf("Empty sketch, B = %d: %d, %d, %d, resets: %d\n", uint32(m/2), originalEst, finalEst, nItermInserted, nResets)
 
 	//S2 scenario
-
+	fmt.Printf("S2 Scenario: (original, final, number of items inserted)\n")
 	originalEst, finalEst, nItermInserted, _ = runAttack("S2", empty, false, defaultB, defaultT)
-	fmt.Printf("Results in the S2 scenario: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
-
-	originalEst, finalEst, nItermInserted, _ = runAttack("S2", 500, false, defaultB, defaultT)
-	fmt.Printf("Results in the S2 scenario with 500 initial items: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
-
-	originalEst, finalEst, nItermInserted, _ = runAttack("S2", 1000, false, defaultB, defaultT)
-	fmt.Printf("Results in the S2 scenario with 1000 initial items: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
-
[…]

### 2022-02-05 docs/HLL_exp_bad/main.tex

@@
 	\textbf{Original Card. Est.} & 0 & 0 &  0  & 0   \\ \hline
-	\textbf{\# Targeted buckets $B$} & 1 & 128 (= m/2) & 1 &  1 \\ \hline
+	\textbf{\# Targeted buckets $B$} & 1 & 128 ($m/2$) & 1 &  1 \\ \hline
 	\textbf{\# Items added} & 4,132 & 523,791  & 4,106 & 4,106 \\ \hline
@@
       \textbf{Original Card. Est.} & 0 & 0 & 0  \\ \hline
-       \textbf{\# Targeted buckets $B$} & 1 & 128 (= m/2) & - \\ \hline
+       \textbf{\# Targeted buckets $B$} & 1 & 128 ($m/2$) & - \\ \hline
      \textbf{\# Items added} & 977 & 125,091 & 74,390 \\ \hline
@@
         \centering
-        \caption{Attack results as function of t in the S2 scenario.}
+        \caption{Attack results as function of t in scenario S2 with initially non-empty HLL sketch.}
         \includegraphics[width=0.9\textwidth]{figures/graph_S2.pdf}
@@
         \centering
-        \caption{Attack results as function of t in the S4 scenario.}
+        \caption{Attack results as function of t in scenario S4 with initially non-empty HLL sketch.}
         \includegraphics[width=0.9\textwidth]{figures/graph_S4.pdf} % second figure itself

### 2022-02-05 docs/HLL_exp_bad/main.tex
update results text

@@
 %(note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
-We create an HLL sketch using $n=8$ and hence $m=256$ buckets and, for attacks in the setting of an initially non-empty target HLL sketch, initialise it with 1,000 random items representing honest users' data.
+We create an HLL sketch using $n=8$ and hence $m=256$ buckets and, for attacks in the setting of an initially non-empty target HLL sketch, initialise it with $t \cdot m$ log $m$ random items representing honest users' data.
 
@@
 
-As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from ${\mathcal{X}}$, the adversary has to choose what to insert from a subset of 250,000 random items that are given to it.
+As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mount our attack in their own setting: instead of picking from ${\mathcal{X}}$, the adversary has to choose what to insert from a subset of 250,000 random items that are given to it.
 
@@
 
-First, when attacking a sketch already containing 1000 random items, we can see that the adversary is on average able to insert about half of the $2^{20}$ items in our starting set in scenario S2, and about 83\% in scenario S4.
-The cardinality estimate is, on average, increased slightly in S2 (from 1,001 to 1,025) and remains at the same starting value (1,007 on average) in S4.
-In both cases, we inserted more than half a million items whilst increasing the cardinality estimate by about $0.2\%$!
+First, when attacking a sketch already containing $t \cdot m$ log $m$ random items, we can see that the adversary is on average able to insert a fraction $1-1/2^t$ of the $2^{20}$ items in our starting set in scenario S2, and a fraction $1-H_M^{-1}$ in scenario S4.
+For example for $t = 1$, the adversary can insert half of the items in S2, and about 83\% in S4.
+The cardinality estimate is, on average, increased slightly in S2 (from 1,411 to 1,430) and remains at the same starting value (1,416 on average) in S4.
+In both cases, we inserted more than half a million items whilst increasing the cardinality estimate by about $1\%$!
 These results align well with our theoretical analysis.

### 2022-02-06 docs/HLL_exp_bad/main.tex
clean up to results section and Redis more generally

--- 2022-02-05 docs/HLL_exp_bad/main.tex
@@
 
-We assume that at least $2^{t-1}\cdot m\ln m$ items have already been honestly inserted into the target HLL, for some integer $t$ with $1 \leq t \leq \ell$ which will be a parameter of our attack.
+We assume that at least $2^{t-1} m\ln m$ items have already been honestly inserted into the target HLL, for some integer $t$ with $1 \leq t \leq \ell$ which will be a parameter of our attack.
 We assume the S2 adversary somehow knows that this condition holds (e.g.\ through knowledge of when the HLL was initialised and the average rate at which items are inserted into it).
@@
 
-Under the condition that at least $2^{t-1}\cdot m\ln m$ items have already been honestly inserted, it is highly likely that every one of the $m=2^n$ buckets has counter value $M[j]$ with $M[j] \ge t$.
-This follows from an extension of the standard analysis of the coupon collector problem~\cite[Section 3.6]{MR95} under our usual assumption that $h$ behaves as a random function.
-To see this, note that when an item $x$ is inserted into a specific bucket $j$, the condition $M[j] \ge t$ holds with probability $2^{-t+1}$; we then need about $m\ln m$ such events to occur to ensure the desired condition holds in every bucket.
-
-With $2^{t-1}\cdot m \ln m$ honest insertions, it is also highly likely that the small range correction (Hit Counting algorithm) will not be used, and the target HLL's cardinality estimate will be computed using the raw estimate.
+Under the condition that at least $2^{t-1} m\ln m$ items have already been honestly inserted, it is highly likely that every one of the $m=2^n$ buckets has counter value $M[j]$ with $M[j] \ge t$.
+This follows from the standard analysis of the coupon collector problem~\cite[Section 3.6]{MR95} under our usual assumption that $h$ behaves as a random function.
+To see this, note that when an item $x$ is inserted into a specific bucket $j$, the condition $M[j] \ge t$ holds with probability $q:=1/2^{t-1}$; so with a total of $N$ attempted insertions, we expect $Nq$ usable insertions, with a sharp concentration around this number.
+We need about $m\ln m$ usable insertions to occur to ensure we get one in every bucket; setting $Nq = m\ln m$ completes the argument.
+
+With $2^{t-1}m \ln m$ honest insertions, it is also highly likely that the small range correction (Hit Counting algorithm) will not be used, and the target HLL's cardinality estimate will be computed using the raw estimate.
 %\[
@@
 The S2 adversary knows $h$ so it selects items $x$ from ${\mathcal{X}}$ for which the $\ell$ rightmost bits of $h(x)$ start with up to $t-1$ ``0'' bits.
-In this case, since the relevant bucket $j$ already has $M[j] \ge t$, the bucket's counter is not updated, and the cardinality estimate of the HLL sketch remains unchanged.
+In this case, since the relevant bucket $j$ already has $M[j] \ge t$ (with high probability), the bucket's counter is not updated, and the cardinality estimate of the HLL sketch remains unchanged.
 
@@
 
-In scenarios S2 and S4 with an initially non-empty HLL sketch, we assume as for our earlier analysis of the standard HLL algorithm that all buckets have already been hit (this will be the case with high probability as soon as $m \log m$ items have been honestly inserted).
-Hence $C_0 = 0$ and $\sigma$ is evaluated at 0 in (\ref{eqn:ertl}), where its value is 0.
+In scenarios S2 and S4 with an initially non-empty HLL sketch, we assume that $2^{t-1}m \ln m$ items have been honestly inserted, as per our analysis in Section~\ref{sec:nonempty}.
+This ensures that, with high probability, we have $C_0 = C_1 = \cdots = C_{t-1} = 0$.
+In particular, since $C_0 = 0$, $\sigma$ is evaluated at 0 in (\ref{eqn:ertl}), where its value is 0.
 This means that the estimate $\hat{\lambda}$ is computed using a denominator of the form:
 \[
-\sum_{k=1}^{\ell}C_k 2^{-k} + m \tau(1-C_{\ell+1}/m)2^{-\ell}.
-\]
-Now we can proceed exactly as in Sections~\ref{sec:nonemptyHLLS2} and~~\ref{sec:nonemptyHLLS4}, inserting items $x$ whose hashes have rightmost $\ell$ bits beginning with a ``1'' bit (and with a more relaxed condition in S4).
-These insertions do not alter the values of the $C_k$, and hence $\hat{\lambda}$ is unaffected.
-In S2, we expect to insert half the items in ${\mathcal{X}}$ without increasing the cardinality estimate $\hat{\lambda}$ at all.
+\sum_{k=t}^{\ell}C_k 2^{-k} + m \tau(1-C_{\ell+1}/m)2^{-\ell}.
+\]
+Now we can proceed exactly as in Section~\ref{sec:nonempty}, inserting items $x$ whose hashes have rightmost $\ell$ bits beginning with up to $t-1$ ``0'' bits (and with a more relaxed condition in S4).
+With high probability, these insertions do not alter the values of the $C_k$, and hence $\hat{\lambda}$ should be unaffected.
+In S2, we expect to insert a fraction $1-1/2^t$ of the items in ${\mathcal{X}}$ without increasing the cardinality estimate $\hat{\lambda}$ at all.
 In S4, the analysis is the same as in Section~\ref{sec:nonemptyHLLS4}.
@@
 %(note that the library allows the hash function to be specified; we pick Murmur3 for our attack and consider it reasonable that the adversary should know the hash function in use).
-We create an HLL sketch using $n=8$ and hence $m=256$ buckets and, for attacks in the setting of an initially non-empty target HLL sketch, initialise it with 1,000 random items representing honest users' data.
-
-We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $2^{20}$ possible inputs while trying to keep the cardinality estimate as low as possible, in each of the scenarios S1, S2 and S4.
-
-As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mount our attack in their more restrictive setting: instead of picking from ${\mathcal{X}}$, the adversary has to choose what to insert from a subset of 250,000 random items that are given to it.
-
-Finally, for completeness, we also run the attacks against HLL sketches filled with adversarial inputs only, thus that are initially empty.
+We create an HLL sketch using $n=8$ and hence $m=256$ buckets.
+We run attacks in each of the scenarios S1, S2 and S4, with initially empty and non-empty HLL sketches.
+In the latter case, we initialise the sketch with $2^{t-1} m \ln m$ random items, where $t$ is an integer parameter that we vary in our experiments (cf.\ Section~\ref{sec:nonempty}).
+We then challenge the adversary to pick the largest possible number of items to add to the sketch from our set of  $2^{20}$ possible inputs while trying to keep the cardinality estimate as low as possible.
+As a second experiment, and in order to get results that can be compared to the work of~\cite{hllvuln}, we mount our attack in their chosen setting for scenario S1: instead of picking from ${\mathcal{X}}$, the adversary has to choose what to insert from a subset of 250,000 random items that are given to it.
+All attacks were run 50 times and we report average outcomes.
+
+Our results are reported in Tables~\ref{table:tab1},~\ref{table:tab2} and Figures~\ref{fig:graphS2},~\ref{fig:graphS4}.
+The tables detail our results for the case of initially empty HLL sketches while the figures show results in S2 and S4 for initially non-empty HLL sketches.
+
+%Finally, for completeness, we also run the attacks against HLL sketches filled with adversarial inputs only, thus that are initially empty.
 We vary the value of $B$ in the attack for scenario S1 to demonstrate some of the available trade-offs.
@@
 All attacks were run 50 times and we report average outcomes.
-Our results are reported in Tables~\ref{table:tab1} and~\ref{table:tab2}, which detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary, and the cardinality estimate after adding the adversary's items.
-
-These empirical results confirm our analysis on several points.
-
-First, when attacking a sketch already containing 1000 random items, we can see that the adversary is on average able to insert about half of the $2^{20}$ items in our starting set in scenario S2, and about 83\% in scenario S4.
-The cardinality estimate is, on average, increased slightly in S2 (from 1,001 to 1,025) and remains at the same starting value (1,007 on average) in S4.
-In both cases, we inserted more than half a million items whilst increasing the cardinality estimate by about $0.2\%$!
-These results align well with our theoretical analysis.
-%In scenario S1 we run the attack designed for an initially empty target HLL sketch. It performs well in practice, for example, increasing the HLL cardinality estimate from 100 to 126????
-%( $205$'$000 \approx 250'000 \times 0.82 \approx 250$'$000\times(1-\frac{1}{2\times H_c})$, with $H_c\approx\frac{1'000}{256}=3.9$)
-%Here $0.83 \approx (1-\frac{1}{2\cdot H_c})$ with $H_c\approx\frac{1000}{256}=3.9$, so the results are also consistent with our theoretical analysis.
-%% That analysis was now updated -- does it still match?
-Second, when attacking an empty HLL sketch, we can see that in all scenarios, the final cardinality estimate is always exactly a rounded version of the Hit Counting estimator $m \cdot \ln (m/(m-B))$ after inserting a fraction $\frac{B}{m}$ of the items, as predicted by our analysis.
+Our results are reported in Tables~\ref{table:tab1} and~\ref{table:tab2}.
+These tables detail the estimated cardinality at the beginning of the attack, the number of items added by the adversary, and the cardinality estimate after adding the adversary's items.
+
+These experimental results broadly confirm our analyses.
+First, when attacking an empty HLL sketch (cf.\ Table~\ref{table:tab1}), we can see that in all scenarios, the final cardinality estimate is always exactly a rounded version of the Hit Counting estimator $m \cdot \ln (m/(m-B))$ after inserting a fraction $\frac{B}{m}$ of the items, as predicted by our analysis.
 For example, in scenario S1 with $B=128$, the final cardinality estimate is 177 (in agreement with the formula) while with $B=1$ in scenarios S2, S4, the final cardinality estimate is just 1 (again, as predicted by the formula).
-
-We can also compare our results from Table~\ref{table:tab2} with results from the attack presented in~\cite[Section 5.1]{hllvuln}.
+Moreover, the number of items we can insert on average aligns well with our theoretical predictions (e.g.\ $4106$ in S2 and S4, compared with a prediction of $2^{20}/2^8 = 4096$).
+
+Second, Figures~\ref{fig:graphS2},~\ref{fig:graphS4} show very close agreement between the predictions made by our analysis in Section~\ref{sec:nonempty} and our experimental results for the situation where the HLL sketch is initially non-empty.
+The only significant difference is in scenario S2, where there is a small increase in the cardinality estimate, around 1-2\% on average, as a result of the attack, whereas the theoretical analysis suggests there should be no increase.
+The reason for this is that the insertion of the initial random $2^{t-1} m\ln m$ items is not guaranteed to produce a situation in which every bucket contains a value $t$ or greater -- this condition holds only with high probability.
+
+We can also compare with the attack presented in~\cite[Section 5.1]{hllvuln}.
 In that paper, the authors target the Redis implementation of HLL and perform the attack against an empty sketch filled with adversarial inputs only.
@@
 This makes it difficult to make a precise comparison with the results of~\cite{hllvuln}.
-Nevertheless, the S1 adversary following our strategy against the classical HLL algorithm is able to add 977 and 125,091 distinct items while holding the cardinality estimate at 1 and 177, respectively.
+Nevertheless, the S1 adversary following our strategy against the classical HLL algorithm is able to add 977 and 125,091 distinct items while holding the cardinality estimate at 1 and 177, respectively (see Table~\ref{table:tab2}).
 Compared to~\cite{hllvuln}, we can reduce the cardinality estimate by a factor of almost 1,000 instead of 5.
-This comes at the cost of performing more resets compared to~\cite{hllvuln}.
+This comes at the cost of performing significantly more resets compared to~\cite{hllvuln} (more than 1,000 compared to 3).
 We recall that in practice each reset may cause a delay (for example, it may involve rebooting a network device); we also recall that we consider S1 to be a somewhat artificial attack scenario.
@@
         \centering
-        \caption{Attack results as function of t in scenario S2 with initially non-empty HLL sketch.}
+        \caption{Attack results as function of t in scenario S2 with initially non-empty HLL sketch.}\label{fig:graphS2}
         \includegraphics[width=0.9\textwidth]{figures/graph_S2.pdf}
@@
         \centering
-        \caption{Attack results as function of t in scenario S4 with initially non-empty HLL sketch.}
+        \caption{Attack results as function of t in scenario S4 with initially non-empty HLL sketch.}\label{fig:graphS4}
         \includegraphics[width=0.9\textwidth]{figures/graph_S4.pdf} % second figure itself
      \end{minipage}
-    \label{fig:graphs}
 \end{figure*}
@@
 The results are averaged over 50 iterations and reported in Table~\ref{table:tab3}.
-Similarly to the case for classic HLL, our empirical results match with our theoretical analysis.
+Similarly to the case for classic HLL, our empirical results match closely with our theoretical analysis.
 
@@
 \begin{center}
-\caption{Attack results for HLL in Redis, averaged over 50 iterations.}
+\caption{Attack results for HLL in Redis, averaged over 50 iterations.}\label{table:tab3}
 \begin{tabular}{| m{11em} | m{4em} | m{4.2em} |}
@@
 \end{center}
-\label{table:tab3}
 \end{table*}
@@
 
-The authors thank Dennis Hofheinz and Anupama Unnikrishnan for useful discussions on simulation-based definitions, Tom Shrimpton for suggesting to look at GoogleBigQuery, and Eylon Yogev for drawing our attention to related work on streaming algorithms with adversarial inputs.
+The authors thank Dennis Hofheinz and Anupama Unnikrishnan for useful discussions on simulation-based definitions, Tom Shrimpton for suggesting to look at GoogleBigQuery, Eylon Yogev for drawing our attention to related work on streaming algorithms with adversarial inputs, and Angelika Steger for help with balls-into-bins arguments.
 \fi

### 2022-02-06 docs/HLL_exp_bad/main.tex
???

@@
 Such an adversary is quite realistic.
-For example, both the Redis and Presto implementations of HLL use a fixed hash function and parameters that can be read off from source code.\footnote{See \url{https://github.com/redis/redis/blob/unstable/src/hyperloglog.c\#L396}, \url{
-https://github.com/redis/redis/blob/unstable/src/hyperloglog.c\#L466} for Redis and 
+For example, both the Redis and Presto implementations of HLL use a fixed hash function and parameters that can be read off from source code.\footnote{See \url{https://github.com/redis/redis/blob/unstable/src/hyperloglog.c\#L397}, \url{
+https://github.com/redis/redis/blob/unstable/src/hyperloglog.c\#L467} for Redis and 
 \url{https://github.com/prestodb/presto/blob/2ad67dcf000be86ebc5ff7732bbb9994c8e324a8/presto-main/src/main/java/com/facebook/presto/type/khyperloglog/KHyperLogLog.java\#L20}

### 2022-02-07 docs/HLL_exp_bad/main.tex
right aligning table 4 and updating URL for presto

@@
 https://github.com/redis/redis/blob/unstable/src/hyperloglog.c\#L467} for Redis and 
-\url{https://github.com/prestodb/presto/blob/2ad67dcf000be86ebc5ff7732bbb9994c8e324a8/presto-main/src/main/java/com/facebook/presto/type/khyperloglog/KHyperLogLog.java\#L20}
+\url{https://github.com/airlift/airlift/blob/master/stats/src/main/java/io/airlift/stats/cardinality/HyperLogLog.java\#L58}
 for Presto.}
@@
 \caption{Attack results for HLL in Redis, averaged over 50 iterations.}\label{table:tab3}
-\begin{tabular}{| m{11em} | m{4em} | m{4.2em} |}
+%\begin{tabular}{| m{11em} | m{4em} | m{4.2em} |}
+\begin{tabular}{| c | r | r |}
     \hline

### 2022-02-07 docs/HLL_exp_bad/main.tex
compressing table 4
better explanation of anomaly in S2 attack, Fig 3

@@
 The reason for this is that the insertion of the initial random $2^{t-1} m\ln m$ items is not guaranteed to produce a situation in which every bucket contains a value $t$ or greater -- this condition holds only with high probability.
+The result is that some bucket contents are \emph{increased} to $t$ as a result of insertions during the course of the attack, leading to a slightly increased cardinality estimate.
 
@@
 \begin{center}
-\caption{Attack results for HLL in Redis, averaged over 50 iterations.}\label{table:tab3}
+\caption{Attack results for HLL in Redis, scenario S2, averaged over 50 iterations.}\label{table:tab3}
 %\begin{tabular}{| m{11em} | m{4em} | m{4.2em} |}
@@
     \hline
-    \textbf{Scenario} &  \multicolumn{2}{c|}{S2} \\ \hline
-    %\textbf{HLL initially empty?} & no & yes \\ \hline
     \textbf{Original Card. Est.} & 0 & 139,967 \\ \hline

### 2022-02-08 docs/HLL_exp_bad/main.tex
reduce eq (1) size
fix ln/log

@@
 \textbf{if} $E > \frac{1}{30}2^{32}$ \textbf{then}\\
-    \hspace{2em}  \textbf{set} $E^*:= -2^{32}$ln$(1 - E/2^{32})$; \hfill \textit{\{large range correction\}}\\
+    \hspace{2em}  \textbf{set} $E^*:= -2^{32}\ln(1 - E/2^{32})$; \hfill \textit{\{large range correction\}}\\
 \textbf{return} \textit{cardinality estimate $E^*$ with typical relative error} $\pm1.04/\sqrt{m}$.
@@
 So to roughly estimate the cardinality of a stream, we need only to store a representation of $k$, the length of the largest observed string of leading zero-bits for hashes of items in the stream.
-This provides a very compact mechanism for estimating a stream's cardinality -- if the true cardinality is $N$, then just $\mathcal{O}(\log\log(N))$ bits of storage are needed to compute the cardinality estimate.
+This provides a very compact mechanism for estimating a stream's cardinality -- if the true cardinality is $N$, then just $\mathcal{O}(\log_2\log_2(N))$ bits of storage are needed to compute the cardinality estimate.
 
@@
 \begin{equation}\label{eqn:ertl}
-\hat{\lambda}  = \frac{\alpha_\infty m^2}{m\sigma(C_0/m) + \sum_{k=1}^{\ell}C_k 2^{-k} + m \tau(1-C_{\ell+1}/m)2^{-\ell}} \, .
+\resizebox{.9\hsize}{!}{$
+\hat{\lambda}  = \frac{\alpha_\infty m^2}{m\sigma(C_0/m) + \sum_{k=1}^{\ell}C_k 2^{-k} + m \tau(1-C_{\ell+1}/m)2^{-\ell}} \, .$}
 \end{equation}

### 2022-02-14 docs/HLL_exp_bad/main.tex
redis graph

@@
 \centering
+\caption{The HyperLogLog Algorithm from~\cite{hll}}
      \makebox[\linewidth]{
@@
 }}
-\caption{The HyperLogLog Algorithm from~\cite{hll}}
+
 \label{fig:hll}
@@
 We assume the S2 adversary somehow knows that this condition holds (e.g.\ through knowledge of when the HLL was initialised and the average rate at which items are inserted into it).
+If it were to be the case that our assumption does not hold and the adversary cannot estimate the number of items in the sketch, they can use a downward estimate such as $t=1$.
+As this lower-bound on $t$ is highly likely to happen, the adversary is taking only very small risk.
 Notice that an S3 adversary can use its ability to get cardinality estimates to obtain this knowledge; this is one situation where an S3 adversary can make good use of the extra information available to it as compared to an S2 adversary.
@@
 \begin{table*}[tb!]
-\caption{Comparison of attack results in the S1 adversarial scenario, in the setting of~\cite{hllvuln}.}
+\caption{Comparison of attack results for a classic HLL implementation in the S1 adversarial scenario, \\ in the setting of~\cite{hllvuln}.}
 \begin{center}
@@
         \includegraphics[width=0.9\textwidth]{figures/graph_S4.pdf} % second figure itself
+     \end{minipage}
+
+     \begin{minipage}{0.45\textwidth}
+        \centering
+        \caption{Attack results as function of t for HLL in Redis (scenario S2) with initially non-empty HLL sketch.}\label{fig:graphRedis}
+        \includegraphics[width=0.9\textwidth]{figures/graph_redis.pdf} % second figure itself
      \end{minipage}
@@
 Thus, we only consider scenario S2 for Redis.
-For simplicity, we set attack parameter $\epsilon$ to $\frac{1}{2}$ in the attack on an initially empty HLL sketch.
-The results are averaged over 50 iterations and reported in Table~\ref{table:tab3}.
+We set attack parameter $\epsilon$ to $\frac{1}{2}$ and then $\frac{1}{4}$ in the attack on an initially empty HLL sketch.
+We recall that this will lead to the number of empty buckets $C_0$ to be equal to $\frac{m}{2} (=8,192)$ and $\frac{m}{4} (=4,096)$ respectively.
+The results are averaged over 50 iterations and reported in Table~\ref{table:tab3} and Figure~\ref{fig:graphRedis}.
 Similarly to the case for classic HLL, our empirical results match closely with our theoretical analysis.
@@
 \begin{center}
-\caption{Attack results for HLL in Redis, scenario S2, averaged over 50 iterations.}\label{table:tab3}
+\caption{Attack results on initially empty sketches for HLL in Redis (scenario S2), \\ averaged over 50 iterations.}\label{table:tab3}
 %\begin{tabular}{| m{11em} | m{4em} | m{4.2em} |}
-\begin{tabular}{| c | r | r |}
+\begin{tabular}{| l | r | r |}
     \hline
-    \textbf{Original Card. Est.} & 0 & 139,967 \\ \hline
-     \textbf{$\epsilon$} &  0.5 & - \\ \hline
-    \textbf{\# Items added} & 524,210 & 524,454\\ \hline
-    \textbf{Final Card. Est.} & 13,185 & 139,967 \\ \hline
+    \textbf{Original Card. Est.} & 0 & 0 \\ \hline
+     \textbf{$\epsilon$} &  0.5 & 0.25  \\ \hline
+    \textbf{\# Items added} &  524,2104 & 785,920 \\ \hline
+    \textbf{Final Card. Est.} & 13,185 & 35,938\\ \hline
 \end{tabular}
@@
 %\todo[inline]{Results from Redis or GBQ if we get them.}
+
+%    \textbf{Original Card. Est.} & 0 & 139,967 \\ \hline
+%     \textbf{$\epsilon$} &  0.5 & - \\ \hline
+%    \textbf{\# Items added} & 524,210 & 524,454\\ \hline
+%    \textbf{Final Card. Est.} & 13,185 & 139,967 \\ \hline
 

### 2022-02-14 attacks/redis/src/hyperloglog.c

@@ -204,9 +204,14 @@ struct hllhdr {
 #define HLL_SPARSE 1 /* Sparse encoding. */
 #define HLL_RAW 255 /* Only used internally, never exposed. */
 #define HLL_MAX_ENCODING 1
-#define RAND_ITEMS 150000 /*0*/
-#define B 1<<(HLL_P-1)
+
 #define MAX_VALUE (1<<20)-1
+
+#define EMPTY true
+#define EPSILON 0.25
+
+int const B = (1-EPSILON)*HLL_REGISTERS;
+int const T =1; /* in {1, 2, 3} */
 
 
 static char *invalid_hll_err = "-INVALIDOBJ Corrupted HLL object detected";
@@ -1226,11 +1231,7 @@ void pfattackCommand(client *c) {
     robj *o = lookupKeyWrite(c->db,c->argv[1]);
     struct hllhdr *hdr;
     int updated = 0;
-
     if (o == NULL) {
-        /* Create the key with a string value of the exact length to
-         * hold our HLL data structure. sdsnewlen() when NULL is passed
-         * is guaranteed to return bytes initialized to zero. */
         o = createHLLObject();
         dbAdd(c->db,c->argv[1],o);
         updated++;
@@ -1239,50 +1240,49 @@ void pfattackCommand(client *c) {
         o = dbUnshareStringValue(c->db,c->argv[1],o);
     }
 
-    /* Insert RAND_ITEMS honest initial items */
-    srand(time(NULL));
+    /* Insert number_items honest initial items */
+    if (!EMPTY) {
+      srand(time(NULL));
+      // 2^{t-1}*m*ln(m)
+      int number_items = (1<<(T-1))*158991;
+      for (int j = 0; j < number_items; j++) {
+        int r = rand() % MAX_VALUE;
+        hllAdd(o, (unsigned char*)&r, sizeof(r));
+      }
 
-    for (int j = 0; j < RAND_ITEMS; j++) {
-      int r = rand() % MAX_VALUE;
-      hllAdd(o, (unsigned char*)&r, sizeof(r));
+      /* Prints the initial cardinality (code from PFCOUNT) */
+
+      hdr = o->ptr;
+      uint64_t card;
+      if (HLL_VALID_CACHE(hdr)) {
+          card = (uint64_t)hdr->card[0];
+          card |= (uint64_t)hdr->card[1] << 8;
+          card |= (uint64_t)hdr->card[2] << 16;
[…]

### 2022-02-14 attacks/redis/src/release.h

@@ -1,3 +1,3 @@
-#define REDIS_GIT_SHA1 "b36391c5"
-#define REDIS_GIT_DIRTY "   12469"
-#define REDIS_BUILD_ID "Mathildes-MacBook-Air.local-1643356764"
+#define REDIS_GIT_SHA1 "7677981b"
+#define REDIS_GIT_DIRTY "     984"
+#define REDIS_BUILD_ID "IC-SPRING-LMC03.local-1644853386"

### 2022-02-14 attacks/former_code/draft.go
redis graph

@@ -1,97 +1,0 @@
-package main
-
-import (
-	"fmt"
-	"hash"
-	"math/rand"
-	"time"
-)
-
-//Attack under scenario S1
-//Note: h should not be given but it is nessecary to have the Add oracle for CreateBatch and CheckItem.
-
-//Attack under scenario S1
-//Note: h should not be given but it is nessecary to have the Add oracle for CreateBatch and CheckItem.
-func AttackS12(nBuckets int, h hash.Hash32, rt20 bool, all []string) []string {
-	var allItems []string
-	if rt20 {
-		allItems = CreateItems(250000)
-	} else {
-		allItems = all
-	}
-
-	itemsCand := make(map[string]bool, len(allItems))
-	for _, s := range allItems {
-		itemsCand[s] = true
-	}
-	itemsCandList := Map2List(itemsCand)
-	var itemsDisc []string
-
-	//goal := len(allItems) / 2
-	//We want to keep the 50% best
-	step := 0
-	for step < 1 {
-		fmt.Printf("Step %d\n", step)
-		step++
-		//shuffle list
-		rand.Seed(time.Now().UnixNano())
-		rand.Shuffle(len(itemsCandList), func(i, j int) { itemsCandList[i], itemsCandList[j] = itemsCandList[j], itemsCandList[i] })
-		//hll with registers filled
-		hll := FillRegisters(nBuckets, itemsCandList, h)
-		//for each item, we check if it increases the count, if it does, we rm it
-		for _, item := range itemsCandList {
-			if CheckBadItem(item, nBuckets, h, hll) {
-				delete(itemsCand, item)
-				itemsDisc = append(itemsDisc, item)
-			}
-		}
-		itemsCandList = Map2List(itemsCand)
-		//fmt.Printf("itemsCand size %d\n", len(itemsCand))
-		//fmt.Printf("disc size %d\n", len(itemsDisc))
-	}
-	return itemsCandList
-}
-
-//Attack under S1 but with less resets
-func AttackS1LessResets(nBuckets int, h hash.Hash32, rt20 bool, all []string, resetsAllowed int) []string {
-	var allItems []string
-	if rt20 {
-		allItems = CreateItems(250000)
-	} else {
-		allItems = all
-	}
-
-	itemsCand := make(map[string]bool, len(allItems))
-	for _, s := range allItems {
-		itemsCand[s] = true
-	}
[…]

### 2022-02-14 attacks/former_code/main_former.go

@@ -1,246 +1,0 @@
-package main
-
-import (
-	"flag"
-	"fmt"
-	"hash"
-
-	"./clarkduvall/hyperloglog"
-
-	"./spaolacci/murmur3"
-)
-
-var p = uint8(8)
-var nBuckets = 1 << p
-
-func main() {
-
-	//Setup
-	hll, _ := hyperloglog.New(p)
-	hash := murmur3.New32
-
-	//Parse the command line arguments
-	scenario := flag.String("scenario", "S2", "a string")
-	userData := flag.Int("userData", 0, "an int")
-	iterations := flag.Int("iterations", 1, "an int")
-	RT20 := flag.Bool("RT20", true, "a bool")
-	log := flag.Int("log", 0, "an int")
-	resets := flag.Int("resets", -1, "an int")
-	flag.Parse()
-
-	var originalEst uint64
-	var finalEst uint64
-	var itemsPicked int
-	all := AllItems()
-
-	for i := 0; i < *iterations; i++ {
-
-		//we initialize with 1000 items from a random user if flag is set
-		var userItems []string
-		userItems = CreateItems(*userData)
-		for _, i := range userItems {
-			element := hash()
-			element.Write([]byte(i))
-			hll.Add(element)
-		}
-
-		//Initial estimated cardinality
-		originalEstTemp := hll.Count()
-		if *log != 0 {
-			fmt.Printf("HLL cardinality approximation at start: %d.\n", originalEstTemp)
-		}
-
-		regBeg := make([]uint8, 256)
-		copy(regBeg, hll.Reg)
-
-		//Craft and add attacker's packets
-		var attackerItems []string
-		switch *scenario {
-		case "S3":
-			attackerItems = AttackS3(hll, nBuckets, hash(), *RT20, all)
-		case "S1":
-			if *resets == -1 {
-				attackerItems = AttackS1(nBuckets, hash(), *RT20, all)
-			} else {
-				attackerItems = AttackS1LessResets(nBuckets, hash(), *RT20, all, *resets)
-			}
-		default:
-			attackerItems = AttackS2(nBuckets, hash(), *RT20, all)
-		}
-		itemsPickedTemp := len(attackerItems)
-		for _, i := range attackerItems {
-			element := hash()
-			element.Write([]byte(i))
-			hll.Add(element)
-		}
-		if *log == 2 {
[…]

### 2022-02-14 attacks/former_code/utils.go

@@ -1,230 +1,0 @@
-package main
-
-import (
-	"hash"
-	"math"
-	"math/rand"
-	"time"
-
-	"./clarkduvall/hyperloglog"
-)
-
-//==================utils==================
-//CreateItems outputs n random strings of length 4 (52^4 > 7'000'000 possibilities)
-func CreateItems(n int) []string {
-	itemsMap := make(map[string]bool)
-	rand.Seed(time.Now().UnixNano())
-	var letters = []rune("abcdefghijklmnopqrstuvwxyzABCDEFGHIJKLMNOPQRSTUVWXYZ")
-	sLen := 4
-	for len(itemsMap) != n {
-		b := make([]rune, sLen)
-		for i := range b {
-			b[i] = letters[rand.Intn(len(letters))]
-		}
-		if !itemsMap[string(b)] {
-			itemsMap[string(b)] = true
-		}
-	}
-	var items []string
-	for i := range itemsMap {
-		items = append(items, i)
-	}
-	return items
-}
-
-//AllItems outputs all ASCII strings of length 4 (52^4 > 7'000'000 possibilities)
-func AllItems() []string {
-	var items []string
-	var letters = []rune("abcdefghijklmnopqrstuvwxyzABCDEFGHIJKLMNOPQRSTUVWXYZ")
-	sLen := 4
-	for i1 := 0; i1 < 52; i1++ {
-		for i2 := 0; i2 < 52; i2++ {
-			for i3 := 0; i3 < 52; i3++ {
-				for i4 := 0; i4 < 52; i4++ {
-					b := make([]rune, sLen)
-					b[0] = letters[i1]
-					b[1] = letters[i2]
-					b[2] = letters[i3]
-					b[3] = letters[i4]
-					items = append(items, string(b))
-				}
-			}
-		}
-	}
-	return items
-}
-
-//GenMask generates a mask to check whether a string has less then ci leading 0s
-func GenMask(ci uint8) uint32 {
-	mask := uint32(0)
-	for i := uint8(0); i < ci; i++ {
-		mask = mask << 1
-		mask += 1
-	}
-	for i := ci; i < 24; i++ {
-		mask = mask << 1
-	}
-	return mask
-}
-
-//EmptyHLL returns an emtpy HLL, it is used to simulate the reset function
-func EmptyHLL() *hyperloglog.HyperLogLog {
-	hll, _ := hyperloglog.New(p)
-	return hll
-}
-
-//Removes items of rm from items
[…]

### 2022-02-14 latex/main.tex
new struct

--- 2022-02-14 docs/HLL_exp_bad/main.tex
@@
 We assume the S2 adversary somehow knows that this condition holds (e.g.\ through knowledge of when the HLL was initialised and the average rate at which items are inserted into it).
-If it were to be the case that our assumption does not hold and the adversary cannot estimate the number of items in the sketch, they can use a downward estimate such as $t=1$.
-As this lower-bound on $t$ is highly likely to happen, the adversary is taking only very small risk.
-Notice that an S3 adversary can use its ability to get cardinality estimates to obtain this knowledge; this is one situation where an S3 adversary can make good use of the extra information available to it as compared to an S2 adversary.
+An S2 adversary who cannot estimate the number of items in the sketch can always conservatively set $t=1$ and minimise the risk that the required condition does not hold.
+Notice also that an S3 adversary can use its ability to get cardinality estimates to obtain an estimate for $t$; this is one situation where an S3 adversary can make good use of the extra information available to it as compared to an S2 adversary.
 In S4, the adversary can use its access to a snapshot to mount a more powerful attack; see Section~\ref{sec:nonemptyHLLS4}.
@@
 This follows from the standard analysis of the coupon collector problem~\cite[Section 3.6]{MR95} under our usual assumption that $h$ behaves as a random function.
-To see this, note that when an item $x$ is inserted into a specific bucket $j$, the condition $M[j] \ge t$ holds with probability $q:=1/2^{t-1}$; so with a total of $N$ attempted insertions, we expect $Nq$ usable insertions, with a sharp concentration around this number.
+To see this, note that when an item $x$ is inserted into a specific bucket $j$, the condition $M[j] \ge t$ holds with probability $q:=1/2^{t-1}$; so with a total of $N$ attempted insertions, we expect $Nq$ usable insertions, with a sharp concentration around the expected number.
 We need about $m\ln m$ usable insertions to occur to ensure we get one in every bucket; setting $Nq = m\ln m$ completes the argument.
@@
         \centering
-        \caption{Attack results as function of t in scenario S2 with initially non-empty HLL sketch.}\label{fig:graphS2}
+        \caption{Attack results as function of $t$ in scenario S2 with initially non-empty HLL sketch.}\label{fig:graphS2}
         \includegraphics[width=0.9\textwidth]{figures/graph_S2.pdf}
@@
         \centering
-        \caption{Attack results as function of t in scenario S4 with initially non-empty HLL sketch.}\label{fig:graphS4}
+        \caption{Attack results as function of $t$ in scenario S4 with initially non-empty HLL sketch.}\label{fig:graphS4}
         \includegraphics[width=0.9\textwidth]{figures/graph_S4.pdf} % second figure itself
@@
         \centering
-        \caption{Attack results as function of t for HLL in Redis (scenario S2) with initially non-empty HLL sketch.}\label{fig:graphRedis}
+        \caption{Attack results as function of $t$ for HLL in Redis (scenario S2) with initially non-empty HLL sketch.}\label{fig:graphRedis}
         \includegraphics[width=0.9\textwidth]{figures/graph_redis.pdf} % second figure itself
@@
 We recall that this will lead to the number of empty buckets $C_0$ to be equal to $\frac{m}{2} (=8,192)$ and $\frac{m}{4} (=4,096)$ respectively.
-The results are averaged over 50 iterations and reported in Table~\ref{table:tab3} and Figure~\ref{fig:graphRedis}.
-Similarly to the case for classic HLL, our empirical results match closely with our theoretical analysis.
+The results are averaged over 50 iterations and reported in Table~\ref{table:tab3}.
+Meanwhile, Figure~\ref{fig:graphRedis} shows our results for an initially non-empty sketch as a function of $t$, where $2^{t-1}m\ln m$ items are honestly inserted into the sketch.
+Similarly to the case for classic HLL, our empirical results in  Table~\ref{table:tab3} and Figure~\ref{fig:graphRedis} match closely with our theoretical analyses.
 

### 2022-02-21 latex/main.tex
bit of latex hackery to deal with IEEE appendices
more hackery for Redis eqn

@@
 Ertl~\cite{hllnew} proposed using the following cardinality estimator in place of the raw estimator in Figure~\ref{fig:hll}:
+\ifIEEEversion
 \begin{equation}\label{eqn:ertl}
@@
 \end{equation}
+\else
+\begin{equation}\label{eqn:ertl}
+\hat{\lambda}  = \frac{\alpha_\infty m^2}{m\sigma(C_0/m) + \sum_{k=1}^{\ell}C_k 2^{-k} + m \tau(1-C_{\ell+1}/m)2^{-\ell}} \, .
+\end{equation}
+\fi
+
 Here, in the denominator, $C_k$ denotes the number of buckets containing value $k$ in the HLL sketch, so $C_0$ is in particular the number of empty buckets; $\sigma(\cdot)$ and $\tau(\cdot)$ are specific functions defined as:
@@
 
+\ifIEEEversion
 \appendices
+\else
+\appendix
+\fi
 

### 2022-03-07 paper/code copy/redis/src/hyperloglog.c
Cleaning repo before going public

@@ -1,0 +1,1692 @@
+/* hyperloglog.c - Redis HyperLogLog probabilistic cardinality approximation.
+ * This file implements the algorithm and the exported Redis commands.
+ *
+ * Copyright (c) 2014, Salvatore Sanfilippo <antirez at gmail dot com>
+ * All rights reserved.
+ *
+ * Redistribution and use in source and binary forms, with or without
+ * modification, are permitted provided that the following conditions are met:
+ *
+ *   * Redistributions of source code must retain the above copyright notice,
+ *     this list of conditions and the following disclaimer.
+ *   * Redistributions in binary form must reproduce the above copyright
+ *     notice, this list of conditions and the following disclaimer in the
+ *     documentation and/or other materials provided with the distribution.
+ *   * Neither the name of Redis nor the names of its contributors may be used
+ *     to endorse or promote products derived from this software without
+ *     specific prior written permission.
+ *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+ * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+ * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
+ * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
+ * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
+ * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
+ * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
+ * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
+ * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
+ * POSSIBILITY OF SUCH DAMAGE.
+ */
+
+#include "server.h"
+
+#include <stdint.h>
+#include <math.h>
+
[…]

### 2022-03-07 paper/code/classic_hll/clarkduvall/hyperloglog/hyperloglog.go

@@ -1,0 +1,118 @@
+// Package hyperloglog implements the HyperLogLog and HyperLogLog++ cardinality
+// estimation algorithms.
+// These algorithms are used for accurately estimating the cardinality of a
+// multiset using constant memory. HyperLogLog++ has multiple improvements over
+// HyperLogLog, with a much lower error rate for smaller cardinalities.
+//
+// HyperLogLog is described here:
+// http://algo.inria.fr/flajolet/Publications/FlFuGaMe07.pdf
+//
+// HyperLogLog++ is described here:
+// http://research.google.com/pubs/pub40671.html
+package hyperloglog
+
+import (
+	"bytes"
+	"encoding/gob"
+	"errors"
+	"math"
+)
+
+const two32 = 1 << 32
+
+type HyperLogLog struct {
+	Reg []uint8
+	m   uint32
+	p   uint8
+}
+
+// New returns a new initialized HyperLogLog.
+func New(precision uint8) (*HyperLogLog, error) {
+	if precision > 16 || precision < 4 {
+		return nil, errors.New("precision must be between 4 and 16")
+	}
+
+	h := &HyperLogLog{}
+	h.p = precision
+	h.m = 1 << precision
+	h.Reg = make([]uint8, h.m)
+	return h, nil
+}
+
+// Clear sets HyperLogLog h back to its initial state.
+func (h *HyperLogLog) Clear() {
+	h.Reg = make([]uint8, h.m)
+}
+
+// Add adds a new item to HyperLogLog h.
+func (h *HyperLogLog) Add(item Hash32) {
+	x := item.Sum32()
+	i := eb32(x, 32, 32-h.p) // {x31,...,x32-p}
+	w := x<<h.p | 1<<(h.p-1) // {x32-p,...,x0}
+
+	zeroBits := clz32(w) + 1
+	if zeroBits > h.Reg[i] {
+		h.Reg[i] = zeroBits
+	}
+}
+
+// Merge takes another HyperLogLog and combines it with HyperLogLog h.
+func (h *HyperLogLog) Merge(other *HyperLogLog) error {
+	if h.p != other.p {
+		return errors.New("precisions must be equal")
+	}
+
+	for i, v := range other.Reg {
+		if v > h.Reg[i] {
+			h.Reg[i] = v
+		}
+	}
+	return nil
+}
+
+// Count returns the cardinality estimate.
[…]

### 2022-03-07 paper/code/redis/src/acl.c

@@ -1,0 +1,2937 @@
+/*
+ * Copyright (c) 2018, Salvatore Sanfilippo <antirez at gmail dot com>
+ * All rights reserved.
+ *
+ * Redistribution and use in source and binary forms, with or without
+ * modification, are permitted provided that the following conditions are met:
+ *
+ *   * Redistributions of source code must retain the above copyright notice,
+ *     this list of conditions and the following disclaimer.
+ *   * Redistributions in binary form must reproduce the above copyright
+ *     notice, this list of conditions and the following disclaimer in the
+ *     documentation and/or other materials provided with the distribution.
+ *   * Neither the name of Redis nor the names of its contributors may be used
+ *     to endorse or promote products derived from this software without
+ *     specific prior written permission.
+ *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+ * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+ * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
+ * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
+ * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
+ * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
+ * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
+ * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
+ * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
+ * POSSIBILITY OF SUCH DAMAGE.
+ */
+
+#include "server.h"
+#include "sha256.h"
+#include <fcntl.h>
+#include <ctype.h>
+
+/* =============================================================================
+ * Global state for ACLs
+ * ==========================================================================*/
+
+rax *Users; /* Table mapping usernames to user structures. */
+
[…]

### 2022-03-07 paper/code/redis/src/config.c

@@ -1,0 +1,2984 @@
+/* Configuration file parsing and CONFIG GET/SET commands implementation.
+ *
+ * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
+ * All rights reserved.
+ *
+ * Redistribution and use in source and binary forms, with or without
+ * modification, are permitted provided that the following conditions are met:
+ *
+ *   * Redistributions of source code must retain the above copyright notice,
+ *     this list of conditions and the following disclaimer.
+ *   * Redistributions in binary form must reproduce the above copyright
+ *     notice, this list of conditions and the following disclaimer in the
+ *     documentation and/or other materials provided with the distribution.
+ *   * Neither the name of Redis nor the names of its contributors may be used
+ *     to endorse or promote products derived from this software without
+ *     specific prior written permission.
+ *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+ * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+ * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
+ * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
+ * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
+ * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
+ * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
+ * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
+ * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
+ * POSSIBILITY OF SUCH DAMAGE.
+ */
+
+#include "server.h"
+#include "cluster.h"
+
+#include <fcntl.h>
+#include <sys/stat.h>
+#include <glob.h>
+#include <string.h>
+
[…]

### 2022-03-07 paper/code/redis/src/hyperloglog.c

@@ -1,0 +1,1692 @@
+/* hyperloglog.c - Redis HyperLogLog probabilistic cardinality approximation.
+ * This file implements the algorithm and the exported Redis commands.
+ *
+ * Copyright (c) 2014, Salvatore Sanfilippo <antirez at gmail dot com>
+ * All rights reserved.
+ *
+ * Redistribution and use in source and binary forms, with or without
+ * modification, are permitted provided that the following conditions are met:
+ *
+ *   * Redistributions of source code must retain the above copyright notice,
+ *     this list of conditions and the following disclaimer.
+ *   * Redistributions in binary form must reproduce the above copyright
+ *     notice, this list of conditions and the following disclaimer in the
+ *     documentation and/or other materials provided with the distribution.
+ *   * Neither the name of Redis nor the names of its contributors may be used
+ *     to endorse or promote products derived from this software without
+ *     specific prior written permission.
+ *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
+ * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
+ * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
+ * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
+ * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
+ * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
+ * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
+ * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
+ * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
+ * POSSIBILITY OF SUCH DAMAGE.
+ */
+
+#include "server.h"
+
+#include <stdint.h>
+#include <math.h>
+
[…]

### 2022-03-07 scripts/classic_hll/attack_hll.go

@@ -1,0 +1,325 @@
+package main
+
+import (
+	"encoding/binary"
+	"fmt"
+	"hash"
+	"math"
+	"math/rand"
+	"time"
+
+	"./clarkduvall/hyperloglog"
+	"./spaolacci/murmur3"
+)
+
+var n = uint8(8)
+var m = 1 << n
+var maxValue = int(math.Pow(2, 20)) - 1
+var iterations = 50
+
+var defaultB = uint32(1)
+var defaultT = uint8(1)
+var empty = 0
+
+func main() {
+
+	fmt.Printf("For %d iterations:\n", iterations)
+
+	//S1 scenario
+
+	fmt.Printf("S1 Scenario: (original, final, number of items inserted, number of resets)\n")
+	var originalEst, finalEst, nItermInserted, nResets, expected int
+
+	originalEst, finalEst, nItermInserted, nResets = runAttack("S1", empty, false, defaultB, defaultT)
+	fmt.Printf("Empty sketch, B = 1: %d, %d, %d, resets: %d\n", originalEst, finalEst, nItermInserted, nResets)
+
+	originalEst, finalEst, nItermInserted, nResets = runAttack("S1", empty, false, uint32(m/2), defaultT)
+	fmt.Printf("Empty sketch, B = %d: %d, %d, %d, resets: %d\n", uint32(m/2), originalEst, finalEst, nItermInserted, nResets)
+
+	//S2 scenario
+
+	fmt.Printf("S2 Scenario: (original, final, number of items inserted)\n")
+	originalEst, finalEst, nItermInserted, _ = runAttack("S2", empty, false, defaultB, defaultT)
+	fmt.Printf("Empty sketch, B = 1: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
+
+	kmlogm := []int{}
+	for t := float64(1); t < 11; t++ {
+		kmlogm = append(kmlogm, int(math.Pow(2, t-1)*float64(m)*math.Log(float64(m))))
+	}
+
+	for i, val := range kmlogm {
+		originalEst, finalEst, nItermInserted, _ = runAttack("S2Preload", val, false, defaultB, uint8(i+1))
+		fmt.Printf("Non-empty sketch, t = %d: %d, %d, %d\n", i+1, originalEst, finalEst, nItermInserted)
+	}
+
+	// S4 scenario
+
+	fmt.Printf("S4 Scenario: (original, final, number of items inserted, expected fraction to insert)\n")
[…]

### 2022-03-07 scripts/redis/src/release.h

@@ -1,0 +1,3 @@
+#define REDIS_GIT_SHA1 "00000000"
+#define REDIS_GIT_DIRTY "       0"
+#define REDIS_BUILD_ID "IC-SPRING-LMC03.local-1644943718"

### 2022-03-07 paper/code copy/redis/src/hyperloglog.c
keeping only scripts

@@ -1,1692 +1,0 @@
-/* hyperloglog.c - Redis HyperLogLog probabilistic cardinality approximation.
- * This file implements the algorithm and the exported Redis commands.
- *
- * Copyright (c) 2014, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-
-#include <stdint.h>
-#include <math.h>
-
-/* The Redis HyperLogLog implementation is based on the following ideas:
- *
[…]

### 2022-03-07 paper/code/classic_hll/attack_hll.go

@@ -1,324 +1,0 @@
-package main
-
-import (
-	"encoding/binary"
-	"fmt"
-	"hash"
-	"math"
-	"math/rand"
-	"time"
-
-	"./clarkduvall/hyperloglog"
-	"./spaolacci/murmur3"
-)
-
-var n = uint8(8)
-var m = 1 << n
-var maxValue = int(math.Pow(2, 20)) - 1
-var iterations = 50
-
-var defaultB = uint32(1)
-var defaultT = uint8(1)
-var empty = 0
-
-func main() {
-
-	fmt.Printf("For %d iterations:\n", iterations)
-
-	//S1 scenario
-
-	fmt.Printf("S1 Scenario: (original, final, number of items inserted, number of resets)\n")
-	var originalEst, finalEst, nItermInserted, nResets, expected int
-
-	originalEst, finalEst, nItermInserted, nResets = runAttack("S1", empty, false, defaultB, defaultT)
-	fmt.Printf("Empty sketch, B = 1: %d, %d, %d, resets: %d\n", originalEst, finalEst, nItermInserted, nResets)
-
-	originalEst, finalEst, nItermInserted, nResets = runAttack("S1", empty, false, uint32(m/2), defaultT)
-	fmt.Printf("Empty sketch, B = %d: %d, %d, %d, resets: %d\n", uint32(m/2), originalEst, finalEst, nItermInserted, nResets)
-
-	//S2 scenario
-	fmt.Printf("S2 Scenario: (original, final, number of items inserted)\n")
-	originalEst, finalEst, nItermInserted, _ = runAttack("S2", empty, false, defaultB, defaultT)
-	fmt.Printf("Empty sketch, B = 1: %d, %d, %d\n", originalEst, finalEst, nItermInserted)
-
-	// S2 Preload
-
-	kmlogm := []int{}
-	for t := float64(1); t < 11; t++ {
-		kmlogm = append(kmlogm, int(math.Pow(2, t-1)*float64(m)*math.Log(float64(m))))
-	}
-
-	for i, val := range kmlogm {
-		originalEst, finalEst, nItermInserted, _ = runAttack("S2Preload", val, false, defaultB, uint8(i+1))
-		fmt.Printf("t = %d: %d, %d, %d\n", i+1, originalEst, finalEst, nItermInserted)
-	}
-
-	//S4
-	fmt.Printf("S4 Scenario: (original, final, number of items inserted, expected fraction to insert)\n")
[…]

### 2022-03-07 paper/code/classic_hll/clarkduvall/hyperloglog/hyperloglog.go

@@ -1,118 +1,0 @@
-// Package hyperloglog implements the HyperLogLog and HyperLogLog++ cardinality
-// estimation algorithms.
-// These algorithms are used for accurately estimating the cardinality of a
-// multiset using constant memory. HyperLogLog++ has multiple improvements over
-// HyperLogLog, with a much lower error rate for smaller cardinalities.
-//
-// HyperLogLog is described here:
-// http://algo.inria.fr/flajolet/Publications/FlFuGaMe07.pdf
-//
-// HyperLogLog++ is described here:
-// http://research.google.com/pubs/pub40671.html
-package hyperloglog
-
-import (
-	"bytes"
-	"encoding/gob"
-	"errors"
-	"math"
-)
-
-const two32 = 1 << 32
-
-type HyperLogLog struct {
-	Reg []uint8
-	m   uint32
-	p   uint8
-}
-
-// New returns a new initialized HyperLogLog.
-func New(precision uint8) (*HyperLogLog, error) {
-	if precision > 16 || precision < 4 {
-		return nil, errors.New("precision must be between 4 and 16")
-	}
-
-	h := &HyperLogLog{}
-	h.p = precision
-	h.m = 1 << precision
-	h.Reg = make([]uint8, h.m)
-	return h, nil
-}
-
-// Clear sets HyperLogLog h back to its initial state.
-func (h *HyperLogLog) Clear() {
-	h.Reg = make([]uint8, h.m)
-}
-
-// Add adds a new item to HyperLogLog h.
-func (h *HyperLogLog) Add(item Hash32) {
-	x := item.Sum32()
-	i := eb32(x, 32, 32-h.p) // {x31,...,x32-p}
-	w := x<<h.p | 1<<(h.p-1) // {x32-p,...,x0}
-
-	zeroBits := clz32(w) + 1
-	if zeroBits > h.Reg[i] {
-		h.Reg[i] = zeroBits
-	}
-}
-
-// Merge takes another HyperLogLog and combines it with HyperLogLog h.
-func (h *HyperLogLog) Merge(other *HyperLogLog) error {
-	if h.p != other.p {
-		return errors.New("precisions must be equal")
-	}
-
-	for i, v := range other.Reg {
-		if v > h.Reg[i] {
-			h.Reg[i] = v
-		}
-	}
-	return nil
-}
-
-// Count returns the cardinality estimate.
[…]

### 2022-03-07 paper/code/classic_hll/clarkduvall/hyperloglog/hyperloglog_test.go

@@ -1,204 +1,0 @@
-package hyperloglog
-
-import (
-	"bytes"
-	"encoding/gob"
-	"reflect"
-	"testing"
-)
-
-type fakeHash32 uint32
-
-func (f fakeHash32) Sum32() uint32 { return uint32(f) }
-
-func TestHLLAdd(t *testing.T) {
-	h, _ := New(16)
-
-	h.Add(fakeHash32(0x00010fff))
-	n := h.Reg[1]
-	if n != 5 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash32(0x0002ffff))
-	n = h.Reg[2]
-	if n != 1 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash32(0x00030000))
-	n = h.Reg[3]
-	if n != 17 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash32(0x00030001))
-	n = h.Reg[3]
-	if n != 17 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash32(0xff037000))
-	n = h.Reg[0xff03]
-	if n != 2 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash32(0xff030800))
-	n = h.Reg[0xff03]
-	if n != 5 {
-		t.Error(n)
-	}
-}
-
-func TestHLLCount(t *testing.T) {
-	h, _ := New(16)
-
-	n := h.Count()
-	if n != 0 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash32(0x00010fff))
-	h.Add(fakeHash32(0x00020fff))
-	h.Add(fakeHash32(0x00030fff))
-	h.Add(fakeHash32(0x00040fff))
-	h.Add(fakeHash32(0x00050fff))
-	h.Add(fakeHash32(0x00050fff))
-
-	n = h.Count()
-	if n != 5 {
-		t.Error(n)
-	}
-}
-
-func TestHLLMergeError(t *testing.T) {
-	h, _ := New(16)
-	h2, _ := New(10)
-
-	err := h.Merge(h2)
-	if err == nil {
-		t.Error("different precision should return error")
-	}
-}
-
-func TestHLLMerge(t *testing.T) {
-	h, _ := New(16)
-	h.Add(fakeHash32(0x00010fff))
-	h.Add(fakeHash32(0x00020fff))
-	h.Add(fakeHash32(0x00030fff))
[…]

### 2022-03-07 paper/code/classic_hll/clarkduvall/hyperloglog/hyperloglogplus_test.go

@@ -1,596 +1,0 @@
-package hyperloglog
-
-import (
-	"bytes"
-	"encoding/gob"
-	"math"
-	"reflect"
-	"testing"
-)
-
-type fakeHash64 uint64
-
-func (f fakeHash64) Sum64() uint64 { return uint64(f) }
-
-func abs(x int64) int64 {
-	if x < 0 {
-		return -x
-	}
-	return x
-}
-
-func TestHLLPPAddNoSparse(t *testing.T) {
-	h, _ := NewPlus(16)
-	h.toNormal()
-
-	h.Add(fakeHash64(0x00010fffffffffff))
-	n := h.reg[1]
-	if n != 5 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash64(0x0002ffffffffffff))
-	n = h.reg[2]
-	if n != 1 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash64(0x0003000000000000))
-	n = h.reg[3]
-	if n != 49 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash64(0x0003000000000001))
-	n = h.reg[3]
-	if n != 49 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash64(0xff03700000000000))
-	n = h.reg[0xff03]
-	if n != 2 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash64(0xff03080000000000))
-	n = h.reg[0xff03]
-	if n != 5 {
-		t.Error(n)
-	}
-}
-
-func TestHLLPPPrecisionNoSparse(t *testing.T) {
-	h, _ := NewPlus(4)
-	h.toNormal()
-
-	h.Add(fakeHash64(0x1fffffffffffffff))
-	n := h.reg[1]
-	if n != 1 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash64(0xffffffffffffffff))
-	n = h.reg[0xf]
-	if n != 1 {
-		t.Error(n)
-	}
-
-	h.Add(fakeHash64(0x00ffffffffffffff))
-	n = h.reg[0]
-	if n != 5 {
-		t.Error(n)
-	}
-}
-
-func TestHLLPPToNormal(t *testing.T) {
-	h, _ := NewPlus(16)
-	h.Add(fakeHash64(0x00010fffffffffff))
-	h.mergeSparseAndToNormal()
-	c := h.Count()
-	if c != 1 {
-		t.Error(c)
-	}
-
-	if h.sparse {
[…]

### 2022-03-07 paper/code/redis/src/acl.c

@@ -1,2937 +1,0 @@
-/*
- * Copyright (c) 2018, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "sha256.h"
-#include <fcntl.h>
-#include <ctype.h>
-
-/* =============================================================================
- * Global state for ACLs
- * ==========================================================================*/
-
-rax *Users; /* Table mapping usernames to user structures. */
-
-user *DefaultUser;  /* Global reference to the default user.
[…]

### 2022-03-07 paper/code/redis/src/anet.h

@@ -1,77 +1,0 @@
-/* anet.c -- Basic TCP socket stuff made a bit less boring
- *
- * Copyright (c) 2006-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#ifndef ANET_H
-#define ANET_H
-
-#include <sys/types.h>
-
-#define ANET_OK 0
-#define ANET_ERR -1
-#define ANET_ERR_LEN 256
-
-/* Flags used with certain functions. */
-#define ANET_NONE 0
[…]

### 2022-03-07 paper/code/redis/src/aof.c

@@ -1,2541 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "bio.h"
-#include "rio.h"
-#include "functions.h"
-
-#include <signal.h>
-#include <fcntl.h>
-#include <sys/stat.h>
-#include <sys/types.h>
-#include <sys/time.h>
-#include <sys/resource.h>
[…]

### 2022-03-07 paper/code/redis/src/bitops.c

@@ -1,1263 +1,0 @@
-/* Bit operations.
- *
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-
-/* -----------------------------------------------------------------------------
- * Helpers and low level bit functions.
- * -------------------------------------------------------------------------- */
-
-/* Count number of bits set in the binary array pointed by 's' and long
- * 'count' bytes. The implementation of this function is required to
[…]

### 2022-03-07 paper/code/redis/src/blocked.c

@@ -1,818 +1,0 @@
-/* blocked.c - generic support for blocking operations like BLPOP & WAIT.
- *
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- *
- * ---------------------------------------------------------------------------
- *
- * API:
- *
- * blockClient() set the CLIENT_BLOCKED flag in the client, and set the
- * specified block type 'btype' filed to one of BLOCKED_* macros.
- *
- * unblockClient() unblocks the client doing the following:
[…]

### 2022-03-07 paper/code/redis/src/call_reply.c

@@ -1,516 +1,0 @@
-/*
- * Copyright (c) 2009-2021, Redis Labs Ltd.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "call_reply.h"
-
-#define REPLY_FLAG_ROOT (1<<0)
-#define REPLY_FLAG_PARSED (1<<1)
-#define REPLY_FLAG_RESP3 (1<<2)
-
-/* --------------------------------------------------------
- * An opaque struct used to parse a RESP protocol reply and
[…]

### 2022-03-07 paper/code/redis/src/call_reply.h

@@ -1,56 +1,0 @@
-/*
- * Copyright (c) 2009-2021, Redis Labs Ltd.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#ifndef SRC_CALL_REPLY_H_
-#define SRC_CALL_REPLY_H_
-
-#include "resp_parser.h"
-
-typedef struct CallReply CallReply;
-
-CallReply *callReplyCreate(sds reply, void *private_data);
-int callReplyType(CallReply *rep);
-const char *callReplyGetString(CallReply *rep, size_t *len);
[…]

### 2022-03-07 paper/code/redis/src/cli_common.c

@@ -1,373 +1,0 @@
-/* CLI (command line interface) common methods
- * 
- * Copyright (c) 2020, Redis Labs
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "fmacros.h"
-#include "cli_common.h"
-#include <stdio.h>
-#include <stdlib.h>
-#include <fcntl.h>
-#include <errno.h>
-#include <hiredis.h>
-#include <sdscompat.h> /* Use hiredis' sds compat header that maps sds calls to their hi_ variants */
[…]

### 2022-03-07 paper/code/redis/src/cli_common.h

@@ -1,51 +1,0 @@
-#ifndef __CLICOMMON_H
-#define __CLICOMMON_H
-
-#include <hiredis.h>
-#include <sdscompat.h> /* Use hiredis' sds compat header that maps sds calls to their hi_ variants */
-
-typedef struct cliSSLconfig {
-    /* Requested SNI, or NULL */
-    char *sni;
-    /* CA Certificate file, or NULL */
-    char *cacert;
-    /* Directory where trusted CA certificates are stored, or NULL */
-    char *cacertdir;
-    /* Skip server certificate verification. */
-    int skip_cert_verify;
-    /* Client certificate to authenticate with, or NULL */
-    char *cert;
-    /* Private key file to authenticate with, or NULL */
-    char *key;
-    /* Preferred cipher list, or NULL (applies only to <= TLSv1.2) */
-    char* ciphers;
-    /* Preferred ciphersuites list, or NULL (applies only to TLSv1.3) */
-    char* ciphersuites;
-} cliSSLconfig;
-
-
-/* server connection information object, used to describe an ip:port pair, db num user input, and user:pass. */
-typedef struct cliConnInfo {
-    char *hostip;
-    int hostport;
-    int input_dbnum;
-    char *auth;
-    char *user;
-} cliConnInfo;
-
-int cliSecureConnection(redisContext *c, cliSSLconfig config, const char **err);
-
-ssize_t cliWriteConn(redisContext *c, const char *buf, size_t buf_len);
-
-int cliSecureInit();
-
-sds readArgFromStdin(void);
-
-sds *getSdsArrayFromArgv(int argc,char **argv, int quoted);
-
-sds unquoteCString(char *str);
-
-void parseRedisUri(const char *uri, const char* tool_name, cliConnInfo *connInfo, int *tls_flag);
-
-void freeCliConnInfo(cliConnInfo connInfo);
-#endif /* __CLICOMMON_H */

### 2022-03-07 paper/code/redis/src/cluster.h

@@ -1,369 +1,0 @@
-#ifndef __CLUSTER_H
-#define __CLUSTER_H
-
-/*-----------------------------------------------------------------------------
- * Redis cluster data structures, defines, exported API.
- *----------------------------------------------------------------------------*/
-
-#define CLUSTER_SLOTS 16384
-#define CLUSTER_OK 0            /* Everything looks ok */
-#define CLUSTER_FAIL 1          /* The cluster can't work */
-#define CLUSTER_NAMELEN 40      /* sha1 hex length */
-#define CLUSTER_PORT_INCR 10000 /* Cluster port = baseport + PORT_INCR */
-
-/* The following defines are amount of time, sometimes expressed as
- * multiplicators of the node timeout value (when ending with MULT). */
-#define CLUSTER_FAIL_REPORT_VALIDITY_MULT 2 /* Fail report validity. */
-#define CLUSTER_FAIL_UNDO_TIME_MULT 2 /* Undo fail if master is back. */
-#define CLUSTER_MF_TIMEOUT 5000 /* Milliseconds to do a manual failover. */
-#define CLUSTER_MF_PAUSE_MULT 2 /* Master pause manual failover mult. */
-#define CLUSTER_SLAVE_MIGRATION_DELAY 5000 /* Delay for slave migration. */
-
-/* Redirection errors returned by getNodeByQuery(). */
-#define CLUSTER_REDIR_NONE 0          /* Node can serve the request. */
-#define CLUSTER_REDIR_CROSS_SLOT 1    /* -CROSSSLOT request. */
-#define CLUSTER_REDIR_UNSTABLE 2      /* -TRYAGAIN redirection required */
-#define CLUSTER_REDIR_ASK 3           /* -ASK redirection required. */
-#define CLUSTER_REDIR_MOVED 4         /* -MOVED redirection required. */
-#define CLUSTER_REDIR_DOWN_STATE 5    /* -CLUSTERDOWN, global state. */
-#define CLUSTER_REDIR_DOWN_UNBOUND 6  /* -CLUSTERDOWN, unbound slot. */
-#define CLUSTER_REDIR_DOWN_RO_STATE 7 /* -CLUSTERDOWN, allow reads. */
-
-struct clusterNode;
-
-/* clusterLink encapsulates everything needed to talk with a remote node. */
-typedef struct clusterLink {
-    mstime_t ctime;             /* Link creation time */
-    connection *conn;           /* Connection to remote node */
[…]

### 2022-03-07 paper/code/redis/src/config.c

@@ -1,2984 +1,0 @@
-/* Configuration file parsing and CONFIG GET/SET commands implementation.
- *
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "cluster.h"
-
-#include <fcntl.h>
-#include <sys/stat.h>
-#include <glob.h>
-#include <string.h>
-
-/*-----------------------------------------------------------------------------
- * Config file name-value maps.
[…]

### 2022-03-07 paper/code/redis/src/connection.c

@@ -1,433 +1,0 @@
-/*
- * Copyright (c) 2019, Redis Labs
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "connhelpers.h"
-
-/* The connections module provides a lean abstraction of network connections
- * to avoid direct socket and async event management across the Redis code base.
- *
- * It does NOT provide advanced connection features commonly found in similar
- * libraries such as complete in/out buffer management, throttling, etc. These
- * functions remain in networking.c.
- *
- * The primary goal is to allow transparent handling of TCP and TLS based
[…]

### 2022-03-07 paper/code/redis/src/connection.h

@@ -1,244 +1,0 @@
-
-/*
- * Copyright (c) 2019, Redis Labs
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#ifndef __REDIS_CONNECTION_H
-#define __REDIS_CONNECTION_H
-
-#include <errno.h>
-
-#define CONN_INFO_LEN   32
-
-struct aeEventLoop;
-typedef struct connection connection;
-
-typedef enum {
[…]

### 2022-03-07 paper/code/redis/src/defrag.c

@@ -1,1221 +1,0 @@
-/* 
- * Active memory defragmentation
- * Try to find key / value allocations that need to be re-allocated in order 
- * to reduce external fragmentation.
- * We do that by scanning the keyspace and for each pointer we have, we can try to
- * ask the allocator if moving it to a new address will help reduce fragmentation.
- *
- * Copyright (c) 2020, Redis Labs, Inc
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "cluster.h"
-#include <time.h>
[…]

### 2022-03-07 paper/code/redis/src/eval.c

@@ -1,1649 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "sha1.h"
-#include "rand.h"
-#include "cluster.h"
-#include "monotonic.h"
-#include "resp_parser.h"
-#include "script_lua.h"
-
-#include <lua.h>
-#include <lauxlib.h>
-#include <lualib.h>
-#include <ctype.h>
[…]

### 2022-03-07 paper/code/redis/src/evict.c

@@ -1,769 +1,0 @@
-/* Maxmemory directive handling (LRU eviction and other policies).
- *
- * ----------------------------------------------------------------------------
- *
- * Copyright (c) 2009-2016, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "bio.h"
-#include "atomicvar.h"
-#include "script.h"
-#include <math.h>
-
-/* ----------------------------------------------------------------------------
- * Data structures
- * --------------------------------------------------------------------------*/
-
[…]

### 2022-03-07 paper/code/redis/src/functions.c

@@ -1,1050 +1,0 @@
-/*
- * Copyright (c) 2021, Redis Ltd.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "functions.h"
-#include "sds.h"
-#include "dict.h"
-#include "adlist.h"
-#include "atomicvar.h"
-
-typedef enum {
-    restorePolicy_Flush, restorePolicy_Append, restorePolicy_Replace
-} restorePolicy;
-
-static size_t engine_cache_memory = 0;
-
-/* Forward declaration */
[…]

### 2022-03-07 paper/code/redis/src/functions.h

@@ -1,133 +1,0 @@
-/*
- * Copyright (c) 2021, Redis Ltd.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#ifndef __FUNCTIONS_H_
-#define __FUNCTIONS_H_
-
-/*
- * functions.c unit provides the Redis Functions API:
- * * FUNCTION CREATE
- * * FUNCTION CALL
- * * FUNCTION DELETE
- * * FUNCTION KILL
- * * FUNCTION INFO
- *
- * Also contains implementation for:
- * * Save/Load function from rdb
[…]

### 2022-03-07 paper/code/redis/src/geohash.h

@@ -1,136 +1,0 @@
-/*
- * Copyright (c) 2013-2014, yinqiwen <yinqiwen@gmail.com>
- * Copyright (c) 2014, Matt Stancliff <matt@genges.com>.
- * Copyright (c) 2015, Salvatore Sanfilippo <antirez@gmail.com>.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *  * Redistributions of source code must retain the above copyright notice,
- *    this list of conditions and the following disclaimer.
- *  * Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- *  * Neither the name of Redis nor the names of its contributors may be used
- *    to endorse or promote products derived from this software without
- *    specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS
- * BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF
- * THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-#ifndef GEOHASH_H_
-#define GEOHASH_H_
-
-#include <stddef.h>
-#include <stdint.h>
[…]

### 2022-03-07 paper/code/redis/src/geohash_helper.c

@@ -1,263 +1,0 @@
-/*
- * Copyright (c) 2013-2014, yinqiwen <yinqiwen@gmail.com>
- * Copyright (c) 2014, Matt Stancliff <matt@genges.com>.
- * Copyright (c) 2015-2016, Salvatore Sanfilippo <antirez@gmail.com>.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *  * Redistributions of source code must retain the above copyright notice,
- *    this list of conditions and the following disclaimer.
- *  * Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- *  * Neither the name of Redis nor the names of its contributors may be used
- *    to endorse or promote products derived from this software without
- *    specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS
- * BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF
- * THE POSSIBILITY OF SUCH DAMAGE.
- */
-
-/* This is a C++ to C conversion from the ardb project.
- * This file started out as:
[…]

### 2022-03-07 paper/code/redis/src/hyperloglog.c

@@ -1,1692 +1,0 @@
-/* hyperloglog.c - Redis HyperLogLog probabilistic cardinality approximation.
- * This file implements the algorithm and the exported Redis commands.
- *
- * Copyright (c) 2014, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-
-#include <stdint.h>
-#include <math.h>
-
-/* The Redis HyperLogLog implementation is based on the following ideas:
- *
[…]

### 2022-03-07 paper/code/redis/src/listpack.c

@@ -1,2417 +1,0 @@
-/* Listpack -- A lists of strings serialization format
- *
- * This file implements the specification you can find at:
- *
- *  https://github.com/antirez/listpack
- *
- * Copyright (c) 2017, Salvatore Sanfilippo <antirez at gmail dot com>
- * Copyright (c) 2020, Redis Labs, Inc
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
[…]

### 2022-03-07 paper/code/redis/src/listpack.h

@@ -1,98 +1,0 @@
-/* Listpack -- A lists of strings serialization format
- *
- * This file implements the specification you can find at:
- *
- *  https://github.com/antirez/listpack
- *
- * Copyright (c) 2017, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#ifndef __LISTPACK_H
-#define __LISTPACK_H
-
-#include <stdlib.h>
-#include <stdint.h>
-
[…]

### 2022-03-07 paper/code/redis/src/modules/hellocluster.c

@@ -1,119 +1,0 @@
-/* Helloworld cluster -- A ping/pong cluster API example.
- *
- * -----------------------------------------------------------------------------
- *
- * Copyright (c) 2018, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#define REDISMODULE_EXPERIMENTAL_API
-#include "../redismodule.h"
-#include <stdio.h>
[…]

### 2022-03-07 paper/code/redis/src/modules/hellodict.c

@@ -1,132 +1,0 @@
-/* Hellodict -- An example of modules dictionary API
- *
- * This module implements a volatile key-value store on top of the
- * dictionary exported by the Redis modules API.
- *
- * -----------------------------------------------------------------------------
- *
- * Copyright (c) 2018, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#define REDISMODULE_EXPERIMENTAL_API
[…]

### 2022-03-07 paper/code/redis/src/multi.c

@@ -1,428 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-
-/* ================================ MULTI/EXEC ============================== */
-
-/* Client state initialization for MULTI/EXEC */
-void initClientMultiState(client *c) {
-    c->mstate.commands = NULL;
-    c->mstate.count = 0;
-    c->mstate.cmd_flags = 0;
[…]

### 2022-03-07 paper/code/redis/src/networking.c

@@ -1,4193 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "atomicvar.h"
-#include "cluster.h"
-#include "script.h"
-#include <sys/socket.h>
-#include <sys/uio.h>
-#include <math.h>
-#include <ctype.h>
-
-static void setProtocolError(const char *errstr, client *c);
-int postponeClientRead(client *c);
[…]

### 2022-03-07 paper/code/redis/src/notify.c

@@ -1,143 +1,0 @@
-/*
- * Copyright (c) 2013, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-
-/* This file implements keyspace events notification via Pub/Sub and
- * described at https://redis.io/topics/notifications. */
-
-/* Turn a string representing notification classes into an integer
- * representing notification classes flags xored.
- *
- * The function returns -1 if the input contains characters not mapping to
- * any class. */
[…]

### 2022-03-07 paper/code/redis/src/rdb.c

@@ -1,3506 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "lzf.h"    /* LZF compression library */
-#include "zipmap.h"
-#include "endianconv.h"
-#include "stream.h"
-#include "functions.h"
-
-#include <math.h>
-#include <fcntl.h>
-#include <sys/types.h>
[…]

### 2022-03-07 paper/code/redis/src/rdb.h

@@ -1,177 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#ifndef __RDB_H
-#define __RDB_H
-
-#include <stdio.h>
-#include "rio.h"
-
-/* TBD: include only necessary headers. */
-#include "server.h"
-
[…]

### 2022-03-07 paper/code/redis/src/redis-benchmark.c

@@ -1,2016 +1,0 @@
-/* Redis benchmark utility.
- *
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "fmacros.h"
-#include "version.h"
-
-#include <stdio.h>
-#include <string.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <errno.h>
-#include <time.h>
-#include <sys/time.h>
-#include <signal.h>
-#include <assert.h>
[…]

### 2022-03-07 paper/code/redis/src/redis-check-aof.c

@@ -1,301 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Pieter Noordhuis <pcnoordhuis at gmail dot com>
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include <sys/stat.h>
-
-#define ERROR(...) { \
-    char __buf[1024]; \
-    snprintf(__buf, sizeof(__buf), __VA_ARGS__); \
[…]

### 2022-03-07 paper/code/redis/src/redis-check-rdb.c

@@ -1,432 +1,0 @@
-/*
- * Copyright (c) 2016, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "mt19937-64.h"
-#include "server.h"
-#include "rdb.h"
-
-#include <stdarg.h>
-#include <sys/time.h>
-#include <unistd.h>
-#include <sys/stat.h>
-
-void createSharedObjects(void);
[…]

### 2022-03-07 paper/code/redis/src/redismodule.h

@@ -1,1353 +1,0 @@
-#ifndef REDISMODULE_H
-#define REDISMODULE_H
-
-#include <sys/types.h>
-#include <stdint.h>
-#include <stdio.h>
-#include <stdlib.h>
-
-/* ---------------- Defines common between core and modules --------------- */
-
-/* Error status return values. */
-#define REDISMODULE_OK 0
-#define REDISMODULE_ERR 1
-
-/* API versions. */
-#define REDISMODULE_APIVER_1 1
-
-/* Version of the RedisModuleTypeMethods structure. Once the RedisModuleTypeMethods 
- * structure is changed, this version number needs to be changed synchronistically. */
-#define REDISMODULE_TYPE_METHOD_VERSION 4
-
-/* API flags and constants */
-#define REDISMODULE_READ (1<<0)
-#define REDISMODULE_WRITE (1<<1)
-
-/* RedisModule_OpenKey extra flags for the 'mode' argument.
- * Avoid touching the LRU/LFU of the key when opened. */
-#define REDISMODULE_OPEN_KEY_NOTOUCH (1<<16)
-
-/* List push and pop */
-#define REDISMODULE_LIST_HEAD 0
-#define REDISMODULE_LIST_TAIL 1
-
-/* Key types. */
-#define REDISMODULE_KEYTYPE_EMPTY 0
-#define REDISMODULE_KEYTYPE_STRING 1
-#define REDISMODULE_KEYTYPE_LIST 2
-#define REDISMODULE_KEYTYPE_HASH 3
-#define REDISMODULE_KEYTYPE_SET 4
-#define REDISMODULE_KEYTYPE_ZSET 5
-#define REDISMODULE_KEYTYPE_MODULE 6
-#define REDISMODULE_KEYTYPE_STREAM 7
-
-/* Reply types. */
-#define REDISMODULE_REPLY_UNKNOWN -1
-#define REDISMODULE_REPLY_STRING 0
-#define REDISMODULE_REPLY_ERROR 1
-#define REDISMODULE_REPLY_INTEGER 2
-#define REDISMODULE_REPLY_ARRAY 3
-#define REDISMODULE_REPLY_NULL 4
-#define REDISMODULE_REPLY_MAP 5
-#define REDISMODULE_REPLY_SET 6
-#define REDISMODULE_REPLY_BOOL 7
-#define REDISMODULE_REPLY_DOUBLE 8
[…]

### 2022-03-07 paper/code/redis/src/release.h

@@ -1,3 +1,0 @@
-#define REDIS_GIT_SHA1 "7677981b"
-#define REDIS_GIT_DIRTY "     984"
-#define REDIS_BUILD_ID "IC-SPRING-LMC03.local-1644853386"

### 2022-03-07 paper/code/redis/src/script.c

@@ -1,540 +1,0 @@
-/*
- * Copyright (c) 2009-2021, Redis Ltd.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "script.h"
-#include "cluster.h"
-
-scriptFlag scripts_flags_def[] = {
-    {.flag = SCRIPT_FLAG_NO_WRITES, .str = "no-writes"},
-    {.flag = SCRIPT_FLAG_ALLOW_OOM, .str = "allow-oom"},
[…]

### 2022-03-07 paper/code/redis/src/script_lua.c

@@ -1,1431 +1,0 @@
-/*
- * Copyright (c) 2009-2021, Redis Ltd.
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "script_lua.h"
-
-#include "server.h"
-#include "sha1.h"
-#include "rand.h"
-#include "cluster.h"
-#include "monotonic.h"
-#include "resp_parser.h"
-#include "version.h"
-#include <lauxlib.h>
-#include <lualib.h>
-#include <ctype.h>
-#include <math.h>
-
[…]

### 2022-03-07 paper/code/redis/src/server.h

@@ -1,3432 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#ifndef __REDIS_H
-#define __REDIS_H
-
-#include "fmacros.h"
-#include "config.h"
-#include "solarisfixes.h"
-#include "rio.h"
-#include "atomicvar.h"
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
[…]

### 2022-03-07 paper/code/redis/src/stream.h

@@ -1,137 +1,0 @@
-#ifndef STREAM_H
-#define STREAM_H
-
-#include "rax.h"
-#include "listpack.h"
-
-/* Stream item ID: a 128 bit number composed of a milliseconds time and
- * a sequence counter. IDs generated in the same millisecond (or in a past
- * millisecond if the clock jumped backward) will use the millisecond time
- * of the latest generated ID and an incremented sequence. */
-typedef struct streamID {
-    uint64_t ms;        /* Unix time in milliseconds. */
-    uint64_t seq;       /* Sequence number. */
-} streamID;
-
-typedef struct stream {
-    rax *rax;               /* The radix tree holding the stream. */
-    uint64_t length;        /* Number of elements inside this stream. */
-    streamID last_id;       /* Zero if there are yet no items. */
-    rax *cgroups;           /* Consumer groups dictionary: name -> streamCG */
-} stream;
-
-/* We define an iterator to iterate stream items in an abstract way, without
- * caring about the radix tree + listpack representation. Technically speaking
- * the iterator is only used inside streamReplyWithRange(), so could just
- * be implemented inside the function, but practically there is the AOF
- * rewriting code that also needs to iterate the stream to emit the XADD
- * commands. */
-typedef struct streamIterator {
-    stream *stream;         /* The stream we are iterating. */
-    streamID master_id;     /* ID of the master entry at listpack head. */
-    uint64_t master_fields_count;       /* Master entries # of fields. */
-    unsigned char *master_fields_start; /* Master entries start in listpack. */
-    unsigned char *master_fields_ptr;   /* Master field to emit next. */
-    int entry_flags;                    /* Flags of entry we are emitting. */
-    int rev;                /* True if iterating end to start (reverse). */
-    uint64_t start_key[2];  /* Start key as 128 bit big endian. */
-    uint64_t end_key[2];    /* End key as 128 bit big endian. */
-    raxIterator ri;         /* Rax iterator. */
-    unsigned char *lp;      /* Current listpack. */
-    unsigned char *lp_ele;  /* Current listpack cursor. */
-    unsigned char *lp_flags; /* Current entry flags pointer. */
-    /* Buffers used to hold the string of lpGet() when the element is
-     * integer encoded, so that there is no string representation of the
-     * element inside the listpack itself. */
-    unsigned char field_buf[LP_INTBUF_SIZE];
-    unsigned char value_buf[LP_INTBUF_SIZE];
-} streamIterator;
-
-/* Consumer group. */
-typedef struct streamCG {
[…]

### 2022-03-07 paper/code/redis/src/t_hash.c

@@ -1,1158 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include <math.h>
-
-/*-----------------------------------------------------------------------------
- * Hash type API
- *----------------------------------------------------------------------------*/
-
-/* Check the length of a number of objects to see if we need to convert a
- * listpack to a real hash. Note that we only check string encoded objects
- * as their string length can be queried in constant time. */
[…]

### 2022-03-07 paper/code/redis/src/t_stream.c

@@ -1,3705 +1,0 @@
-/*
- * Copyright (c) 2017, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include "endianconv.h"
-#include "stream.h"
-
-/* Every stream item inside the listpack, has a flags field that is used to
- * mark the entry as deleted, or having the same field as the "master"
- * entry at the start of the listpack> */
[…]

### 2022-03-07 paper/code/redis/src/t_string.c

@@ -1,943 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "server.h"
-#include <math.h> /* isnan(), isinf() */
-
-/* Forward declarations */
-int getGenericCommand(client *c);
-
-/*-----------------------------------------------------------------------------
- * String Commands
- *----------------------------------------------------------------------------*/
-
[…]

### 2022-03-07 paper/code/redis/src/t_zset.c

@@ -1,4342 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * Copyright (c) 2009-2012, Pieter Noordhuis <pcnoordhuis at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-/*-----------------------------------------------------------------------------
- * Sorted set API
- *----------------------------------------------------------------------------*/
-
-/* ZSETs are ordered sets using two data structures to hold the same elements
[…]

### 2022-03-07 paper/code/redis/src/tls.c

@@ -1,1042 +1,0 @@
-/*
- * Copyright (c) 2019, Redis Labs
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-
-#include "server.h"
-#include "connhelpers.h"
-#include "adlist.h"
-
-#ifdef USE_OPENSSL
-
-#include <openssl/conf.h>
-#include <openssl/ssl.h>
[…]

### 2022-03-07 paper/code/redis/src/util.c

@@ -1,1066 +1,0 @@
-/*
- * Copyright (c) 2009-2012, Salvatore Sanfilippo <antirez at gmail dot com>
- * All rights reserved.
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions are met:
- *
- *   * Redistributions of source code must retain the above copyright notice,
- *     this list of conditions and the following disclaimer.
- *   * Redistributions in binary form must reproduce the above copyright
- *     notice, this list of conditions and the following disclaimer in the
- *     documentation and/or other materials provided with the distribution.
- *   * Neither the name of Redis nor the names of its contributors may be used
- *     to endorse or promote products derived from this software without
- *     specific prior written permission.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
- * AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
- * IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
- * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
- * LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
- * CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
- * SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
- * INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
- * CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
- * ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
- * POSSIBILITY OF SUCH DAMAGE.
- */
-
-#include "fmacros.h"
-#include <stdlib.h>
-#include <stdio.h>
-#include <string.h>
-#include <ctype.h>
-#include <limits.h>
-#include <math.h>
-#include <unistd.h>
-#include <sys/time.h>
-#include <float.h>
-#include <stdint.h>
-#include <errno.h>
-#include <time.h>
[…]

### 2022-03-07 paper/code/redis/src/ziplist.c

@@ -1,2666 +1,0 @@
-/* The ziplist is a specially encoded dually linked list that is designed
- * to be very memory efficient. It stores both strings and integer values,
- * where integers are encoded as actual integers instead of a series of
- * characters. It allows push and pop operations on either side of the list
- * in O(1) time. However, because every operation requires a reallocation of
- * the memory used by the ziplist, the actual complexity is related to the
- * amount of memory used by the ziplist.
- *
- * ----------------------------------------------------------------------------
- *
- * ZIPLIST OVERALL LAYOUT
- * ======================
- *
- * The general layout of the ziplist is as follows:
- *
- * <zlbytes> <zltail> <zllen> <entry> <entry> ... <entry> <zlend>
- *
- * NOTE: all fields are stored in little endian, if not specified otherwise.
- *
- * <uint32_t zlbytes> is an unsigned integer to hold the number of bytes that
- * the ziplist occupies, including the four bytes of the zlbytes field itself.
- * This value needs to be stored to be able to resize the entire structure
- * without the need to traverse it first.
- *
- * <uint32_t zltail> is the offset to the last entry in the list. This allows
- * a pop operation on the far side of the list without the need for full
- * traversal.
- *
- * <uint16_t zllen> is the number of entries. When there are more than
- * 2^16-2 entries, this value is set to 2^16-1 and we need to traverse the
- * entire list to know how many items it holds.
- *
- * <uint8_t zlend> is a special entry representing the end of the ziplist.
- * Is encoded as a single byte equal to 255. No other normal entry starts
- * with a byte set to the value of 255.
- *
- * ZIPLIST ENTRIES
- * ===============
- *
- * Every entry in the ziplist is prefixed by metadata that contains two pieces
- * of information. First, the length of the previous entry is stored to be
- * able to traverse the list from back to front. Second, the entry encoding is
- * provided. It represents the entry type, integer or string, and in the case
- * of strings it also represents the length of the string payload.
- * So a complete entry is stored like this:
- *
- * <prevlen> <encoding> <entry-data>
- *
[…]

### 2022-03-07 paper/code/redis/tests/modules/aclcheck.c

@@ -1,192 +1,0 @@
-
-#include "redismodule.h"
-#include <errno.h>
-#include <assert.h>
-#include <string.h>
-#include <strings.h>
-
-/* A wrap for SET command with ACL check on the key. */
-int set_aclcheck_key(RedisModuleCtx *ctx, RedisModuleString **argv, int argc) {
-    if (argc < 4) {
-        return RedisModule_WrongArity(ctx);
-    }
-
-    int permissions;
-    const char *flags = RedisModule_StringPtrLen(argv[1], NULL);
-
-    if (!strcasecmp(flags, "W")) {
-        permissions = REDISMODULE_KEY_PERMISSION_WRITE;
-    } else if (!strcasecmp(flags, "R")) {
-        permissions = REDISMODULE_KEY_PERMISSION_READ;
-    } else if (!strcasecmp(flags, "*")) {
-        permissions = REDISMODULE_KEY_PERMISSION_ALL;
-    } else {
-        RedisModule_ReplyWithError(ctx, "INVALID FLAGS");
-        return REDISMODULE_OK;
-    }
-
-    /* Check that the key can be accessed */
-    RedisModuleString *user_name = RedisModule_GetCurrentUserName(ctx);
-    RedisModuleUser *user = RedisModule_GetModuleUserFromUserName(user_name);
-    int ret = RedisModule_ACLCheckKeyPermissions(user, argv[2], permissions);
-    if (ret != 0) {
-        RedisModule_ReplyWithError(ctx, "DENIED KEY");
-        RedisModule_FreeModuleUser(user);
-        RedisModule_FreeString(ctx, user_name);
-        return REDISMODULE_OK;
-    }
-
-    RedisModuleCallReply *rep = RedisModule_Call(ctx, "SET", "v", argv + 2, argc - 2);
-    if (!rep) {
-        RedisModule_ReplyWithError(ctx, "NULL reply returned");
-    } else {
-        RedisModule_ReplyWithCallReply(ctx, rep);
-        RedisModule_FreeCallReply(rep);
-    }
-
-    RedisModule_FreeModuleUser(user);
-    RedisModule_FreeString(ctx, user_name);
-    return REDISMODULE_OK;
-}
-
-/* A wrap for PUBLISH command with ACL check on the channel. */
-int publish_aclcheck_channel(RedisModuleCtx *ctx, RedisModuleString **argv, int argc) {
-    if (argc != 3) {
[…]

### 2022-03-07 paper/code/redis/tests/modules/auth.c

@@ -1,100 +1,0 @@
-#include "redismodule.h"
-
-#define UNUSED(V) ((void) V)
-
-// A simple global user
-static RedisModuleUser *global = NULL;
-static long long client_change_delta = 0;
-
-void UserChangedCallback(uint64_t client_id, void *privdata) {
-    REDISMODULE_NOT_USED(privdata);
-    REDISMODULE_NOT_USED(client_id);
-    client_change_delta++;
-}
-
-int Auth_CreateModuleUser(RedisModuleCtx *ctx, RedisModuleString **argv, int argc) {
-    REDISMODULE_NOT_USED(argv);
-    REDISMODULE_NOT_USED(argc);
-
-    if (global) {
-        RedisModule_FreeModuleUser(global);
-    }
-
-    global = RedisModule_CreateModuleUser("global");
-    RedisModule_SetModuleUserACL(global, "allcommands");
-    RedisModule_SetModuleUserACL(global, "allkeys");
-    RedisModule_SetModuleUserACL(global, "on");
-
-    return RedisModule_ReplyWithSimpleString(ctx, "OK");
-}
-
-int Auth_AuthModuleUser(RedisModuleCtx *ctx, RedisModuleString **argv, int argc) {
-    REDISMODULE_NOT_USED(argv);
-    REDISMODULE_NOT_USED(argc);
-    uint64_t client_id;
-    RedisModule_AuthenticateClientWithUser(ctx, global, UserChangedCallback, NULL, &client_id);
-
-    return RedisModule_ReplyWithLongLong(ctx, (uint64_t) client_id);
-}
-
-int Auth_AuthRealUser(RedisModuleCtx *ctx, RedisModuleString **argv, int argc) {
-    if (argc != 2) return RedisModule_WrongArity(ctx);
-
-    size_t length;
-    uint64_t client_id;
-
-    RedisModuleString *user_string = argv[1];
-    const char *name = RedisModule_StringPtrLen(user_string, &length);
-
-    if (RedisModule_AuthenticateClientWithACLUser(ctx, name, length, 
-            UserChangedCallback, NULL, &client_id) == REDISMODULE_ERR) {
-        return RedisModule_ReplyWithError(ctx, "Invalid user");   
-    }
-
-    return RedisModule_ReplyWithLongLong(ctx, (uint64_t) client_id);
-}
-
[…]

### 2022-03-07 paper/code/redis/tests/modules/blockedclient.c

@@ -1,325 +1,0 @@
-/* define macros for having usleep */
-#define _BSD_SOURCE
-#define _DEFAULT_SOURCE
-#include <unistd.h>
-
-#include "redismodule.h"
-#include <assert.h>
-#include <stdio.h>
-#include <pthread.h>
-
-#define UNUSED(V) ((void) V)
-
-/* used to test processing events during slow bg operation */
-static volatile int g_slow_bg_operation = 0;
-static volatile int g_is_in_slow_bg_operation = 0;
-
-void *sub_worker(void *arg) {
-    // Get Redis module context
-    RedisModuleCtx *ctx = (RedisModuleCtx *)arg;
-
-    // Try acquiring GIL
-    int res = RedisModule_ThreadSafeContextTryLock(ctx);
-
-    // GIL is already taken by the calling thread expecting to fail.
-    assert(res != REDISMODULE_OK);
-
-    return NULL;
-}
-
-void *worker(void *arg) {
-    // Retrieve blocked client
-    RedisModuleBlockedClient *bc = (RedisModuleBlockedClient *)arg;
-
-    // Get Redis module context
-    RedisModuleCtx *ctx = RedisModule_GetThreadSafeContext(bc);
-
-    // Acquire GIL
-    RedisModule_ThreadSafeContextLock(ctx);
-
-    // Create another thread which will try to acquire the GIL
-    pthread_t tid;
-    int res = pthread_create(&tid, NULL, sub_worker, ctx);
-    assert(res == 0);
-
-    // Wait for thread
-    pthread_join(tid, NULL);
-
-    // Release GIL
-    RedisModule_ThreadSafeContextUnlock(ctx);
-
-    // Reply to client
-    RedisModule_ReplyWithSimpleString(ctx, "OK");
-
-    // Unblock client
-    RedisModule_UnblockClient(bc, NULL);
-
-    // Free the Redis module context
-    RedisModule_FreeThreadSafeContext(ctx);
-
-    return NULL;
-}
-
-int acquire_gil(RedisModuleCtx *ctx, RedisModuleString **argv, int argc)
-{
-    UNUSED(argv);
-    UNUSED(argc);
-
-    int flags = RedisModule_GetContextFlags(ctx);
-    int allFlags = RedisModule_GetContextFlagsAll();
-    if ((allFlags & REDISMODULE_CTX_FLAGS_MULTI) &&
[…]

### 2022-03-07 paper/code/redis/tests/modules/eventloop.c

@@ -1,277 +1,0 @@
-/* This module contains four tests :
- * 1- test.sanity    : Basic tests for argument validation mostly.
- * 2- test.sendbytes : Creates a pipe and registers its fds to the event loop,
- *                     one end of the pipe for read events and the other end for
- *                     the write events. On writable event, data is written. On
- *                     readable event data is read. Repeated until all data is
- *                     received.
- * 3- test.iteration : A test for BEFORE_SLEEP and AFTER_SLEEP callbacks.
- *                     Counters are incremented each time these events are
- *                     fired. They should be equal and increment monotonically.
- * 4- test.oneshot   : Test for oneshot API
- */
-
-#define REDISMODULE_EXPERIMENTAL_API
-#include "redismodule.h"
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <memory.h>
-#include <errno.h>
-
-int fds[2];
-long long buf_size;
-char *src;
-long long src_offset;
-char *dst;
-long long dst_offset;
-
-RedisModuleBlockedClient *bc;
-RedisModuleCtx *reply_ctx;
-
-void onReadable(int fd, void *user_data, int mask) {
-    REDISMODULE_NOT_USED(mask);
-
-    RedisModule_Assert(strcmp(user_data, "userdataread") == 0);
-
-    while (1) {
-        int rd = read(fd, dst + dst_offset, buf_size - dst_offset);
-        if (rd <= 0)
-            return;
-        dst_offset += rd;
-
-        /* Received all bytes */
-        if (dst_offset == buf_size) {
-            if (memcmp(src, dst, buf_size) == 0)
-                RedisModule_ReplyWithSimpleString(reply_ctx, "OK");
-            else
-                RedisModule_ReplyWithError(reply_ctx, "ERR bytes mismatch");
-
-            RedisModule_EventLoopDel(fds[0], REDISMODULE_EVENTLOOP_READABLE);
-            RedisModule_EventLoopDel(fds[1], REDISMODULE_EVENTLOOP_WRITABLE);
-            RedisModule_Free(src);
[…]

### 2022-03-07 paper/code/redis/tests/modules/getkeys.c

@@ -1,121 +1,0 @@
-
-#include "redismodule.h"
-#include <strings.h>
-#include <assert.h>
-#include <unistd.h>
-#include <errno.h>
-
-#define UNUSED(V) ((void) V)
-
-/* A sample movable keys command that returns a list of all
- * arguments that follow a KEY argument, i.e.
- */
-int getkeys_command(RedisModuleCtx *ctx, RedisModuleString **argv, int argc)
-{
-    int i;
-    int count = 0;
-
-    /* Handle getkeys-api introspection */
-    if (RedisModule_IsKeysPositionRequest(ctx)) {
-        for (i = 0; i < argc; i++) {
-            size_t len;
-            const char *str = RedisModule_StringPtrLen(argv[i], &len);
-
-            if (len == 3 && !strncasecmp(str, "key", 3) && i + 1 < argc)
-                RedisModule_KeyAtPos(ctx, i + 1);
-        }
-
-        return REDISMODULE_OK;
-    }
-
-    /* Handle real command invocation */
-    RedisModule_ReplyWithArray(ctx, REDISMODULE_POSTPONED_LEN);
-    for (i = 0; i < argc; i++) {
-        size_t len;
-        const char *str = RedisModule_StringPtrLen(argv[i], &len);
-
-        if (len == 3 && !strncasecmp(str, "key", 3) && i + 1 < argc) {
-            RedisModule_ReplyWithString(ctx, argv[i+1]);
-            count++;
-        }
-    }
-    RedisModule_ReplySetArrayLength(ctx, count);
-
-    return REDISMODULE_OK;
-}
-
-int getkeys_fixed(RedisModuleCtx *ctx, RedisModuleString **argv, int argc)
-{
-    int i;
-
-    RedisModule_ReplyWithArray(ctx, argc - 1);
-    for (i = 1; i < argc; i++) {
-        RedisModule_ReplyWithString(ctx, argv[i]);
-    }
-    return REDISMODULE_OK;
-}
-
-/* Introspect a command using RM_GetCommandKeys() and returns the list
- * of keys. Essentially this is COMMAND GETKEYS implemented in a module.
- */
-int getkeys_introspect(RedisModuleCtx *ctx, RedisModuleString **argv, int argc)
-{
-    UNUSED(argv);
-    UNUSED(argc);
-
-    if (argc < 3) {
-        RedisModule_WrongArity(ctx);
-        return REDISMODULE_OK;
-    }
-
[…]

### 2022-03-07 paper/code/redis/tests/modules/keyspecs.c

@@ -1,116 +1,0 @@
-#include "redismodule.h"
-#ifdef INCLUDE_UNRELEASED_KEYSPEC_API
-
-#define UNUSED(V) ((void) V)
-
-int kspec_legacy(RedisModuleCtx *ctx, RedisModuleString **argv, int argc) {
-    UNUSED(argv);
-    UNUSED(argc);
-    RedisModule_ReplyWithSimpleString(ctx, "OK");
-    return REDISMODULE_OK;
-}
-
-int kspec_complex1(RedisModuleCtx *ctx, RedisModuleString **argv, int argc) {
-    UNUSED(argv);
-    UNUSED(argc);
-    RedisModule_ReplyWithSimpleString(ctx, "OK");
-    return REDISMODULE_OK;
-}
-
-int kspec_complex2(RedisModuleCtx *ctx, RedisModuleString **argv, int argc) {
-    UNUSED(argv);
-    UNUSED(argc);
-    RedisModule_ReplyWithSimpleString(ctx, "OK");
-    return REDISMODULE_OK;
-}
-
-int RedisModule_OnLoad(RedisModuleCtx *ctx, RedisModuleString **argv, int argc) {
-    REDISMODULE_NOT_USED(argv);
-    REDISMODULE_NOT_USED(argc);
-
-    int spec_id;
-
-    if (RedisModule_Init(ctx, "keyspecs", 1, REDISMODULE_APIVER_1)== REDISMODULE_ERR)
-        return REDISMODULE_ERR;
-
-    /* Test legacy range "gluing" */
-    if (RedisModule_CreateCommand(ctx,"kspec.legacy",kspec_legacy,"",0,0,0) == REDISMODULE_ERR)
-        return REDISMODULE_ERR;
-    RedisModuleCommand *legacy = RedisModule_GetCommand(ctx,"kspec.legacy");
-
-    if (RedisModule_AddCommandKeySpec(legacy,"RO ACCESS",&spec_id) == REDISMODULE_ERR)
-        return REDISMODULE_ERR;
-    if (RedisModule_SetCommandKeySpecBeginSearchIndex(legacy,spec_id,1) == REDISMODULE_ERR)
-        return REDISMODULE_ERR;
-    if (RedisModule_SetCommandKeySpecFindKeysRange(legacy,spec_id,0,1,0) == REDISMODULE_ERR)
-        return REDISMODULE_ERR;
-
[…]

### 2022-03-07 paper/code/redis/utils/generate-command-code.py

@@ -1,425 +1,0 @@
-#!/usr/bin/env python3
-
-import os
-import glob
-import json
-
-ARG_TYPES = {
-    "string": "ARG_TYPE_STRING",
-    "integer": "ARG_TYPE_INTEGER",
-    "double": "ARG_TYPE_DOUBLE",
-    "key": "ARG_TYPE_KEY",
-    "pattern": "ARG_TYPE_PATTERN",
-    "unix-time": "ARG_TYPE_UNIX_TIME",
-    "pure-token": "ARG_TYPE_PURE_TOKEN",
-    "oneof": "ARG_TYPE_ONEOF",
-    "block": "ARG_TYPE_BLOCK",
-}
-
-GROUPS = {
-    "generic": "COMMAND_GROUP_GENERIC",
-    "string": "COMMAND_GROUP_STRING",
-    "list": "COMMAND_GROUP_LIST",
-    "set": "COMMAND_GROUP_SET",
-    "sorted_set": "COMMAND_GROUP_SORTED_SET",
-    "hash": "COMMAND_GROUP_HASH",
-    "pubsub": "COMMAND_GROUP_PUBSUB",
-    "transactions": "COMMAND_GROUP_TRANSACTIONS",
-    "connection": "COMMAND_GROUP_CONNECTION",
-    "server": "COMMAND_GROUP_SERVER",
-    "scripting": "COMMAND_GROUP_SCRIPTING",
-    "hyperloglog": "COMMAND_GROUP_HYPERLOGLOG",
-    "cluster": "COMMAND_GROUP_CLUSTER",
-    "sentinel": "COMMAND_GROUP_SENTINEL",
-    "geo": "COMMAND_GROUP_GEO",
-    "stream": "COMMAND_GROUP_STREAM",
-    "bitmap": "COMMAND_GROUP_BITMAP",
-}
-
-RESP2_TYPES = {
-    "simple-string": "RESP2_SIMPLE_STRING",
-    "error": "RESP2_ERROR",
-    "integer": "RESP2_INTEGER",
-    "bulk-string": "RESP2_BULK_STRING",
-    "null-bulk-string": "RESP2_NULL_BULK_STRING",
-    "array": "RESP2_ARRAY",
-    "null-array": "RESP2_NULL_ARRAY",
-}
-
-RESP3_TYPES = {
-    "simple-string": "RESP3_SIMPLE_STRING",
-    "error": "RESP3_ERROR",
[…]

### 2022-03-07 scripts/RedisSim/kp.c

@@ -1,44 +1,0 @@
-#include<stdio.h>
-#include<math.h>
-#include<stdint.h>
-
-int main()
-{
-	uint64_t n=  14; 							// number of bits to determine bucket - this setting is for Redis
-	uint64_t m = ((uint64_t) 1) << n; 	// number of buckets
-	uint64_t ell = 50; 							// number of bits to determine bucket updates - this setting also for Redis
-	int ell_test; 									// difference to make to D
-	double alpha = 0.721; 					// magic constant
-	printf("m: %6llu\n",m);
-	double CE = 100000000.0; 				// starting cardinality estimate; vary this to simulate different conditions but don't go too small or the maths falls apart
-	printf("CE: %6f\n",CE);
-	double N = alpha*m*m; 				// numerator in CE equation, a fixed function of alpha and m
-	printf("N: %6f\n",N);
-	double D = N/CE; 							// denominator - CE = N/D
-	printf("D: %6f\n",D);
-	double delta = 1.0;
-	for (ell_test = 0; ell_test <= ell; ell_test++) {
-		delta = 0.5*delta; 						//update delta by halving it each time around the loop 
-		printf("ell_test: %6d\n",ell_test);
-		printf("delta: %.10e\n",delta);
-		double newD = D - delta; 			// newD updates D with the change // here we can get underflows; a real implementation of HLL would not die here because it computes D afresh each time, though it's clear the Redis implementation will have accuracy issues for \ell = 50 unless it uses quad accuracy?
-		if (newD < 0) 		{					// this can't occur in a real implementation but can in our simulation; this simply means ell_test is too small to be consistent with the starting value of CE
-			printf(" Value ell_test = %6d not consistent with starting CE\n",ell_test);  // this is not an error, just a limitation of the simulation
-			continue; 
-			}
-		printf("oldD: %.10e\n",D);
-		printf("newD: %.10e\n",newD);
-		double newCE = N/newD; 		// newCE updates the value of CE = N/D to the new value
-		printf("new CE: %.10e\n",newCE);
-		printf("CE difference: %.10e\n",newCE-CE);
-		if (newCE-CE > 1)
-			printf("Success for ell_test = %6d\n",ell_test); // we want a difference larger than 1 between old and new CE so as to survive any rounding
-		else 
-			{
-			printf("Failure for ell_test = %6d\n",ell_test); // 
-			break; //lets not go any further on failure
-			}
-		printf("\n");
-	}
-
-}
